### 2022-01-21 writing/investigating-GPP.tex
created paper

\documentclass[11pt,fleqn]{article}
\usepackage[margin=1in,top=1in,bottom=1in]{geometry}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{hyperref}
%\usepackage[dvips]{graphics}
%\usepackage[table]{xcolor}
%\usepackage{amssymb}
\usepackage{float}
%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{subcaption}

\usepackage[normalem]{ulem}

\usepackage{multicol}
\usepackage{txfonts}
\usepackage{amsfonts}
\usepackage{natbib}
\usepackage{gb4e}
\usepackage[all]{xy}
\usepackage{rotating}
\usepackage{tipa}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{url}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{array}

\def\bad{{\leavevmode\llap{*}}}
\def\marginal{{\leavevmode\llap{?}}}
\def\verymarginal{{\leavevmode\llap{??}}}
\def\swmarginal{{\leavevmode\llap{4}}}
\def\infelic{{\leavevmode\llap{\#}}}

\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
%\definecolor{gray}{rgb}{0.36, 0.54, 0.66}

\definecolor{Pink}{RGB}{240,0,120}
\newcommand{\red}[1]{\textcolor{Pink}{#1}}
\newcommand{\jd}[1]{\textbf{\textcolor{Pink}{[jd: #1]}}}

\newcommand{\dashrule}[1][black]{%
  \color{#1}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}\xleaders\hbox{\rule{4pt}{0pt}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}}\hfill\kern0pt%
}

\setlength{\parindent}{.3in}
\setlength{\parskip}{0ex}

\newcommand{\yi}{\'{\symbol{16}}}
\newcommand{\nasi}{\~{\symbol{16}}}
\newcommand{\hina}{h\nasi na}
\newcommand{\ina}{\nasi na}

\newcommand{\foc}{$_{\mbox{\small F}}$}

\hyphenation{par-ti-ci-pa-tion}

\setlength{\bibhang}{0.5in}
\setlength{\bibsep}{0mm}
\bibpunct[:]{(}{)}{,}{a}{}{,}

\newcommand{\6}{\mbox{$[\hspace*{-.6mm}[$}} 
\newcommand{\9}{\mbox{$]\hspace*{-.6mm}]$}}
\newcommand{\sem}[2]{\6#1\9$^{#2}$}
\renewcommand{\ni}{\~{\i}}

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{#1}}
\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{#1})}

\newcolumntype{R}[2]{%
    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
    l%
    <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{90}{0em}}}% no optional argument here, please!

\title{Does at-issueness predict projection? Further investigations of the Gradient Projection Principle}

%\thanks{For helpful comments on the research presented here, we thank David Beaver, Cleo Condoravdi, Kai von Fintel, Lauri Karttunen, Mandy Simons, Greg Scontras, the anonymous reviewers for {\em Semantics and Linguistic Theory} 2018, as well as the audiences at the MIT Linguistics colloquium, the 2018 Annual Meeting of XPRAG.de and at the University of T\"ubingen. We gratefully acknowledge financial support for this research from {\em National Science Foundation} grant BCS-1452674 (JT) and the Targeted Investment for Excellence Initiative at The Ohio State University (JT). IGOR Tuebingen, SALT TALK}}

\author{Author(s)}

%\author[$\circ$]{Judith Tonhauser}
%\author[$\bullet$]{Judith Degen}
%\affil[$\circ$]{The Ohio State University / University of Stuttgart}
%\affil[$\bullet$]{Stanford University}
%
%\renewcommand\Authands{ and }

\newcommand{\jt}[1]{\textbf{\color{blue}JT: #1}}

\begin{document}

%\tableofcontents
%\newpage

\maketitle

\vspace*{-1cm}

\begin{abstract}

\citealt{tbd-variability} hypothesized that at-issueness is one factor that modulates the projection of content (see also \citealt{brst-salt10,brst-ar}). Specifically, according to their Gradient Projection Principle, if content $C$ is expressed by a constituent embedded under an entailment-canceling operator, then $C$ projects to the extent that it is not at-issue. Their work provided evidence for the GPP from xx contents that had been described as projective in the literature, at-issueness was measured in two different ways, only one consistently brought evidence in support (question, asking whether, are you sure? mixed results 2b). This paper provides further investigate the GPP by investigating a) more varied items, b) with more types of embedding, and c) other measures of at-issueness. These experiments provide additional support for the GPP, and they also show: comparison of projection across entailment-canceling operators, comparison of at-issueness diagnostics. Suggests that the diagnostics may not all measure the same (\citealt{snider2017}).

\end{abstract}

\section{Introduction}\label{s1}

\begin{exe}
\ex\label{gpp} {\bf Gradient Projection Principle} \hfill (\citealt[400]{tbd-variability}) \\ If content $C$ is expressed by a constituent embedded under an entailment-canceling operator, then $C$ projects to the extent that it is not at-issue.
\end{exe}

\begin{exe}
\ex\label{rqs} {\bf Main research question} \\
Are not-at-issueness and projection positively correlated, as predicted by the Gradient Projection Principle?
\end{exe}

\begin{exe}
\ex\label{rqs} {\bf Ancillary research questions}
\begin{xlist}
\ex Is the projection of contents embedded under different entailment-canceling operators positively correlated, as assumed in the semantics/pragmatics literature?

\ex Is the at-issueness of contents embedded under different entailment-canceling operators positively correlated?

\ex Is the at-issueness of contents positively correlated under different measures of at-issueness?
\end{xlist}
\end{exe}

\section{Prior literature}\label{s2}

\section{Experiments 1-3: Methods}\label{s3}

\section{Experiments 1-3: Results}\label{s4}

\subsection{Comparing projection}

\subsection{Comparing at-isueness}

\subsection{GPP}

Further analyses:

\begin{itemize}
\item look at how many unique participants

\item compare projection across experiments/embeddings

compare projection across embeddings: research assumes that there are no differences, but there is some experimental evidence that there may be (Kathleen/Elizabeth, CommitmentBank), so it?s good to check (compare 4 experiments)

\item negative correlation between Q/A and assent might be compound of a) embedding (negation) and b) assent, because we don?t see negative correlation, but no correlation, with other embeddings (modal, conditional)

\item  compare variance for both projection and at-issueness, under the different embeddings and diagnostics: very little variance for assent diagnostic/negation embedding

\item have a look at the predicates that are exceptional to the GPP in Exps1 and 2, how do they behave across the other experiments? Exp3: looks like they are also their own little group

\item  when we compare negation, modal, conditional with assent diagnostic, we can see what effect embedding has

\item next experiments: no embedding of the 20 predicates, with assent diagnostic, to compare to prior literature who didn?t use embedding with the assent diagnostic; do assent diagnostic without ?that?s true? anaphor to engage with Snider?s assumption that ?yes? is not really anaphoric in the assent diagnostic and hence also not in the Q/A diagnostic

\end{itemize}

\section{General discussion}\label{s4}

\section{Conclusions}\label{s5}

% end document here for word count
%\end{document}

\bibliographystyle{cslipubs-natbib}
\bibliography{bibliography}

\newpage

\section*{Supplemental materials}

\appendix

\setcounter{page}{1}
%\renewcommand{\thetable}{A\arabic{table}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\section{20 complement clauses}\label{a-clauses}

The following clauses realized the complements of the predicates in Exps.~1-3:

\begin{enumerate}[leftmargin=3ex,itemsep=-2pt]

\begin{multicols}{2}

\item Mary is pregnant.
\item Josie went on vacation to France.
\item Emma studied on Saturday morning.
\item Olivia sleeps until noon.
\item Sophia got a tattoo.
\item Mia drank 2 cocktails last night.
\item Isabella ate a steak on Sunday.
\item  Emily bought a car yesterday.
\item  Grace visited her sister.
\item Zoe calculated the tip.

\columnbreak

\item  Danny ate the last cupcake.
\item  Frank got a cat.
\item  Jackson ran 10 miles.
\item  Jayden rented a car.
\item  Tony had a drink last night.
\item  Josh learned to ride a bike yesterday.
\item  Owen shoveled snow last winter.
\item  Julian dances salsa.
\item  Jon walks to work.
\item  Charley speaks Spanish.

\end{multicols}

\end{enumerate}

\section{Control stimuli in Exps.~1-3}\label{a-control}

\begin{exe}
\exi{(1)}  Control stimuli in Exps.~1
\begin{xlist}

\ex   Is Zack coming to the meeting tomorrow?

\ex Is Mary's aunt sick?

\ex Did Todd play football in high school?

\ex Is Vanessa good at math?

\ex Did Madison have a baby?

\ex Was Hendrick's car expensive?

\end{xlist}
\end{exe}

\begin{exe}
\exi{(2)} Control stimuli in Exps.~2
\begin{xlist}
\ex Entailing control stimuli
\begin{xlist}
\ex {\bf What is true:} Frederick managed to solve the problem. (Tested inference: Frederick solved the problem.)
\ex {\bf What is true:} Zack bought himself a car this morning. (Tested inference: Zack owns a car.)
\ex {\bf What is true:} Tara broke the window with a bat. (Tested inference: The window broke.)
\ex {\bf What is true:} Vanessa happened to look into the mirror. (Tested inference: Vanessa looked into the mirror.)
\end{xlist}
\ex Non-entailing control stimuli
\begin{xlist}
\ex {\bf What is true:} Dana watched a movie last night. (Tested inference: Dana wears a wig.)
\ex {\bf What is true:} Hendrick is renting an apartment. (Tested inference: The apartment has a balcony.)
\ex {\bf What is true:} Madison was unsuccessful in closing the window. (Tested inference:  Madison closed the window.)
\ex {\bf What is true:} Sebastian failed the exam. (Tested inference: Sebastian did really well on the exam.)
\end{xlist}
\end{xlist}
\end{exe}

\begin{exe}
\exi{(3)} Control stimuli in Exps.~3
\begin{xlist}
\ex Contradictory control stimuli
\begin{xlist}
\ex Madison laughed loudly and she didn't laugh.

\ex Dana has never smoked in her life and she stopped smoking recently.
\ex Hendrick's car is completely red and his car is not red.

\ex Sebastian lives in the USA and has never been to the USA.
\end{xlist}

\ex Non-contradictory control stimuli
\begin{xlist}
\ex Vanessa is really good at math, but I'm not.
\ex Zack believes that I'm married, but I'm actually single.
\ex Tara wants me to cook for her and I'm a terrific cook.
\ex Frederick is both smarter and taller than I am.

\end{xlist}
\end{xlist}
\end{exe}

\section{Data exclusion}\label{a-excl}

Table \ref{f-exclusion} presents how many participants' data were excluded from the analysis based on the exclusion criteria. The first column records the experiment, the second (`recruited') how many participants were recruited, and the final column (`remaining') how many participants' data entered the analysis. The `Exclusion criteria' columns show how many participants' data were excluded based on the four exclusion criteria:

\begin{itemize}[topsep = -1ex,itemsep=-2pt]

\item `multiple': Due to an experimental glitch, some participants participated in Exps.~1b, 2b or 3b more than once. Of these participants, we only analyzed the data from the first time they participated.

\item `language': Participants' data were excluded if they did not self-identify as native speakers of American English.

\item `controls': Participants' data were excluded if their response mean on the 6 control items was more than 2 sd above the group mean (Exp.~1a), if they gave a wrong rating (`yes') to more than one of the six controls (Exp.~1b), if their response mean on the entailing or the non-entailing controls was more than 2 sd below or above, respectively, the group mean (Exp.~2a), if they gave more than one wrong rating to one of the eight controls, where a wrong rating is a `yes' to a non-entailing control and a `no' to an entailing one (Exp.~2b), if their response means on the contradictory or non-contradictory controls were more than 2 sd below or above, respectively, the group mean (Exp.~3a), and if they gave more than one wrong response to one of the eight control sentences, where a wrong response was a `yes' to a non-contradictory control or a `no' to a contradictory one (Exp.~3b).

\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale, that is, if the variance of their response distribution was more than 2 sd below the group mean variance.

\end{itemize}

\begin{table}[h!]
\centering
\begin{tabular}{l r | r r r r | r}
&  & \multicolumn{4}{c|}{Exclusion criteria} &  \\ 
&  recruited & multiple & language & controls & variance & remaining \\ 
\hline
Exp.~1q & xx & xx & xx & xx & xx & xx \\
Exp.~1n & xx & xx & xx & xx & xx & xx \\
Exp.~1m & xx & xx & xx & xx & xx & xx \\
Exp.~1c & xx & xx & xx & xx & xx & xx \\
Exp.~2q & xx & xx & xx & xx & xx & xx \\
Exp.~2n & xx & xx & xx & xx & xx & xx \\
Exp.~2m & xx & xx & xx & xx & xx & xx \\
Exp.~2c & xx & xx & xx & xx & xx & xx \\
Exp.~3q & xx & xx & xx & xx & xx & xx \\
Exp.~3n & xx & xx & xx & xx & xx & xx \\
Exp.~3m & xx & xx & xx & xx & xx & xx \\
Exp.~3c & xx & xx & xx & xx & xx & xx \\
\end{tabular}
\caption{Data exclusion in Exps.~1, 2 and 3}\label{f-exclusion}
\end{table}

\end{document}

### 2022-01-31 writing/investigating-GPP.tex
updated analyses and paper

@@
 
+\begin{exe}
+\ex\label{pred} 20 clause-embedding predicates
+
+\begin{xlist}
+
+\ex canonically factive: {\em be annoyed, discover, know, reveal, see}
+
+\ex nonfactive:
+
+\begin{xlist}
+
+\ex nonveridical nonfactive: {\em pretend, suggest, say, think}
+
+\ex veridical nonfactive: {\em be right, demonstrate}
+
+\end{xlist}
+
+\ex optionally factive: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove}
+
+\end{xlist}
+
+\end{exe}
+
 \section{Experiments 1-3: Methods}\label{s3}
+
+Exps.~1a and 1b were designed to investigate which of the CCs of the 20 clause-embedding predicates in \ref{pred} are presupposed.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
+All experiments were conducted with IRB approval.
+Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.}
+
+The experiments employed the {\sc certain that} diagnostic for projection (see also, e.g.\ \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}): on this diagnostic, participants are presented with an utterance in which the content to be investigated occurs in an entailment-canceling environment, like a polar question, as illustrated in \ref{stim} for the content of the appositive, that Martha's new car is a BMW.\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}
+
+\begin{exe}
+
+\ex\label{stim} Patrick asks: {\em Was Martha's new car, a BMW, expensive?}
+
+\end{exe}
+To assess projection, participants are asked whether the speaker is certain of the content; for instance, in \ref{stim}, participants are asked whether Patrick is certain that Martha's new car is a BMW.
+If a participant takes the speaker to be certain of the content, we assume that the content projects, that is, the content is presupposed by the speaker; if a participant does not take the speaker to be certain of the content, we assume that the content does not project, that is, the content is not presupposed.
+
+Following \citealt{tbd-variability}, Exp.~1a implemented the certain that diagnostic with a gradient response scale: participants gave their responses on a slider marked `no' at one end and `yes' at the other.
+We assume that the closer to `yes' a participant's response is, the more the speaker is certain of the content, that is, the more projective the content is.
+In other words, we assume that gradient certainty ratings reflect gradience in the speaker's commitment to the truth of the content.\footnote{Strictly speaking,  gradient certainty ratings reflect gradience in the degree to which participants \emph{perceive} the speaker to be committed to the truth of the content.
+That is, as in any experiment, the quantity of interest, in this case speaker commitment, is only indirectly measured.} As discussed in \citealt{tbd-variability}, a second interpretation of gradient certainty ratings is that they reflect a participant's degree of belief in the speaker's (binary) commitment to the truth of the content.
+While we adopt the first interpretation in our discussion, we remain agnostic about the interpretation of gradient certainty ratings: either interpretation gives rise to the expectation,  on the definition of factive predicates in \ref{def}a, that certainty ratings for factive predicates are categorically higher than for optionally factive and nonfactive ones.
+
+Exp.~1b was identical to Exp.~1a but employed a two-alternative forced choice task where participants categorically responded `yes' or `no'.
+
+\subsection{Participants}
+
+We recruited 250-300 participants for each of Exps.~1-3.
+Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs.
+Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
+Information on recruited participants (total number, ages, gender) and the payments they received can be found in Supplement \ref{a-participants}.
+
+\subsection{Materials}
+
+Exps.~1-3 investigated the projection and at-issueness of the CC of the 20 clause-embedding predicates in (\ref{pred}).
+The clausal complements of the 20 predicates were realized by 20 clauses (provided in Supplement \ref{a-clauses}), for a total of 400 predicate/clause combinations.
+These 400 predicate/clause combinations were combined with proper name subjects (a unique name per complement clause) to realize polar questions (in Exps.
+1q, 2q, and 3q), negated declaratives (in Exps.
+1n, 2n, and 3n), declaratives with the modal adverb {\em perhaps} (in Exps.
+1m, 2m, and 3m), and the antecedents of conditionals (in Exps.
+1c, 2c, and 3c).
+These sentences were presented as utterances by a named speaker.
+A sample set of target utterances is given in (\ref{sample-stims}): here, the named speaker is {\em Daniel}, the subject of the clause-embedding predicate is {\em Cole}, the predicate is {\em discover}, and the complement clause is {\em Julian dances salsa}.\footnote{The indirect object of {\em inform} was realized by the proper name {\em Sam}.
+The predicates differed in the tense in which they were realized.
+In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.
+In the antecedents of conditionals, all predicates were realized in the present tense.
+For the consequents of the conditional sentences see Supplement \ref{a-target}.}
+
+\begin{exe}
+\ex\label{sample-stims}
+\begin{xlist}
+\ex {\bf Daniel:} ``{\em Did Cole discover that Julian dances salsa?}"
+\ex {\bf Daniel:} ``{\em Cole didn't discover that Julian dances salsa.}"
+\ex {\bf Daniel:} ``{\em Perhaps Cole discovered that Julian dances salsa.}"
+\ex {\bf Daniel:} ``{\em If Cole discovered that Julian dances salsa, Logan will be joyful.}"
+\end{xlist}
+\end{exe}
+The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique.
+
+Projection and at-issueness of the contents of the clausal complements were measured in separate blocks.
+In the projection blocks across Exps.~1-3, projection was measured with the `certain that' diagnostic.
+The target stimuli consisted of the target utterances, as shown in Figure \ref{f-projection-trials}.
+
+\begin{figure}[h!]
+\centering
+
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[width=7.55cm]{figures/exp1q-projection-trial}}
+\caption{Target trial in projection block of Exp.~1q.}\label{fig-exp1q-projection}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[width=7.9cm]{figures/exp1n-projection-trial}} 
+\caption{Target trial in projection block of Exp.~1n.}\label{fig-exp1n-projection}
+ \end{subfigure}
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/exp1m-projection-trial}}
+\caption{Target trial in projection block of Exp.~1m}\label{fig-exp1m-projection}
+ \end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[height=5.4cm,width=7.9cm]{figures/exp1c-projection-trial}} 
+\caption{Target trial in projection block of Exp.~1c}\label{fig-exp1c-projection}
+\end{subfigure}
+
+\caption{Target trials in projection blocks of Exps.~1.}\label{f-projection-trials}
+\end{figure}
+
+The at-issueness diagnostics differed between Exps.~1, 2, and 3: at-issueness was measured in Exp.~1q with the `asking whether' diagnostic, with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, with the `yes' diagnostic in Exps.~2q and 3q, with the `assent with positive continuation' diagnostic in Exps.~2n, 2m, and 2c,  and with the `assent with adversative continuation' diagnostic in Exps.~3n, 3m, and 3c.
+
+, as shown in Figure \ref{f-ai-trials}.
+
+To assess whether participants were attending to the task, each experiment also included six control stimuli, which were also utterances made by a named speaker.
+(The controls are provided in Supplement \ref{a-control}.) Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same 6 control stimuli.
+Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
+Block order and within-block trial order was randomized.
+
+In the projection block, target stimuli consisted of a fact and a polar question that was ut- tered by a named speaker, as shown in Figure 1B.
+The polar questions were formed by real- izing the 20 clauses as the complements of the 20 clause-embedding predicates in Figure 1C.
+Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question.
+Projection was measured using the ?certain that?
+diagnostic (Djärv \& Bacovcin, 2017; Lorson, 2018; Mahler, 2020; Tonhauser et al., 2018): participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
+They gave their responses on a slider marked ?no?
+at one end (coded as 0) and ?yes?
+at the other (coded as 1).
+Greater speaker commitment to the CC should result in higher slider ratings.
+
+\subsection{Procedure}
+
+Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask a question.
+Participants were asked to rate whether the speaker was certain of the CC.
+They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1), as shown in Figure \ref{fig-trial-exp1}.
+
+\begin{figure}[h!]
+\begin{center}
+\fbox{\includegraphics[width=10cm]{Language-figures/color/Figure1}}
+\end{center}
+\caption{A sample trial in Experiment 1a}\label{fig-trial-exp1}
+\end{figure}
+
+%After completing the experiment (as well as the other five experiments we report on), participants filled out a short, optional survey about their age, their gender, their native language(s) and, if English is their native language, whether they are a speaker of American English (as opposed to, e.g.\ Australian or Indian English).
+To encourage them to respond truthfully, participants were told that they would be paid no matter what answers they gave in the survey.
+
+After completing the experiment, participants filled out a short optional demographic sur- vey.
+To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.
+
+\subsection{Data exclusion}
+
+We excluded data from participants who took any experiment more than once and who did not self-declare to be native speakers of American English.
+We also excluded data from participants based on their ratings on the main clause controls and other criteria given in Supplement \ref{a-participants}.
+In each experiment, the data from at least {\bf FIX} participants were analyzed.
+Information on the participants whose data entered into the analysis (total number, ages, gender) can be found in Supplement \ref{a-participants}.
 
@@
 
+\section{Consequents for target stimuli in Exps.~1c, 2c, and 3c}\label{a-target}
+
+10 positive valuence, 10 negative valuence {\bf which other considerations?}
+
+\begin{enumerate}[leftmargin=3ex,itemsep=-2pt]
+
+\item \ldots that Mary is pregnant, Esther will be mad.
+\item \ldots that Josie went on vacation to France, Arnold will be frustrated.
+\item \ldots that Emma studied on Saturday morning, Liam will be proud.
+\item \ldots that Olivia sleeps until noon, Elijah will be embarrassed.
+\item \ldots that Sophia got a tattoo, Ariel will be giddy.
+\item \ldots that Mia drank 2 cocktails last night, Mariela will be worried.
+\item \ldots that Isabella ate a steak on Sunday, Liz will be delighted.
+\item  \ldots that Emily bought a car yesterday, Kate will be excited.
+\item \ldots that  Grace visited her sister, Henry will be surprised.
+\item \ldots that Zoe calculated the tip, Alex will be astonished.
+\item  \ldots that Danny ate the last cupcake, Harper will be disgusted.
+\item  \ldots that Frank got a cat, Lucas will be grouchy.
+\item  \ldots that Jackson ran 10 miles, Kayla will be cheerful.
+\item  \ldots that Jayden rented a car, Brittany will be furious.
+\item  \ldots that Tony had a drink last night, Victoria will be ashamed.
+\item  \ldots that Josh learned to ride a bike yesterday, Mason will be envious.
+\item  \ldots that Owen shoveled snow last winter, Bianca will be jealous.
+\item  \ldots that Julian dances salsa, Logan will be joyful.
+\item  \ldots that Jon walks to work, Caleb will be suspicious.
+\item  \ldots that Charley speaks Spanish, Jay will be happy.
+
+\end{enumerate}
+
 \section{Control stimuli in Exps.~1-3}\label{a-control}
 
+The control stimuli in Exps.~1-3 were formed from the sentences in (1).
+In Exps.~1, we used the polar question variants of the sentences in (1) in Exp.~1q, and the (positive declarative) sentences in (1) in Exps.~1n, 1m, and 1c; in Exp.~1c, we added the non-restrictive relative clauses (NRRCs) provided in (1) to the respective subjects, so that the control stimuli have two clauses, like the target stimuli.
+
+We used positive declarative control sentences in Exps.~1n, 1m, and 1c so that we could have clear expectations about participants' responses.
+Thus, we expected participants to give low responses in the `certain that' and `asking whether' diagnostic in Exp.~1q (indicating that the main clause content does not project and is at-issue), and we expected participants to give low responses in the `certain that' diagnostic and high responses in the `sure that' diagnostic in Exps.~1n, 1m, and 1c (indicating, again, that the main clause content does not project and is at-issue).
+
 \begin{exe}
-\exi{(1)}  Control stimuli in Exps.~1
+\exi{(1)}  Sentences for control stimuli in Exps.~1
 \begin{xlist}
-
-\ex   Is Zack coming to the meeting tomorrow?
-
-\ex Is Mary's aunt sick?
-
-\ex Did Todd play football in high school?
-
-\ex Is Vanessa good at math?
-
-\ex Did Madison have a baby?
-
-\ex Was Hendrick's car expensive?
-
+\ex These muffins have blueberries in them.  (NRRC: , which are really delicious, )
+\ex This pizza has mushrooms on it. (NRRC: , which I just made from scratch, )
+\ex Jack was playing outside with the kids. (NRRC: , who is my long-time neighbor, )
+\ex Ann dances ballet. (NRRC: , who is a local performer, )
+\ex John's kids were in the garage. (NRRC: , who are very well-behaved, )
+\ex Samantha has a new hat. (NRRC: , who is really into fashion, )
 \end{xlist}
@@
 
-\begin{exe}
-\exi{(2)} Control stimuli in Exps.~2
-\begin{xlist}
-\ex Entailing control stimuli
-\begin{xlist}
-\ex {\bf What is true:} Frederick managed to solve the problem. (Tested inference: Frederick solved the problem.)
-\ex {\bf What is true:} Zack bought himself a car this morning. (Tested inference: Zack owns a car.)
-\ex {\bf What is true:} Tara broke the window with a bat. (Tested inference: The window broke.)
-\ex {\bf What is true:} Vanessa happened to look into the mirror. (Tested inference: Vanessa looked into the mirror.)
-\end{xlist}
-\ex Non-entailing control stimuli
-\begin{xlist}
-\ex {\bf What is true:} Dana watched a movie last night. (Tested inference: Dana wears a wig.)
-\ex {\bf What is true:} Hendrick is renting an apartment. (Tested inference: The apartment has a balcony.)
-\ex {\bf What is true:} Madison was unsuccessful in closing the window. (Tested inference:  Madison closed the window.)
-\ex {\bf What is true:} Sebastian failed the exam. (Tested inference: Sebastian did really well on the exam.)
-\end{xlist}
-\end{xlist}
-\end{exe}
+In Exps.~2 and 3, the NRRCs were included in all control stimuli: the carrier sentences were the polar question variants of the sentences in (1) in Exps.~2q and 3q, and the sentences in (1) for the remaining experiments.
+The NRRCs were included in Exps.~2 and 3 to make the use of the assent diagnostics in the at-issueness blocks more natural: like the target stimuli, the control stimuli consist of two clauses, one of which the relevant speaker assents with.
 
-\begin{exe}
-\exi{(3)} Control stimuli in Exps.~3
-\begin{xlist}
-\ex Contradictory control stimuli
-\begin{xlist}
-\ex Madison laughed loudly and she didn't laugh.
+We had the same expectations as for the control stimuli in Exps.~1: low responses in the `certain that' and `asking whether' diagnostic in Exps.~2q and 3q (indicating that the main clause content does not project and is at-issue), and low responses in the `certain that' diagnostic and high responses in the `sure that' diagnostic in the remaining experiments (indicating, again, that the main clause content does not project and is at-issue).
 
-\ex Dana has never smoked in her life and she stopped smoking recently.
-\ex Hendrick's car is completely red and his car is not red.
+{\bf Provide information on whether these expectations were met -- THEY WERE NOT ALWAYS}
 
-\ex Sebastian lives in the USA and has never been to the USA.
-\end{xlist}
+\newpage
 
-\ex Non-contradictory control stimuli
-\begin{xlist}
-\ex Vanessa is really good at math, but I'm not.
-\ex Zack believes that I'm married, but I'm actually single.
-\ex Tara wants me to cook for her and I'm a terrific cook.
-\ex Frederick is both smarter and taller than I am.
+\section{Data exclusion}\label{a-participants}
 
-\end{xlist}
-\end{xlist}
-\end{exe}
+{\bf Include information on payment}
 
-\section{Data exclusion}\label{a-excl}
+Gender: no data collected in Exp 1q
+
+Relevant manual inspection on `variance': Exp 1q, Exp 2q, Exp 3q (participants automatically identified, but not all excluded after manual inspection)
 
@@
 
+{\bf comment on why so many women}
+
 \begin{itemize}[topsep = -1ex,itemsep=-2pt]
 
-\item `multiple': Due to an experimental glitch, some participants participated in Exps.~1b, 2b or 3b more than once.
-Of these participants, we only analyzed the data from the first time they participated.
+\item `multiple': Due to an experimental glitch, some participants participated more than once.
+Since no information was available on which one was their first take, those participants' data was removed (which means that for each participant who took the experiment twice, at least two data sets were removed).
 
@@
 
-\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale, that is, if the variance of their response distribution was more than 2 sd below the group mean variance.
+\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the non-control trials.
+We identified such participants by calculating the variance of the response distribution for each participant, and manually inspecting those whose variance was more than 2 sd below the group mean variance.
 
@@
 
+{\bf Exp 3n, 3m, 3c: need to look at not-at-issue ratings on controls to see how to exclude}
+
 \begin{table}[h!]
 \centering
-\begin{tabular}{l r | r r r r | r}
-&  & \multicolumn{4}{c|}{Exclusion criteria} &  \\ 
-&  recruited & multiple & language & controls & variance & remaining \\ 
+\begin{tabular}{l | r r r | r r r r | r r r }
+&  \multicolumn{3}{c|}{Recruited participants} & \multicolumn{4}{c|}{Exclusion criteria} & \multicolumn{3}{c}{Remaining participants}  \\ 
+Exp. & total & ages (mean) & f/m/o/u & mult & lang & cntr & var & total & ages (mean) & f/m/o/u  \\ 
 \hline
-Exp.~1q & xx & xx & xx & xx & xx & xx \\
-Exp.~1n & xx & xx & xx & xx & xx & xx \\
-Exp.~1m & xx & xx & xx & xx & xx & xx \\
-Exp.~1c & xx & xx & xx & xx & xx & xx \\
-Exp.~2q & xx & xx & xx & xx & xx & xx \\
-Exp.~2n & xx & xx & xx & xx & xx & xx \\
-Exp.~2m & xx & xx & xx & xx & xx & xx \\
-Exp.~2c & xx & xx & xx & xx & xx & xx \\
-Exp.~3q & xx & xx & xx & xx & xx & xx \\
-Exp.~3n & xx & xx & xx & xx & xx & xx \\
-Exp.~3m & xx & xx & xx & xx & xx & xx \\
-Exp.~3c & xx & xx & xx & xx & xx & xx \\
+1q & 300  & 19-74 (38.2)  & --  & 5  & 7  & 35  & 0  & 242   & 21-74 (39.2) &  -- \\
+1n & 300  & 18-74 (33.2)  &  150/145/5/0 & 0  & 8 & 17 & 1 & 274  & 18-74 (33.3) & 141/128/5/0 \\
+1m & 300  & 18-74 (32.7) & 150/141/7/2  &  0 & 0  & 19 & 0   & 281   & 18-74 (32.7) & 144/129/7/1 \\
+1c &  300 & 18-58 (25.9)  & 249/45/6/0 & 0 & 6  & 26 & 2  & 266  &  18-58 (24.8) & 235/25/6/0  \\
+2q & 250  &  18-58 (25.5) &  201/43/6/0  & 0 & 4  & 24  & 1 & 220  & 18-58 (24.8)  & 187/28/5/0  \\
+2n & 250  &  18-69 (33.2)  &  127/114/6/1 & 1  & 4  & 29 & 0& 215  &  18-69 (33.1) & 113/95/6/1    \\
+2m & 251  & 18-74 (31.7)  & 132/113/6/0  & 0  & 4  & 27 &  0 & 220  & 18-70 (31.9) & 116/98/6/0 \\
+2c &  250 &  18-56 (24.5)  & 212/30/8/0  & 0  & 0  & 26 & 0  & 224  & 18-56 (24.4) & 195/24/5/0 \\
+3q & 250  &  18-66 (32.4) &  140/102/7/1 &  0 & 4  & 20  & 0  &  225 & 18-66 (32.6) & 125/93//7/0  \\
+3n & 250  &  18-70 (24.6) & 114/31/5/0  & 0  & 5  &  &   &   &  &  \\
+3m & 250  & 18-63 (25.5)  & 205/40/5/0 &  0 & 3  &  &   &   &  &  \\
+3c & 250  & 18-59  (27.5) & 182/64/4/0  & 0  & 3   &  &   &   &  &  \\
 \end{tabular}
-\caption{Data exclusion in Exps.~1, 2 and 3}\label{f-exclusion}
+\caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3}\label{f-exclusion}
 \end{table}

### 2022-01-31 experiments/main/4_projaiC-past-tense/js/experiment.js

@@ -861,486 +861,486 @@ var items = _.shuffle([
    }
  ]);
 
-  var contents = {
+   var contents = {
    "mary": {
      "question":"Mary is pregnant",
      "MC":"Mary is pregnant.",
-      "be_annoyed":"If Mandy is annoyed that Mary is pregnant, Esther will be mad.",
-     "discover":"If Mandy discovers that Mary is pregnant, Esther will be mad.",
+     "be_annoyed":"If Mandy is annoyed that Mary is pregnant, Esther will be mad.",
+     "discover":"If Mandy discovered that Mary is pregnant, Esther will be mad.",
      "know":"If Mandy knows that Mary is pregnant, Esther will be mad.",
-     "reveal":"If Mandy reveals that Mary is pregnant, Esther will be mad.",
-     "see":"If Mandy sees that Mary is pregnant, Esther will be mad.",
-     "pretend":"If Mandy pretends that Mary is pregnant, Esther will be mad.",
-     "suggest":"If Mandy suggests that Mary is pregnant, Esther will be mad.",
-     "say":"If Mandy says that Mary is pregnant, Esther will be mad.",
+     "reveal":"If Mandy revealed that Mary is pregnant, Esther will be mad.",
+     "see":"If Mandy saw that Mary is pregnant, Esther will be mad.",
+     "pretend":"If Mandy pretended that Mary is pregnant, Esther will be mad.",
+     "suggest":"If Mandy suggested that Mary is pregnant, Esther will be mad.",
+     "say":"If Mandy said that Mary is pregnant, Esther will be mad.",
      "think":"If Mandy thinks that Mary is pregnant, Esther will be mad.",
      "be_right":"If Mandy is right that Mary is pregnant, Esther will be mad.",
-     "demonstrate":"If Mandy demonstrates that Mary is pregnant, Esther will be mad.",
-     "acknowledge":"If Mandy acknowledges that Mary is pregnant, Esther will be mad.",
-     "admit":"If Mandy admits that Mary is pregnant, Esther will be mad.",
-     "announce":"If Mandy announces that Mary is pregnant, Esther will be mad.",
-     "confess":"If Mandy confesses that Mary is pregnant, Esther will be mad.",
-     "confirm":"If Mandy confirms that Mary is pregnant, Esther will be mad.",
-     "establish":"If Mandy establishes that Mary is pregnant, Esther will be mad.",
[…]

### 2022-01-31 experiments/main/4_projaiC/js/experiment.js

@@ -861,486 +861,486 @@ var items = _.shuffle([
    }
  ]);
 
-   var contents = {
+  var contents = {
    "mary": {
      "question":"Mary is pregnant",
      "MC":"Mary is pregnant.",
-     "be_annoyed":"If Mandy is annoyed that Mary is pregnant, Esther will be mad.",
-     "discover":"If Mandy discovered that Mary is pregnant, Esther will be mad.",
+      "be_annoyed":"If Mandy is annoyed that Mary is pregnant, Esther will be mad.",
+     "discover":"If Mandy discovers that Mary is pregnant, Esther will be mad.",
      "know":"If Mandy knows that Mary is pregnant, Esther will be mad.",
-     "reveal":"If Mandy revealed that Mary is pregnant, Esther will be mad.",
-     "see":"If Mandy saw that Mary is pregnant, Esther will be mad.",
-     "pretend":"If Mandy pretended that Mary is pregnant, Esther will be mad.",
-     "suggest":"If Mandy suggested that Mary is pregnant, Esther will be mad.",
-     "say":"If Mandy said that Mary is pregnant, Esther will be mad.",
+     "reveal":"If Mandy reveals that Mary is pregnant, Esther will be mad.",
+     "see":"If Mandy sees that Mary is pregnant, Esther will be mad.",
+     "pretend":"If Mandy pretends that Mary is pregnant, Esther will be mad.",
+     "suggest":"If Mandy suggests that Mary is pregnant, Esther will be mad.",
+     "say":"If Mandy says that Mary is pregnant, Esther will be mad.",
      "think":"If Mandy thinks that Mary is pregnant, Esther will be mad.",
      "be_right":"If Mandy is right that Mary is pregnant, Esther will be mad.",
-     "demonstrate":"If Mandy demonstrated that Mary is pregnant, Esther will be mad.",
-     "acknowledge":"If Mandy acknowledged that Mary is pregnant, Esther will be mad.",
-     "admit":"If Mandy admitted that Mary is pregnant, Esther will be mad.",
-     "announce":"If Mandy announced that Mary is pregnant, Esther will be mad.",
-     "confess":"If Mandy confessed that Mary is pregnant, Esther will be mad.",
-     "confirm":"If Mandy confirmed that Mary is pregnant, Esther will be mad.",
-     "establish":"If Mandy established that Mary is pregnant, Esther will be mad.",
[…]

### 2022-01-31 results/main/10_projaiN/rscripts/preprocessing.R

@@ -1,6 +1,6 @@
-# 6_projaiN
-# Exp 2n
-# projection and at-issueness (assent with positive continuation)
+# 10_projaiN
+# Exp 3n
+# projection and at-issueness (assent with adversative continuation)
 # preprocessing
 
 # set working directory to directory of script
@@ -22,17 +22,14 @@ nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)
 head(d)
 summary(d)
 
-length(unique(d$workerid)) #249
-
-# participant 1877 took the experiment twice:
-# View(table(d$workerid))
+length(unique(d$workerid)) #250
 
 # read in the subject information
 ds = read_csv("../data/experiment-subject_information.csv")
-length(unique(ds$workerid)) #249
+length(unique(ds$workerid)) #250
 nrow(ds) #250
 head(ds)
-summary(ds) 
+summary(d) 
 
 # look at participants' comments
 unique(ds$comments)
@@ -41,30 +38,17 @@ unique(ds$comments)
 d = d %>%
   left_join(ds, by=c("workerid"))
 
-# remove data from participant 1877 since no information on which was first take
-d <- d %>%
-  filter(workerid != "1877") %>%
-  droplevels()
-length(unique(d$workerid)) #248
-
 # participant info
-table(d$age) #18-69
-
-str(d$age)
-# turn non-numeric age information into NA
-d$age <- as.numeric(as.character(d$age))
-table(d$age)
-
-length(which(is.na(d$age))) # 0 missing values
-mean(d$age,na.rm=TRUE) #33.2
+table(d$age) #18-70
+length(which(is.na(d$age))) #0 missing values
+mean(d$age,na.rm=TRUE) #24.6
 
 d %>% 
   select(gender, workerid) %>% 
   unique() %>% 
   group_by(gender) %>% 
   summarize(count=n())
-#127 female, 114 male, 6 other, 1 undeclared
-
+#114 female, 31 male, 5 other, 0 undeclared
 
 # change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
 # by subtracting the ai responses from 1
@@ -81,26 +65,25 @@ unique(d$trial) # trials from 1 to 52
 ## exclude participants' data ----
 
 ### exclude non-American English speakers
-length(unique(d$workerid)) #248
-length(which(is.na(d$language))) #no missing responses
+length(unique(d$workerid)) #250
[…]

### 2022-01-31 results/main/11_projaiM/rscripts/preprocessing.R

@@ -1,6 +1,6 @@
-# 6_projaiN
-# Exp 2n
-# projection and at-issueness (assent with positive continuation)
+# 11_projaiM
+# Exp 3m
+# projection and at-issueness (assent with adversative continuation)
 # preprocessing
 
 # set working directory to directory of script
@@ -22,17 +22,14 @@ nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)
 head(d)
 summary(d)
 
-length(unique(d$workerid)) #249
-
-# participant 1877 took the experiment twice:
-# View(table(d$workerid))
+length(unique(d$workerid)) #250
 
 # read in the subject information
 ds = read_csv("../data/experiment-subject_information.csv")
-length(unique(ds$workerid)) #249
+length(unique(ds$workerid)) #250
 nrow(ds) #250
 head(ds)
-summary(ds) 
+summary(d) 
 
 # look at participants' comments
 unique(ds$comments)
@@ -41,30 +38,17 @@ unique(ds$comments)
 d = d %>%
   left_join(ds, by=c("workerid"))
 
-# remove data from participant 1877 since no information on which was first take
-d <- d %>%
-  filter(workerid != "1877") %>%
-  droplevels()
-length(unique(d$workerid)) #248
-
 # participant info
-table(d$age) #18-69
-
-str(d$age)
-# turn non-numeric age information into NA
-d$age <- as.numeric(as.character(d$age))
-table(d$age)
-
-length(which(is.na(d$age))) # 0 missing values
-mean(d$age,na.rm=TRUE) #33.2
+table(d$age) #18-63
+length(which(is.na(d$age))) #0 missing values
+mean(d$age,na.rm=TRUE) #25.5
 
 d %>% 
   select(gender, workerid) %>% 
   unique() %>% 
   group_by(gender) %>% 
   summarize(count=n())
-#127 female, 114 male, 6 other, 1 undeclared
-
+#205 female, 40 male, 5 other, 0 undeclared
 
 # change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
 # by subtracting the ai responses from 1
@@ -81,26 +65,25 @@ unique(d$trial) # trials from 1 to 52
 ## exclude participants' data ----
 
 ### exclude non-American English speakers
-length(unique(d$workerid)) #248
-length(which(is.na(d$language))) #no missing responses
+length(unique(d$workerid)) #250
[…]

### 2022-01-31 results/main/12_projaiC/rscripts/preprocessing.R

@@ -1,6 +1,6 @@
-# 6_projaiN
-# Exp 2n
-# projection and at-issueness (assent with positive continuation)
+# 12_projaiC
+# Exp 3c
+# projection and at-issueness (assent with adversative continuation)
 # preprocessing
 
 # set working directory to directory of script
@@ -22,17 +22,14 @@ nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)
 head(d)
 summary(d)
 
-length(unique(d$workerid)) #249
-
-# participant 1877 took the experiment twice:
-# View(table(d$workerid))
+length(unique(d$workerid)) #250
 
 # read in the subject information
 ds = read_csv("../data/experiment-subject_information.csv")
-length(unique(ds$workerid)) #249
+length(unique(ds$workerid)) #250
 nrow(ds) #250
 head(ds)
-summary(ds) 
+summary(d) 
 
 # look at participants' comments
 unique(ds$comments)
@@ -41,30 +38,19 @@ unique(ds$comments)
 d = d %>%
   left_join(ds, by=c("workerid"))
 
-# remove data from participant 1877 since no information on which was first take
-d <- d %>%
-  filter(workerid != "1877") %>%
-  droplevels()
-length(unique(d$workerid)) #248
-
 # participant info
-table(d$age) #18-69
+table(d$age) #18-59
+length(which(is.na(d$age))) #0 missing values
 
-str(d$age)
-# turn non-numeric age information into NA
 d$age <- as.numeric(as.character(d$age))
-table(d$age)
-
-length(which(is.na(d$age))) # 0 missing values
-mean(d$age,na.rm=TRUE) #33.2
+mean(d$age,na.rm=TRUE) #27.5
 
 d %>% 
   select(gender, workerid) %>% 
   unique() %>% 
   group_by(gender) %>% 
   summarize(count=n())
-#127 female, 114 male, 6 other, 1 undeclared
-
+#182 female, 64 male, 4 other, 0 undeclared
 
 # change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
 # by subtracting the ai responses from 1
@@ -81,26 +67,25 @@ unique(d$trial) # trials from 1 to 52
 ## exclude participants' data ----
 
 ### exclude non-American English speakers
-length(unique(d$workerid)) #248
-length(which(is.na(d$language))) #no missing responses
+length(unique(d$workerid)) #250
+length(which(is.na(d$language))) #0 missing responses
[…]

### 2022-01-31 results/main/1_projaiQ/rscripts/preprocessing.R

@@ -1,15 +1,20 @@
-# questions: projection and at-issueness
+# 1_projaiQ
+# Exp 1q
+# projection and at-issueness (asking whether)
 # preprocessing
 
 # set working directory to directory of script
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
-source('../../helpers.R')
-theme_set(theme_bw())
+# load required packages
+library(tidyverse)
 
-# load required packages for pre-processing data
-require(tidyverse)
+# load helpers
+source('../../helpers.R')
+
+# set theme
+theme_set(theme_bw())
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
@@ -26,8 +31,10 @@ count = d %>%
   group_by(workerid) %>%
   tally(sort=T)
 count
-View(count)
+#View(count)
 # 5 Turkers did the experiment more than once, namely 16 times together
+# workerid 26, 16, 44, 55, 119
+
 
 # remove data from Turkers who took the experiment more than once 
 workers = d %>%
@@ -36,7 +43,7 @@ workers = d %>%
   filter(n == 52) %>%
   droplevels()
 workers
-View(workers)
+#View(workers)
 
 d = d %>%
   filter(workerid %in% as.character(workers$workerid))
@@ -60,7 +67,7 @@ workers = ds %>%
   filter(n < 2) %>%
   droplevels()
 workers
-View(workers)
+#View(workers)
 
 ds = ds %>%
   filter(workerid %in% as.character(workers$workerid))
@@ -73,12 +80,14 @@ d = d %>%
   left_join(ds, by=c("workerid"))
 
 # age info (for all 284 unique Turkers that participated)
-median(d$age) #35
+table(d$age) #19-74
 mean(d$age) #38
 ggplot(d, aes(x=age)) +
   geom_histogram()
 
 nrow(d) #14768 / 52 = 284 
+
+# no gender information available
 
 # change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
 # by subtracting the ai responses from 1
@@ -91,6 +100,8 @@ d$trial = d$slide_number_in_experiment - 4
 unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
+
+## exclude participants' data ----
 
 ### exclude non-American English speakers
[…]

### 2022-01-31 results/main/2_projaiN/rscripts/preprocessing.R

@@ -1,22 +1,32 @@
-# negation: projection and at-issueness
+# 2_projaiN
+# Exp 1n
+# projection and at-issueness (sure that)
 # preprocessing
 
 # set working directory to directory of script
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
+# load required packages
+library(tidyverse)
+
+# load helpers
 source('../../helpers.R')
 
-
-# load required packages for pre-processing data
-library(tidyverse)
+# set theme
 theme_set(theme_bw())
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
+nrow(d) #15600 / 300 = 52 trials
+head(d)
+summary(d)
 
 # read in the subject information
 ds = read_csv("../data/experiment-subject_information.csv")
+nrow(ds) #300
+head(ds)
+summary(ds)
 
 # merge subject information into data
 d = d %>%
@@ -31,7 +41,7 @@ unique(ds$comments)
 # participant info
 table(d$age) #18-74
 length(which(is.na(d$age))) # 0 missing values
-median(d$age,na.rm=TRUE) #31
+mean(d$age,na.rm=TRUE) #33.2
 
 d %>% 
   select(gender, workerid) %>% 
@@ -52,11 +62,12 @@ unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
 
+## exclude participants' data ----
+
 ### exclude non-English speakers and non-American English speakers
 # exclude non-English speakers
 length(which(is.na(ds$language))) #no missing responses
-table(ds$language) 
-
+table(d$language) 
 
 d <- d %>%
   filter(language != "Spanish" & language != "Vietnamese") %>%  droplevels()
@@ -148,37 +159,46 @@ d <- d %>%
   droplevels()
 length(unique(d$workerid)) # 275 remaining Turkers (17 Turkers excluded)
 
-# exclude turkers who always clicked on roughly the same point on the scale 
-# ie turkers whose variance in overall response distribution is more 
+# variance
+
+# exclude participants who always clicked on roughly the same point on the scale 
+# ie participants whose variance in overall response distribution is more 
 # than 2 sd below mean by-participant variance
+table(d$trigger)
[…]

### 2022-01-31 results/main/3_projaiM/rscripts/preprocessing.R

@@ -1,23 +1,32 @@
-# modal: projection and at-issueness
+# 3_projaiM
+# Exp 1m
+# projection and at-issueness (sure that)
 # preprocessing
 
 # set working directory to directory of script
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
+# load required packages
+library(tidyverse)
+
+# load helpers
 source('../../helpers.R')
 
+# set theme
 theme_set(theme_bw())
-
-# load required packages for pre-processing data
-library(tidyverse)
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
-
+nrow(d) #15600 / 300 = 52 trials
+head(d)
+summary(d)
 
 # read in the subject information
 ds = read_csv("../data/experiment-subject_information.csv")
+nrow(ds) #300
+head(ds)
+summary(ds)
 
 # merge subject information into data
 d = d %>%
@@ -32,7 +41,7 @@ unique(ds$comments)
 # participant info
 table(d$age) #18-74
 length(which(is.na(d$age))) # 0 missing values
-median(d$age,na.rm=TRUE) #30
+mean(d$age,na.rm=TRUE) #32.7
 
 d %>% 
   select(gender, workerid) %>% 
@@ -52,6 +61,8 @@ d$trial = d$slide_number_in_experiment - 4
 unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
+
+## exclude participants' data ----
 
 ### exclude non-English speakers and non-American English speakers
 # exclude non-English speakers
@@ -147,35 +158,46 @@ d <- d %>%
   droplevels()
 length(unique(d$workerid)) #281 remaining Turkers (19 Turkers excluded)
 
-# exclude turkers who always clicked on roughly the same point on the scale 
-# ie turkers whose variance in overall response distribution is more 
+# variance
+
+# exclude participants who always clicked on roughly the same point on the scale 
+# ie participants whose variance in overall response distribution is more 
 # than 2 sd below mean by-participant variance
+table(d$trigger)
+table(d$question_type)
+
 variances = d %>%
-  filter(short_trigger != "MC") %>%
+  filter(trigger != "MC") %>%
   group_by(workerid) %>%
[…]

### 2022-01-31 results/main/4_projaiC/rscripts/preprocessing.R

@@ -1,25 +1,36 @@
-# conditional: projection and at-issueness
+# 4_projaiC
+# Exp 1c
+# projection and at-issueness (sure that)
 # preprocessing
 
 # set working directory to directory of script
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
+# load required packages
+library(tidyverse)
+
+# load helpers
 source('../../helpers.R')
 
+# set theme
 theme_set(theme_bw())
-
-# load required packages for pre-processing data
-library(tidyverse)
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
-
+nrow(d) #15600 / 300 = 52 trials
+head(d)
+summary(d)
 
 # read in the subject information
 ds = read_csv("../data/experiment-subject_information.csv")
+nrow(ds)
 
-#remove comments column (something in it creates a problem)
+# look at Turkers' comments
+summary(ds)
+unique(ds$comments)
+
+# remove comments column (something in it creates a problem)
 #ds <- ds %>% select(-comments)
 #names(ds)
 
@@ -30,21 +41,21 @@ d = d %>%
 # how many participants?
 length(unique(d$workerid)) #300
 
-# look at Turkers' comments
-summary(ds)
-unique(ds$comments)
-
 # participant info
-table(d$age) #18-79
-length(which(is.na(d$age))) # 0 missing values
-median(d$age,na.rm=TRUE) #32
+table(d$age) #18-58 
+
+# remove wacky ages from consideration
+d <-  d %>% mutate(age = replace(age, age>150, NA))
+
+length(which(is.na(d$age))) # 104 missing values
+mean(d$age,na.rm=TRUE) #25.9
 
 d %>% 
   select(gender, workerid) %>% 
   unique() %>% 
   group_by(gender) %>% 
   summarize(count=n())
-#158 female, 135 male, 6 other, 1 undeclared
+#249 female, 45 male, 6 other, 0 undeclared
 
 # change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
 # by subtracting the ai responses from 1
@@ -58,25 +69,26 @@ unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
 
+## exclude participants' data ----
+
 ### exclude non-English speakers and non-American English speakers
[…]

### 2022-01-31 results/main/5_projaiQ/rscripts/preprocessing.R

@@ -1,4 +1,5 @@
 # 5_projaiQ
+# Exp 2q
 # projection and at-issueness (assent with positive continuation)
 # preprocessing
 
@@ -6,11 +7,14 @@
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
-source('../../helpers.R')
-theme_set(theme_bw())
+# load required packages
+library(tidyverse)
 
-# load required packages for pre-processing data
-require(tidyverse)
+# load helpers
+source('../../helpers.R')
+
+# set theme
+theme_set(theme_bw())
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
@@ -37,7 +41,7 @@ d = d %>%
 # participant info
 table(d$age) #18-58
 length(which(is.na(d$age))) # 0 missing values
-median(d$age,na.rm=TRUE) #23
+mean(d$age,na.rm=TRUE) #25.5
 
 d %>% 
   select(gender, workerid) %>% 
@@ -57,6 +61,8 @@ d$trial = d$slide_number_in_experiment - 4
 unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
+
+## exclude participants' data ----
 
 ### exclude non-American English speakers
 length(unique(d$workerid)) #250
@@ -156,5 +162,48 @@ d <- d %>%
   droplevels()
 length(unique(d$workerid)) # 221 remaining participants (24 participants excluded)
 
-# write cleaned dataset to file
-write_csv(d, path="../data/data_preprocessed.csv")
+# variance
+
+# exclude participants who always clicked on roughly the same point on the scale 
+# ie participants whose variance in overall response distribution is more 
+# than 2 sd below mean by-participant variance
+table(d$trigger)
+table(d$question_type)
+
+variances = d %>%
+  filter(trigger != "MC") %>%
+  group_by(workerid) %>%
+  summarize(Variance = var(response)) %>%
+  mutate(TooSmall = Variance < mean(Variance) - 2*sd(Variance))
+
+lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
+summary(variances)
+lowvarworkers # 3 participants had lower mean variance
+
+lvw = d %>%
[…]

### 2022-01-31 results/main/7_projaiM/rscripts/preprocessing.R

@@ -1,4 +1,5 @@
 # 7_projaiM
+# Exp 2m
 # projection and at-issueness (assent with positive continuation)
 # preprocessing
 
@@ -6,11 +7,14 @@
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
-source('../../helpers.R')
-theme_set(theme_bw())
+# load required packages
+library(tidyverse)
 
-# load required packages for pre-processing data
-require(tidyverse)
+# load helpers
+source('../../helpers.R')
+
+# set theme
+theme_set(theme_bw())
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
@@ -37,7 +41,7 @@ d = d %>%
 # participant info
 table(d$age) #18-74
 length(which(is.na(d$age))) # 52 missing values
-median(d$age,na.rm=TRUE) #29
+mean(d$age,na.rm=TRUE) #31.7
 
 d %>% 
   select(gender, workerid) %>% 
@@ -57,6 +61,8 @@ d$trial = d$slide_number_in_experiment - 4
 unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
+
+## exclude participants' data ----
 
 ### exclude non-American English speakers
 length(unique(d$workerid)) #251
@@ -156,5 +162,47 @@ d <- d %>%
   droplevels()
 length(unique(d$workerid)) # 220 remaining participants (27 participants excluded)
 
-# write cleaned dataset to file
-write_csv(d, path="../data/data_preprocessed.csv")
+# variance
+
+# exclude participants who always clicked on roughly the same point on the scale 
+# ie participants whose variance in overall response distribution is more 
+# than 2 sd below mean by-participant variance
+table(d$trigger)
+table(d$question_type)
+
+variances = d %>%
+  filter(trigger != "MC") %>%
+  group_by(workerid) %>%
+  summarize(Variance = var(response)) %>%
+  mutate(TooSmall = Variance < mean(Variance) - 2*sd(Variance))
+
+lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
+summary(variances)
+lowvarworkers # 0 participants had lower mean variance
+
+lvw = d %>%
[…]

### 2022-01-31 results/main/8_projaiC/rscripts/preprocessing.R

@@ -1,4 +1,5 @@
 # 8_projaiC
+# Exp 2c
 # projection and at-issueness (assent with positive continuation)
 # preprocessing
 
@@ -6,11 +7,14 @@
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
-source('../../helpers.R')
-theme_set(theme_bw())
+# load required packages
+library(tidyverse)
 
-# load required packages for pre-processing data
-require(tidyverse)
+# load helpers
+source('../../helpers.R')
+
+# set theme
+theme_set(theme_bw())
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
@@ -40,7 +44,7 @@ length(which(is.na(d$age))) # 0 missing values
 
 # turn non-numeric value into NA
 d$age <- as.numeric(as.character(d$age))
-median(d$age,na.rm=TRUE) #22
+mean(d$age,na.rm=TRUE) #22
 
 d %>% 
   select(gender, workerid) %>% 
@@ -60,6 +64,8 @@ d$trial = d$slide_number_in_experiment - 4
 unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
+
+## exclude participants' data ----
 
 ### exclude non-American English speakers
 length(unique(d$workerid)) #250
@@ -159,5 +165,47 @@ d <- d %>%
   droplevels()
 length(unique(d$workerid)) # 224 remaining participants (26 participants excluded)
 
-# write cleaned dataset to file
-write_csv(d, path="../data/data_preprocessed.csv")
+# variance
+
+# exclude participants who always clicked on roughly the same point on the scale 
+# ie participants whose variance in overall response distribution is more 
+# than 2 sd below mean by-participant variance
+table(d$trigger)
+table(d$question_type)
+
+variances = d %>%
+  filter(trigger != "MC") %>%
+  group_by(workerid) %>%
+  summarize(Variance = var(response)) %>%
+  mutate(TooSmall = Variance < mean(Variance) - 2*sd(Variance))
+
+lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
+summary(variances)
+lowvarworkers # 0 participants had lower mean variance
+
+lvw = d %>%
[…]

### 2022-01-31 results/main/9_projaiQ/rscripts/preprocessing.R

@@ -1,4 +1,5 @@
 # 9_projaiQ
+# Exp 3q
 # projection and at-issueness (assent with adversative continuation)
 # preprocessing
 
@@ -6,11 +7,14 @@
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
-source('../../helpers.R')
-theme_set(theme_bw())
+# load required packages
+library(tidyverse)
 
-# load required packages for pre-processing data
-require(tidyverse)
+# load helpers
+source('../../helpers.R')
+
+# set theme
+theme_set(theme_bw())
 
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
@@ -37,7 +41,7 @@ d = d %>%
 # participant info
 table(d$age) #18-66
 length(which(is.na(d$age))) #52 missing values
-median(d$age,na.rm=TRUE) #30
+mean(d$age,na.rm=TRUE) #32.4
 
 d %>% 
   select(gender, workerid) %>% 
@@ -57,6 +61,8 @@ d$trial = d$slide_number_in_experiment - 4
 unique(d$trial) # trial numbers from 1 to 53 (27 missing because instruction)
 d[d$trial > 26,]$trial = d[d$trial > 26,]$trial - 1
 unique(d$trial) # trials from 1 to 52
+
+## exclude participants' data ----
 
 ### exclude non-American English speakers
 length(unique(d$workerid)) #250
@@ -156,5 +162,48 @@ d <- d %>%
   droplevels()
 length(unique(d$workerid)) # 225 remaining participants (20 participants excluded)
 
-# write cleaned dataset to file
-write_csv(d, path="../data/data_preprocessed.csv")
+# variance
+
+# exclude participants who always clicked on roughly the same point on the scale 
+# ie participants whose variance in overall response distribution is more 
+# than 2 sd below mean by-participant variance
+table(d$trigger)
+table(d$question_type)
+
+variances = d %>%
+  filter(trigger != "MC") %>%
+  group_by(workerid) %>%
+  summarize(Variance = var(response)) %>%
+  mutate(TooSmall = Variance < mean(Variance) - 2*sd(Variance))
+
+lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
+summary(variances)
+lowvarworkers # 2 participants had lower mean variance
+
+lvw = d %>%
[…]

### 2022-02-04 writing/investigating-GPP.tex
updated paper

@@
         \centering
-\fbox{\includegraphics[width=7.55cm]{figures/exp1q-projection-trial}}
-\caption{Target trial in projection block of Exp.~1q.}\label{fig-exp1q-projection}
+\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/2q-proj}}
+\caption{Exps.~1q, 2q, and 3q.}\label{fig-exp1q-projection}
 \end{subfigure}%
@@
 \centering
-\fbox{\includegraphics[width=7.9cm]{figures/exp1n-projection-trial}} 
-\caption{Target trial in projection block of Exp.~1n.}\label{fig-exp1n-projection}
+\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/1n-proj}} 
+\caption{Exps.~1n, 2n, and 3n.}\label{fig-exp1n-projection}
  \end{subfigure}
@@
         \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/exp1m-projection-trial}}
-\caption{Target trial in projection block of Exp.~1m}\label{fig-exp1m-projection}
+\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/1m-proj}}
+\caption{Exps.~1m, 2m, and 3m}\label{fig-exp1m-projection}
  \end{subfigure}%
@@
 \centering
-\fbox{\includegraphics[height=5.4cm,width=7.9cm]{figures/exp1c-projection-trial}} 
-\caption{Target trial in projection block of Exp.~1c}\label{fig-exp1c-projection}
+\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/1c-proj}} 
+\caption{Exps.~1c, 2c, and 3c}\label{fig-exp1c-projection}
 \end{subfigure}
 
-\caption{Target trials in projection blocks of Exps.~1.}\label{f-projection-trials}
+\caption{Target trials in projection blocks of Exps.~1, 2, and 3 for the complement {\em Julian dances salsa}.}\label{f-projection-trials}
 \end{figure}
 
-The at-issueness diagnostics differed between Exps.~1, 2, and 3: at-issueness was measured in Exp.~1q with the `asking whether' diagnostic, with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, with the `yes' diagnostic in Exps.~2q and 3q, with the `assent with positive continuation' diagnostic in Exps.~2n, 2m, and 2c,  and with the `assent with adversative continuation' diagnostic in Exps.~3n, 3m, and 3c.
+The at-issueness diagnostics differed between Exps.~1, 2, and 3: at-issueness was measured in Exp.~1q, 2q, and 3q, with variants of the question-based diagnostic of at-issueness, namely the `asking whether' diagnostic in Exp.~1q, and the `yes, $p$' diagnostic in Exps.~2q and 3q, as shown in Figure \ref{fig-ai-trials-q}.
 
-, as shown in Figure \ref{f-ai-trials}.
+, with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, with the `yes' diagnostic in Exps.~2q and 3q, with the `assent with positive continuation' diagnostic in Exps.~2n, 2m, and 2c,  and with the `assent with adversative continuation' diagnostic in Exps.~3n, 3m, and 3c.
 
@@
 Block order and within-block trial order was randomized.
+
+\begin{figure}[h!]
+\centering
+
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1q-ai}}
+\caption{Exp.~1q.}\label{fig-exp1q-ai}
+\end{subfigure}
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2q-ai}} 
+\caption{Exp.~2q.}\label{fig-exp2q-ai}
+ \end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3q-ai}}
+\caption{Exp.~3q}\label{fig-exp3q-ai}
+ \end{subfigure}
+
+\caption{Target trials in at-issueness blocks of Exps.~1q, 2q, and 3q for the complement {\em Julian dances salsa}.}\label{f-ai-trialsq}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2n-ai}}
+\caption{Exp.~2n.}\label{fig-exp1q-ai}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2m-ai}} 
+\caption{Exp.~2m.}\label{fig-exp2q-ai}
+ \end{subfigure}
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=9cm]{figures/2c-ai}}
+\caption{Exp.~2c}\label{fig-exp3q-ai}
+ \end{subfigure}
+
+\caption{Target trials in at-issueness blocks of Exps.~2n, 2m, and 2c for the complement {\em Julian dances salsa}.}\label{f-ai-trialsq}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3n-ai}}
+\caption{Exp.~3n.}\label{fig-exp1q-ai}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3m-ai}} 
+\caption{Exp.~3m.}\label{fig-exp2q-ai}
+ \end{subfigure}
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=9cm]{figures/3c-ai}}
+\caption{Exp.~3c}\label{fig-exp3q-ai}
+ \end{subfigure}
+
+\caption{Target trials in at-issueness blocks of Exps.~3n, 3m, and 3c for the complement {\em Julian dances salsa}.}\label{f-ai-trialsq}
+\end{figure}
 
@@
 They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1), as shown in Figure \ref{fig-trial-exp1}.
-
-\begin{figure}[h!]
-\begin{center}
-\fbox{\includegraphics[width=10cm]{Language-figures/color/Figure1}}
-\end{center}
-\caption{A sample trial in Experiment 1a}\label{fig-trial-exp1}
-\end{figure}
 
@@
 
-\section{Control stimuli in Exps.~1-3}\label{a-control}
+\section{Control stimuli in the projection blocks of Exps.~1-3}\label{a-control}
 
-The control stimuli in Exps.~1-3 were formed from the sentences in (1).
+The control stimuli in the projection blocks of Exps.~1-3 were formed from the sentences in (1).
 In Exps.~1, we used the polar question variants of the sentences in (1) in Exp.~1q, and the (positive declarative) sentences in (1) in Exps.~1n, 1m, and 1c; in Exp.~1c, we added the non-restrictive relative clauses (NRRCs) provided in (1) to the respective subjects, so that the control stimuli have two clauses, like the target stimuli.
-
-We used positive declarative control sentences in Exps.~1n, 1m, and 1c so that we could have clear expectations about participants' responses.
-Thus, we expected participants to give low responses in the `certain that' and `asking whether' diagnostic in Exp.~1q (indicating that the main clause content does not project and is at-issue), and we expected participants to give low responses in the `certain that' diagnostic and high responses in the `sure that' diagnostic in Exps.~1n, 1m, and 1c (indicating, again, that the main clause content does not project and is at-issue).
 
@@
 \end{exe}
+We expected participants to give low responses in the `certain that'  diagnostic  in Exp.~1q (indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question) and high responses in the `certain that'  diagnostic in Exps.~1n, 1m, and 1c (indicating that the speaker is certain of the main clause content).
+These expectations were borne out, as shown in the first three rows of Table \ref{t-proj-controls}.
+
+%We used positive declarative control sentences in Exps.~1n, 1m, and 1c so that we could have clear expectations about participants' responses.
+Thus, we expected participants to give low responses in the `certain that' and `asking whether' diagnostic in Exp.~1q (indicating that the main clause content does not project and is at-issue), and we expected participants to give low responses in the `certain that' diagnostic and high responses in the `sure that' diagnostic in Exps.~1n, 1m, and 1c (indicating, again, that the main clause content does not project and is at-issue).
 
@@
 
-We had the same expectations as for the control stimuli in Exps.~1: low responses in the `certain that' and `asking whether' diagnostic in Exps.~2q and 3q (indicating that the main clause content does not project and is at-issue), and low responses in the `certain that' diagnostic and high responses in the `sure that' diagnostic in the remaining experiments (indicating, again, that the main clause content does not project and is at-issue).
+We had the same expectations as for the control stimuli in Exps.~1: low responses in the `certain that' diagnostic in Exps.~2q and 3q (indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question), and high responses in the `certain that' diagnostic in the remaining experiments (indicating, again, that the speaker is certain of the main clause content).
+These expectations were borne out, as shown in the lower nine rows of Table \ref{t-proj-controls}.
 
-{\bf Provide information on whether these expectations were met -- THEY WERE NOT ALWAYS}
+\begin{table}[h!]
+\centering
+\begin{tabular}{r r}
+Experiment & mean certainty rating \\ 
+\hline
+1q & .14 \\
+2q & .18 \\
+3q & .17 \\
+\hline
+1n & .95 \\
+1m & .96 \\
+1c & .94 \\
+\hline
+2n & .96 \\
+2m & .96 \\
+2c & .96 \\
+\hline
+3n & .94 \\
+3m & .93 \\
+3c & .93 \\
+\hline
+\end{tabular}
+\caption{Mean certainty ratings for control stimuli, for self-declared American English participants}\label{t-proj-controls}
+\end{table}
 
 \newpage
+
+\section{Ratings for main clause contents in at-issueness blocks in Exps.~1-3}\label{a-control-ai}
+
+In the at-issueness blocks of Exps.~1-3, we collected at-issueness ratings for the main clause contents of the control stimuli of the projection blocks described in Supplement \ref{a-control}.
+Originally, the intent was to exclude participants' data on the basis of their at-issueness ratings, in parallel to their certainty ratings.
+We expected the main clause content to be at-issue across the diagnostics, which means that we expected low responses across all at-issueness diagnostics (recall that participants' responses were coded so that the higher a participants' response, the more not-at-issue the content was hypothesized to be, to investigate whether there is a positive correlation between projection and not-at-issueness).
+As shown in Table \ref{t-ai-controls}, these expectations were not always borne out, which is why we decided to not use participants' ratings of the main clause contents in the at-issueness blocks as an exclusion criterion.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{r l}
+Experiment & mean rating \\ 
+\hline
+1q & .05 \\
+2q & .07 \\
+3q & .28 \\
+\hline
+1n & .04 \\
+1m & .03 \\
+1c & .08 \\
+\hline
+2n & .22 (16 participants excluded by specified exclusion criterion) \\
+2m & .25 \\
+2c & .22 \\
+\hline
+3n & .44  (nobody excluded by specified exclusion criterion) \\
+3m & .5 \\
+3c & .53 \\
+\hline
+\end{tabular}
+\caption{Mean ratings on at-issueness diagnostics for main clause contents in Exps.~1-3, for self-declared American English participants}\label{t-ai-controls}
+\end{table}
 

### 2022-02-04 results/main/10_projaiN/rscripts/preprocessing.R

@@ -142,13 +142,13 @@ ggplot(ai.means, aes(x=workerid,y=Mean)) +
 # look at participants whose response mean on projection and ainess of main clauses is more than 2
 # standard deviations away from the overall mean
 
-# get the participants who are more than 2 standard deviations above the mean on projection 
-p <- p.means[p.means$Mean > (mean(p.means$Mean) + 2*sd(p.means$Mean)),]
-p #0 participants 
+# get the participants who are more than 2 standard deviations below the mean on projection 
+p <- p.means[p.means$Mean < (mean(p.means$Mean) - 2*sd(p.means$Mean)),]
+p #13 participants 
 
 # get the participants who are more than 2 standard deviations above the mean on ai 
 ai <- ai.means[ai.means$Mean > (mean(ai.means$Mean) + 2*sd(ai.means$Mean)),]
-ai #10 participants
+ai #0 participants
 
 # make data subset of just the outliers
 outliers <- d.MC %>%

### 2022-02-04 writing/investigating-GPP.tex
updated paper and preprocessing

@@
 \begin{xlist}
-\ex Is the projection of contents embedded under different entailment-canceling operators positively correlated, as assumed in the semantics/pragmatics literature?
 
@@
 
-\section{Prior literature}\label{s2}
+Previous investigations of the relationship between projection and at-issueness:
+
+\begin{itemize}
+
+\item \citealt{brst-salt10,brst-ar}
+
+\item \citealt{xue-onea11}
+
+\item \citealt{tbd-variability}
+
+\end{itemize}
+
+Limited how?
+(only presuppositions + CCs of factive predicates: \citealt{degen-tonhauser-factive} / few measures of at-issueness / some supporting results, some not / some embedded, other matrix clause)
 
@@
 
+\section{Defining and diagnosing at-issueness}\label{s2}
+
+\begin{itemize}
+
+\item Todor Koev
+
+\item Natasha Korotkova
+
+\item Murray, Anderbois et al
+
+\item \citealt{snider2017,snider-paper}
+
+\item \citealt{syrett-koev2015}
+
+\item \citealt{tonhauser-sula6}
+
+\item Diagnostics for at-issueness are either suitable for polar questions or for declaratives.
+
+\begin{itemize}
+
+\item Polar questions: at-issue content partitions the context set
+
+`asking whether'
+
+`yes, $c$'
+
+`yes, but not-c''
+
+\item Declaratives: at-issue content is the target of anaphora
+
+`sure that'
+
+`yes, that's true, $c$'
+
+`yes, that's true, but not-c''
+
+\end{itemize}
+
+\end{itemize}
+
 \section{Experiments 1-3: Methods}\label{s3}
 
-Exps.~1a and 1b were designed to investigate which of the CCs of the 20 clause-embedding predicates in \ref{pred} are presupposed.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
+Exps.~1-3 were designed to investigate whether not-at-issueness and projection are positively correlated, as predicted by the Gradient Projection Principle, as well as to compare various measures of at-issueness.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
 All experiments were conducted with IRB approval.
@@
 
-The experiments employed the {\sc certain that} diagnostic for projection (see also, e.g.\ \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}): on this diagnostic, participants are presented with an utterance in which the content to be investigated occurs in an entailment-canceling environment, like a polar question, as illustrated in \ref{stim} for the content of the appositive, that Martha's new car is a BMW.\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}
+To measure projection, Exps.~1-3 employed the `certain that' diagnostic for projection, which is suitable to measure projection for both polar questions and declarative sentences (see also, e.g.\ \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of the utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) occurs embedded under an entailment-canceling operator:
 
@@
 
-\ex\label{stim} Patrick asks: {\em Was Martha's new car, a BMW, expensive?}
-
+\ex\label{proj-stim} `certain that' diagnostic for projection
+\begin{xlist}
+\ex Daniel: Did Cole discover that Julian dances salsa?
+\ex Daniel: Cole didn't discover that Julian dances salsa.
+\ex Daniel: Perhaps Cole discovered that Julian dances salsa.
+\ex Daniel: If Cole discovered that Julian dances salsa, Logan will be joyful.
+\end{xlist}
 \end{exe}
-To assess projection, participants are asked whether the speaker is certain of the content; for instance, in \ref{stim}, participants are asked whether Patrick is certain that Martha's new car is a BMW.
-If a participant takes the speaker to be certain of the content, we assume that the content projects, that is, the content is presupposed by the speaker; if a participant does not take the speaker to be certain of the content, we assume that the content does not project, that is, the content is not presupposed.
-
-Following \citealt{tbd-variability}, Exp.~1a implemented the certain that diagnostic with a gradient response scale: participants gave their responses on a slider marked `no' at one end and `yes' at the other.
+To assess projection, participants are asked whether the speaker is certain of the content; for instance, in \ref{proj-stim}, participants are asked whether Daniel is certain that Julian dances salsa.
+If a participant takes the speaker to be certain of the content, we assume that the content projects; if a participant does not take the speaker to be certain of the content, we assume that the content does not project.
+Following \citealt{tbd-variability}, the `certain that' diagnostic was implemented in Exps.~1-3 with a gradient response scale: participants gave their responses on a slider marked `no' at one end and `yes' at the other.
 We assume that the closer to `yes' a participant's response is, the more the speaker is certain of the content, that is, the more projective the content is.
@@
 That is, as in any experiment, the quantity of interest, in this case speaker commitment, is only indirectly measured.} As discussed in \citealt{tbd-variability}, a second interpretation of gradient certainty ratings is that they reflect a participant's degree of belief in the speaker's (binary) commitment to the truth of the content.
-While we adopt the first interpretation in our discussion, we remain agnostic about the interpretation of gradient certainty ratings: either interpretation gives rise to the expectation,  on the definition of factive predicates in \ref{def}a, that certainty ratings for factive predicates are categorically higher than for optionally factive and nonfactive ones.
+While we adopt the first interpretation in our discussion, we remain agnostic about the interpretation of gradient certainty ratings.
 
-Exp.~1b was identical to Exp.~1a but employed a two-alternative forced choice task where participants categorically responded `yes' or `no'.
+At-issueness was measured with different diagnostics across Exps.~1-3. For polar questions (Exps.~1q, 2q, 3q), at-issueness was assessed with the three measures in (\ref{ai-question}):
+
+\begin{exe}
+\ex\label{ai-question} At-issueness diagnostics for polar questions
+\begin{xlist}
+\ex `asking whether $c$' 
+\\ {\bf Christopher:} Did Cole discover that Julian dances salsa?
+\\ Is Christopher asking whether Julian dances salsa?
+
+\ex {\em yes}, $c$
+\\ {\bf Christopher:} Did Cole discover that Julian dances salsa?
+\\ {\bf Sandy:} Yes, Julian dances salsa.
+\\ Does Sandy's response to Christoper sound good?
+
+\ex {\em yes, but} $\neg c'$
+\\ {\bf Christopher:} Did Cole discover that Julian dances salsa?
+\\ {\bf Sandy:} Yes, but Cole didn't discover it. 
+\\ Does Sandy's response to Christoper sound good?
+
+\end{xlist}
+\end{exe}
+
+For declaratives, at-issueness was assessed with the three measures in (\ref{ai-decl}):
+
+\begin{exe}
+\ex\label{ai-decl} At-issueness diagnostics for declaratives
+
+\begin{xlist}
+
+\ex `sure that $c$'
+\\ {\bf Christopher:} Cole didn't discover that Julian dances salsa.
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.
+\\ {\bf Sandy:} Are you sure?
+\\ {\bf Christopher:} Yes, I'm sure that Julian dances salsa.
+\\ Did Christopher answer Sandy's question?
+
+\ex {\em yes, that's true}, $c$
+\\ {\bf Christopher:} Cole didn't discover that Julian dances salsa.
+\\ {\bf Sandy:} Yes, that's true, Julian dances salsa.
+\\ Does Sandy's response to Christopher sound good?
+
+\ex {\em yes, that's true, but} $\neg c'$
+\\ {\bf Christopher:} Cole didn't discover that Julian dances salsa.
+\\ {\bf Sandy:} Yes, that's true, but Cole discovered it. 
+\\ Does Sandy's response to Christopher sound good?
+
+\end{xlist}
+
+\end{exe}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{r | r | r | r}
+
+Exp. & Embedding & Projection measure & At-issueness measure \\ 
+\hline
+1q & question & certain that $c$ & asking whether $c$ \\
+1n & negation & certain that $c$ & sure that $c$ \\
+1m & modal & certain that $c$ & sure that $c$ \\
+1c & conditional & certain that $c$ & sure that $c$ \\
+\hline
+2q & question & certain that $c$ & {\em yes}, $c$ \\ 
+2n & negation & certain that $c$& {\em yes, that's true}, $c$ \\ 
+2m & modal & certain that $c$& {\em yes, that's true}, $c$ \\ 
+2c & conditional & certain that $c$& {\em yes, that's true}, $c$ \\ 
+\hline
+3q & question & certain that $c$& {\em yes}, but $\neg c'$ \\ 
+3n & negation & certain that $c$& {\em yes, that's true}, but $\neg c'$ \\ 
+3m & modal & certain that $c$& {\em yes, that's true}, but $\neg c'$ \\ 
+3c & conditional & certain that $c$& {\em yes, that's true}, but $\neg c'$ \\ 
+\hline
+\end{tabular}
+\caption{Overview of the entailment-canceling embeddings, projection measures, and at-issueness measures in Exps.~1-3, where $c$ is the content under investigation, and $c'$ is a salient, different content.}\label{t-overview}
+\end{table}
+
+{\bf Comparisons for at-issueness:}
+
+\begin{itemize}
+
+\item Across Exps 1q, 2q, 3q (same assumption about at-issue content partitioning context set; reveals effect of embedding)
+
+\item Across Exps n, m, c (same measure of at-issueness in Exps 1, 2, 3; reveals effect of embedding)
+
+\item 2q -- 2n, 2m, 2c (yes...$c$; i.e., anaphor with $c$)
+
+\item 3q -- 3n, 3m, 3c (yes...but $\neg c$, i.e., anaphor with $c$)
+
+\end{itemize}
+
+\footnote{Exps.~1-3 also allow us to address the question of whether the projection of contents embedded under different entailment-canceling operators is positively correlated, as assumed in the semantics/pragmatics literature (e.g., REFERENCES).
+Prior experimental literature: \citealt{smith-hall11,smith-hall-cls,demarneffe-etal-sub23}.
+As shown in Supplement \ref{a-projection}, this is the case.}
 
@@
 
-\subsection{Materials}
+\subsection{Materials and procedure}
 
@@
 
-\subsection{Procedure}
-
 Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask a question.
@@
 
-%After completing the experiment (as well as the other five experiments we report on), participants filled out a short, optional survey about their age, their gender, their native language(s) and, if English is their native language, whether they are a speaker of American English (as opposed to, e.g.\ Australian or Indian English).
-To encourage them to respond truthfully, participants were told that they would be paid no matter what answers they gave in the survey.
-
-After completing the experiment, participants filled out a short optional demographic sur- vey.
+After completing the experiment, participants filled out a short, optional survey about their age, their gender, their native language(s) and, if English is their native language, whether they are a speaker of American English (as opposed to, e.g.\ Australian or Indian English).
 To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.
@@
 We also excluded data from participants based on their ratings on the main clause controls and other criteria given in Supplement \ref{a-participants}.
-In each experiment, the data from at least {\bf FIX} participants were analyzed.
+In each experiment, the data from between 215-281 remaining participants were analyzed.
 Information on the participants whose data entered into the analysis (total number, ages, gender) can be found in Supplement \ref{a-participants}.
@@
 
-\subsection{Comparing projection}
-
-\subsection{Comparing at-isueness}
+\subsection{Comparing at-issueness}
 
@@
 \item look at how many unique participants
-
-\item compare projection across experiments/embeddings
-
-compare projection across embeddings: research assumes that there are no differences, but there is some experimental evidence that there may be (Kathleen/Elizabeth, CommitmentBank), so it?s good to check (compare 4 experiments)
 
@@
 
-\section{Consequents for target stimuli in Exps.~1c, 2c, and 3c}\label{a-target}
+\section{Consequents for conditional target stimuli in Exps.~1c, 2c, and 3c}\label{a-target}
 
-10 positive valuence, 10 negative valuence {\bf which other considerations?}
+The consequents for the conditional target stimuli in Exps.~1c, 2c, and 3c were created with the following considerations in mind.
+Each complement clause was paired with a unique consequent clause, as shown in the list below for the 20 complement clauses.
+Each consequent clause consisted of a uniquely named subject and an adjectival predication in the future tense ({\em will be}); the adjectives all denoted an emotion.
+We selected the 20 emotion-denoting adjectives based on the following criteria: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5 (based on the valence and arousal values reported in \citealt{warriner-etal2013}).
 
@@
 
-{\bf Include information on payment}
+This supplement provides information on the recruited participants, the criteria by which participants' data were excluded, the remaining participants (that is, the participants whose data entered into the analysis), and the payment for each of Exps.~1-3.
+Table \ref{f-exclusion} provides information on the recruited and remaining participants, including the total number, the range of the participants' ages, the mean ages of the participants, and their self-reported gender (`f' = female, `m' = male, `o' = other, `u' = undeclared).
+No gender data was collected in Exp.~1q.
+Participants' data were excluded based on the following criteria:
 
-Gender: no data collected in Exp 1q
-
-Relevant manual inspection on `variance': Exp 1q, Exp 2q, Exp 3q (participants automatically identified, but not all excluded after manual inspection)
-
-Table \ref{f-exclusion} presents how many participants' data were excluded from the analysis based on the exclusion criteria.
-The first column records the experiment, the second (`recruited') how many participants were recruited, and the final column (`remaining') how many participants' data entered the analysis.
-The `Exclusion criteria' columns show how many participants' data were excluded based on the four exclusion criteria:
-
-{\bf comment on why so many women}
-
-\begin{itemize}[topsep = -1ex,itemsep=-2pt]
+\begin{itemize}[itemsep=-2pt]
 
@@
 
-\item `controls': Participants' data were excluded if their response mean on the 6 control items was more than 2 sd above the group mean (Exp.~1a), if they gave a wrong rating (`yes') to more than one of the six controls (Exp.~1b), if their response mean on the entailing or the non-entailing controls was more than 2 sd below or above, respectively, the group mean (Exp.~2a), if they gave more than one wrong rating to one of the eight controls, where a wrong rating is a `yes' to a non-entailing control and a `no' to an entailing one (Exp.~2b), if their response means on the contradictory or non-contradictory controls were more than 2 sd below or above, respectively, the group mean (Exp.~3a), and if they gave more than one wrong response to one of the eight control sentences, where a wrong response was a `yes' to a non-contradictory control or a `no' to a contradictory one (Exp.~3b).
+\item `controls': Participants' data were excluded if their mean rating on the 6 main clause control items in the projection block was more than 2 sd above the group mean (in Exps.~1q, 2q, and 3q) or more than 2 sd below the group mean (in the remaining experiments).
+Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across Exps.~1-3).
 
-\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the non-control trials.
-We identified such participants by calculating the variance of the response distribution for each participant, and manually inspecting those whose variance was more than 2 sd below the group mean variance.
-
+\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli.
+We identified such participants by identifying participants whose mean variance on the target stimuli was more than 2sd below the group mean variance on the target stimuli, and manually inspecting their response patterns.
+%Exps 1q, 2q, 3q: participants with low variance not excluded after manual inspection
 \end{itemize}
 
-{\bf Exp 3n, 3m, 3c: need to look at not-at-issue ratings on controls to see how to exclude}
+Participants took around 9-11 minutes to complete the various experiments.
+Participants were paid more in Exps.~1c, 2c, and 3c than the remaining experiments because the target stimuli in those experiments were longer (as they consisted of conditionals).
+More women than men were recruited in many of the experiments because they were run at a time when Prolific apparently went viral on TikTok, resulting in a large number of young women registering for the service (around July 24, 2021; see \url{https://blog.prolific.co/we-recently-went-viral-on-tiktok-heres-what-we-learned/}, \\ last accessed February 4, 2022).
 
-\begin{table}[h!]
+\begin{sidewaystable}[h!]
 \centering
-\begin{tabular}{l | r r r | r r r r | r r r }
-&  \multicolumn{3}{c|}{Recruited participants} & \multicolumn{4}{c|}{Exclusion criteria} & \multicolumn{3}{c}{Remaining participants}  \\ 
-Exp. & total & ages (mean) & f/m/o/u & mult & lang & cntr & var & total & ages (mean) & f/m/o/u  \\ 
+\begin{tabular}{l | r r r | r r r r | r r r | r }
+&  \multicolumn{3}{c|}{Recruited participants} & \multicolumn{4}{c|}{Exclusion criteria} & \multicolumn{3}{c|}{Remaining participants} &  \\ 
+Exp. & total & ages (mean) & f/m/o/u & multiple & language & controls & variance & total & ages (mean) & f/m/o/u & payment   \\ 
 \hline
-1q & 300  & 19-74 (38.2)  & --  & 5  & 7  & 35  & 0  & 242   & 21-74 (39.2) &  -- \\
-1n & 300  & 18-74 (33.2)  &  150/145/5/0 & 0  & 8 & 17 & 1 & 274  & 18-74 (33.3) & 141/128/5/0 \\
-1m & 300  & 18-74 (32.7) & 150/141/7/2  &  0 & 0  & 19 & 0   & 281   & 18-74 (32.7) & 144/129/7/1 \\
-1c &  300 & 18-58 (25.9)  & 249/45/6/0 & 0 & 6  & 26 & 2  & 266  &  18-58 (24.8) & 235/25/6/0  \\
-2q & 250  &  18-58 (25.5) &  201/43/6/0  & 0 & 4  & 24  & 1 & 220  & 18-58 (24.8)  & 187/28/5/0  \\
-2n & 250  &  18-69 (33.2)  &  127/114/6/1 & 1  & 4  & 29 & 0& 215  &  18-69 (33.1) & 113/95/6/1    \\
-2m & 251  & 18-74 (31.7)  & 132/113/6/0  & 0  & 4  & 27 &  0 & 220  & 18-70 (31.9) & 116/98/6/0 \\
-2c &  250 &  18-56 (24.5)  & 212/30/8/0  & 0  & 0  & 26 & 0  & 224  & 18-56 (24.4) & 195/24/5/0 \\
-3q & 250  &  18-66 (32.4) &  140/102/7/1 &  0 & 4  & 20  & 0  &  225 & 18-66 (32.6) & 125/93//7/0  \\
-3n & 250  &  18-70 (24.6) & 114/31/5/0  & 0  & 5  &  &   &   &  &  \\
-3m & 250  & 18-63 (25.5)  & 205/40/5/0 &  0 & 3  &  &   &   &  &  \\
-3c & 250  & 18-59  (27.5) & 182/64/4/0  & 0  & 3   &  &   &   &  &  \\
+1q & 300  & 19-74 (38.2)  & --  & 5  & 7  & 35  & 0  & 242   & 21-74 (39.2) &  -- & \\
+1n & 300  & 18-74 (33.2)  &  150/145/5/0 & 0  & 8 & 17 & 1 & 274  & 18-74 (33.3) & 141/128/5/0 & \$1.70\\
+1m & 300  & 18-74 (32.7) & 150/141/7/2  &  0 & 0  & 19 & 0   & 281   & 18-74 (32.7) & 144/129/7/1 & \$1.70 \\
+1c &  300 & 18-58 (25.9)  & 249/45/6/0 & 0 & 6  & 26 & 2  & 266  &  18-58 (24.8) & 235/25/6/0  & \$1.70 \\
+2q & 250  &  18-58 (25.5) &  201/43/6/0  & 0 & 4  & 24  & 1 & 220  & 18-58 (24.8)  & 187/28/5/0  & \$2.15 \\
+2n & 250  &  18-69 (33.2)  &  127/114/6/1 & 1  & 4  & 29 & 0& 215  &  18-69 (33.1) & 113/95/6/1    & \$1.70\\
+2m & 251  & 18-74 (31.7)  & 132/113/6/0  & 0  & 4  & 27 &  0 & 220  & 18-70 (31.9) & 116/98/6/0 & \$1.70\\
+2c &  250 &  18-56 (24.5)  & 212/30/8/0  & 0  & 0  & 26 & 0  & 224  & 18-56 (24.4) & 195/24/5/0 & \$2.15\\
+3q & 250  &  18-66 (32.4) &  140/102/7/1 &  0 & 4  & 20  & 0  &  225 & 18-66 (32.6) & 125/93//7/0  & \$1.70 \\
+3n & 250  &  18-70 (24.6) & 114/31/5/0  & 0  & 5  &  13 & 4  & 228  & 18-70 (24.3) &  198/25/5/0 &\$1.70 \\
+3m & 250  & 18-63 (25.5)  & 205/40/5/0 &  0 & 3  & 14 &  0 & 233  &  18-63 (24.8) & 197/31/5/0 &\$1.70 \\
+3c & 250  & 18-59  (27.5) & 182/64/4/0  & 0  & 3  & 17 & 0 & 230  &  18-59 (26.7)  & 177/49/4/0  & \$2.15\\
 \end{tabular}
 \caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3}\label{f-exclusion}
-\end{table}
+\end{sidewaystable}
+
+\section{Comparing projection with different entailment-canceling environments}\label{a-projection}
 

### 2022-02-04 results/main/11_projaiM/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #9.2 minutes
+median(time$time_in_minutes) #8.2 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)
@@ -142,25 +147,25 @@ ggplot(ai.means, aes(x=workerid,y=Mean)) +
 # look at participants whose response mean on projection and ainess of main clauses is more than 2
 # standard deviations away from the overall mean
 
-# get the participants who are more than 2 standard deviations above the mean on projection 
-p <- p.means[p.means$Mean > (mean(p.means$Mean) + 2*sd(p.means$Mean)),]
-p #0 participants 
+# get the participants who are more than 2 standard deviations below the mean on projection 
+p <- p.means[p.means$Mean < (mean(p.means$Mean) - 2*sd(p.means$Mean)),]
+p #14 participants 
 
 # get the participants who are more than 2 standard deviations above the mean on ai 
 ai <- ai.means[ai.means$Mean > (mean(ai.means$Mean) + 2*sd(ai.means$Mean)),]
-ai #10 participants
+ai #0 participants
 
 # make data subset of just the outliers
 outliers <- d.MC %>%
   filter(workerid %in% p$workerid | workerid %in% ai$workerid)
 outliers = droplevels(outliers)
-nrow(outliers) #240 / 12 = 20 outlier participants
+nrow(outliers) #168 / 12 = 14 outlier participants
 
 # exclude all outlier participants identified above
 d <- d %>%
   filter(!(workerid %in% p$workerid | workerid %in% ai$workerid)) %>%
   droplevels()
-length(unique(d$workerid)) # 225 remaining participants (20 participants excluded)
+length(unique(d$workerid)) # 233 remaining participants (20 participants excluded)
 
 # variance
 
@@ -178,7 +183,7 @@ variances = d %>%
 
 lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
 summary(variances)
-lowvarworkers # 3 participants had lower mean variance
+lowvarworkers # 0 participants had lower mean variance
 
 lvw = d %>%
[…]

### 2022-02-04 results/main/12_projaiC/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #10.9 minutes
+median(time$time_in_minutes) #9.3 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)
@@ -144,25 +149,25 @@ ggplot(ai.means, aes(x=workerid,y=Mean)) +
 # look at participants whose response mean on projection and ainess of main clauses is more than 2
 # standard deviations away from the overall mean
 
-# get the participants who are more than 2 standard deviations above the mean on projection 
-p <- p.means[p.means$Mean > (mean(p.means$Mean) + 2*sd(p.means$Mean)),]
-p #0 participants 
+# get the participants who are more than 2 standard deviations below the mean on projection 
+p <- p.means[p.means$Mean < (mean(p.means$Mean) - 2*sd(p.means$Mean)),]
+p #17 participants 
 
 # get the participants who are more than 2 standard deviations above the mean on ai 
 ai <- ai.means[ai.means$Mean > (mean(ai.means$Mean) + 2*sd(ai.means$Mean)),]
-ai #10 participants
+ai #0 participants
 
 # make data subset of just the outliers
 outliers <- d.MC %>%
   filter(workerid %in% p$workerid | workerid %in% ai$workerid)
 outliers = droplevels(outliers)
-nrow(outliers) #240 / 12 = 20 outlier participants
+nrow(outliers) #204 / 12 = 17 outlier participants
 
 # exclude all outlier participants identified above
 d <- d %>%
   filter(!(workerid %in% p$workerid | workerid %in% ai$workerid)) %>%
   droplevels()
-length(unique(d$workerid)) # 225 remaining participants (20 participants excluded)
+length(unique(d$workerid)) # 230 remaining participants (20 participants excluded)
 
 # variance
 
@@ -180,7 +185,7 @@ variances = d %>%
 
 lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
 summary(variances)
-lowvarworkers # 3 participants had lower mean variance
+lowvarworkers # 0 participants had lower mean variance
 
 lvw = d %>%
[…]

### 2022-02-04 results/main/1_projaiQ/rscripts/preprocessing.R

@@ -87,9 +87,7 @@ ggplot(d, aes(x=age)) +
 
 nrow(d) #14768 / 52 = 284 
 
-# no gender information available
-
-# change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
+# no gender information a# change the response for ai condition so that what was 0/not-at-issue is now 1/not-at-issue
 # by subtracting the ai responses from 1
 table(d$question_type,d$response)
 d[d$question_type == "ai",]$response = 1 - d[d$question_type == "ai",]$response
@@ -216,6 +214,11 @@ variances = d %>%
   summarize(Variance = var(response)) %>%
   mutate(TooSmall = Variance < mean(Variance) - 2*sd(Variance))
 
+View(variances)
+
+ggplot(variances,aes(x=workerid,y=Variance)) +
+  geom_point()
+
 lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
 summary(variances)
 lowvarworkers # 3 participants had lower mean variance
@@ -225,8 +228,10 @@ lvw = d %>%
   droplevels() %>%
   mutate(Participant = as.factor(as.character(workerid)))
 
-ggplot(lvw,aes(x=Participant,y=response)) +
+ggplot(lvw,aes(x=Participant,y=response,color=trigger_class)) +
   geom_jitter()
+
+# manual inspection shows that these three participants used the whole scale, so no reason to exclude them
 
 # exclude 0 participant with really low variance 
 #d <- droplevels(subset(d, !(d$workerid %in% lowvarworkers)))

### 2022-02-04 results/main/2_projaiN/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #9.8 minutes
+median(time$time_in_minutes) #8.5 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #15600 / 300 = 52 trials
@@ -140,11 +145,11 @@ ggplot(ai.means, aes(x=workerid,y=Mean)) +
 
 # get the Turkers who are more than 2 standard deviations below the mean on projection 
 p <- p.means[p.means$Mean < (mean(p.means$Mean) - 2*sd(p.means$Mean)),]
-p
+p #11 participants
 
 # get the Turkers who are more than 2 standard deviations above the mean on ai 
 ai <- ai.means[ai.means$Mean > (mean(ai.means$Mean) + 2*sd(ai.means$Mean)),]
-ai
+ai #11 participants
 
 # look at the main clauses that these "outlier" Turkers did
 # make data subset of just the outliers

### 2022-02-04 results/main/3_projaiM/rscripts/preprocessing.R

@@ -139,11 +139,11 @@ ggplot(ai.means, aes(x=workerid,y=Mean)) +
 
 # get the Turkers who are more than 2 standard deviations below the mean on projection 
 p <- p.means[p.means$Mean < (mean(p.means$Mean) - 2*sd(p.means$Mean)),]
-p
+p #14 participants
 
 # get the Turkers who are more than 2 standard deviations above the mean on ai 
 ai <- ai.means[ai.means$Mean > (mean(ai.means$Mean) + 2*sd(ai.means$Mean)),]
-ai
+ai #12 participants
 
 # look at the main clauses that these "outlier" Turkers did
 # make data subset of just the outliers

### 2022-02-04 results/main/4_projaiC/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #10.4 minutes
+median(time$time_in_minutes) #9.1 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #15600 / 300 = 52 trials

### 2022-02-04 results/main/5_projaiQ/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #9.5 minutes
+median(time$time_in_minutes) #8.4 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)
@@ -180,12 +185,17 @@ lowvarworkers = as.character(variances[variances$TooSmall,]$workerid)
 summary(variances)
 lowvarworkers # 3 participants had lower mean variance
 
+#View(variances)
+
+ggplot(variances,aes(x=workerid,y=Variance)) +
+  geom_point()
+
 lvw = d %>%
   filter(as.character(workerid) %in% lowvarworkers) %>%
   droplevels() %>%
   mutate(Participant = as.factor(as.character(workerid)))
 
-ggplot(lvw,aes(x=Participant,y=response)) +
+ggplot(lvw,aes(x=Participant,y=response,color=trigger_class)) +
   geom_jitter()
 
 # exclude 1 participant with really low variance (#1547)

### 2022-02-04 results/main/6_projaiN/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #10.4 minutes
+median(time$time_in_minutes) #8.6 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)

### 2022-02-04 results/main/7_projaiM/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #10.8 minutes
+median(time$time_in_minutes) #8.5 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13052 / 251 = 52 trials (the experiment was done 251 times)

### 2022-02-04 results/main/8_projaiC/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #9.8 minutes
+median(time$time_in_minutes) #8.6 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)

### 2022-02-04 results/main/9_projaiQ/rscripts/preprocessing.R

@@ -16,6 +16,11 @@ source('../../helpers.R')
 # set theme
 theme_set(theme_bw())
 
+# how long did the experiment take?
+time = read_csv("../data/experiment-merged.csv")
+mean(time$time_in_minutes) #9.9 minutes
+median(time$time_in_minutes) #8.8 minutes
+
 # read in the raw data
 d = read_csv("../data/experiment-trials.csv")
 nrow(d) #13000 / 250 = 52 trials (the experiment was done 250 times, as planned)

### 2022-02-28 writing/investigating-GPP.tex
updated literature review

@@
 
-\section{Defining and diagnosing at-issueness}\label{s2}
+\section{Analyzing and diagnosing at-issueness}\label{s2}
+
+\subsection{Formal analyses of at-issueness}
+
+backgrounded content
+
+\begin{exe}
+\ex Formal analyses of at-issueness
+
+\begin{xlist}
+\ex Dimension-based \\ At-issue content interacts with operators
+
+\ex Question-based \\ At-issue content addresses the question under discussion
+
+\ex Anaphora-based \\ At-issue content introduces a discourse referent for subsequent reference
+
+\end{xlist}
+\end{exe}
 
@@
 
-\item Todor Koev
+\item \citealt{potts05}
 
-\item Natasha Korotkova
+\begin{itemize}
 
-\item Murray, Anderbois et al
+\item p.23/Figure 2.1: at-issue entailments vs.
+backgrounded content (conventional presuppositions, CIs, conversationally-triggered presuppositions) vs.
+conversational implicatures (not backgrounded!!) {\bf 3-way distinction}
 
-\item \citealt{snider2017,snider-paper}
+\item p.24: But the supplementary relative who lived in a working-class suburb of Boston plays a secondary role relative to the information conveyed by the main clause.
+The issue is not where the grandmother lived, but rather the fact that the speaker summered with her as a child.
+
+\item p.30: the utterance might be followed in the discourse by ?Hey, she?ll do both!?
+Or it might be preceded by an agreement that if Mary does one, then she does the other.
+The maxims of quality and quantity would then conspire to en- sure that (2.35b) disappeared.
+Thus, the conversational-implicature dimension is a negotiable part of denotations.
+
+\item p.31: a CI is never relativized to the beliefs of an entity other than the speaker.
+But at-issue content certainly is; in Sue wrongly believes Conner got promoted, the at-issue proposition that Conner got promoted is asserted to hold only in Sue?s belief worlds.
+this embedded proposition is not speaker-oriented, and hence not classifiable as a CI contri- bution,
+
+\item p.32:  CIs are independent of the at-issue con- tent.
+In contrast, the fundamental goal of almost all presupposition logics is to create a dependency between the presuppositions and the at-issue entailments
+
+\item p.34: Neither at- issue content nor CI content should be presupposed.
+
+\item p.43: everything about the definition says that CIs contribute new information. In this sense, they are like at-issue content.
+
+\item {\bf can we diagnose not-at-issue content independent of a particular type of not-at-issue content?
+that is, is there a diagnostic for not-at-issueness that applies to all such content, regardless of its specific type?}
+
+\item p.51: discourse-level phenomena. It seems that none are sensitive to the distinction between these two types of content. (at-issue vs CI): VP ellipsis and do it/that pronominalization
+
+\item p.53: Overall, they are an indication that we distinguish the at-issue and CI dimensions in the composition. From the point of view of a discourse as a whole, they are identical.
+
+\item p.54: A similar, and less fraught, test involves discourse referents.
+Because noth- ing can scope over supplementary CI content, it is quite good at establishing discourse referents.
+...
+Here again, presupposed material patterns differently.
+It alone cannot establish a discourse referent, at least not without a significant amount of contextual priming and some creative inferences on the part of the hearer;
+
+\item p.54: Speaker-oriented adverbs, for instance, suggest a new test. Exam- ples (3.16)?(3.17) show that their content can be queried just as at-issue content can.
+
+\item p.54: The upshot of these examples is that CI content is model-theoretically the same as at-issue content.
+...
+Thus, we are led to the con- clusion that the distinction exists only at the level of the meaning language.
+{\bf potts assumes that ai and nai content are not distinct on a discourse level!}
+
+\item p.70: in the CI logic employed here, translation is a nontrivial step, as it is the locus of the at-issue/CI distinction.
+If we interpreted natural language expres- sions directly, we would be left with only two options, both undesirable: we could locate the distinction in the natural language syntax or in the models.
+The syntactic view is initially unpromising given the semantic nature of the defini- tion of CIs, and it receives sustained criticism in chapter 6.
+The model-theoretic option is to locate the differences in the models.
+But section 3.3 shows that in- tersentential phenomena treat the two classes of meaning as model-theoretically identical.
+Moreover, it is hard to see what the distinctions could be.
+
+\end{itemize}
+
+\item \citealt{potts2007-compass}
+
+\begin{itemize}
+
+\item p.666: ``typically the content that speakers offer as primary and also the content that they are most expecting to have to negotiate with their interlocutors before it is accepted into the common ground''
+
+\end{itemize}
+
+\item \citealt{koev2013,koev2018,koev2021}
+
+\begin{itemize}
+
+\item \citealt{koev2018}
+
+\begin{itemize}
+
+\item p.1: At-issue content is that part of the sentence meaning that is intuitively felt to express the ?main point?
+when the sentence is uttered in a given context.
+Potts (2005; 2007) was among the first to use the term ?at-issue content?
+in the relevant sense, characterizing it as content that ?carr[ies] the main themes of a discourse?
+(Potts, 2005, 7), and that speakers are ?most expect- ing to have to negotiate with their interlocutors before it is accepted into the common ground?
+(Potts, 2007, 666).
+The first property, i.e.
+the ability to contribute to the discourse topic, is often expressed by saying that at-issue content is what is relevant to the question under discussion.
+The second property, i.e.
+the ability to represent what is being negotiated among interlocu- tors, is typically captured by saying that at-issue content constitutes a proposal to update the common ground.
+
+\item p.2: There exist several empirical tests for at-issueness, e.g.
+question/answer pairs and accessibility to direct denials or other discourse continuations.
+Exist- ing accounts have employed such tests rather indiscriminately or without elaborating on how these tests relate to a given theoretical construal.
+In order to be able to judge a given theory on its own merits, in each case I will consider only those diagnostics that transparently follow from what counts as at issue in the relevant sense.
+
+\item p.3: q-at-issueness:  A proposition p is Q-atissue relative to a question under discussion Q and a context c iff ? p is relevant to Q in c and
+? p is appropriately conventionally marked relative to Q.
+
+\item p.4: Above, Q-at-issueness was diagnosed in question/answer pairs.
+This is consistent with the question-based definition in 2, provided that an overt question is assumed to manifest the current question under discussion.
+
+\item p.5: several authors identify at-issue content with the proposal introduced by a declarative utterance, i.e.
+with its asserted content (AnderBois et al., 2015; Farkas \& Bruce, 2010; Koev, 2013; Murray, 2014; Potts, 2005).
+{\bf potts 2005 does not assume that CIs are not asserted!}
+
+\item p.5: A proposition p is P-at-issue in a context c iff ? p is a proposal in c and ? p has not been accepted or rejected in c.
+
+\item p.5: to test a proposition for proposalhood, we need a lin- guistic tool that tracks what is being negotiated at any given point.
+One widely used diagnostic to this effect is the assent/dissent test, which probes the amenability of semantic content to direct responses
+
+\item p.6: This test brings up the question of what counts as a direct response.
+One peculiarity about the responses in 10?11 is that they are anaphoric to parts of the target utterance through propositional anaphors like that or no.7 This is important because nonanaphoric denials behave differently: They can felicitously target nominal appositives (see also Hunter & Asher, 2016).
+
+\item p.6: One possible reaction to the contrast in 10?11 v.
+12?13 is to say that the assent/dissent test tracks anaphoric potential rather than P-at-issueness.
+This is reasonable, although it has also been claimed that P-at-issue content obligatorily establishes a propositional discourse referent (Murray, 2014).
+
+\item p.6: I will thus follow the bulk of the literature in assuming that such responses target the assertion and will employ them as the main diagnostic for P-at-issueness, leaving the precise link between anaphoricity and at-issueness to further research.
+
+\item p.7: A third notion of at-issueness is situated within theories which view discourse as structured by coherence relations, also called ?rhetorical?
+or ?discourse?
+relations (Asher \& Lascarides, 2003; Hobbs, 1979, 1985; Kehler, 2002; Mann & Thompson, 1988).
+
+\item p.8:  A proposition expressed by a node ?? of a discourse tree D is C-atissue iff ?? is on the right frontier of D.
+
+\item p.8: A proposition is C-atissue iff a freshly uttered segment can attach to it by some appropriate coherence relation.
+
+\item p.9: Coherence-based accounts thus view the assent/dissent test as a special case of discourse attachment, i.e.
+one in which the response is connected to the target by an Agreement/Denial relation.
+
+\item p.9: Q-at-issueness is primarily a backward-looking notion or relates to existing discourse. By contrast, P-at-issueness and C-at-issueness are forward-looking or relate to what is to come.
+
+\item p.10: I first compare Q-at-issueness and P-at-issueness, arguing that neither property entails the other.
+
+\item p.10: Koev argues that 26 shows that the appositive is not P-at-issue but that 27 shows that it is Q-at-issue: ``appositive relative clauses are often taken into consideration when answering the question under discussion.
+Here, the crucial information that Mary experienced firsthand the terrible consequences of cancer comes from the appositive; the main clause alone does not answer the question, at least not without loss of information.''
+
+(26) A: Mary took care of her husband, who had prostate cancer, for almost a year. B: ? No, he didn't.
+
+(27) Q: Why is Mary fundraising for the upcoming Walk For Cancer?
+A: She took care of her husband, who had prostate cancer, for almost a year.
+
+\item p.10: Relevance implicatures. They are not P-at-issue (as diagnosed by dissent) but Q-at-issue (as diagnosed by Q/A pairs).
+
+\item p.10f: slifting constructions. Can be P-at-issue (dissented with) but don't answer Q (diagnosed by Q/A pairs)
+
+\item p.11:  Simons et al. (2010) and Beaver et al. (2017) claim that projectivity is perfectly correlated with (the lack of) Q-at-issueness {\bf NOT TRUE}.
+
+\item p.12: The Projection Principle then appears to be problematic in both directions: Projecting content can be Q-at issue, and nonprojecting content can be not Q-at issue.
+{\bf Need to engage with Koev 2018 as somebody who doesn't believe in the GPP}
+
+\item p.12: The second observation is that the Projection Principle is stated in terms of Q-at-issueness {\bf NO!} and other notions of at-issueness may lead to different predictions.
+The question of which notion of at-issueness (if any) makes the right predictions about projection is yet to be systematically investigated.
+
+\item endnote 1: I also limit my attention to declarative sentences. The question of how or whether at-issueness interacts with other sentence types will have to await another study.
+
+\item endnote 5: The assent/dissent test is known under different names in the literature and has been used to distinguish asserted content from implications triggered by constructions as different as presupposition markers (Chierchia & McConnell-Ginet, 2000; Karttunen & Peters, 1979; Shanon, 1976; Strawson, 1950), epistemic modals (Lyons, 1977; Papafragou, 2006; von Fintel & Gillies, 2007), evidentials (Faller, 2002; Matthewson et al., 2007; Murray, 2014), and appositives (AnderBois et al., 2015; Koev, 2013; Syrett & Koev, 2015; Tonhauser, 2012).
+
+\item endnote 16: Appositive content typically projects, so it could be argued that it is not ?intended? as being at issue (see Simons et al., 2010).
+However, given its significance in the above context and the fact that conventional marking of at-issueness is often overruled, the information that the director fired Jon would very likely be taken into consideration when answering the question under discussion.
+
+\end{itemize}
+
+\end{itemize}
+
+\item \citealt{amaral-etal2007}
+
+\item \citealt{gutzmann}
+
+\item \citealt{korotkova2020}
+
+\item \citealt{murray2014}
+
+\item \citealt{anderbois-etal2015}
+
+\item \citealt{snider-thesis, snider2017, snider2018}
 
@@
 \item \citealt{tonhauser-sula6}
+
+\item \citealt{brst-salt10}
+
+\item \citealt{tbd-variability}
+
+\item \citealt{brst-ar}
+
+\end{itemize}
+
+\subsection{Diagnosing at-issueness}
+
+\begin{itemize}
 
@@
 \end{itemize}
+
+\subsection{Next steps}
 
@@
 
-Exp. & Embedding & Projection measure & At-issueness measure \\ 
+{\bf Exp.} & {\bf Embedding} & {\bf Projection measure} & {\bf At-issueness measure} \\ 
 \hline
@@
 \end{tabular}
-\caption{Overview of the entailment-canceling embeddings, projection measures, and at-issueness measures in Exps.~1-3, where $c$ is the content under investigation, and $c'$ is a salient, different content.}\label{t-overview}
+\caption{Overview of the entailment-canceling embeddings, projection measures, and at-issueness measures in Exps.~1-3, where $c$ is the content under investigation (in the target trials: the content of the clausal complement) and $c'$ is a different, salient content (in the target trials: the main clause content).}\label{t-overview}
 \end{table}
@@
 \hline
-1q & 300  & 19-74 (38.2)  & --  & 5  & 7  & 35  & 0  & 242   & 21-74 (39.2) &  -- & \\
+1q & 300  & 19-74 (38.2)  & --  & 5  & 7  & 35  & 0  & 242   & 21-74 (39.2) &  -- & \$1.70\\
 1n & 300  & 18-74 (33.2)  &  150/145/5/0 & 0  & 8 & 17 & 1 & 274  & 18-74 (33.3) & 141/128/5/0 & \$1.70\\
 1m & 300  & 18-74 (32.7) & 150/141/7/2  &  0 & 0  & 19 & 0   & 281   & 18-74 (32.7) & 144/129/7/1 & \$1.70 \\
-1c &  300 & 18-58 (25.9)  & 249/45/6/0 & 0 & 6  & 26 & 2  & 266  &  18-58 (24.8) & 235/25/6/0  & \$1.70 \\
-2q & 250  &  18-58 (25.5) &  201/43/6/0  & 0 & 4  & 24  & 1 & 220  & 18-58 (24.8)  & 187/28/5/0  & \$2.15 \\
+1c &  300 & 18-58 (25.9)  & 249/45/6/0 & 0 & 6  & 26 & 2  & 266  &  18-58 (24.8) & 235/25/6/0  & \$2.15 \\
+2q & 250  &  18-58 (25.5) &  201/43/6/0  & 0 & 4  & 24  & 1 & 220  & 18-58 (24.8)  & 187/28/5/0  & \$1.70 \\
 2n & 250  &  18-69 (33.2)  &  127/114/6/1 & 1  & 4  & 29 & 0& 215  &  18-69 (33.1) & 113/95/6/1    & \$1.70\\

### 2022-03-14 writing/investigating-GPP.tex
small change

@@
 Each consequent clause consisted of a uniquely named subject and an adjectival predication in the future tense ({\em will be}); the adjectives all denoted an emotion.
-We selected the 20 emotion-denoting adjectives based on the following criteria: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5 (based on the valence and arousal values reported in \citealt{warriner-etal2013}).
+We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner-etal2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
 
@@
 
-The control stimuli in the projection blocks of Exps.~1-3 were formed from the sentences in (1).
+The control stimuli in Exps.~1-3 were main clauses. In Exps.~1n and 1m, we used the sentences in
+
+%in the projection blocks of were formed from the sentences in (1).
 In Exps.~1, we used the polar question variants of the sentences in (1) in Exp.~1q, and the (positive declarative) sentences in (1) in Exps.~1n, 1m, and 1c; in Exp.~1c, we added the non-restrictive relative clauses (NRRCs) provided in (1) to the respective subjects, so that the control stimuli have two clauses, like the target stimuli.
@@
 3n & .44  (nobody excluded by specified exclusion criterion) \\
-3m & .5 \\
+3m & .50 \\
 3c & .53 \\

### 2022-03-14 writing/investigating-GPP.tex
started merging appendix C and D

@@
 
-\section{Control stimuli in the projection blocks of Exps.~1-3}\label{a-control}
+\section{Control stimuli in Exps.~1-3}\label{a-control}
 
-The control stimuli in Exps.~1-3 were main clauses. In Exps.~1n and 1m, we used the sentences in
+The control stimuli in Exps.~1-3 were main clauses.
+For the question embedding (Exps.~1q, 2q and 3q), we used the polar questions in (2).
+The NRRCs in parentheses are only used in Exps.~2q and 3q to make the use of the assent diagnostics in the at-issueness blocks more natural: like the target stimuli, the control stimuli consist of two clauses, one of which the relevant speaker assents with.
+
+\begin{exe}
+\exi{(1)} Sentences for control stimuli in question embedding (Exps.~1q, 2q and 3q)
+\begin{xlist}
+\ex Do these muffins (, which are really delicious,) have blueberries in them?
+\ex Does this pizza (, which I just made from scratch,) have mushrooms on it? 
+\ex Was Jack (, who is my long-time neighbor,) playing outside with the kids? 
+\ex Does Ann (, who is a local performer,) dance ballet?
+\ex Were John's kids (, who are very well-behaved,) in the garage?
+\ex Does Samantha (, who is really into fashion,) have a new hat?
+\end{xlist}
+\end{exe}
+
+In the negation, modal and conditional embeddings, we used the positive declarative variants of (2), given in (1).
+Exps.~1n and 1m, we used the sentences in (1) without the NRRCs.
+We used the sentences in (1) with the NRRCs for Exp.~1c and for Exps.~2 and 3.
 
@@
 \begin{exe}
-\exi{(1)}  Sentences for control stimuli in Exps.~1
+\exi{(2)}  Sentences for control stimuli in negation, modal and conditional embeddings
 \begin{xlist}
-\ex These muffins have blueberries in them.  (NRRC: , which are really delicious, )
-\ex This pizza has mushrooms on it. (NRRC: , which I just made from scratch, )
-\ex Jack was playing outside with the kids. (NRRC: , who is my long-time neighbor, )
-\ex Ann dances ballet. (NRRC: , who is a local performer, )
-\ex John's kids were in the garage. (NRRC: , who are very well-behaved, )
-\ex Samantha has a new hat. (NRRC: , who is really into fashion, )
+\ex These muffins (, which are really delicious,) have blueberries in them.
+\ex This pizza (, which I just made from scratch,) has mushrooms on it. 
+\ex Jack (, who is my long-time neighbor,) was playing outside with the kids. 
+\ex Ann (, who is a local performer,) dances ballet.
+\ex John's kids (, who are very well-behaved,) were in the garage.
+\ex Samantha (, who is really into fashion,) has a new hat.
 \end{xlist}
 \end{exe}
-We expected participants to give low responses in the `certain that'  diagnostic  in Exp.~1q (indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question) and high responses in the `certain that'  diagnostic in Exps.~1n, 1m, and 1c (indicating that the speaker is certain of the main clause content).
-These expectations were borne out, as shown in the first three rows of Table \ref{t-proj-controls}.
 
-%We used positive declarative control sentences in Exps.~1n, 1m, and 1c so that we could have clear expectations about participants' responses.
-Thus, we expected participants to give low responses in the `certain that' and `asking whether' diagnostic in Exp.~1q (indicating that the main clause content does not project and is at-issue), and we expected participants to give low responses in the `certain that' diagnostic and high responses in the `sure that' diagnostic in Exps.~1n, 1m, and 1c (indicating, again, that the main clause content does not project and is at-issue).
+\paragraph{Expectations of answers to certainty and at-issueness diagnostics} For the control stimuli in (1) used in Exps.~1q, 2q and 3q, which are polar questions, we expected participants to give low responses in the `certain that' diagnostic, indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question.
+For the control stimuli in (2), which are positive declarative sentences, we expected high responses in the `certain that' diagnostic for the control stimuli, indicating that the speaker is certain of the main clause content.
+These expectations were borne out, as shown in the third column of Table~\ref{t-controls}.
 
-In Exps.~2 and 3, the NRRCs were included in all control stimuli: the carrier sentences were the polar question variants of the sentences in (1) in Exps.~2q and 3q, and the sentences in (1) for the remaining experiments.
-The NRRCs were included in Exps.~2 and 3 to make the use of the assent diagnostics in the at-issueness blocks more natural: like the target stimuli, the control stimuli consist of two clauses, one of which the relevant speaker assents with.
-
-We had the same expectations as for the control stimuli in Exps.~1: low responses in the `certain that' diagnostic in Exps.~2q and 3q (indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question), and high responses in the `certain that' diagnostic in the remaining experiments (indicating, again, that the speaker is certain of the main clause content).
+%In Exps.~2 and 3, the NRRCs were included in all control stimuli: the carrier sentences were the polar question variants of the sentences in (1) in Exps.~2q and 3q, and the sentences in (1) for the remaining experiments.
+%We had the same expectations as for the control stimuli in Exps.~1: low responses in the `certain that' diagnostic in Exps.~2q and 3q (indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question), and high responses in the `certain that' diagnostic in the remaining experiments (indicating, again, that the speaker is certain of the main clause content).
 These expectations were borne out, as shown in the lower nine rows of Table \ref{t-proj-controls}.
@@
 \centering
-\begin{tabular}{r r}
-Experiment & mean certainty rating \\ 
+\begin{tabular}{r r r r l }
+Experiment & Control stimuli & Certainty &  \multicolumn{2}{l}{At-issueness} \\ & & Mean rating & Mean rating & Measure \\ 
 \hline
-1q & .14 \\
-2q & .18 \\
-3q & .17 \\
+1q & (1) & .14 & .05  & asking whether $c$ \\
+2q & (1) with NRRC & .18 & .07 & {\em yes}, $c$\\
+3q & (1) with NRRC & .17 &  .28 & {\em yes}, but $\neg c'$\\
 \hline
-1n & .95 \\
-1m & .96 \\
-1c & .94 \\
+1n & (2) &  .95 & .04 & sure that $c$\\
+1m & (2) & .96 & .03 & sure that $c$\\
+1c & (2) with NRRC & .94  & .08 & sure that $c$\\
 \hline
-2n & .96 \\
-2m & .96 \\
-2c & .96 \\
+2n & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
+2m & (2) with NRRC& .96 & .25 & {\em yes, that's true}, $c$\\
+2c & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
 \hline
-3n & .94 \\
-3m & .93 \\
-3c & .93 \\
+3n & (2) with NRRC & .94 & .44 & {\em yes, that's true}, but $\neg c'$\\
+3m & (2) with NRRC & .93 & .50 & {\em yes, that's true}, but $\neg c'$\\
+3c & (2) with NRRC & .93 & .53 & {\em yes, that's true}, but $\neg c'$\\
 \hline
 \end{tabular}
-\caption{Mean certainty ratings for control stimuli, for self-declared American English participants}\label{t-proj-controls}
+\caption{Mean certainty and at-issueness ratings for control stimuli, for self-declared American English participants}\label{t-controls}
 \end{table}
 
-\newpage
+For the at-issueness diagnostics, we expected the main clause content to be at-issue across the diagnostics.
+We thus expected low responses across all at-issueness diagnostics, given that participants' responses were coded so that the higher a participants' response, the more not-at-issue the content was hypothesized to be (since we investigate whether there is a positive correlation between projection and not-at-issueness).
+However, as shown in the fourth column of Table~\ref{t-proj-controls}, the at-issueness ratings are not always low.
+In particular, xx
 
-\section{Ratings for main clause contents in at-issueness blocks in Exps.~1-3}\label{a-control-ai}
+Exp.~3q\\
+A: Do these muffins, which are really delicious, have blueberries in them?\\
+B: Yes, but they aren't really delicious.
 
-In the at-issueness blocks of Exps.~1-3, we collected at-issueness ratings for the main clause contents of the control stimuli of the projection blocks described in Supplement \ref{a-control}.
-Originally, the intent was to exclude participants' data on the basis of their at-issueness ratings, in parallel to their certainty ratings.
+Exp.~2n, 2m, 2c\\
+A: These muffins, which are really delicious, have blueberries in them.\\
+B: Yes, that's true, they have blueberries in them.
+
+Exp.~3n, 3m, 3c\\
+A: These muffins, which are really delicious, have blueberries in them.\\
+B: Yes, that's true, but they aren't really delicious.
+
+%Originally, the intent was to exclude participants' data on the basis of their at-issueness ratings, in parallel to their certainty ratings.
 We expected the main clause content to be at-issue across the diagnostics, which means that we expected low responses across all at-issueness diagnostics (recall that participants' responses were coded so that the higher a participants' response, the more not-at-issue the content was hypothesized to be, to investigate whether there is a positive correlation between projection and not-at-issueness).
@@
 
-\begin{table}[h!]
-\centering
-\begin{tabular}{r l}
-Experiment & mean rating \\ 
-\hline
-1q & .05 \\
-2q & .07 \\
-3q & .28 \\
-\hline
-1n & .04 \\
-1m & .03 \\
-1c & .08 \\
-\hline
-2n & .22 (16 participants excluded by specified exclusion criterion) \\
-2m & .25 \\
-2c & .22 \\
-\hline
-3n & .44  (nobody excluded by specified exclusion criterion) \\
-3m & .50 \\
-3c & .53 \\
-\hline
-\end{tabular}
-\caption{Mean ratings on at-issueness diagnostics for main clause contents in Exps.~1-3, for self-declared American English participants}\label{t-ai-controls}
-\end{table}
+%\begin{table}[h!]
+%\centering
+%\begin{tabular}{r l}
+%Experiment & mean rating \\ 
+%\hline
+%1q & .05 \\
+%2q & .07 \\
+%3q & .28 \\
+%\hline
+%1n & .04 \\
+%1m & .03 \\
+%1c & .08 \\
+%\hline
+%2n & .22 (16 participants excluded by specified exclusion criterion) \\
+%2m & .25 \\
+%2c & .22 \\
+%\hline
+%3n & .44  (nobody excluded by specified exclusion criterion) \\
+%3m & .50 \\
+%3c & .53 \\
+%\hline
+%\end{tabular}
+%\caption{Mean ratings on at-issueness diagnostics for main clause contents in %Exps.~1-3, for self-declared American English participants}\label{t-ai-controls}
+%\end{table}
 

### 2022-07-06 writing/investigating-GPP.tex
update paper

--- 2022-02-28 writing/investigating-GPP.tex
@@
 \begin{exe}
+\ex Context: A and B live near a dog that regularly barks for extended periods of time. It is barking right now. A says:  \\ Ugh. The damn dog is barking again.
+\end{exe}
+
+Contents:
+
+\begin{itemize}
+\item there is a unique salient dog (backgrounded, ps)
+\item the dog has barked before (backgrounded, ps)
+\item the speaker is annoyed at the dog/barking (new, could be what the speaker wants to contribute)
+\item the dog is barking (``at-issue'', ordinary entailment)
+\end{itemize}
+
+First dimension: Role in discourse context: some anchor to context, some convey relevant backgrounded information, some are what the speaker wants to communicate.
+
+Second dimension: Compositional semantics: some interact, some don't.
+
+Third dimension: Role in subsequent discourse: some are what carries the conversation on, some do not.
+
+One of the contents is privileged in that it is the content that is put up for negotiation.
+It is often new information, though not necessarily.
+It is what is ``linguistically relevant'': it interacts with operators and is accessible for anaphoric reference.
+
+\begin{exe}
 \ex\label{gpp} {\bf Gradient Projection Principle} \hfill (\citealt[400]{tbd-variability}) \\ If content $C$ is expressed by a constituent embedded under an entailment-canceling operator, then $C$ projects to the extent that it is not at-issue.
 \end{exe}
+
+We already know that the GPP is stated too strongly: prior influences projection (\citealt{degen-tonhauser-openmind}) {\bf independently of at-issueness (unpublished)}.
+At-issueness is one factor among many.
 
@@
 
-Previous investigations of the relationship between projection and at-issueness:
+Previous investigations of/claims about the relationship between projection and at-issueness:
 
@@
 \item \citealt{tbd-variability}
+
+\item \citealt{koev2018}
 
@@
 \subsection{Formal analyses of at-issueness}
+
+- Esipova, Maria. 2018. QUD-addressing appositives don?t have to be clause-final. Snippets 33. 7?8. https://doi.org/10.7358/snip-2018-033-esip.
+
+- Djärv, Kajsa. 2019. Factive and assertive attitude reports. Philadelphia, PA: Uni- versity of Pennsylvania dissertation. https://repository.upenn.edu/edissertations/ 3645.
 
@@
 \item p.6: This test brings up the question of what counts as a direct response.
-One peculiarity about the responses in 10?11 is that they are anaphoric to parts of the target utterance through propositional anaphors like that or no.7 This is important because nonanaphoric denials behave differently: They can felicitously target nominal appositives (see also Hunter & Asher, 2016).
+One peculiarity about the responses in 10?11 is that they are anaphoric to parts of the target utterance through propositional anaphors like that or no.7 This is important because nonanaphoric denials behave differently: They can felicitously target nominal appositives (see also Hunter \& Asher, 2016).
 
@@
 or ?discourse?
-relations (Asher \& Lascarides, 2003; Hobbs, 1979, 1985; Kehler, 2002; Mann & Thompson, 1988).
+relations (Asher \& Lascarides, 2003; Hobbs, 1979, 1985; Kehler, 2002; Mann \& Thompson, 1988).
 
@@
 
-\item endnote 5: The assent/dissent test is known under different names in the literature and has been used to distinguish asserted content from implications triggered by constructions as different as presupposition markers (Chierchia & McConnell-Ginet, 2000; Karttunen & Peters, 1979; Shanon, 1976; Strawson, 1950), epistemic modals (Lyons, 1977; Papafragou, 2006; von Fintel & Gillies, 2007), evidentials (Faller, 2002; Matthewson et al., 2007; Murray, 2014), and appositives (AnderBois et al., 2015; Koev, 2013; Syrett & Koev, 2015; Tonhauser, 2012).
+\item endnote 5: The assent/dissent test is known under different names in the literature and has been used to distinguish asserted content from implications triggered by constructions as different as presupposition markers (Chierchia \& McConnell-Ginet, 2000; Karttunen \& Peters, 1979; Shanon, 1976; Strawson, 1950), epistemic modals (Lyons, 1977; Papafragou, 2006; von Fintel \& Gillies, 2007), evidentials (Faller, 2002; Matthewson et al., 2007; Murray, 2014), and appositives (AnderBois et al., 2015; Koev, 2013; Syrett \& Koev, 2015; Tonhauser, 2012).
 
@@
 
-\item \citealt{snider-thesis, snider2017, snider2018}
+\item \citealt{snider-thesis}
+
+\item \citealt{snider2017} (superseded for the most part by 2018 paper)
+
+\begin{itemize}
+
+\item In this paper, I argue that there is no tight link between at-issueness and anaphoric availability?
+that these two properties of propositions are distinct.
+I present data which demonstrates that a proposition?s at-issue status alone is not sufficient to determine whether it will be available for anaphoric reference (including but not limited to direct assent/dissent), as well as data which shows that a proposition?s being available for anaphoric reference cannot be used to diagnose its at-issue status.
+
+\item  Tonhauser asserts that they are all indeed testing for at-issueness: ?they are all useful to diagnose (not-)at-issueness with at least one kind of content? (p. 252). {\bf I was wrong}
+
+\item p.6: The question/answer tests return the same results, because their use of anaphors (response particles) is only incidental.
+In contrast, the direct dissent tests rely on explicit anaphors by design; an anaphor-less equivalent to the direct dissent test in (11) ceases to be the same diagnostic.
+
+\end{itemize}
+
+\item \citealt{snider2018}
+
+\begin{itemize}
+
+\item p.374: at-issueness and the anaphoric potential of a proposition are distinct notions; one class of tests commonly used to diagnose at-issue status in fact diagnoses only anaphoric potential
+
+\item NRRCs: anaphoric potential varies, especially with position, but never good to address question under discussion
+
+\item at-issue content: certain content conveyed by an utterance has a distinguished status as the main point of the utterance, while other content does not.
+Potts 2005 discusses ?at-issue entailments?
+as distinct from presuppositions and conventional implicatures, but does not describe how to identify content as at-issue.
+
+\item p.375: Properties of at-issue content (from \citealt{tonhauser-sula6})
+
+At-issue content can be directly assented or dissented with {\bf third dimension}
+
+At-issue content addresses the question under discussion {\bf first dimension}
+
+At-issue content determines the relevant set of alternatives {\bf second dimension, compositionality}
+
+Some content may overlap on all three dimensions: e.g., that Fritz is happy in {\em Fritz is happy}.
+But sentence-final appositives can be assented/dissented with, but they do not address the QUD or determine the relevant set of alternatives.
+And that the speaker is annoyed may address the QUD, but does not determine the relevant set of alternatives.
+
+\item p.376f: Snider argues against \citealt{syrett-koev2015} claim that appositives can address the QUD.
+``Why questions are crucially different from these other questions in that they seek explanations.
+Because they seek explanations, such questions have no single comprehensive answer: any answer given is defeasible2 or can be considered insufficiently precise, and additional contextual information can always contribute to an explanation (see discussion in Bjorndahl \& Snider, 2016).
+The appositive might contribute information which bolsters an explanation, supporting the matrix clause?s answer to the QUD, but that alone does not constitute addressing the QUD.
+If the matrix clause does not address the QUD, then an appositive alone cannot be taken to address the QUD, even in a why question, as in (7)''
+
+Syrett \& Koev 2015:587 further points to an example where an appositive appears to ?address one part of a coordinated QUD?, inspired by an example from Koev 2013.
+But the concept of a ?coordinated QUD?
+is novel: Roberts 1996 describes no such structure, instead describing sequences of ?relevant sub-questions to some super-question?
+as an ?enumeration, suggesting a plan for how to attack the super-question?, and says that one ?can only address such sub-questions one at a time?
+(5).
+The response in (8A) thus addresses the first sub-question?as it must, because the second requires an answer to the first?and not both at once.
+
+\item p.378: The crucial difference between the classes of diagnostics illustrated above is that the first is anaphora-based, where the other two are QUD-based, as they test what can address or establish a Question Under Discussion.
+
+Snider assumes that the two QUD diagnostics (addressing QUD, establishing QUD) are QUD-based, i.e., tap into the same property?
+But establishing a QUD is about compositionality, and addressing a QUD is about the role of content in discourse.
+
+\item p.380: Alternatively, this observation might lead one to move away from the Simons et al.
+2010 definition entirely.
+Hunter \& Asher 2016 proposes a discourse based account using SDRT (Asher, 1993; Asher \& Lascarides, 2003) in which at-issueness and anaphoric potential are indeed closely linked?which isn?t at odds with the data discussed here, because they use an entirely different notion of at-issueness.
+On this account, ?content is at-issue at a certain point of the discourse just in case it is on the [Right Frontier (RF)] of the discourse at that time.
+Content that is AI becomes NAI when knocked off the RF; content that is NAI can become AI if targeted through discourse subordination?
+and thus returned to the Right Frontier (Hunter \& Asher, 2016:1036).
+And, because discourse referents (propositional and otherwise) are only accessible for anaphora resolution in SDRT if they are on the Right Frontier, propositional content which is available for anaphoric reference is necessarily at-issue on this account.
+
+\item p.380: However one chooses to define or model at-issueness, it is important to recognize the difference between the at-issue status of a proposition and its anaphoric potential.
+The two are at least conceptually distinct: at-issueness is about what the main point of an utterance is, while anaphoric potential is about what propositions are available for subsequent anaphoric reference.
+
+\end{itemize}
 

### 2022-08-03 notes lisa/notes.tex
Added missing graphs and my notes

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{xunicode,xltxtra, polyglossia}
\usepackage{fontspec} %(include if mathspec is not loaded)
\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage[dvipsnames]{xcolor}
\usepackage{enumitem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}

\usepackage{fancyhdr}
\pagestyle{fancy}

\usepackage{linguex}
\usepackage{multicol}

\usepackage[colorlinks, hyperfootnotes=false]{hyperref} % hyperlinks
\hypersetup{linkcolor=MidnightBlue,citecolor=MidnightBlue,filecolor=MidnightBlue,urlcolor=MidnightBlue}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%

\lhead{\opt \Large GPP  experiments notes}
\rhead{\opt \large Lisa Hofmann\\ \today}
\cfoot{\thepage}

%language
\setdefaultlanguage[variant=american]{english}
% \def\hyph{-\penalty0\hskip0pt\relax}

% fonts
\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Libertine}
\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
% \setmonofont[Scale=MatchLowercase]{Andale Mono}
\newfontfamily\opt{Optima}

\begin{document}
\tableofcontents

% \section{Questions} % (fold)
% 	\begin{itemize}
% 		\item Were the questions for projection / at-issueness diagnostics asked in randomized order?
% 		\item What were the thoughts on arousal? How would this come into play? Is the idea that it would interact with the QUD?
% 		\item How exactly are the pilots different from the main experiments?
% 	\end{itemize}

% % section questions (end)

\pagebreak
\section{Experiments with Declaratives and Negation}

	\paragraph{Summary} % (fold)

		\begin{itemize}
			\item We don't see the correlation between non-at-issueness measures and projection measures here that are predicted by the GPP
			\item This seems to be due to problems with the intended at-issueness diagnostics: They confound at-issueness with speaker commitments, which does make them sensitive to projection as well
		\end{itemize}

	% paragraph summary (end)

	\subsection{2n---Negation embedding, yes, that's true+complement diagnostic (+ acceptability rating)}
			\ex. \a. \textbf{Christopher:} Cole didn’t discover that Julian dances salsa.
				\b. \textbf{Sandy:} Yes, that’s true, Julian dances salsa.
				\z.
			\z.

			\begin{itemize}

				\item yes, that's true+complement diagnostic for at-issueness (?):
				\begin{itemize}
					\item Tests whether you can agree with the previous assertion and assert the complement
					\item so, in a sense this is a diagnostic of projection as well
				\end{itemize}

				\item Certain-that diagnostic for projection: Tests whether participant assumes that speaker is committed to complement

				\item Projection does not correlate with the (intended) non-at-issueness measures
			\end{itemize}

			\begin{figure}[h]
				\centering
				\includegraphics[]{figures/n2-correl.pdf}
				\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 2n}
				\label{fig:figure1}
			\end{figure}

			\begin{itemize}
				\item In fact, it looks like there might be an inverse relationship
				\item I would expect that projection-measure correlates better with the opposite (i.e. the intended measure of at-issueness)
				\item For declaratives, a different at-issueness diagnostic may be needed

			\end{itemize}

	\subsection{How does the correlation for 2n look for projection vs. the positive responses to the assent-diagnostic?} % (fold)	
		Plot for correlation between projectivity-measure and high acceptability ratings for assent w/ assertion of complement.

			\begin{figure}[h]
				\centering
				\includegraphics[]{figures/n2-inv-correl.pdf}
				\caption{Mean at-issueness ratings vs mean certainty ratings by predicate, Exp. 2n}
				\label{fig:figure4}
			\end{figure}

		\begin{itemize}
			\item This looks more like a non-linear correlation
		\end{itemize}
	% subsection how_does_the_correlation_for_2n_look_with_at_issueness_ (end)

	\subsection{3n---Negation embedding, yes, that's true + negation of main clause (+ acceptability rating)} % (fold)
		\ex. \a. \textbf{Christopher:} Cole didn’t discover that Julian dances salsa.
				\b. \textbf{Sandy:} Yes, that’s true, but Cole discovered it.
				\z.
			\z.

			\begin{itemize}

				\item yes, that's true+negation of main clause diagnostic for at-issueness:
				\begin{itemize}
					\item Tests whether you can agree with the previous assertion and assert the negation of the main clause
					\item agreement with main point of previous utterance
					\item If the main point was the complement, and the first speaker is committed to truth of the complement, then assent should be possible while rejecting the main clause.
					\item This addresses at-issueness of the complement.
					\item But potentially also confounds speaker commitment to complement (i.e. projection) with at-issueness.
					\item This is assuming that assent signals that the second speaker takes on (some of) the commitments of the first speaker.
					\item Based on this assumption, we can also assume the acceptability ratings to be relatively low throughout. %(see \textbf{Figure \ref{fig:figure5}})
				\end{itemize}

				% \begin{figure}[h]
				% 	\centering
				% 	\includegraphics[width=.75\textwidth]{figures/n3-nai.pdf}
				% 	\caption{Mean not-at-issueness ratings by predicate, Exp. 3n}
				% 	\label{fig:figure5}
				% \end{figure}

				% \pagebreak
				\item Certain-that diagnostic for projection: Tests whether participant assumes that speaker is committed to complement

				\item Does projection correlate with the (intended) non-at-issueness measures? If there is a confound, we would again expect something more to the opposite effect. (see \textbf{Figure \ref{fig:figure6}})
			\end{itemize}

			\begin{figure}[h]
				\centering
				\includegraphics[scale=0.90]{figures/n3-correl.pdf}
				\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 3n}
				\label{fig:figure6}
			\end{figure}

	% subsection 3n_negation_embedding_yes_that_s_true_negation_of_main_clause_acceptability_rating (end)

	\pagebreak
	\subsection{1n---Negation embedding, are you sure?, assent + assertion of complement + Q-A judgment} % (fold)
		\ex. \a. \textbf{Christopher:} Cole didn’t discover that Julian dances salsa.
				\b. \textbf{Sandy:} Are you sure?
				\b. \textbf{Christopher:} Yes, I'm sure that Julian dances Salsa.
				\z.
			\z.

		\begin{itemize}
			\item This diagnostic also conflates non-at-issueness with projection: Only if Christopher is committed to truth of the complement can they respond in this way

			\item Here, also we don't see the correlation come out (\textbf{Figure \ref{fig:1n-correl}})
		\end{itemize}

			\begin{figure}[h]
					\centering
					\includegraphics[scale=0.9]{figures/n1-correl.pdf}
					\caption{Mean at-issueness ratings vs mean certainty ratings by predicate, Exp. 1n \label{fig:1n-correl}}
				\end{figure}

	% subsection 1n_negation_embedding_are_you_sure_q_a_judgment (end)

	\pagebreak
	\subsection{A conclusion about at-issueness diagnostics} % (fold)
		\begin{itemize}
			\item A diagnostic of at-issueness should not interact with speaker commitment
			\item Problem with negation: If the complement is at issue, it will be targeted by entailment-cancelling operators
			\item On the other hand, if the complement is not at issue, it is predicted to project
			\item A diagnostic should target the question of whether the first utterance addresses the quesion $\{c, \neg c\}$, while leaving open how exactly the question is resolved

			\item Here is one idea:
		\end{itemize}

		\begin{itemize}
			\item Asking what it was about without commitment:
				\ex. \a. \textbf{Christopher:} Cole didn’t discover that Julian dances salsa.
					\b. \textbf{Sandy:} Oh right, Mary told me that as well.\\
					\textbf{Forced Choice:} What did Mary talk to Sandy about?
						\a. Whether Julian dances salsa.
						\b. Whether Cole discovered that Julian dances salsa.
						\z.
					\z.
				\z.

			% \pagebreak
			\item Question-answer based: Can the declarative address a question about the complement?
				\ex. \a. \textbf{Sandy:} Does Julian dance salsa?
					\b. \textbf{Christopher:} Cole didn’t discover that Julian dances salsa.\\
						\textbf{Q-A judgment:} Did Christopher answer Sandy's question?
					\z.
				\z.

			\item Anaphora-based:

				\ex. \a. \textbf{Christopher:} Cole didn’t discover that Julian dances salsa.
					\b. \textbf{Sandy:} That would be quite surprising.\\
					\textbf{Forced Choice:} What would Sandy find surprising?
						\a. Julian dancing salsa.
						\b. Cole discovering that Julian dances salsa.
						\z.
					\z.
				\z.

		\end{itemize}

		\noindent These tests separate at-issueness diagnostics from commitment, as well as the question of whether or not the second speaker (dis)agrees with the first.

		\begin{itemize}
			\item This is important, especially under negation, because if we expect that at-issueness correlates with non-projection, we expect the complement to be targeted by negation, and therefore counterfactual.

			\item Separating at-issueness and commitment of the complement I think that we can expect a correlation between non-at-issueness and projection again.

				\ex. \textbf{Context:} Cole is a child that is being picked up from daycare
					\a. \textbf{Parent:} What did Cole do today?
					\b. \textbf{Caretaker:} Cole didn't see that Julian danced salsa.\\
					$\Rightarrow$ Probable inference that Julian danced salsa
					\z.
				\z.

				\ex. \textbf{Context:} Julian is a child that is being picked up from daycare 
					\a. \textbf{Parent:} What did Julian do today?
					\b. \textbf{Caretaker:} Cole didn't see that Julian danced salsa.\\
					$\Rightarrow$ Probable inference that Julian did not dance salsa (?)\\
					$\Rightarrow$ Additional evidential inference Cole is a reliable source on whether Julian danced Salsa
					\z.
				\z.

			\item That should be the case, at least for certain predicates, namely veridical ones (see discussion below)

			\item Some open questions:
			\begin{itemize}
				\item The experiment assumes that the probability of complement at-issueness is influenced by the lexical embedding. That seems to be the case

				\item However, there are other factors that can influence at-issueness
				\begin{itemize}
					\item \textbf{Discourse context (QUD):} An utterance is more likely to be about a topic that has already been brought up/a question that has been raised
					\item \textbf{Epistemic context (common ground):} An utterance is more likely to be about an issue that hasn't been settled
				\end{itemize}

			\end{itemize}

		\end{itemize}

	% subsubsection a_conclusion_about_at_issueness_diagnostics (end)

% \pagebreak
\section{Experiments with question embedding} % (fold)

	\paragraph{Summary} % (fold)
		\begin{itemize}
			\item For questions, the correlation comes out more clearly
			\item This suggests that the at-issueness diagnostics for questions work better.
				\begin{itemize}
					\item This seems clear for asking-whether diagnostic
					\item Although we questioned this last time, I think that it applies to yes+$c$ diagnostic
					\item It doesn't work for yes but not $c^\prime$ diagnostic
				\end{itemize}

			\item There are some lexical exceptions wrt the correlation
			\item The exceptions are:
			\begin{itemize}
				\item Mostly non-veridical predicates (\emph{say, suggest, think, pretend})
					\begin{itemize}
						\item This suggests lexical contibution of the verb:
						\item Only if the verb is associated with an inference about the truth of the complement, that inference can project
					\end{itemize}

				\item Maybe some veridical assertives(?) (\emph{prove, \textcolor{gray!50}{establish, demonstrate, confirm}})
				\begin{itemize}
					\item c.f. Anand \& Hacquard 2014 on epistemic vs assertive attitudes: veridical epistemics are factive, while veridical assertives are non-presuppositional
					\item Interestingly, though, other veridical assertives behave differently (\emph{inform, acknowledge, confess, admit})
					\item Not sure if the latter group actually is characterized as in the same way in A\&H: do they take (non-agentive/non-intentional) repository of information subjects (e.g. \emph{\#/? The test results informed us/ acknowledged/confessed/admitted that Mary is free of contagious diseases.} vs. \emph{The test results proved/ established/demonstrated/confirmed that Mary is free of contagious diseases.})
					\item These two groups are discussed in A\&H as apparent counterexamples
					\item Djärv (2019): analysis of non-projection as involving evidential interpretations
				\end{itemize}
			\end{itemize}
		\end{itemize}

	% paragraph summary (end)

	\pagebreak
	\subsection{1q---Question embedding, asking-whether diagnostic} % (fold)

			\ex. \a. \textbf{Christopher:} Did Cole discover that Julian dances salsa?
				\b. \textbf{Question-based at-issueness diagnostic:} Is Christopher asking whether Julian dances salsa?
				\z.
			\z.

			\begin{itemize}

				\item Asking-whether diagnostic for at-issueness:
				\begin{itemize}
					\item Tests whether the utterance is \lq about\rq\ the truth of the complement
					\item Question-based at-issueness diagnostic
					\item What is the main point of the utterance?
				\end{itemize}

				\item Certain-that diagnostic for projection: Tests whether participant assumes that speaker is committed to complement

				\item Not-at-issueness correlates with projection (\textbf{Figure \ref{fig:q1-correl}})
				\begin{itemize}
					\item but only for veridical predicates
					\item the lexical exceptions are non-veridical predicates
					\item Updated GPP (?): given an inference is associated with a predicate, it will project to the extent that it is not at issue
				\end{itemize}

			\end{itemize}

			\begin{figure}[h]
				\centering
				\includegraphics[]{figures/q1-correl.pdf}
				\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 1q}
				\label{fig:q1-correl}
			\end{figure}

	% subsection q1_question_embedding_asking_whether_diagnostic (end)

	\pagebreak
	\subsection{2q---Question embedding, yes+complement diagnostic (+ acceptability rating)} % (fold)
		\ex. \a. \textbf{Christopher:} Did Cole discover that Julian dances salsa?
				\b. \textbf{Sandy:} Yes, Julian dances salsa.
				\z.
			\z.

			\begin{itemize}
				\item yes+c diagnostic for at-issueness (acceptablity rating):
				\begin{itemize}
					\item Tests whether an affirmative response to the question can be accompanied with an assertion of the complement (there is an intuitive sense in which they should be \lq about\rq\ the same thing)

					\item Therefore, this can also be understood as a diagnostic of the main point of the question

				\end{itemize}

				\item Certain-that diagnostic for projection: Tests whether participant assumes that speaker is committed to complement

				\item Not-at-issueness correlates with projection (\textbf{Figure \ref{fig:q2-correl}})
				\begin{itemize}
					\item Similar pattern of lexical exceptions
					\item They are more spaced out here
				\end{itemize}
			\end{itemize}

			\begin{figure}[h]
				\centering
				\includegraphics[]{figures/q2-correl.pdf}
				\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 2q}
				\label{fig:q2-correl}
			\end{figure}

	% subsection q1_question_embedding_asking_whether_diagnostic (end)

	\subsection{3q---Question embedding, yes but not $c^\prime$ (+ acceptability rating)}
		\ex. \a. \textbf{Christopher:} Did Cole discover that Julian dances salsa?
				\b. \textbf{Sandy:} Yes, Julian dances salsa.
				\z.
			\z.

		\begin{itemize}
				\item yes but not $c^\prime$ diagnostic for at-issueness (acceptablity rating):
				\begin{itemize}
					\item Tests whether an affirmative response to the question can be accompanied with a denial of the main clause

					\item Expectation: If the question is about the complement, we might be able to affirm the complement while denying the main clause

					\item This is, again, based on the assumption that the complement and main clause are interpreted independent of each other

					\item However, if Djärv (2019) is right that projective interpretations are interpreted as evidentials, it is expected that we cannot affirm the complement while rejecting the main clause, in cases where there is projection.

					\item In fact all ratings for this test are quite low overall.

				\end{itemize}

				\item Certain-that diagnostic for projection: Tests whether participant assumes that speaker is committed to complement

				\item (Intended) not-at-issueness measure does not correlate with projection, see \textbf{Figure \ref{fig:q3-correl}}
			\end{itemize}

			\begin{figure}[h]
				\centering
				\includegraphics[]{figures/q3-correl.pdf}
				\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 3q
				\label{fig:q3-correl}}
			\end{figure}

	% subsection

% section three_experiment_setups (end)

\section{Experiments with modals}
	\subsection{1m---Modal embedding (perhaps),  are you sure?, assent + assertion of complement + Q-A judgment} % (fold)
		\ex. \a. \textbf{Christopher:} Perhaps Cole discovered that Julian dances salsa.
			\b. \textbf{Sandy:} Are you sure?
			\b. \textbf{Christopher:} Yes, I'm sure that Julian dances Salsa.
			\z.
		\z.

		\begin{itemize}
			\item Like with negation above, this test clearly conflates at-issueness and projection: Only of Christopher is committed to $c$, can they respond in this way

			\item Correlation does not exactly come out see \textbf{Figure \ref{fig:m1-corr}}
			\begin{itemize}
				\item However, more so than for negation
				\item Interestingly, the lexical exceptions are similar here: \emph{prove, say, suggest, think, pretend}
			\end{itemize}
		\end{itemize}

		\begin{figure}[h]
			\centering
			\includegraphics[scale=0.85]{figures/m1-correl.pdf}
			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 1m}
			\label{fig:m1-corr}
		\end{figure}

	% subsection 1m_modal_embedding_ (end)

	\pagebreak
	\subsection{2m---Modal embedding (perhaps) yes, that's true+complement diagnostic (+ acceptability rating)} % (fold)
		\ex. \a. \textbf{Christopher:} Perhaps Cole discovered that Julian dances salsa.
			\b. \textbf{Sandy:} Yes, that’s true, Julian dances salsa.
			\z.
		\z.

		\begin{itemize}
			\item Same test as above: Intended at-issueness measure also tests whether agreeing is possible while asserting the complement: confound with commitment

			\item Overall acceptabiliy ratings are higher for this test compared to 1m

			\item But the overall pattern in correlation with projection-measure is similar \textbf{Figure \ref{fig:m2-corr}}
			\begin{itemize}
				\item More of a correlation compared to negation experiments
				\item Again same predicates outside of the main cluster: \emph{prove, say, suggest, think, pretend}
			\end{itemize}
		\end{itemize}

		\begin{figure}[h]
			\centering
			\includegraphics[scale=0.85]{figures/m2-correl.pdf}
			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 2m}
			\label{fig:m2-corr}
		\end{figure}

	% subsection 1m_modal_embedding_ (end)

	\pagebreak
	\subsection{3m---Modal embedding (perhaps), yes, that's true + negation of main clause (+ acceptability rating)} % (fold)
		\ex. \a. \textbf{Christopher:} Perhaps Cole discovered that Julian dances salsa.
			\b. \textbf{Sandy:} Yes, that’s true, but Cole didn't discover it.
			\z.
		\z.

		\begin{itemize}
			\item Similar pattern to other modal experiments above
		\end{itemize}

		\begin{figure}[h]
			\centering
			\includegraphics[]{figures/m3-correl.pdf}
			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 3m}
			\label{fig:m3-corr}
		\end{figure}

	% subsubsection 3m_ (end)

\pagebreak
\section{To do} % (fold)

	\begin{itemize}
		\item 1q+2q work nicely w lexical exceptions (exceptions have to do with veridicality)
		\item What is diagnosed here?
		\item What kinds of diagnostics correspond to theories of at-issueness?
		\item Polar quesitons vs assertions as different discourse moves; also reponses to these interact with commitment in different ways
		\item Can we find a reliable diagnostic for at-issueness, about the same thing and across contexts?
		\item Aaron White and colleagues have data looking at negative inferences
		\item how do the different diagnostics and confounds w commitment shine a light on lexical differences?
	\end{itemize}

% section to_do (end)

% section today_further_questions (end)

\end{document}

### 2022-08-03 results/main/11_projaiM/rscripts/graphs.R

@@ -1,0 +1,179 @@
+# 11_projaiM
+# modal: projection and at-issueness (assent with positive continuation)
+# graphs
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load required packages
+require(tidyverse)
+library(ggrepel)
+library(dichromat)
+library(forcats)
+library(ggrepel)
+theme_set(theme_bw())
+
+# load helper functions
+source('../../helpers.R')
+
+d = read_csv("../data/data_preprocessed.csv")
+
+# spread responses over separate columns for projectivity and at-issueness
+t = d %>%
+  mutate(block_ai = ifelse(question_type == "ai", ifelse(block == "block1", "block1", "block2"), ifelse(block == "block1", "block2", "block1"))) %>%
+  select(workerid,content,short_trigger,question_type,response,block_ai) %>%
+  spread(question_type,response)
+
+cd = t
+
+table(cd$short_trigger)
+# change cd verb names to match veridicality names
+cd = cd %>%
+  mutate(verb=recode(short_trigger, control = "MC", annoyed = "be_annoyed", be_right_that = "be_right", inform_Sam = "inform"))
+
+# mean not-at-issueness by predicate, including the main clause controls
+ai.means = cd %>%
+  group_by(short_trigger) %>%
+  summarize(Mean = mean(ai), CILow = ci.low(ai), CIHigh = ci.high(ai)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, verb = fct_reorder(as.factor(short_trigger),Mean))
+ai.means
+
+# define colors for the predicates
+cols = data.frame(V=levels(ai.means$verb))
+
+cols$VeridicalityGroup = as.factor(
+  ifelse(cols$V %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
+         
+levels(cols$V)
+cols$V <- factor(cols$V, levels = cols[order(as.character(ai.means$verb)),]$V, ordered = TRUE)
+
+cols$Colors =  ifelse(cols$VeridicalityGroup == "F", "tomato1", "black")
+                      
+cols$Colors
[…]

### 2022-08-11 writing/investigating-GPP.tex
updated paper with discussion of controls

@@
 
+\newpage
+
 \section{Control stimuli in Exps.~1-3}\label{a-control}
@@
 The control stimuli in Exps.~1-3 were main clauses.
-For the question embedding (Exps.~1q, 2q and 3q), we used the polar questions in (2).
-The NRRCs in parentheses are only used in Exps.~2q and 3q to make the use of the assent diagnostics in the at-issueness blocks more natural: like the target stimuli, the control stimuli consist of two clauses, one of which the relevant speaker assents with.
-
-\begin{exe}
-\exi{(1)} Sentences for control stimuli in question embedding (Exps.~1q, 2q and 3q)
+In the question embedding experiments (Exps.~1q, 2q and 3q), the control stimuli consisted of the polar questions in (1).
+The NRRCs in parentheses were realized in Exps.~2q and 3q, where at-issueness was measured with an assent diagnostic: the control stimuli here consisted of two clauses (like the target stimuli), allowing the relevant speaker to assent with one of the two clauses.
+
+\begin{exe}
+\exi{(1)} Sentences for control stimuli in in question embedding experiments (Exps.~1q, 2q and 3q)
 \begin{xlist}
@@
 
-In the negation, modal and conditional embeddings, we used the positive declarative variants of (2), given in (1).
-Exps.~1n and 1m, we used the sentences in (1) without the NRRCs.
-We used the sentences in (1) with the NRRCs for Exp.~1c and for Exps.~2 and 3.
-
-%in the projection blocks of were formed from the sentences in (1).
-In Exps.~1, we used the polar question variants of the sentences in (1) in Exp.~1q, and the (positive declarative) sentences in (1) in Exps.~1n, 1m, and 1c; in Exp.~1c, we added the non-restrictive relative clauses (NRRCs) provided in (1) to the respective subjects, so that the control stimuli have two clauses, like the target stimuli.
+We expected participants to give low responses on the `certain that' diagnostic for the control stimuli in (1), indicating that the speaker is not certain of the main clause content, because main clause content is hypothesized to not project out of the question.
+These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for Exps.~1q, 2q, and 3q.
+We also expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (1), indicating that the main clause content is at-issue.
+These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but not for Exp.~3q.
+
+In the remaining experiments, the control stimuli consisted of the positive declarative variants of (1) given in (2).
+The NRRCs in parentheses were realized in Exps.~2 and 3 for the reason explained above, as well as in Exp.~1c, to make the control stimuli more similar to the target stimuli (which also consisted of two clauses: the antecedent and the consequent).
 
@@
 
-\paragraph{Expectations of answers to certainty and at-issueness diagnostics} For the control stimuli in (1) used in Exps.~1q, 2q and 3q, which are polar questions, we expected participants to give low responses in the `certain that' diagnostic, indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question.
-For the control stimuli in (2), which are positive declarative sentences, we expected high responses in the `certain that' diagnostic for the control stimuli, indicating that the speaker is certain of the main clause content.
-These expectations were borne out, as shown in the third column of Table~\ref{t-controls}.
-
-%In Exps.~2 and 3, the NRRCs were included in all control stimuli: the carrier sentences were the polar question variants of the sentences in (1) in Exps.~2q and 3q, and the sentences in (1) for the remaining experiments.
-%We had the same expectations as for the control stimuli in Exps.~1: low responses in the `certain that' diagnostic in Exps.~2q and 3q (indicating that the speaker is not certain of the main clause content, i.e., that the content does not project out of the question), and high responses in the `certain that' diagnostic in the remaining experiments (indicating, again, that the speaker is certain of the main clause content).
-These expectations were borne out, as shown in the lower nine rows of Table \ref{t-proj-controls}.
+We expected participants to give high responses on the `certain that' diagnostic for the control stimuli in (2), indicating that the speaker is certain of the main clause content, because speakers are hypothesized to be committed to the asserted main clause content.
+These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for the `n', `m' and `c' variants of Exps.~1, 2, and 3.
+We expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (2), indicating that the main clause content is at-issue.
+These expectations were borne out for the `n', `m', and `c' variants of Exps.~1, as shown in the fourth column of Table~\ref{t-controls}, but not for the `n', `m', and `c' variants of Exps.~2 and 3.
 
@@
 \begin{tabular}{r r r r l }
-Experiment & Control stimuli & Certainty &  \multicolumn{2}{l}{At-issueness} \\ & & Mean rating & Mean rating & Measure \\ 
+& &  \multicolumn{2}{c}{Mean ratings} &  \\ 
+Exp. & Control stimuli & Certainty & At-issueness & At-issueness measure \\ 
 \hline
 1q & (1) & .14 & .05  & asking whether $c$ \\
-2q & (1) with NRRC & .18 & .07 & {\em yes}, $c$\\
-3q & (1) with NRRC & .17 &  .28 & {\em yes}, but $\neg c'$\\
-\hline
 1n & (2) &  .95 & .04 & sure that $c$\\
@@
 \hline
+2q & (1) with NRRC & .18 & .07 & {\em yes}, $c$\\
 2n & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
@@
 \hline
+3q & (1) with NRRC & .17 &  .28 & {\em yes}, but $\neg c'$\\
 3n & (2) with NRRC & .94 & .44 & {\em yes, that's true}, but $\neg c'$\\
@@
 
-For the at-issueness diagnostics, we expected the main clause content to be at-issue across the diagnostics.
-We thus expected low responses across all at-issueness diagnostics, given that participants' responses were coded so that the higher a participants' response, the more not-at-issue the content was hypothesized to be (since we investigate whether there is a positive correlation between projection and not-at-issueness).
-However, as shown in the fourth column of Table~\ref{t-proj-controls}, the at-issueness ratings are not always low.
-In particular, xx
-
-Exp.~3q\\
-A: Do these muffins, which are really delicious, have blueberries in them?\\
-B: Yes, but they aren't really delicious.
-
-Exp.~2n, 2m, 2c\\
-A: These muffins, which are really delicious, have blueberries in them.\\
-B: Yes, that's true, they have blueberries in them.
-
-Exp.~3n, 3m, 3c\\
-A: These muffins, which are really delicious, have blueberries in them.\\
-B: Yes, that's true, but they aren't really delicious.
+\newpage
+
+The at-issueness measures that did not work out as expected are illustrated in the examples in (3), (4), and (5).
+In (3), B responds in the affirmative to A's question, but then denies the truth of the NRRC.
+The group mean is comparatively higher (at .28) than for the other `q' experiments, including Exp.~2q.
+This result suggests that B's affirmative response to A's question is taken by the participants to affirm not only the main clause content, but also the content of the NRRC.
+Subsequent denial of the truth of the NRRC results in degradation in acceptability.
+
+\begin{exe}
+\exi{(3)} Sample control stimulus in Exp.~3q
+\begin{xlist}
+\exi{A:} Do these muffins, which are really delicious, have blueberries in them?
+\exi{B:} Yes, but they aren't really delicious.
+\end{xlist}
+\end{exe}
+
+In (4), B assents with A's indicative assertion, and then re-affirms the truth of the main clause content.
+The group mean is comparatively higher (at .22 or .25) than for Exp.~2q.
+This result suggests that responding in the affirmative to A's question is a different speech act than assenting with A's indicative assertion: whereas it does not appear to be odd, in the latter case, to spell out the propositional antecedent of {\em yes}, it does seem to be odd to do so in the latter case.
+
+\begin{exe}
+\exi{(4)} Sample control stimulus in Exps.~2n, 2m, and 2c
+\begin{xlist}
+\exi{A:} These muffins, which are really delicious, have blueberries in them.
+\exi{B:} Yes, that's true, they have blueberries in them.
+\end{xlist}
+\end{exe}
+
+In (5), B assents with A's indicative assertion, but then denies the truth of the NRRC.
+Here, the group mean is quite a bit higher (at .44, .50, .53) than for Exp.~3q or for Exps.~2n, 3m, and 3c.
+This result suggests, again, that responding in the affirmative to A's question is a different speech act than assenting with A's indicative assertion: while denying the truth of the NRRC after responding in the affirmative to the question is already odd (Exp.~3q), denying the truth of the NRRC after assenting with A's assertion with the NRRC is worse.
+
+\begin{exe}
+\exi{(5)} Sample control stimulus in Exps.~3n, 3m, and 3c
+\begin{xlist}
+\exi{A:} These muffins, which are really delicious, have blueberries in them.
+\exi{B:} Yes, that's true, but they aren't really delicious.
+\end{xlist}
+\end{exe}
+
+The responses to the control stimuli with NRRCs that used either one of the assent diagnostics suggest that assenting with a speaker's indicative utterance that includes an NRRC includes the content of the NRRC.
+Two possibilities: either the content of the NRRC is already entered into the CG when A asserts the sentence (thus making reaffirmation or denial odd); or B's {\em yes, that's true} assents not just with the main clause content, but also (though perhaps a bit less?) with the content of the NRRC (thereby making reaffirmation or denial odd).
 

### 2022-08-12 notes lisa/notes.tex
include graphs for data from conditionals

@@
 
+\section{Experiments with Condstionals}
+
+	\subsection{1c---Conditional embedding,  are you sure?, assent + assertion of complement + Q-A judgment}
+		\ex. \a. \textbf{Christopher:} If Cole discovered that Julian dances salsa, Logan will be joyful.
+			\b. \textbf{Sandy:} Are you sure?
+			\b. \textbf{Christopher:} Yes, I'm sure that Julian dances Salsa.
+			\z.
+		\z.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[scale=0.85]{figures/c1-correl.pdf}
+			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 1c}
+			\label{fig:m1-corr}
+		\end{figure}
+
+	\subsection{2c---Conditional embedding yes, that's true+complement diagnostic (+ acceptability rating)} % (fold)
+		\ex. \a. \textbf{Christopher:} If Cole discovered that Julian dances salsa, Logan will be joyful.
+			\b. \textbf{Sandy:} Yes, that’s true, Julian dances salsa.
+			\z.
+		\z.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[scale=0.85]{figures/c2-correl.pdf}
+			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 2c}
+			\label{fig:m1-corr}
+		\end{figure}
+
+	\subsection{3c---Conditional embedding, yes, that's true + negation of main clause (+ acceptability rating)} % (fold)
+		\ex. \a. \textbf{Christopher:} If Cole discovered that Julian dances salsa, Logan will be joyful.
+			\b. \textbf{Sandy:} Yes, that’s true, but Cole didn't discover it.
+			\z.
+		\z.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[scale=0.85]{figures/c3-correl.pdf}
+			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. 3c}
+			\label{fig:m1-corr}
+		\end{figure}
+
 \pagebreak

### 2022-08-12 results/main/12_projaiC/rscripts/graphs.R

@@ -1,0 +1,180 @@
+# 12_projaiC
+# conditional: projection and at-issueness (assent with positive continuation)
+# graphs
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load required packages
+require(tidyverse)
+library(ggrepel)
+library(dichromat)
+library(forcats)
+library(ggrepel)
+theme_set(theme_bw())
+
+# load helper functions
+source('../../helpers.R')
+
+d = read_csv("../data/data_preprocessed.csv")
+
+# spread responses over separate columns for projectivity and at-issueness
+t = d %>%
+  mutate(block_ai = ifelse(question_type == "ai", ifelse(block == "block1", "block1", "block2"), ifelse(block == "block1", "block2", "block1"))) %>%
+  select(workerid,content,short_trigger,question_type,response,block_ai) %>%
+  spread(question_type,response)
+
+cd = t
+
+table(cd$short_trigger)
+# change cd verb names to match veridicality names
+cd = cd %>%
+  mutate(verb=recode(short_trigger, control = "MC", annoyed = "be_annoyed", be_right_that = "be_right", inform_Sam = "inform"))
+
+# mean not-at-issueness by predicate, including the main clause controls
+ai.means = cd %>%
+  group_by(short_trigger) %>%
+  summarize(Mean = mean(ai), CILow = ci.low(ai), CIHigh = ci.high(ai)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, verb = fct_reorder(as.factor(short_trigger),Mean))
+ai.means
+
+# define colors for the predicates
+cols = data.frame(V=levels(ai.means$verb))
+
+cols$VeridicalityGroup = as.factor(
+  ifelse(cols$V %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
+         
+levels(cols$V)
+cols$V <- factor(cols$V, levels = cols[order(as.character(ai.means$verb)),]$V, ordered = TRUE)
+
+cols$Colors =  ifelse(cols$VeridicalityGroup == "F", "tomato1", "black")
+                      
+cols$Colors
[…]

### 2022-10-21 writing/investigating-GPP.tex
changed paper structure
updated section 2
testing rebase
added notes from 10/21 meeting

@@
 
+\newpage
+
+\tableofcontents
+
+\newpage
+
 \section{Introduction}\label{s1}
@@
 
-\section{Analyzing and diagnosing at-issueness}\label{s2}
-
 \subsection{Formal analyses of at-issueness}
@@
 
-- Djärv, Kajsa. 2019. Factive and assertive attitude reports. Philadelphia, PA: Uni- versity of Pennsylvania dissertation. https://repository.upenn.edu/edissertations/ 3645.
+- Dj\"arv, Kajsa. 2019. Factive and assertive attitude reports. Philadelphia, PA: Uni- versity of Pennsylvania dissertation. https://repository.upenn.edu/edissertations/ 3645.
 
@@
 
-\section{Experiments 1-3: Methods}\label{s3}
+\section{Experiments 1-3: Methods}\label{s2}
+
+hallo
 
@@
 Projection was measured using the ?certain that?
-diagnostic (Djärv \& Bacovcin, 2017; Lorson, 2018; Mahler, 2020; Tonhauser et al., 2018): participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
+diagnostic (Dj\"arv \& Bacovcin, 2017; Lorson, 2018; Mahler, 2020; Tonhauser et al., 2018): participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
 They gave their responses on a slider marked ?no?
@@
 
-\section{Experiments 1-3: Results}\label{s4}
-
-\subsection{Comparing at-issueness}
-
-\subsection{GPP}
-
-Further analyses:
-
+\section{Projection}
+
+Compare projectivity ratings across entailment-cancelling operators. 
+\begin{itemize}
+  \item is there a difference between projectivity across contexts? 
+  \item is there a difference between verbs in terms of how much embedding context makes a difference (semi-factive v factive)?
+  \item previous point that non-factive (i.e. veridical) predicates can also project can be made here for other embedding operators as well
+\end{itemize}
+
+\section{At-issueness}
+
+At-issueness diagnostics (kind of) work for questions:
+\begin{itemize}
+  \item asking whether diagnostic (exp. 1q)
+  \item Q-A diagnostic (exp. 2q)
+\end{itemize}
+
+\section{GPP}
+
+Replicating results from \cite{tbd-variability}: Correlation between NAI \& projection for PolQs found for:
+\begin{itemize}
+  \item asking whether diagnostic (exp. 1q)
+  \item Q-A diagnostic (exp. 2q)
+\end{itemize}
+
+\noindent Both of these also show some interesting lexical exceptions:
+\begin{itemize}
+  \item non-veridical predicates
+  \item some specific kinds of reportatives
+  \item how can we systematically identify exceptions based on data analysis?
+  \item are these the same kinds of exceptions as in the glossa paper?
+\end{itemize}
+
+\noindent Further analyses:
 \begin{itemize}

### 2022-10-26 writing/investigating-GPP.tex
updated paper section 2

@@
 
+projection of contents embedded under different entailment-canceling operators is positively correlated, as assumed in the semantics/pragmatics literature (e.g., REFERENCES).
+Prior experimental literature: \citealt{smith-hall11,smith-hall-cls,demarneffe-etal-sub23}.
+
+\newpage
+
 \section{Experiments 1-3: Methods}\label{s2}
 
-hallo
-
-Exps.~1-3 were designed to investigate whether not-at-issueness and projection are positively correlated, as predicted by the Gradient Projection Principle, as well as to compare various measures of at-issueness.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
+Exps.~1-3 were designed to investigate whether not-at-issueness and projection are positively correlated, as predicted by the Gradient Projection Principle, to compare the projection of contents from different entailment-canceling environments, and to compare different measures of at-issueness.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
 All experiments were conducted with IRB approval.
-Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.}
-
-To measure projection, Exps.~1-3 employed the `certain that' diagnostic for projection, which is suitable to measure projection for both polar questions and declarative sentences (see also, e.g.\ \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of the utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) occurs embedded under an entailment-canceling operator:
-
-\begin{exe}
-
+Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.} To be able to compare across the experiments, the experiments investigated the same contents and each of the experiments consisted of two blocks: a projection block, in which the projection of the contents was measured, and an at-issueness block, in which the at-issueness of the contents was measured.
+The at-issueness blocks of Exps.~1-3 implemented different measures of at-issueness, which allows for a comparison of at-issueness measures.
+To compare projection from different entailment-canceling environments, each of Exps.~1-3 consisted of four experiments (for a total of 12 experiments) where the contents were realized by expressions in a polar question, in the scope of sentential negation, in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.
+For an overview of the 12 experiments see Table \ref{t:overview}: the name of each of the 12 experiments identifies whether belongs to the set of Exps.~1, 2, or 3 (which differ in at-issueness measure) and whether the embedding-environment is a polar question (`q'), negation (`n'), the epistemic possibility modal (`m'), or the antecedent of a conditional (`c').
+
+Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of Daniel's utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) occurs in one of the four entailment-canceling environments:
+
+\begin{exe}
 \ex\label{proj-stim} `certain that' diagnostic for projection
 \begin{xlist}
-\ex Daniel: Did Cole discover that Julian dances salsa?
-\ex Daniel: Cole didn't discover that Julian dances salsa.
-\ex Daniel: Perhaps Cole discovered that Julian dances salsa.
-\ex Daniel: If Cole discovered that Julian dances salsa, Logan will be joyful.
-\end{xlist}
-\end{exe}
-To assess projection, participants are asked whether the speaker is certain of the content; for instance, in \ref{proj-stim}, participants are asked whether Daniel is certain that Julian dances salsa.
-If a participant takes the speaker to be certain of the content, we assume that the content projects; if a participant does not take the speaker to be certain of the content, we assume that the content does not project.
-Following \citealt{tbd-variability}, the `certain that' diagnostic was implemented in Exps.~1-3 with a gradient response scale: participants gave their responses on a slider marked `no' at one end and `yes' at the other.
-We assume that the closer to `yes' a participant's response is, the more the speaker is certain of the content, that is, the more projective the content is.
-In other words, we assume that gradient certainty ratings reflect gradience in the speaker's commitment to the truth of the content.\footnote{Strictly speaking,  gradient certainty ratings reflect gradience in the degree to which participants \emph{perceive} the speaker to be committed to the truth of the content.
-That is, as in any experiment, the quantity of interest, in this case speaker commitment, is only indirectly measured.} As discussed in \citealt{tbd-variability}, a second interpretation of gradient certainty ratings is that they reflect a participant's degree of belief in the speaker's (binary) commitment to the truth of the content.
-While we adopt the first interpretation in our discussion, we remain agnostic about the interpretation of gradient certainty ratings.
-
-At-issueness was measured with different diagnostics across Exps.~1-3. For polar questions (Exps.~1q, 2q, 3q), at-issueness was assessed with the three measures in (\ref{ai-question}):
-
-\begin{exe}
-\ex\label{ai-question} At-issueness diagnostics for polar questions
-\begin{xlist}
-\ex `asking whether $c$' 
-\\ {\bf Christopher:} Did Cole discover that Julian dances salsa?
-\\ Is Christopher asking whether Julian dances salsa?
-
+\ex Daniel: Did Cole discover that Julian dances salsa? \hfill [Exps.~1q, 2q, 3q]
+\ex Daniel: Cole didn't discover that Julian dances salsa. \hfill [Exps.~1n, 2n, 3n]
+\ex Daniel: Perhaps Cole discovered that Julian dances salsa. \hfill [Exps.~1m, 2m, 3m]
+\ex Daniel: If Cole discovered that Julian dances salsa, Logan will be joyful. \hfill [Exps.~1c, 2c, 3c]
+\end{xlist}
+Question to participants: Is Daniel certain that Julian dances salsa?
+\end{exe}
+To measure projection, participants are asked whether the speaker is certain of the relevant content.
+In \ref{proj-stim}, for instance, participants are asked whether Daniel is certain that Julian dances salsa.
+If a participant takes the speaker to be certain of the content, we assume that the content projects out of the entailment-canceling environment; if a participant does not take the speaker to be certain of the content, we assume that the content does not project.
+
+At-issueness was measured with different diagnostics across Exps.~1-3.
+In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c.
+
+\begin{exe}
+\ex\label{ai-exps1} At-issueness measures in Exps.~1
+\begin{xlist}
+\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
+\\ Question to participants: Is Christopher asking whether Julian dances salsa?
+
+\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
+\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
+\\ Sandy: Are you sure?
+\\ Christopher: Yes, I'm sure that Julian dances salsa.
+\\ Question to participants: Did Christopher answer Sandy's question?
+
+\end{xlist}
+\end{exe}
+
+In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed.
+Participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
+The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps2}a)) or assents (for declaratives, as shown in (\ref{ai-exps2}b).
+The affirmation/assent is followed by the content to be diagnosed, here that Julian dances salsa.
+
+\begin{exe}
+\ex\label{ai-exps2} At-issueness measures in Exps.~2
+\begin{xlist}
 \ex {\em yes}, $c$
-\\ {\bf Christopher:} Did Cole discover that Julian dances salsa?
-\\ {\bf Sandy:} Yes, Julian dances salsa.
-\\ Does Sandy's response to Christoper sound good?
-
-\ex {\em yes, but} $\neg c'$
-\\ {\bf Christopher:} Did Cole discover that Julian dances salsa?
-\\ {\bf Sandy:} Yes, but Cole didn't discover it. 
-\\ Does Sandy's response to Christoper sound good?
-
-\end{xlist}
-\end{exe}
-
-For declaratives, at-issueness was assessed with the three measures in (\ref{ai-decl}):
-
-\begin{exe}
-\ex\label{ai-decl} At-issueness diagnostics for declaratives
-
-\begin{xlist}
-
-\ex `sure that $c$'
-\\ {\bf Christopher:} Cole didn't discover that Julian dances salsa.
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.
-\\ {\bf Sandy:} Are you sure?
-\\ {\bf Christopher:} Yes, I'm sure that Julian dances salsa.
-\\ Did Christopher answer Sandy's question?
+\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
+\\ Sandy: Yes, Julian dances salsa.
+\\ Question to participants: Does Sandy's response to Christoper sound good?
 
 \ex {\em yes, that's true}, $c$
-\\ {\bf Christopher:} Cole didn't discover that Julian dances salsa.
-\\ {\bf Sandy:} Yes, that's true, Julian dances salsa.
-\\ Does Sandy's response to Christopher sound good?
-
-\ex {\em yes, that's true, but} $\neg c'$
-\\ {\bf Christopher:} Cole didn't discover that Julian dances salsa.
-\\ {\bf Sandy:} Yes, that's true, but Cole discovered it. 
-\\ Does Sandy's response to Christopher sound good?
-
-\end{xlist}
-
-\end{exe}
+\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
+\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
+\\ Sandy: Yes, that's true, Julian dances salsa.
+\\ Question to participants: Does Sandy's response to Christopher sound good?
+
+\end{xlist}
+\end{exe}
+Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
+If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
+If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
+
+In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content.
+As in the other assent diagnostic, participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
+The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps3}a)) or assents (for declaratives, as shown in (\ref{ai-exps3}b).
+The affirmation/assent is followed by the denial of another salient content, here, that Cole (didn't) discover it.
+
+\begin{exe}
+\ex\label{ai-exps3} At-issueness measures in Exps.~3
+\begin{xlist}
+\ex {\em yes}, $c$
+\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
+\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
+\\ Question to participants: Does Sandy's response to Christoper sound good?
+
+\ex {\em yes, that's true}, $c$
+\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
+\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
+\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
+\\ Question to participants: Does Sandy's response to Christopher sound good?
+
+\end{xlist}
+\end{exe}
+As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
+If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
+If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
 
@@
 \end{tabular}
-\caption{Overview of the entailment-canceling embeddings, projection measures, and at-issueness measures in Exps.~1-3, where $c$ is the content under investigation (in the target trials: the content of the clausal complement) and $c'$ is a different, salient content (in the target trials: the main clause content).}\label{t-overview}
+\caption{Overview of Exps.~1-3, which differ in the entailment-canceling embedding (q = polar question, n = negation, m = epistemic possibility modal, c = antecedent of conditional) and measures for not-at-issueness.
+$c$ is the content under investigation (in the target trials: the content of the clausal complement) and $c'$ is a different, salient content (in the target trials: the main clause content).}\label{t:overview}
 \end{table}
-
-{\bf Comparisons for at-issueness:}
-
-\begin{itemize}
-
-\item Across Exps 1q, 2q, 3q (same assumption about at-issue content partitioning context set; reveals effect of embedding)
-
-\item Across Exps n, m, c (same measure of at-issueness in Exps 1, 2, 3; reveals effect of embedding)
-
-\item 2q -- 2n, 2m, 2c (yes...$c$; i.e., anaphor with $c$)
-
-\item 3q -- 3n, 3m, 3c (yes...but $\neg c$, i.e., anaphor with $c$)
-
-\end{itemize}
-
-\footnote{Exps.~1-3 also allow us to address the question of whether the projection of contents embedded under different entailment-canceling operators is positively correlated, as assumed in the semantics/pragmatics literature (e.g., REFERENCES).
-Prior experimental literature: \citealt{smith-hall11,smith-hall-cls,demarneffe-etal-sub23}.
-As shown in Supplement \ref{a-projection}, this is the case.}
 
@@
 
-Exps.~1-3 investigated the projection and at-issueness of the CC of the 20 clause-embedding predicates in (\ref{pred}).
-The clausal complements of the 20 predicates were realized by 20 clauses (provided in Supplement \ref{a-clauses}), for a total of 400 predicate/clause combinations.
-These 400 predicate/clause combinations were combined with proper name subjects (a unique name per complement clause) to realize polar questions (in Exps.
-1q, 2q, and 3q), negated declaratives (in Exps.
-1n, 2n, and 3n), declaratives with the modal adverb {\em perhaps} (in Exps.
-1m, 2m, and 3m), and the antecedents of conditionals (in Exps.
-1c, 2c, and 3c).
-These sentences were presented as utterances by a named speaker.
-A sample set of target utterances is given in (\ref{sample-stims}): here, the named speaker is {\em Daniel}, the subject of the clause-embedding predicate is {\em Cole}, the predicate is {\em discover}, and the complement clause is {\em Julian dances salsa}.\footnote{The indirect object of {\em inform} was realized by the proper name {\em Sam}.
+The projection and at-issueness of the contents of 20 clauses (such as {\em Julian dances salsa}) were measured in separate blocks in Exps.~1-3.
+For the full set of clauses, see Appendix \ref{a-clauses}.
+
+In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}.
+Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique.
+The indirect object of {\em inform} was realized by the proper name {\em Sam}.
 The predicates differed in the tense in which they were realized.
@@
 In the antecedents of conditionals, all predicates were realized in the present tense.
-For the consequents of the conditional sentences see Supplement \ref{a-target}.}
-
-\begin{exe}
-\ex\label{sample-stims}
-\begin{xlist}
-\ex {\bf Daniel:} ``{\em Did Cole discover that Julian dances salsa?}"
-\ex {\bf Daniel:} ``{\em Cole didn't discover that Julian dances salsa.}"
-\ex {\bf Daniel:} ``{\em Perhaps Cole discovered that Julian dances salsa.}"
-\ex {\bf Daniel:} ``{\em If Cole discovered that Julian dances salsa, Logan will be joyful.}"
-\end{xlist}
-\end{exe}
-The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique.
-
-Projection and at-issueness of the contents of the clausal complements were measured in separate blocks.
-In the projection blocks across Exps.~1-3, projection was measured with the `certain that' diagnostic.
-The target stimuli consisted of the target utterances, as shown in Figure \ref{f-projection-trials}.
+For the consequents of the conditional sentences see Supplement \ref{a-target}.} Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question.
+Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
+They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1).
+Greater speaker commitment to the CC should result in higher slider ratings.
+
+%Following \citealt{tbd-variability}, the `certain that' diagnostic was implemented in Exps.~1-3 with a gradient response scale: participants gave their responses on a slider marked `no' at one end and `yes' at the other.
+We assume that the closer to `yes' a participant's response is, the more the speaker is certain of the content, that is, the more projective the content is.
+In other words, we assume that gradient certainty ratings reflect gradience in the speaker's commitment to the truth of the content.\footnote{Strictly speaking,  gradient certainty ratings reflect gradience in the degree to which participants \emph{perceive} the speaker to be committed to the truth of the content.
+That is, as in any experiment, the quantity of interest, in this case speaker commitment, is only indirectly measured.} As discussed in \citealt{tbd-variability}, a second interpretation of gradient certainty ratings is that they reflect a participant's degree of belief in the speaker's (binary) commitment to the truth of the content.
+While we adopt the first interpretation in our discussion, we remain agnostic about the interpretation of gradient certainty ratings.
+
+\begin{figure}[h!]
+\centering
+\fbox{\begin{minipage}{15cm}{\em acknowledge, admit, announce, be annoyed, be right, confess, confirm, demonstrate, discover, establish, hear, inform, know, pretend, prove, reveal, say, see, suggest, think.
+}\end{minipage}}
+\caption{20 clause-embedding predicates in Exps.~1-3.}\label{fig:predicates}
+\end{figure}
 
@@
 \end{subfigure}
-
-\caption{Target trials in projection blocks of Exps.~1, 2, and 3 for the complement {\em Julian dances salsa}.}\label{f-projection-trials}
+\caption{Target trials in projection blocks of Exps.~1, 2, and 3 for the complement {\em Julian dances salsa}.}\label{fig:projection-trials}
 \end{figure}
+
+In the at-issueness blocks of Exps.~1, target
+
+In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}.
+Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique.
+The indirect object of {\em inform} was realized by the proper name {\em Sam}.
+The predicates differed in the tense in which they were realized.
+In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.
+In the antecedents of conditionals, all predicates were realized in the present tense.
+For the consequents of the conditional sentences see Supplement \ref{a-target}.} Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question.
+Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
+They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1).
+Greater speaker commitment to the CC should result in higher slider ratings.
 
@@
 \fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1q-ai}}
-\caption{Exp.~1q.}\label{fig-exp1q-ai}
+\caption{Exp.~1q.}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1n-ai}} 
+\caption{Exp.~1n.}
  \end{subfigure}
@@
         \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1m-ai}}
+\caption{Exp.~1m}
+ \end{subfigure}%
+ \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1c-ai}}
+\caption{Exp.~1c}
+ \end{subfigure}
+
+\caption{Target trials in at-issueness blocks of Exps.~1 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps1}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
 \fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2q-ai}}
-\caption{Exp.~2q.}\label{fig-exp2q-ai}
+\caption{Exp.~2q.}
 \end{subfigure}%
@@
 \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2n-ai}} 
+\caption{Exp.~2n.}
+ \end{subfigure}
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2m-ai}}
+\caption{Exp.~2m}
+ \end{subfigure}%
+ \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2c-ai}}
+\caption{Exp.~2c}
+ \end{subfigure}
+
+\caption{Target trials in at-issueness blocks of Exps.~2 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps3}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
 \fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3q-ai}}
-\caption{Exp.~3q}\label{fig-exp3q-ai}
+\caption{Exp.~3q.}
+\end{subfigure}%
+\begin{subfigure}[t]{0.5\textwidth}
+\centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3n-ai}} 
+\caption{Exp.~3n.}
  \end{subfigure}
-
-\caption{Target trials in at-issueness blocks of Exps.~1q, 2q, and 3q for the complement {\em Julian dances salsa}.}\label{f-ai-trialsq}
+\begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3m-ai}}
+\caption{Exp.~3m}
+ \end{subfigure}%
+ \begin{subfigure}[t]{0.5\textwidth}
+        \centering
+\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3c-ai}}
+\caption{Exp.~3c}
+ \end{subfigure}
+
+\caption{Target trials in at-issueness blocks of Exps.~3 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps3}
 \end{figure}
 
-\begin{figure}[h!]
-\centering
-
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2n-ai}}
-\caption{Exp.~2n.}\label{fig-exp1q-ai}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2m-ai}} 
-\caption{Exp.~2m.}\label{fig-exp2q-ai}
- \end{subfigure}
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=9cm]{figures/2c-ai}}
-\caption{Exp.~2c}\label{fig-exp3q-ai}
- \end{subfigure}
-
-\caption{Target trials in at-issueness blocks of Exps.~2n, 2m, and 2c for the complement {\em Julian dances salsa}.}\label{f-ai-trialsq}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3n-ai}}
-\caption{Exp.~3n.}\label{fig-exp1q-ai}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3m-ai}} 
-\caption{Exp.~3m.}\label{fig-exp2q-ai}
- \end{subfigure}
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=9cm]{figures/3c-ai}}
-\caption{Exp.~3c}\label{fig-exp3q-ai}
- \end{subfigure}
-
-\caption{Target trials in at-issueness blocks of Exps.~3n, 3m, and 3c for the complement {\em Julian dances salsa}.}\label{f-ai-trialsq}
-\end{figure}
-
-In the projection block, target stimuli consisted of a fact and a polar question that was ut- tered by a named speaker, as shown in Figure 1B.
-The polar questions were formed by real- izing the 20 clauses as the complements of the 20 clause-embedding predicates in Figure 1C.
-Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question.
-Projection was measured using the ?certain that?
-diagnostic (Dj\"arv \& Bacovcin, 2017; Lorson, 2018; Mahler, 2020; Tonhauser et al., 2018): participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
-They gave their responses on a slider marked ?no?
-at one end (coded as 0) and ?yes?
-at the other (coded as 1).
-Greater speaker commitment to the CC should result in higher slider ratings.
-
-Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask a question.
-Participants were asked to rate whether the speaker was certain of the CC.
-They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1), as shown in Figure \ref{fig-trial-exp1}.
-
-After completing the experiment, participants filled out a short, optional survey about their age, their gender, their native language(s) and, if English is their native language, whether they are a speaker of American English (as opposed to, e.g.\ Australian or Indian English).
+====
+
+The projection block also included 6 control trials, which functioned as attention checks.
+The content of these stimuli was expected not to project: For example, in Fig.~\ref{fig-exp1-projection-control}, the speaker is not committed to the main clause content, that Samantha has a new hat.
+The same 6 main clauses were also used to form 6 filler trials in the prior block; a sample stimulus is given in Fig.~\ref{fig-exp1-prior-filler}.
+These filler stimuli were not used to assess participants' attention.
+For the full set of stimuli see Supplement \ref{a-stim}.
+
+Each participant's stimulus set was semi-randomly generated by first randomly pairing up the 20 predicates and clauses.
+Half of the stimuli were then randomly assigned the respective clause's higher-probability fact, and half its lower-probability fact.
+Participants completed a total of 52 trials: 20 target trials in each block, 6 control trials in the projection block, and 6 filler trials in the prior block.
+Each participant completed the same 6 filler and control trials.
+Block order and within-block trial order were randomized.
+
+After completing the experiment, participants filled out a short optional demographic survey.
 To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.
@@
 Information on the participants whose data entered into the analysis (total number, ages, gender) can be found in Supplement \ref{a-participants}.
+
+\newpage
 
@@
 
-\section{Data exclusion}\label{a-participants}
-
-This supplement provides information on the recruited participants, the criteria by which participants' data were excluded, the remaining participants (that is, the participants whose data entered into the analysis), and the payment for each of Exps.~1-3.
-Table \ref{f-exclusion} provides information on the recruited and remaining participants, including the total number, the range of the participants' ages, the mean ages of the participants, and their self-reported gender (`f' = female, `m' = male, `o' = other, `u' = undeclared).
+\section{Participant information and data exclusion criteria}\label{a-participants}
+
+This supplement provides information on the participants of the 12 experiments and the criteria by which participants' data were excluded.
+The columns of Table \ref{t:exclusion} show, for each of the 12 experiments (Exp.~1q, \ldots, Exp.~3c) how many participants were recruited, their age range and mean age, and their self-reported gender (`f' = female, `m' = male, `o' = other, `u' = undeclared).
 No gender data was collected in Exp.~1q.
@@
 
-\item `multiple': Due to an experimental glitch, some participants participated more than once.
-Since no information was available on which one was their first take, those participants' data was removed (which means that for each participant who took the experiment twice, at least two data sets were removed).
+\item `multiple': Due to an experimental glitch, some participants participated more than once in Exp.~1q.
+Since no information was available on which one was their first take, those participants' data was removed.
 
@@
 \item `controls': Participants' data were excluded if their mean rating on the 6 main clause control items in the projection block was more than 2 sd above the group mean (in Exps.~1q, 2q, and 3q) or more than 2 sd below the group mean (in the remaining experiments).
-Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across Exps.~1-3).
-
-\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli.
-We identified such participants by identifying participants whose mean variance on the target stimuli was more than 2sd below the group mean variance on the target stimuli, and manually inspecting their response patterns.
-%Exps 1q, 2q, 3q: participants with low variance not excluded after manual inspection
-\end{itemize}
-
+Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments).
+
+\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli, that is, if their mean variance on the target stimuli was more than 2 sd below the group mean variance.
+{\bf and manually inspecting their response patterns.
+Exps 1q, 2q, 3q: participants with low variance not excluded after manual inspection NEED TO CONSIDER THIS AGAIN}
+\end{itemize}
+
+The remaining columns of Table \ref{t:exclusion} provide information on the remaining participants, that is, the participants' data that entered into the analysis.
 Participants took around 9-11 minutes to complete the various experiments.
 Participants were paid more in Exps.~1c, 2c, and 3c than the remaining experiments because the target stimuli in those experiments were longer (as they consisted of conditionals).
-More women than men were recruited in many of the experiments because they were run at a time when Prolific apparently went viral on TikTok, resulting in a large number of young women registering for the service (around July 24, 2021; see \url{https://blog.prolific.co/we-recently-went-viral-on-tiktok-heres-what-we-learned/}, \\ last accessed February 4, 2022).
+More women than men were recruited in many of the experiments because the experiments were run at a time when Prolific went viral on TikTok, resulting in a large number of young women registering for the service (around July 24, 2021; see \url{https://blog.prolific.co/we-recently-went-viral-on-tiktok-heres-what-we-learned/}, \\ last accessed February 4, 2022).
 
@@
 \end{tabular}
-\caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3}\label{f-exclusion}
+\caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3}\label{t:exclusion}
 \end{sidewaystable}

### 2022-10-27 writing/investigating-GPP.tex
worked on section 2 and appendices

@@
 
+\subsection{At-issueness diagnostics}\label{s:ai-diagnostics}
+
+In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q, illustrated in (\ref{ai-exps1}a), and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, illustrated in (\ref{ai-exps1}b).
+The `asking whether' diagnostic requires that the content to be diagnosed is realized by an expression in a polar question; this diagnostic was also used in \citealt{tbd-variability} Exps.~1.
+A speaker (Christopher in (\ref{ai-exps1}a) asks a polar question and the participant is asked whether the speaker is asking about the content to be diagnosed (here, that Julian dances salsa).
+If a participant takes the speaker to be asking of the content, the content is assumed to be at-issue; if a participant does not take the speaker to be asking about the content, the content is assumed to be not-at-issue.
+The `sure that' diagnostic requires that the content to be diagnosed is realized by an expression in declarative sentence; this diagnostic was also used in \citealt{tbd-variability} Exps.~2.
+A speaker utters the declarative sentence and the addressee (here, Sandy) asks {\em Are you sure?}.
+The speaker response with {\em Yes, I am sure that} followed by the expression that denotes the content to be diagnosed.
+Participants are asked whether the speaker answered the addressee's question.
+If a participant takes the speaker to have answered the question, the content is assumed to be at-issue; if a participant does not take the speaker to have answered the question, the content is assumed to be not-at-issue.
+{\bf move all this into section 1}
+
+\begin{exe}
+\ex\label{ai-exps1} At-issueness measures in Exps.~1
+\begin{xlist}
+\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
+\\ Question to participants: Is Christopher asking whether Julian dances salsa?
+
+\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
+\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
+\\ Sandy: Are you sure?
+\\ Christopher: Yes, I'm sure that Julian dances salsa.
+\\ Question to participants: Did Christopher answer Sandy's question?
+
+\end{xlist}
+\end{exe}
+
+In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed.
+Participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
+The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps2}a)) or assents (for declaratives, as shown in (\ref{ai-exps2}b).
+The affirmation/assent is followed by the content to be diagnosed, here that Julian dances salsa.
+
+\begin{exe}
+\ex\label{ai-exps2} At-issueness measures in Exps.~2
+\begin{xlist}
+\ex {\em yes}, $c$
+\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
+\\ Sandy: Yes, Julian dances salsa.
+\\ Question to participants: Does Sandy's response to Christoper sound good?
+
+\ex {\em yes, that's true}, $c$
+\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
+\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
+\\ Sandy: Yes, that's true, Julian dances salsa.
+\\ Question to participants: Does Sandy's response to Christopher sound good?
+
+\end{xlist}
+\end{exe}
+Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
+If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
+If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
+
+In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content.
+As in the other assent diagnostic, participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
+The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps3}a)) or assents (for declaratives, as shown in (\ref{ai-exps3}b).
+The affirmation/assent is followed by the denial of another salient content, here, that Cole (didn't) discover it.
+
+\begin{exe}
+\ex\label{ai-exps3} At-issueness measures in Exps.~3
+\begin{xlist}
+\ex {\em yes}, $c$
+\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
+\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
+\\ Question to participants: Does Sandy's response to Christoper sound good?
+
+\ex {\em yes, that's true}, $c$
+\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
+\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
+\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
+\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
+\\ Question to participants: Does Sandy's response to Christopher sound good?
+
+\end{xlist}
+\end{exe}
+As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
+If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
+If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
+
 \subsection{Formal analyses of at-issueness}
@@
 
-Exps.~1-3 were designed to investigate whether not-at-issueness and projection are positively correlated, as predicted by the Gradient Projection Principle, to compare the projection of contents from different entailment-canceling environments, and to compare different measures of at-issueness.\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
+Exps.~1-3 were designed to investigate the three research questions in (\ref{rqs}), repeated here for convenience:\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
 All experiments were conducted with IRB approval.
-Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.} To be able to compare across the experiments, the experiments investigated the same contents and each of the experiments consisted of two blocks: a projection block, in which the projection of the contents was measured, and an at-issueness block, in which the at-issueness of the contents was measured.
-The at-issueness blocks of Exps.~1-3 implemented different measures of at-issueness, which allows for a comparison of at-issueness measures.
-To compare projection from different entailment-canceling environments, each of Exps.~1-3 consisted of four experiments (for a total of 12 experiments) where the contents were realized by expressions in a polar question, in the scope of sentential negation, in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.
-For an overview of the 12 experiments see Table \ref{t:overview}: the name of each of the 12 experiments identifies whether belongs to the set of Exps.~1, 2, or 3 (which differ in at-issueness measure) and whether the embedding-environment is a polar question (`q'), negation (`n'), the epistemic possibility modal (`m'), or the antecedent of a conditional (`c').
-
-Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of Daniel's utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) occurs in one of the four entailment-canceling environments:
+Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.}
+
+\begin{exe}
+\ex\label{rqs} Research questions
+\begin{xlist}
+\ex Does the projection of content differ across entailment-canceling environments?
+\ex Are the results of different measures of at-issueness positively correlated?
+\ex Is the projection of content positively correlated with its not-at-issueness, as predicted by \citepos{tbd-variability} Gradient Projection Principle?
+\end{xlist}
+\end{exe}
+To address these questions, Exps.~1-3 investigated the projection and at-issueness of same 20 contents.
+Each of the experiments consisted of two blocks: a projection block, in which the projection of the contents was measured, and an at-issueness block, in which the at-issueness of the contents was measured.
+To address the research question in (\ref{rqs}b), different measures of at-issueness were implemented in the at-issueness blocks of Exps.~1-3, respectively.
+To address the research question in (\ref{rqs}a), each of Exps.~1-3 consisted of four experiments (for a total of 12 experiments) where the contents were realized by expressions in a polar question, in the scope of sentential negation, in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.
+The research question (\ref{rqs}c) is addressed by comparing the projection and at-issueness ratings of each of the 12 experiments.
+For an overview of the 12 experiments see Table \ref{t:overview}, where the name of each experiment identifies whether belongs to the set of Exps.~1, 2, or 3 (which differ in at-issueness measures) and whether the embedding-environment is a polar question (`q'), negation (`n'), the epistemic possibility modal (`m'), or the antecedent of a conditional (`c').
+
+Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of Daniel's utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) is realized by an expression that occurs in one of the four entailment-canceling environments, that is, in a polar question (\ref{proj-stim}a), in the scope of negation (\ref{proj-stim}b), in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.
 
@@
 In \ref{proj-stim}, for instance, participants are asked whether Daniel is certain that Julian dances salsa.
-If a participant takes the speaker to be certain of the content, we assume that the content projects out of the entailment-canceling environment; if a participant does not take the speaker to be certain of the content, we assume that the content does not project.
-
-At-issueness was measured with different diagnostics across Exps.~1-3.
-In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c.
-
-\begin{exe}
-\ex\label{ai-exps1} At-issueness measures in Exps.~1
-\begin{xlist}
-\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
-\\ Question to participants: Is Christopher asking whether Julian dances salsa?
-
-\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
-\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
-\\ Sandy: Are you sure?
-\\ Christopher: Yes, I'm sure that Julian dances salsa.
-\\ Question to participants: Did Christopher answer Sandy's question?
-
-\end{xlist}
-\end{exe}
-
-In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed.
+If a participant takes the speaker to be certain of the content, the content is assumed to project out of the entailment-canceling environment; if a participant does not take the speaker to be certain of the content, the content is assumed to not project.
+
+As shown in Table \ref{t:overview}, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q (as this diagnostic can only applied to content in polar questions) and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c (as this diagnostic can only be applied to content in declarative sentences).
+As discussed in section \ref{s:ai-diagnostics}, these two measures of at-issueness differ not just with respect to the required sentence type but also in whether it is a question- or anaphora-based diagnostic.
+In Exps.~2, at-issueness was measured a version of the `affirmation diagnostic' in Exp.~2q and by a version of the `assent diagnostic' in Exps.~2n, 2m, and 2c.
+Specifically, it was measured by versions of these diagnostics in which the affirmation ({\em yes}) or assent ({\em yes, that's true}) is followed by the content to be diagnosed.
+In Exps.~3, at-issueness was measured by a different version of these diagnostics, namely by the version in which the affirmation or assent is followed by a denial of another salient content.
+
+%
+%In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q, illustrated in (\ref{ai-exps1}a), and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, illustrated in (\ref{ai-exps1}b).
+The `asking whether' diagnostic requires that the content to be diagnosed is realized by an expression in a polar question; this diagnostic was also used in \citealt{tbd-variability} Exps.~1.
+A speaker (Christopher in (\ref{ai-exps1}a) asks a polar question and the participant is asked whether the speaker is asking about the content to be diagnosed (here, that Julian dances salsa).
+If a participant takes the speaker to be asking of the content, the content is assumed to be at-issue; if a participant does not take the speaker to be asking about the content, the content is assumed to be not-at-issue.
+The `sure that' diagnostic requires that the content to be diagnosed is realized by an expression in declarative sentence; this diagnostic was also used in \citealt{tbd-variability} Exps.~2.
+A speaker utters the declarative sentence and the addressee (here, Sandy) asks {\em Are you sure?}.
+The speaker response with {\em Yes, I am sure that} followed by the expression that denotes the content to be diagnosed.
+Participants are asked whether the speaker answered the addressee's question.
+If a participant takes the speaker to have answered the question, the content is assumed to be at-issue; if a participant does not take the speaker to have answered the question, the content is assumed to be not-at-issue.
+{\bf move all this into section 1}
+%
+%\begin{exe}
+%\ex\label{ai-exps1} At-issueness measures in Exps.~1
+%\begin{xlist}
+%\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
+%\\ Question to participants: Is Christopher asking whether Julian dances salsa?
+%
+%\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
+%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
+%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
+%\\ Sandy: Are you sure?
+%\\ Christopher: Yes, I'm sure that Julian dances salsa.
+%\\ Question to participants: Did Christopher answer Sandy's question?
+%
+%\end{xlist}
+%\end{exe}
+%
+%In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed.
 Participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
@@
 The affirmation/assent is followed by the content to be diagnosed, here that Julian dances salsa.
-
-\begin{exe}
-\ex\label{ai-exps2} At-issueness measures in Exps.~2
-\begin{xlist}
-\ex {\em yes}, $c$
-\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
-\\ Sandy: Yes, Julian dances salsa.
-\\ Question to participants: Does Sandy's response to Christoper sound good?
-
-\ex {\em yes, that's true}, $c$
-\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
-\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
-\\ Sandy: Yes, that's true, Julian dances salsa.
-\\ Question to participants: Does Sandy's response to Christopher sound good?
-
-\end{xlist}
-\end{exe}
-Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
+%
+%\begin{exe}
+%\ex\label{ai-exps2} At-issueness measures in Exps.~2
+%\begin{xlist}
+%\ex {\em yes}, $c$
+%\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
+%\\ Sandy: Yes, Julian dances salsa.
+%\\ Question to participants: Does Sandy's response to Christoper sound good?
+%
+%\ex {\em yes, that's true}, $c$
+%\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
+%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
+%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
+%\\ Sandy: Yes, that's true, Julian dances salsa.
+%\\ Question to participants: Does Sandy's response to Christopher sound good?
+%
+%\end{xlist}
+%\end{exe}
+%Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
 If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
 If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
-
-In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content.
+%
+%In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content.
 As in the other assent diagnostic, participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
@@
 The affirmation/assent is followed by the denial of another salient content, here, that Cole (didn't) discover it.
-
-\begin{exe}
-\ex\label{ai-exps3} At-issueness measures in Exps.~3
-\begin{xlist}
-\ex {\em yes}, $c$
-\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
-\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
-\\ Question to participants: Does Sandy's response to Christoper sound good?
-
-\ex {\em yes, that's true}, $c$
-\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
-\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
-\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
-\\ Question to participants: Does Sandy's response to Christopher sound good?
-
-\end{xlist}
-\end{exe}
-As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
+%
+%\begin{exe}
+%\ex\label{ai-exps3} At-issueness measures in Exps.~3
+%\begin{xlist}
+%\ex {\em yes}, $c$
+%\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
+%\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
+%\\ Question to participants: Does Sandy's response to Christoper sound good?
+%
+%\ex {\em yes, that's true}, $c$
+%\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
+%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
+%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
+%\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
+%\\ Question to participants: Does Sandy's response to Christopher sound good?
+%
+%\end{xlist}
+%\end{exe}
+%As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
 If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
@@
 Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
-Information on recruited participants (total number, ages, gender) and the payments they received can be found in Supplement \ref{a-participants}.
+Information on the participants and the exclusion criteria is provided in Supplement \ref{a-participants}.
 
@@
 The projection and at-issueness of the contents of 20 clauses (such as {\em Julian dances salsa}) were measured in separate blocks in Exps.~1-3.
-For the full set of clauses, see Appendix \ref{a-clauses}.
+For the full set of clauses, see Supplement \ref{a-clauses}.
 
@@
 In the antecedents of conditionals, all predicates were realized in the present tense.
-For the consequents of the conditional sentences see Supplement \ref{a-target}.} Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question.
-Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
+For the consequents of the conditional sentences see Supplement \ref{a-target}.} Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
 They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1).
@@
 
-In the at-issueness blocks of Exps.~1, target
-
-In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}.
-Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique.
-The indirect object of {\em inform} was realized by the proper name {\em Sam}.
-The predicates differed in the tense in which they were realized.
-In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.
-In the antecedents of conditionals, all predicates were realized in the present tense.
-For the consequents of the conditional sentences see Supplement \ref{a-target}.} Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody ask somebody else a question.
-Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
-They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1).
-Greater speaker commitment to the CC should result in higher slider ratings.
-
-The at-issueness diagnostics differed between Exps.~1, 2, and 3: at-issueness was measured in Exp.~1q, 2q, and 3q, with variants of the question-based diagnostic of at-issueness, namely the `asking whether' diagnostic in Exp.~1q, and the `yes, $p$' diagnostic in Exps.~2q and 3q, as shown in Figure \ref{fig-ai-trials-q}.
-
-, with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, with the `yes' diagnostic in Exps.~2q and 3q, with the `assent with positive continuation' diagnostic in Exps.~2n, 2m, and 2c,  and with the `assent with adversative continuation' diagnostic in Exps.~3n, 3m, and 3c.
-
-To assess whether participants were attending to the task, each experiment also included six control stimuli, which were also utterances made by a named speaker.
-(The controls are provided in Supplement \ref{a-control}.) Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same 6 control stimuli.
-Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
-Block order and within-block trial order was randomized.
+In the at-issueness blocks of Exps.~1-3, the 20 clauses were also realized as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates} and the target stimuli realized either a polar question sentence (in Exps.~1q, 2q, and 3q), a negated sentence (in Exps.~1n, 2n, and 3n), a sentence with {\em perhaps} (in Exps.~1m, 2m, 3m), or a conditional sentence (in Exps.~1c, 2c, 3c).
+Sample trials of the at-issueness measures implemented in the 12 experiments are provided in Fig.~\ref{fig:ai-exps1} for Exps.~1, in Fig.~\ref{fig:ai-exps2} for Exps.~2, and in Fig.~\ref{fig:ai-exps3} for Exps.~3.
+Across all these measures, participants gave their responses on a slider marked `no' at one end (coded as 1) and `yes' at the other (coded as 0).
+Greater not-at-issueness of the CC should result in higher slider ratings.
 
@@
 
-\caption{Target trials in at-issueness blocks of Exps.~2 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps3}
+\caption{Target trials in at-issueness blocks of Exps.~2 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps2}
 \end{figure}
@@
 
-====
-
-The projection block also included 6 control trials, which functioned as attention checks.
-The content of these stimuli was expected not to project: For example, in Fig.~\ref{fig-exp1-projection-control}, the speaker is not committed to the main clause content, that Samantha has a new hat.
-The same 6 main clauses were also used to form 6 filler trials in the prior block; a sample stimulus is given in Fig.~\ref{fig-exp1-prior-filler}.
-These filler stimuli were not used to assess participants' attention.
-For the full set of stimuli see Supplement \ref{a-stim}.
-
-Each participant's stimulus set was semi-randomly generated by first randomly pairing up the 20 predicates and clauses.
-Half of the stimuli were then randomly assigned the respective clause's higher-probability fact, and half its lower-probability fact.
-Participants completed a total of 52 trials: 20 target trials in each block, 6 control trials in the projection block, and 6 filler trials in the prior block.
-Each participant completed the same 6 filler and control trials.
-Block order and within-block trial order were randomized.
-
+To assess whether participants were attending to the task, each experiment also included six control stimuli, which were also utterances made by a named speaker.
+Depending on the experiment, the content of these stimuli was expected to receive either high or low certainty ratings, and either high or low ratings on the at-issueness measure; see Supplement \ref{a-control} for details and the control stimuli.
+
+Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else.
+Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates) and the same 6 (experiment-specific) control stimuli.
+Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
+Block order and within-block trial order was randomized.
 After completing the experiment, participants filled out a short optional demographic survey.
@@
 
-\section{20 complement clauses}\label{a-clauses}
-
-The following clauses realized the complements of the predicates in Exps.~1-3:
+\section{20 clauses}\label{a-clauses}
+
+The content of the following 20 clauses, which realized the complements of the 20 clause-embedding predicates, were investigated in Exps.~1-3:
 
@@
 
-\section{Consequents for conditional target stimuli in Exps.~1c, 2c, and 3c}\label{a-target}
+\section{Consequents for conditional target stimuli in Exps.~1c, 2c, and 3c}\label{a-consequents}
 
@@
 
-\newpage
-
 \section{Control stimuli in Exps.~1-3}\label{a-control}
 
-The control stimuli in Exps.~1-3 were main clauses.
-In the question embedding experiments (Exps.~1q, 2q and 3q), the control stimuli consisted of the polar questions in (1).
-The NRRCs in parentheses were realized in Exps.~2q and 3q, where at-issueness was measured with an assent diagnostic: the control stimuli here consisted of two clauses (like the target stimuli), allowing the relevant speaker to assent with one of the two clauses.
+The control stimuli in Exps.~1-3 were the contents of main clauses.
+In Exps.~1q, 2q and 3q, the control stimuli consisted of the polar questions in (1).
+The non-restrictive relative clauses (NRRCs), given in parentheses in (1), were included in Exps.~2q and 3q, where at-issueness was measured with an assent diagnostic.
+The control stimuli here consisted of two clauses (like the target stimuli), to allow the relevant speaker to assent with one of two clauses.
 
@@
 We also expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (1), indicating that the main clause content is at-issue.
-These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but not for Exp.~3q.
+These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but the ratings were higher than expected for Exp.~3q.
+{\bf NEED TO DISCUSS THIS}
 
 In the remaining experiments, the control stimuli consisted of the positive declarative variants of (1) given in (2).
-The NRRCs in parentheses were realized in Exps.~2 and 3 for the reason explained above, as well as in Exp.~1c, to make the control stimuli more similar to the target stimuli (which also consisted of two clauses: the antecedent and the consequent).
+The NRRCs in parentheses were realized in Exps.~2 and 3 for the reason explained above, as well as in Exp.~1c, to make the control stimuli more similar to the target stimuli (which also consisted of two clauses, namely the antecedent and the consequent).
 
@@
 These expectations were borne out for the `n', `m', and `c' variants of Exps.~1, as shown in the fourth column of Table~\ref{t-controls}, but not for the `n', `m', and `c' variants of Exps.~2 and 3.
+{\bf NEED TO DISCUSS THIS}
 
@@
 & &  \multicolumn{2}{c}{Mean ratings} &  \\ 
-Exp. & Control stimuli & Certainty & At-issueness & At-issueness measure \\ 
+Exp. & Control stimuli & Certainty & Not-at-issueness & At-issueness measure \\ 
 \hline
@@
 Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments).
+{\bf REVIEW THIS, MAKE SURE WE REALLY DO THIS}
 
@@
 
-\section{Comparing projection with different entailment-canceling environments}\label{a-projection}
-
 \end{document}

### 2022-10-28 writing/investigating-GPP.tex
worked on section 2 and appendices

@@
 
-The content of the following 20 clauses, which realized the complements of the 20 clause-embedding predicates, were investigated in Exps.~1-3:
+The contents of the following 20 clauses, which realized the complements of the 20 clause-embedding predicates, were investigated in Exps.~1-3:
 
@@
 
-The consequents for the conditional target stimuli in Exps.~1c, 2c, and 3c were created with the following considerations in mind.
-Each complement clause was paired with a unique consequent clause, as shown in the list below for the 20 complement clauses.
-Each consequent clause consisted of a uniquely named subject and an adjectival predication in the future tense ({\em will be}); the adjectives all denoted an emotion.
+We created 20 consequent clauses for the three experiments in which the 20 clauses were embedded in the antecedent of a conditional (Exps.~1c, 2c, 3c).
+Each of the 20 clauses was paired with a unique consequent clause, as shown in the list below.
+To minimize the variability of the effect of the contents of these consequent clauses on the projection of the contents of the 20 complement clauses, the consequent clauses all consist of a uniquely named subject and an adjectival predication in the future tense ({\em will be}), and the adjectives all denote an emotion.
 We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner-etal2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
@@
 
-We expected participants to give low responses on the `certain that' diagnostic for the control stimuli in (1), indicating that the speaker is not certain of the main clause content, because main clause content is hypothesized to not project out of the question.
+We expected participants to give low responses on the `certain that' diagnostic for the control stimuli in (1), indicating that the speaker is not certain of the main clause content, because main clause content is hypothesized to not project out of polar questions.
 These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for Exps.~1q, 2q, and 3q.
@@
 These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but the ratings were higher than expected for Exp.~3q.
-{\bf NEED TO DISCUSS THIS}
+See below for discussion.
 
@@
 These expectations were borne out for the `n', `m', and `c' variants of Exps.~1, as shown in the fourth column of Table~\ref{t-controls}, but not for the `n', `m', and `c' variants of Exps.~2 and 3.
-{\bf NEED TO DISCUSS THIS}
+See below for discussion.
 
@@
 
-\newpage
-
-The at-issueness measures that did not work out as expected are illustrated in the examples in (3), (4), and (5).
-In (3), B responds in the affirmative to A's question, but then denies the truth of the NRRC.
-The group mean is comparatively higher (at .28) than for the other `q' experiments, including Exp.~2q.
-This result suggests that B's affirmative response to A's question is taken by the participants to affirm not only the main clause content, but also the content of the NRRC.
-Subsequent denial of the truth of the NRRC results in degradation in acceptability.
-
-\begin{exe}
-\exi{(3)} Sample control stimulus in Exp.~3q
+As mentioned above, the mean at-issueness ratings were higher than expected in the `n', `m', and `c' variants of Exps.~2 and all four of the Exps.~3.
+We discuss these exceptions in more detail here because they allow us to further understand the various measures for at-issueness investigated in this paper.
+The example in (3a) illustrates the version of the assent diagnostic applied in Exps.~2n, 2m, and 2c: the assent particle {\em yes} is followed by the clause whose content is diagnosed, here, the content of the main clause.
+The related affirmation diagnostic applied in Exp.~2q (where the mean at-issueness rating was as low as expected) is illustrated in (3b).
+
+\begin{exe}
+\exi{(3)} 
+\begin{xlist}
+\ex Sample control stimulus in Exps.~2n, 2m, and 2c
+\begin{xlist}
+\exi{A:} These muffins, which are really delicious, have blueberries in them.
+\exi{B:} Yes, that's true, they have blueberries in them.
+\end{xlist}
+Question to participants: Does A's response to B sound good?
+\ex Sample control stimulus in Exp.~2q
+\begin{xlist}
+\exi{A:} Do these muffins, which are really delicious, have blueberries in them?
+\exi{B:} Yes, they have blueberries in them.
+\end{xlist}
+Question to participants: Does A's response to B sound good?
+\end{xlist}
+\end{exe}
+As shown in Table \ref{t-controls},  the group mean on the control stimuli is numerically higher (at .22 or .25) than for Exp.~2q (.07).
+We hypothesize that a possible explanation for this difference is that participants take A to assert both the content of the main clause and of the NRRC in (3a) but is only asking about the main clause content in (3b).
+Whereas B's affirmation in (3b) specifies the one content that is affirmed, B's assent in (3a) specifies only one of the two contents that A asserted, seemingly leaving out a specification of the second content, that of the NRRC.
+Participants may judge B's response in (3b) to be less acceptable than in (3a) because of this missing content specification.
+This hypothesis is consistent with the results of \citepos{syrett-koev2015} Exp.~3, which suggests that content of a sentence-medial NRRCs can be the target of a direct denial, though the main clause content is preferred as the target of such a denial.
+That the group means on the control stimuli in our Exps.~2n, 2m, and 2c are still relatively low may be due to the fact that the one content that was specified is the at-issue main clause content.
+One would expect lower acceptability ratings in a version of this diagnostic in which the one content that was specified is the NRRC.
+
+The mean at-issueness ratings for the control stimuli were also higher than expected in Exps.~3.
+The examples in (4) illustrate the versions of the affirmation and assent diagnostics used in these experiments:
+
+\begin{exe}
+\exi{(4)} 
+\begin{xlist}
+\ex Sample control stimulus in Exp.~3q
 \begin{xlist}
@@
 \end{xlist}
-\end{exe}
-
-In (4), B assents with A's indicative assertion, and then re-affirms the truth of the main clause content.
-The group mean is comparatively higher (at .22 or .25) than for Exp.~2q.
-This result suggests that responding in the affirmative to A's question is a different speech act than assenting with A's indicative assertion: whereas it does not appear to be odd, in the latter case, to spell out the propositional antecedent of {\em yes}, it does seem to be odd to do so in the latter case.
-
-\begin{exe}
-\exi{(4)} Sample control stimulus in Exps.~2n, 2m, and 2c
+Question to participants: Does A's response to B sound good?
+\ex Sample control stimulus in Exps.~3n, 3m, and 3c
 \begin{xlist}
 \exi{A:} These muffins, which are really delicious, have blueberries in them.
-\exi{B:} Yes, that's true, they have blueberries in them.
-\end{xlist}
-\end{exe}
-
-In (5), B assents with A's indicative assertion, but then denies the truth of the NRRC.
-Here, the group mean is quite a bit higher (at .44, .50, .53) than for Exp.~3q or for Exps.~2n, 3m, and 3c.
-This result suggests, again, that responding in the affirmative to A's question is a different speech act than assenting with A's indicative assertion: while denying the truth of the NRRC after responding in the affirmative to the question is already odd (Exp.~3q), denying the truth of the NRRC after assenting with A's assertion with the NRRC is worse.
-
-\begin{exe}
-\exi{(5)} Sample control stimulus in Exps.~3n, 3m, and 3c
-\begin{xlist}
-\exi{A:} These muffins, which are really delicious, have blueberries in them.
-\exi{B:} Yes, that's true, but they aren't really delicious.
-\end{xlist}
-\end{exe}
-
-The responses to the control stimuli with NRRCs that used either one of the assent diagnostics suggest that assenting with a speaker's indicative utterance that includes an NRRC includes the content of the NRRC.
-Two possibilities: either the content of the NRRC is already entered into the CG when A asserts the sentence (thus making reaffirmation or denial odd); or B's {\em yes, that's true} assents not just with the main clause content, but also (though perhaps a bit less?) with the content of the NRRC (thereby making reaffirmation or denial odd).
-
-%Originally, the intent was to exclude participants' data on the basis of their at-issueness ratings, in parallel to their certainty ratings.
-We expected the main clause content to be at-issue across the diagnostics, which means that we expected low responses across all at-issueness diagnostics (recall that participants' responses were coded so that the higher a participants' response, the more not-at-issue the content was hypothesized to be, to investigate whether there is a positive correlation between projection and not-at-issueness).
-As shown in Table \ref{t-ai-controls}, these expectations were not always borne out, which is why we decided to not use participants' ratings of the main clause contents in the at-issueness blocks as an exclusion criterion.
-
-%\begin{table}[h!]
-%\centering
-%\begin{tabular}{r l}
-%Experiment & mean rating \\ 
-%\hline
-%1q & .05 \\
-%2q & .07 \\
-%3q & .28 \\
-%\hline
-%1n & .04 \\
-%1m & .03 \\
-%1c & .08 \\
-%\hline
-%2n & .22 (16 participants excluded by specified exclusion criterion) \\
-%2m & .25 \\
-%2c & .22 \\
-%\hline
-%3n & .44  (nobody excluded by specified exclusion criterion) \\
-%3m & .50 \\
-%3c & .53 \\
-%\hline
-%\end{tabular}
-%\caption{Mean ratings on at-issueness diagnostics for main clause contents in %Exps.~1-3, for self-declared American English participants}\label{t-ai-controls}
-%\end{table}
+\exi{B:} Yes, that's true, they but they aren't really delicious.
+\end{xlist}
+Question to participants: Does A's response to B sound good?
+\end{xlist}
+\end{exe}
+The mean at-issueness rating for the control stimuli in Exp.~3q was comparatively higher (at .28) than for Exp.~1q (.05) or Exp.~2q (.07).
+We hypothesize that that this difference (especially to Exp.~2q) is due the content of the NRRC being directly denied, even though it was presented as backgrounded content in A's question (as shown in (2), the content of none of the other NRRCs was a matter of personal taste).
+
+The mean at-issueness ratings for the control stimuli in Exps.~3n, 3m, and 3c were numerically even higher than for Exp.~3q and, in fact, the highest across all 12 experiments (at .44, .50, and .53, respectively).
+There are two factors that could be implicated in the difference between Exp.~3q and Exps.~3n, 3m, and 3c: first, the NRRC is included in a polar question in the former but a declarative assertion in the latter; second, B utters an affirmation {\em yes} in the former but an assent {\em yes, that's true} in the latter.
+For instance, participants might judge B's direct denial of the content of the NRRC as less acceptable when A uttered a declarative assertion and B assented with that assertion using {\em yes, that's true} than when A uttered a polar question and B responded in the affirmative with {\em yes}.
+Both factors must be considered in future research on at-issueness measures.
 
@@
 This supplement provides information on the participants of the 12 experiments and the criteria by which participants' data were excluded.
-The columns of Table \ref{t:exclusion} show, for each of the 12 experiments (Exp.~1q, \ldots, Exp.~3c) how many participants were recruited, their age range and mean age, and their self-reported gender (`f' = female, `m' = male, `o' = other, `u' = undeclared).
-No gender data was collected in Exp.~1q.
-Participants' data were excluded based on the following criteria:
+The first three columns of Table \ref{t:exclusion} show, for each of the 12 experiments how many participants were recruited, their age range and mean age, and their self-reported gender; no gender data was collected in Exp.~1q.
+The next three columns provide information on the number of participants whose data were excluded based on the following criteria:
 
@@
 Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments).
-{\bf REVIEW THIS, MAKE SURE WE REALLY DO THIS}
-
-\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli, that is, if their mean variance on the target stimuli was more than 2 sd below the group mean variance.
-{\bf and manually inspecting their response patterns.
-Exps 1q, 2q, 3q: participants with low variance not excluded after manual inspection NEED TO CONSIDER THIS AGAIN}
+
+\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli.
+To identify such participants, we first identified participants whose mean variance on the target stimuli was more than 2 sd below the group mean variance and then manually inspecting their response patterns.
+The data of participants who used the full scale was not excluded.
 \end{itemize}
@@
 &  \multicolumn{3}{c|}{Recruited participants} & \multicolumn{4}{c|}{Exclusion criteria} & \multicolumn{3}{c|}{Remaining participants} &  \\ 
-Exp. & total & ages (mean) & f/m/o/u & multiple & language & controls & variance & total & ages (mean) & f/m/o/u & payment   \\ 
+Exp. & total & ages (mean) & gender (f/m/o/u) & multiple & language & controls & variance & total & ages (mean) & gender (f/m/o/u) & payment   \\ 
 \hline
@@
 \end{tabular}
-\caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3}\label{t:exclusion}
+\caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3.
+Gender distinctions were `f' = female, `m' = male, `o' = other, and `u' = undeclared.}\label{t:exclusion}
 \end{sidewaystable}

### 2022-10-28 results/main/1_projaiQ/rscripts/preprocessing.R

@@ -214,7 +214,7 @@ variances = d %>%
   summarize(Variance = var(response)) %>%
   mutate(TooSmall = Variance < mean(Variance) - 2*sd(Variance))
 
-View(variances)
+#View(variances)
 
 ggplot(variances,aes(x=workerid,y=Variance)) +
   geom_point()
@@ -228,8 +228,8 @@ lvw = d %>%
   droplevels() %>%
   mutate(Participant = as.factor(as.character(workerid)))
 
-ggplot(lvw,aes(x=Participant,y=response,color=trigger_class)) +
-  geom_jitter()
+ggplot(lvw[lvw$trigger != "MC",],aes(x=Participant,y=response,color=trigger_class)) +
+  geom_point()
 
 # manual inspection shows that these three participants used the whole scale, so no reason to exclude them
 
@@ -246,3 +246,4 @@ length(which(is.na(d$age))) # 0 missing values
 mean(d$age,na.rm=TRUE) #39.2
 
 # no gender information available
+

### 2022-10-28 results/main/2_projaiN/rscripts/preprocessing.R

@@ -211,3 +211,4 @@ d %>%
   group_by(gender) %>% 
   summarize(count=n())
 #141 female, 128 male, 5 other, 0 undeclared
+

### 2022-10-28 results/main/3_projaiM/rscripts/preprocessing.R

@@ -205,3 +205,4 @@ d %>%
   group_by(gender) %>% 
   summarize(count=n())
 #144 female, 129 male, 7 other, 1 undeclared
+

### 2022-10-28 results/main/4_projaiC/rscripts/preprocessing.R

@@ -221,3 +221,4 @@ d %>%
   group_by(gender) %>% 
   summarize(count=n())
 #235 female, 25 male, 6 other, 0 undeclared
+

### 2022-11-06 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
added data summary and graphs for projection data

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%% document settings
\documentclass[10pt]{article}
\usepackage{euler}
\usepackage{xunicode,xltxtra}
\usepackage{fontspec}

\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}

% fonts
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Biolinum}
	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
	\setmonofont[Scale=MatchLowercase]{Andale Mono}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{mathtools}
\usepackage{linguex}
\usepackage{booktabs}

\usepackage[colorlinks=true,citecolor=MidnightBlue,urlcolor=MidnightBlue,linkcolor=MidnightBlue]{hyperref}

\title{Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis}
\author{Lisa Hofmann}
\date{\today}

\begin{document}
\setkeys{Gin}{width=1.0\textwidth}
<<setup, include=FALSE, echo=FALSE>>=
library(knitr)
# library(highr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = xfun::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo=TRUE, prompt=TRUE, comment=NA, tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/figures', fig.align = "center", out.width="\\linewidth")
options(tidy=TRUE, linewidth = 100)
@

\maketitle
\setcounter{tocdepth}{2}
\tableofcontents

\section{Introducing the dataset}
<<load-data, echo=FALSE, results='hide'>>=
# set working directory to directory of script
# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd("/Users/lisahofmann/Dropbox/Mein Mac (MBP-von-Lisa)/Documents/Academic/Research Projects/projective content/experiments-github/materials-lisa/analysis/3_projection-across-operators")

# load data from exps into list
df_list <- list(data_1q = read.csv("../../../results/main/1_projaiQ/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_1n = read.csv("../../../results/main/2_projaiN/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_1m = read.csv("../../../results/main/3_projaiM/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_1c = read.csv("../../../results/main/4_projaiC/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2q = read.csv("../../../results/main/5_projaiQ/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2n = read.csv("../../../results/main/6_projaiN/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2m = read.csv("../../../results/main/7_projaiM/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2c = read.csv("../../../results/main/8_projaiC/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3q = read.csv("../../../results/main/9_projaiQ/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3n = read.csv("../../../results/main/10_projaiN/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3m = read.csv("../../../results/main/11_projaiM/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3c = read.csv("../../../results/main/12_projaiC/data/data_preprocessed.csv", 
                    header=TRUE, sep=",")
                )

@

<<clean-data, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=
require(tidyverse)
library(stringr)
library(rlist)

# spread responses over separate columns for projectivity and at-issueness
# add label for which position at block appeared in
df_list <- lapply(df_list, function(df) {
  df = df %>%
    mutate(ai_block = ifelse(question_type == "ai", 
                             ifelse(block == "block1", "block1", "block2"), 
                             ifelse(block == "block1", "block2", "block1"))) %>%
    select(workerid,content,short_trigger,question_type,response,ai_block) %>%
    spread(question_type,response)
  df
})

# change cd verb names to match veridicality names
df_list <- lapply(df_list, function(df) {
  df = df %>%
    mutate(verb=recode(short_trigger, control = "MC", annoyed = "be_annoyed", 
                       be_right_that = "be_right", inform_Sam = "inform"))
  df
})

# add labels for operator, and experiment block
df_list = map2(df_list, names(df_list), function(df, name) {
  operator <- str_sub(name, -1, -1)
  exp_block <- str_sub(name, -2, -2)
  df$op <- rep(operator, length(df[,1]))
  df$exp_block <- rep(exp_block, length(df[,1]))
  return(df)
})

# recode relevant vectors as factors
df_list <- lapply(df_list, function(df) {
  df$content <- as.factor(df$content)
  df$verb <- as.factor(df$verb)
  df$workerid <- as.factor(df$workerid)
  df$op <- as.factor(df$op)
  df$exp_block <- as.factor(df$exp_block)
  df
})

# remove data from MC controls
df_list <- lapply(df_list, function(df) {
  df = filter(df, verb != "MC")
  df$verb = droplevels(df$verb)
  df$content = droplevels(df$content)
  df
})

# combine data from all experiments
data_all <- list.rbind(df_list)
@

<<data-overview>>=
str(data_all)
length(levels(data_all$workerid))
@

\noindent The dataset consists of $57160$ observations from $2682$ participants (recruited on the online platforms Prolific and Amazon Mechanical Turk), across 12 experiments.

We are interested in how highly participants rate speaker commitment to the truth of an embedded complement clause, coded as \texttt{projective} on a real-numbered sliding scale between $0-1$.\\

\noindent The complement clause was embedded under an attitude verb, which in turn was embedded under an entailment-cancelling operator. Our fixed effects factors manipulate the following:

\begin{enumerate}
  \item The choice of attitude verb (coded as \texttt{verb})
  \item The entailment-cancelling operator (coded as \texttt{op})
\end{enumerate}

\noindent The levels for our fixed effects factors are the following:
<<fixed-efcts-lvls>>=
levels(data_all$verb)
length(levels(data_all$verb))
levels(data_all$op)
length(levels(data_all$op))
@

\noindent We are interested in the effect on \texttt{projective} of \texttt{verb} and \texttt{op}, as well as their interaction, corresponding to a $20 \times 4$ factorial design, yielding
<<ncond>>=
length(levels(data_all$verb))*length(levels(data_all$op))
@
conditions.\\

\noindent We have $20$ items, corresponding to the content of the complement clause.
<<nitems>>=
levels(data_all$content)
length(levels(data_all$content))
@

\noindent We have roughly $36$ observations by item and condition. This is an approximate number, because the \texttt{op} manipulation is a between-studies manipulation, and the number of participants differs by experiment:
<<nobs>>=
# n observations
length(data_all[,1])

# observations by item
length(data_all[,1])/length(levels(data_all$content))
table(data_all$content)

# observations by verb
length(data_all[,1])/length(levels(data_all$verb))
table(data_all$verb)

# observations by operator
length(data_all[,1])/length(levels(data_all$op))
table(data_all$op)

# observations by item and condition
length(data_all[,1])/length(levels(data_all$content))/
  (length(levels(data_all$verb))*length(levels(data_all$op)))

@

\newpage
\section{Data Overview and Statistical Summaries}
\subsection{Distribution of projectivity ratings by operator:}
<<project-by-op-distr, echo = FALSE>>=
data_all %>% 
  # mutate(verb = fct_reorder(cond, as.numeric(dataf$response), .fun = 'mean')) %>%
  ggplot(aes(x = projective, group = op)) +
  facet_grid(rows = vars(op)) +
  geom_histogram(binwidth = .01, fill = "lightblue") +
  # coord_flip() +
  xlab("") +
  ylab("") +
  labs(title = "")+
  theme_bw()
  # scale_fill_brewer(palette = "Blues")
@

\begin{itemize}
  \item These definitely do not look normal
  \item Maybe a beta-regression would be useful?
  \item But even that would be relying on some simplifying assumptions, since we might be ignoring the little bump in the middle 
\end{itemize}

\newpage
\subsection{Distribution of projectivity ratings by verb:}
<<project-by-v-distr, echo = FALSE>>=
data_all %>% 
  # mutate(verb = fct_reorder(cond, as.numeric(dataf$response), .fun = 'mean')) %>%
  ggplot(aes(x = projective, group = verb)) +
  facet_wrap(vars(verb)) +
  geom_histogram(binwidth = .01, fill = "lightblue") +
  # coord_flip() +
  xlab("") +
  ylab("") +
  labs(title = "")+
  theme_bw()
  # scale_fill_brewer(palette = "Blues")
@

\begin{itemize}
  \item Some of these also show a higher mass around the middle of the scale
  \item but it looks the beta-distribution could be a useful approximation
\end{itemize}

\newpage
\subsection{Means and confidence intervals for projectivity rating by operator}
<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".6\\textwidth">>=
# load helper functions
source('../../../results/main/helpers.R')

proj_means_op = data_all %>% group_by(op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         op = fct_reorder(as.factor(op),Mean)) %>% ungroup()

proj_means_op %>% mutate(op = fct_reorder(op, Mean, 
                                         .fun = mean)) %>% 
  ggplot(aes(x = op, y=Mean, color = op)) +
  # coord_cartesian(ylim = c(0,1)) +
  geom_point(aes(shape = op), size = 4) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
  geom_line() + 
  labs(title = "Mean projectivity by operator")+
  theme_bw() +
  scale_color_brewer(palette = "PRGn")
@
\noindent The following generalizations emerge:
\begin{itemize}
  \item Conditionals have the highest projectivity ratings
  \item Projectivity ratings for questions are higher than those for modals and negation, but lower than those for conditionals
  \item Modals and negation have the lowest projectivity ratings
  \item The ratings for negation look a little higher than for modals, but error bars overlap
\end{itemize}

\noindent Although these differences appear to be significant, they are quite small.\\

\newpage
\subsection{Means and confidence intervals for projectivity rating by verb:}
<<summary-v, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
# load helper functions
source('../../../results/main/helpers.R')

proj_means_v = data_all %>% group_by(verb) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

proj_means_v %>% mutate(verb = fct_reorder(verb, Mean, 
                                         .fun = mean)) %>% 
  ggplot(aes(x=verb, y=Mean)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "blue") +
  geom_line() + 
  labs(title = "Mean projectivity by predicate")+
  theme_bw()
@
\begin{itemize}
  \item We see gradual differences in projectivity between verbs
\end{itemize}

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
proj_means = data_all %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

# kable(proj_means, "latex", longtable = TRUE)

proj_means %>% mutate(verb = fct_reorder(verb, Mean, 
                                         .fun = mean)) %>% 
  mutate(op = fct_reorder(op, Mean, .fun = mean)) %>% 
  ggplot(aes(x=verb, y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 4) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() + 
  labs(title = "Mean projectivity for predicate by operator")+
  theme_bw() +
  scale_color_brewer(palette = "PRGn")

@

\begin{itemize}
  \item We see interactions between verb and operator
  \item However, we do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen. Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.

  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.

  \item What else\dots?
\end{itemize}

% \section{Exploring a couple of models}
% \subsection{Linear, frequentist}
% Let's try a linear model first, and check error distributions.
%

\end{document}

### 2022-11-06 materials-lisa/notes/notes.tex
added data summary and graphs for projection data

--- 2022-08-03 notes lisa/notes.tex
@@
 \pagebreak
+\section{Some more notes} % (fold)
+
+	\subsection{At-issueness diagnostics} % (fold)
+
+		\begin{itemize}
+			\item some diagnostics of at-issueness for some propositions $p$, $q$ associated with $U$  may work fine for certain types of declaratives, based on
+			\begin{itemize}
+				\item whether $p, q$ are logically independent of each other
+				\item if $p, q$ are both true according to speaker, if at-issueness of content can be diagnosed given the assumption that speaker is committed to content (or not)
+			\end{itemize}
+
+			\item because of point about commitment definitely does not work for at-issueness of content under entailment-cancelling operators (point about commitment)
+
+			\item may run into problems for embedded / matrix content, because these are related, unlike e.g. appositives (but like evidentials)
+
+		\end{itemize}
+
+	% subsection at_issueness_diagnostics (end)
+
+% section some_more_notes (end)
+
 \section{To do} % (fold)

### 2022-11-06 materials-lisa/analysis/3_projection-across-operators/operators.R

@@ -1,0 +1,266 @@
+
+### comparing projectivity across entailment-cancelling operators
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# for exps 1: compare Q / N / M / C; 1-4
+
+# load data from exps 1
+data_1q <- read.csv("../../../results/main/1_projaiQ/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_1n <- read.csv("../../../results/main/2_projaiN/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_1m <- read.csv("../../../results/main/3_projaiM/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_1c <- read.csv("../../../results/main/4_projaiC/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+
+# load data from exps 2
+data_2q <- read.csv("../../../results/main/5_projaiQ/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_2n <- read.csv("../../../results/main/6_projaiN/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_2m <- read.csv("../../../results/main/7_projaiM/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_2c <- read.csv("../../../results/main/8_projaiC/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+
+# load data from exps 3
+data_3q <- read.csv("../../../results/main/9_projaiQ/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_3n <- read.csv("../../../results/main/10_projaiN/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_3m <- read.csv("../../../results/main/11_projaiM/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+data_3c <- read.csv("../../../results/main/12_projaiC/data/data_preprocessed.csv", 
+                    header=TRUE, sep=",")
+
+df_list <- list(data_1q=data_1q, data_1m=data_1m, data_1n=data_1n, data_1c=data_1c,
[…]

### 2022-11-06 writing/investigating-GPP.tex
made small changes to paper
added data summary and graphs for projection data

@@
 
+Prijection comparison hasn't been done, except in CommitmentBank paper (right?, perhaps A'dam colloquium presentation?).
+
 \begin{exe}
@@
 \ex Is the at-issueness of contents positively correlated under different measures of at-issueness?
+
+\ex Compare projection under different entailment-cancelling operators
+
+\ex Do commonly used at-issueness diagnostics work for content embedded under entailment-cancelling operators?
+
 \end{xlist}
@@
 
-We recruited 250-300 participants for each of Exps.~1-3.
+We recruited 250-300 participants for each of the 12 experiments.
 Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs.

### 2022-11-18 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
try this model

@@
 <<project-by-op-distr, echo = FALSE>>=
+library(ggplot2)
 data_all %>% 
@@
   scale_color_brewer(palette = "PRGn")
+
+proj_means_op %>% mutate(op = fct_reorder(op, Mean, 
+                                         .fun = mean)) %>% 
+  ggplot(aes(x = op, y=Mean, color = op, group = verb)) +
+  # coord_cartesian(ylim = c(0,1)) +
+  geom_point(aes(shape = op), size = 4) + 
+  scale_shape_manual(values = c("M", "N", "Q", "C")) +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
+  facet_wrap(vars(verb)) +
+  geom_line() + 
+  labs(title = "Mean projectivity by operator")+
+  theme_bw() +
+  scale_color_brewer(palette = "PRGn")
+
 @
@@
 <<summary-v, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
-# load helper functions
-source('../../../results/main/helpers.R')
 
@@
 
-% \section{Exploring a couple of models}
-% \subsection{Linear, frequentist}
-% Let's try a linear model first, and check error distributions.
-%
+\section{Analysis}
+
+We are interested in how the response (projectivity ratings) depends on the verb and embedding operator.
+
+<<>>=
+# main analysis 4: predict projection from not-at-issueness, predicate, and their interaction ----
+
+t = data_all %>% 
+  mutate(verb = fct_relevel(verb, "suggest"))
+
+m.dummy = lm(projective ~ ai*short_trigger, data=t)
+summary(m.dummy)
+
+# reference level set to "prove"
+
+# model with short_trigger slope did not converge
+m.4 = lmer(projective ~ cai*short_trigger + (1+cai|content) + (1+cai|workerid), 
+           data = t, REML=F,control = lmerControl(optimizer ='optimx', optCtrl=list(method='L-BFGS-B')))
+saveRDS(m.4, "../models/m.4.rds")
+m.4 <- readRDS("../models/m.4.rds")
+summary(m.4)
+texreg(m.4,single.row = TRUE, leading.zero = FALSE, booktabs = TRUE)
+
+@
+
+<<>>=
+formula <- response ~ op * verb
+# formula <- response ~ op * verb + (1 | workerid) + (1 | item)
+# formula <- response ~ op * verb + (1 + op + verb | workerid) + (1 + op + verb | item)
+@
+
+Trying a beta-regression.
+Our response variable has observations from a sliding scale between $(0,1)$.
+For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).
+
+<<>>=
+data_all$response = (data_all$projective*(nrow(data_all)-1) + .5)/nrow(data_all)
+@
+
+<<>>=
+# library(rstan)
+# options(mc.cores=parallel::detectCores())
+# rstan_options(auto_write=TRUE)
+# library(brms)
+# 
+# model1 <- brm(formula = bf(response ~ op * verb ,
+#                            phi ~ op * verb,
+#                            family = Beta()),
+#               family = Beta(),
+#               data = data_all,
+#               iter=10000, warmup=2000,chains=6)
+
+@
 

### 2022-11-18 materials-lisa/analysis/3_projection-across-operators/operators.R

@@ -41,9 +41,10 @@ df_list <- list(data_1q=data_1q, data_1m=data_1m, data_1n=data_1n, data_1c=data_
                   data_2q=data_2q, data_2m=data_2m, data_2n=data_2n, data_2c=data_2c,
                   data_3q=data_3q, data_3m=data_3m, data_3n=data_3n, data_3c=data_3c)
 
-# spread responses over separate columns for projectivity and at-issueness
+
 # require(tidyverse)
 
+# spread responses over separate columns for projectivity and at-issueness
 df_list <- lapply(df_list, function(df) {
   df = df %>%
     mutate(block_ai = ifelse(question_type == "ai", 
@@ -68,6 +69,7 @@ df_list <- lapply(df_list, function(df) {
 
 # library(stringr)
 
+# add labels for operator, and experiment block
 df_list = map2(df_list, names(df_list), function(df, name) {
   # print(df$content[1])
   operator <- str_sub(name, -1, -1)

### 2022-11-18 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
updated repo structure

--- 2022-11-06 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
@@
 # set working directory to directory of script
-# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd("/Users/lisahofmann/Dropbox/Mein Mac (MBP-von-Lisa)/Documents/Academic/Research Projects/projective content/experiments-github/materials-lisa/analysis/3_projection-across-operators")
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+#setwd("/Users/lisahofmann/Dropbox/Mein Mac (MBP-von-Lisa)/Documents/Academic/Research Projects/projective content/experiments-github/materials-lisa/analysis/3_projection-across-operators")
 
@@
 
+<<>>=
+t = data_all %>% 
+  mutate(verb = fct_relevel(verb, "suggest")) %>%
+  mutate(op = fct_relevel(op, "q"))
+
+#m.dummy = lm(projective ~ ai*short_trigger, data=t)
+#summary(m.dummy)
+
+# reference level set to "prove"
+
+# model with short_trigger slope did not converge
+m.4 = lmer(projective ~ cai*short_trigger + (1+cai|content) + (1+cai|workerid), 
+           data = t, REML=F,control = lmerControl(optimizer ='optimx', optCtrl=list(method='L-BFGS-B')))
+saveRDS(m.4, "../models/m.4.rds")
+m.4 <- readRDS("../models/m.4.rds")
+summary(m.4)
+texreg(m.4,single.row = TRUE, leading.zero = FALSE, booktabs = TRUE)
+@
+
 \end{document}

### 2022-11-23 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
updated graphs and generalizations w verb-profiles

--- 2022-11-18 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
@@
 # set working directory to directory of script
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-#setwd("/Users/lisahofmann/Dropbox/Mein Mac (MBP-von-Lisa)/Documents/Academic/Research Projects/projective content/experiments-github/materials-lisa/analysis/3_projection-across-operators")
+# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+# setwd(this.dir)
+setwd("/Users/lisahofmann/Dropbox/Mein Mac (MBP-von-Lisa)/Documents/Academic/Research Projects/projective content/experiments-github/materials-lisa/analysis/3_projection-across-operators")
 
-# load data from exps into list
-df_list <- list(data_1q = read.csv("../../../results/main/1_projaiQ/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_1n = read.csv("../../../results/main/2_projaiN/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_1m = read.csv("../../../results/main/3_projaiM/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_1c = read.csv("../../../results/main/4_projaiC/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_2q = read.csv("../../../results/main/5_projaiQ/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_2n = read.csv("../../../results/main/6_projaiN/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_2m = read.csv("../../../results/main/7_projaiM/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_2c = read.csv("../../../results/main/8_projaiC/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_3q = read.csv("../../../results/main/9_projaiQ/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_3n = read.csv("../../../results/main/10_projaiN/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_3m = read.csv("../../../results/main/11_projaiM/data/data_preprocessed.csv", 
-                    header=TRUE, sep=","),
-                data_3c = read.csv("../../../results/main/12_projaiC/data/data_preprocessed.csv", 
+# load data
+data <- read.csv("../../../results/main/meta-analyses/projectivity/data/data_combined.csv", 
                     header=TRUE, sep=",")
-                )
 
@@
 
-<<clean-data, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=
-require(tidyverse)
-library(stringr)
-library(rlist)
-
-# spread responses over separate columns for projectivity and at-issueness
-# add label for which position at block appeared in
-df_list <- lapply(df_list, function(df) {
-  df = df %>%
-    mutate(ai_block = ifelse(question_type == "ai", 
-                             ifelse(block == "block1", "block1", "block2"), 
-                             ifelse(block == "block1", "block2", "block1"))) %>%
-    select(workerid,content,short_trigger,question_type,response,ai_block) %>%
-    spread(question_type,response)
-  df
-})
-
-# change cd verb names to match veridicality names
-df_list <- lapply(df_list, function(df) {
-  df = df %>%
-    mutate(verb=recode(short_trigger, control = "MC", annoyed = "be_annoyed", 
-                       be_right_that = "be_right", inform_Sam = "inform"))
-  df
-})
-
-# add labels for operator, and experiment block
-df_list = map2(df_list, names(df_list), function(df, name) {
-  operator <- str_sub(name, -1, -1)
-  exp_block <- str_sub(name, -2, -2)
-  df$op <- rep(operator, length(df[,1]))
-  df$exp_block <- rep(exp_block, length(df[,1]))
-  return(df)
-})
-
-# recode relevant vectors as factors
-df_list <- lapply(df_list, function(df) {
-  df$content <- as.factor(df$content)
-  df$verb <- as.factor(df$verb)
-  df$workerid <- as.factor(df$workerid)
-  df$op <- as.factor(df$op)
-  df$exp_block <- as.factor(df$exp_block)
-  df
-})
-
-# remove data from MC controls
-df_list <- lapply(df_list, function(df) {
-  df = filter(df, verb != "MC")
-  df$verb = droplevels(df$verb)
-  df$content = droplevels(df$content)
-  df
-})
-
-# combine data from all experiments
-data_all <- list.rbind(df_list)
-@
-
 <<data-overview>>=
-str(data_all)
-length(levels(data_all$workerid))
+str(data)
+data$workerid <- as.factor(data$workerid)
+length(levels(data$workerid))
 @
@@
 <<fixed-efcts-lvls>>=
-levels(data_all$verb)
-length(levels(data_all$verb))
-levels(data_all$op)
-length(levels(data_all$op))
+data$verb <- as.factor(data$verb)
+levels(data$verb)
+length(levels(data$verb))
+data$op <- as.factor(data$op)
+levels(data$op)
+length(levels(data$op))
 @
@@
 <<ncond>>=
-length(levels(data_all$verb))*length(levels(data_all$op))
+length(levels(data$verb))*length(levels(data$op))
 @
@@
 <<nitems>>=
-levels(data_all$content)
-length(levels(data_all$content))
+data$content <- as.factor(data$content)
+levels(data$content)
+length(levels(data$content))
 @
@@
 # n observations
-length(data_all[,1])
+length(data[,1])
 
 # observations by item
-length(data_all[,1])/length(levels(data_all$content))
-table(data_all$content)
+length(data[,1])/length(levels(data$content))
+table(data$content)
 
 # observations by verb
-length(data_all[,1])/length(levels(data_all$verb))
-table(data_all$verb)
+length(data[,1])/length(levels(data$verb))
+table(data$verb)
 
 # observations by operator
-length(data_all[,1])/length(levels(data_all$op))
-table(data_all$op)
+length(data[,1])/length(levels(data$op))
+table(data$op)
 
 # observations by item and condition
-length(data_all[,1])/length(levels(data_all$content))/
-  (length(levels(data_all$verb))*length(levels(data_all$op)))
+length(data[,1])/length(levels(data$content))/
+  (length(levels(data$verb))*length(levels(data$op)))
 
@@
 \subsection{Distribution of projectivity ratings by operator:}
-<<project-by-op-distr, echo = FALSE>>=
-data_all %>% 
-  # mutate(verb = fct_reorder(cond, as.numeric(dataf$response), .fun = 'mean')) %>%
+<<project-by-op-distr, echo = FALSE, message=FALSE>>=
+require(ggplot2)
+require(dplyr)
+data %>% 
   ggplot(aes(x = projective, group = op)) +
@@
   geom_histogram(binwidth = .01, fill = "lightblue") +
-  # coord_flip() +
   xlab("") +
@@
   theme_bw()
-  # scale_fill_brewer(palette = "Blues")
 @
@@
 <<project-by-v-distr, echo = FALSE>>=
-data_all %>% 
-  # mutate(verb = fct_reorder(cond, as.numeric(dataf$response), .fun = 'mean')) %>%
+data %>% 
   ggplot(aes(x = projective, group = verb)) +
@@
   \item Some of these also show a higher mass around the middle of the scale
-  \item but it looks the beta-distribution could be a useful approximation
+  \item but it looks the beta-distribution could be useful
 \end{itemize}
@@
 source('../../../results/main/helpers.R')
+require(forcats)
 
-proj_means_op = data_all %>% group_by(op) %>%
+proj_means_op = data %>% group_by(op) %>%
   summarize(Mean = mean(projective), CILow = ci.low(projective),
@@
                                           .fun = mean)) %>%
-  ggplot(aes(x = op, y=Mean, color = op)) +
+  ggplot(aes(x = op, y=Mean)) +
   # coord_cartesian(ylim = c(0,1)) +
-  geom_point(aes(shape = op), size = 4) + 
-  scale_shape_manual(values = c("M", "N", "Q", "C")) +
-  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
-  geom_line() + 
+  geom_point(size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") +
   labs(title = "Mean projectivity by operator")+
@@
 @
+
 \noindent The following generalizations emerge:
@@
 \subsection{Means and confidence intervals for projectivity rating by verb:}
-<<summary-v, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
-# load helper functions
-source('../../../results/main/helpers.R')
-
-proj_means_v = data_all %>% group_by(verb) %>%
-  summarize(Mean = mean(projective), CILow = ci.low(projective), 
-            CIHigh = ci.high(projective)) %>%
-  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
-         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()
-
-proj_means_v %>% mutate(verb = fct_reorder(verb, Mean, 
-                                         .fun = mean)) %>% 
-  ggplot(aes(x=verb, y=Mean)) +
-  coord_cartesian(ylim=c(0,1)) +
-  geom_point(color = "blue") +
-  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "blue") +
-  geom_line() + 
-  labs(title = "Mean projectivity by predicate")+
-  theme_bw()
-@
+This will be replaced by violin-plot.
+% <<summary-v, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
+% 
+% proj_means_v = data %>% group_by(verb) %>%
+%   summarize(Mean = mean(projective), CILow = ci.low(projective),
+%             CIHigh = ci.high(projective)) %>%
+%   mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
+%          verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()
+% 
+% proj_means_v %>% mutate(verb = fct_reorder(verb, Mean,
+%                                          .fun = mean)) %>%
+%   ggplot(aes(x=verb, y=Mean)) +
+%   coord_cartesian(ylim=c(0,1)) +
+%   geom_point(color = "lightblue") +
+%   geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
+%   labs(title = "Mean projectivity by predicate")+
+%   theme_bw()
+% @
 \begin{itemize}
@@
 <<summary-combined, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
-proj_means = data_all %>% group_by(verb, op) %>%
+
+proj_means = data %>% group_by(verb, op) %>%
   summarize(Mean = mean(projective), CILow = ci.low(projective),
@@
          verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()
-
-# kable(proj_means, "latex", longtable = TRUE)
 
@@
   \item We see interactions between verb and operator
-  \item However, we do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
-Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
 
@@
 
-  \item What else\dots?
+  \item More projective verbs (\emph{\lq announce\rq} and above) have $\texttt{C} > \texttt{M}$
+
+  \item Highly projective verbs (\emph{\lq hear \rq} and above) have  $\texttt{N} > \texttt{M}$
+
+  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
+  \begin{itemize}
+    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
+    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
+    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here
   \end{itemize}
 
-% \section{Exploring a couple of models}
-% \subsection{Linear, frequentist}
-% Let's try a linear model first, and check error distributions.
+  \item For more generalizations, let's look at the same information plotted differently: By-operator projectivity for each verb
+\end{itemize}
+
+<<verb-profiles, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
+proj_means %>% mutate(op = fct_reorder(op, Mean,
+                                          .fun = mean)) %>%
+  mutate(verb = fct_reorder(verb, Mean,
+                            .fun = mean)) %>%
+  ggplot(aes(x = op, y=Mean, group = verb)) +
+  coord_cartesian(ylim = c(0,1)) +
+  facet_wrap(vars(verb)) +
+  geom_point(size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") +
+  labs(title = "Mean projectivity by operator, for each verb")+
+  theme_bw()
+
+@
+Some more generalizations: We find different \lq profiles\rq for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:
+
+\begin{itemize}
+  \item \emph{pretend, think}: \textbf{anti-veridical profile}\\
+    \texttt{N} $>$ \texttt{M, Q, C}, overall low projectivity
+
+  \item \emph{acknowledge, hear, inform, see, discover, know, be annoyed}: \textbf{\lq factive\rq\ profile}\\
+    \texttt{Q} $>$ \texttt{N, C} \textcolor{gray}{$>_?$} \texttt{M}, overall high projectivity (it may be possible to find further subgroups here)
+
+  \item \emph{prove, confirm, establish, demonstrate, (announce), confess, admit, reveal}: \textbf{veridical profile}\\
+    \texttt{M, C} $>$ \texttt{Q} $>$ \texttt{N}, overall med-lo to med-hi projectivity
+
+  \item \emph{be right, suggest}: \textbf{reportative profile}\\
+    \texttt{M, N, C} $>$ \texttt{Q}
+
+\end{itemize}
+
+Maybe these can have better names, not trying to suggest that verbs can neatly divided in factive v non-factive, but potentially this class / profile is what prompted intuitions in previous literature, and naming in this tradition could make sense, but can be changed depending on our rhetoric, of course.
+
+% \section{Analysis}
+% 
+% We are interested in how the response (projectivity ratings) depends on the verb and embedding operator, and whether the above generalizations can be supported statistically.
+% 
+% 
+% 
+% <<>>=
+% formula <- response ~ op * verb
+% # formula <- response ~ op * verb + (1 | workerid) + (1 | item)
+% # formula <- response ~ op * verb + (1 + op + verb | workerid) + (1 + op + verb | item)
+% @
+% 
+% Trying a beta-regression.
+Our response variable has observations from a sliding scale between $(0,1)$.
+For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).
+% 
+% <<>>=
+% data$response = (data$projective*(nrow(data)-1) + .5)/nrow(data)
+% @
+% 
+% 
+% <<>>=
+% # library(rstan)
+% # options(mc.cores=parallel::detectCores())
+% # rstan_options(auto_write=TRUE)
+% # library(brms)
+% # 
+% # model1 <- brm(formula = bf(response ~ op * verb ,
+% #                            phi ~ op * verb,
+% #                            family = Beta()),
+% #               family = Beta(),
+% #               data = data,
+% #               iter=10000, warmup=2000,chains=6)
+% 
+% 
+% @
+% 
 %
 
-<<>>=
-t = data_all %>% 
-  mutate(verb = fct_relevel(verb, "suggest")) %>%
-  mutate(op = fct_relevel(op, "q"))
-
-#m.dummy = lm(projective ~ ai*short_trigger, data=t)
-#summary(m.dummy)
-
-# reference level set to "prove"
-
-# model with short_trigger slope did not converge
-m.4 = lmer(projective ~ cai*short_trigger + (1+cai|content) + (1+cai|workerid), 
-           data = t, REML=F,control = lmerControl(optimizer ='optimx', optCtrl=list(method='L-BFGS-B')))
-saveRDS(m.4, "../models/m.4.rds")
-m.4 <- readRDS("../models/m.4.rds")
-summary(m.4)
-texreg(m.4,single.row = TRUE, leading.zero = FALSE, booktabs = TRUE)
-@
-
 \end{document}

### 2022-11-23 materials-lisa/analysis/3_projection-across-operators/operators.R

@@ -1,268 +1,0 @@
-
-### comparing projectivity across entailment-cancelling operators
-
-# set working directory to directory of script
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-# for exps 1: compare Q / N / M / C; 1-4
-
-# load data from exps 1
-data_1q <- read.csv("../../../results/main/1_projaiQ/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_1n <- read.csv("../../../results/main/2_projaiN/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_1m <- read.csv("../../../results/main/3_projaiM/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_1c <- read.csv("../../../results/main/4_projaiC/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-
-# load data from exps 2
-data_2q <- read.csv("../../../results/main/5_projaiQ/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_2n <- read.csv("../../../results/main/6_projaiN/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_2m <- read.csv("../../../results/main/7_projaiM/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_2c <- read.csv("../../../results/main/8_projaiC/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-
-# load data from exps 3
-data_3q <- read.csv("../../../results/main/9_projaiQ/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_3n <- read.csv("../../../results/main/10_projaiN/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_3m <- read.csv("../../../results/main/11_projaiM/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-data_3c <- read.csv("../../../results/main/12_projaiC/data/data_preprocessed.csv", 
-                    header=TRUE, sep=",")
-
-df_list <- list(data_1q=data_1q, data_1m=data_1m, data_1n=data_1n, data_1c=data_1c,
[…]

### 2022-11-23 results/main/meta-analyses/projectivity/rscripts/graphs.R

@@ -24,12 +24,11 @@ proj_means_op = data %>% group_by(op) %>%
 
 proj_means_op %>% mutate(op = fct_reorder(op, Mean, 
                                           .fun = mean)) %>% 
-  ggplot(aes(x = op, y=Mean, color = op)) +
+  ggplot(aes(x = op, y=Mean)) +
   # coord_cartesian(ylim = c(0,1)) +
-  geom_point(aes(shape = op), size = 4) + 
-  scale_shape_manual(values = c("M", "N", "Q", "C")) +
-  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
-  geom_line() + 
+  geom_point(aes(shape = op), size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") + 
   labs(title = "Mean projectivity by operator")+
   theme_bw() +
   scale_color_brewer(palette = "PRGn")
@@ -87,12 +86,11 @@ proj_means %>% mutate(op = fct_reorder(op, Mean,
   mutate(verb = fct_reorder(verb, Mean, 
                             .fun = mean)) %>% 
   ggplot(aes(x = op, y=Mean, group = verb)) +
-  # coord_cartesian(ylim = c(0,1)) +
+  coord_cartesian(ylim = c(0,1)) +
   facet_wrap(vars(verb)) +
-  geom_point(size = 1) + 
-  scale_shape_manual(values = c("M", "N", "Q", "C")) +
-  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
-  geom_line() + 
+  geom_point(size = 1, color = "lightblue") + 
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") + 
   labs(title = "Mean projectivity by operator, for each verb")+
   theme_bw()
 
@@ -102,7 +100,7 @@ proj_means %>% mutate(verb = fct_reorder(verb, Mean,
                                        .fun = mean)) %>% 
   
   ggplot(aes(x = verb, y=Mean, color = op, group = op)) +
-  # coord_cartesian(ylim = c(0,1)) +
+  coord_cartesian(ylim = c(0,1)) +
   facet_grid(rows = vars(op)) +
   geom_point(aes(shape = op), size = 4) + 
   scale_shape_manual(values = c("M", "N", "Q", "C")) +

### 2022-11-28 SALT-abstract/SALT-abstract.tex
updates to analysis and added abstract file

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\documentclass[12pt, a4paper]{scrartcl}

%% Page Layout
\usepackage[margin=1in]{geometry}

\usepackage{euler} % math font package needs to be loaded before others
\usepackage{xunicode,xltxtra, polyglossia}
\setdefaultlanguage[variant=american]{english}

%% fonts, symbols, text
	%%% fonts
	\usepackage{fontspec} %(include if mathspec is not loaded)
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	%%% text decoration
	\usepackage[normalem]{ulem} % sout
	%%% semantics symbols
	\usepackage{stmaryrd}
	\usepackage{amsmath,amssymb}
	\newcommand{\transl}{\rightsquigarrow \ensuremath}
	%%% other symbols
	\usepackage{pifont}% http://ctan.org/pkg/pifont
	\newcommand{\cmark}{\ding{51}}%
	\newcommand{\xmark}{\ding{55}}%

%% layout
%%% page layout
\usepackage{multicol}

%% bibliography
\usepackage[round]{natbib}
\newcommand{\posscite}[1]{\citeauthor{#1}'s (\citeyear{#1})}

%% figures, examples, diagrams
%%% examples
\usepackage{linguex}
\renewcommand{\firstrefdash}{}
%%% tables 
\usepackage{booktabs}
%%% figures
\usepackage{graphics}

%% decoration and features
%%% colors
\usepackage[dvipsnames]{xcolor}

%% bibliography
\renewcommand*{\refname}{\normalsize\textbf{References}\\ \vspace{-.5\baselineskip}}

%% fonts
\setmainfont[Scale=MatchLowercase,Mapping=tex-text,SmallCapsFont={TeX Gyre Termes}, SmallCapsFeatures={Letters=SmallCaps}]{Times New Roman}
\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Times New Roman}

\usepackage{setspace}
\begin{document}

\bibliographystyle{plainnat}
% \enablehyphenation
% \vspace{-2em}
% \maketitle
\textcolor{white}{.} \vspace{-2.9\baselineskip} \\
\begin{center}
	\textbf{\large%\thetitle
		A diverse family (of sentences)\\ Projectivity differs across embedding operators---but not like you think}
\end{center}
% \vspace{-.6\baselineskip}
Attitude verbs can often be associated with an inference that their complement clause is true, even when embedded under an entailment cancelling operator (in which case the inference to the truth of the complement is said to \emph{project}, see karttunen, kiparski kiparski, on factive verbs). This is illustrated for \emph{\lq discover\rq} in \ref{ex:family}:

\ex. \label{ex:family} Projection across various entailment-cancelling operators:
	\a. Polar Questions:\\
		\emph{\lq Did Cole discover that Julian dances Salsa?\rq}
	\b. Negation:\\
		\emph{\lq Cole didn't discover that Julian dances Salsa.\rq}
	\b. Modals:\\
		\emph{\lq Perhaps Cole discovered that Julian dances Salsa.\rq}
	\b. Conditionals:\\
		\emph{\lq If Cole discovered that Julian dances Salsa, Logan will be joyful.\rq}
	\z.
\z.

Tonhauser (ETC, REFERENCES) showed that whether or not this inference projects is not a categorial property of lexical triggers (references); but a gradient property, which is also affected by various contextual factors (references). In light of this gradience, and the complex interaction of projection with various contextual factors, we may also expect the semantically quite hetergeneous entailment-cancelling operators in \ref{ex:family} to affect projection in differential ways.

Our work presents experimental data from a study addressing the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments? \textbf{(ii)} Are these effects the same or different for different verbs? In an experimental task designed to assess speaker commitment to the complement clause, and therefore projection, we find that the 
differences in entailment-cancelling operator differentially affect projection, and that there are differences between triggers in their interaction with embedding operators.

Karttunen: \emph{\lq discover\rq} is a semi-factive verb: generalizations 
\begin{itemize}
	\item more projective under negation than questions
	\item factive + semi-factive: two classes of projective verbs with divergent projection behavior under different embedding operators
	\item Karttunen's classification of factive/semi-factive verbs is often assumed
	\item Kajsa Djärv: cognitive vs emotive predicates
	\item experimental work by (see email)
\end{itemize}

\noindent To investigate differences in projection behavior across various entailment-cancelling operators as in \ref{ex:family}, and between various types of attitude verbs, we used a

speaker commitment about the embedded clause. E.g. one of the sentences in \ref{ex:family} would be presented as asserted by a named speaker (on screen as \textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}). Participants were then asked to provide a judgment in response to a prompt like: \emph{Is Daniel certain that Julian dances Salsa?}), by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).

This work presents experimental data about inferences about the truth of clauses that are embedded by attudes and various entailment-cancelling operators, .
\begin{itemize}
	\item speaker commitment about clauses embedded by attitude verb
	\item manipulating choice of attitude verbs, and embedding operators
\end{itemize}

\begin{itemize}
	\item projective content, presupposition projection, 
\end{itemize}

\vspace{-1.2\baselineskip}
\paragraph{Method.} \hspace{-0.95em}, 
	dssdf
% paragraph i_contextual_i (end)

\pagebreak
\bibliography{../discourse-negativity}

\end{document}

### 2022-11-28 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
updates to analysis and added abstract file

@@
     \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
-    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here
+    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive.
+The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
   \end{itemize}
@@
 @
-Some more generalizations: We find different \lq profiles\rq for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:
+Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:
 
@@
 
-% \section{Analysis}
+\section{Analysis}
+
+We are interested in how the response (projectivity ratings) depends on the verb and embedding operator, and whether the above generalizations can be supported statistically.\\
+
+% \subsection{Frequentist modeling}
+Exploring a couple of models
+
+\subsection{Linear frequentist with \texttt{suggest/m} as baseline}
+
+<<freq-linear-suggest-m, message=FALSE, warning=FALSE>>=
+
+# our independent variables
+## inference-triggering predicate, coded as "verb"
+data$verb <- as.factor(data$verb)
+levels(data$verb)
+# contrasts(data$verb)
+### dummy coding with "suggest" as baseline
+data$verb <- relevel(data$verb, ref = "suggest")
+# contrasts(data$verb)
+
+## embedding operator, coded as "op"
+data$op <- as.factor(data$op)
+levels(data$op)
+# contrasts(data$op)
+### dummy coding with "m" as baseline
+data$op <- relevel(data$op, ref = "m")
+# contrasts(data$op)
+
+# coding random effects as factors
+data$workerid <- as.factor(data$workerid)
+data$content <- as.factor(data$content)
+
+library(lme4)
+library(lmerTest)
+
+# glmm1 <- lmer(projective ~ verb * op + (1 | workerid) +
+#                  (1 | content), data=data)
+
+# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+print(summary(glmm1), cor=F, dig=3)
+@
+
+\subsection{Linear frequentist with \texttt{annoyed/q} as baseline}
+<<freq-linear-annoyed-q, message=FALSE>>=
+
+# dummy coding for "verb" with "suggest" as baseline
+data$verb <- relevel(data$verb, ref = "be_annoyed")
+
+### dummy coding for "op" with "q" as baseline
+data$op <- relevel(data$op, ref = "q")
+
+glmm2a <- lmer(projective ~ verb * op + (1 | workerid) +
+                 (1 | content), data=data)
+
+# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+print(summary(glmm2a), cor=F, dig=3)
+
+@
+
+\subsection{Linear frequentist with \texttt{prove/n} as baseline}
+<<freq-linear-annoyed-q, message=FALSE>>=
+
+# dummy coding for "verb" with "prove" as baseline
+data$verb <- relevel(data$verb, ref = "prove")
+
+### dummy coding for "op" with "n" as baseline
+data$op <- relevel(data$op, ref = "n")
+
+# glmm3 <- lmer(projective ~ verb * op + (1 | workerid) +
+#                  (1 | content), data=data)
+# 
+# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+print(summary(glmm3), cor=F, dig=3)
+
+@
+
+\subsection{Linear frequentist with \texttt{know/m} as baseline}
+<<freq-linear-annoyed-q, message=FALSE>>=
+
+# dummy coding for "verb" with "prove" as baseline
+data$verb <- relevel(data$verb, ref = "know")
+
+### dummy coding for "op" with "n" as baseline
+data$op <- relevel(data$op, ref = "m")
+
+# glmm4 <- lmer(projective ~ verb * op + (1 | workerid) +
+#                  (1 | content), data=data)
+# 
+# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+print(summary(glmm4), cor=F, dig=3)
+
+@
+
+% \subsection{Bayesian beta-regression with \texttt{suggest/m} as baseline}
 % 
-% We are interested in how the response (projectivity ratings) depends on the verb and embedding operator, and whether the above generalizations can be supported statistically.
+% Our response variable has observations from a sliding scale between $(0,1)$.
+For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).
+% 
+% <<transform-for-beta>>=
+% data$response = (data$projective*(nrow(data)-1) + .5)/nrow(data)
+% @
+% 
+% <<betareg>>=
+% library(rstan)
+% options(mc.cores=parallel::detectCores())
+% rstan_options(auto_write=TRUE)
+% library(brms)
 % 
@@
 % 
+% # bbmm1 <- brm(formula = bf(response ~ op * verb, # beta model mean parameter (mu)
+% #                             phi ~ op * verb, # beta model precision parameter (phi)
+% #                             family = Beta()),
+% #                family = Beta(),
+% #                data = data,
+% #                iter=10000, warmup=2000,chains=6)
+% 
+% @
+
+% \subsubsection{Linear frequentist with difference coding}
+% Make sure I understand how difference coding works
+% This should help
+% https://stats.oarc.ucla.edu/r/library/r-library-contrast-coding-systems-for-categorical-variables/#backward
+
+% https://rdrr.io/github/tsostarics/contrastable/src/R/forward_difference_code.R
+
 % <<>>=
+% # frequentist fixed effects only, input error: no matrix
+% # bm1 <- betareg(
+% #   response ~ verb * op | # beta model mean parameter (mu)
+% #     response ~ verb * op, # beta model precision parameter
+% #   data = data,
+% #   link = "logit"
+% # )
 % formula <- response ~ op * verb
@@
 % Trying a beta-regression. 
-Our response variable has observations from a sliding scale between $(0,1)$.
-For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).
-% 
-% <<>>=
-% data$response = (data$projective*(nrow(data)-1) + .5)/nrow(data)
-% @
-% 
-% 
 % <<>>=

### 2022-11-28 results/main/meta-analyses/projectivity/rscripts/analysis.R

@@ -18,7 +18,6 @@ data$projective <- numeric(data$projective)
 data$verb <- as.factor(data$verb)
 levels(data$verb)
 contrasts(data$verb)
-
 ### dummy coding with "suggest" as baseline
 data$verb <- relevel(data$verb, ref = "suggest")
 contrasts(data$verb)
@@ -30,7 +29,9 @@ contrasts(data$verb)
 data$op <- as.factor(data$op)
 levels(data$op)
 contrasts(data$op)
-### dummy coding with "c" (conditionals) as baseline
+### dummy coding with "m" (conditionals) as baseline
+data$op <- relevel(data$op, ref = "m")
+contrasts(data$op)
 
 # coding random effects as factors
 data$workerid <- as.factor(data$workerid)
@@ -42,13 +43,15 @@ data$content <- as.factor(data$content)
 library(lme4)
 library(lmertest)
 
-# glmm1 <- lmer(projective ~ op * verb + (op | workerid) + 
+# glmm1 <- lmer(projective ~ verb * op + (op | workerid) +
 #                  (op | content), data=data)
 
 # save.image("linear-models.RData")
 load("linear-models.RData")
 print(summary(glmm1), cor=F, dig=3)
 
+# glmm2 <- lmer(projective ~ verb * op + (op | workerid) +
+#                 (op | content), data=data)
 
-
-
+# save.image("linear-models.RData")
+print(summary(glmm2), cor=F, dig=3)

### 2022-11-30 SALT-abstract/SALT-abstract.tex
updates to abstract draft

@@
 \end{center}
-% \vspace{-.6\baselineskip}
-Attitude verbs can often be associated with an inference that their complement clause is true, even when embedded under an entailment cancelling operator (in which case the inference to the truth of the complement is said to \emph{project}, see karttunen, kiparski kiparski, on factive verbs).
+
+\noindent This work investigates the projection of inferences to the truth of an embedded complement clause across different entailment-cancelling operators, for a variety of attitude verbs.
+We present experimental evidence that projection does vary by operator, and further, that the effect of entailment-cancelling operators differs between various attitude verb triggers.
+However, our findings about the interaction of operator and trigger do not support \posscite{karttunen71b} classification of factive vs.
+semi-factive verbs.
+
+\paragraph{Projection across various entailment-cancelling operators.}
+	Attitude verbs can often be associated with an inference that their complement clause is true, even when embedded under an entailment cancelling operator (in which case the inference to the truth of the complement is said to \emph{project}, see e.g.
+\citealp{karttunen71b}).
 This is illustrated for \emph{\lq discover\rq} in \ref{ex:family}:
@@
 
-Tonhauser (ETC, REFERENCES) showed that whether or not this inference projects is not a categorial property of lexical triggers (references); but a gradient property, which is also affected by various contextual factors (references).
-In light of this gradience, and the complex interaction of projection with various contextual factors, we may also expect the semantically quite hetergeneous entailment-cancelling operators in \ref{ex:family} to affect projection in differential ways.
-
-Our work presents experimental data from a study addressing the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
-\textbf{(ii)} Are these effects the same or different for different verbs?
-In an experimental task designed to assess speaker commitment to the complement clause, and therefore projection, we find that the
-differences in entailment-cancelling operator differentially affect projection, and that there are differences between triggers in their interaction with embedding operators.
+	Previous work on projective content showed that whether or not this inference projects is not a categorial property of lexical triggers \citep{tbd-variability}; but a gradient property, which is also affected by various contextual factors \citep{brst-salt10,brst-ar,degen-tonhauser-openmind}.
+In light of this gradience, and the complex interaction of projection with various contextual factors, we may also expect the semantically hetergeneous entailment-cancelling operators in \ref{ex:family} to affect projection in differential ways.\\
 
@@
 		\item Kajsa Djärv: cognitive vs emotive predicates
-	\item experimental work by (see email)
 	\end{itemize}
 
-\noindent To investigate differences in projection behavior across various entailment-cancelling operators as in \ref{ex:family}, and between various types of attitude verbs, we used a
+	Our work presents experimental data from a study addressing the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
+\textbf{(ii)} Do these effects vary for different verbs?
+In an experimental task designed to assess speaker commitment to the complement clause, we find that the choice of entailment-cancelling operator affects projection, and that there are differences between triggers in their interaction with embedding operators.
 
-speaker commitment about the embedded clause.
-E.g.
-one of the sentences in \ref{ex:family} would be presented as asserted by a named speaker (on screen as \textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}).
+\paragraph{Experimental investigation.} % (fold)
+	To investigate differences in projection behavior across various entailment-cancelling operators as in \ref{ex:family}, and between various types of attitude verbs, we used a response task which elicitied participant judgments about how strongly they thought a speaker would be committed towards the embedded clause.
+We presented sentences like in \ref{ex:family} as asserted by a named speaker (on screen as \textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}).
 Participants were then asked to provide a judgment in response to a prompt like: \emph{Is Daniel certain that Julian dances Salsa?}), by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
 
-This work presents experimental data about inferences about the truth of clauses that are embedded by attudes and various entailment-cancelling operators, .
+% paragraph experiment (end)
+
+\paragraph{Design and Expectations.} % (fold)
+	\ex. 20 clause-embedding predicates 
+		\a. canonically factive: {\em be annoyed, discover, know, reveal, see}
+		\b. nonfactive:
+			\a. nonveridical nonfactive: {\em pretend, suggest, say, think}
+			\b. veridical nonfactive: {\em be right, demonstrate}
+			\z.
+		\b. optionally factive: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove}
+		\z.
+	\z.
+
 	\begin{itemize}
-	\item speaker commitment about clauses embedded by attitude verb
-	\item manipulating choice of attitude verbs, and embedding operators
+		\item Expectations based on Karttunen:
+		\begin{itemize}
+			\item no variation for factive verbs
+			\item semi-factives (like \emph{discover}) should project more strongly across negation than questions
 		\end{itemize}
 
+		\item expectations based on Djärv:
 		\begin{itemize}
-	\item projective content, presupposition projection, 
+			\item emotive predicates (like \emph{be annoyed}) behave like Karttunen's factives
+			\item cognitive predicates (like \emph{discover, see\dots}) behave like Karttunen's semi-factives
+		\end{itemize}
 	\end{itemize}
 
-\vspace{-1.2\baselineskip}
-\paragraph{Method.} \hspace{-0.95em}, 
-	dssdf
-% paragraph i_contextual_i (end)
+% paragraph design_and_expectations_ (end)
+
+\paragraph{Method.} % (fold)
+	\dots
+% paragraph method (end)
+
+\paragraph{Results and Analysis.} % (fold)
+	\begin{itemize}
+		\item pooled data across experiments
+	\end{itemize}
+
+	\textbf{Figure \ref{fig:figure1}} shows the mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals.
+
+	\begin{figure}[h]
+		\centering
+		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}
+		\caption{Mean certainty ratings by operator and predicate}
+		\label{fig:figure1}
+	\end{figure}
+
+	\begin{itemize}
+		\item The embedding operator affects projection (main effect of embedding operator)
+		\item This effect differs by verb (interaction effects of operator/verb)
+	\end{itemize}
+% paragraph method (end)
+
+\paragraph{Discussion and Conclusion.} % (fold)
+	\dots
+% paragraph method (end)
 
 \pagebreak
-\bibliography{../discourse-negativity}
+\bibliography{projective.bib}
 

### 2022-11-30 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
updated analysis files

@@
 
-glmm2a <- lmer(projective ~ verb * op + (1 | workerid) +
-                 (1 | content), data=data)
+# glmm2a <- lmer(projective ~ verb * op + (1 | workerid) +
+#                  (1 | content), data=data)
 
 # save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
-load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+# load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
 print(summary(glmm2a), cor=F, dig=3)
@@
 @
-
-\subsection{Linear frequentist with \texttt{prove/n} as baseline}
-<<freq-linear-annoyed-q, message=FALSE>>=
-
-# dummy coding for "verb" with "prove" as baseline
-data$verb <- relevel(data$verb, ref = "prove")
-
-### dummy coding for "op" with "n" as baseline
-data$op <- relevel(data$op, ref = "n")
-
-# glmm3 <- lmer(projective ~ verb * op + (1 | workerid) +
-#                  (1 | content), data=data)
-# 
-# save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
-load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
-print(summary(glmm3), cor=F, dig=3)
-
-@
+% 
+% \subsection{Linear frequentist with \texttt{prove/n} as baseline}
+% % <<freq-linear-prove-n, message=FALSE>>=
+% # dummy coding for "verb" with "prove" as baseline
+% data$verb <- relevel(data$verb, ref = "prove")
+% 
+% ### dummy coding for "op" with "n" as baseline
+% data$op <- relevel(data$op, ref = "n")
+% 
+% # glmm3 <- lmer(projective ~ verb * op + (1 | workerid) +
+% #                  (1 | content), data=data)
+% #
+% # save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+% # load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+% print(summary(glmm3), cor=F, dig=3)
+% 
+% @
 
 \subsection{Linear frequentist with \texttt{know/m} as baseline}
-<<freq-linear-annoyed-q, message=FALSE>>=
+<<freq-linear-know-m, message=FALSE>>=
 
@@
 # save.image("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
-load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
+# load("../../../results/main/meta-analyses/projectivity/rscripts/linear-models.RData")
 print(summary(glmm4), cor=F, dig=3)

### 2022-11-30 results/main/meta-analyses/projectivity/rscripts/analysis.R

@@ -55,3 +55,5 @@ print(summary(glmm1), cor=F, dig=3)
 
 # save.image("linear-models.RData")
 print(summary(glmm2), cor=F, dig=3)
+
+print(summary(glmm1), cor=F, dig=3)

### 2022-11-30 SALT-abstract/SALT-abstract.tex
condensed abstract intro

@@
 
-\noindent This work investigates the projection of inferences to the truth of an embedded complement clause across different entailment-cancelling operators, for a variety of attitude verbs.
-We present experimental evidence that projection does vary by operator, and further, that the effect of entailment-cancelling operators differs between various attitude verb triggers.
-However, our findings about the interaction of operator and trigger do not support \posscite{karttunen71b} classification of factive vs.
+\vspace{-.5\baselineskip}
+\noindent We present experimental evidence that the projectivity of attitude complements varies across different entailment-cancelling operators, and that this effect of entailment-cancelling operator differs between various attitude verb triggers.
+However, our findings about the interaction of operator and trigger do not support \posscite{karttunen_observations_1971} classification of factive vs.
 semi-factive verbs.
 
+\vspace{-\baselineskip}
 \paragraph{Projection across various entailment-cancelling operators.}
-	Attitude verbs can often be associated with an inference that their complement clause is true, even when embedded under an entailment cancelling operator (in which case the inference to the truth of the complement is said to \emph{project}, see e.g.
-\citealp{karttunen71b}).
-This is illustrated for \emph{\lq discover\rq} in \ref{ex:family}:
+	Certain attitude ascriptions come with an inference to the truth of their complement, even if embedded under entailment-cancelling operators (illustrated for \emph{\lq discover\rq} in \ref{ex:family}), in which case this inference is said to \emph{project}, (e.g.
+\citealp{karttunen_observations_1971}).
 
-	\ex. \label{ex:family} Projection across various entailment-cancelling operators:
-		\a. Polar Questions:\\
+	\vspace{-.25\baselineskip}
+	\ex. \label{ex:family}
+		\a. Polar Questions: \hfill
 			\emph{\lq Did Cole discover that Julian dances Salsa?\rq}
-		\b. Negation:\\
+		\b. Negation: \hfill
 			\emph{\lq Cole didn't discover that Julian dances Salsa.\rq}
-		\b. Modals:\\
+		\b. Modals: \hfill
 			\emph{\lq Perhaps Cole discovered that Julian dances Salsa.\rq}
-		\b. Conditionals:\\
+		\b. Conditionals: \hfill
 			\emph{\lq If Cole discovered that Julian dances Salsa, Logan will be joyful.\rq}
@@
 
-	Previous work on projective content showed that whether or not this inference projects is not a categorial property of lexical triggers \citep{tbd-variability}; but a gradient property, which is also affected by various contextual factors \citep{brst-salt10,brst-ar,degen-tonhauser-openmind}.
-In light of this gradience, and the complex interaction of projection with various contextual factors, we may also expect the semantically hetergeneous entailment-cancelling operators in \ref{ex:family} to affect projection in differential ways.\\
+	\vspace{-.25\baselineskip}
+	Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}.
+In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.
 
-	Karttunen: \emph{\lq discover\rq} is a semi-factive verb: generalizations 
-	\begin{itemize}
-		\item more projective under negation than questions
-		\item factive + semi-factive: two classes of projective verbs with divergent projection behavior under different embedding operators
-		\item Karttunen's classification of factive/semi-factive verbs is often assumed
-		\item Kajsa Djärv: cognitive vs emotive predicates
-	\end{itemize}
+	Generalizations to this effect were suggested by \citet{karttunen_observations_1971}, who distinguishes between \emph{factive} verbs (\emph{regret, forget, resent}) and \emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}), suggesting that factives always project, while semi-factives always project across negation, but not always in polar questions or conditionals.
+SMITH study, Djärv on conitive/emotive difference.
+	We experimentally test these generaliztions in a study addressing the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
+\textbf{(ii)} Do these effects vary for different verbs (and in what way)?
 
-	Our work presents experimental data from a study addressing the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
-\textbf{(ii)} Do these effects vary for different verbs?
-In an experimental task designed to assess speaker commitment to the complement clause, we find that the choice of entailment-cancelling operator affects projection, and that there are differences between triggers in their interaction with embedding operators.
+\vspace{-\baselineskip}
+\paragraph{Experimental investigation.} % (fold)
+	% Using an experimental task designed to assess speaker commitment to the complement clause, we find that the choice of entailment-cancelling operator affects projection, and that there are differences between triggers in their interaction with embedding operators.
 
-\paragraph{Experimental investigation.} % (fold)
-	To investigate differences in projection behavior across various entailment-cancelling operators as in \ref{ex:family}, and between various types of attitude verbs, we used a response task which elicitied participant judgments about how strongly they thought a speaker would be committed towards the embedded clause.
-We presented sentences like in \ref{ex:family} as asserted by a named speaker (on screen as \textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}).
-Participants were then asked to provide a judgment in response to a prompt like: \emph{Is Daniel certain that Julian dances Salsa?}), by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
+	To investigate differences in projection behavior across various en- tailment-cancelling operators as in \ref{ex:family}, and between various types of attitude verbs, we used a response task which elicitied participant judgments about how strongly they thought a speaker would be committed towards the embedded clause.
+We presented sentences like in \ref{ex:family} as asserted by a named speaker (on screen as e.g.
+“\textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}”).
+Participants were then asked to provide a judgment in response to a prompt like: \emph{“Is Daniel certain that Julian dances Salsa?”}), by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
 
@@
 \pagebreak
-\bibliography{projective.bib}
+\bibliography{projective-content.bib}
 

### 2022-12-05 SALT-abstract/SALT-abstract.tex
abstract updates

@@
 % \maketitle
-\textcolor{white}{.} \vspace{-2.9\baselineskip} \\
+\textcolor{white}{.} \vspace{-3.9\baselineskip} \\
 \begin{center}
@@
 
-\vspace{-.5\baselineskip}
+\vspace{-.4\baselineskip}
 \noindent We present experimental evidence that the projectivity of attitude complements varies across different entailment-cancelling operators, and that this effect of entailment-cancelling operator differs between various attitude verb triggers.
@@
 \vspace{-\baselineskip}
-\paragraph{Projection across various entailment-cancelling operators.}
-	Certain attitude ascriptions come with an inference to the truth of their complement, even if embedded under entailment-cancelling operators (illustrated for \emph{\lq discover\rq} in \ref{ex:family}), in which case this inference is said to \emph{project}, (e.g.
+\paragraph{Projection across entailment-cancelling operators.}  \hspace{-1em}
+	Certain attitude ascriptions come with an inference to the truth of their complement, even if embedded under entailment-cancelling operators (shown for \emph{\lq discover\rq} in \ref{ex:family}), in which case the inference is said to \emph{project} (e.g.
 \citealp{karttunen_observations_1971}).
 
-	\vspace{-.25\baselineskip}
+	\vspace{-.3\baselineskip}
 	\ex. \label{ex:family}
-		\a. Polar Questions: \hfill
+		\a. \label{ex:mod}
+			Modals: \hfill
+			\emph{\lq Perhaps Cole discovered that Julian dances Salsa.\rq}
+		\b. \label{ex:neg}
+			Negation: \hfill
+			\emph{\lq Cole didn't discover that Julian dances Salsa.\rq}
+		\b. \label{ex:q}
+			Polar Questions: \hfill
 			\emph{\lq Did Cole discover that Julian dances Salsa?\rq}
-		\b. Negation: \hfill
-			\emph{\lq Cole didn't discover that Julian dances Salsa.\rq}
-		\b. Modals: \hfill
-			\emph{\lq Perhaps Cole discovered that Julian dances Salsa.\rq}
-		\b. Conditionals: \hfill
+		\b. \label{ex:cond}
+			Conditionals: \hfill
 			\emph{\lq If Cole discovered that Julian dances Salsa, Logan will be joyful.\rq}
@@
 
-	\vspace{-.25\baselineskip}
-	Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}.
+	\vspace{-.4\baselineskip}
+	Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}.
 In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.
 
-	Generalizations to this effect were suggested by \citet{karttunen_observations_1971}, who distinguishes between \emph{factive} verbs (\emph{regret, forget, resent}) and \emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}), suggesting that factives always project, while semi-factives always project across negation, but not always in polar questions or conditionals.
-SMITH study, Djärv on conitive/emotive difference.
-	We experimentally test these generaliztions in a study addressing the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
+	\citet{karttunen_observations_1971} proposed generalizations to this effect, distinguishing between \emph{factive} verbs (\emph{regret, forget, resent}) and \emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}), suggesting that factives always project, while semi-factives always project across negation, but not always in polar questions or conditionals.
+To provide a systematic way of distinguishing these classes, \cite{djarv_cognitive_2018} suggest that they correspond to emotive and cognitive predicates, respectively.
+Regarding the effect of operators on projection, \cite{smith_relationship_2014} find an effect of ..
+on ...
+(?) Here, we experimentally address the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
 \textbf{(ii)} Do these effects vary for different verbs (and in what way)?
@@
 \vspace{-\baselineskip}
-\paragraph{Experimental investigation.} % (fold)
-	% Using an experimental task designed to assess speaker commitment to the complement clause, we find that the choice of entailment-cancelling operator affects projection, and that there are differences between triggers in their interaction with embedding operators.
-
-	To investigate differences in projection behavior across various en- tailment-cancelling operators as in \ref{ex:family}, and between various types of attitude verbs, we used a response task which elicitied participant judgments about how strongly they thought a speaker would be committed towards the embedded clause.
-We presented sentences like in \ref{ex:family} as asserted by a named speaker (on screen as e.g.
+\paragraph{Experimentally investigating projection,} \hspace{-1em}
+	we used a response task to elicit judgments about how strongly a speaker would be committed towards the embedded clause \citep[from][]{tonhauser_prosodic_2016}.
+We presented sentences like in \ref{ex:family} as asserted by a named speaker (e.g.
 “\textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}”).
-Participants were then asked to provide a judgment in response to a prompt like: \emph{“Is Daniel certain that Julian dances Salsa?”}), by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
+Participants then provided a certainty-rating in response to a prompt like: \emph{“Is Daniel certain that Julian dances Salsa?”}, by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
 
@@
 
-\paragraph{Design and Expectations.} % (fold)
-	\ex. 20 clause-embedding predicates 
-		\a. canonically factive: {\em be annoyed, discover, know, reveal, see}
-		\b. nonfactive:
-			\a. nonveridical nonfactive: {\em pretend, suggest, say, think}
-			\b. veridical nonfactive: {\em be right, demonstrate}
-			\z.
-		\b. optionally factive: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove}
-		\z.
-	\z.
-
-	\begin{itemize}
-		\item Expectations based on Karttunen:
-		\begin{itemize}
-			\item no variation for factive verbs
-			\item semi-factives (like \emph{discover}) should project more strongly across negation than questions
-		\end{itemize}
-
-		\item expectations based on Djärv:
-		\begin{itemize}
-			\item emotive predicates (like \emph{be annoyed}) behave like Karttunen's factives
-			\item cognitive predicates (like \emph{discover, see\dots}) behave like Karttunen's semi-factives
-		\end{itemize}
-	\end{itemize}
+\vspace{-\baselineskip}
+\paragraph{Design and Expectations.} \hspace{-1em}
+	We compared certainty-ratings for the four entailment-canceling operators in \ref{ex:family}, and 20 clause-embedding predicates (\texttt{verb}: {\em be annoyed, discover, know, reveal, see, acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, pretend, suggest, say, think, be right, demonstrate}).
+	Based on the Karttunen-Djärv generalization about emotive factives vs.
+cognitive semi-factives, we would expect emotive factives (\emph{be annoyed}), and verbs normally taken to be factive (\emph{know}), to be highly projective in a way that is indifferent to the embedding context, and cognitive semi-factives (\emph{discover, see}) to show higher projectivity ratings under negation compared to questions and conditionals.
 
@@
 
-\paragraph{Method.} % (fold)
-	\dots
+\vspace{-\baselineskip}
+\paragraph{Method.} \hspace{-1em}
+	The study consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause.
+Experiments 1--3 used polar question embedding, exps.
+4--6 used negation, 7--9 modals, and 10--12 conditionals, making the \texttt{operator} manipulation a between-subjects factor.
+We analyzed data from 2682 self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk).
+Participants saw items Latin-squared and randomized with six control stimuli.
+
 % paragraph method (end)
 
-\paragraph{Results and Analysis.} % (fold)
-	\begin{itemize}
-		\item pooled data across experiments
-	\end{itemize}
-
-	\textbf{Figure \ref{fig:figure1}} shows the mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals.
+\paragraph{Results and Analysis.} \hspace{-1em}
+	Pooling the data across our 12 experiments, we examined the effect on certainty-ratings of \texttt{operator}, and \texttt{verb}, as well as their interaction.
+Mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals are shown in  \textbf{Figure \ref{fig:figure1}}.
 
 	\begin{figure}[h]
+		\vspace{-.8\baselineskip}
 		\centering
-		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}
-		\caption{Mean certainty ratings by operator and predicate}
+		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
+		\caption{\small Mean certainty ratings by predicate and operator}
 		\label{fig:figure1}
@@
 
+	\vspace{-.3\baselineskip}
+	% \textbf{version w baseline know/m}
+	% \noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{know} and \texttt{modal} as reference levels, and random intercepts for participants and items.
+We found highly significant main effects of \texttt{operator} (\texttt{conditional}: $+0.142$, \texttt{negation}: $+0.143$, ; \texttt{question}: $+0.225$, where $p < 2e^{-16}$ in all three cases), as well as many interactions of \texttt{operator} and  \texttt{verb} across the board (where $p < 0.001$ in $38$ cases, $p < 0.01$ in one, and $p < 0.05$ in three out of $57$ possible interactions).\\
+
+	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
+The mean for this baseline (intercept) is $0.867$.
+We found highly significant main effects of \texttt{operator} for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$).
+We also found many interactions of \texttt{operator} and \texttt{verb} across the board (where $p < 0.001$ in $43$ cases, $p < 0.01$ in three, and $p < 0.05$ in one out of $57$ possible interactions).\\
+
+	Notably, the interactions of:
 	\begin{itemize}
-		\item The embedding operator affects projection (main effect of embedding operator)
-		\item This effect differs by verb (interaction effects of operator/verb)
+		\item discover and question, conditional: significant positive interaction effects (can we really argue that for discover: $Q, C > N$, based on this?
+I am not really sure how to do this.
+what I would normally do instead is to look at the posterior samples in a bayesian model to compare each condition and see if they are really different, for $Q, C > N$)
+
+		\item know and question, modal: how can we make argument from model output to differences seen in graph ($Q > C, N > M$)
+
 	\end{itemize}
+
 % paragraph method (end)
@@
 \paragraph{Discussion and Conclusion.} % (fold)
-	\dots
+	\begin{itemize}
+		\item Main effect of operator: choice of operator has an effect on projectivity
+		\item For \emph{be annoyed}, no significant effect is found for \texttt{question} (vs \texttt{negation}), as would be expected based on Karttunen, but we do find unexpected differences between $N > C, M$
+		\item for \emph{discover}, we find $Q, C > N$, an effect in the opposite direction from expectation based on Karttunen
+		\item \emph{know} shows effects that would be incompatible with characterization as factive or semi-factive: $Q > C, N > M$.
+if semi-factive, we would expect $N > C, Q$, if factive, no difference would be expected
+	\end{itemize}
 % paragraph method (end)

### 2022-12-05 materials-lisa/analysis/3_projection-across-operators/data-analysis.Rnw
removed tex gyre termes font

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%% document settings
\documentclass[10pt]{article}
\usepackage{euler}
\usepackage{xunicode,xltxtra}
\usepackage{fontspec}

% fonts
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Biolinum}
	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
	\setmonofont[Scale=MatchLowercase]{Andale Mono}

\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{mathtools}
\usepackage{linguex}
\usepackage{booktabs}

\usepackage[colorlinks=true,citecolor=MidnightBlue,urlcolor=MidnightBlue,linkcolor=MidnightBlue]{hyperref}

\title{Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis}
\author{Hofmann, deMarneffe, Tonhauser}
\date{\today}

\begin{document}
\setkeys{Gin}{width=1.0\textwidth}

<<setup, include=FALSE, echo=FALSE>>=
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = xfun::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo=TRUE, prompt=TRUE, comment=NA, tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/', fig.align = "center", out.width="\\linewidth")
options(tidy=TRUE, linewidth = 100)
@

\maketitle
\setcounter{tocdepth}{2}
\tableofcontents

\section{Introducing the dataset}
<<load-data, echo=FALSE, results='hide'>>=
# set working directory to directory of script
# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
# setwd(this.dir)
setwd("/Users/lisa/Documents/research/GPP-experiments/materials-lisa/analysis/3_projection-across-operators")

# load data
df <- read.csv("../../../results/main/meta-analyses/projectivity/data/data_combined.csv", 
                    header=TRUE, sep=",")

@

<<data-overview>>=
str(df)
df$workerid <- as.factor(df$workerid)
length(levels(df$workerid))
@

\noindent The dataset consists of $57160$ observations from $2682$ participants (recruited on the online platforms Prolific and Amazon Mechanical Turk), across 12 experiments.

We are interested in how highly participants rate speaker commitment to the truth of an embedded complement clause, coded as \texttt{projective} on a real-numbered sliding scale between $0-1$.\\

\noindent The complement clause was embedded under an attitude verb, which in turn was embedded under an entailment-cancelling operator. Our fixed effects factors manipulate the following:

\begin{enumerate}
  \item The choice of attitude verb (coded as \texttt{verb})
  \item The entailment-cancelling operator (coded as \texttt{op})
\end{enumerate}

\noindent The levels for our fixed effects factors are the following:
<<fixed-efcts-lvls>>=
df$verb <- as.factor(df$verb)
levels(df$verb)
length(levels(df$verb))
df$op <- as.factor(df$op)
levels(df$op)
length(levels(df$op))
@

\noindent We are interested in the effect on \texttt{projective} of \texttt{verb} and \texttt{op}, as well as their interaction, corresponding to a $20 \times 4$ factorial design, yielding
<<ncond>>=
length(levels(df$verb))*length(levels(df$op))
@
conditions.\\

\noindent We have $20$ items, corresponding to the content of the complement clause.
<<nitems>>=
df$content <- as.factor(df$content)
levels(df$content)
length(levels(df$content))
@

\noindent We have roughly $36$ observations by item and condition. This is an approximate number, because the \texttt{op} manipulation is a between-studies manipulation, and the number of participants differs by experiment:
<<nobs>>=
# n observations
length(df[,1])

# observations by item
length(df[,1])/length(levels(df$content))
table(df$content)

# observations by verb
length(df[,1])/length(levels(df$verb))
table(df$verb)

# observations by operator
length(df[,1])/length(levels(df$op))
table(df$op)

# observations by item and condition
length(df[,1])/length(levels(df$content))/
  (length(levels(df$verb))*length(levels(df$op)))

@

\newpage
\section{Statistical Summaries and Graphs}
\subsection{Projectivity rating by operator}

Distribution of projectivity ratings by operator with means and $95\%$ bootstrapped confidence intervals.

<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".6\\textwidth">>=
# packages for data manipulation and plots
require(ggplot2)
require(dplyr)
require(forcats)

# load helper functions
source('../../../results/main/helpers.R')

pmeans_op = df %>% group_by(op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         op = fct_reorder(as.factor(op),Mean)) %>% ungroup()

pmeans_op %>% mutate(op = fct_reorder(op, Mean,
                                          .fun = mean)) %>%
  ggplot(aes(x = op, y=Mean)) +
  geom_violin(data=data.frame(op=df$op, Mean=df$projective),scale="width",color="gray80") +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  # labs(title = "Projectivity by operator")+
  theme_bw()

@

\noindent The following generalizations emerge:
\begin{itemize}
  \item Conditionals have the highest projectivity ratings
  \item Projectivity ratings for questions are higher than those for modals and negation, but lower than those for conditionals
  \item Modals and negation have the lowest projectivity ratings
  \item The ratings for negation look a little higher than for modals, but error bars overlap
\end{itemize}

\noindent Although these differences appear to be significant, they are quite small.\\

\newpage
\subsection{Distributions of projectivity rating by verb with means and 95\% bootstrapped confidence interval:}

<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=

pmeans_verb = df %>% group_by(verb) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean))

# plot of projectivity means, 95% CIs and participants' ratings 
pmeans_verb %>% 
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.3,color="black") +
  geom_point(size=0.5,color="black") +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
  xlab("") + ylab("")

@

\begin{itemize}
  \item We see gradual differences in projectivity between verbs
\end{itemize}

\subsection{Distributions, means, and CIs for projectivity ratings by verb and operator:}
<<summary-combined-1>>=
pmeans = df %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

ggplot(pmeans, aes(x=verb, y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width = 0.5, color="black") +
  geom_point(size=0.5,color="black") +
  facet_wrap(~op, labeller = labeller(op = c("m" = "Modal", "n" = "Negation", "q" = "Question", "c" = "Conditional"))) +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ylab("Mean certainty rating") +
  xlab("Predicate")
@

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=

pmeans %>% mutate(verb = fct_reorder(verb, Mean,
                                         .fun = mean)) %>%
  mutate(op = fct_reorder(op, Mean, .fun = mean)) %>%
  ggplot(aes(x=verb, y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 3) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() +
  scale_y_continuous(limits = c(0,1)) +
  ylab("") +
  xlab("") +
  guides(shape = "none") +
  scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) +
  labs(title = "", color = "Operator")

@

\end{document}

### 2022-12-05 materials-lisa/analysis/3_projection-across-operators/data-analysis.tex
removed tex gyre termes font

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%% document settings
\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[dvipsnames]{xcolor}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{euler}
\usepackage{xunicode,xltxtra}
\usepackage{fontspec}

% fonts
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Biolinum}
	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
	\setmonofont[Scale=MatchLowercase]{Andale Mono}

\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{mathtools}
\usepackage{linguex}
\usepackage{booktabs}

\usepackage[colorlinks=true,citecolor=MidnightBlue,urlcolor=MidnightBlue,linkcolor=MidnightBlue]{hyperref}

\title{Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis}
\author{Hofmann, deMarneffe, Tonhauser}
\date{\today}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\setkeys{Gin}{width=1.0\textwidth}

\maketitle
\setcounter{tocdepth}{2}
\tableofcontents

\section{Introducing the dataset}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{str}\hlstd{(df)}
\end{alltt}
\begin{verbatim}
'data.frame':	57160 obs. of  9 variables:
 $ workerid     : int  1 1 1 1 1 1 1 1 1 1 ...
 $ content      : chr  "charley" "danny" "emily" "emma" ...
 $ short_trigger: chr  "acknowledge" "hear" "reveal" "discover" ...
 $ ai_block     : chr  "block1" "block1" "block1" "block1" ...
 $ ai           : num  0.98 0.99 0.99 0.99 0.98 0.98 1 0.99 0.99 0.99 ...
 $ projective   : num  0.3 0.98 0.01 0.99 0.98 0.99 0.01 0.01 0.27 0.01 ...
 $ verb         : chr  "acknowledge" "hear" "reveal" "discover" ...
 $ op           : chr  "q" "q" "q" "q" ...
 $ exp_block    : int  1 1 1 1 1 1 1 1 1 1 ...
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlstd{df}\hlopt{$}\hlstd{workerid} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(df}\hlopt{$}\hlstd{workerid)}
\hlstd{> }\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{workerid))}
\end{alltt}
\begin{verbatim}
[1] 2682
\end{verbatim}
\end{kframe}
\end{knitrout}

\noindent The dataset consists of $57160$ observations from $2682$ participants (recruited on the online platforms Prolific and Amazon Mechanical Turk), across 12 experiments.

We are interested in how highly participants rate speaker commitment to the truth of an embedded complement clause, coded as \texttt{projective} on a real-numbered sliding scale between $0-1$.\\

\noindent The complement clause was embedded under an attitude verb, which in turn was embedded under an entailment-cancelling operator. Our fixed effects factors manipulate the following:

\begin{enumerate}
  \item The choice of attitude verb (coded as \texttt{verb})
  \item The entailment-cancelling operator (coded as \texttt{op})
\end{enumerate}

\noindent The levels for our fixed effects factors are the following:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{df}\hlopt{$}\hlstd{verb} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(df}\hlopt{$}\hlstd{verb)}
\hlstd{> }\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{verb)}
\end{alltt}
\begin{verbatim}
 [1] "acknowledge" "admit"       "announce"    "be_annoyed"  "be_right"   
 [6] "confess"     "confirm"     "demonstrate" "discover"    "establish"  
[11] "hear"        "inform"      "know"        "pretend"     "prove"      
[16] "reveal"      "say"         "see"         "suggest"     "think"      
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{verb))}
\end{alltt}
\begin{verbatim}
[1] 20
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlstd{df}\hlopt{$}\hlstd{op} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(df}\hlopt{$}\hlstd{op)}
\hlstd{> }\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{op)}
\end{alltt}
\begin{verbatim}
[1] "c" "m" "n" "q"
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{op))}
\end{alltt}
\begin{verbatim}
[1] 4
\end{verbatim}
\end{kframe}
\end{knitrout}

\noindent We are interested in the effect on \texttt{projective} of \texttt{verb} and \texttt{op}, as well as their interaction, corresponding to a $20 \times 4$ factorial design, yielding
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{verb))}\hlopt{*}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{op))}
\end{alltt}
\begin{verbatim}
[1] 80
\end{verbatim}
\end{kframe}
\end{knitrout}
conditions.\\

\noindent We have $20$ items, corresponding to the content of the complement clause.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlstd{df}\hlopt{$}\hlstd{content} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(df}\hlopt{$}\hlstd{content)}
\hlstd{> }\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{content)}
\end{alltt}
\begin{verbatim}
 [1] "charley"  "danny"    "emily"    "emma"     "frank"    "grace"   
 [7] "isabella" "jackson"  "jayden"   "jon"      "josh"     "josie"   
[13] "julian"   "mary"     "mia"      "olivia"   "owen"     "sophia"  
[19] "tony"     "zoe"     
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{content))}
\end{alltt}
\begin{verbatim}
[1] 20
\end{verbatim}
\end{kframe}
\end{knitrout}

\noindent We have roughly $36$ observations by item and condition. This is an approximate number, because the \texttt{op} manipulation is a between-studies manipulation, and the number of participants differs by experiment:
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{> }\hlcom{# n observations}
\hlstd{> }\hlkwd{length}\hlstd{(df[,}\hlnum{1}\hlstd{])}
\end{alltt}
\begin{verbatim}
[1] 57160
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlcom{# observations by item}
\hlstd{> }\hlkwd{length}\hlstd{(df[,}\hlnum{1}\hlstd{])}\hlopt{/}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{content))}
\end{alltt}
\begin{verbatim}
[1] 2858
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlkwd{table}\hlstd{(df}\hlopt{$}\hlstd{content)}
\end{alltt}
\begin{verbatim}

 charley    danny    emily     emma    frank    grace isabella  jackson 
    2858     2858     2858     2858     2858     2858     2858     2858 
  jayden      jon     josh    josie   julian     mary      mia   olivia 
    2858     2858     2858     2858     2858     2858     2858     2858 
    owen   sophia     tony      zoe 
    2858     2858     2858     2858 
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlcom{# observations by verb}
\hlstd{> }\hlkwd{length}\hlstd{(df[,}\hlnum{1}\hlstd{])}\hlopt{/}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{verb))}
\end{alltt}
\begin{verbatim}
[1] 2858
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlkwd{table}\hlstd{(df}\hlopt{$}\hlstd{verb)}
\end{alltt}
\begin{verbatim}

acknowledge       admit    announce  be_annoyed    be_right     confess 
       2858        2858        2858        2858        2858        2858 
    confirm demonstrate    discover   establish        hear      inform 
       2858        2858        2858        2858        2858        2858 
       know     pretend       prove      reveal         say         see 
       2858        2858        2858        2858        2858        2858 
    suggest       think 
       2858        2858 
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlcom{# observations by operator}
\hlstd{> }\hlkwd{length}\hlstd{(df[,}\hlnum{1}\hlstd{])}\hlopt{/}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{op))}
\end{alltt}
\begin{verbatim}
[1] 14290
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlkwd{table}\hlstd{(df}\hlopt{$}\hlstd{op)}
\end{alltt}
\begin{verbatim}

    c     m     n     q 
14400 14680 14340 13740 
\end{verbatim}
\begin{alltt}
\hlstd{> }\hlcom{# observations by item and condition}
\hlstd{> }\hlkwd{length}\hlstd{(df[,}\hlnum{1}\hlstd{])}\hlopt{/}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{content))}\hlopt{/}
\hlstd{+ }  \hlstd{(}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{verb))}\hlopt{*}\hlkwd{length}\hlstd{(}\hlkwd{levels}\hlstd{(df}\hlopt{$}\hlstd{op)))}
\end{alltt}
\begin{verbatim}
[1] 35.725
\end{verbatim}
\end{kframe}
\end{knitrout}

\newpage
\section{Statistical Summaries and Graphs}
\subsection{Projectivity rating by operator}

Distribution of projectivity ratings by operator with means and $95\%$ bootstrapped confidence intervals.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=.6\textwidth]{figures/summary-op-1}

}

\end{knitrout}

\noindent The following generalizations emerge:
\begin{itemize}
  \item Conditionals have the highest projectivity ratings
  \item Projectivity ratings for questions are higher than those for modals and negation, but lower than those for conditionals
  \item Modals and negation have the lowest projectivity ratings
  \item The ratings for negation look a little higher than for modals, but error bars overlap
\end{itemize}

\noindent Although these differences appear to be significant, they are quite small.\\

\newpage
\subsection{Distributions of projectivity rating by verb with means and 95\% bootstrapped confidence interval:}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/summary-verb-1}

}

\end{knitrout}

\begin{itemize}
  \item We see gradual differences in projectivity between verbs
\end{itemize}

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/summary-combined-1}

}

\end{knitrout}

\end{document}

### 2022-12-06 writing/investigating-GPP.tex
updated paper

@@
 \end{exe}
+
+\subsection{Factive, semi-factive, etc. references}
+
+{\bf IDEA:} Look for factive and semi-factive (cognitive and emotive) predicates in CommitmentBank, VerbVeridicality, and MegaVeridicality
+
+\begin{itemize}
+
+\item \citealt[65]{karttunen71b}
+
+``Finally, there are at least two distinct types of factive verbs, factives and semi-factives, that differ from each other in contexts that involve certain modal operators, although they are not distinct in simple negative or affirmative assertions.''
+
+\item \citealt[63ff.]{karttunen71b}
+
+{\tt Whatever a sentence with a factive predicate presupposes, the presupposition ought to remain no matter whether the
+main sentence is a negative assertion, an interrogative sentence, or the antecedent of a conditional construction.
+
+For most of the verbs listed in (1) this is true. It is clear that all the examples in (22) presuppose that John had not told the truth.
+
+(22) John didn't regret / realize / discover that he had not told the truth.
+
+The only circumstance in which the examples would not carry along a commitment to the view that John had not told the truth is a case where the assertions in (22) are uttered as an emphatic denial of somebody else's previous assertion.
+That is, one might say something like (23)
+
+(23) John DIDN'T regret that he had not told the truth. How could he have done that when he knew that what he had said was true?
+
+If we leave examples of this sort out of consideration, there are no problems with negative assertions.7 All of the verbs listed in (1) meet this test for factivity.
+However, it seems that, in questions, some of the verbs behave differ- ently.
+Consider the examples in (24).
+
+(24) Did you regret / realize / discover that you had not told the truth?
+
+Most of my informants agree that with regret as the main verb, and possibly also with realize, the question in (24) commits the speaker to the view that the addressee has not told the truth.
+However, (24c) can also be understood as a sincere request for information.
+The speaker is not sure about the truth of the complement and is prepared to accept the address- ee's discovery as a fact.
+In addition to the verb discover, verbs like find out, and see differ from the other factive verbs in this respect.
+In questions, they permit both a factive and non-factive interpretation.
+
+In conditionals, the verbs that lose their factivity certainly also include realize. Consider the examples in
+(25).
+
+(25) If I regret / realize / discover later later that I have not told the truth, I will confess it to everyone.
+
+In uttering (25b) and (25c) one is not making a confession. With realize and discover as predicate, one only admits that
+there is a possibility that one has not told the truth, but one does not concede anything more than that.
+With regret as the main verb, the first clause of (25) contains an admission that the complement is true.
+
+As I suggested earlier in connection with the examples in (13) and (14), it seems that a conditional sentence conversationally implies that, in the view of the speaker, it is at least possible for the antecedent to turn out to be true.
+By this analysis, the examples in (25) conversa- tionally imply the corresponding sentences in (26).
+(26) It is possible that I will regret / realize / discover later that I have not told the truth.
+
+Given (11'), the stronger pair of meaning postulates for regret, it is clear that from (26a) one can infer the truth of the complement.
+Since there is no such necessary relation ship in (26b) and (26c), the corresponding stronger meaning
+postulates for realize and discover would fail. From the fact that it is possible that I may discover something I cannot conclude that this something is in fact the case.
+However, for regret, forget, resent, and all the factives that take sentential subjects this kind of reasoning is valid.
+The meaning postulates for discover, realize, find out, see, notice, and other similar verbs - call them 'semi- factives' - have to be of the weaker form given in (11).
+}
+
+\item \citealt[479f.]{hooper-thompson1973} follow Karttunen 1971.
+
+\item \citealt{djaerv-etal2018}
+
+\begin{itemize}
+
+\item Footnote 2: Here we follow H\&T in our use of `factive' and `semifactive'; these two classes are now commonly referred to as `emotive factives' and `cognitive factives' respectively.
+
+\item p.6: 
+{\tt Class D predicates e.g.
+resent, regret, be sorry, be surprised, be interesting.
+These are the (emotive) factive predicates.
+Given H\&T?s assumptions about the relation between factivity, presupposition, and assertion, the use of these verbs entails that the com- plement of these verbs is presupposed, and cannot be asserted.
+
+Class E predicates e.g.
+realize, learn, discover, know.3 This group constitutes the semi- factives, which presuppose/entail the truth of their complements only in some environments ?
+in particular, this presupposition can be lost in questions and con- ditionals (Kiparsky \& Kiparsky 1970).}
+
+\item Footnote 3: Whether know should be grouped together with the semifactives is highly contentious; here we take no position on this.
+
+\item p.24: EV2 was most acceptable for semifactives and communicative assertives.
+Compared with assertive predicates that explicitly convey a communicative act, epis- temic (or cognitive) assertives yielded much lower ratings for EV2.
+And finally, EV2 was least acceptable under factives.
+
+\end{itemize}
+
+\item Dj\"arv Factivity, Assertion, and Clausal Definiteness
+
+p.170: ``As many readers will no doubt be aware, Karttunen (1971) referred to cognitive factives like discover
+and realize as ?semi-factives?; a term adopted by Hooper & Thompson (1973) and subsequent work on
+MCP, and generally contrasted with ?true? (emotive) factives like resent and regret. These terms suggest
+that factives like discover are therefore somehow ?less factive? than factives like resent. This, however,
+is not what Karttunen?s (1971) distinction refers to.
+Rather, the distinction has to do with the relative sensitivity of the presupposition to entailment-cancelling operators like modals and negation, in contexts
+that establish speaker-ignorance with respect to p (see for instance Heim 1983; Simons 2001).''
+
+\end{itemize}
 

### 2022-12-06 SALT-abstract/SALT-abstract.tex
removed tex gyre termes font
abstract updates 12/6

@@
 %% fonts
-\setmainfont[Scale=MatchLowercase,Mapping=tex-text,SmallCapsFont={TeX Gyre Termes}, SmallCapsFeatures={Letters=SmallCaps}]{Times New Roman}
+\setmainfont[Scale=MatchLowercase,Mapping=tex-text, SmallCapsFeatures={Letters=SmallCaps}]{Times New Roman}
 \setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Times New Roman}
@@
 	\noindent We present experimental evidence that the projectivity of attitude complements varies across different entailment-cancelling operators, and that this effect of entailment-cancelling operator differs between various attitude verb triggers.
-However, our findings about the interaction of operator and trigger do not support \posscite{karttunen_observations_1971} classification of factive vs.
-semi-factive verbs.
+We find that verbs show patterns of between-operator projection behavior which group in interesting ways, but
+	do not support \posscite{karttunen_observations_1971} long-standing classification of factive vs.
+semi-factive verbs, which has been assumed throughout the literature \citep[e.g.][]{djarv_cognitive_2018} + OTHERS.
 
@@
 	\vspace{-.4\baselineskip}
-	Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}.
+	% Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}.
 In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.
+	%
+	\citet{karttunen_observations_1971} suggested that the entailment-cancelling operators in \ref{ex:family} affect projection differentially and proposed a generalization distinguishing \emph{factive} verbs (\emph{regret, forget, resent}) vs.
+\emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}): Factives always project, while semi-factives always project across negation, but not always in polar questions or conditional antecedents.
+Although widely assumed, experimental evaluations of this distinction have been little.
 
-	\citet{karttunen_observations_1971} proposed generalizations to this effect, distinguishing between \emph{factive} verbs (\emph{regret, forget, resent}) and \emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}), suggesting that factives always project, while semi-factives always project across negation, but not always in polar questions or conditionals.
 	To provide a systematic way of distinguishing these classes, \cite{djarv_cognitive_2018} suggest that they correspond to emotive and cognitive predicates, respectively.
-Regarding the effect of operators on projection, \cite{smith_relationship_2014} find an effect of ..
-on ...
-(?) Here, we experimentally address the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments?
-\textbf{(ii)} Do these effects vary for different verbs (and in what way)?
+In a study assessing the acceptability of affirmative responses to a target utterance while explicitly denying a potentially projective inference \citep[based on][]{cummins_experimental_2012}, they find that this kind of affirmative cancellation is more readily available for emotive than cognitive predicates, suggesting that the main clause content and projective content are logically more independent from each other for emotive predicates.
+\cite{smith_relationship_2014}, investigating the effect of operator on the projection of various types of projective content, find an effect on ratings of participant surprisal about the inferences in question, suggesting that inferences triggered by \emph{know} and epithets project more from under negation than conditional antecedents whereas non-restrictive relative clauses show the opposite pattern.
+
+	In our work, we analyze measures from a task designed to measure speaker commitment more directly to address the questions: \textbf{(i)} Is projection affected by differences in entailment-canceling environments?
+\textbf{(ii)} Do these effects vary for different triggers (and in what way)?
 
@@
 \paragraph{Method.} \hspace{-1em}
-	The study consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause.
+	The study, originally designed to address a different research question, consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause.
 Experiments 1--3 used polar question embedding, exps.
@@
 
-\paragraph{Results and Analysis.} \hspace{-1em}
+\vspace{-\baselineskip}
+\paragraph{Results and Discussion.} \hspace{-1em}
 	Pooling the data across our 12 experiments, we examined the effect on certainty-ratings of \texttt{operator}, and \texttt{verb}, as well as their interaction.
@@
 
-	\begin{figure}[h]
-		\vspace{-.8\baselineskip}
-		\centering
-		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
-		\caption{\small Mean certainty ratings by predicate and operator}
-		\label{fig:figure1}
-	\end{figure}
-
-	\vspace{-.3\baselineskip}
+	% \vspace{-.3\baselineskip}
 	% \textbf{version w baseline know/m}
@@
 
-	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
-The mean for this baseline (intercept) is $0.867$.
-We found highly significant main effects of \texttt{operator} for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$).
+	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
+	% The mean for this baseline (intercept) is $0.867$.
+	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter.
+We also found many interactions of \texttt{operator} and \texttt{verb} across the board, suggesting that the effect of embedding context diffes by verb.
+Notably, \emph{\lq discover\rq} is more projective in polar questions and conditional antecedents than under negation, patterning opposite to Karttunen's claims about semi-factives.
+For the emotive predicate \emph{\lq be annoyed\rq} no significant effect is found for \texttt{question} (vs \texttt{negation}), as would be expected based on Karttunen, but we do find unexpected differences between \texttt{negation} $>$ \texttt{conditional, modal}.
+\emph{know} shows effects that would be incompatible with a characterization as either factive or semi-factive: $\texttt{question} > \texttt{conditional, negation} > \texttt{modal}$.
+If \emph{\lq know\rq} is a factive predicate, no difference would be expected.
+If it is semi-factive, we would, again, expect higher projectivity under negation than in questions and conditionals.
+
+	% for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$).
 We also found many interactions of \texttt{operator} and \texttt{verb} across the board (where $p < 0.001$ in $43$ cases, $p < 0.01$ in three, and $p < 0.05$ in one out of $57$ possible interactions).\\
 
-	Notably, the interactions of:
-	\begin{itemize}
-		\item discover and question, conditional: significant positive interaction effects (can we really argue that for discover: $Q, C > N$, based on this?
-I am not really sure how to do this.
-what I would normally do instead is to look at the posterior samples in a bayesian model to compare each condition and see if they are really different, for $Q, C > N$)
+% paragraph results (end)
 
-		\item know and question, modal: how can we make argument from model output to differences seen in graph ($Q > C, N > M$)
+\vspace{-\baselineskip}
+\paragraph{Verb profiles.} % (fold)
+	add discussion about verb classes here
+% paragraph profiles (end)
 
-	\end{itemize}
-
-% paragraph method (end)
-
-\paragraph{Discussion and Conclusion.} % (fold)
-	\begin{itemize}
-		\item Main effect of operator: choice of operator has an effect on projectivity
-		\item For \emph{be annoyed}, no significant effect is found for \texttt{question} (vs \texttt{negation}), as would be expected based on Karttunen, but we do find unexpected differences between $N > C, M$
-		\item for \emph{discover}, we find $Q, C > N$, an effect in the opposite direction from expectation based on Karttunen
-		\item \emph{know} shows effects that would be incompatible with characterization as factive or semi-factive: $Q > C, N > M$.
-if semi-factive, we would expect $N > C, Q$, if factive, no difference would be expected
-	\end{itemize}
-% paragraph method (end)
+\vspace{-\baselineskip}
+\paragraph{Conclusion.} % (fold)
+	add conclusion here
+% paragraph conclusion (end)
 
 \pagebreak
+\begin{figure}[t]
+		\vspace{-.8\baselineskip}
+		\centering
+		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
+		\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals (y-axis), by verb (x-axis), and operator (color/grouping, where: \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions).}
+		\label{fig:figure1}
+	\end{figure}
+
 \bibliography{projective-content.bib}

### 2022-12-06 materials-lisa/analysis/3_projection-across-operators/data-analysis.Rnw
abstract updates 12/6

@@
 
-<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".6\\textwidth">>=
+<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".5\\textwidth">>=
 # packages for data manipulation and plots
@@
   geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
-  geom_line(color = "lightblue") +
-  # labs(title = "Projectivity by operator")+
-  theme_bw()
+  geom_line(color = "lightblue")
+  # labs(title = "Projectivity by operator")
 
@@
 
-<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=
+<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=4.5>>=
 
@@
 
-\subsection{Distributions, means, and CIs for projectivity ratings by verb and operator:}
-<<summary-combined-1>>=
+\newpage
+\subsection{Distributions, means, and CIs for projectivity ratings by verb, for each operator:}
+<<summary-combined-1, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=
 pmeans = df %>% group_by(verb, op) %>%
@@
 
-ggplot(pmeans, aes(x=verb, y=Mean)) +
-  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
+pmeans %>%
+  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean)) +
+  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective, op=df$op),scale="width",color="gray80") +
   geom_errorbar(aes(ymin=YMin,ymax=YMax),width = 0.5, color="black") +
@@
   theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
-  ylab("Mean certainty rating") +
-  xlab("Predicate")
+  ylab("") +
+  xlab("")
 @
 
+\newpage
 \subsection{Means and confidence intervals for projectivity rating by verb and operator:}
-<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
+<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=9, fig.height=4.7>>=
 
-pmeans %>% mutate(verb = fct_reorder(verb, Mean,
-                                         .fun = mean)) %>%
-  mutate(op = fct_reorder(op, Mean, .fun = mean)) %>%
-  ggplot(aes(x=verb, y=Mean, group = op, color = op)) +
+pmeans %>% 
+  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean, group = op, color = op)) +
   coord_cartesian(ylim=c(0,1)) +
   geom_point(aes(shape = op), size = 3) + 
-  scale_shape_manual(values = c("M", "N", "Q", "C")) +
+  scale_shape_manual(values = c("C", "M", "N", "Q")) +
   geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
@@
   xlab("") +
-  guides(shape = "none") +
-  scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
-  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) +
-  labs(title = "", color = "Operator")
+  # guides(shape = "none") +
+  # scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
+  theme_bw() +
+  theme(legend.position = "none") +
+  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1))
 
+@
+\begin{itemize}
+  \item We see interactions between verb and operator
+
+  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.
+
+  \item More projective verbs (\emph{\lq announce\rq} and above) have $\texttt{C} > \texttt{M}$
+
+  \item Highly projective verbs (\emph{\lq hear \rq} and above) have  $\texttt{N} > \texttt{M}$
+
+  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
+  \begin{itemize}
+    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
+    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
+    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive.
+The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
+  \end{itemize}
+
+  \item For more generalizations, let's look at the same information plotted differently: By-operator projectivity for each verb
+\end{itemize}
+
+\newpage
+\subsection{Projectivity ratings by operator, for each verb (Verb profiles)}
+Version 1: With distributions of projectivity ratings as violin-plot
+<<verb-profiles-1, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
+pmeans %>%
+  ggplot(aes(x = fct_reorder(op, Mean), y = Mean)) +
+  coord_cartesian(ylim = c(0,1)) +
+  facet_wrap(vars(fct_reorder(verb, Mean))) +
+  geom_point(size = 1, color = "lightblue") +
+  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective, op=df$op),scale="width",color="gray80") +
+  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") +
+  xlab("") + ylab("") +
+  labs(title = "Mean projectivity by operator, for each verb")
+@
+Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:
+
+\newpage
+\noindent Version 2: With lines to show profiles
+<<verb-profiles-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
+pmeans %>%
+  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb)) +
+  coord_cartesian(ylim = c(0,1)) +
+  facet_wrap(vars(fct_reorder(verb, Mean))) +
+  geom_point(size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") +
+  xlab("") + ylab("") +
+  theme_bw() +
+  labs(title = "Mean projectivity by operator, for each verb")
+@
+
+\begin{itemize}
+  \item \emph{pretend, think}: \textbf{anti-veridical profile}\\
+    \texttt{N} $>$ \texttt{M, Q, C}, overall low projectivity
+
+  \item \emph{acknowledge, hear, inform, see, discover, know, be annoyed}: \textbf{\lq factive\rq\ profile}\\
+    \texttt{Q} $>$ \texttt{N, C} \textcolor{gray}{$>_?$} \texttt{M}, overall high projectivity (it may be possible to find further subgroups here)
+
+  \item \emph{prove, confirm, establish, demonstrate, (announce), confess, admit, reveal}: \textbf{veridical profile}\\
+    \texttt{M, C} $>$ \texttt{Q} $>$ \texttt{N}, overall med-lo to med-hi projectivity
+
+  \item \emph{be right, suggest}: \textbf{reportative profile}\\
+    \texttt{M, N, C} $>$ \texttt{Q}
+
+\end{itemize}
+
+Maybe these can have better names, not trying to suggest that verbs can neatly divided in factive v non-factive, but potentially this class / profile is what prompted intuitions in previous literature, and naming in this tradition could make sense, but can be changed depending on our rhetoric, of course.
+
+\newpage
+
+<<verb-profiles-3, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
+vgroups <- data.frame(verb = levels(pmeans$verb), profile = 
+                       c("0+0",  # acknowledge
+                         "00+",   # admit 
+                         "00+",  # announce
+                         "-+-",  # be annoyed
+                         "+-+",    # be right
+                         "00+",   # confess
+                         "+-+",   # confirm
+                         "+0+",   # demonstrate
+                         "0+0",  # discover
+                         "+-+",   # establish
+                         "-+0",  # hear
+                         "-+-",  # inform
+                         "-+-",  # know
+                         "-00",    # pretend
+                         "+-+",   # prove
+                         "0++",   # reveal
+                         "0-+",    # say
+                         "-+-",  # see
+                         "0-+",   # suggest
+                         "-00"     # think
+                         ))
+
+lookup_group <- function(verb_name){
+  return(as.character(filter(vgroups, verb == verb_name)[2]))
+}
+
+pmeans = pmeans %>% rowwise() %>% mutate(groups = lookup_group(verb))
+
+pmeans %>%
+  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb, color = verb)) +
+  coord_cartesian(ylim = c(0,1)) +
+  facet_wrap(~groups) +
+  geom_point(size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1) +
+  geom_line() +
+  xlab("") + ylab("") +
+  theme_bw() +
+  labs(title = "Mean projectivity by operator, for each verb")
 @

### 2022-12-06 materials-lisa/analysis/3_projection-across-operators/data-analysis.tex
abstract updates 12/6

@@
 
-{\centering \includegraphics[width=.6\textwidth]{figures/summary-op-1}
+{\centering \includegraphics[width=.5\textwidth]{figures/summary-op-1}
 
@@
 
+\newpage
+\subsection{Distributions, means, and CIs for projectivity ratings by verb, for each operator:}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\linewidth]{figures/summary-combined-1-1}
+
+}
+
+\end{knitrout}
+
+\newpage
 \subsection{Means and confidence intervals for projectivity rating by verb and operator:}
@@
 
-{\centering \includegraphics[width=\linewidth]{figures/summary-combined-1}
+{\centering \includegraphics[width=\linewidth]{figures/summary-combined-2-1}
+
+}
+
+\end{knitrout}
+\begin{itemize}
+  \item We see interactions between verb and operator
+
+  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.
+
+  \item More projective verbs (\emph{\lq announce\rq} and above) have $\texttt{C} > \texttt{M}$
+
+  \item Highly projective verbs (\emph{\lq hear \rq} and above) have  $\texttt{N} > \texttt{M}$
+
+  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
+  \begin{itemize}
+    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
+    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
+    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive.
+The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
+  \end{itemize}
+
+  \item For more generalizations, let's look at the same information plotted differently: By-operator projectivity for each verb
+\end{itemize}
+
+\newpage
+\subsection{Projectivity ratings by operator, for each verb (Verb profiles)}
+Version 1: With distributions of projectivity ratings as violin-plot
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\linewidth]{figures/verb-profiles-1-1}
+
+}
+
+\end{knitrout}
+Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:
+
+\newpage
+\noindent Version 2: With lines to show profiles
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\linewidth]{figures/verb-profiles-2-1}
+
+}
+
+\end{knitrout}
+
+\begin{itemize}
+  \item \emph{pretend, think}: \textbf{anti-veridical profile}\\
+    \texttt{N} $>$ \texttt{M, Q, C}, overall low projectivity
+
+  \item \emph{acknowledge, hear, inform, see, discover, know, be annoyed}: \textbf{\lq factive\rq\ profile}\\
+    \texttt{Q} $>$ \texttt{N, C} \textcolor{gray}{$>_?$} \texttt{M}, overall high projectivity (it may be possible to find further subgroups here)
+
+  \item \emph{prove, confirm, establish, demonstrate, (announce), confess, admit, reveal}: \textbf{veridical profile}\\
+    \texttt{M, C} $>$ \texttt{Q} $>$ \texttt{N}, overall med-lo to med-hi projectivity
+
+  \item \emph{be right, suggest}: \textbf{reportative profile}\\
+    \texttt{M, N, C} $>$ \texttt{Q}
+
+\end{itemize}
+
+Maybe these can have better names, not trying to suggest that verbs can neatly divided in factive v non-factive, but potentially this class / profile is what prompted intuitions in previous literature, and naming in this tradition could make sense, but can be changed depending on our rhetoric, of course.
+
+\newpage
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\linewidth]{figures/verb-profiles-3-1}
 

### 2022-12-08 SALT-abstract/SALT-abstract.tex
updated SALT abstract

@@
 	\textbf{\large%\thetitle
-		A diverse family (of sentences)\\ Projectivity differs across embedding operators---but not like you think}
+		%A diverse family (of sentences)\\ 
+		Projection differs across embedding operators---but not like you think}
 \end{center}
 
-\vspace{-.4\baselineskip}
-	\noindent We present experimental evidence that the projectivity of attitude complements varies across different entailment-cancelling operators, and that this effect of entailment-cancelling operator differs between various attitude verb triggers.
-We find that verbs show patterns of between-operator projection behavior which group in interesting ways, but
-	do not support \posscite{karttunen_observations_1971} long-standing classification of factive vs.
-semi-factive verbs, which has been assumed throughout the literature \citep[e.g.][]{djarv_cognitive_2018} + OTHERS.
+\vspace*{-.4cm}
 
-\vspace{-\baselineskip}
-\paragraph{Projection across entailment-cancelling operators.}  \hspace{-1em}
-	Certain attitude ascriptions come with an inference to the truth of their complement, even if embedded under entailment-cancelling operators (shown for \emph{\lq discover\rq} in \ref{ex:family}), in which case the inference is said to \emph{project} (e.g.
+\noindent 
+We present experimental evidence that i) the projection of the content of the clausal complement of attitude predicates varies across entailment-canceling operators (negation, question, modal, conditional) and ii) this by-operator variation differs across attitude predicates.
+Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealt{karttunen_observations_1971,djarv_cognitive_2018}).
+Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.
+
+\noindent
+{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters one of the attitude ascriptions in  \ref{ex:family} is committed to the truth of the content of the complement (CC), that Julian dances salsa, even though the complement occurs under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, a modal \ref{ex:mod}, or a conditional \ref{ex:cond}.
+
+%	Certain attitude ascriptions come with an inference to the truth of their complement, even if embedded under entailment-cancelling operators (shown for \emph{discover} in \ref{ex:family}), in which case the inference is said to \emph{project} (e.g.
 \citealp{karttunen_observations_1971}).
@@
 	\ex. \label{ex:family}
-		\a. \label{ex:mod}
-			Modals: \hfill
-			\emph{\lq Perhaps Cole discovered that Julian dances Salsa.\rq}
-		\b. \label{ex:neg}
+		\a. \label{ex:neg}
 			Negation: \hfill
-			\emph{\lq Cole didn't discover that Julian dances Salsa.\rq}
+			\emph{\lq Cole didn't discover that Julian dances salsa.\rq}
 		\b. \label{ex:q}
-			Polar Questions: \hfill
-			\emph{\lq Did Cole discover that Julian dances Salsa?\rq}
-		\b. \label{ex:cond}
-			Conditionals: \hfill
-			\emph{\lq If Cole discovered that Julian dances Salsa, Logan will be joyful.\rq}
+			Polar Question: \hfill
+			\emph{\lq Did Cole discover that Julian dances salsa?\rq}
+		\c. \label{ex:mod}
+			Modal: \hfill
+			\emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
+		\d. \label{ex:cond}
+			Conditional: \hfill
+			\emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
 		\z.
@@
 	%
-	\citet{karttunen_observations_1971} suggested that the entailment-cancelling operators in \ref{ex:family} affect projection differentially and proposed a generalization distinguishing \emph{factive} verbs (\emph{regret, forget, resent}) vs.
-\emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}): Factives always project, while semi-factives always project across negation, but not always in polar questions or conditional antecedents.
-Although widely assumed, experimental evaluations of this distinction have been little.
+	\citealt{karttunen_observations_1971} suggested that the CC of factive predicates (e.g., \emph{regret, forget}) projects from under all four operators, whereas that of semi-factive predicates (e.g., \emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
 
-	To provide a systematic way of distinguishing these classes, \cite{djarv_cognitive_2018} suggest that they correspond to emotive and cognitive predicates, respectively.
-In a study assessing the acceptability of affirmative responses to a target utterance while explicitly denying a potentially projective inference \citep[based on][]{cummins_experimental_2012}, they find that this kind of affirmative cancellation is more readily available for emotive than cognitive predicates, suggesting that the main clause content and projective content are logically more independent from each other for emotive predicates.
-\cite{smith_relationship_2014}, investigating the effect of operator on the projection of various types of projective content, find an effect on ratings of participant surprisal about the inferences in question, suggesting that inferences triggered by \emph{know} and epithets project more from under negation than conditional antecedents whereas non-restrictive relative clauses show the opposite pattern.
+There has been, to date, one investigation of by-operator variation: \citealt{smith_relationship_2014}, who investigated by-operator variation (negation, conditional) for the projective content of six expressions (\emph{know, the, win}, epithets, clefs, non-restrictive relative clauses (NRRCs)), observed by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefs) and that this by-operator variation differs across contents (e.g., the content of NRRCs was more projective under conditionals than negation, the opposite pattern was observed for \emph{win}).
+It is not clear, however, whether the response task used by \citealt{smith_relationship_2014} measured projection, as participants were asked to rate how surprised they would be to learn the content under investigation after observing the utterance.
 
-	In our work, we analyze measures from a task designed to measure speaker commitment more directly to address the questions: \textbf{(i)} Is projection affected by differences in entailment-canceling environments?
+There has not yet been an experimental investigation of by-operator variation between factive and semi-factive predicates.
+\citealt{djarv_cognitive_2018} and \citealt{tonhauser_how_2018} did, however, observe by-predicate projection variation under polar questions.
+\citealt{djarv_cognitive_2018} observed a difference between \emph{be happy} and \emph{appreciate} (which they assumed are factive predicates) and \emph{be aware} and \emph{realize} (which they assumed are semi-factive predicates).
+However, the response task (acceptability of an affirmation of the CC while the main clause content was denied) did not measure projection of the CC.
+\citealt{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates from under polar questions: The by-predicate projection differences they observed did not align with what would be expected from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more projective than that of semi-factive \emph{discover}).
+
+\noindent
+{\bf Experiment.} We present the results of an experiment designed to investigate by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
+Projection was measured for the same contents across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see also, e.g., \citealt{tonhauser_how_2018, djaerv-bacovcin-salt27, mahler2020}).
+
+%In our work, we analyze measures from a task designed to measure speaker commitment more directly to address the questions: \textbf{(i)} Is projection affected by differences in entailment-canceling environments?
 \textbf{(ii)} Do these effects vary for different triggers (and in what way)?
-
-\vspace{-\baselineskip}
-\paragraph{Experimentally investigating projection,} \hspace{-1em}
-	we used a response task to elicit judgments about how strongly a speaker would be committed towards the embedded clause \citep[from][]{tonhauser_prosodic_2016}.
+We used a response task to elicit judgments about how strongly a speaker would be committed towards the embedded clause \citep[from][]{tonhauser_prosodic_2016}.
 We presented sentences like in \ref{ex:family} as asserted by a named speaker (e.g.
 “\textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}”).
-Participants then provided a certainty-rating in response to a prompt like: \emph{“Is Daniel certain that Julian dances Salsa?”}, by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
+Participants then provided a certainty-rating in response to a prompt like: \emph{“Is Daniel certain that Julian dances salsa?”}, by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
 
 % paragraph experiment (end)
+\noindent
+{\bf Methods and expectations.} Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The attitude predicates were embedded under polar questions in Exps.~1, under negation in Exps.~2, under {\em perhaps} in Exps.~3, and in the antecedent of a conditional in Exps.~4.
+(Each set of experiments consisted of three experiments that differed in the at-issueness measure used in a separate block.
+We focus on the projection ratings here.) In each experiment, participants were asked to read utterances like those in \ref{ex:family} and judge whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
+Participants gave their response on a slider marked `no' (coded as 0) at one and and `yes' (coded as 1) on the other end.
+Each participant rated the projection of the CC of all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
+We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
+We expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and possibly lower ratings under the other operators.
 
-\vspace{-\baselineskip}
-\paragraph{Design and Expectations.} \hspace{-1em}
-	We compared certainty-ratings for the four entailment-canceling operators in \ref{ex:family}, and 20 clause-embedding predicates (\texttt{verb}: {\em be annoyed, discover, know, reveal, see, acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, pretend, suggest, say, think, be right, demonstrate}).
-	Based on the Karttunen-Djärv generalization about emotive factives vs.
+%We compared certainty-ratings for the four entailment-canceling operators in \ref{ex:family}, and 20 clause-embedding predicates (\texttt{verb}: {\em be annoyed, discover, know, reveal, see, acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, pretend, suggest, say, think, be right, demonstrate}).
+%	Based on the Karttunen-Djärv generalization about emotive factives vs.
 cognitive semi-factives, we would expect emotive factives (\emph{be annoyed}), and verbs normally taken to be factive (\emph{know}), to be highly projective in a way that is indifferent to the embedding context, and cognitive semi-factives (\emph{discover, see}) to show higher projectivity ratings under negation compared to questions and conditionals.
 
-% paragraph design_and_expectations_ (end)
-
-\vspace{-\baselineskip}
-\paragraph{Method.} \hspace{-1em}
-	The study, originally designed to address a different research question, consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause.
+%The study, originally designed to address a different research question, consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause.
 Experiments 1--3 used polar question embedding, exps.
 4--6 used negation, 7--9 modals, and 10--12 conditionals, making the \texttt{operator} manipulation a between-subjects factor.
-We analyzed data from 2682 self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk).
+
+%We analyzed data from 2682 self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk).
 Participants saw items Latin-squared and randomized with six control stimuli.
@@
 
-\vspace{-\baselineskip}
-\paragraph{Results and Discussion.} \hspace{-1em}
-	Pooling the data across our 12 experiments, we examined the effect on certainty-ratings of \texttt{operator}, and \texttt{verb}, as well as their interaction.
-Mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals are shown in  \textbf{Figure \ref{fig:figure1}}.
+\noindent
+{\bf Results.} Fig.~\ref{fig:figure1} plots the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
+We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under questions and negation than from under conditionals and modals, the CC of \emph{know} projects more from under questions than from under conditionals and negation, and the CC of \emph{discover} projects more from under questions and conditionals than from under negation and modals.
+These results are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
+For {\em know}, we observe a third pattern -- one that is not aligned with a classification as a factive nor as a semi-factive predicate.
+
+==== THIS IS AS FAR AS I've COME ====
 
@@
 
-\vspace{-\baselineskip}
-\paragraph{Verb profiles.} % (fold)
+\noindent
+{\bf Discussion.}
+
+We observe that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}; replicating the results of  \citealt{degen-tonhauser-language}).
+
+No categorical distinction between factive and non-factive predicates either
+
 add discussion about verb classes here
-% paragraph profiles (end)
 
-\vspace{-\baselineskip}
-\paragraph{Conclusion.} % (fold)
-	add conclusion here
-% paragraph conclusion (end)
+\newpage
 
-\pagebreak
 \begin{figure}[t]
@@
 		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
-		\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals (y-axis), by verb (x-axis), and operator (color/grouping, where: \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions).}
+		\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
+Embedding operator coded by letter and color:  \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions.}
 		\label{fig:figure1}
 	\end{figure}
+
+\newpage
 

### 2022-12-08 SALT-abstract/SALT-abstract.tex
added model summaries, fixed references + formatting issues

@@
 
-\documentclass[12pt, a4paper]{scrartcl}
+\documentclass[12pt, a4paper]{article}
 
@@
 We present experimental evidence that i) the projection of the content of the clausal complement of attitude predicates varies across entailment-canceling operators (negation, question, modal, conditional) and ii) this by-operator variation differs across attitude predicates.
-Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealt{karttunen_observations_1971,djarv_cognitive_2018}).
+Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}, though see \citealp{beaver_have_2010}).
 Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.
@@
 	%
-	\citealt{karttunen_observations_1971} suggested that the CC of factive predicates (e.g., \emph{regret, forget}) projects from under all four operators, whereas that of semi-factive predicates (e.g., \emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
+	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g.
+\emph{regret, forget}) projects from under all four operators, whereas that of semi-factive predicates (e.g.
+\emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
 
@@
 	{\bf Experiment.} We present the results of an experiment designed to investigate by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
-Projection was measured for the same contents across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see also, e.g., \citealt{tonhauser_how_2018, djaerv-bacovcin-salt27, mahler2020}).
+Projection was measured for the same contents across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see also, e.g., \citealt{tonhauser_how_2018,djarv_prosodic_2017, mahler2020}).
 
@@
 \noindent
-{\bf Methods and expectations.} Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The attitude predicates were embedded under polar questions in Exps.~1, under negation in Exps.~2, under {\em perhaps} in Exps.~3, and in the antecedent of a conditional in Exps.~4.
+{\bf Methods and expectations.}
+	Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The attitude predicates were embedded under polar questions in Exps.~1, under negation in Exps.~2, under {\em perhaps} in Exps.~3, and in the antecedent of a conditional in Exps.~4.
 (Each set of experiments consisted of three experiments that differed in the at-issueness measure used in a separate block.
@@
 
-We observe that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}; replicating the results of  \citealt{degen-tonhauser-language}).
+We observe that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}; replicating the results of \citealp{degen_are_2022}).
 
@@
 
+\begin{table}[ht]
+	\caption{Model output w \texttt{be\_annoyed/n} as baseline}
+	\label{tab:annoyed-n}
+	\centering
+
+	\begin{tabular}{lrrrrr}
+		\toprule
+		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
+		\midrule
+		(Intercept) & 0.8669958 & 0.0114402 & 1479.204 & 75.7850523 & 0.0000000\\
+		opc & -0.1155850 & 0.0156641 & 26227.368 & -7.3789836 & 0.0000000\\
+		opm & -0.1564480 & 0.0155858 & 26288.139 & -10.0378702 & 0.0000000\\
+		opq & 0.0249138 & 0.0143090 & 56826.915 & 1.7411232 & 0.0816674\\
+		verbsuggest & -0.6162683 & 0.0134941 & 54372.895 & -45.6694112 & 0.0000000\\
+		\addlinespace
+		verbacknowledge & -0.2204417 & 0.0134948 & 54373.777 & -16.3352724 & 0.0000000\\
+		verbadmit & -0.3402430 & 0.0134933 & 54371.874 & -25.2156750 & 0.0000000\\
+		verbannounce & -0.3607504 & 0.0134931 & 54371.547 & -26.7359958 & 0.0000000\\
+		verbbe\_right & -0.6638924 & 0.0134936 & 54372.293 & -49.2003974 & 0.0000000\\
+		verbconfess & -0.3681695 & 0.0134949 & 54373.904 & -27.2820517 & 0.0000000\\
+		\addlinespace
+		verbconfirm & -0.6338363 & 0.0134931 & 54371.657 & -46.9746939 & 0.0000000\\
+		verbdemonstrate & -0.5307026 & 0.0134935 & 54372.157 & -39.3301466 & 0.0000000\\
+		verbdiscover & -0.1864716 & 0.0134942 & 54372.992 & -13.8186520 & 0.0000000\\
+		verbestablish & -0.5542561 & 0.0134933 & 54371.853 & -41.0764079 & 0.0000000\\
+		verbhear & -0.1956794 & 0.0134938 & 54372.473 & -14.5014554 & 0.0000000\\
+		\addlinespace
+		verbinform & -0.1374428 & 0.0134945 & 54373.345 & -10.1851106 & 0.0000000\\
+		verbknow & -0.0748953 & 0.0134936 & 54372.288 & -5.5504197 & 0.0000000\\
+		verbpretend & -0.5049533 & 0.0134944 & 54373.218 & -37.4195409 & 0.0000000\\
+		verbprove & -0.6364950 & 0.0134935 & 54372.138 & -47.1704273 & 0.0000000\\
+		verbreveal & -0.3057814 & 0.0134937 & 54372.418 & -22.6609884 & 0.0000000\\
+		\addlinespace
+		verbsay & -0.5419725 & 0.0134925 & 54370.770 & -40.1685159 & 0.0000000\\
+		verbsee & -0.1827512 & 0.0134932 & 54371.791 & -13.5438968 & 0.0000000\\
+		verbthink & -0.5101198 & 0.0134930 & 54371.491 & -37.8062103 & 0.0000000\\
+		opc:verbsuggest & 0.1594325 & 0.0190634 & 54372.626 & 8.3632939 & 0.0000000\\
+		opm:verbsuggest & 0.1766113 & 0.0189718 & 54372.046 & 9.3091498 & 0.0000000\\
+		\addlinespace
+		opq:verbsuggest & -0.0737648 & 0.0192906 & 54372.717 & -3.8238781 & 0.0001315\\
+		opc:verbacknowledge & 0.1945569 & 0.0190630 & 54372.333 & 10.2059764 & 0.0000000\\
+		opm:verbacknowledge & 0.1100553 & 0.0189729 & 54373.070 & 5.8006461 & 0.0000000\\
+		opq:verbacknowledge & 0.0672935 & 0.0192909 & 54372.991 & 3.4883599 & 0.0004864\\
+		opc:verbadmit & 0.2803349 & 0.0190626 & 54371.933 & 14.7060204 & 0.0000000\\
+		\addlinespace
+		opm:verbadmit & 0.1810708 & 0.0189717 & 54371.994 & 9.5442376 & 0.0000000\\
+		opq:verbadmit & 0.0440050 & 0.0192896 & 54371.832 & 2.2812856 & 0.0225354\\
+		opc:verbannounce & 0.2503929 & 0.0190639 & 54373.132 & 13.1343759 & 0.0000000\\
+		opm:verbannounce & 0.1633026 & 0.0189721 & 54372.338 & 8.6075049 & 0.0000000\\
+		opq:verbannounce & 0.0017532 & 0.0192899 & 54372.121 & 0.0908892 & 0.9275810\\
+		\addlinespace
+		opc:verbbe\_right & 0.1424317 & 0.0190623 & 54371.698 & 7.4718895 & 0.0000000\\
+		opm:verbbe\_right & 0.2582399 & 0.0189717 & 54371.967 & 13.6118412 & 0.0000000\\
+		opq:verbbe\_right & -0.0776846 & 0.0192897 & 54371.974 & -4.0272512 & 0.0000565\\
+		opc:verbconfess & 0.3111230 & 0.0190643 & 54373.455 & 16.3196605 & 0.0000000\\
+		opm:verbconfess & 0.1837114 & 0.0189741 & 54374.125 & 9.6821973 & 0.0000000\\
+		\addlinespace
+		opq:verbconfess & 0.0393640 & 0.0192890 & 54371.320 & 2.0407459 & 0.0412809\\
+		opc:verbconfirm & 0.2963833 & 0.0190629 & 54372.237 & 15.5476205 & 0.0000000\\
+		opm:verbconfirm & 0.3663333 & 0.0189721 & 54372.324 & 19.3090505 & 0.0000000\\
+		opq:verbconfirm & 0.0312870 & 0.0192902 & 54372.377 & 1.6219108 & 0.1048282\\
+		opc:verbdemonstrate & 0.2854366 & 0.0190647 & 54373.829 & 14.9719656 & 0.0000000\\
+		\addlinespace
+		opm:verbdemonstrate & 0.2783816 & 0.0189716 & 54371.830 & 14.6736292 & 0.0000000\\
+		opq:verbdemonstrate & 0.0753554 & 0.0192892 & 54371.520 & 3.9066023 & 0.0000937\\
+		opc:verbdiscover & 0.2270099 & 0.0190632 & 54372.450 & 11.9082962 & 0.0000000\\
+		opm:verbdiscover & 0.1002424 & 0.0189726 & 54372.795 & 5.2835252 & 0.0000001\\
+		opq:verbdiscover & 0.0762282 & 0.0192898 & 54372.014 & 3.9517423 & 0.0000777\\
+		\addlinespace
+		opc:verbestablish & 0.2570817 & 0.0190633 & 54372.539 & 13.4857080 & 0.0000000\\
+		opm:verbestablish & 0.2879581 & 0.0189714 & 54371.677 & 15.1785426 & 0.0000000\\
+		opq:verbestablish & 0.0061604 & 0.0192893 & 54371.583 & 0.3193687 & 0.7494481\\
+		opc:verbhear & 0.2016309 & 0.0190630 & 54372.273 & 10.5770983 & 0.0000000\\
+		opm:verbhear & -0.0046313 & 0.0189714 & 54371.681 & -0.2441216 & 0.8071376\\
+		\addlinespace
+		opq:verbhear & 0.0756021 & 0.0192902 & 54372.373 & 3.9192008 & 0.0000890\\
+		opc:verbinform & 0.1278740 & 0.0190645 & 54373.631 & 6.7074362 & 0.0000000\\
+		opm:verbinform & 0.0141486 & 0.0189741 & 54374.122 & 0.7456761 & 0.4558664\\
+		opq:verbinform & 0.0587564 & 0.0192913 & 54373.322 & 3.0457537 & 0.0023221\\
+		opc:verbknow & 0.1142699 & 0.0190629 & 54372.206 & 5.9943629 & 0.0000000\\
+		\addlinespace
+		opm:verbknow & 0.0130032 & 0.0189709 & 54371.181 & 0.6854301 & 0.4930755\\
+		opq:verbknow & 0.0569598 & 0.0192897 & 54371.975 & 2.9528579 & 0.0031498\\
+		opc:verbpretend & -0.0902047 & 0.0190640 & 54373.155 & -4.7316842 & 0.0000022\\
+		opm:verbpretend & -0.0194245 & 0.0189738 & 54373.822 & -1.0237558 & 0.3059552\\
+		opq:verbpretend & -0.2196823 & 0.0192900 & 54372.250 & -11.3883795 & 0.0000000\\
+		\addlinespace
+		opc:verbprove & 0.2579585 & 0.0190627 & 54372.031 & 13.5320999 & 0.0000000\\
+		opm:verbprove & 0.3580525 & 0.0189720 & 54372.223 & 18.8726898 & 0.0000000\\
+		opq:verbprove & 0.0230805 & 0.0192908 & 54372.899 & 1.1964509 & 0.2315259\\
+		opc:verbreveal & 0.2902650 & 0.0190644 & 54373.511 & 15.2255216 & 0.0000000\\
+		opm:verbreveal & 0.1767476 & 0.0189723 & 54372.513 & 9.3160804 & 0.0000000\\
+		\addlinespace
+		opq:verbreveal & 0.0665887 & 0.0192906 & 54372.717 & 3.4518763 & 0.0005571\\
+		opc:verbsay & 0.1985351 & 0.0190618 & 54371.193 & 10.4153412 & 0.0000000\\
+		opm:verbsay & 0.1451974 & 0.0189723 & 54372.495 & 7.6531273 & 0.0000000\\
+		opq:verbsay & -0.1815738 & 0.0192889 & 54371.195 & -9.4133926 & 0.0000000\\
+		opc:verbsee & 0.2175191 & 0.0190629 & 54372.234 & 11.4105822 & 0.0000000\\
+		\addlinespace
+		opm:verbsee & 0.0594092 & 0.0189723 & 54372.499 & 3.1313655 & 0.0017409\\
+		opq:verbsee & 0.1108274 & 0.0192896 & 54371.840 & 5.7454526 & 0.0000000\\
+		opc:verbthink & 0.0023197 & 0.0190619 & 54371.272 & 0.1216913 & 0.9031440\\
+		opm:verbthink & 0.0145577 & 0.0189717 & 54371.965 & 0.7673367 & 0.4428847\\
+		opq:verbthink & -0.1738342 & 0.0192903 & 54372.438 & -9.0115053 & 0.0000000\\
+		\bottomrule
+	\end{tabular}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+	\caption{Model output w \texttt{discover/n} as baseline}
+	\label{tab:discover-n}
+	\centering
+
+	\begin{tabular}{lrrrrr}
+		\toprule
+		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
+		\midrule
+		(Intercept) & 0.6805242 & 0.0114403 & 1479.217 & 59.4849565 & 0.0000000\\
+		opc & 0.1114249 & 0.0156639 & 26226.686 & 7.1134644 & 0.0000000\\
+		opm & -0.0562056 & 0.0155855 & 26286.826 & -3.6062779 & 0.0003112\\
+		opq & 0.1011420 & 0.0143103 & 56828.295 & 7.0677830 & 0.0000000\\
+		verbbe\_annoyed & 0.1864716 & 0.0134942 & 54372.998 & 13.8186521 & 0.0000000\\
+		\addlinespace
+		verbsuggest & -0.4297967 & 0.0134937 & 54372.361 & -31.8516905 & 0.0000000\\
+		verbacknowledge & -0.0339702 & 0.0134936 & 54372.308 & -2.5174950 & 0.0118221\\
+		verbadmit & -0.1537714 & 0.0134942 & 54372.985 & -11.3953857 & 0.0000000\\
+		verbannounce & -0.1742788 & 0.0134947 & 54373.649 & -12.9145926 & 0.0000000\\
+		verbbe\_right & -0.4774208 & 0.0134935 & 54372.079 & -35.3816304 & 0.0000000\\
+		\addlinespace
+		verbconfess & -0.1816980 & 0.0134936 & 54372.192 & -13.4655376 & 0.0000000\\
+		verbconfirm & -0.4473648 & 0.0134944 & 54373.311 & -33.1517710 & 0.0000000\\
+		verbdemonstrate & -0.3442310 & 0.0134928 & 54371.247 & -25.5121517 & 0.0000000\\
+		verbestablish & -0.3677846 & 0.0134928 & 54371.167 & -27.2579106 & 0.0000000\\
+		verbhear & -0.0092079 & 0.0134940 & 54372.798 & -0.6823676 & 0.4950095\\
+		\addlinespace
+		verbinform & 0.0490288 & 0.0134942 & 54373.066 & 3.6333114 & 0.0002801\\
+		verbknow & 0.1115762 & 0.0134931 & 54371.668 & 8.2691029 & 0.0000000\\
+		verbpretend & -0.3184818 & 0.0134943 & 54373.176 & -23.6011440 & 0.0000000\\
+		verbprove & -0.4500234 & 0.0134934 & 54372.042 & -33.3512833 & 0.0000000\\
+		verbreveal & -0.1193099 & 0.0134941 & 54372.880 & -8.8416323 & 0.0000000\\
+		\addlinespace
+		verbsay & -0.3555009 & 0.0134943 & 54373.168 & -26.3444678 & 0.0000000\\
+		verbsee & 0.0037204 & 0.0134930 & 54371.450 & 0.2757288 & 0.7827574\\
+		verbthink & -0.3236483 & 0.0134936 & 54372.197 & -23.9853913 & 0.0000000\\
+		opc:verbbe\_annoyed & -0.2270099 & 0.0190632 & 54372.453 & -11.9082963 & 0.0000000\\
+		opm:verbbe\_annoyed & -0.1002424 & 0.0189726 & 54372.798 & -5.2835253 & 0.0000001\\
+		\addlinespace
+		opq:verbbe\_annoyed & -0.0762282 & 0.0192898 & 54372.017 & -3.9517423 & 0.0000777\\
+		opc:verbsuggest & -0.0675773 & 0.0190625 & 54371.851 & -3.5450396 & 0.0003929\\
+		opm:verbsuggest & 0.0763690 & 0.0189709 & 54371.211 & 4.0255864 & 0.0000569\\
+		opq:verbsuggest & -0.1499930 & 0.0192910 & 54373.069 & -7.7752981 & 0.0000000\\
+		opc:verbacknowledge & -0.0324529 & 0.0190617 & 54371.116 & -1.7025196 & 0.0886637\\
+		\addlinespace
+		opm:verbacknowledge & 0.0098129 & 0.0189711 & 54371.417 & 0.5172570 & 0.6049789\\
+		opq:verbacknowledge & -0.0089347 & 0.0192902 & 54372.396 & -0.4631724 & 0.6432426\\
+		opc:verbadmit & 0.0533251 & 0.0190632 & 54372.508 & 2.7972739 & 0.0051554\\
+		opm:verbadmit & 0.0808284 & 0.0189728 & 54372.916 & 4.2602367 & 0.0000205\\
+		opq:verbadmit & -0.0322232 & 0.0192908 & 54372.883 & -1.6703960 & 0.0948468\\
+		\addlinespace
+		opc:verbannounce & 0.0233831 & 0.0190642 & 54373.367 & 1.2265427 & 0.2199998\\
+		opm:verbannounce & 0.0630602 & 0.0189726 & 54372.754 & 3.3237576 & 0.0008887\\
+		opq:verbannounce & -0.0744750 & 0.0192926 & 54374.465 & -3.8602856 & 0.0001134\\
+		opc:verbbe\_right & -0.0845782 & 0.0190620 & 54371.360 & -4.4370099 & 0.0000091\\
+		opm:verbbe\_right & 0.1579976 & 0.0189719 & 54372.163 & 8.3279685 & 0.0000000\\
+		\addlinespace
+		opq:verbbe\_right & -0.1539128 & 0.0192901 & 54372.283 & -7.9788617 & 0.0000000\\
+		opc:verbconfess & 0.0841131 & 0.0190632 & 54372.437 & 4.4123417 & 0.0000102\\
+		opm:verbconfess & 0.0834691 & 0.0189719 & 54372.142 & 4.3996174 & 0.0000109\\
+		opq:verbconfess & -0.0368642 & 0.0192905 & 54372.632 & -1.9110088 & 0.0560087\\
+		opc:verbconfirm & 0.0693734 & 0.0190644 & 54373.529 & 3.6388992 & 0.0002741\\
+		\addlinespace
+		opm:verbconfirm & 0.2660910 & 0.0189728 & 54372.917 & 14.0248948 & 0.0000000\\
+		opq:verbconfirm & -0.0449413 & 0.0192914 & 54373.472 & -2.3295974 & 0.0198311\\
+		opc:verbdemonstrate & 0.0584268 & 0.0190633 & 54372.586 & 3.0648793 & 0.0021786\\
+		opm:verbdemonstrate & 0.1781393 & 0.0189714 & 54371.707 & 9.3898743 & 0.0000000\\
+		opq:verbdemonstrate & -0.0008729 & 0.0192906 & 54372.707 & -0.0452486 & 0.9639093\\
+		\addlinespace
+		opc:verbestablish & 0.0300718 & 0.0190618 & 54371.161 & 1.5775980 & 0.1146638\\
+		opm:verbestablish & 0.1877158 & 0.0189720 & 54372.212 & 9.8943713 & 0.0000000\\
+		opq:verbestablish & -0.0700678 & 0.0192899 & 54372.103 & -3.6323635 & 0.0002811\\
+		opc:verbhear & -0.0253789 & 0.0190637 & 54372.970 & -1.3312665 & 0.1831069\\
+		opm:verbhear & -0.1048737 & 0.0189721 & 54372.368 & -5.5277707 & 0.0000000\\
+		\addlinespace
+		opq:verbhear & -0.0006261 & 0.0192909 & 54372.991 & -0.0324581 & 0.9741069\\
+		opc:verbinform & -0.0991359 & 0.0190633 & 54372.605 & -5.2003420 & 0.0000002\\
+		opm:verbinform & -0.0860938 & 0.0189733 & 54373.368 & -4.5376355 & 0.0000057\\
+		opq:verbinform & -0.0174718 & 0.0192931 & 54374.843 & -0.9055994 & 0.3651520\\
+		opc:verbknow & -0.1127399 & 0.0190628 & 54372.127 & -5.9141311 & 0.0000000\\
+		\addlinespace
+		opm:verbknow & -0.0872392 & 0.0189719 & 54372.149 & -4.5983336 & 0.0000043\\
+		opq:verbknow & -0.0192684 & 0.0192895 & 54371.814 & -0.9989027 & 0.3178463\\
+		opc:verbpretend & -0.3172145 & 0.0190639 & 54373.134 & -16.6395086 & 0.0000000\\
+		opm:verbpretend & -0.1196669 & 0.0189725 & 54372.658 & -6.3073967 & 0.0000000\\
+		opq:verbpretend & -0.2959105 & 0.0192916 & 54373.627 & -15.3388172 & 0.0000000\\
+		\addlinespace
+		opc:verbprove & 0.0309486 & 0.0190626 & 54371.908 & 1.6235265 & 0.1044827\\
+		opm:verbprove & 0.2578102 & 0.0189728 & 54372.987 & 13.5883819 & 0.0000000\\
+		opq:verbprove & -0.0531478 & 0.0192907 & 54372.839 & -2.7550974 & 0.0058694\\
+		opc:verbreveal & 0.0632551 & 0.0190641 & 54373.234 & 3.3180321 & 0.0009071\\
+		opm:verbreveal & 0.0765052 & 0.0189711 & 54371.383 & 4.0327309 & 0.0000552\\
+		\addlinespace
+		opq:verbreveal & -0.0096396 & 0.0192917 & 54373.669 & -0.4996755 & 0.6173056\\
+		opc:verbsay & -0.0284748 & 0.0190633 & 54372.590 & -1.4936938 & 0.1352615\\
+		opm:verbsay & 0.0449550 & 0.0189713 & 54371.586 & 2.3696339 & 0.0178092\\
+		opq:verbsay & -0.2578020 & 0.0192911 & 54373.227 & -13.3637456 & 0.0000000\\
+		opc:verbsee & -0.0094907 & 0.0190633 & 54372.613 & -0.4978528 & 0.6185898\\
+		\addlinespace
+		opm:verbsee & -0.0408332 & 0.0189714 & 54371.728 & -2.1523480 & 0.0313743\\
+		opq:verbsee & 0.0345992 & 0.0192916 & 54373.623 & 1.7934827 & 0.0729013\\
+		opc:verbthink & -0.2246902 & 0.0190628 & 54372.159 & -11.7868151 & 0.0000000\\
+		opm:verbthink & -0.0856847 & 0.0189719 & 54372.175 & -4.5163904 & 0.0000063\\
+		opq:verbthink & -0.2500624 & 0.0192911 & 54373.178 & -12.9625865 & 0.0000000\\
+		\bottomrule
+	\end{tabular}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+	\caption{Model output w \texttt{know/n} as baseline}
+	\label{tab:discover-n}
+	\centering
+
+	\begin{tabular}{lrrrrr}
+		\toprule
+		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
+		\midrule
+		(Intercept) & 0.7921005 & 0.0114401 & 1479.182 & 69.2391632 & 0.0000000\\
+		opc & -0.0013151 & 0.0156638 & 26226.099 & -0.0839558 & 0.9330922\\
+		opm & -0.1434448 & 0.0155855 & 26286.727 & -9.2037503 & 0.0000000\\
+		opq & 0.0818736 & 0.0143093 & 56827.205 & 5.7217106 & 0.0000000\\
+		verbdiscover & -0.1115762 & 0.0134931 & 54371.661 & -8.2691029 & 0.0000000\\
+		\addlinespace
+		verbbe\_annoyed & 0.0748953 & 0.0134936 & 54372.286 & 5.5504198 & 0.0000000\\
+		verbsuggest & -0.5413729 & 0.0134928 & 54371.215 & -40.1230775 & 0.0000000\\
+		verbacknowledge & -0.1455464 & 0.0134940 & 54372.778 & -10.7859911 & 0.0000000\\
+		verbadmit & -0.2653476 & 0.0134936 & 54372.243 & -19.6647018 & 0.0000000\\
+		verbannounce & -0.2858550 & 0.0134935 & 54372.051 & -21.1847253 & 0.0000000\\
+		\addlinespace
+		verbbe\_right & -0.5889971 & 0.0134936 & 54372.234 & -43.6501264 & 0.0000000\\
+		verbconfess & -0.2932742 & 0.0134933 & 54371.896 & -21.7347532 & 0.0000000\\
+		verbconfirm & -0.5589410 & 0.0134943 & 54373.134 & -41.4204989 & 0.0000000\\
+		verbdemonstrate & -0.4558072 & 0.0134931 & 54371.605 & -33.7807521 & 0.0000000\\
+		verbestablish & -0.4793608 & 0.0134938 & 54372.501 & -35.5245129 & 0.0000000\\
+		\addlinespace
+		verbhear & -0.1207841 & 0.0134931 & 54371.608 & -8.9515418 & 0.0000000\\
+		verbinform & -0.0625474 & 0.0134940 & 54372.811 & -4.6351853 & 0.0000036\\
+		verbpretend & -0.4300580 & 0.0134939 & 54372.685 & -31.8704317 & 0.0000000\\
+		verbprove & -0.5615997 & 0.0134934 & 54372.014 & -41.6202578 & 0.0000000\\
+		verbreveal & -0.2308861 & 0.0134943 & 54373.179 & -17.1098402 & 0.0000000\\
+		\addlinespace
+		verbsay & -0.4670771 & 0.0134939 & 54372.578 & -34.6140392 & 0.0000000\\
+		verbsee & -0.1078558 & 0.0134932 & 54371.715 & -7.9933523 & 0.0000000\\
+		verbthink & -0.4352245 & 0.0134933 & 54371.896 & -32.2547883 & 0.0000000\\
+		opc:verbdiscover & 0.1127399 & 0.0190628 & 54372.122 & 5.9141311 & 0.0000000\\
+		opm:verbdiscover & 0.0872392 & 0.0189719 & 54372.145 & 4.5983336 & 0.0000043\\
+		\addlinespace
+		opq:verbdiscover & 0.0192684 & 0.0192895 & 54371.810 & 0.9989027 & 0.3178463\\
+		opc:verbbe\_annoyed & -0.1142699 & 0.0190629 & 54372.206 & -5.9943630 & 0.0000000\\
+		opm:verbbe\_annoyed & -0.0130032 & 0.0189709 & 54371.180 & -0.6854301 & 0.4930755\\
+		opq:verbbe\_annoyed & -0.0569598 & 0.0192897 & 54371.974 & -2.9528580 & 0.0031498\\
+		opc:verbsuggest & 0.0451626 & 0.0190614 & 54370.865 & 2.3693169 & 0.0178244\\
+		\addlinespace
+		opm:verbsuggest & 0.1636081 & 0.0189713 & 54371.561 & 8.6239932 & 0.0000000\\
+		opq:verbsuggest & -0.1307246 & 0.0192896 & 54371.847 & -6.7769511 & 0.0000000\\
+		opc:verbacknowledge & 0.0802870 & 0.0190639 & 54373.059 & 4.2114778 & 0.0000254\\
+		opm:verbacknowledge & 0.0970521 & 0.0189731 & 54373.215 & 5.1152465 & 0.0000003\\
+		opq:verbacknowledge & 0.0103337 & 0.0192900 & 54372.183 & 0.5357031 & 0.5921659\\
+		\addlinespace
+		opc:verbadmit & 0.1660650 & 0.0190633 & 54372.568 & 8.7112405 & 0.0000000\\
+		opm:verbadmit & 0.1680676 & 0.0189725 & 54372.674 & 8.8584875 & 0.0000000\\
+		opq:verbadmit & -0.0129548 & 0.0192897 & 54371.901 & -0.6715939 & 0.5018451\\
+		opc:verbannounce & 0.1361230 & 0.0190629 & 54372.249 & 7.1407107 & 0.0000000\\
+		opm:verbannounce & 0.1502994 & 0.0189716 & 54371.859 & 7.9223393 & 0.0000000\\
+		\addlinespace
+		opq:verbannounce & -0.0552066 & 0.0192897 & 54371.911 & -2.8619784 & 0.0042117\\
+		opc:verbbe\_right & 0.0281618 & 0.0190625 & 54371.883 & 1.4773357 & 0.1395915\\
+		opm:verbbe\_right & 0.2452367 & 0.0189720 & 54372.236 & 12.9262410 & 0.0000000\\
+		opq:verbbe\_right & -0.1346444 & 0.0192901 & 54372.308 & -6.9799743 & 0.0000000\\
+		opc:verbconfess & 0.1968531 & 0.0190627 & 54372.016 & 10.3266157 & 0.0000000\\
+		\addlinespace
+		opm:verbconfess & 0.1707082 & 0.0189737 & 54373.731 & 8.9971020 & 0.0000000\\
+		opq:verbconfess & -0.0175959 & 0.0192891 & 54371.441 & -0.9122160 & 0.3616590\\
+		opc:verbconfirm & 0.1821133 & 0.0190626 & 54371.955 & 9.5534239 & 0.0000000\\
+		opm:verbconfirm & 0.3533301 & 0.0189718 & 54372.046 & 18.6239655 & 0.0000000\\
+		opq:verbconfirm & -0.0256729 & 0.0192917 & 54373.707 & -1.3307729 & 0.1832693\\
+		\addlinespace
+		opc:verbdemonstrate & 0.1711667 & 0.0190647 & 54373.791 & 8.9782020 & 0.0000000\\
+		opm:verbdemonstrate & 0.2653784 & 0.0189716 & 54371.911 & 13.9881581 & 0.0000000\\
+		opq:verbdemonstrate & 0.0183955 & 0.0192903 & 54372.499 & 0.9536136 & 0.3402835\\
+		opc:verbestablish & 0.1428117 & 0.0190634 & 54372.667 & 7.4914038 & 0.0000000\\
+		opm:verbestablish & 0.2749549 & 0.0189719 & 54372.156 & 14.4927310 & 0.0000000\\
+		\addlinespace
+		opq:verbestablish & -0.0507994 & 0.0192892 & 54371.529 & -2.6335636 & 0.0084518\\
+		opc:verbhear & 0.0873610 & 0.0190622 & 54371.557 & 4.5829467 & 0.0000046\\
+		opm:verbhear & -0.0176345 & 0.0189709 & 54371.193 & -0.9295579 & 0.3526042\\
+		opq:verbhear & 0.0186422 & 0.0192902 & 54372.432 & 0.9664076 & 0.3338446\\
+		opc:verbinform & 0.0136041 & 0.0190628 & 54372.117 & 0.7136441 & 0.4754504\\
+		\addlinespace
+		opm:verbinform & 0.0011454 & 0.0189738 & 54373.859 & 0.0603655 & 0.9518647\\
+		opq:verbinform & 0.0017966 & 0.0192921 & 54374.044 & 0.0931259 & 0.9258039\\
+		opc:verbpretend & -0.2044746 & 0.0190640 & 54373.227 & -10.7256652 & 0.0000000\\
+		opm:verbpretend & -0.0324277 & 0.0189726 & 54372.773 & -1.7091875 & 0.0874219\\
+		opq:verbpretend & -0.2766421 & 0.0192909 & 54373.044 & -14.3405213 & 0.0000000\\
+		\addlinespace
+		opc:verbprove & 0.1436885 & 0.0190622 & 54371.604 & 7.5378609 & 0.0000000\\
+		opm:verbprove & 0.3450493 & 0.0189712 & 54371.500 & 18.1880552 & 0.0000000\\
+		opq:verbprove & -0.0338794 & 0.0192909 & 54373.040 & -1.7562330 & 0.0790543\\
+		opc:verbreveal & 0.1759951 & 0.0190637 & 54372.960 & 9.2319255 & 0.0000000\\
+		opm:verbreveal & 0.1637444 & 0.0189722 & 54372.374 & 8.6307727 & 0.0000000\\
+		\addlinespace
+		opq:verbreveal & 0.0096288 & 0.0192902 & 54372.406 & 0.4991553 & 0.6176720\\
+		opc:verbsay & 0.0842652 & 0.0190632 & 54372.497 & 4.4203002 & 0.0000099\\
+		opm:verbsay & 0.1321942 & 0.0189719 & 54372.094 & 6.9679113 & 0.0000000\\
+		opq:verbsay & -0.2385336 & 0.0192898 & 54371.999 & -12.3658155 & 0.0000000\\
+		opc:verbsee & 0.1032492 & 0.0190630 & 54372.324 & 5.4162003 & 0.0000001\\
+		\addlinespace
+		opm:verbsee & 0.0464060 & 0.0189718 & 54372.089 & 2.4460454 & 0.0144464\\
+		opq:verbsee & 0.0538675 & 0.0192901 & 54372.298 & 2.7924979 & 0.0052321\\
+		opc:verbthink & -0.1119503 & 0.0190627 & 54372.030 & -5.8727378 & 0.0000000\\
+		opm:verbthink & 0.0015545 & 0.0189713 & 54371.557 & 0.0819391 & 0.9346955\\
+		opq:verbthink & -0.2307940 & 0.0192903 & 54372.447 & -11.9642778 & 0.0000000\\
+		\bottomrule
+	\end{tabular}
+\end{table}
+
 \bibliography{projective-content.bib}

### 2022-12-08 results/main/meta-analyses/projectivity/rscripts/analysis.R

@@ -11,26 +11,23 @@ data <- read.csv("../data/data_combined.csv", header = TRUE, sep = ",")
 
 # variables and contrasts for data analysis: ----
 ## our dependent variable: projective (sliding-scale ratings of speaker certainty that complement clause is true)
-data$projective <- numeric(data$projective)
+data$projective <- as.numeric(data$projective)
 
 # our independent variables
 ## inference-triggering predicate, coded as "verb"
 data$verb <- as.factor(data$verb)
 levels(data$verb)
 contrasts(data$verb)
-### dummy coding with "suggest" as baseline
-data$verb <- relevel(data$verb, ref = "suggest")
+### dummy coding with "be annoyed" as baseline
+data$verb <- relevel(data$verb, ref = "be_annoyed")
 contrasts(data$verb)
-
-#### this was our choice for a baseline, because this is the verb with the 
-#### smallest differences between by-operator projectivity (ostensibly, based on graph)
 
 ## embedding operator, coded as "op"
 data$op <- as.factor(data$op)
 levels(data$op)
 contrasts(data$op)
-### dummy coding with "m" (conditionals) as baseline
-data$op <- relevel(data$op, ref = "m")
+### dummy coding with "n" (negation) as baseline
+data$op <- relevel(data$op, ref = "n")
 contrasts(data$op)
 
 # coding random effects as factors
@@ -41,19 +38,31 @@ data$content <- as.factor(data$content)
 # exploring a couple of models (linear)
 ## library for linear mixed models
 library(lme4)
-library(lmertest)
+library(lmerTest)
+library(knitr)
 
-# glmm1 <- lmer(projective ~ verb * op + (op | workerid) +
-#                  (op | content), data=data)
+# glmm1 <- lmer(projective ~ op * verb + (1 | workerid) +
+#                  (1 | content), data=data)
 
 # save.image("linear-models.RData")
 load("linear-models.RData")
-print(summary(glmm1), cor=F, dig=3)
 
-# glmm2 <- lmer(projective ~ verb * op + (op | workerid) +
-#                 (op | content), data=data)
+summary1 <- summary(glmm1)
+print(summary1, cor=F, dig=3)
+kable(summary1$coefficients, format = "latex", booktabs = TRUE)
+
+
+### dummy coding with "discover" as baseline
+data$verb <- relevel(data$verb, ref = "discover")
+contrasts(data$verb)
+
[…]

### 2022-12-08 SALT-abstract/SALT-abstract.tex
updated SALT abstract

@@
 \usepackage{setspace}
+
+\parindent=2ex
+
 \begin{document}
@@
 {\bf Results.} Fig.~\ref{fig:figure1} plots the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
-We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under questions and negation than from under conditionals and modals, the CC of \emph{know} projects more from under questions than from under conditionals and negation, and the CC of \emph{discover} projects more from under questions and conditionals than from under negation and modals.
-These results are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
+We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than from under conditionals and modals, the CC of \emph{know} projects less from under negation than under questions, but more from under negation than from under modals, and the CC of \emph{discover} projects less from negation than from under conditionals and questions, and more from under negation than modals.
+These results (which are confirmed by linear mixed effects models, see Table \ref{t:models}) are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
 For {\em know}, we observe a third pattern -- one that is not aligned with a classification as a factive nor as a semi-factive predicate.
-
-==== THIS IS AS FAR AS I've COME ====
 
@@
 
-	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
+%	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
 	% The mean for this baseline (intercept) is $0.867$.
-	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter.
+%	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter.
 We also found many interactions of \texttt{operator} and \texttt{verb} across the board, suggesting that the effect of embedding context diffes by verb.
@@
 	\end{figure}
+
+\begin{table}[h]
+\centering
+\begin{tabular}{llrrrr}
+Model & Estimate & Std. Error & t value & Pr(>|t|)\\
+\midrule
+\#1 & Intercept: \emph{be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
+& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
+& operator: modal & -0.16 & 0.02  & -10.04 & *** \\
+& operator: question & 0.02 & 0.01  & 1.74 & n.s. \\
+\midrule
+\#2 & Intercept: \emph{know}/negation & 0.79 & 0.01 &  69.24 & *** \\
+&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
+&		operator: modal  & -0.14 & 0.02  & -9.2 & *** \\
+&		operator: question & 0.08 & 0.01  & 5.72 & *** \\
+\midrule
+\#3 & Intercept: \emph{discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
+& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
+&		operator: modal & -0.06 & 0.02  & -3.6 & *** \\
+&		operator: question & 0.1 & 0.01  & 7.07 & *** \\
+
+\bottomrule
+\end{tabular}
+\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC.
+Models were fit with \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}.}\label{t:models}
+\end{table}
 

### 2022-12-09 SALT-abstract/SALT-abstract.tex
updates

--- 2022-12-08 SALT-abstract/SALT-abstract.tex
@@
 % \maketitle
-\textcolor{white}{.} \vspace{-3.9\baselineskip} \\
+
 \begin{center}
@@
 
-\vspace*{-.4cm}
-
+\vspace*{-.6\baselineskip}
 \noindent 
 	We present experimental evidence that i) the projection of the content of the clausal complement of attitude predicates varies across entailment-canceling operators (negation, question, modal, conditional) and ii) this by-operator variation differs across attitude predicates.
-Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealt{karttunen_observations_1971,djarv_cognitive_2018}).
+Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}, though see \citealp{beaver_have_2010}).
 Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.
@@
 \noindent
-{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters one of the attitude ascriptions in  \ref{ex:family} is committed to the truth of the content of the complement (CC), that Julian dances salsa, even though the complement occurs under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, a modal \ref{ex:mod}, or a conditional \ref{ex:cond}.
+	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters one of the attitude ascriptions in  \ref{ex:family} is committed to the truth of the content of the complement (CC),
+	% that Julian dances salsa,
+	even though it occurs under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, a modal \ref{ex:mod}, or a conditional \ref{ex:cond}.
 
@@
 \noindent
-{\bf Results.} Fig.~\ref{fig:figure1} plots the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
-We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under questions and negation than from under conditionals and modals, the CC of \emph{know} projects more from under questions than from under conditionals and negation, and the CC of \emph{discover} projects more from under questions and conditionals than from under negation and modals.
-These results are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
+{\bf Results.}
+	Fig.~\ref{fig:figure1} shows the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
+We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than from under conditionals and modals, the CC of \emph{know} projects less from under negation than under questions, but more from under negation than from under modals, and the CC of \emph{discover} projects less from negation than from under conditionals and questions, and more from under negation than modals.
+These results (which are supported by linear mixed effects models, see Table \ref{t:models}) are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
 For {\em know}, we observe a third pattern -- one that is not aligned with a classification as a factive nor as a semi-factive predicate.
-
-==== THIS IS AS FAR AS I've COME ====
 
@@
 
-	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
+	%	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
 		% The mean for this baseline (intercept) is $0.867$.
-	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter.
+	%	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter.
 We also found many interactions of \texttt{operator} and \texttt{verb} across the board, suggesting that the effect of embedding context diffes by verb.
@@
 \noindent
-{\bf Discussion.}
+{\bf Discussion and Conclusion.} 
+	Our results clearly show that different entailment-cancelling operators affect projection in different ways, and that the effect of entailment-cancelling operator on projection differs for different verbs.
+The picture that emerges could not be characterized in terms of Karttunen's characterization of (semi)factives.
+Our results indicate a need for a more nuanced view of the family-of-sentences diagnostics of projection, as well as a better understanding of the effect of the interaction between lexical semantics and operators on projection.
 
-We observe that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}; replicating the results of  \citealt{degen-tonhauser-language}).
+	Further, we were able to replicate the results of \citealp{degen_are_2022}), observing that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}, and finding no categorical distinction between factive and non-factive predicates either.
 
-No categorical distinction between factive and non-factive predicates either
+	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
+\textbf{Figure \ref{fig:figure2}} gives the mean certainty ratings for the four operators by verb, for some groups of our verbs that group together in the way that they show by-operator variation.
+We highlight four \lq projection-profiles\rq here: the verbs \emph{pretend, think} are the only verbs that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
+\emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}).
+\emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}).
+Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).
 
-add discussion about verb classes here
+	The picture that emerges is complex and requires a nuanced view of lexical effects on projection.
+I raises the question of how lexical differences between verbs could explain why certain verbs project more from under certain operators and less from others.
+While much previous work on this question has based their generalizations on impressionistic judgments, empirically establishing different projection-profiles associated with verbs therefore provides a basis for future research on lexical effects on projection.
 
@@
 
-\begin{figure}[t]
-		\vspace{-.8\baselineskip}
+\begin{figure}[h]
+	% \vspace{-\baselineskip}
 	\centering
-		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
+	\scalebox{.9}{\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}}\vspace{-1.7\baselineskip}
 	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
@@
 
+\begin{table}[h]
+	\centering
+	\vspace{-.7\baselineskip}
+	\scalebox{.85}{
+		\begin{tabular}{llrrrr}
+			Model & & Estimate & Std. Error & t-value\\
+			\midrule
+			\#1 & Intercept: \emph{be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
+			& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
+			& operator: modal & -0.16 & 0.02  & -10.04 & *** \\
+			& operator: question & 0.02 & 0.01  & 1.74 & n.s. \\
+			\midrule
+			\#2 & Intercept: \emph{know}/negation & 0.79 & 0.01 &  69.24 & *** \\
+			&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
+			&		operator: modal  & -0.14 & 0.02  & -9.2 & *** \\
+			&		operator: question & 0.08 & 0.01  & 5.72 & *** \\
+			\midrule
+			\#3 & Intercept: \emph{discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
+			& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
+			&		operator: modal & -0.06 & 0.02  & -3.6 & *** \\
+			&		operator: question & 0.1 & 0.01  & 7.07 & *** \\
+
+			\bottomrule
+		\end{tabular}
+	}
+	\vspace{-.3\baselineskip}
+	\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC.
+Models were fit with \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}.}\label{t:models}
+\end{table}
+
+\begin{figure}[h]
+	\hspace{-2em}
+	\begin{minipage}{1.1\textwidth}\centering
+		\includegraphics[width=.24\linewidth]{graphs/profile1.pdf}
+		\includegraphics[width=.24\linewidth]{graphs/profile2.pdf}
+		% \includegraphics[width=.32\linewidth]{graphs/profile3.pdf}\\
+		\includegraphics[width=.24\linewidth]{graphs/profile4.pdf}
+		\includegraphics[width=.24\linewidth]{graphs/profile5.pdf}
+	\end{minipage}
+	\vspace{-\baselineskip}
+	\caption{Mean certainty-ratings by operator for each verb, showing different \lq verb-profiles\rq.}
+	\label{fig:figure2}
+\end{figure}
+
 \newpage
+
+% \begin{table}[ht]
+% 	\caption{Model output w \texttt{be\_annoyed/n} as baseline}
+% 	\label{tab:annoyed-n}
+% 	\centering
+
+% 	\begin{tabular}{lrrrrr}
+% 		\toprule
+% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
+% 		\midrule
+% 		(Intercept) & 0.8669958 & 0.0114402 & 1479.204 & 75.7850523 & 0.0000000\\
+% 		opc & -0.1155850 & 0.0156641 & 26227.368 & -7.3789836 & 0.0000000\\
+% 		opm & -0.1564480 & 0.0155858 & 26288.139 & -10.0378702 & 0.0000000\\
+% 		opq & 0.0249138 & 0.0143090 & 56826.915 & 1.7411232 & 0.0816674\\
+% 		verbsuggest & -0.6162683 & 0.0134941 & 54372.895 & -45.6694112 & 0.0000000\\
+% 		\addlinespace
+% 		verbacknowledge & -0.2204417 & 0.0134948 & 54373.777 & -16.3352724 & 0.0000000\\
+% 		verbadmit & -0.3402430 & 0.0134933 & 54371.874 & -25.2156750 & 0.0000000\\
+% 		verbannounce & -0.3607504 & 0.0134931 & 54371.547 & -26.7359958 & 0.0000000\\
+% 		verbbe\_right & -0.6638924 & 0.0134936 & 54372.293 & -49.2003974 & 0.0000000\\
+% 		verbconfess & -0.3681695 & 0.0134949 & 54373.904 & -27.2820517 & 0.0000000\\
+% 		\addlinespace
+% 		verbconfirm & -0.6338363 & 0.0134931 & 54371.657 & -46.9746939 & 0.0000000\\
+% 		verbdemonstrate & -0.5307026 & 0.0134935 & 54372.157 & -39.3301466 & 0.0000000\\
+% 		verbdiscover & -0.1864716 & 0.0134942 & 54372.992 & -13.8186520 & 0.0000000\\
+% 		verbestablish & -0.5542561 & 0.0134933 & 54371.853 & -41.0764079 & 0.0000000\\
+% 		verbhear & -0.1956794 & 0.0134938 & 54372.473 & -14.5014554 & 0.0000000\\
+% 		\addlinespace
+% 		verbinform & -0.1374428 & 0.0134945 & 54373.345 & -10.1851106 & 0.0000000\\
+% 		verbknow & -0.0748953 & 0.0134936 & 54372.288 & -5.5504197 & 0.0000000\\
+% 		verbpretend & -0.5049533 & 0.0134944 & 54373.218 & -37.4195409 & 0.0000000\\
+% 		verbprove & -0.6364950 & 0.0134935 & 54372.138 & -47.1704273 & 0.0000000\\
+% 		verbreveal & -0.3057814 & 0.0134937 & 54372.418 & -22.6609884 & 0.0000000\\
+% 		\addlinespace
+% 		verbsay & -0.5419725 & 0.0134925 & 54370.770 & -40.1685159 & 0.0000000\\
+% 		verbsee & -0.1827512 & 0.0134932 & 54371.791 & -13.5438968 & 0.0000000\\
+% 		verbthink & -0.5101198 & 0.0134930 & 54371.491 & -37.8062103 & 0.0000000\\
+% 		opc:verbsuggest & 0.1594325 & 0.0190634 & 54372.626 & 8.3632939 & 0.0000000\\
+% 		opm:verbsuggest & 0.1766113 & 0.0189718 & 54372.046 & 9.3091498 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbsuggest & -0.0737648 & 0.0192906 & 54372.717 & -3.8238781 & 0.0001315\\
+% 		opc:verbacknowledge & 0.1945569 & 0.0190630 & 54372.333 & 10.2059764 & 0.0000000\\
+% 		opm:verbacknowledge & 0.1100553 & 0.0189729 & 54373.070 & 5.8006461 & 0.0000000\\
+% 		opq:verbacknowledge & 0.0672935 & 0.0192909 & 54372.991 & 3.4883599 & 0.0004864\\
+% 		opc:verbadmit & 0.2803349 & 0.0190626 & 54371.933 & 14.7060204 & 0.0000000\\
+% 		\addlinespace
+% 		opm:verbadmit & 0.1810708 & 0.0189717 & 54371.994 & 9.5442376 & 0.0000000\\
+% 		opq:verbadmit & 0.0440050 & 0.0192896 & 54371.832 & 2.2812856 & 0.0225354\\
+% 		opc:verbannounce & 0.2503929 & 0.0190639 & 54373.132 & 13.1343759 & 0.0000000\\
+% 		opm:verbannounce & 0.1633026 & 0.0189721 & 54372.338 & 8.6075049 & 0.0000000\\
+% 		opq:verbannounce & 0.0017532 & 0.0192899 & 54372.121 & 0.0908892 & 0.9275810\\
+% 		\addlinespace
+% 		opc:verbbe\_right & 0.1424317 & 0.0190623 & 54371.698 & 7.4718895 & 0.0000000\\
+% 		opm:verbbe\_right & 0.2582399 & 0.0189717 & 54371.967 & 13.6118412 & 0.0000000\\
+% 		opq:verbbe\_right & -0.0776846 & 0.0192897 & 54371.974 & -4.0272512 & 0.0000565\\
+% 		opc:verbconfess & 0.3111230 & 0.0190643 & 54373.455 & 16.3196605 & 0.0000000\\
+% 		opm:verbconfess & 0.1837114 & 0.0189741 & 54374.125 & 9.6821973 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbconfess & 0.0393640 & 0.0192890 & 54371.320 & 2.0407459 & 0.0412809\\
+% 		opc:verbconfirm & 0.2963833 & 0.0190629 & 54372.237 & 15.5476205 & 0.0000000\\
+% 		opm:verbconfirm & 0.3663333 & 0.0189721 & 54372.324 & 19.3090505 & 0.0000000\\
+% 		opq:verbconfirm & 0.0312870 & 0.0192902 & 54372.377 & 1.6219108 & 0.1048282\\
+% 		opc:verbdemonstrate & 0.2854366 & 0.0190647 & 54373.829 & 14.9719656 & 0.0000000\\
+% 		\addlinespace
+% 		opm:verbdemonstrate & 0.2783816 & 0.0189716 & 54371.830 & 14.6736292 & 0.0000000\\
+% 		opq:verbdemonstrate & 0.0753554 & 0.0192892 & 54371.520 & 3.9066023 & 0.0000937\\
+% 		opc:verbdiscover & 0.2270099 & 0.0190632 & 54372.450 & 11.9082962 & 0.0000000\\
+% 		opm:verbdiscover & 0.1002424 & 0.0189726 & 54372.795 & 5.2835252 & 0.0000001\\
+% 		opq:verbdiscover & 0.0762282 & 0.0192898 & 54372.014 & 3.9517423 & 0.0000777\\
+% 		\addlinespace
+% 		opc:verbestablish & 0.2570817 & 0.0190633 & 54372.539 & 13.4857080 & 0.0000000\\
+% 		opm:verbestablish & 0.2879581 & 0.0189714 & 54371.677 & 15.1785426 & 0.0000000\\
+% 		opq:verbestablish & 0.0061604 & 0.0192893 & 54371.583 & 0.3193687 & 0.7494481\\
+% 		opc:verbhear & 0.2016309 & 0.0190630 & 54372.273 & 10.5770983 & 0.0000000\\
+% 		opm:verbhear & -0.0046313 & 0.0189714 & 54371.681 & -0.2441216 & 0.8071376\\
+% 		\addlinespace
+% 		opq:verbhear & 0.0756021 & 0.0192902 & 54372.373 & 3.9192008 & 0.0000890\\
+% 		opc:verbinform & 0.1278740 & 0.0190645 & 54373.631 & 6.7074362 & 0.0000000\\
+% 		opm:verbinform & 0.0141486 & 0.0189741 & 54374.122 & 0.7456761 & 0.4558664\\
+% 		opq:verbinform & 0.0587564 & 0.0192913 & 54373.322 & 3.0457537 & 0.0023221\\
+% 		opc:verbknow & 0.1142699 & 0.0190629 & 54372.206 & 5.9943629 & 0.0000000\\
+% 		\addlinespace
+% 		opm:verbknow & 0.0130032 & 0.0189709 & 54371.181 & 0.6854301 & 0.4930755\\
+% 		opq:verbknow & 0.0569598 & 0.0192897 & 54371.975 & 2.9528579 & 0.0031498\\
+% 		opc:verbpretend & -0.0902047 & 0.0190640 & 54373.155 & -4.7316842 & 0.0000022\\
+% 		opm:verbpretend & -0.0194245 & 0.0189738 & 54373.822 & -1.0237558 & 0.3059552\\
+% 		opq:verbpretend & -0.2196823 & 0.0192900 & 54372.250 & -11.3883795 & 0.0000000\\
+% 		\addlinespace
+% 		opc:verbprove & 0.2579585 & 0.0190627 & 54372.031 & 13.5320999 & 0.0000000\\
+% 		opm:verbprove & 0.3580525 & 0.0189720 & 54372.223 & 18.8726898 & 0.0000000\\
+% 		opq:verbprove & 0.0230805 & 0.0192908 & 54372.899 & 1.1964509 & 0.2315259\\
+% 		opc:verbreveal & 0.2902650 & 0.0190644 & 54373.511 & 15.2255216 & 0.0000000\\
+% 		opm:verbreveal & 0.1767476 & 0.0189723 & 54372.513 & 9.3160804 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbreveal & 0.0665887 & 0.0192906 & 54372.717 & 3.4518763 & 0.0005571\\
+% 		opc:verbsay & 0.1985351 & 0.0190618 & 54371.193 & 10.4153412 & 0.0000000\\
+% 		opm:verbsay & 0.1451974 & 0.0189723 & 54372.495 & 7.6531273 & 0.0000000\\
+% 		opq:verbsay & -0.1815738 & 0.0192889 & 54371.195 & -9.4133926 & 0.0000000\\
+% 		opc:verbsee & 0.2175191 & 0.0190629 & 54372.234 & 11.4105822 & 0.0000000\\
+% 		\addlinespace
+% 		opm:verbsee & 0.0594092 & 0.0189723 & 54372.499 & 3.1313655 & 0.0017409\\
+% 		opq:verbsee & 0.1108274 & 0.0192896 & 54371.840 & 5.7454526 & 0.0000000\\
+% 		opc:verbthink & 0.0023197 & 0.0190619 & 54371.272 & 0.1216913 & 0.9031440\\
+% 		opm:verbthink & 0.0145577 & 0.0189717 & 54371.965 & 0.7673367 & 0.4428847\\
+% 		opq:verbthink & -0.1738342 & 0.0192903 & 54372.438 & -9.0115053 & 0.0000000\\
+% 		\bottomrule
+% 	\end{tabular}
+% \end{table}
+
+% \newpage
+
+% \begin{table}[ht]
+% 	\caption{Model output w \texttt{discover/n} as baseline}
+% 	\label{tab:discover-n}
+% 	\centering
+
+% 	\begin{tabular}{lrrrrr}
+% 		\toprule
+% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
+% 		\midrule
+% 		(Intercept) & 0.6805242 & 0.0114403 & 1479.217 & 59.4849565 & 0.0000000\\
+% 		opc & 0.1114249 & 0.0156639 & 26226.686 & 7.1134644 & 0.0000000\\
+% 		opm & -0.0562056 & 0.0155855 & 26286.826 & -3.6062779 & 0.0003112\\
+% 		opq & 0.1011420 & 0.0143103 & 56828.295 & 7.0677830 & 0.0000000\\
+% 		verbbe\_annoyed & 0.1864716 & 0.0134942 & 54372.998 & 13.8186521 & 0.0000000\\
+% 		\addlinespace
+% 		verbsuggest & -0.4297967 & 0.0134937 & 54372.361 & -31.8516905 & 0.0000000\\
+% 		verbacknowledge & -0.0339702 & 0.0134936 & 54372.308 & -2.5174950 & 0.0118221\\
+% 		verbadmit & -0.1537714 & 0.0134942 & 54372.985 & -11.3953857 & 0.0000000\\
+% 		verbannounce & -0.1742788 & 0.0134947 & 54373.649 & -12.9145926 & 0.0000000\\
+% 		verbbe\_right & -0.4774208 & 0.0134935 & 54372.079 & -35.3816304 & 0.0000000\\
+% 		\addlinespace
+% 		verbconfess & -0.1816980 & 0.0134936 & 54372.192 & -13.4655376 & 0.0000000\\
+% 		verbconfirm & -0.4473648 & 0.0134944 & 54373.311 & -33.1517710 & 0.0000000\\
+% 		verbdemonstrate & -0.3442310 & 0.0134928 & 54371.247 & -25.5121517 & 0.0000000\\
+% 		verbestablish & -0.3677846 & 0.0134928 & 54371.167 & -27.2579106 & 0.0000000\\
+% 		verbhear & -0.0092079 & 0.0134940 & 54372.798 & -0.6823676 & 0.4950095\\
+% 		\addlinespace
+% 		verbinform & 0.0490288 & 0.0134942 & 54373.066 & 3.6333114 & 0.0002801\\
+% 		verbknow & 0.1115762 & 0.0134931 & 54371.668 & 8.2691029 & 0.0000000\\
+% 		verbpretend & -0.3184818 & 0.0134943 & 54373.176 & -23.6011440 & 0.0000000\\
+% 		verbprove & -0.4500234 & 0.0134934 & 54372.042 & -33.3512833 & 0.0000000\\
+% 		verbreveal & -0.1193099 & 0.0134941 & 54372.880 & -8.8416323 & 0.0000000\\
+% 		\addlinespace
+% 		verbsay & -0.3555009 & 0.0134943 & 54373.168 & -26.3444678 & 0.0000000\\
+% 		verbsee & 0.0037204 & 0.0134930 & 54371.450 & 0.2757288 & 0.7827574\\
+% 		verbthink & -0.3236483 & 0.0134936 & 54372.197 & -23.9853913 & 0.0000000\\
+% 		opc:verbbe\_annoyed & -0.2270099 & 0.0190632 & 54372.453 & -11.9082963 & 0.0000000\\
+% 		opm:verbbe\_annoyed & -0.1002424 & 0.0189726 & 54372.798 & -5.2835253 & 0.0000001\\
+% 		\addlinespace
+% 		opq:verbbe\_annoyed & -0.0762282 & 0.0192898 & 54372.017 & -3.9517423 & 0.0000777\\
+% 		opc:verbsuggest & -0.0675773 & 0.0190625 & 54371.851 & -3.5450396 & 0.0003929\\
+% 		opm:verbsuggest & 0.0763690 & 0.0189709 & 54371.211 & 4.0255864 & 0.0000569\\
+% 		opq:verbsuggest & -0.1499930 & 0.0192910 & 54373.069 & -7.7752981 & 0.0000000\\
+% 		opc:verbacknowledge & -0.0324529 & 0.0190617 & 54371.116 & -1.7025196 & 0.0886637\\
+% 		\addlinespace
+% 		opm:verbacknowledge & 0.0098129 & 0.0189711 & 54371.417 & 0.5172570 & 0.6049789\\
+% 		opq:verbacknowledge & -0.0089347 & 0.0192902 & 54372.396 & -0.4631724 & 0.6432426\\
+% 		opc:verbadmit & 0.0533251 & 0.0190632 & 54372.508 & 2.7972739 & 0.0051554\\
+% 		opm:verbadmit & 0.0808284 & 0.0189728 & 54372.916 & 4.2602367 & 0.0000205\\
+% 		opq:verbadmit & -0.0322232 & 0.0192908 & 54372.883 & -1.6703960 & 0.0948468\\
+% 		\addlinespace
+% 		opc:verbannounce & 0.0233831 & 0.0190642 & 54373.367 & 1.2265427 & 0.2199998\\
+% 		opm:verbannounce & 0.0630602 & 0.0189726 & 54372.754 & 3.3237576 & 0.0008887\\
+% 		opq:verbannounce & -0.0744750 & 0.0192926 & 54374.465 & -3.8602856 & 0.0001134\\
+% 		opc:verbbe\_right & -0.0845782 & 0.0190620 & 54371.360 & -4.4370099 & 0.0000091\\
+% 		opm:verbbe\_right & 0.1579976 & 0.0189719 & 54372.163 & 8.3279685 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbbe\_right & -0.1539128 & 0.0192901 & 54372.283 & -7.9788617 & 0.0000000\\
+% 		opc:verbconfess & 0.0841131 & 0.0190632 & 54372.437 & 4.4123417 & 0.0000102\\
+% 		opm:verbconfess & 0.0834691 & 0.0189719 & 54372.142 & 4.3996174 & 0.0000109\\
+% 		opq:verbconfess & -0.0368642 & 0.0192905 & 54372.632 & -1.9110088 & 0.0560087\\
+% 		opc:verbconfirm & 0.0693734 & 0.0190644 & 54373.529 & 3.6388992 & 0.0002741\\
+% 		\addlinespace
+% 		opm:verbconfirm & 0.2660910 & 0.0189728 & 54372.917 & 14.0248948 & 0.0000000\\
+% 		opq:verbconfirm & -0.0449413 & 0.0192914 & 54373.472 & -2.3295974 & 0.0198311\\
+% 		opc:verbdemonstrate & 0.0584268 & 0.0190633 & 54372.586 & 3.0648793 & 0.0021786\\
+% 		opm:verbdemonstrate & 0.1781393 & 0.0189714 & 54371.707 & 9.3898743 & 0.0000000\\
+% 		opq:verbdemonstrate & -0.0008729 & 0.0192906 & 54372.707 & -0.0452486 & 0.9639093\\
+% 		\addlinespace
+% 		opc:verbestablish & 0.0300718 & 0.0190618 & 54371.161 & 1.5775980 & 0.1146638\\
+% 		opm:verbestablish & 0.1877158 & 0.0189720 & 54372.212 & 9.8943713 & 0.0000000\\
+% 		opq:verbestablish & -0.0700678 & 0.0192899 & 54372.103 & -3.6323635 & 0.0002811\\
+% 		opc:verbhear & -0.0253789 & 0.0190637 & 54372.970 & -1.3312665 & 0.1831069\\
+% 		opm:verbhear & -0.1048737 & 0.0189721 & 54372.368 & -5.5277707 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbhear & -0.0006261 & 0.0192909 & 54372.991 & -0.0324581 & 0.9741069\\
+% 		opc:verbinform & -0.0991359 & 0.0190633 & 54372.605 & -5.2003420 & 0.0000002\\
+% 		opm:verbinform & -0.0860938 & 0.0189733 & 54373.368 & -4.5376355 & 0.0000057\\
+% 		opq:verbinform & -0.0174718 & 0.0192931 & 54374.843 & -0.9055994 & 0.3651520\\
+% 		opc:verbknow & -0.1127399 & 0.0190628 & 54372.127 & -5.9141311 & 0.0000000\\
+% 		\addlinespace
+% 		opm:verbknow & -0.0872392 & 0.0189719 & 54372.149 & -4.5983336 & 0.0000043\\
+% 		opq:verbknow & -0.0192684 & 0.0192895 & 54371.814 & -0.9989027 & 0.3178463\\
+% 		opc:verbpretend & -0.3172145 & 0.0190639 & 54373.134 & -16.6395086 & 0.0000000\\
+% 		opm:verbpretend & -0.1196669 & 0.0189725 & 54372.658 & -6.3073967 & 0.0000000\\
+% 		opq:verbpretend & -0.2959105 & 0.0192916 & 54373.627 & -15.3388172 & 0.0000000\\
+% 		\addlinespace
+% 		opc:verbprove & 0.0309486 & 0.0190626 & 54371.908 & 1.6235265 & 0.1044827\\
+% 		opm:verbprove & 0.2578102 & 0.0189728 & 54372.987 & 13.5883819 & 0.0000000\\
+% 		opq:verbprove & -0.0531478 & 0.0192907 & 54372.839 & -2.7550974 & 0.0058694\\
+% 		opc:verbreveal & 0.0632551 & 0.0190641 & 54373.234 & 3.3180321 & 0.0009071\\
+% 		opm:verbreveal & 0.0765052 & 0.0189711 & 54371.383 & 4.0327309 & 0.0000552\\
+% 		\addlinespace
+% 		opq:verbreveal & -0.0096396 & 0.0192917 & 54373.669 & -0.4996755 & 0.6173056\\
+% 		opc:verbsay & -0.0284748 & 0.0190633 & 54372.590 & -1.4936938 & 0.1352615\\
+% 		opm:verbsay & 0.0449550 & 0.0189713 & 54371.586 & 2.3696339 & 0.0178092\\
+% 		opq:verbsay & -0.2578020 & 0.0192911 & 54373.227 & -13.3637456 & 0.0000000\\
+% 		opc:verbsee & -0.0094907 & 0.0190633 & 54372.613 & -0.4978528 & 0.6185898\\
+% 		\addlinespace
+% 		opm:verbsee & -0.0408332 & 0.0189714 & 54371.728 & -2.1523480 & 0.0313743\\
+% 		opq:verbsee & 0.0345992 & 0.0192916 & 54373.623 & 1.7934827 & 0.0729013\\
+% 		opc:verbthink & -0.2246902 & 0.0190628 & 54372.159 & -11.7868151 & 0.0000000\\
+% 		opm:verbthink & -0.0856847 & 0.0189719 & 54372.175 & -4.5163904 & 0.0000063\\
+% 		opq:verbthink & -0.2500624 & 0.0192911 & 54373.178 & -12.9625865 & 0.0000000\\
+% 		\bottomrule
+% 	\end{tabular}
+% \end{table}
+
+% \newpage
+
+% \begin{table}[ht]
+% 	\caption{Model output w \texttt{know/n} as baseline}
+% 	\label{tab:discover-n}
+% 	\centering
+
+% 	\begin{tabular}{lrrrrr}
+% 		\toprule
+% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
+% 		\midrule
+% 		(Intercept) & 0.7921005 & 0.0114401 & 1479.182 & 69.2391632 & 0.0000000\\
+% 		opc & -0.0013151 & 0.0156638 & 26226.099 & -0.0839558 & 0.9330922\\
+% 		opm & -0.1434448 & 0.0155855 & 26286.727 & -9.2037503 & 0.0000000\\
+% 		opq & 0.0818736 & 0.0143093 & 56827.205 & 5.7217106 & 0.0000000\\
+% 		verbdiscover & -0.1115762 & 0.0134931 & 54371.661 & -8.2691029 & 0.0000000\\
+% 		\addlinespace
+% 		verbbe\_annoyed & 0.0748953 & 0.0134936 & 54372.286 & 5.5504198 & 0.0000000\\
+% 		verbsuggest & -0.5413729 & 0.0134928 & 54371.215 & -40.1230775 & 0.0000000\\
+% 		verbacknowledge & -0.1455464 & 0.0134940 & 54372.778 & -10.7859911 & 0.0000000\\
+% 		verbadmit & -0.2653476 & 0.0134936 & 54372.243 & -19.6647018 & 0.0000000\\
+% 		verbannounce & -0.2858550 & 0.0134935 & 54372.051 & -21.1847253 & 0.0000000\\
+% 		\addlinespace
+% 		verbbe\_right & -0.5889971 & 0.0134936 & 54372.234 & -43.6501264 & 0.0000000\\
+% 		verbconfess & -0.2932742 & 0.0134933 & 54371.896 & -21.7347532 & 0.0000000\\
+% 		verbconfirm & -0.5589410 & 0.0134943 & 54373.134 & -41.4204989 & 0.0000000\\
+% 		verbdemonstrate & -0.4558072 & 0.0134931 & 54371.605 & -33.7807521 & 0.0000000\\
+% 		verbestablish & -0.4793608 & 0.0134938 & 54372.501 & -35.5245129 & 0.0000000\\
+% 		\addlinespace
+% 		verbhear & -0.1207841 & 0.0134931 & 54371.608 & -8.9515418 & 0.0000000\\
+% 		verbinform & -0.0625474 & 0.0134940 & 54372.811 & -4.6351853 & 0.0000036\\
+% 		verbpretend & -0.4300580 & 0.0134939 & 54372.685 & -31.8704317 & 0.0000000\\
+% 		verbprove & -0.5615997 & 0.0134934 & 54372.014 & -41.6202578 & 0.0000000\\
+% 		verbreveal & -0.2308861 & 0.0134943 & 54373.179 & -17.1098402 & 0.0000000\\
+% 		\addlinespace
+% 		verbsay & -0.4670771 & 0.0134939 & 54372.578 & -34.6140392 & 0.0000000\\
+% 		verbsee & -0.1078558 & 0.0134932 & 54371.715 & -7.9933523 & 0.0000000\\
+% 		verbthink & -0.4352245 & 0.0134933 & 54371.896 & -32.2547883 & 0.0000000\\
+% 		opc:verbdiscover & 0.1127399 & 0.0190628 & 54372.122 & 5.9141311 & 0.0000000\\
+% 		opm:verbdiscover & 0.0872392 & 0.0189719 & 54372.145 & 4.5983336 & 0.0000043\\
+% 		\addlinespace
+% 		opq:verbdiscover & 0.0192684 & 0.0192895 & 54371.810 & 0.9989027 & 0.3178463\\
+% 		opc:verbbe\_annoyed & -0.1142699 & 0.0190629 & 54372.206 & -5.9943630 & 0.0000000\\
+% 		opm:verbbe\_annoyed & -0.0130032 & 0.0189709 & 54371.180 & -0.6854301 & 0.4930755\\
+% 		opq:verbbe\_annoyed & -0.0569598 & 0.0192897 & 54371.974 & -2.9528580 & 0.0031498\\
+% 		opc:verbsuggest & 0.0451626 & 0.0190614 & 54370.865 & 2.3693169 & 0.0178244\\
+% 		\addlinespace
+% 		opm:verbsuggest & 0.1636081 & 0.0189713 & 54371.561 & 8.6239932 & 0.0000000\\
+% 		opq:verbsuggest & -0.1307246 & 0.0192896 & 54371.847 & -6.7769511 & 0.0000000\\
+% 		opc:verbacknowledge & 0.0802870 & 0.0190639 & 54373.059 & 4.2114778 & 0.0000254\\
+% 		opm:verbacknowledge & 0.0970521 & 0.0189731 & 54373.215 & 5.1152465 & 0.0000003\\
+% 		opq:verbacknowledge & 0.0103337 & 0.0192900 & 54372.183 & 0.5357031 & 0.5921659\\
+% 		\addlinespace
+% 		opc:verbadmit & 0.1660650 & 0.0190633 & 54372.568 & 8.7112405 & 0.0000000\\
+% 		opm:verbadmit & 0.1680676 & 0.0189725 & 54372.674 & 8.8584875 & 0.0000000\\
+% 		opq:verbadmit & -0.0129548 & 0.0192897 & 54371.901 & -0.6715939 & 0.5018451\\
+% 		opc:verbannounce & 0.1361230 & 0.0190629 & 54372.249 & 7.1407107 & 0.0000000\\
+% 		opm:verbannounce & 0.1502994 & 0.0189716 & 54371.859 & 7.9223393 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbannounce & -0.0552066 & 0.0192897 & 54371.911 & -2.8619784 & 0.0042117\\
+% 		opc:verbbe\_right & 0.0281618 & 0.0190625 & 54371.883 & 1.4773357 & 0.1395915\\
+% 		opm:verbbe\_right & 0.2452367 & 0.0189720 & 54372.236 & 12.9262410 & 0.0000000\\
+% 		opq:verbbe\_right & -0.1346444 & 0.0192901 & 54372.308 & -6.9799743 & 0.0000000\\
+% 		opc:verbconfess & 0.1968531 & 0.0190627 & 54372.016 & 10.3266157 & 0.0000000\\
+% 		\addlinespace
+% 		opm:verbconfess & 0.1707082 & 0.0189737 & 54373.731 & 8.9971020 & 0.0000000\\
+% 		opq:verbconfess & -0.0175959 & 0.0192891 & 54371.441 & -0.9122160 & 0.3616590\\
+% 		opc:verbconfirm & 0.1821133 & 0.0190626 & 54371.955 & 9.5534239 & 0.0000000\\
+% 		opm:verbconfirm & 0.3533301 & 0.0189718 & 54372.046 & 18.6239655 & 0.0000000\\
+% 		opq:verbconfirm & -0.0256729 & 0.0192917 & 54373.707 & -1.3307729 & 0.1832693\\
+% 		\addlinespace
+% 		opc:verbdemonstrate & 0.1711667 & 0.0190647 & 54373.791 & 8.9782020 & 0.0000000\\
+% 		opm:verbdemonstrate & 0.2653784 & 0.0189716 & 54371.911 & 13.9881581 & 0.0000000\\
+% 		opq:verbdemonstrate & 0.0183955 & 0.0192903 & 54372.499 & 0.9536136 & 0.3402835\\
+% 		opc:verbestablish & 0.1428117 & 0.0190634 & 54372.667 & 7.4914038 & 0.0000000\\
+% 		opm:verbestablish & 0.2749549 & 0.0189719 & 54372.156 & 14.4927310 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbestablish & -0.0507994 & 0.0192892 & 54371.529 & -2.6335636 & 0.0084518\\
+% 		opc:verbhear & 0.0873610 & 0.0190622 & 54371.557 & 4.5829467 & 0.0000046\\
+% 		opm:verbhear & -0.0176345 & 0.0189709 & 54371.193 & -0.9295579 & 0.3526042\\
+% 		opq:verbhear & 0.0186422 & 0.0192902 & 54372.432 & 0.9664076 & 0.3338446\\
+% 		opc:verbinform & 0.0136041 & 0.0190628 & 54372.117 & 0.7136441 & 0.4754504\\
+% 		\addlinespace
+% 		opm:verbinform & 0.0011454 & 0.0189738 & 54373.859 & 0.0603655 & 0.9518647\\
+% 		opq:verbinform & 0.0017966 & 0.0192921 & 54374.044 & 0.0931259 & 0.9258039\\
+% 		opc:verbpretend & -0.2044746 & 0.0190640 & 54373.227 & -10.7256652 & 0.0000000\\
+% 		opm:verbpretend & -0.0324277 & 0.0189726 & 54372.773 & -1.7091875 & 0.0874219\\
+% 		opq:verbpretend & -0.2766421 & 0.0192909 & 54373.044 & -14.3405213 & 0.0000000\\
+% 		\addlinespace
+% 		opc:verbprove & 0.1436885 & 0.0190622 & 54371.604 & 7.5378609 & 0.0000000\\
+% 		opm:verbprove & 0.3450493 & 0.0189712 & 54371.500 & 18.1880552 & 0.0000000\\
+% 		opq:verbprove & -0.0338794 & 0.0192909 & 54373.040 & -1.7562330 & 0.0790543\\
+% 		opc:verbreveal & 0.1759951 & 0.0190637 & 54372.960 & 9.2319255 & 0.0000000\\
+% 		opm:verbreveal & 0.1637444 & 0.0189722 & 54372.374 & 8.6307727 & 0.0000000\\
+% 		\addlinespace
+% 		opq:verbreveal & 0.0096288 & 0.0192902 & 54372.406 & 0.4991553 & 0.6176720\\
+% 		opc:verbsay & 0.0842652 & 0.0190632 & 54372.497 & 4.4203002 & 0.0000099\\
+% 		opm:verbsay & 0.1321942 & 0.0189719 & 54372.094 & 6.9679113 & 0.0000000\\
+% 		opq:verbsay & -0.2385336 & 0.0192898 & 54371.999 & -12.3658155 & 0.0000000\\
+% 		opc:verbsee & 0.1032492 & 0.0190630 & 54372.324 & 5.4162003 & 0.0000001\\
+% 		\addlinespace
+% 		opm:verbsee & 0.0464060 & 0.0189718 & 54372.089 & 2.4460454 & 0.0144464\\
+% 		opq:verbsee & 0.0538675 & 0.0192901 & 54372.298 & 2.7924979 & 0.0052321\\
+% 		opc:verbthink & -0.1119503 & 0.0190627 & 54372.030 & -5.8727378 & 0.0000000\\
+% 		opm:verbthink & 0.0015545 & 0.0189713 & 54371.557 & 0.0819391 & 0.9346955\\
+% 		opq:verbthink & -0.2307940 & 0.0192903 & 54372.447 & -11.9642778 & 0.0000000\\
+% 		\bottomrule
+% 	\end{tabular}
+% \end{table}
 

### 2022-12-09 results/main/meta-analyses/projectivity/rscripts/analysis.R

@@ -66,3 +66,18 @@ summary2 <- summary(glmm2)
 print(summary2, cor=F, dig=3)
 kable(summary2$coefficients, format = "latex", booktabs = TRUE)
 
+
+### dummy coding with "know" as baseline
+data$verb <- relevel(data$verb, ref = "know")
+contrasts(data$verb)
+
+# glmm3 <- lmer(projective ~ op * verb + (1 | workerid) +
+#                  (1 | content), data=data)
+
+# save.image("linear-models.RData")
+load("linear-models.RData")
+
+summary3 <- summary(glmm3)
+print(summary3, cor=F, dig=3)
+kable(summary3$coefficients, format = "latex", booktabs = TRUE)
+

### 2022-12-09 results/main/meta-analyses/projectivity/rscripts/graphs.R

@@ -12,7 +12,8 @@ data <- read.csv("../data/data_combined.csv", header = TRUE, sep = ",")
 source('../../../helpers.R')
 
 # libraries for manipulating dataframes, and plotting
-library(tidyverse)
+library(dplyr)
+library(forcats)
 library(ggplot2)
 
 # projectivity by operator -----
@@ -79,21 +80,21 @@ proj_means = data %>% group_by(verb, op) %>%
 proj_means <- mutate(proj_means, verb = fct_reorder(verb, Mean, .fun = mean))
 proj_means <- mutate(proj_means, op = fct_reorder(op, Mean, .fun = mean))
 
-ggplot(proj_means, aes(x=verb, y=Mean, group = op, color = op)) +
+proj_means %>% 
+  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean, group = op, color = op)) +
   coord_cartesian(ylim=c(0,1)) +
   geom_point(aes(shape = op), size = 3) + 
   scale_shape_manual(values = c("M", "N", "Q", "C")) +
   geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
   geom_line() +
   scale_y_continuous(limits = c(0,1)) +
-  ylab("Mean certainty rating") +
-  xlab("Predicate") +
-  guides(shape = "none") +
-  scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
-  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) +
-  labs(title = "Mean certainty for predicate by operator", color = "Operator")
+  ylab("") +
+  xlab("") +
+  theme_bw() +
+  theme(legend.position = "none") +
+  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1))
 
-ggsave("../graphs/proj-by-both.pdf",height=8,width=14)
+ggsave("../graphs/proj-by-both.pdf",height=4.7,width=9)
 
 
 
@@ -171,4 +172,148 @@ ggplot(pmeans, aes(x=verb, y=Mean)) +
   ylab("Mean certainty rating") +
   xlab("Predicate")
 ggsave(f="../graphs/projectivity-verb-participant.pdf",height=6,width=8)
+
+
+# VERB PROFILES ---- 
+
+proj_means %>%
+  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb)) +
[…]

### 2022-12-09 SALT-abstract/SALT-abstract.tex
updates

@@
 \noindent 
-	We present experimental evidence that i) the projection of the content of the clausal complement of attitude predicates varies across entailment-canceling operators (negation, question, modal, conditional) and ii) this by-operator variation differs across attitude predicates.
-Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}, though see \citealp{beaver_have_2010}).
+	We present experimental evidence that \textbf{(i)} the projection of the content of attitude complements varies across entailment-canceling operators (negation, question, modal, conditional) and \textbf{(ii)} this by-operator variation differs across attitude predicates.
+Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
 Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.
@@
 
-	There has been, to date, one investigation of by-operator variation: \citealt{smith_relationship_2014}, who investigated by-operator variation (negation, conditional) for the projective content of six expressions (\emph{know, the, win}, epithets, clefs, non-restrictive relative clauses (NRRCs)), observed by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefs) and that this by-operator variation differs across contents (e.g., the content of NRRCs was more projective under conditionals than negation, the opposite pattern was observed for \emph{win}).
-It is not clear, however, whether the response task used by \citealt{smith_relationship_2014} measured projection, as participants were asked to rate how surprised they would be to learn the content under investigation after observing the utterance.
+	There has been, to date, one investigation of by-operator variation: \citet{smith_relationship_2014}, who investigated by-operator variation (negation, conditional) for projective content, observed by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefts) and that this by-operator variation differs across contents.
+E.g., the content of NRRCs was more projective under conditionals than negation, the opposite pattern was observed for \emph{know, win} and epithets.
+However, the response task used asked participants to rate how surprised they would be to learn the content under investigation after observing the utterance.
+Since many factors other than speaker commitments may influence surprisal, it is not clear, whether they (only) measured projection.
 
 	There has not yet been an experimental investigation of by-operator variation between factive and semi-factive predicates.
-\citealt{djarv_cognitive_2018} and \citealt{tonhauser_how_2018} did, however, observe by-predicate projection variation under polar questions.
-\citealt{djarv_cognitive_2018} observed a difference between \emph{be happy} and \emph{appreciate} (which they assumed are factive predicates) and \emph{be aware} and \emph{realize} (which they assumed are semi-factive predicates).
-However, the response task (acceptability of an affirmation of the CC while the main clause content was denied) did not measure projection of the CC.
+\citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} did, however, observe by-predicate projection variation in polar questions.
+\citet{djarv_cognitive_2018} found a difference between \emph{be happy} and \emph{appreciate} (which they assumed are factive predicates) and \emph{be aware} and \emph{realize} (which they assumed are semi-factive predicates).
+However, the response task (acceptability of an affirmation of the main clause while the CC was denied) did not measure projection of the CC.
 \citealt{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates from under polar questions: The by-predicate projection differences they observed did not align with what would be expected from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more projective than that of semi-factive \emph{discover}).

### 2022-12-09 SALT-abstract/SALT-abstract.tex
updated SALT abstract

@@
 
+\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{\#1}}
+\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
+\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{\#1})}
+
 \parindent=2ex
@@
 
-\vspace*{-.6\baselineskip}
+\vspace*{-.8\baselineskip}
 \noindent 
-	We present experimental evidence that \textbf{(i)} the projection of the content of attitude complements varies across entailment-canceling operators (negation, question, modal, conditional) and \textbf{(ii)} this by-operator variation differs across attitude predicates.
+	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements varies across four entailment-canceling operators (negation, question, modal, antecedent of conditional) and \textbf{(ii)} that this by-operator variation differs across attitude predicates.
 Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
@@
 
-	There has been, to date, one investigation of by-operator variation: \citet{smith_relationship_2014}, who investigated by-operator variation (negation, conditional) for projective content, observed by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefts) and that this by-operator variation differs across contents.
-E.g., the content of NRRCs was more projective under conditionals than negation, the opposite pattern was observed for \emph{know, win} and epithets.
-However, the response task used asked participants to rate how surprised they would be to learn the content under investigation after observing the utterance.
-Since many factors other than speaker commitments may influence surprisal, it is not clear, whether they (only) measured projection.
+	There has been, to date, one investigation of by-operator variation: \citet{smith_relationship_2014}, who investigated by-operator projection variation for negation and conditional embeddings, observed that there is by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefts) and that this by-operator variation differs across contents: The content of NRRCs and \emph{win} was more projective under conditionals than negation, the opposite pattern was observed for \emph{know} and epithets.
+However, the response task they used asked participants to rate how surprised they would be to learn the content under investigation after observing the utterance.
+Since factors other than speaker commitment modulate surprisal, it is not clear that they (only) measured projection.
 
@@
 {\bf Results.}
-	Fig.~\ref{fig:figure1} shows the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
+	Fig.\ref{fig:figure1} shows the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
 We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than from under conditionals and modals, the CC of \emph{know} projects less from under negation than under questions, but more from under negation than from under modals, and the CC of \emph{discover} projects less from negation than from under conditionals and questions, and more from under negation than modals.
 These results (which are supported by linear mixed effects models, see Table \ref{t:models}) are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
-For {\em know}, we observe a third pattern -- one that is not aligned with a classification as a factive nor as a semi-factive predicate.
+For {\em know}, we observe a third pattern, one that is neither that of a factive nor that of a semi-factive predicate.
+Similar considerations apply to \emph{see} and \emph{reveal}, which are also considered (semi-)factive in the literature.
 
@@
 \noindent
-{\bf Discussion and Conclusion.} 
-	Our results clearly show that different entailment-cancelling operators affect projection in different ways, and that the effect of entailment-cancelling operator on projection differs for different verbs.
-The picture that emerges could not be characterized in terms of Karttunen's characterization of (semi)factives.
-Our results indicate a need for a more nuanced view of the family-of-sentences diagnostics of projection, as well as a better understanding of the effect of the interaction between lexical semantics and operators on projection.
+{\bf Discussion:  Empirical and theoretical implications.} 
+Our results --- that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection --- have several empirical and theoretical implications.
+First, the results for the predicates typically considered (semi-)factive, namely \emph{be annoyed, know, see, discover} and \emph{reveal}, call into question the assumed distinction between factive and semi-factive predicates (see also \citealt{beaver_have_2010}).
+Future research that appeals to these two categories must clarify their definition.
+Second, claims about projection variability must be relativized to the entailment-canceling operator.
+While our data replicate the result from \citealt{tonhauser_how_2018} that, in a polar question, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals.
+Finally, our results provide further support (from embedding under negation, modals, and conditionals) for the result of \citealt{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
 
-	Further, we were able to replicate the results of \citealp{degen_are_2022}), observing that projection varies by predicate (e.g., the CC of \emph{be annoyed} is more projective than that of \emph{discover}, and finding no categorical distinction between factive and non-factive predicates either.
+\noindent
+{\bf Discussion: Novel research question.} Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
+This is a pressing question for future research, to which our data offer some tentative answers.
+We identify four major patterns.
+The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of Fig.\ref{fig:figure2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates.
+For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in panel (b), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
+The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in panel (c): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not operators), to result in lower projection ratings under negation.
+Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in panel (d).
+The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
 
 	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
-\textbf{Figure \ref{fig:figure2}} gives the mean certainty ratings for the four operators by verb, for some groups of our verbs that group together in the way that they show by-operator variation.
+Fig.~\ref{fig:figure2} gives the mean certainty ratings for the four operators by verb, for some groups of our verbs that group together in the way that they show by-operator variation.
 We highlight four \lq projection-profiles\rq here: the verbs \emph{pretend, think} are the only verbs that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
@@
 			\midrule
-			\#1 & Intercept: \emph{be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
+			\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
 			& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
@@
 			\midrule
-			\#2 & Intercept: \emph{know}/negation & 0.79 & 0.01 &  69.24 & *** \\
+			\#2 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 &  69.24 & *** \\
 			&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
@@
 			\midrule
-			\#3 & Intercept: \emph{discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
+			\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
 			& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
@@
 	\vspace{-\baselineskip}
-	\caption{Mean certainty-ratings by operator for each verb, showing different \lq verb-profiles\rq.}
+	\caption{Mean certainty-ratings by operator for four predicate patterns in our data.}
 	\label{fig:figure2}

### 2022-12-09 SALT-abstract/SALT-abstract.tex
abstract final

@@
 	\newcommand{\posscite}[1]{\citeauthor{#1}'s (\citeyear{#1})}
+	\renewcommand*{\refname}{\normalsize\textbf{References}\\ \vspace{-.5\baselineskip}}
+	\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{\#1}}
+	\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
+	\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{\#1})}
 
@@
 
-%% bibliography
-\renewcommand*{\refname}{\normalsize\textbf{References}\\ \vspace{-.5\baselineskip}}
-
 %% fonts
@@
 	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Times New Roman}
+	\setmonofont[Scale=MatchLowercase]{Andale Mono}
 
-\usepackage{setspace}
-
-\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{\#1}}
-\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
-\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{\#1})}
-
+% spacing
 	\parindent=2ex
@@
 \begin{document}
+	\setlength{\Exlabelsep}{.7em}
+	% \setlength{\Exindent}{0em}
 
@@
 
-\vspace*{-.8\baselineskip}
+\vspace{-.6\baselineskip}
 \noindent 
-	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements varies across four entailment-canceling operators (negation, question, modal, antecedent of conditional) and \textbf{(ii)} that this by-operator variation differs across attitude predicates.
-Our results do not align with the long-standing distinction between factive and semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
-Instead, the observed by-operator variation groups predicates by lexical semantic/pragmatic properties that raise important questions for future research on projection.
+	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs across attitude predicates.
+Our results do not align with the long-standing distinction of factive vs.
+semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
+	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragma-tic properties which raises important questions for future research on projection.
 
 \noindent
-	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters one of the attitude ascriptions in  \ref{ex:family} is committed to the truth of the content of the complement (CC),
-	% that Julian dances salsa,
+	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the truth of the content of the complement (CC),
 	even though it occurs under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, a modal \ref{ex:mod}, or a conditional \ref{ex:cond}.
-
-	%	Certain attitude ascriptions come with an inference to the truth of their complement, even if embedded under entailment-cancelling operators (shown for \emph{discover} in \ref{ex:family}), in which case the inference is said to \emph{project} (e.g.
-\citealp{karttunen_observations_1971}).
-
-	\vspace{-.3\baselineskip}
+	\vspace{-.5\baselineskip}
 	\ex. \label{ex:family}
 		\a. \label{ex:neg}
-			Negation: \hfill
+			{\bf Negation:} \hfill
 			\emph{\lq Cole didn't discover that Julian dances salsa.\rq}
 		\b. \label{ex:q}
-			Polar Question: \hfill
+			{\bf Polar Question:} \hfill
 			\emph{\lq Did Cole discover that Julian dances salsa?\rq}
 		\c. \label{ex:mod}
-			Modal: \hfill
+			{\bf Modal:} \hfill
 			\emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
 		\d. \label{ex:cond}
-			Conditional: \hfill
+			{\bf Conditional:} \hfill
 			\emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
@@
 \vspace{-.4\baselineskip}
-	% Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}.
-In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.
-	%
 	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g.
-\emph{regret, forget}) projects from under all four operators, whereas that of semi-factive predicates (e.g.
+\emph{regret, forget}) projects across all four operators, whereas that of semi-factive predicates (e.g.
 \emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
+	To date, systematic investigations of Karttunen's intuition are scarce.
 
-	There has been, to date, one investigation of by-operator variation: \citet{smith_relationship_2014}, who investigated by-operator projection variation for negation and conditional embeddings, observed that there is by-operator variation for some contents (e.g., that of \emph{know}, but not that of clefts) and that this by-operator variation differs across contents: The content of NRRCs and \emph{win} was more projective under conditionals than negation, the opposite pattern was observed for \emph{know} and epithets.
-However, the response task they used asked participants to rate how surprised they would be to learn the content under investigation after observing the utterance.
-Since factors other than speaker commitment modulate surprisal, it is not clear that they (only) measured projection.
-
-	There has not yet been an experimental investigation of by-operator variation between factive and semi-factive predicates.
-\citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} did, however, observe by-predicate projection variation in polar questions.
-\citet{djarv_cognitive_2018} found a difference between \emph{be happy} and \emph{appreciate} (which they assumed are factive predicates) and \emph{be aware} and \emph{realize} (which they assumed are semi-factive predicates).
-However, the response task (acceptability of an affirmation of the main clause while the CC was denied) did not measure projection of the CC.
-\citealt{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates from under polar questions: The by-predicate projection differences they observed did not align with what would be expected from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more projective than that of semi-factive \emph{discover}).
+	\citet{smith_relationship_2014} investigated by-operator variation (negation, conditional) for various types of projective content, observing that the content of non-restrictive relative clauses and \emph{win} was more projective under conditionals than negation, \emph{know} and epithets exhibited the opposite pattern, whereas clefts showed no difference.
+However, their task asked participants how surprised they would be to learn the content under investigation after observing the utterance.
+Since factors other than speaker commitment modulate (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is not clear that they (only) measured projection.
+So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
+However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
+\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
+Here, it is also not obvious how exactly this task relates to projection.
+\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, by collecting ratings about speaker certainty about the CC.
+The observed by-predicate differences did not align the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).
 
 \noindent
-	{\bf Experiment.} We present the results of an experiment designed to investigate by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
-Projection was measured for the same contents across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see also, e.g., \citealt{tonhauser_how_2018,djarv_prosodic_2017, mahler2020}).
-
-	%In our work, we analyze measures from a task designed to measure speaker commitment more directly to address the questions: \textbf{(i)} Is projection affected by differences in entailment-canceling environments?
-\textbf{(ii)} Do these effects vary for different triggers (and in what way)?
-We used a response task to elicit judgments about how strongly a speaker would be committed towards the embedded clause \citep[from][]{tonhauser_prosodic_2016}.
-We presented sentences like in \ref{ex:family} as asserted by a named speaker (e.g.
-“\textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}”).
-Participants then provided a certainty-rating in response to a prompt like: \emph{“Is Daniel certain that Julian dances salsa?”}, by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).
+{\bf Experiment.}
+	We present the results of a study investigating by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
+Projection was measured for the same contents across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}).
 
@@
 \noindent
-{\bf Methods and expectations.}
+{\bf Methods and Expectations.}
 	Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The attitude predicates were embedded under polar questions in Exps.~1, under negation in Exps.~2, under {\em perhaps} in Exps.~3, and in the antecedent of a conditional in Exps.~4.
@@
 We focus on the projection ratings here.) In each experiment, participants were asked to read utterances like those in \ref{ex:family} and judge whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
-Participants gave their response on a slider marked `no' (coded as 0) at one and and `yes' (coded as 1) on the other end.
+Participants gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
 Each participant rated the projection of the CC of all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
 We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
-We expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and possibly lower ratings under the other operators.
-
-	%We compared certainty-ratings for the four entailment-canceling operators in \ref{ex:family}, and 20 clause-embedding predicates (\texttt{verb}: {\em be annoyed, discover, know, reveal, see, acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, pretend, suggest, say, think, be right, demonstrate}).
-	%	Based on the Karttunen-Djärv generalization about emotive factives vs.
-cognitive semi-factives, we would expect emotive factives (\emph{be annoyed}), and verbs normally taken to be factive (\emph{know}), to be highly projective in a way that is indifferent to the embedding context, and cognitive semi-factives (\emph{discover, see}) to show higher projectivity ratings under negation compared to questions and conditionals.
-
-	%The study, originally designed to address a different research question, consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause.
-Experiments 1--3 used polar question embedding, exps.
-4--6 used negation, 7--9 modals, and 10--12 conditionals, making the \texttt{operator} manipulation a between-subjects factor.
-
-	%We analyzed data from 2682 self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk).
-Participants saw items Latin-squared and randomized with six control stimuli.
-
+Karttunen's generalization leads us to expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and possibly lower ratings under the other operators.
 	% paragraph method (end)
@@
 {\bf Results.}
-	Fig.\ref{fig:figure1} shows the mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
-We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than from under conditionals and modals, the CC of \emph{know} projects less from under negation than under questions, but more from under negation than from under modals, and the CC of \emph{discover} projects less from negation than from under conditionals and questions, and more from under negation than modals.
-These results (which are supported by linear mixed effects models, see Table \ref{t:models}) are not aligned with the distinction between factive and semi-factive predicates proposed in \citealt{karttunen_observations_1971} or assumed in \citealt{djarv_cognitive_2018}: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
-For {\em know}, we observe a third pattern, one that is neither that of a factive nor that of a semi-factive predicate.
+	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
+We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+	%
+	These results (which are supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}) are unexpected 
+	based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
+The pattern observed for {\em know} is neither that of a factive nor that of a semi-factive predicate.
 Similar considerations apply to \emph{see} and \emph{reveal}, which are also considered (semi-)factive in the literature.
@@
 	Our results --- that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection --- have several empirical and theoretical implications.
-First, the results for the predicates typically considered (semi-)factive, namely \emph{be annoyed, know, see, discover} and \emph{reveal}, call into question the assumed distinction between factive and semi-factive predicates (see also \citealt{beaver_have_2010}).
+First, the results for the predicates typically considered (semi-)factive, namely \emph{be annoyed, know, see, discover} and \emph{reveal}, call into question the assumed distinction between factive and semi-factive predicates (see also \cite{beaver_have_2010}).
 Future research that appeals to these two categories must clarify their definition.
 Second, claims about projection variability must be relativized to the entailment-canceling operator.
-While our data replicate the result from \citealt{tonhauser_how_2018} that, in a polar question, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals.
-Finally, our results provide further support (from embedding under negation, modals, and conditionals) for the result of \citealt{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
+While our data replicate the result from \citet{tonhauser_how_2018} that, in a polar question, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals.
+Finally, our results provide further support (from embedding under negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
 
 \noindent
-{\bf Discussion: Novel research question.} Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
-This is a pressing question for future research, to which our data offer some tentative answers.
-We identify four major patterns.
-The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of Fig.\ref{fig:figure2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates.
-For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in panel (b), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
-The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in panel (c): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not operators), to result in lower projection ratings under negation.
-Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in panel (d).
+{\bf Discussion: Novel research question.}
+	Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
+	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
+	%
+	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates.
+	%
+	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
+	%
+	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in panel (c), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
+	%
+	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in (d).
 The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
 
-	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
-Fig.~\ref{fig:figure2} gives the mean certainty ratings for the four operators by verb, for some groups of our verbs that group together in the way that they show by-operator variation.
-We highlight four \lq projection-profiles\rq here: the verbs \emph{pretend, think} are the only verbs that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
+	% In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
+\textbf{Fig.~\ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, identifying some groups of predicates that show similar by-operator variation.
+We highlight four \lq projection-profiles\rq here: \emph{pretend} and \emph{think} are the only predicates that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
 \emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}).
@@
 
-	The picture that emerges is complex and requires a nuanced view of lexical effects on projection.
-I raises the question of how lexical differences between verbs could explain why certain verbs project more from under certain operators and less from others.
-While much previous work on this question has based their generalizations on impressionistic judgments, empirically establishing different projection-profiles associated with verbs therefore provides a basis for future research on lexical effects on projection.
+% \noindent
+% {\bf Conclusion.}
+% 	The picture that emerges is complex and requires a nuanced view of lexical effects on projection.
+It raises the question of how lexical differences between predicates could explain why the CC projects more from under certain operators and less from others.
+Much previous work on this question has based its generalizations on impressionistic judgments, empirically establishing different projection-profiles associated with verbs therefore provides a basis for future research on lexical effects on projection.
 
 \newpage
+{\scriptsize
+\noindent {\bf (Selected) References:}
+%
+{\bf Beaver (2010).} Have you noticed that your belly button lint colour is related to the colour of your clothing? {\em Presuppositions and Discourse: Essays offered to Hans Kamp.}~\textbullet~
+{\bf Degen \& Tonhauser (2022).} Are there factive predicates? An empirical investigation. {\em Language.}~\textbullet~
+{\bf Djärv \& Bacovcin (2017).} Prosodic effects on factive presupposition projection. {\em Semantics and Linguistic Theory.}~\textbullet~
+{\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
+{\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
+{\bf Karttunen (1971).} Some observations on factivity. {\em Research on Language \& Social Interaction.} ~\textbullet~
+{\bf Mahler (2020).} The social component of projection behavior of clausal complements. {\em Linguistic Society of America.}~\textbullet~
+{\bf Smith and Hall (2014).} The relationship between projection and embedding environment. {\em Proceedings of the 48th Meeting of the Chicago Linguistics Society.}~\textbullet~
+{\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}~\textbullet~
+{\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
 
 \begin{figure}[h]
-	% \vspace{-\baselineskip}
+	\vspace{-.8\baselineskip}
 	\centering
@@
 	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
-Embedding operator coded by letter and color:  \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions.}
+Embedding operator coded by letter and color:  \texttt{N} (orange): negation, \texttt{M} (black): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (blue): polar questions.}
 	\label{fig:figure1}
@@
 \begin{table}[h]
-	\centering
-	\vspace{-.7\baselineskip}
+	\vspace{-1.2\baselineskip}\phantom{.}
+	\hspace{-1.2em}
+	\begin{tabular}{p{.7\linewidth} p{.02\linewidth} p{.25\linewidth}}
 		\scalebox{.85}{
@@
 		}
-	\vspace{-.3\baselineskip}
-	\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC.
-Models were fit with \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}.}\label{t:models}
+		&&
+		\parbox{\linewidth}{\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC.
+Models were fit with \texttt{lme4, lmertest} in \texttt{R}.}\label{t:models}}
+
+	\end{tabular}
 \end{table}
 
-\begin{figure}[h]
-	\hspace{-2em}
-	\begin{minipage}{1.1\textwidth}\centering
-		\includegraphics[width=.24\linewidth]{graphs/profile1.pdf}
-		\includegraphics[width=.24\linewidth]{graphs/profile2.pdf}
-		% \includegraphics[width=.32\linewidth]{graphs/profile3.pdf}\\
-		\includegraphics[width=.24\linewidth]{graphs/profile4.pdf}
-		\includegraphics[width=.24\linewidth]{graphs/profile5.pdf}
+\hspace{-3.3em}\begin{minipage}{\textwidth}\centering
+	\begin{tabular}{p{.235\linewidth} p{.235\linewidth} p{.235\linewidth} p{.235\linewidth}}
+		\parbox{\linewidth}{\centering (a) \textbf{Negation high}\\
+		\includegraphics[width=.26\textwidth]{graphs/profile1.pdf}}
+		&
+		\parbox{\linewidth}{\centering (b) \textbf{Negation low}\\
+		\includegraphics[width=.26\textwidth]{graphs/profile4.pdf}}
+		&
+		\parbox{\linewidth}{\centering (c) \textbf{Conditional high}\\
+		\includegraphics[width=.26\textwidth]{graphs/profile2.pdf}}
+		&
+		\parbox{\linewidth}{\centering (d) \textbf{Modal low}\\
+		\includegraphics[width=.26\textwidth]{graphs/profile5.pdf}}
+	\end{tabular}
 \end{minipage}
-	\vspace{-\baselineskip}
-	\caption{Mean certainty-ratings by operator for four predicate patterns in our data.}
-	\label{fig:figure2}
-\end{figure}
+
+\normalsize \noindent Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
+
+% \begin{figure}[h]
+% 	\hspace{-2.5em}
+
+% 	\vspace{-\baselineskip}
+% 	% \caption{Mean certainty-ratings by operator for four predicate patterns in our data.}
+% 	\label{fig:figure2}
+% \end{figure}
 

### 2022-12-09 results/main/meta-analyses/projectivity/rscripts/graphs.R

@@ -80,15 +80,21 @@ proj_means = data %>% group_by(verb, op) %>%
 proj_means <- mutate(proj_means, verb = fct_reorder(verb, Mean, .fun = mean))
 proj_means <- mutate(proj_means, op = fct_reorder(op, Mean, .fun = mean))
 
+# Color blind friendly palette with black (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)):
+cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
 proj_means %>% 
   ggplot(aes(x=fct_reorder(verb, Mean), y=Mean, group = op, color = op)) +
   coord_cartesian(ylim=c(0,1)) +
   geom_point(aes(shape = op), size = 3) + 
   scale_shape_manual(values = c("M", "N", "Q", "C")) +
+  scale_colour_manual(values=cbbPalette) +
   geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
   geom_line() +
-  scale_y_continuous(limits = c(0,1)) +
-  ylab("") +
+  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1), 
+                     labels = c("0","0.2","0.4","0.6","0.8","1"), 
+                     name = "Mean certainty rating") +
+  # ylab("") +
   xlab("") +
   theme_bw() +
   theme(legend.position = "none") +
@@ -97,7 +103,7 @@ proj_means %>%
 ggsave("../graphs/proj-by-both.pdf",height=4.7,width=9)
 
 
-
+# other plots ----
 proj_means %>% mutate(op = fct_reorder(op, Mean, 
                                           .fun = mean)) %>% 
   mutate(verb = fct_reorder(verb, Mean, 
@@ -174,6 +180,7 @@ ggplot(pmeans, aes(x=verb, y=Mean)) +
 ggsave(f="../graphs/projectivity-verb-participant.pdf",height=6,width=8)
 
 
+
 # VERB PROFILES ---- 
 
 proj_means %>%
@@ -233,64 +240,111 @@ proj_means %>%
   theme_bw()
 ggsave(f="../graphs/profiles-grouped.pdf",height=6,width=8)
 
-# group 1 
+# individual profiles ----
+
+# Color blind friendly palette with black (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)):
[…]

### 2022-12-09 SALT-abstract/SALT-abstract.tex
SALT abstract final final

@@
 \noindent 
-	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs across attitude predicates.
+	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
 Our results do not align with the long-standing distinction of factive vs.
@@
 \emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
-	To date, systematic investigations of Karttunen's intuition are scarce.
+	To date, systematic empirical investigations of Karttunen's intuition are scarce.
 
@@
 	%
-	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
+	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
 	%
-	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in panel (c), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
+	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in (c), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
 	%
@@
 {\scriptsize
-\noindent {\bf (Selected) References:}
+\noindent {\bf References:}
 %
@@
 {\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
+{\bf Zimmermann (2011).} {\em The grammatical expression of focus in West Chadic: Variation and uniformity in and across languages.} Walter de Gruyter.
 
 \begin{figure}[h]
-	\vspace{-.8\baselineskip}
+	\vspace{-.6\baselineskip}
 	\centering

### 2022-12-09 SALT-abstract/SALT-abstract.tex
SALT submitted

@@
 
-\documentclass[12pt, a4paper]{article}
+\documentclass[12pt, letterpaper]{article}
 
@@
 \begin{document}
-	\setlength{\Exlabelsep}{.7em}
-	% \setlength{\Exindent}{0em}
 
@@
 
-\vspace{-.6\baselineskip}
+\vspace{-.7\baselineskip}
 \noindent 
@@
 semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
-	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragma-tic properties which raises important questions for future research on projection.
+	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragmatic properties which raises important questions for future research on projection.
 
 \noindent
-	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the truth of the content of the complement (CC),
-	even though it occurs under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, a modal \ref{ex:mod}, or a conditional \ref{ex:cond}.
-	\vspace{-.5\baselineskip}
+	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
+	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}.
+	\vspace{-.6\baselineskip}
 	\ex. \label{ex:family}
@@
 
-\vspace{-.4\baselineskip}
+\vspace{-.6\baselineskip}
 	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g.
@@
 
-	\citet{smith_relationship_2014} investigated by-operator variation (negation, conditional) for various types of projective content, observing that the content of non-restrictive relative clauses and \emph{win} was more projective under conditionals than negation, \emph{know} and epithets exhibited the opposite pattern, whereas clefts showed no difference.
-However, their task asked participants how surprised they would be to learn the content under investigation after observing the utterance.
-Since factors other than speaker commitment modulate (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is not clear that they (only) measured projection.
+	\citet{smith_relationship_2014} investigated by-operator variation (negation, conditional) for various types of projective content, observing that the content of non-restrictive relative clauses and \emph{win} was more projective under conditionals than negation, while \emph{know} and epithets exhibited the opposite pattern.
+	% , whereas clefts showed no difference.
+	Their task asked participants how surprised they would be to learn the content under investigation after observing the utterance.
+Since other factors besides speaker commitment modulate (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.
 So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
@@
 Here, it is also not obvious how exactly this task relates to projection.
-\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, by collecting ratings about speaker certainty about the CC.
-The observed by-predicate differences did not align the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).
+\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
+The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).
 
@@
 {\bf Experiment.}
-	We present the results of a study investigating by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
-Projection was measured for the same contents across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}).
+	We present results of a study investigating by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
+Projection was measured  across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}).
 
@@
 {\bf Methods and Expectations.}
-	Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The attitude predicates were embedded under polar questions in Exps.~1, under negation in Exps.~2, under {\em perhaps} in Exps.~3, and in the antecedent of a conditional in Exps.~4.
-(Each set of experiments consisted of three experiments that differed in the at-issueness measure used in a separate block.
-We focus on the projection ratings here.) In each experiment, participants were asked to read utterances like those in \ref{ex:family} and judge whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
+	Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The predicates were embedded under polar questions in Exps.~1, under negation (Exps.~2), under {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
+(Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block.
+We focus on the projection ratings here.) In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
 Participants gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
-Each participant rated the projection of the CC of all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
+Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
 We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
@@
 {\bf Results.}
-	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by embedding operator; the predicates are ordered by their mean projection across all operators (\emph{be annoyed} has the highest overall projection mean).
+	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
 We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
 	%
-	These results (which are supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}) are unexpected 
-	based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: Contrary to assumption, the CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
-The pattern observed for {\em know} is neither that of a factive nor that of a semi-factive predicate.
+	These results (supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}) are unexpected 
+	based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
+The pattern observed for {\em know} is neither that of a factive nor a semi-factive predicate.
 Similar considerations apply to \emph{see} and \emph{reveal}, which are also considered (semi-)factive in the literature.
@@
 {\bf Discussion:  Empirical and theoretical implications.} 
-	Our results --- that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection --- have several empirical and theoretical implications.
-First, the results for the predicates typically considered (semi-)factive, namely \emph{be annoyed, know, see, discover} and \emph{reveal}, call into question the assumed distinction between factive and semi-factive predicates (see also \cite{beaver_have_2010}).
-Future research that appeals to these two categories must clarify their definition.
+	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---have several empirical and theoretical implications.
+First, the results for the predicates typically considered (semi-)factive, ({\em viz.} \emph{be annoyed, know, see, discover} and \emph{reveal}), call into question the assumed distinction between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
+Future research appealing to these categories must clarify their definition.
 Second, claims about projection variability must be relativized to the entailment-canceling operator.
-While our data replicate the result from \citet{tonhauser_how_2018} that, in a polar question, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals.
-Finally, our results provide further support (from embedding under negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
+While our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals.
+Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
 
@@
 	%
-	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality contributed by these predicates.
+	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality associated with these predicates.
 	%
-	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
+	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf `Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
 	%
-	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in the antecedent of a conditional: This {\bf `Conditional high'} pattern, shown in (c), may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
+	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This {\bf `Conditional high'} pattern (c) may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
 	%
-	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern, shown in (d).
+	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern (d).
 The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
@@
 {\scriptsize
-\noindent {\bf References:}
+\noindent {\bf (Selected) References:}
 %
@@
 {\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
-{\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
+% {\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
 {\bf Karttunen (1971).} Some observations on factivity. {\em Research on Language \& Social Interaction.} ~\textbullet~
@@
 {\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}~\textbullet~
-{\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
-{\bf Zimmermann (2011).} {\em The grammatical expression of focus in West Chadic: Variation and uniformity in and across languages.} Walter de Gruyter.
+% {\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
+% {\bf Zimmermann (2011).} {\em The grammatical expression of focus in West Chadic: Variation and uniformity in and across languages.} Walter de Gruyter.
 
 \begin{figure}[h]
-	\vspace{-.6\baselineskip}
+	\vspace{-.5\baselineskip}
 	\centering
-	\scalebox{.9}{\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}}\vspace{-1.7\baselineskip}
+	\scalebox{.9}{\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}}\vspace{-1.8\baselineskip}
 	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
@@
 \begin{table}[h]
-	\vspace{-1.2\baselineskip}\phantom{.}
-	\hspace{-1.2em}
-	\begin{tabular}{p{.7\linewidth} p{.02\linewidth} p{.25\linewidth}}
-		\scalebox{.85}{
+	\vspace{-1\baselineskip}\phantom{.}
+	\centering
+	% \hspace{-1.2em}
+	\begin{tabular}{p{.65\linewidth} p{.3\linewidth}}
+		\scalebox{.72}{
 			\begin{tabular}{llrrrr}
@@
 		}
-		&&
+		&
 		\parbox{\linewidth}{\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC.
@@
 
-\hspace{-3.3em}\begin{minipage}{\textwidth}\centering
+\hspace{-2.7em}
+	\begin{minipage}{\textwidth}\centering
 	\begin{tabular}{p{.235\linewidth} p{.235\linewidth} p{.235\linewidth} p{.235\linewidth}}
@@
 
-\normalsize \noindent Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
+\vspace{-.4\baselineskip}
+\noindent \normalsize Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
 

### 2023-02-08 SALT-abstract.tex
added some notes and graphs

\vspace{-\baselineskip}
\paragraph{Projection across entailment-cancelling operators.}  \hspace{-1em}
	Attitude ascriptions may come with an inference to the truth of their complement, which can project across entailment-cancelling operators (illustrated for \emph{\lq discover\rq} in \ref{ex:family}) \citep[et seq.]{kiparsky_fact_1970}.

	\vspace{-.4\baselineskip}
	% The semantically heterogeneous entailment-cancelling operators are still often assumed to have similar effects on projection (xxx references xxx).
	\citet{karttunen_observations_1971} proposed that projective triggers differ in how their projection behavior is affected by the different embedding contexts in \ref{ex:family}, distinguishing \emph{factive} predicates (\emph{regret, forget, resent}) from \emph{semi-factive} predicates (\emph{discover, realize, see, find out, notice}), suggesting that the content of factive complements always projects, while for semi-factives it does not always do so in polar questions or conditional antecedents \citep[see also][]{stalnaker_pragmatic_1977}. Throughout the literature, this distinction is often understood in terms of emotive factive and cognitive semi-factive predicates \citep{klein_two_1975,djarv_cognitive_2018}, suggesting that emotive factives indifferently project across entailment-cancelling operators, while cognitive semi-factives project only when asserted or denied.
	Although widely assumed, there has been little experimental evidence or systematic investigation to support Karttunen's intuition.

	% Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}. In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.
	%

	%
	% To provide a systematic way of distinguishing these classes, \cite{djarv_cognitive_2018} suggest that they correspond to emotive and cognitive predicates, respectively. In a study assessing the acceptability of affirmative responses to a target utterance while explicitly denying a potentially projective inference \citep[based on][]{cummins_experimental_2012}, they find that this kind of affirmative cancellation is more readily available for emotive than cognitive predicates, suggesting that the main clause content and projective content are logically more independent from each other for emotive predicates.
	%

	\cite{smith_relationship_2014}, investigating the effect of operator for various types of projection triggers on ratings of surprisal about the inferences in question, conclude that inferences triggered by \emph{know} and epithets project more from under negation than conditional antecedents whereas non-restrictive relative clauses show the opposite pattern.

	Here, we analyze measures from a task designed to assess speaker commitment more directly to address the questions: \textbf{(i)} Is projection affected by differences in entailment-canceling environments? \textbf{(ii)} Do these effects vary for different attitude verb triggers (and in what way)? A complex picture of projection emerges, with interactions beween entailment-canceling operators and triggers, not yet suggested in the literature. Our results indicate a need for a more nuanced view of the family-of-sentences diagnostics of projection, as well as a better understanding of the effect of the interaction between lexical semantics and operators on projection.

\vspace{-\baselineskip}
\paragraph{Experiments.} \hspace{-1em}
	We use the data from 12 experiments, which consisted of two blocks each: a projection task and a at-issueness task. These experiments were initially designed to investigate three at-issueness diagnostics (with sentences under each of the four entailment-canceling operations, for a total of 12 experiments) and their correlation to projection ratings. Here, we focus only on the projection data. To assess projection, we used a task eliciting responses about how strongly a speaker would be committed towards the embedded clause \citep[from][]{tonhauser_prosodic_2016}. We presented sentences like in \ref{ex:family} as uttered by a named speaker (e.g. “\textbf{Daniel:} \emph{\lq Did Cole\dots?\rq}”). Participants then provided a certainty-rating in response to a prompt like: \emph{“Is Daniel certain that Julian dances Salsa?”}, by moving a slider on a scale from \lq no\rq\ (coded as \texttt{0}), to \lq yes\rq\ (coded as \texttt{1}).

% paragraph experiment (end)

\vspace{-\baselineskip}
\paragraph{Comparisons and Expectations.} \hspace{-1em}
	We compared certainty-ratings for the four entailment-canceling operators in \ref{ex:family} (factor \texttt{operator}), and 20 clause-embedding predicates (factor \texttt{verb}: {\em be annoyed, discover, know, reveal, see, acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, pretend, suggest, say, think, be right, demonstrate}).
	Based on the Karttunen-Klein hypothesis about emotive factives vs. cognitive semi-factives, we would expect emotive factives (\emph{be annoyed}), and verbs normally taken to be factive (\emph{know}), to be highly projective in a way that is indifferent to the embedding context, and cognitive semi-factives (\emph{discover, see}) to show higher projectivity ratings under negation compared to questions and conditionals.

% paragraph design_and_expectations_ (end)

\vspace{-\baselineskip}
\paragraph{Method.} \hspace{-1em}
	The 12 experiments in our study all manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause. Experiments 1--3 used polar question embedding, exps. 4--6 used negation, 7--9 modals, and 10--12 conditionals, making the \texttt{operator} manipulation a between-subjects factor. We analyzed data from $2682$ self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk). Participants saw items Latin-squared and randomized with six control stimuli.

% paragraph method (end)

\vspace{-\baselineskip}
\paragraph{Results and Discussion.} \hspace{-1em}
	Pooling the data across our experiments, we examined the effect on certainty-ratings of \texttt{operator}, and \texttt{verb}, as well as their interaction. Mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals are given in  \textbf{Figure \ref{fig:figure1}}.

	% \vspace{-.3\baselineskip}
	% \textbf{version w baseline know/m}
	% \noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{know} and \texttt{modal} as reference levels, and random intercepts for participants and items. We found highly significant main effects of \texttt{operator} (\texttt{conditional}: $+0.142$, \texttt{negation}: $+0.143$, ; \texttt{question}: $+0.225$, where $p < 2e^{-16}$ in all three cases), as well as many interactions of \texttt{operator} and  \texttt{verb} across the board (where $p < 0.001$ in $38$ cases, $p < 0.01$ in one, and $p < 0.05$ in three out of $57$ possible interactions).\\

	The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
	% The mean for this baseline (intercept) is $0.867$.
	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter. We also found many interactions of \texttt{operator} and \texttt{verb} across the board, suggesting that the effect of embedding context diffes by verb.

	Notably, for the emotive predicate \emph{\lq be annoyed\rq}, we found no significant difference between question and negation contexts, as would be expected based on Karttunen, but unexpectedly, we find lower ratings for conditionals and modals compared to negation. The cognitive predicate \emph{\lq discover\rq}, is more projective in polar questions and conditional antecedents than under negation, patterning opposite to Karttunen's claims about semi-factives. The predicate \emph{know}, which is often regarded a canonical factive, shows effects that would be incompatible with a characterization as either factive or semi-factive: $\texttt{question} > \texttt{conditional, negation} > \texttt{modal}$. If \emph{\lq know\rq} is a factive predicate, no difference would be expected here. If it is semi-factive, we would, again, expect higher projectivity under negation than in questions and conditionals.

	% for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$). We also found many interactions of \texttt{operator} and \texttt{verb} across the board (where $p < 0.001$ in $43$ cases, $p < 0.01$ in three, and $p < 0.05$ in one out of $57$ possible interactions).\\

% paragraph results (end)

\vspace{-\baselineskip}
\paragraph{Verb profiles.} % (fold)
	\textbf{Figure \ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, showing different predicate profiles. xxx analyses of the profiles xxx
% paragraph profiles (end)

\vspace{-\baselineskip}
\paragraph{Conclusion.} % (fold)
	\begin{itemize}
		\item Entailment-cancelling operators affect projection in different ways: calling into question the uniform applicability of the family-of-sentences diagnostic
		\item The effect of entailment-cancelling operator on projection differs for different verbs. The picture that emerges could not be characterized in terms of Karttunen's characterization of (semi)factives.
		\item We find some preliminary verb classes.
	\end{itemize}
% paragraph conclusion (end)

\pagebreak
\begin{figure}[t]
		\vspace{-.8\baselineskip}
		\centering
		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
		\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals (y-axis), by verb (x-axis), and operator (color/grouping, where: \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions).}
		\label{fig:figure1}
	\end{figure}

	\begin{figure}[ht]
		\vspace{-.8\baselineskip}
		\centering
		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
		\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals (y-axis), by verb (x-axis), and operator (color/grouping, where: \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions).}
		\label{fig:figure2}
	\end{figure}

\bibliography{projective-content.bib}

\end{document}

### 2023-02-08 materials-lisa/notes/GPP-AIness-notes.tex
added some notes and graphs

--- 2022-11-06 materials-lisa/notes/notes.tex
@@
 \tableofcontents
-
-% \section{Questions} % (fold)
-% 	\begin{itemize}
-% 		\item Were the questions for projection / at-issueness diagnostics asked in randomized order?
-% 		\item What were the thoughts on arousal? How would this come into play? Is the idea that it would interact with the QUD?
-% 		\item How exactly are the pilots different from the main experiments?
-% 	\end{itemize}
-
-% % section questions (end)
 
@@
 		\ex. \a. \textbf{Christopher:} Did Cole discover that Julian dances salsa?
-				\b. \textbf{Sandy:} Yes, Julian dances salsa.
+				\b. \textbf{Sandy:} Yes, but Cole didn't discover that.
 				\z.
@@
 \pagebreak
+\section{Experiments with Conditionals} % (fold)
+	\subsection{1c---Conditional embedding,  are you sure?, assent + assertion of complement + Q-A judgment}
+
+		\ex. \a. \textbf{Christopher:} If Cole discovered that Julian dances salsa, Logan will be joyful.
+			\b. \textbf{Sandy:} Are you sure?
+			\b. \textbf{Christopher:} Yes, I'm sure that Julian dances Salsa.
+			\z.
+		\z.
+
+		\ex. The two follow-ups will be acceptable, (only) if:
+			\a. \textbf{At-issueness:}\\ \label{cond:ai}
+				Christophers utterance was \lq about\rq\ whether Julian dances Salsa, \textbf{and}
+			\b. \textbf{Commitment:}\\ \label{cond:commit}
+				Christopher is committed that Julian dances Salsa
+			\z.
+		\z.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[]{figures/c1-correl.pdf}
+			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. c1}
+			\label{fig:c1-corr}
+		\end{figure}
+
+		Three points are relevant here:
+		\begin{itemize}
+			\item Regarding at-issueness:\\
+				We can assume that any content from a sentence-initial conditional antecedent, is hard to interpret as at-issue.
+The discourse has already progressed, and the sentence-final main clause is more likely to determine the relevant set of alternatives for the subsequent discourse.
+(Maybe it is worth taking into account a more dynamic view of at-issueness.
+Look clause-by-clause; Adressing the previous QUD vs determining relevant set of alternatives for subsequent discourse)
+			\item Regarding speaker commitment:\\
+				Based on literal content, the speaker may or may not be committed to Julian dancing salsa.
+			\item Regarding interaction of the two:
+			\begin{itemize}
+				\item If content is at issue, it will be targeted by entailment cancelling operator, and now the speaker is not committed to the content.
+In that case, the follow-up is unacceptable by \ref{cond:commit}.
+				\item If content is not at-issue, it will project, and the speaker is committed. Now, however, the follow-up is unacceptable by \ref{cond:ai}.
+			\end{itemize}
+
+		\end{itemize}
+
+	% subsection 1c_conditional_ (end)
+
+	\subsection{2c---Conditional embedding,  yes, that's true+complement diagnostic (+ acceptability rating)} 
+		\ex. \a. \textbf{Christopher:} If Cole discovered that Julian dances salsa, Logan will be joyful.
+			\b. \textbf{Sandy:} Yes, that’s true, Julian dances salsa.
+			\z.
+		\z.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[]{figures/c2-correl.pdf}
+			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. c2}
+			\label{fig:c2-corr}
+		\end{figure}
+
+	% subsection 2c_ (end)
+
+	\subsection{3c---Conditional embedding, yes, that's true + negation of main clause (+ acceptability rating)}
+		\ex. \a. \textbf{Christopher:} If Cole discovered that Julian dances salsa, Logan will be joyful.
+			\b. \textbf{Sandy:} Yes, that’s true, but Cole didn't discover it.
+			\z.
+		\z.
+
+		\begin{figure}[h]
+			\centering
+			\includegraphics[]{figures/c3-correl.pdf}
+			\caption{Mean not-at-issueness ratings vs mean certainty ratings by predicate, Exp. c3}
+			\label{fig:c3-corr}
+		\end{figure}
+
+	% subsection 3c_conc (end)
+
+% section experiments_with_conditionals (end)
+
 \section{Some more notes} % (fold)

### 2023-02-08 results/main/meta-analyses/new/0_data prep/preprocessing.R

@@ -89,5 +89,5 @@ df_list <- lapply(df_list, function(df) {
 data_all <- list.rbind(df_list)
 
 # save combined data frame into csv file
-write.csv(data_all, "../data/data_combined.csv", row.names=FALSE)
+write.csv(data_all, "../data_combined.csv", row.names=FALSE)
 

### 2023-02-08 results/main/meta-analyses/new/2_ai-ness/rscripts/graphs.R

@@ -1,0 +1,246 @@
+
+
+### summary statistics and graphs for projection data ------
+
+# set wd to script dir
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load data
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
+
+# load helper functions
+source('../../../../helpers.R')
+
+# libraries for manipulating dataframes, and plotting
+library(dplyr)
+library(forcats)
+library(ggplot2)
+
+### summary statistics and graphs for projection data ------
+
+# set wd to script dir
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load data
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
+
+# load helper functions
+source('../../../../helpers.R')
+
+# libraries for manipulating dataframes, and plotting
+library(dplyr)
+library(forcats)
+library(ggplot2)
+
+# revert ai ratings back to original scale orientation
+data$ai <- 1-data$ai
+
+
+# add labels for ai-ness diagnostics ----
+data$test = as.factor(paste(data$op, data$exp_block))
+levels(data$test)
+new_levels <- c("sure?", "assent", "A/R")
+new_levels <- rep(new_levels, 3)
+new_levels <- append(new_levels, c("whether", "assent", "A/R"))
+levels(data$test) <- new_levels
+
+# mean ainess by diagnostic -----
+
+ai_means_test = data %>% group_by(test) %>%
+  summarize(Mean = mean(ai), CILow = ci.low(ai), 
+            CIHigh = ci.high(ai)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
+         test = fct_reorder(as.factor(test),Mean)) %>% ungroup()
+
+ai_means_test %>% mutate(test = fct_reorder(test, Mean, 
+                                          .fun = mean)) %>% 
+  ggplot(aes(x = test, y=Mean)) +
+  coord_cartesian(ylim = c(0,1)) +
+  geom_point(aes(shape = test), size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
[…]

### 2023-03-04 writing/SuB-abstract/SuB-abstract.tex
created SuB abstract folder
updated SuB abstract

--- 2022-12-09 SALT-abstract/SALT-abstract.tex
@@
 \usepackage[margin=1in]{geometry}
+
+\usepackage{enumitem}
 
@@
 	\textbf{\large%\thetitle
-		%A diverse family (of sentences)\\ 
-		Projection differs across embedding operators---but not like you think}
+		Projection differs across embedding operators}
 \end{center}
@@
 	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
-Our results do not align with the long-standing distinction of factive vs.
-semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
-	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragmatic properties which raises important questions for future research on projection.
+{JT: I don't think point (i) is true when we aggregate across the predicates, no?
+And when we don't aggregate, point (i) is just point (ii).
+So perhaps we should just focus on point (ii)?
+I've written the theoretical discussion that way.} The observed variation is not captured by contemporary projection analyses (e.g., \citealp{heim83,vds92,abrusan2011,schlenker2021}) and does not align with the long-standing distinction of factive vs.
+semi-factive predicates (e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
+Rather, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragmatic properties which raises important questions for future research on projection.
 
@@
 \vspace{-.6\baselineskip}
-	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g.
-\emph{regret, forget}) projects across all four operators, whereas that of semi-factive predicates (e.g.
-\emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
-	To date, systematic empirical investigations of Karttunen's intuition are scarce.
-
-	\citet{smith_relationship_2014} investigated by-operator variation (negation, conditional) for various types of projective content, observing that the content of non-restrictive relative clauses and \emph{win} was more projective under conditionals than negation, while \emph{know} and epithets exhibited the opposite pattern.
-	% , whereas clefts showed no difference.
-	Their task asked participants how surprised they would be to learn the content under investigation after observing the utterance.
-Since other factors besides speaker commitment modulate (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.
-So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
-However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
-\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-Here, it is also not obvious how exactly this task relates to projection.
-\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
-The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).
+Contemporary research rarely considers potential differences in projection from under these four operators, with very few exceptions and conflicting results:  
+\begin{itemize}[topsep = 0pt, itemsep=-5pt, leftmargin = 12pt]
+\item \citet{smith_relationship_2014} observed that the content of non-restrictive relative clauses and the preparatory content of \emph{win} was more projective under conditionals than negation, while the content of the clausal complement (CC) of English \emph{know} and the projective content of epithets exhibited the opposite pattern.
+In contrast, \citet{siekers-solstad2022} did not observe that the CC of German \emph{wissen} `know' was more projective under conditionals than negation.
+\item \citet{karttunen_observations_1971} suggested that the CC of English factive predicates (e.g.
+\emph{be annoyed, regret}) projects across all four operators, whereas that of English semi-factive predicates (e.g.
+\emph{discover, realize, see, notice}) always projects from under negation, but not always from the other three operators.
+This suggestion does not find empirical support from the experiment reported on in \citealt{siekers-solstad2022} for German, where the CC of the factive predicates \emph{bereuen} `regret', \emph{wissen} `know', and \emph{enth\"ullen} `reveal' did not project more from under negation than that of the semi-factive predicates \emph{entdecken} `discover', \emph{bemerken} `notice', and {\em herausfinden} `find out'.
+\end{itemize}
 
 \noindent
-{\bf Experiment.}
-	We present results of a study investigating by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
-Projection was measured  across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}).
+To investigate whether these diverging results are due to differences between English and German or differences in the methods of investigation, we ran a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}.
+Our experiment used the same projection measure as \citealt{siekers-solstad2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied this diagnostic to the contents of the complements of 20 English clause-embedding predicates, including purported factive predicates (e.g., \emph{be annoyed, know, reveal}) and purported semi-factive predicates (e.g., \emph{discover, see}).
+Given recent results that the CC of non-factive predicates is also projective (\citealt{degen-tonhauser-language}), albeit to varying degrees, we also included 15 non-factive predicates.
 
-% paragraph experiment (end)
 \noindent
-{\bf Methods and Expectations.}
+{\bf Experiment: Methods and Expectations.}
 	Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The predicates were embedded under polar questions in Exps.~1, under negation (Exps.~2), under {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
@@
 {\bf Results.}
+{\bf The results need to be rewritten to make the points (i) and (ii).
+Right now, they are too focused on the factive/semi-factive distinction.
+No wonder did the SALT-reviewers think that this is the main point of our abstract...
+We should also say that our results replicate the result of Degen \& Tonhauser 2022 (Language) that projection is gradient, made there for interrogative embedding and here shown also for the other operators.}
+
 	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
@@
 \noindent
+{\bf Discussion:  Implications for projection analyses.}\\
+Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
+The first reason is that contemporary analyses do not lead us to expect interactions with operator meanings.
+In \citealt{heim83}, for instance, the CC of (semi-)factive predicates projects to the global context, except when global projection would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
+While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out.
+The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates (e.g., \citealt{heim83,vds92}, whose CCs are analyzed as presuppositions) and or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan2011,simons-etal2017}).
+A possible exception is the analysis of \citealt{schlenker2021}, which predicts the potential for projection for CCs that are contextually entailed.
+In the full talk, we discuss how this analysis might be able to capture the gradient projection observed in our experiment.
+The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
+Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
+
 {\bf Discussion:  Empirical and theoretical implications.}
+
+{\bf We need to relate our results to those of Siekers and Solstad wrt the question of whether there is xling variation in projection from under different operators.}
+
+{\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
+However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
+\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
+Here, it is also not obvious how exactly this task relates to projection.
+\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
+The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
+
 	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---have several empirical and theoretical implications.
@@
 
-\bibliography{projective-content.bib}
+\bibliography{../bibliography.bib}
 

### 2023-03-13 writing/SuB-abstract/SuB-abstract.tex
abstract edits sections 1+2

@@
 	\textbf{\large%\thetitle
-		Projection differs across embedding operators}
+		Projection variability across embedding operators and triggers}
 \end{center}
@@
 \noindent 
-	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
-{JT: I don't think point (i) is true when we aggregate across the predicates, no?
-And when we don't aggregate, point (i) is just point (ii).
-So perhaps we should just focus on point (ii)?
-I've written the theoretical discussion that way.} The observed variation is not captured by contemporary projection analyses (e.g., \citealp{heim83,vds92,abrusan2011,schlenker2021}) and does not align with the long-standing distinction of factive vs.
-semi-factive predicates (e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
-Rather, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragmatic properties which raises important questions for future research on projection.
+	We present experimental data which suggests that \textbf{(i)} the projectivity of the content of attitude complements varies depending on the entailment-canceling operator used, and \textbf{(ii)} that this effect of entailment-cancelling operator between attitude predicates.
+	%
+	Our finding of (ii) has not been previously observed and presents new theoretical challenges.
+The variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), extending this result to by-predicate variability in interaction with embedding contexts.
+	%
+	Existing theoretical accounts of projection across different entailment-canceling operators cannot explain the observed by-trigger differences (e.g.,
+	\citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
+	% \citealp{heim83,vds92,abrusan2011,schlenker2021}).
+	To our knowledge, the only approach that considers interactions between predicates and embedding contexts is the traditional classification of factive versus semi-factive predicates (e.g., \citealt{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
+	%
+	Our experimental data suggests that any account of projection must consider interactions between triggers and embedding operators, and that this interaction cannot be fully determined based on categorical lexical classes.
+Instead, the projection variability reveals a more nuanced picture of lexical semantic/pragmatic properties which raises important questions for future research on projection.
 
 \noindent
-	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
-	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}.
-	\vspace{-.6\baselineskip}
+	{\bf Investigating Projectivity across entailment-cancelling operators.} Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
+	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.
+	% \vspace{-.6\baselineskip}
 	\ex. \label{ex:family}
@@
 
-\vspace{-.6\baselineskip}
-Contemporary research rarely considers potential differences in projection from under these four operators, with very few exceptions and conflicting results:  
-\begin{itemize}[topsep = 0pt, itemsep=-5pt, leftmargin = 12pt]
-\item \citet{smith_relationship_2014} observed that the content of non-restrictive relative clauses and the preparatory content of \emph{win} was more projective under conditionals than negation, while the content of the clausal complement (CC) of English \emph{know} and the projective content of epithets exhibited the opposite pattern.
-In contrast, \citet{siekers-solstad2022} did not observe that the CC of German \emph{wissen} `know' was more projective under conditionals than negation.
-\item \citet{karttunen_observations_1971} suggested that the CC of English factive predicates (e.g.
-\emph{be annoyed, regret}) projects across all four operators, whereas that of English semi-factive predicates (e.g.
-\emph{discover, realize, see, notice}) always projects from under negation, but not always from the other three operators.
-This suggestion does not find empirical support from the experiment reported on in \citealt{siekers-solstad2022} for German, where the CC of the factive predicates \emph{bereuen} `regret', \emph{wissen} `know', and \emph{enth\"ullen} `reveal' did not project more from under negation than that of the semi-factive predicates \emph{entdecken} `discover', \emph{bemerken} `notice', and {\em herausfinden} `find out'.
-\end{itemize}
+	\noindent Following the methodology and approach from \citet{tonhauser_how_2018,degen_are_2022}, who find gradient differences in projection among different clause-embedding predicates, we talk about \emph{projectivity} as the degree or frequency with which a projective interpretation is assigned.
+%
+	Contemporary empirical research rarely considers potential differences in projectivity from under these four operators, with very few exceptions and conflicting results: \citet{smith_relationship_2014} investigated by-operator variation (negation, conditional) for various types of projective content, observing that the content of non-restrictive relative clauses and the preparatory content of \emph{win} was more projective under conditionals than negation, while the projective content of epithets and the clausal complement (CC) of English \emph{know} exhibited the opposite pattern.
+In contrast, \citet{sieker_projective_2022}, investigating projective variabiltiy of the CC German attitude predicates across the four operators in \Last, they found no significant projection variability across operators, and did not report interactions by-predicate.
+In particular they did not find a difference between factive and semi-factive verbs, nor did they replicate \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}).
+This also raises the question whether these diverging results are due to differences between English and German or differences in the methods of investigation.
+	%
 
-\noindent
-To investigate whether these diverging results are due to differences between English and German or differences in the methods of investigation, we ran a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}.
-Our experiment used the same projection measure as \citealt{siekers-solstad2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied this diagnostic to the contents of the complements of 20 English clause-embedding predicates, including purported factive predicates (e.g., \emph{be annoyed, know, reveal}) and purported semi-factive predicates (e.g., \emph{discover, see}).
-Given recent results that the CC of non-factive predicates is also projective (\citealt{degen-tonhauser-language}), albeit to varying degrees, we also included 15 non-factive predicates.
+	To investigate projection variability across operators and attitude predicates, we ran a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}.
+Our experiment used the same projection measure as \citealt{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied this diagnostic to the contents of the complements of 20 English clause-embedding predicates, including purported factive predicates (e.g., \emph{be annoyed, know, reveal}) and purported semi-factive predicates (e.g., \emph{discover, see}).
+Given recent results that the CC of non-factive predicates is also projective (\citealt{degen_are_2022}), albeit to varying degrees, we also included 15 non-factive predicates.
 
@@
 
+\begin{itemize}
+	\item \citet{karttunen_observations_1971} suggested that the CC of English factive predicates (e.g.
+\emph{be annoyed, regret}) projects across all four operators, whereas that of English semi-factive predicates (e.g.
+\emph{discover, realize, see, notice}) always projects from under negation, but not always from the other three operators.
+This suggestion does not find empirical support from the experiment reported on in \citealt{siekers-solstad2022} for German, where the CC of the factive predicates \emph{bereuen} `regret', \emph{wissen} `know', and \emph{enth\"ullen} `reveal' did not project more from under negation than that of the semi-factive predicates \emph{entdecken} `discover', \emph{bemerken} `notice', and {\em herausfinden} `find out'.
+\end{itemize}
+
 {\bf Discussion:  Empirical and theoretical implications.}
@@
 
-{\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
+% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
 However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
@@
 
-\bibliography{../bibliography.bib}
+% \bibliography{../bibliography.bib}
+\bibliography{../projective-content.bib}
 

### 2023-03-14 writing/SuB-abstract/SuB-abstract.tex
abstract updates

--- 2022-12-09 SALT-abstract/SALT-abstract.tex
@@
 
-\documentclass[12pt, letterpaper]{article}
+\documentclass[12pt, a4paper]{article}
 
 %% Page Layout
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=2.54cm]{geometry}
 
-\usepackage{euler} % math font package needs to be loaded before others
+\usepackage{enumitem}
+
+% \usepackage{euler} % math font package needs to be loaded before others
 \usepackage{xunicode,xltxtra, polyglossia}
@@
 	\textbf{\large%\thetitle
-		%A diverse family (of sentences)\\ 
-		Projection differs across embedding operators---but not like you think}
+		Projection variability for attitude complements across entia}
 \end{center}
@@
 \noindent 
-	We present experimental evidence \textbf{(i)} that the projection of the content of attitude complements differs between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
-Our results do not align with the long-standing distinction of factive vs.
-semi-factive predicates (see, e.g., \citealp{karttunen_observations_1971,hooper_applicability_1973,djarv_cognitive_2018}).
-	Instead, the observed by-operator variation reveals a more nuanced picture of lexical semantic/pragmatic properties which raises important questions for future research on projection.
+	We present experimental data that the projection of the content of attitude complements \textbf{(i)} varies between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
+	%
+	The observed variability is not captured by existing theoretical accounts of projection (e.g.,
+	\citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), but suggests that an account of projection must consider interactions between predicates and operators, which cannot be fully determined based on categorical lexical classes.
+Instead, our results reveal a more nuanced picture of lexical semantic and pragmatic properties, which raises important questions for future research on projection.
+	%
 
-\noindent
-	{\bf Projection across entailment-cancelling operators.} Interpreters may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
-	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}.
-	\vspace{-.6\baselineskip}
+\noindent {\bf Projection across entailment-cancelling operators.}
+	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
+	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.
+
+	\vspace{-.5\baselineskip}
 	\ex. \label{ex:family}
@@
 
-\vspace{-.6\baselineskip}
-	\citet{karttunen_observations_1971} proposed that the CC of factive predicates (e.g.
-\emph{regret, forget}) projects across all four operators, whereas that of semi-factive predicates (e.g.
-\emph{discover, realize, see, notice}) always projects from under negation, but not always from under polar questions, modals, or conditionals.
-	To date, systematic empirical investigations of Karttunen's intuition are scarce.
-
-	\citet{smith_relationship_2014} investigated by-operator variation (negation, conditional) for various types of projective content, observing that the content of non-restrictive relative clauses and \emph{win} was more projective under conditionals than negation, while \emph{know} and epithets exhibited the opposite pattern.
-	% , whereas clefts showed no difference.
-	Their task asked participants how surprised they would be to learn the content under investigation after observing the utterance.
-Since other factors besides speaker commitment modulate (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.
-So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
-However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
-\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-Here, it is also not obvious how exactly this task relates to projection.
-\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
-The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).
+	\vspace{-.5\baselineskip}
+	\noindent  Current research only rarely examines differences in projectivity across these four operators, with conflicting findings.
+	%
+	\citet{smith_relationship_2014} found that that the content of non-restrictive relative clauses and the preparatory content of \emph{win} was more projective under conditionals than negation, whereas the projective content of epithets and the clausal complement of \emph{know} showed the opposite pattern.
+	%
+	\citet{karttunen_observations_1971} proposed that the CC of English factive predicates (e.g.
+\emph{be annoyed, regret}) projects across all four operators, whereas that of semi-factives (e.g.
+\emph{discover, realize, see, notice}) only projects across negation.
+	%
+	In contrast, \posscite{sieker_projective_2022} investigation of the projection of the CC of German attitude predicates found no significant projection variability across the four operators in \Last, regardless of predicate type.
+In particular, they did not find a difference between factive and semi-factive verbs, nor did they replicate \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}).
+	%
+	The divergent results raise the question of whether they are due to variation between English and German or methodological differences.
 
 \noindent
-{\bf Experiment.}
-	We present results of a study investigating by-operator projection variation for the CCs of 20 attitude predicates, including purported factive and semi-factive predicates (e.g., \emph{be annoyed, discover}).
-Projection was measured  across all four operators in \ref{ex:family} using the `certain that' diagnostic for projection (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}).
+{\bfseries Experiment.}
+	%
+	We conducted a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}, using the `certain that' diagnostic (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020,sieker_projective_2022}).
+We applied this diagnostic to the contents of the complements of 20 English clause-embedding predicates, including purported factive predicates (e.g., \emph{be annoyed, know, reveal}) and purported semi-factive predicates (e.g., \emph{discover, see}),
+	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
 
-% paragraph experiment (end)
 \noindent 
 {\bf Methods and Expectations.}
-	Projection of the CC of the 20 attitude predicate was measured in four sets of experiments: The predicates were embedded under polar questions in Exps.~1, under negation (Exps.~2), under {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
+	Projection of the CC of the 20 attitude predicates was measured in four sets of experiments: The predicates were embedded in polar questions in Exps.~1, under negation (Exps.~2), under {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
 (Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block.
@@
 Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
-We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
-Karttunen's generalization leads us to expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and possibly lower ratings under the other operators.
-	% paragraph method (end)
+We analyze data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
+Based on Karttunen's generalization we would expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and lower ratings under the other operators.
 
 \noindent
-{\bf Results.}
-	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
-We observe by-operator variation in projection means as well as differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+{\bf Results and Analysis.} 
+	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
 	%
-	These results (supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}) are unexpected 
-	based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of the purportedly factive predicate \emph{be annoyed} does not project invariably from all four operators, and the CC of the purportedly semi-factive predicate \emph{discover} does not project more from under negation than the other three operators.
-The pattern observed for {\em know} is neither that of a factive nor a semi-factive predicate.
-Similar considerations apply to \emph{see} and \emph{reveal}, which are also considered (semi-)factive in the literature.
+	We find \textbf{(i)} differences in projection between operators, even if we aggregate across predicates, and \textbf{(ii)} different differences (interactions) between operators and predicates.
+\textbf{(i)} the differences between operators, when pooling predicates are small, but signifgicant.
+This is supported by a linear mixed effects model, with fixed effect: \texttt{operator}, reference level: \texttt{question}; random effect: \texttt{participant} intercepts; Model output: \texttt{question} (baseline): $0.515$ (Maximum Likelihood Estimate), $0.006$ Standard Error; lower projectivity for \texttt{modal}: $-0.04$ MLE, $0.009$ SE, $|t| = 4.55$, $p < 0.001$; \texttt{negation}: $-0.027$ MLE, $0.006$ SE, $|t| = 4.3$, $p < 0.001$; higher projectivity for \texttt{conditional}: $0.044$ MLE, $0.009$ SE, $|t| = 5.07$, $p < 0.001$.\\
+
+	\noindent
+	{\bfseries LH: } I still wonder if it might be enough to appeal to a main effect of operator in a full model (with both predictor variables) to support point (i).
+I know it's not as thourough, but shows that this effect is present at least in the baseline, and saves space.\\
+
+	\noindent \textbf{(ii)} differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+These results (supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}).\\
+
+	\noindent
+	{\bfseries LH: } Here it may be enough to report interaction terms in a full model? Unless we explicitly want to make the point against the factive / semi-factive distinction.\\
+	% and significant interaction effects in these models (e.g. the first model has $n$?? significant interaction terms with $|t| > x$, and $p < y$...)
+
+	\noindent Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.\\
+
+	%
 
@@
 \noindent
-{\bf Discussion:  Empirical and theoretical implications.} 
-	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---have several empirical and theoretical implications.
-First, the results for the predicates typically considered (semi-)factive, ({\em viz.} \emph{be annoyed, know, see, discover} and \emph{reveal}), call into question the assumed distinction between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
+{\bf Discussion:  Implications for projection analyses.}
+	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
+	%
+	The first reason is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
+In \citealt{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
+While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out.
+	%
+	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}, whose CCs are analyzed as presuppositions) and or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
+A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
+In the full talk, we discuss how this analysis might be able to capture the gradient projection observed in our experiment.
+	%
+	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
+Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
+
+\noindent {\bf Discussion:  Empirical implications and lexical properties.} \\
+	%
+	% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
+However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
+\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
+Here, it is also not obvious how exactly this task relates to projection.
+\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
+The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
+	%
+	\noindent
+	{\bfseries LH: } Here, JT said before that we should discuss cross-linguistic implications re Sieker \& Solstad. What would our conlusion be here?.\\
+
+	The predicate / operator interactions we find for the selected English attitude predicates are not in line with \cite{sieker_projective_2022}, who find no such effects for the selected German attitudes, but used the same task as the study presented here.
+This difference in findings indicates potential cross-linguistic differences in projection variability, which provides an avenue for future research.\\
+
+	Our findings contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+The pattern we observed for {\em know} does not fit into either category.
+Similar considerations apply to \emph{see} and \emph{reveal}, which are also categorized as (semi-)factive in the literature, thus questioning the defined difference between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
 Future research appealing to these categories must clarify their definition.
-Second, claims about projection variability must be relativized to the entailment-canceling operator.
-While our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, this result does not carry over to conditionals.
-Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically distinguish between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
+Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
+Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
+Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
 
-\noindent
-{\bf Discussion: Novel research question.}
-	Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
+% \noindent
+% {\bf Discussion: Novel research question.}
+
+	So, can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
 	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
@@
 
-\bibliography{projective-content.bib}
+% \bibliography{../bibliography.bib}
+\bibliography{../projective-content.bib}
 

### 2023-03-14 results/main/meta-analyses/new/1_projectivity/rscripts/analysis.R

@@ -6,7 +6,7 @@ this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
 
 # load data
-data <- read.csv("../data/data_combined.csv", header = TRUE, sep = ",")
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
 
 
 # variables and contrasts for data analysis: ----
@@ -14,26 +14,28 @@ data <- read.csv("../data/data_combined.csv", header = TRUE, sep = ",")
 data$projective <- as.numeric(data$projective)
 
 # our independent variables
-## inference-triggering predicate, coded as "verb"
-data$verb <- as.factor(data$verb)
-levels(data$verb)
-contrasts(data$verb)
+## inference-triggering predicate, coded as "pred"
+data$pred <- as.factor(data$verb)
+levels(data$pred)
+contrasts(data$pred)
 ### dummy coding with "be annoyed" as baseline
-data$verb <- relevel(data$verb, ref = "be_annoyed")
-contrasts(data$verb)
+data$pred <- relevel(data$pred, ref = "be_annoyed")
+contrasts(data$pred)
 
 ## embedding operator, coded as "op"
 data$op <- as.factor(data$op)
 levels(data$op)
 contrasts(data$op)
-### dummy coding with "n" (negation) as baseline
-data$op <- relevel(data$op, ref = "n")
+### op dummy coding with "n" (negation) as baseline
+data$op <- relevel(data$op, ref = "q")
 contrasts(data$op)
+
 
 # coding random effects as factors
 data$workerid <- as.factor(data$workerid)
-data$content <- as.factor(data$content)
-
+# data$content <- as.factor(data$content) 
+## content should not be a random effect, since contents are uniquely paired w
+## preds, which are independent variable
 
 # exploring a couple of models (linear)
 ## library for linear mixed models
@@ -41,7 +43,16 @@ library(lme4)
 library(lmerTest)
 library(knitr)
 
-# glmm1 <- lmer(projective ~ op * verb + (1 | workerid) +
+# linear model w op as predictor only
+glmm0 <- lmer(projective ~ op + (1 + op | workerid))
+
+
+### op dummy coding with "n" (negation) as baseline
+data$op <- relevel(data$op, ref = "n")
+contrasts(data$op)
+
+
+# glmm1 <- lmer(projective ~ op * pred + (1 | workerid) +
[…]

### 2023-03-14 writing/SuB-abstract/SuB-abstract.tex
fixed title

@@
 	\textbf{\large%\thetitle
-		Projection variability for attitude complements across entia}
+		Projection variability of attitude complements across different operators}
 \end{center}
@@
 	%
-	The observed variability is not captured by existing theoretical accounts of projection (e.g.,
-	\citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), but suggests that an account of projection must consider interactions between predicates and operators, which cannot be fully determined based on categorical lexical classes.
+	The observed variability is not captured by existing theoretical accounts of projection
+	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), but suggests that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes.
 Instead, our results reveal a more nuanced picture of lexical semantic and pragmatic properties, which raises important questions for future research on projection.
@@
 	%
-	In contrast, \posscite{sieker_projective_2022} investigation of the projection of the CC of German attitude predicates found no significant projection variability across the four operators in \Last, regardless of predicate type.
+	In contrast, \posscite{sieker_projective_2022} investigation of the projection of the CC of German attitude predicates found higher projectivity under negation than other operators (e.g.
+modals), but no interaction with predicate type.
 In particular, they did not find a difference between factive and semi-factive verbs, nor did they replicate \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}).
@@
 Based on Karttunen's generalization we would expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and lower ratings under the other operators.
+
+	% the projection of the complement clause (CC) of 20 attitude predicates was measured in four sets of experiments: polar questions (Exps.
+1), negation (Exps.
+2), perhaps (Exps.
+3), and conditional antecedents (Exps.
+4).
+Each set had three experiments, using different measures.
+Participants judged whether the speaker was certain of the CC and responded on a slider from 0 to 1.
+Each participant saw all 20 attitude predicates paired with a unique content under one operator.
+Data was analyzed from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon’s MT platform.
+Factive predicates were expected to receive high projection ratings under all four operators, while semi-factive predicates were expected to have high ratings under negation and lower ratings under the other operators.
 

### 2023-03-14 writing/SuB-abstract/SuB-abstract.tex
section 2 updates

@@
 
+% 1
 \vspace{-.7\baselineskip}
@@
 
+% 2
 \noindent {\bf Projection across entailment-cancelling operators.}
@@
 	\vspace{-.5\baselineskip}
-	\noindent  Current research only rarely examines differences in projectivity across these four operators, with conflicting findings.
+	\noindent  Current research only rarely examines projectivity differences across these operators, with conflicting findings.
 	%
-	\citet{smith_relationship_2014} found that that the content of non-restrictive relative clauses and the preparatory content of \emph{win} was more projective under conditionals than negation, whereas the projective content of epithets and the clausal complement of \emph{know} showed the opposite pattern.
+	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it only projects across negation.
 	%
-	\citet{karttunen_observations_1971} proposed that the CC of English factive predicates (e.g.
-\emph{be annoyed, regret}) projects across all four operators, whereas that of semi-factives (e.g.
-\emph{discover, realize, see, notice}) only projects across negation.
+	Comparing different types of contents, \citet{smith_relationship_2014} found that that the projective content of epithets and the CC of \emph{know} was more projective under negation than conditionals,
+	whereas the content of non-restrictive relative clauses and the preparatory content of \emph{win} showed the opposite pattern.
+	% whereas the projective contents of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
 	%
-	In contrast, \posscite{sieker_projective_2022} investigation of the projection of the CC of German attitude predicates found higher projectivity under negation than other operators (e.g.
-modals), but no interaction with predicate type.
-In particular, they did not find a difference between factive and semi-factive verbs, nor did they replicate \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}).
-	%
-	The divergent results raise the question of whether they are due to variation between English and German or methodological differences.
+	Further, \posscite{sieker_projective_2022} investigation of the projection of the CC of German purported factive and semi-factive predicates replicated \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}), but as part of a pattern of overall higher projectivity under negation than other operators.
+They found no interaction with predicate type, or evidence for the (semi-)factive distinction.
+
+	All of these studies suggest a comparatively high projectivity across negation, but the results diverge about whether they found interactions between operators and different types of contents.
+To investigate if the different results are due to variation between English and German, differences in contents being tested, or methodological differences, we conducted a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}, using the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive predicates (\emph{be annoyed, know, reveal}) and semi-factives (\emph{discover, see}),
+	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
+Our study aimed to replicate the projection variability across operators of various English projective contents reported in Smith and Hall (2014) for English attitude complements.
 
 \noindent 
-{\bfseries Experiment.}
+{\bf Method.}
+	Projection of the CC of the 20 attitude predicates was measured in four sets of experiments: The predicates were embedded in polar questions in Exps.\ 1, under negation (Exps.\ 2), under {\em perhaps} (Exps.\ 3), and in conditional antecedents (Exps.\ 4).
+(Each set contained three experiments using different at-issueness measure in a separate block.
+Here, we focus on the projection ratings.)
 	%
-	We conducted a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}, using the `certain that' diagnostic (see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020,sieker_projective_2022}).
-We applied this diagnostic to the contents of the complements of 20 English clause-embedding predicates, including purported factive predicates (e.g., \emph{be annoyed, know, reveal}) and purported semi-factive predicates (e.g., \emph{discover, see}),
-	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
-
-\noindent 
-{\bf Methods and Expectations.}
-	Projection of the CC of the 20 attitude predicates was measured in four sets of experiments: The predicates were embedded in polar questions in Exps.~1, under negation (Exps.~2), under {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
-(Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block.
-We focus on the projection ratings here.) In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
+	In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
 Participants gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
@@
 We analyze data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
-Based on Karttunen's generalization we would expect the CC of factive predicates to consistently receive relatively high projection ratings under all four operators, and the CC of semi-factive predicates to exhibit high projection ratings under negation and lower ratings under the other operators.
-
-	% the projection of the complement clause (CC) of 20 attitude predicates was measured in four sets of experiments: polar questions (Exps.
-1), negation (Exps.
-2), perhaps (Exps.
-3), and conditional antecedents (Exps.
-4).
-Each set had three experiments, using different measures.
-Participants judged whether the speaker was certain of the CC and responded on a slider from 0 to 1.
-Each participant saw all 20 attitude predicates paired with a unique content under one operator.
-Data was analyzed from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon’s MT platform.
-Factive predicates were expected to receive high projection ratings under all four operators, while semi-factive predicates were expected to have high ratings under negation and lower ratings under the other operators.
+	% Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.
 
@@
 	%
-	We find \textbf{(i)} differences in projection between operators, even if we aggregate across predicates, and \textbf{(ii)} different differences (interactions) between operators and predicates.
-\textbf{(i)} the differences between operators, when pooling predicates are small, but signifgicant.
-This is supported by a linear mixed effects model, with fixed effect: \texttt{operator}, reference level: \texttt{question}; random effect: \texttt{participant} intercepts; Model output: \texttt{question} (baseline): $0.515$ (Maximum Likelihood Estimate), $0.006$ Standard Error; lower projectivity for \texttt{modal}: $-0.04$ MLE, $0.009$ SE, $|t| = 4.55$, $p < 0.001$; \texttt{negation}: $-0.027$ MLE, $0.006$ SE, $|t| = 4.3$, $p < 0.001$; higher projectivity for \texttt{conditional}: $0.044$ MLE, $0.009$ SE, $|t| = 5.07$, $p < 0.001$.\\
-
-	\noindent
-	{\bfseries LH: } I still wonder if it might be enough to appeal to a main effect of operator in a full model (with both predictor variables) to support point (i).
-I know it's not as thourough, but shows that this effect is present at least in the baseline, and saves space.\\
-
-	\noindent \textbf{(ii)} differences across the predicates in by-operator variation: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
-These results (supported by linear mixed effects models, see \textbf{Table~\ref{t:models}}).\\
-
-	\noindent
-	{\bfseries LH: } Here it may be enough to report interaction terms in a full model? Unless we explicitly want to make the point against the factive / semi-factive distinction.\\
-	% and significant interaction effects in these models (e.g. the first model has $n$?? significant interaction terms with $|t| > x$, and $p < y$...)
-
-	\noindent Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.\\
+	We find \textbf{(i)} by-operator differences in projectivity, and \textbf{(ii)} differences in by-operator variability between predicates.
+	%
+	The generalization \textbf{(i)} is supported by significant main effects of operator in the linear mixed effects models shown in \textbf{Table\ \ref{t:models}}.
+\textbf{(ii)} is supported by the fact that there are at leats $34$ significant interaction terms in these models (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}).
+Specifically, the models in \textbf{Table\ \ref{t:models}} illustrate specific differences across the predicates in by-operator variation with different reference levels for \texttt{predicate}: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals, and the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.
+	%
+	Therefore, we did not replicate the result from \citet{smith_relationship_2014}, that the CC of \emph{know} projects more from negation than conditionals.
+Given that our findings indicate many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from \posscite{sieker_projective_2022} findings for German, where no such interactions were found.
+	%
+	Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.
 
@@
 
-\noindent {\bf Discussion:  Empirical implications and lexical properties.} \\
+\noindent {\bf Discussion:  Empirical implications and lexical properties.}
 	%
@@
 	%
-	\noindent
-	{\bfseries LH: } Here, JT said before that we should discuss cross-linguistic implications re Sieker \& Solstad. What would our conlusion be here?.\\
-
+	% \noindent
+	% {\bfseries LH: } Here, JT said before that we should discuss cross-linguistic implications re Sieker \& Solstad. What would our conlusion be here?.\\
+	%
 	The predicate / operator interactions we find for the selected English attitude predicates are not in line with \cite{sieker_projective_2022}, who find no such effects for the selected German attitudes, but used the same task as the study presented here.
-This difference in findings indicates potential cross-linguistic differences in projection variability, which provides an avenue for future research.\\
+This difference in findings indicates potential cross-linguistic differences in projection variability, which provides an avenue for future research.
 
@@
 	\centering
-	% \hspace{-1.2em}
-	\begin{tabular}{p{.65\linewidth} p{.3\linewidth}}
-		\scalebox{.72}{
+	\hspace{-1.2em}
+	\begin{tabular}{p{.663\linewidth} p{.3\linewidth}}
+		\scalebox{.77}{
 			\begin{tabular}{llrrrr}
@@
 				\midrule
-				\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01  & 75.8 & *** \\
-				& operator: conditional & -0.12 & 0.02  & -7.38 & *** \\
-				& operator: modal & -0.16 & 0.02  & -10.04 & *** \\
-				& operator: question & 0.02 & 0.01  & 1.74 & n.s. \\
+				\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01 & 79.86 & ***\\
+				& operator: conditional & -0.12 & 0.02  & -7.36 & ***\\
+				& operator: modal & -0.16 & 0.02  & -10.01 & ***\\
+				& operator: question & 0.02 & 0.01 & 1.72 & n.s.\\
 				\midrule
-				\#2 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 &  69.24 & *** \\
-				&		operator: conditional & -0.001 & 0.02  & -0.08 & n.s. \\
-				&		operator: modal  & -0.14 & 0.02  & -9.2 & *** \\
-				&		operator: question & 0.08 & 0.01  & 5.72 & *** \\
+				\#2 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01 & 62.70 & ***\\
+				& operator: conditional & 0.11 & 0.02 & 7.11 & ***\\
+				& operator: modal & -0.06 & 0.02 & -3.63 & ***\\
+				& operator: question & 0.10 & 0.01 & 7.08 & ***\\
 				\midrule
-				\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01  & 59.48 & *** \\
-				& operator: conditional & 0.11 & 0.02  & 7.11 & *** \\
-				&		operator: modal & -0.06 & 0.02  & -3.6 & *** \\
-				&		operator: question & 0.1 & 0.01  & 7.07 & *** \\
-
+				\#3 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 & 72.97 & ***\\
+				& operator: conditional & 0.00 & 0.02 & -0.06 & n.s.\\
+				& operator: modal & -0.14 & 0.02 & -9.18 & ***\\
+				& operator: question & 0.08 & 0.01 & 5.67 & ***\\
 				\bottomrule
@@
 		&
-		\parbox{\linewidth}{\caption{\small Relevant parts of three linear mixed effects models that predict certainty ratings from a fixed effect of operator, predicate, and their interaction, with random effects for participant and CC.
-Models were fit with \texttt{lme4, lmertest} in \texttt{R}.}\label{t:models}}
+		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; with fixed effects of operator, predicate, and their interaction; random effect: participant intercepts.
+Models were fit with \texttt{lme4, lmertest} in \texttt{R}.
+All three models also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (Not shown here).}\label{t:models}}
 
@@
 \vspace{-.4\baselineskip}
-\noindent \normalsize Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
+\noindent \small Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
 
@@
 
-% \begin{table}[ht]
-% 	\caption{Model output w \texttt{be\_annoyed/n} as baseline}
-% 	\label{tab:annoyed-n}
-% 	\centering
-
-% 	\begin{tabular}{lrrrrr}
-% 		\toprule
-% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
-% 		\midrule
-% 		(Intercept) & 0.8669958 & 0.0114402 & 1479.204 & 75.7850523 & 0.0000000\\
-% 		opc & -0.1155850 & 0.0156641 & 26227.368 & -7.3789836 & 0.0000000\\
-% 		opm & -0.1564480 & 0.0155858 & 26288.139 & -10.0378702 & 0.0000000\\
-% 		opq & 0.0249138 & 0.0143090 & 56826.915 & 1.7411232 & 0.0816674\\
-% 		verbsuggest & -0.6162683 & 0.0134941 & 54372.895 & -45.6694112 & 0.0000000\\
-% 		\addlinespace
-% 		verbacknowledge & -0.2204417 & 0.0134948 & 54373.777 & -16.3352724 & 0.0000000\\
-% 		verbadmit & -0.3402430 & 0.0134933 & 54371.874 & -25.2156750 & 0.0000000\\
-% 		verbannounce & -0.3607504 & 0.0134931 & 54371.547 & -26.7359958 & 0.0000000\\
-% 		verbbe\_right & -0.6638924 & 0.0134936 & 54372.293 & -49.2003974 & 0.0000000\\
-% 		verbconfess & -0.3681695 & 0.0134949 & 54373.904 & -27.2820517 & 0.0000000\\
-% 		\addlinespace
-% 		verbconfirm & -0.6338363 & 0.0134931 & 54371.657 & -46.9746939 & 0.0000000\\
-% 		verbdemonstrate & -0.5307026 & 0.0134935 & 54372.157 & -39.3301466 & 0.0000000\\
-% 		verbdiscover & -0.1864716 & 0.0134942 & 54372.992 & -13.8186520 & 0.0000000\\
-% 		verbestablish & -0.5542561 & 0.0134933 & 54371.853 & -41.0764079 & 0.0000000\\
-% 		verbhear & -0.1956794 & 0.0134938 & 54372.473 & -14.5014554 & 0.0000000\\
-% 		\addlinespace
-% 		verbinform & -0.1374428 & 0.0134945 & 54373.345 & -10.1851106 & 0.0000000\\
-% 		verbknow & -0.0748953 & 0.0134936 & 54372.288 & -5.5504197 & 0.0000000\\
-% 		verbpretend & -0.5049533 & 0.0134944 & 54373.218 & -37.4195409 & 0.0000000\\
-% 		verbprove & -0.6364950 & 0.0134935 & 54372.138 & -47.1704273 & 0.0000000\\
-% 		verbreveal & -0.3057814 & 0.0134937 & 54372.418 & -22.6609884 & 0.0000000\\
-% 		\addlinespace
-% 		verbsay & -0.5419725 & 0.0134925 & 54370.770 & -40.1685159 & 0.0000000\\
-% 		verbsee & -0.1827512 & 0.0134932 & 54371.791 & -13.5438968 & 0.0000000\\
-% 		verbthink & -0.5101198 & 0.0134930 & 54371.491 & -37.8062103 & 0.0000000\\
-% 		opc:verbsuggest & 0.1594325 & 0.0190634 & 54372.626 & 8.3632939 & 0.0000000\\
-% 		opm:verbsuggest & 0.1766113 & 0.0189718 & 54372.046 & 9.3091498 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbsuggest & -0.0737648 & 0.0192906 & 54372.717 & -3.8238781 & 0.0001315\\
-% 		opc:verbacknowledge & 0.1945569 & 0.0190630 & 54372.333 & 10.2059764 & 0.0000000\\
-% 		opm:verbacknowledge & 0.1100553 & 0.0189729 & 54373.070 & 5.8006461 & 0.0000000\\
-% 		opq:verbacknowledge & 0.0672935 & 0.0192909 & 54372.991 & 3.4883599 & 0.0004864\\
-% 		opc:verbadmit & 0.2803349 & 0.0190626 & 54371.933 & 14.7060204 & 0.0000000\\
-% 		\addlinespace
-% 		opm:verbadmit & 0.1810708 & 0.0189717 & 54371.994 & 9.5442376 & 0.0000000\\
-% 		opq:verbadmit & 0.0440050 & 0.0192896 & 54371.832 & 2.2812856 & 0.0225354\\
-% 		opc:verbannounce & 0.2503929 & 0.0190639 & 54373.132 & 13.1343759 & 0.0000000\\
-% 		opm:verbannounce & 0.1633026 & 0.0189721 & 54372.338 & 8.6075049 & 0.0000000\\
-% 		opq:verbannounce & 0.0017532 & 0.0192899 & 54372.121 & 0.0908892 & 0.9275810\\
-% 		\addlinespace
-% 		opc:verbbe\_right & 0.1424317 & 0.0190623 & 54371.698 & 7.4718895 & 0.0000000\\
-% 		opm:verbbe\_right & 0.2582399 & 0.0189717 & 54371.967 & 13.6118412 & 0.0000000\\
-% 		opq:verbbe\_right & -0.0776846 & 0.0192897 & 54371.974 & -4.0272512 & 0.0000565\\
-% 		opc:verbconfess & 0.3111230 & 0.0190643 & 54373.455 & 16.3196605 & 0.0000000\\
-% 		opm:verbconfess & 0.1837114 & 0.0189741 & 54374.125 & 9.6821973 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbconfess & 0.0393640 & 0.0192890 & 54371.320 & 2.0407459 & 0.0412809\\
-% 		opc:verbconfirm & 0.2963833 & 0.0190629 & 54372.237 & 15.5476205 & 0.0000000\\
-% 		opm:verbconfirm & 0.3663333 & 0.0189721 & 54372.324 & 19.3090505 & 0.0000000\\
-% 		opq:verbconfirm & 0.0312870 & 0.0192902 & 54372.377 & 1.6219108 & 0.1048282\\
-% 		opc:verbdemonstrate & 0.2854366 & 0.0190647 & 54373.829 & 14.9719656 & 0.0000000\\
-% 		\addlinespace
-% 		opm:verbdemonstrate & 0.2783816 & 0.0189716 & 54371.830 & 14.6736292 & 0.0000000\\
-% 		opq:verbdemonstrate & 0.0753554 & 0.0192892 & 54371.520 & 3.9066023 & 0.0000937\\
-% 		opc:verbdiscover & 0.2270099 & 0.0190632 & 54372.450 & 11.9082962 & 0.0000000\\
-% 		opm:verbdiscover & 0.1002424 & 0.0189726 & 54372.795 & 5.2835252 & 0.0000001\\
-% 		opq:verbdiscover & 0.0762282 & 0.0192898 & 54372.014 & 3.9517423 & 0.0000777\\
-% 		\addlinespace
-% 		opc:verbestablish & 0.2570817 & 0.0190633 & 54372.539 & 13.4857080 & 0.0000000\\
-% 		opm:verbestablish & 0.2879581 & 0.0189714 & 54371.677 & 15.1785426 & 0.0000000\\
-% 		opq:verbestablish & 0.0061604 & 0.0192893 & 54371.583 & 0.3193687 & 0.7494481\\
-% 		opc:verbhear & 0.2016309 & 0.0190630 & 54372.273 & 10.5770983 & 0.0000000\\
-% 		opm:verbhear & -0.0046313 & 0.0189714 & 54371.681 & -0.2441216 & 0.8071376\\
-% 		\addlinespace
-% 		opq:verbhear & 0.0756021 & 0.0192902 & 54372.373 & 3.9192008 & 0.0000890\\
-% 		opc:verbinform & 0.1278740 & 0.0190645 & 54373.631 & 6.7074362 & 0.0000000\\
-% 		opm:verbinform & 0.0141486 & 0.0189741 & 54374.122 & 0.7456761 & 0.4558664\\
-% 		opq:verbinform & 0.0587564 & 0.0192913 & 54373.322 & 3.0457537 & 0.0023221\\
-% 		opc:verbknow & 0.1142699 & 0.0190629 & 54372.206 & 5.9943629 & 0.0000000\\
-% 		\addlinespace
-% 		opm:verbknow & 0.0130032 & 0.0189709 & 54371.181 & 0.6854301 & 0.4930755\\
-% 		opq:verbknow & 0.0569598 & 0.0192897 & 54371.975 & 2.9528579 & 0.0031498\\
-% 		opc:verbpretend & -0.0902047 & 0.0190640 & 54373.155 & -4.7316842 & 0.0000022\\
-% 		opm:verbpretend & -0.0194245 & 0.0189738 & 54373.822 & -1.0237558 & 0.3059552\\
-% 		opq:verbpretend & -0.2196823 & 0.0192900 & 54372.250 & -11.3883795 & 0.0000000\\
-% 		\addlinespace
-% 		opc:verbprove & 0.2579585 & 0.0190627 & 54372.031 & 13.5320999 & 0.0000000\\
-% 		opm:verbprove & 0.3580525 & 0.0189720 & 54372.223 & 18.8726898 & 0.0000000\\
-% 		opq:verbprove & 0.0230805 & 0.0192908 & 54372.899 & 1.1964509 & 0.2315259\\
-% 		opc:verbreveal & 0.2902650 & 0.0190644 & 54373.511 & 15.2255216 & 0.0000000\\
-% 		opm:verbreveal & 0.1767476 & 0.0189723 & 54372.513 & 9.3160804 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbreveal & 0.0665887 & 0.0192906 & 54372.717 & 3.4518763 & 0.0005571\\
-% 		opc:verbsay & 0.1985351 & 0.0190618 & 54371.193 & 10.4153412 & 0.0000000\\
-% 		opm:verbsay & 0.1451974 & 0.0189723 & 54372.495 & 7.6531273 & 0.0000000\\
-% 		opq:verbsay & -0.1815738 & 0.0192889 & 54371.195 & -9.4133926 & 0.0000000\\
-% 		opc:verbsee & 0.2175191 & 0.0190629 & 54372.234 & 11.4105822 & 0.0000000\\
-% 		\addlinespace
-% 		opm:verbsee & 0.0594092 & 0.0189723 & 54372.499 & 3.1313655 & 0.0017409\\
-% 		opq:verbsee & 0.1108274 & 0.0192896 & 54371.840 & 5.7454526 & 0.0000000\\
-% 		opc:verbthink & 0.0023197 & 0.0190619 & 54371.272 & 0.1216913 & 0.9031440\\
-% 		opm:verbthink & 0.0145577 & 0.0189717 & 54371.965 & 0.7673367 & 0.4428847\\
-% 		opq:verbthink & -0.1738342 & 0.0192903 & 54372.438 & -9.0115053 & 0.0000000\\
-% 		\bottomrule
-% 	\end{tabular}
-% \end{table}
-
-% \newpage
-
-% \begin{table}[ht]
-% 	\caption{Model output w \texttt{discover/n} as baseline}
-% 	\label{tab:discover-n}
-% 	\centering
-
-% 	\begin{tabular}{lrrrrr}
-% 		\toprule
-% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
-% 		\midrule
-% 		(Intercept) & 0.6805242 & 0.0114403 & 1479.217 & 59.4849565 & 0.0000000\\
-% 		opc & 0.1114249 & 0.0156639 & 26226.686 & 7.1134644 & 0.0000000\\
-% 		opm & -0.0562056 & 0.0155855 & 26286.826 & -3.6062779 & 0.0003112\\
-% 		opq & 0.1011420 & 0.0143103 & 56828.295 & 7.0677830 & 0.0000000\\
-% 		verbbe\_annoyed & 0.1864716 & 0.0134942 & 54372.998 & 13.8186521 & 0.0000000\\
-% 		\addlinespace
-% 		verbsuggest & -0.4297967 & 0.0134937 & 54372.361 & -31.8516905 & 0.0000000\\
-% 		verbacknowledge & -0.0339702 & 0.0134936 & 54372.308 & -2.5174950 & 0.0118221\\
-% 		verbadmit & -0.1537714 & 0.0134942 & 54372.985 & -11.3953857 & 0.0000000\\
-% 		verbannounce & -0.1742788 & 0.0134947 & 54373.649 & -12.9145926 & 0.0000000\\
-% 		verbbe\_right & -0.4774208 & 0.0134935 & 54372.079 & -35.3816304 & 0.0000000\\
-% 		\addlinespace
-% 		verbconfess & -0.1816980 & 0.0134936 & 54372.192 & -13.4655376 & 0.0000000\\
-% 		verbconfirm & -0.4473648 & 0.0134944 & 54373.311 & -33.1517710 & 0.0000000\\
-% 		verbdemonstrate & -0.3442310 & 0.0134928 & 54371.247 & -25.5121517 & 0.0000000\\
-% 		verbestablish & -0.3677846 & 0.0134928 & 54371.167 & -27.2579106 & 0.0000000\\
-% 		verbhear & -0.0092079 & 0.0134940 & 54372.798 & -0.6823676 & 0.4950095\\
-% 		\addlinespace
-% 		verbinform & 0.0490288 & 0.0134942 & 54373.066 & 3.6333114 & 0.0002801\\
-% 		verbknow & 0.1115762 & 0.0134931 & 54371.668 & 8.2691029 & 0.0000000\\
-% 		verbpretend & -0.3184818 & 0.0134943 & 54373.176 & -23.6011440 & 0.0000000\\
-% 		verbprove & -0.4500234 & 0.0134934 & 54372.042 & -33.3512833 & 0.0000000\\
-% 		verbreveal & -0.1193099 & 0.0134941 & 54372.880 & -8.8416323 & 0.0000000\\
-% 		\addlinespace
-% 		verbsay & -0.3555009 & 0.0134943 & 54373.168 & -26.3444678 & 0.0000000\\
-% 		verbsee & 0.0037204 & 0.0134930 & 54371.450 & 0.2757288 & 0.7827574\\
-% 		verbthink & -0.3236483 & 0.0134936 & 54372.197 & -23.9853913 & 0.0000000\\
-% 		opc:verbbe\_annoyed & -0.2270099 & 0.0190632 & 54372.453 & -11.9082963 & 0.0000000\\
-% 		opm:verbbe\_annoyed & -0.1002424 & 0.0189726 & 54372.798 & -5.2835253 & 0.0000001\\
-% 		\addlinespace
-% 		opq:verbbe\_annoyed & -0.0762282 & 0.0192898 & 54372.017 & -3.9517423 & 0.0000777\\
-% 		opc:verbsuggest & -0.0675773 & 0.0190625 & 54371.851 & -3.5450396 & 0.0003929\\
-% 		opm:verbsuggest & 0.0763690 & 0.0189709 & 54371.211 & 4.0255864 & 0.0000569\\
-% 		opq:verbsuggest & -0.1499930 & 0.0192910 & 54373.069 & -7.7752981 & 0.0000000\\
-% 		opc:verbacknowledge & -0.0324529 & 0.0190617 & 54371.116 & -1.7025196 & 0.0886637\\
-% 		\addlinespace
-% 		opm:verbacknowledge & 0.0098129 & 0.0189711 & 54371.417 & 0.5172570 & 0.6049789\\
-% 		opq:verbacknowledge & -0.0089347 & 0.0192902 & 54372.396 & -0.4631724 & 0.6432426\\
-% 		opc:verbadmit & 0.0533251 & 0.0190632 & 54372.508 & 2.7972739 & 0.0051554\\
-% 		opm:verbadmit & 0.0808284 & 0.0189728 & 54372.916 & 4.2602367 & 0.0000205\\
-% 		opq:verbadmit & -0.0322232 & 0.0192908 & 54372.883 & -1.6703960 & 0.0948468\\
-% 		\addlinespace
-% 		opc:verbannounce & 0.0233831 & 0.0190642 & 54373.367 & 1.2265427 & 0.2199998\\
-% 		opm:verbannounce & 0.0630602 & 0.0189726 & 54372.754 & 3.3237576 & 0.0008887\\
-% 		opq:verbannounce & -0.0744750 & 0.0192926 & 54374.465 & -3.8602856 & 0.0001134\\
-% 		opc:verbbe\_right & -0.0845782 & 0.0190620 & 54371.360 & -4.4370099 & 0.0000091\\
-% 		opm:verbbe\_right & 0.1579976 & 0.0189719 & 54372.163 & 8.3279685 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbbe\_right & -0.1539128 & 0.0192901 & 54372.283 & -7.9788617 & 0.0000000\\
-% 		opc:verbconfess & 0.0841131 & 0.0190632 & 54372.437 & 4.4123417 & 0.0000102\\
-% 		opm:verbconfess & 0.0834691 & 0.0189719 & 54372.142 & 4.3996174 & 0.0000109\\
-% 		opq:verbconfess & -0.0368642 & 0.0192905 & 54372.632 & -1.9110088 & 0.0560087\\
-% 		opc:verbconfirm & 0.0693734 & 0.0190644 & 54373.529 & 3.6388992 & 0.0002741\\
-% 		\addlinespace
-% 		opm:verbconfirm & 0.2660910 & 0.0189728 & 54372.917 & 14.0248948 & 0.0000000\\
-% 		opq:verbconfirm & -0.0449413 & 0.0192914 & 54373.472 & -2.3295974 & 0.0198311\\
-% 		opc:verbdemonstrate & 0.0584268 & 0.0190633 & 54372.586 & 3.0648793 & 0.0021786\\
-% 		opm:verbdemonstrate & 0.1781393 & 0.0189714 & 54371.707 & 9.3898743 & 0.0000000\\
-% 		opq:verbdemonstrate & -0.0008729 & 0.0192906 & 54372.707 & -0.0452486 & 0.9639093\\
-% 		\addlinespace
-% 		opc:verbestablish & 0.0300718 & 0.0190618 & 54371.161 & 1.5775980 & 0.1146638\\
-% 		opm:verbestablish & 0.1877158 & 0.0189720 & 54372.212 & 9.8943713 & 0.0000000\\
-% 		opq:verbestablish & -0.0700678 & 0.0192899 & 54372.103 & -3.6323635 & 0.0002811\\
-% 		opc:verbhear & -0.0253789 & 0.0190637 & 54372.970 & -1.3312665 & 0.1831069\\
-% 		opm:verbhear & -0.1048737 & 0.0189721 & 54372.368 & -5.5277707 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbhear & -0.0006261 & 0.0192909 & 54372.991 & -0.0324581 & 0.9741069\\
-% 		opc:verbinform & -0.0991359 & 0.0190633 & 54372.605 & -5.2003420 & 0.0000002\\
-% 		opm:verbinform & -0.0860938 & 0.0189733 & 54373.368 & -4.5376355 & 0.0000057\\
-% 		opq:verbinform & -0.0174718 & 0.0192931 & 54374.843 & -0.9055994 & 0.3651520\\
-% 		opc:verbknow & -0.1127399 & 0.0190628 & 54372.127 & -5.9141311 & 0.0000000\\
-% 		\addlinespace
-% 		opm:verbknow & -0.0872392 & 0.0189719 & 54372.149 & -4.5983336 & 0.0000043\\
-% 		opq:verbknow & -0.0192684 & 0.0192895 & 54371.814 & -0.9989027 & 0.3178463\\
-% 		opc:verbpretend & -0.3172145 & 0.0190639 & 54373.134 & -16.6395086 & 0.0000000\\
-% 		opm:verbpretend & -0.1196669 & 0.0189725 & 54372.658 & -6.3073967 & 0.0000000\\
-% 		opq:verbpretend & -0.2959105 & 0.0192916 & 54373.627 & -15.3388172 & 0.0000000\\
-% 		\addlinespace
-% 		opc:verbprove & 0.0309486 & 0.0190626 & 54371.908 & 1.6235265 & 0.1044827\\
-% 		opm:verbprove & 0.2578102 & 0.0189728 & 54372.987 & 13.5883819 & 0.0000000\\
-% 		opq:verbprove & -0.0531478 & 0.0192907 & 54372.839 & -2.7550974 & 0.0058694\\
-% 		opc:verbreveal & 0.0632551 & 0.0190641 & 54373.234 & 3.3180321 & 0.0009071\\
-% 		opm:verbreveal & 0.0765052 & 0.0189711 & 54371.383 & 4.0327309 & 0.0000552\\
-% 		\addlinespace
-% 		opq:verbreveal & -0.0096396 & 0.0192917 & 54373.669 & -0.4996755 & 0.6173056\\
-% 		opc:verbsay & -0.0284748 & 0.0190633 & 54372.590 & -1.4936938 & 0.1352615\\
-% 		opm:verbsay & 0.0449550 & 0.0189713 & 54371.586 & 2.3696339 & 0.0178092\\
-% 		opq:verbsay & -0.2578020 & 0.0192911 & 54373.227 & -13.3637456 & 0.0000000\\
-% 		opc:verbsee & -0.0094907 & 0.0190633 & 54372.613 & -0.4978528 & 0.6185898\\
-% 		\addlinespace
-% 		opm:verbsee & -0.0408332 & 0.0189714 & 54371.728 & -2.1523480 & 0.0313743\\
-% 		opq:verbsee & 0.0345992 & 0.0192916 & 54373.623 & 1.7934827 & 0.0729013\\
-% 		opc:verbthink & -0.2246902 & 0.0190628 & 54372.159 & -11.7868151 & 0.0000000\\
-% 		opm:verbthink & -0.0856847 & 0.0189719 & 54372.175 & -4.5163904 & 0.0000063\\
-% 		opq:verbthink & -0.2500624 & 0.0192911 & 54373.178 & -12.9625865 & 0.0000000\\
-% 		\bottomrule
-% 	\end{tabular}
-% \end{table}
-
-% \newpage
-
-% \begin{table}[ht]
-% 	\caption{Model output w \texttt{know/n} as baseline}
-% 	\label{tab:discover-n}
-% 	\centering
-
-% 	\begin{tabular}{lrrrrr}
-% 		\toprule
-% 		  & Estimate & Std. Error & df & t value & Pr(>|t|)\\
-% 		\midrule
-% 		(Intercept) & 0.7921005 & 0.0114401 & 1479.182 & 69.2391632 & 0.0000000\\
-% 		opc & -0.0013151 & 0.0156638 & 26226.099 & -0.0839558 & 0.9330922\\
-% 		opm & -0.1434448 & 0.0155855 & 26286.727 & -9.2037503 & 0.0000000\\
-% 		opq & 0.0818736 & 0.0143093 & 56827.205 & 5.7217106 & 0.0000000\\
-% 		verbdiscover & -0.1115762 & 0.0134931 & 54371.661 & -8.2691029 & 0.0000000\\
-% 		\addlinespace
-% 		verbbe\_annoyed & 0.0748953 & 0.0134936 & 54372.286 & 5.5504198 & 0.0000000\\
-% 		verbsuggest & -0.5413729 & 0.0134928 & 54371.215 & -40.1230775 & 0.0000000\\
-% 		verbacknowledge & -0.1455464 & 0.0134940 & 54372.778 & -10.7859911 & 0.0000000\\
-% 		verbadmit & -0.2653476 & 0.0134936 & 54372.243 & -19.6647018 & 0.0000000\\
-% 		verbannounce & -0.2858550 & 0.0134935 & 54372.051 & -21.1847253 & 0.0000000\\
-% 		\addlinespace
-% 		verbbe\_right & -0.5889971 & 0.0134936 & 54372.234 & -43.6501264 & 0.0000000\\
-% 		verbconfess & -0.2932742 & 0.0134933 & 54371.896 & -21.7347532 & 0.0000000\\
-% 		verbconfirm & -0.5589410 & 0.0134943 & 54373.134 & -41.4204989 & 0.0000000\\
-% 		verbdemonstrate & -0.4558072 & 0.0134931 & 54371.605 & -33.7807521 & 0.0000000\\
-% 		verbestablish & -0.4793608 & 0.0134938 & 54372.501 & -35.5245129 & 0.0000000\\
-% 		\addlinespace
-% 		verbhear & -0.1207841 & 0.0134931 & 54371.608 & -8.9515418 & 0.0000000\\
-% 		verbinform & -0.0625474 & 0.0134940 & 54372.811 & -4.6351853 & 0.0000036\\
-% 		verbpretend & -0.4300580 & 0.0134939 & 54372.685 & -31.8704317 & 0.0000000\\
-% 		verbprove & -0.5615997 & 0.0134934 & 54372.014 & -41.6202578 & 0.0000000\\
-% 		verbreveal & -0.2308861 & 0.0134943 & 54373.179 & -17.1098402 & 0.0000000\\
-% 		\addlinespace
-% 		verbsay & -0.4670771 & 0.0134939 & 54372.578 & -34.6140392 & 0.0000000\\
-% 		verbsee & -0.1078558 & 0.0134932 & 54371.715 & -7.9933523 & 0.0000000\\
-% 		verbthink & -0.4352245 & 0.0134933 & 54371.896 & -32.2547883 & 0.0000000\\
-% 		opc:verbdiscover & 0.1127399 & 0.0190628 & 54372.122 & 5.9141311 & 0.0000000\\
-% 		opm:verbdiscover & 0.0872392 & 0.0189719 & 54372.145 & 4.5983336 & 0.0000043\\
-% 		\addlinespace
-% 		opq:verbdiscover & 0.0192684 & 0.0192895 & 54371.810 & 0.9989027 & 0.3178463\\
-% 		opc:verbbe\_annoyed & -0.1142699 & 0.0190629 & 54372.206 & -5.9943630 & 0.0000000\\
-% 		opm:verbbe\_annoyed & -0.0130032 & 0.0189709 & 54371.180 & -0.6854301 & 0.4930755\\
-% 		opq:verbbe\_annoyed & -0.0569598 & 0.0192897 & 54371.974 & -2.9528580 & 0.0031498\\
-% 		opc:verbsuggest & 0.0451626 & 0.0190614 & 54370.865 & 2.3693169 & 0.0178244\\
-% 		\addlinespace
-% 		opm:verbsuggest & 0.1636081 & 0.0189713 & 54371.561 & 8.6239932 & 0.0000000\\
-% 		opq:verbsuggest & -0.1307246 & 0.0192896 & 54371.847 & -6.7769511 & 0.0000000\\
-% 		opc:verbacknowledge & 0.0802870 & 0.0190639 & 54373.059 & 4.2114778 & 0.0000254\\
-% 		opm:verbacknowledge & 0.0970521 & 0.0189731 & 54373.215 & 5.1152465 & 0.0000003\\
-% 		opq:verbacknowledge & 0.0103337 & 0.0192900 & 54372.183 & 0.5357031 & 0.5921659\\
-% 		\addlinespace
-% 		opc:verbadmit & 0.1660650 & 0.0190633 & 54372.568 & 8.7112405 & 0.0000000\\
-% 		opm:verbadmit & 0.1680676 & 0.0189725 & 54372.674 & 8.8584875 & 0.0000000\\
-% 		opq:verbadmit & -0.0129548 & 0.0192897 & 54371.901 & -0.6715939 & 0.5018451\\
-% 		opc:verbannounce & 0.1361230 & 0.0190629 & 54372.249 & 7.1407107 & 0.0000000\\
-% 		opm:verbannounce & 0.1502994 & 0.0189716 & 54371.859 & 7.9223393 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbannounce & -0.0552066 & 0.0192897 & 54371.911 & -2.8619784 & 0.0042117\\
-% 		opc:verbbe\_right & 0.0281618 & 0.0190625 & 54371.883 & 1.4773357 & 0.1395915\\
-% 		opm:verbbe\_right & 0.2452367 & 0.0189720 & 54372.236 & 12.9262410 & 0.0000000\\
-% 		opq:verbbe\_right & -0.1346444 & 0.0192901 & 54372.308 & -6.9799743 & 0.0000000\\
-% 		opc:verbconfess & 0.1968531 & 0.0190627 & 54372.016 & 10.3266157 & 0.0000000\\
-% 		\addlinespace
-% 		opm:verbconfess & 0.1707082 & 0.0189737 & 54373.731 & 8.9971020 & 0.0000000\\
-% 		opq:verbconfess & -0.0175959 & 0.0192891 & 54371.441 & -0.9122160 & 0.3616590\\
-% 		opc:verbconfirm & 0.1821133 & 0.0190626 & 54371.955 & 9.5534239 & 0.0000000\\
-% 		opm:verbconfirm & 0.3533301 & 0.0189718 & 54372.046 & 18.6239655 & 0.0000000\\
-% 		opq:verbconfirm & -0.0256729 & 0.0192917 & 54373.707 & -1.3307729 & 0.1832693\\
-% 		\addlinespace
-% 		opc:verbdemonstrate & 0.1711667 & 0.0190647 & 54373.791 & 8.9782020 & 0.0000000\\
-% 		opm:verbdemonstrate & 0.2653784 & 0.0189716 & 54371.911 & 13.9881581 & 0.0000000\\
-% 		opq:verbdemonstrate & 0.0183955 & 0.0192903 & 54372.499 & 0.9536136 & 0.3402835\\
-% 		opc:verbestablish & 0.1428117 & 0.0190634 & 54372.667 & 7.4914038 & 0.0000000\\
-% 		opm:verbestablish & 0.2749549 & 0.0189719 & 54372.156 & 14.4927310 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbestablish & -0.0507994 & 0.0192892 & 54371.529 & -2.6335636 & 0.0084518\\
-% 		opc:verbhear & 0.0873610 & 0.0190622 & 54371.557 & 4.5829467 & 0.0000046\\
-% 		opm:verbhear & -0.0176345 & 0.0189709 & 54371.193 & -0.9295579 & 0.3526042\\
-% 		opq:verbhear & 0.0186422 & 0.0192902 & 54372.432 & 0.9664076 & 0.3338446\\
-% 		opc:verbinform & 0.0136041 & 0.0190628 & 54372.117 & 0.7136441 & 0.4754504\\
-% 		\addlinespace
-% 		opm:verbinform & 0.0011454 & 0.0189738 & 54373.859 & 0.0603655 & 0.9518647\\
-% 		opq:verbinform & 0.0017966 & 0.0192921 & 54374.044 & 0.0931259 & 0.9258039\\
-% 		opc:verbpretend & -0.2044746 & 0.0190640 & 54373.227 & -10.7256652 & 0.0000000\\
-% 		opm:verbpretend & -0.0324277 & 0.0189726 & 54372.773 & -1.7091875 & 0.0874219\\
-% 		opq:verbpretend & -0.2766421 & 0.0192909 & 54373.044 & -14.3405213 & 0.0000000\\
-% 		\addlinespace
-% 		opc:verbprove & 0.1436885 & 0.0190622 & 54371.604 & 7.5378609 & 0.0000000\\
-% 		opm:verbprove & 0.3450493 & 0.0189712 & 54371.500 & 18.1880552 & 0.0000000\\
-% 		opq:verbprove & -0.0338794 & 0.0192909 & 54373.040 & -1.7562330 & 0.0790543\\
-% 		opc:verbreveal & 0.1759951 & 0.0190637 & 54372.960 & 9.2319255 & 0.0000000\\
-% 		opm:verbreveal & 0.1637444 & 0.0189722 & 54372.374 & 8.6307727 & 0.0000000\\
-% 		\addlinespace
-% 		opq:verbreveal & 0.0096288 & 0.0192902 & 54372.406 & 0.4991553 & 0.6176720\\
-% 		opc:verbsay & 0.0842652 & 0.0190632 & 54372.497 & 4.4203002 & 0.0000099\\
-% 		opm:verbsay & 0.1321942 & 0.0189719 & 54372.094 & 6.9679113 & 0.0000000\\
-% 		opq:verbsay & -0.2385336 & 0.0192898 & 54371.999 & -12.3658155 & 0.0000000\\
-% 		opc:verbsee & 0.1032492 & 0.0190630 & 54372.324 & 5.4162003 & 0.0000001\\
-% 		\addlinespace
-% 		opm:verbsee & 0.0464060 & 0.0189718 & 54372.089 & 2.4460454 & 0.0144464\\
-% 		opq:verbsee & 0.0538675 & 0.0192901 & 54372.298 & 2.7924979 & 0.0052321\\
-% 		opc:verbthink & -0.1119503 & 0.0190627 & 54372.030 & -5.8727378 & 0.0000000\\
-% 		opm:verbthink & 0.0015545 & 0.0189713 & 54371.557 & 0.0819391 & 0.9346955\\
-% 		opq:verbthink & -0.2307940 & 0.0192903 & 54372.447 & -11.9642778 & 0.0000000\\
-% 		\bottomrule
-% 	\end{tabular}
-% \end{table}
-
 % \bibliography{../bibliography.bib}

### 2023-03-14 results/main/meta-analyses/new/1_projectivity/rscripts/analysis.R

@@ -27,8 +27,8 @@ data$op <- as.factor(data$op)
 levels(data$op)
 contrasts(data$op)
 ### op dummy coding with "n" (negation) as baseline
-data$op <- relevel(data$op, ref = "q")
-contrasts(data$op)
+# data$op <- relevel(data$op, ref = "q")
+# contrasts(data$op)
 
 
 # coding random effects as factors
@@ -44,23 +44,22 @@ library(lmerTest)
 library(knitr)
 
 # linear model w op as predictor only
-glmm0 <- lmer(projective ~ op + (1 + op | workerid))
-
+# glmm0 <- lmer(projective ~ op + (1 | workerid), data = data)
+summary0 <- summary(glmm0)
+print(summary0, cor=F, dig=3)
 
 ### op dummy coding with "n" (negation) as baseline
 data$op <- relevel(data$op, ref = "n")
 contrasts(data$op)
 
 
-# glmm1 <- lmer(projective ~ op * pred + (1 | workerid) +
-#                  (1 | content), data=data)
-
+# glmm1 <- lmer(projective ~ op * pred + (1 | workerid), data=data)
 # save.image("linear-models.RData")
 load("linear-models.RData")
 
 summary1 <- summary(glmm1)
 print(summary1, cor=F, dig=3)
-kable(summary1$coefficients, format = "latex", booktabs = TRUE)
+kable(summary1$coefficients, format = "latex", booktabs = TRUE, dig=2)
 
 
 ### dummy coding with "discover" as baseline

### 2023-03-14 writing/SuB-abstract/SuB-abstract.tex
abstract draft update

@@
 	The observed variability is not captured by existing theoretical accounts of projection
-	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), but suggests that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes.
+	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), and suggests that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes.
 Instead, our results reveal a more nuanced picture of lexical semantic and pragmatic properties, which raises important questions for future research on projection.
@@
 \noindent {\bf Projection across entailment-cancelling operators.}
-	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC),
+	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances Salsa.}),
 	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.
@@
 	%
-	Comparing different types of contents, \citet{smith_relationship_2014} found that that the projective content of epithets and the CC of \emph{know} was more projective under negation than conditionals,
-	whereas the content of non-restrictive relative clauses and the preparatory content of \emph{win} showed the opposite pattern.
-	% whereas the projective contents of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
+	Comparing different types of contents, \citet{smith_relationship_2014} found that the projective content of epithets and the CC of \emph{know} is more projective under negation than conditionals,
+	whereas that of non-restrictive relative clauses and \emph{win} shows the opposite pattern.
 	%
-	Further, \posscite{sieker_projective_2022} investigation of the projection of the CC of German purported factive and semi-factive predicates replicated \posscite{smith_relationship_2014} result for the German counterpart of `know' (\emph{wissen}), but as part of a pattern of overall higher projectivity under negation than other operators.
+	\posscite{sieker_projective_2022} investigation of German purported factive and semi-factive predicates replicated \posscite{smith_relationship_2014} result for the German \emph{wissen} (`know'), but as part of an overall pattern of higher projectivity across negation than other operators.
 They found no interaction with predicate type, or evidence for the (semi-)factive distinction.
 
-	All of these studies suggest a comparatively high projectivity across negation, but the results diverge about whether they found interactions between operators and different types of contents.
-To investigate if the different results are due to variation between English and German, differences in contents being tested, or methodological differences, we conducted a series of experiments designed to investigate projection from under the four entailment-canceling operators in \ref{ex:family}, using the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive predicates (\emph{be annoyed, know, reveal}) and semi-factives (\emph{discover, see}),
+	These studies all suggest comparatively high projectivity across negation, but the results diverge about interactions between operators and different types of contents.
+To investigate if this is due to variation between English and German, task-differences, or different contents being tested, we conducted a series of experiments designed to investigate projection across the four entailment-canceling operators in \ref{ex:family}, using the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive predicates (\emph{be annoyed, know, reveal}) and semi-factives (\emph{discover, see}),
 	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
@@
 	%
-	We find \textbf{(i)} by-operator differences in projectivity, and \textbf{(ii)} differences in by-operator variability between predicates.
+	Our analysis reveals two key findings: \textbf{(i)} differences in projectivity across operators, and \textbf{(ii)} differences in by-operator variability between predicates.
 	%
-	The generalization \textbf{(i)} is supported by significant main effects of operator in the linear mixed effects models shown in \textbf{Table\ \ref{t:models}}.
-\textbf{(ii)} is supported by the fact that there are at leats $34$ significant interaction terms in these models (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}).
-Specifically, the models in \textbf{Table\ \ref{t:models}} illustrate specific differences across the predicates in by-operator variation with different reference levels for \texttt{predicate}: For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals, and the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.
+	The first finding is supported by significant main effects of \texttt{operator} in the linear mixed effects models shown in \textbf{Table\ \ref{t:models}}.
+	The second finding is supported by the three models having at least $34$ significant interaction terms (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}).
+
+	The models in \textbf{Table\ \ref{t:models}} also illustrate some specific differences across predicates in by-operator variation (suing different reference levels for \texttt{predicate}): For instance, the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, while the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
 	%
-	Therefore, we did not replicate the result from \citet{smith_relationship_2014}, that the CC of \emph{know} projects more from negation than conditionals.
-Given that our findings indicate many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from \posscite{sieker_projective_2022} findings for German, where no such interactions were found.
+	Additionally, we did not replicate the result from \citet{smith_relationship_2014,sieker_projective_2022}, that the CC of \emph{know} projects more from negation than conditionals.
+Instead, we found that it projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.
 	%
-	Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}), and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.
-
-	%
+	With many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from \citet{sieker_projective_2022}, who did not find such interactions for the German attitudes they investigated.
 
@@
 
+\noindent {\bf Discussion:  Empirical implications and lexical properties.}
+	%
+	% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
+However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
+\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
+Here, it is also not obvious how exactly this task relates to projection.
+\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
+The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
+	%
+	Since our study used the same projection diagnostic, and a similar set of contents as \cite{sieker_projective_2022}, the different results indicate potential cross-linguistic variation in projection variability, suggesting a set of questions for future research.
+	%
+	However, in line with their study, our findings contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+The pattern we observed for {\em know} does not fit into either category,
+	thus questioning the defined difference between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
+	Future research appealing to these categories must clarify their definition. Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
+	% Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
+	%
+	Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
+	%
+	% Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}),
+	% and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.
+	%
+	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting initial generalizations over lexical properties, indicated in \textbf{Figure\ 2}.
+
+% \noindent
+% {\bf Discussion: Novel research question.}
+
+	% So, can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
+	% This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
+	% %
+	% The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality associated with these predicates.
+	% %
+	% The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf `Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
+	% %
+	% For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This {\bf `Conditional high'} pattern (c) may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
+	% %
+	% Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern (d).
+The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
+
+	%%%%%
+	% In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
+\textbf{Fig.~\ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, identifying some groups of predicates that show similar by-operator variation.
+We highlight four \lq projection-profiles\rq here: \emph{pretend} and \emph{think} are the only predicates that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
+\emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}).
+\emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}).
+Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).
+
 \noindent
-{\bf Discussion:  Implications for projection analyses.}
+{\bf Theoretical Implications.}
 	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
@@
 Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
-
-\noindent {\bf Discussion:  Empirical implications and lexical properties.}
-	%
-	% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
-However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
-\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-Here, it is also not obvious how exactly this task relates to projection.
-\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
-The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
-	%
-	% \noindent
-	% {\bfseries LH: } Here, JT said before that we should discuss cross-linguistic implications re Sieker \& Solstad. What would our conlusion be here?.\\
-	%
-	The predicate / operator interactions we find for the selected English attitude predicates are not in line with \cite{sieker_projective_2022}, who find no such effects for the selected German attitudes, but used the same task as the study presented here.
-This difference in findings indicates potential cross-linguistic differences in projection variability, which provides an avenue for future research.
-
-	Our findings contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
-The pattern we observed for {\em know} does not fit into either category.
-Similar considerations apply to \emph{see} and \emph{reveal}, which are also categorized as (semi-)factive in the literature, thus questioning the defined difference between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
-Future research appealing to these categories must clarify their definition.
-Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
-Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
-Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
-
-% \noindent
-% {\bf Discussion: Novel research question.}
-
-	So, can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
-	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
-	%
-	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality associated with these predicates.
-	%
-	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf `Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
-	%
-	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This {\bf `Conditional high'} pattern (c) may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
-	%
-	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern (d).
-The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
-
-	% In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
-\textbf{Fig.~\ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, identifying some groups of predicates that show similar by-operator variation.
-We highlight four \lq projection-profiles\rq here: \emph{pretend} and \emph{think} are the only predicates that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
-\emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}).
-\emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}).
-Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).
 

### 2023-03-15 writing/SuB-abstract/SuB-abstract.tex
minor edit end of background section
restructured results

@@
 \bibliographystyle{plainnat}
-% \enablehyphenation
+\enablehyphenation
 % \vspace{-2em}
@@
 	The observed variability is not captured by existing theoretical accounts of projection
-	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}), and suggests that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes.
-Instead, our results reveal a more nuanced picture of lexical semantic and pragmatic properties, which raises important questions for future research on projection.
+	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
+	%
+	Our results suggest that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes, instead revealing a more nuanced picture of lexical semantic and pragmatic properties.
+This raises important questions for future research on projection.
 	%
@@
 \noindent {\bf Projection across entailment-cancelling operators.}
-	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances Salsa.}),
+	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
 	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.
@@
 	\vspace{-.5\baselineskip}
-	\noindent  Current research only rarely examines projectivity differences across these operators, with conflicting findings.
+	\noindent  Current research rarely examines projection variability across these operators, with conflicting findings.
 	%
-	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it only projects across negation.
+	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it always projects across negation, but not always for the other operators.
 	%
-	Comparing different types of contents, \citet{smith_relationship_2014} found that the projective content of epithets and the CC of \emph{know} is more projective under negation than conditionals,
-	whereas that of non-restrictive relative clauses and \emph{win} shows the opposite pattern.
+	Comparing different types of contents, \citet{smith_relationship_2014} found that the projective content of epithets and the CC of \emph{know} was more projective under negation than conditionals,
+	whereas that of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
 	%
-	\posscite{sieker_projective_2022} investigation of German purported factive and semi-factive predicates replicated \posscite{smith_relationship_2014} result for the German \emph{wissen} (`know'), but as part of an overall pattern of higher projectivity across negation than other operators.
+	\posscite{sieker_projective_2022} study on German attitude predicates replicated \posscite{smith_relationship_2014} result for German \emph{wissen} (`know'), but as part of an overall pattern of higher projectivity across negation than other operators.
 They found no interaction with predicate type, or evidence for the (semi-)factive distinction.
 
-	These studies all suggest comparatively high projectivity across negation, but the results diverge about interactions between operators and different types of contents.
-To investigate if this is due to variation between English and German, task-differences, or different contents being tested, we conducted a series of experiments designed to investigate projection across the four entailment-canceling operators in \ref{ex:family}, using the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive predicates (\emph{be annoyed, know, reveal}) and semi-factives (\emph{discover, see}),
-	as well as 15 non-factive predicates, due to recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
-Our study aimed to replicate the projection variability across operators of various English projective contents reported in Smith and Hall (2014) for English attitude complements.
+	These studies yield divergent results regarding whether they find interactions between operators and various projective contents, therefore raising the question if this is due to cross-linguistic variation, task differences, or different contents being tested.
+To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}.
+We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including purported factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}),
+	and 15 non-factive predicates, given recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
 
@@
 {\bf Results and Analysis.} 
-	\textbf{Figure~\ref{fig:figure1}} shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
 	%
-	Our analysis reveals two key findings: \textbf{(i)} differences in projectivity across operators, and \textbf{(ii)} differences in by-operator variability between predicates.
+	Our first key finding is that there is projection variability by operator: Across predicates, projection ratings were higher under question-embeddings than under negation and modals, but lower than in conditional antecedents.
+These differences, though small, are significant and supported by linear mixed effects model \textbf{\# 1} in \textbf{Table\ \ref{t:models}}.
 	%
-	The first finding is supported by significant main effects of \texttt{operator} in the linear mixed effects models shown in \textbf{Table\ \ref{t:models}}.
-	The second finding is supported by the three models having at least $34$ significant interaction terms (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}).
+	Second, there is by-predicate variation in the effect of operator on projection.
+This is illustrated in \textbf{Figure~\ref{fig:figure1}}, which shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
+	%
+	For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+This finding is further supported by models \textbf{\# 2--4} in \textbf{Table\ \ref{t:models}} each having at least $34$ significant interaction terms (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}).
+	%
+	% The models in \textbf{Table\ \ref{t:models}} also illustrate some specific differences across predicates in by-operator variation (using different reference levels for \texttt{predicate}): For instance, the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, while the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+	%
+	%
+	Contrary to what would be expected based on \posscite{karttunen_observations_1971}
+	distinction between factive and semi-factive predicates,
+	% the CC of assumed semi-factives \emph{discover, see} does not have higher projection-ratings under negation than other operators.
+\textbf{Figure~\ref{fig:figure1}} clearly shows that mean certainty ratings are lower for negation than conditionals and questions.
+	%
+	the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+The pattern we observed for {\em know} does not fit into either category.
+	%
+	%
+	Our findings, that by-operator projection variability differs by content, are in line with \citet{smith_relationship_2014}, but contrary to their findings that the CC of \emph{know} projects more from negation than conditionals, we found no significant difference here (Model \textbf{\# 4} in \textbf{Table\ \ref{t:models}}).\\
 
-	The models in \textbf{Table\ \ref{t:models}} also illustrate some specific differences across predicates in by-operator variation (suing different reference levels for \texttt{predicate}): For instance, the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, while the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
 	%
-	Additionally, we did not replicate the result from \citet{smith_relationship_2014,sieker_projective_2022}, that the CC of \emph{know} projects more from negation than conditionals.
-Instead, we found that it projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.
+	\noindent{\bf LH:} Alternatively, we might only talk about the missed predictions for the semi-factives here? \\
+
+	\noindent Instead, we found that it projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.\\
 	%
-	With many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from \citet{sieker_projective_2022}, who did not find such interactions for the German attitudes they investigated.
+	%
+
+	Like in \posscite{sieker_projective_2022} study of German attitude predicates, we find a presence of a main effect of by-operator projection variation, but the particular effect we find is different: While Sieker and Solstad found highest projection ratings with negation compared to the other operators, we find overall highest projectivity across conditionals, then questions, then modals and negation.
+	%
+	With many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from , who did not find such interactions for the German attitudes they investigated.
+	%
+	Since our study used the same projection diagnostic, and a similar set of contents as \cite{sieker_projective_2022}, the different results indicate potential cross-linguistic variation in projection variability.
+This suggests
+	extending the questions addressed in cross-linguistic investigations of projective variability (\citealt{tonhauser_projection_2020,xue_correlation_2011,tonhauser_how_2018}) to interactions with various entailment-cancelling operators.
 
@@
 
-\noindent {\bf Discussion:  Empirical implications and lexical properties.}
+{\bf Discussion.}
+	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
+	%
+	The first reason is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
+In \citealt{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
+While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out.
+	%
+	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}, whose CCs are analyzed as presuppositions) and or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
+A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
+In the full talk, we discuss how this analysis might be able to capture the gradient projection observed in our experiment.
+	%
+	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
+Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
+
+% \noindent {\bf Discussion:  Empirical implications and lexical properties.}
 	%
@@
 	%
-	Since our study used the same projection diagnostic, and a similar set of contents as \cite{sieker_projective_2022}, the different results indicate potential cross-linguistic variation in projection variability, suggesting a set of questions for future research.
-	%
-	However, in line with their study, our findings contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
-The pattern we observed for {\em know} does not fit into either category,
-	thus questioning the defined difference between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
+	Our results further question the proposed difference between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
 	Future research appealing to these categories must clarify their definition. Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
@@
 \noindent
-{\bf Theoretical Implications.}
-	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
-	%
-	The first reason is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
-In \citealt{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
-While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out.
-	%
-	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}, whose CCs are analyzed as presuppositions) and or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
-A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
-In the full talk, we discuss how this analysis might be able to capture the gradient projection observed in our experiment.
-	%
-	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
-Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
-
-% \noindent
-% {\bf Conclusion.}
-% 	The picture that emerges is complex and requires a nuanced view of lexical effects on projection.
-It raises the question of how lexical differences between predicates could explain why the CC projects more from under certain operators and less from others.
-Much previous work on this question has based its generalizations on impressionistic judgments, empirically establishing different projection-profiles associated with verbs therefore provides a basis for future research on lexical effects on projection.
 
@@
 				\midrule
-				\#1 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01 & 79.86 & ***\\
+				\#1 & Intercept: \emph{question} & 0.52 & 0.01 & 87.31 & ***\\
+				& operator: conditional & 0.04 & 0.01 & 5.07 & ***\\
+				& operator: modal & -0.04 & 0.01 & -4.55 & ***\\
+				& operator: negation & -0.03 & 0.01 & -4.30 & ***\\
+				\midrule
+				\#2 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01 & 79.86 & ***\\
 				& operator: conditional & -0.12 & 0.02  & -7.36 & ***\\
@@
 				\midrule
-				\#2 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01 & 62.70 & ***\\
+				\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01 & 62.70 & ***\\
 				& operator: conditional & 0.11 & 0.02 & 7.11 & ***\\
@@
 				\midrule
-				\#3 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 & 72.97 & ***\\
+				\#4 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 & 72.97 & ***\\
 				& operator: conditional & 0.00 & 0.02 & -0.06 & n.s.\\
@@
 		&
-		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; with fixed effects of operator, predicate, and their interaction; random effect: participant intercepts.
+		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; \textbf{\#1} has fixed effects of operator; random effect: participant intercepts, \textbf{\#2-4} have fixed effect: operator, predicate, and their interaction; random effect: participant intercepts.
 		Models were fit with \texttt{lme4, lmertest} in \texttt{R}.

### 2023-03-15 results/main/meta-analyses/new/1_projectivity/rscripts/analysis.R

@@ -62,32 +62,29 @@ print(summary1, cor=F, dig=3)
 kable(summary1$coefficients, format = "latex", booktabs = TRUE, dig=2)
 
 
-### dummy coding with "discover" as baseline
-data$pred <- relevel(data$pred, ref = "discover")
+### dummy coding with "know" as baseline
+data$pred <- relevel(data$pred, ref = "know")
 contrasts(data$pred)
 
-# glmm2 <- lmer(projective ~ op * pred + (1 | workerid) +
-#                  (1 | content), data=data)
-
+# glmm2 <- lmer(projective ~ op * pred + (1 | workerid), data=data)
 # save.image("linear-models.RData")
 load("linear-models.RData")
 
 summary2 <- summary(glmm2)
 print(summary2, cor=F, dig=3)
-kable(summary2$coefficients, format = "latex", booktabs = TRUE)
+kable(summary2$coefficients, format = "latex", booktabs = TRUE, dig = 2)
 
 
-### dummy coding with "know" as baseline
-data$pred <- relevel(data$pred, ref = "know")
+
+### dummy coding with "discover" as baseline
+data$pred <- relevel(data$pred, ref = "discover")
 contrasts(data$pred)
 
-# glmm3 <- lmer(projective ~ op * pred + (1 | workerid) +
-#                  (1 | content), data=data)
-
+# glmm3 <- lmer(projective ~ op * pred + (1 | workerid), data=data)
 # save.image("linear-models.RData")
 load("linear-models.RData")
 
 summary3 <- summary(glmm3)
 print(summary3, cor=F, dig=3)
-kable(summary3$coefficients, format = "latex", booktabs = TRUE)
+kable(summary3$coefficients, format = "latex", booktabs = TRUE, dig = 2)
 

### 2023-03-15 results/main/meta-analyses/new/1_projectivity/rscripts/graphs.R

@@ -258,7 +258,7 @@ proj_means %>% filter(groups == "1") %>%
   # directlabels::geom_dl(aes(label = verb), position = position_nudge(x = -.1),
   #                       method = list(cex=0.7, "first.bumpup")) +
   directlabels::geom_dl(aes(label = verb), position = position_nudge(x = .1),
-                        method = list(cex=1.3, rot = 50,  "last.bumpup")) +
+                        method = list(cex=1, rot = 50,  "last.bumpup")) +
   scale_shape_manual(values = c("M", "N", "Q", "C")) +
   scale_colour_manual(values=cbbPalette2) +
   geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1) +
@@ -279,8 +279,10 @@ proj_means %>% filter(groups == "2") %>%
                      labels = c("0","0.2","0.4","0.6","0.8","1"), 
                      name = "Mean certainty rating") +
   geom_point(aes(shape = op), size = 2.5) + 
-  directlabels::geom_dl(aes(label = verb), position = position_nudge(x = .075),
-                        method = list(cex=0.54, "last.bumpup")) +
+  # directlabels::geom_dl(aes(label = verb), position = position_nudge(x = .075),
+  #                       method = list(cex=0.54, "last.bumpup")) +
+  directlabels::geom_dl(aes(label = verb), position = position_nudge(x = .1),
+                        method = list(cex=1,  "smart.grid")) +
   scale_shape_manual(values = c("M", "N", "Q", "C")) +
   scale_colour_manual(values=cbbPalette2) +
   geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1) +
@@ -319,8 +321,10 @@ proj_means %>% filter(groups == "4") %>%
                      labels = c("0","0.2","0.4","0.6","0.8","1"), 
                      name = "Mean certainty rating") +
   geom_point(aes(shape = op), size = 3) + 
+  # directlabels::geom_dl(aes(label = verb), position = position_nudge(x = .1),
+  #                       method = list(cex=0.6, "last.bumpup")) +
   directlabels::geom_dl(aes(label = verb), position = position_nudge(x = .1),
-                        method = list(cex=0.6, "last.bumpup")) +
+                        method = list(cex=1,"smart.grid")) +
   scale_shape_manual(values = c("M", "N", "Q", "C")) +
   scale_colour_manual(values=cbbPalette2) +
[…]

### 2023-03-15 writing/SuB-abstract/SuB-abstract.tex
abstract submitted

@@
 
-% 1
-\vspace{-.7\baselineskip}
+\vspace{-.8\baselineskip}
 \noindent 
-	We present experimental data that the projection of the content of attitude complements \textbf{(i)} varies between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
+	We present experimental evidence that the projection of the content of attitude complements \textbf{(i)} varies between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
 	%
@@
 	%
-	Our results suggest that an analysis must consider interactions between predicates and operators, which cannot be fully determined by lexical classes, instead revealing a more nuanced picture of lexical semantic and pragmatic properties.
-This raises important questions for future research on projection.
-	%
+	Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.
 
-% 2
 \noindent {\bf Projection across entailment-cancelling operators.}
@@
 	%
-	\posscite{sieker_projective_2022} study on German attitude predicates replicated \posscite{smith_relationship_2014} result for German \emph{wissen} (`know'), but as part of an overall pattern of higher projectivity across negation than other operators.
-They found no interaction with predicate type, or evidence for the (semi-)factive distinction.
+	This result was replicated for German \emph{wissen} `know' in \posscite{sieker_projective_2022} study on German attitude predicates, but as part of an overall pattern of higher projection ratings with negation than other operators.
+They found no interaction with predicate type, or evidence for the distinction of factives vs.
+semi-factives.
 
-	These studies yield divergent results regarding whether they find interactions between operators and various projective contents, therefore raising the question if this is due to cross-linguistic variation, task differences, or different contents being tested.
+	These studies raise the question if the divergent results are due to cross-linguistic variation, task differences, or different contents being tested.
 To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}.
-We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including purported factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}),
-	and 15 non-factive predicates, given recent findings that their complements are also projective, albeit to varying degrees (\citealt{degen_are_2022}).
+We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}),
+	and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
 
@@
 {\bf Results and Analysis.}
+	Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator.
+Mean projection ratings were higher under question-embeddings (Maximum likelihood estimate: $0.51$) than under negation (MLE: $0.48$) and modals ($0.47$), but lower than in conditional antecedents ($0.56$, see model \#1 in Table\ \ref{t:models}).
 	%
-	Our first key finding is that there is projection variability by operator: Across predicates, projection ratings were higher under question-embeddings than under negation and modals, but lower than in conditional antecedents.
-These differences, though small, are significant and supported by linear mixed effects model \textbf{\# 1} in \textbf{Table\ \ref{t:models}}.
+	This result differs from that of \cite{sieker_projective_2022}, who observed that the CC of German attitude predicates projects more from under negation than the other operators.
 	%
-	Second, there is by-predicate variation in the effect of operator on projection.
-This is illustrated in \textbf{Figure~\ref{fig:figure1}}, which shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
+	\textbf{(ii)} There is by-predicate variation in the effect of operator on projection.
+This is illustrated in Figure~\ref{fig:figure1}, which shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
 	%
-	For instance, whereas the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, the CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
-This finding is further supported by models \textbf{\# 2--4} in \textbf{Table\ \ref{t:models}} each having at least $34$ significant interaction terms (out of $57$ possible interactions of \texttt{operator} and \texttt{predicate}).
+	For instance, the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
 	%
-	% The models in \textbf{Table\ \ref{t:models}} also illustrate some specific differences across predicates in by-operator variation (using different reference levels for \texttt{predicate}): For instance, the CC of \emph{be annoyed} projects more from under negation (and questions) than conditionals and modals, while the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+	These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).
 	%
 	%
-	Contrary to what would be expected based on \posscite{karttunen_observations_1971}
-	distinction between factive and semi-factive predicates,
-	% the CC of assumed semi-factives \emph{discover, see} does not have higher projection-ratings under negation than other operators.
-\textbf{Figure~\ref{fig:figure1}} clearly shows that mean certainty ratings are lower for negation than conditionals and questions.
+	This result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals.
+This differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
 	%
-	the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
-The pattern we observed for {\em know} does not fit into either category.
+	Given that the methods and the set of projective contents of our experiment were very similar to theirs, this difference may suggest cross-linguistic variation in projection variability (cf.
+\citealt{tonhauser_projection_2020}).
 	%
+	On the other hand, in line with \citet{sieker_projective_2022}, our results are contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+The pattern observed for {\em know} does not fit into either category.
 	%
-	Our findings, that by-operator projection variability differs by content, are in line with \citet{smith_relationship_2014}, but contrary to their findings that the CC of \emph{know} projects more from negation than conditionals, we found no significant difference here (Model \textbf{\# 4} in \textbf{Table\ \ref{t:models}}).\\
 
 	%
-	\noindent{\bf LH:} Alternatively, we might only talk about the missed predictions for the semi-factives here? \\
-
-	\noindent Instead, we found that it projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant here.\\
-	%
-	%
-
-	Like in \posscite{sieker_projective_2022} study of German attitude predicates, we find a presence of a main effect of by-operator projection variation, but the particular effect we find is different: While Sieker and Solstad found highest projection ratings with negation compared to the other operators, we find overall highest projectivity across conditionals, then questions, then modals and negation.
-	%
-	With many interaction effects of \texttt{operator} and \texttt{predicate} throughout, our results also differ from , who did not find such interactions for the German attitudes they investigated.
-	%
-	Since our study used the same projection diagnostic, and a similar set of contents as \cite{sieker_projective_2022}, the different results indicate potential cross-linguistic variation in projection variability.
-This suggests
-	extending the questions addressed in cross-linguistic investigations of projective variability (\citealt{tonhauser_projection_2020,xue_correlation_2011,tonhauser_how_2018}) to interactions with various entailment-cancelling operators.
-
-	% \vspace{-.3\baselineskip}
-	% \textbf{version w baseline know/m}
-	% \noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{know} and \texttt{modal} as reference levels, and random intercepts for participants and items.
-We found highly significant main effects of \texttt{operator} (\texttt{conditional}: $+0.142$, \texttt{negation}: $+0.143$, ; \texttt{question}: $+0.225$, where $p < 2e^{-16}$ in all three cases), as well as many interactions of \texttt{operator} and  \texttt{verb} across the board (where $p < 0.001$ in $38$ cases, $p < 0.01$ in one, and $p < 0.05$ in three out of $57$ possible interactions).\\
-
-	%	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
-		% The mean for this baseline (intercept) is $0.867$.
-	%	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter.
-We also found many interactions of \texttt{operator} and \texttt{verb} across the board, suggesting that the effect of embedding context diffes by verb.
-Notably, \emph{\lq discover\rq} is more projective in polar questions and conditional antecedents than under negation, patterning opposite to Karttunen's claims about semi-factives.
-For the emotive predicate \emph{\lq be annoyed\rq} no significant effect is found for \texttt{question} (vs \texttt{negation}), as would be expected based on Karttunen, but we do find unexpected differences between \texttt{negation} $>$ \texttt{conditional, modal}.
-\emph{know} shows effects that would be incompatible with a characterization as either factive or semi-factive: $\texttt{question} > \texttt{conditional, negation} > \texttt{modal}$.
-If \emph{\lq know\rq} is a factive predicate, no difference would be expected.
-If it is semi-factive, we would, again, expect higher projectivity under negation than in questions and conditionals.
-
-	% for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$).
-We also found many interactions of \texttt{operator} and \texttt{verb} across the board (where $p < 0.001$ in $43$ cases, $p < 0.01$ in three, and $p < 0.05$ in one out of $57$ possible interactions).\\
 
@@
 
-{\bf Discussion.}
+\noindent {\bf Discussion.}
 	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
 	%
-	The first reason is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
-In \citealt{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
+	The first is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
+In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
 While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out.
 	%
-	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}, whose CCs are analyzed as presuppositions) and or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
+	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
 A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
-In the full talk, we discuss how this analysis might be able to capture the gradient projection observed in our experiment.
+In a full talk, we would like to discuss how this analysis might be extended to capture the gradient projection observed in our experiment.
 	%
@@
 	%
-	% {\bf THIS IS OLD, FROM ABOVE: So far, there has been no experimental investigation of by-operator variation comparing factive and semi-factive predicates.
-However, \citet{djarv_cognitive_2018} and \citet{tonhauser_how_2018} observed by-predicate variation in polar questions.
-\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, found higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-Here, it is also not obvious how exactly this task relates to projection.
-\citet{tonhauser_how_2018} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
-The observed differences did not match the expectations from Karttunen's classification (e.g., the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}).}
-	%
-	Our results further question the proposed difference between factive and semi-factive predicates (see also \citealp{beaver_have_2010}).
+	Our results further question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates (see also \citealt{beaver_have_2010}).
 	Future research appealing to these categories must clarify their definition. Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
@@
 %
-{\bf Beaver (2010).} Have you noticed that your belly button lint colour is related to the colour of your clothing? {\em Presuppositions and Discourse: Essays offered to Hans Kamp.}~\textbullet~
+{\bf Abrusán (2011).} Predicting the presuppositions of soft triggers. {\em Linguistics and philosophy.} ~\textbullet~
+% {\bf Beaver (2010).} Have you noticed that your belly button lint colour is related to the colour of your clothing? {\em Presuppositions and Discourse: Essays offered to Hans Kamp.}~\textbullet~
 {\bf Degen \& Tonhauser (2022).} Are there factive predicates? An empirical investigation. {\em Language.}~\textbullet~
 {\bf Djärv \& Bacovcin (2017).} Prosodic effects on factive presupposition projection. {\em Semantics and Linguistic Theory.}~\textbullet~
-{\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
+% {\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
+{\bf Heim (1983).} On the Projection Problem for Presuppositions. {\em Proceedings of the Second West Coast Conference on Formal Linguistics.}~\textbullet~
 % {\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
 {\bf Karttunen (1971).} Some observations on factivity. {\em Research on Language \& Social Interaction.} ~\textbullet~
-{\bf Mahler (2020).} The social component of projection behavior of clausal complements. {\em Linguistic Society of America.}~\textbullet~
+% {\bf Mahler (2020).} The social component of projection behavior of clausal complements. {\em Linguistic Society of America.}~\textbullet~
+% {\bf van der Sandt (1992).} Presupposition projection as anaphora resolution. {\em Journal of Semantics.}~\textbullet~
+{\bf Schlenker (2021).} Triggering presuppositions. {\em Glossa: A Journal of General Linguistics.}~\textbullet~
+{\bf Sieker \& Solstad (2022).} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study. {\em Proceedings of the 23rd Amsterdam Colloquium.}~\textbullet~
+% {\bf Simons, Beaver, Roberts \& Tonhauser (2017).} The best question: Explaining the projection behavior of factives. {\em Discourse processes.}~\textbullet~
 {\bf Smith and Hall (2014).} The relationship between projection and embedding environment. {\em Proceedings of the 48th Meeting of the Chicago Linguistics Society.}~\textbullet~
-{\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}~\textbullet~
+{\bf Tonhauser (2020).} Projection variability in Paraguayan Guaraní. {\em Natural Language \& Linguistic Theory.}~\textbullet~
+{\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}
 % {\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
@@
 	\centering
-	\hspace{-1.2em}
-	\begin{tabular}{p{.663\linewidth} p{.3\linewidth}}
-		\scalebox{.77}{
+	\hspace{-1.3em}
+	\begin{tabular}{p{.65\linewidth} p{.31\linewidth}}
+		\scalebox{.75}{
 			\begin{tabular}{llrrrr}
@@
 				\midrule
-				\#1 & Intercept: \emph{question} & 0.52 & 0.01 & 87.31 & ***\\
-				& operator: conditional & 0.04 & 0.01 & 5.07 & ***\\
-				& operator: modal & -0.04 & 0.01 & -4.55 & ***\\
-				& operator: negation & -0.03 & 0.01 & -4.30 & ***\\
+				\#1 & Intercept: \emph{question} & 0.51 & 0.01 & 44.78 & ***\\
+				& operator: conditional & 0.05 & 0.01 & 5.30 & ***\\
+				& operator: modal & -0.04 & 0.01 & -4.45 & ***\\
+				& operator: negation & -0.03 & 0.01 & -4.67 & ***\\
 				\midrule
@@
 		&
-		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; \textbf{\#1} has fixed effects of operator; random effect: participant intercepts, \textbf{\#2-4} have fixed effect: operator, predicate, and their interaction; random effect: participant intercepts.
+		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; \textbf{\#1} has fixed effects of operator; random effects: participant and item intercepts, \textbf{\#2-4} have fixed effect: operator, predicate, and their interaction; random effect: participant intercepts.
 		Models were fit with \texttt{lme4, lmertest} in \texttt{R}.
-All three models also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (Not shown here).}\label{t:models}}
+Models \textbf{\#2--4} also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (not shown here).}\label{t:models}}
 

### 2023-03-15 results/main/meta-analyses/new/1_projectivity/rscripts/analysis.R

@@ -27,15 +27,16 @@ data$op <- as.factor(data$op)
 levels(data$op)
 contrasts(data$op)
 ### op dummy coding with "n" (negation) as baseline
-# data$op <- relevel(data$op, ref = "q")
-# contrasts(data$op)
+data$op <- relevel(data$op, ref = "q")
+contrasts(data$op)
 
 
 # coding random effects as factors
 data$workerid <- as.factor(data$workerid)
-# data$content <- as.factor(data$content) 
+data$item <- as.factor(paste(data$content, data$pred))
 ## content should not be a random effect, since contents are uniquely paired w
 ## preds, which are independent variable
+## nmo
 
 # exploring a couple of models (linear)
 ## library for linear mixed models
@@ -43,10 +44,11 @@ library(lme4)
 library(lmerTest)
 library(knitr)
 
-# linear model w op as predictor only
-# glmm0 <- lmer(projective ~ op + (1 | workerid), data = data)
+# linear model w op as predictor only; q baseline
+glmm0 <- lmer(projective ~ op + (1 | workerid) + (1 | item), data = data)
 summary0 <- summary(glmm0)
 print(summary0, cor=F, dig=3)
+kable(summary0$coefficients, format = "latex", booktabs = TRUE, dig=2)
 
 ### op dummy coding with "n" (negation) as baseline
 data$op <- relevel(data$op, ref = "n")

### 2023-04-17 writing/XPRAG-abstract/XPRAG-abstract.tex
added XPRAG abstract

--- 2023-03-15 writing/SuB-abstract/SuB-abstract.tex
@@
 
-\documentclass[12pt, a4paper]{article}
+\documentclass[11pt, a4paper]{article}
 
@@
 
-% \usepackage{euler} % math font package needs to be loaded before others
+\usepackage{euler} % math font package needs to be loaded before others
 \usepackage{xunicode,xltxtra, polyglossia}
@@
 	\usepackage[dvipsnames]{xcolor}
+	\definecolor{orangee}{HTML}{E69F00}
+	\definecolor{greenn}{HTML}{009E73}
+	\definecolor{bluee}{HTML}{56B4E9}
 
 	%% fonts
-	\setmainfont[Scale=MatchLowercase,Mapping=tex-text, SmallCapsFeatures={Letters=SmallCaps}]{Times New Roman}
-	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Times New Roman}
+	\setmainfont[Scale=MatchLowercase,Mapping=tex-text, SmallCapsFeatures={Letters=SmallCaps}]{Arial}
+	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Arial}
 	\setmonofont[Scale=MatchLowercase]{Andale Mono}
@@
 	\textbf{\large%\thetitle
-		Projection variability of attitude complements across different operators}
+		Projection variability of complement clauses across different operators}
 \end{center}
 
-\vspace{-.8\baselineskip}
-\noindent 
-	We present experimental evidence that the projection of the content of attitude complements \textbf{(i)} varies between entailment-canceling operators, and \textbf{(ii)} that this by-operator variation differs between attitude predicates.
+\paragraph{Short summary:} % (fold)
+	We examine how the content of embedded complement clauses projects across different entailment-canceling operators for various clause-embedding predicates.
 	%
-	The observed variability is not captured by existing theoretical accounts of projection
-	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
+	We measured projection for 20 predicates across four operators: polar questions, negation, modals, and conditionals.
+Results reveal projection variation by operator and by-predicate variation in the effect of operator on projection.
 	%
-	Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.
+	Existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}) do not capture the observed variability.
+	%
+	Our findings highlight the importance of considering interactions between predicates and operators and generate important questions for future research on projection.
 
-\noindent {\bf Projection across entailment-cancelling operators.}
-	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
-	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \emph{projects}.
+% paragraph short_summary (end)
 
-	\vspace{-.5\baselineskip}
+\paragraph{Projection across entailment-cancelling operators.}
+	Interpreters may infer that a speaker uttering a sentence with an embedded complement clause \ref{ex:family} is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
+	even when embedded under an entailment-canceling operator \Next[a-d]. Then we say that the CC \emph{projects}.
+
 	\ex. \label{ex:family}
-		\a. \label{ex:neg}
+		\a. \label{ex:q}
+			{\bf Polar Question:} \hfill
+			\emph{Did Cole discover that Julian dances salsa?}
+		\b. \label{ex:neg}
 			{\bf Negation:} \hfill
-			\emph{\lq Cole didn't discover that Julian dances salsa.\rq}
-		\b. \label{ex:q}
-			{\bf Polar Question:} \hfill
-			\emph{\lq Did Cole discover that Julian dances salsa?\rq}
+			\emph{Cole didn't discover that Julian dances salsa.}
 		\c. \label{ex:mod}
 			{\bf Modal:} \hfill
-			\emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
+			\emph{Perhaps Cole discovered that Julian dances salsa.}
 		\d. \label{ex:cond}
 			{\bf Conditional:} \hfill
-			\emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
+			\emph{If Cole discovered that Julian dances salsa, Logan will be joyful.}
 		\z.
@@
 
-	\vspace{-.5\baselineskip}
-	\noindent  Current research rarely examines projection variability across these operators, with conflicting findings.
+	\noindent
+	Previous research has investigated projection variability across these operators only rarely, and with conflicting findings.
 	%
-	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it always projects across negation, but not always for the other operators.
+	For English clause-embedding predicates, \citet{karttunen_observations_1971} proposed a distinction between factives (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it always projects across negation, but not always across the other operators.
 	%
-	Comparing different types of contents, \citet{smith_relationship_2014} found that the projective content of epithets and the CC of \emph{know} was more projective under negation than conditionals,
-	whereas that of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
+	\cite{sieker_projective_2022}, investigating the projection of German clause-embedding predicates across these operators, found an overall pattern of more projection across negation than the others, but no operator/predicate interaction effect.
 	%
-	This result was replicated for German \emph{wissen} `know' in \posscite{sieker_projective_2022} study on German attitude predicates, but as part of an overall pattern of higher projection ratings with negation than other operators.
-They found no interaction with predicate type, or evidence for the distinction of factives vs.
-semi-factives.
+	%
+	In contrast, \citet{smith_relationship_2014} found different projectivity patterns for different types of English projective contents: That of epithets and the CC of \emph{know} were more projective under negation than conditionals, whereas that of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
+	%
 
-	These studies raise the question if the divergent results are due to cross-linguistic variation, task differences, or different contents being tested.
+	Because the studies diverge about whether they find operator/predicate interaction effects, the question arises whether they differ due to cross-linguistic variation, task differences, or different types of contents being tested.
 To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}.
@@
 
-\noindent 
-{\bf Method.}
-	Projection of the CC of the 20 attitude predicates was measured in four sets of experiments: The predicates were embedded in polar questions in Exps.\ 1, under negation (Exps.\ 2), under {\em perhaps} (Exps.\ 3), and in conditional antecedents (Exps.\ 4).
-(Each set contained three experiments using different at-issueness measure in a separate block.
+\paragraph{Method.}
+	Projection of the CC of the 20 clause-embedding predicates was measured in four sets of experiments: Predicates were embedded in polar questions in Exps.\ 1, under negation (Exps.\ 2), under a modal {\em perhaps} (Exps.\ 3), and in conditional antecedents (Exps.\ 4).
+(Each set contained three experiments using different at-issueness measures in a separate block.
 Here, we focus on the projection ratings.)
 	%
-	In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
-Participants gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
-Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
-We analyze data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform.
-	% Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.
+	Participants read utterances like in \ref{ex:family} and rated whether the speaker (who was named) was certain of the CC (e.g.: \emph{\lq Is [the speaker] certain that Julian dances salsa?\rq}), on a slider marked \emph{`no'} (coded as \texttt{0}) at one end and \emph{`yes'} (coded as \texttt{1}) on the other.
+Each participant saw all 20 clause-embedding predicates (each with a unique content from a set of 20) under one operator.
+We analyze data from $2,682$ self-reported native speakers of American English, recruited on Prolific or Amazon's MT platform.
 
-\noindent
-{\bf Results and Analysis.}
-	Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator.
-Mean projection ratings were higher under question-embeddings (Maximum likelihood estimate: $0.51$) than under negation (MLE: $0.48$) and modals ($0.47$), but lower than in conditional antecedents ($0.56$, see model \#1 in Table\ \ref{t:models}).
+\begin{figure}[ht]
+	\vspace{-.5\baselineskip}
+	\centering
+	\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}
+	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
+Entailment-cancelling operator coded by letter and color:  \textcolor{orangee}{\texttt{N}: negation}, \texttt{M}: modals, \textcolor{greenn}{\texttt{C}: conditional antecedents}, \textcolor{bluee}{\texttt{Q}: polar questions}.}
+	\label{fig:figure1}
+\end{figure}
+
+\paragraph{Results and Analysis.}
+	Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator, with higher mean projection ratings for question-operators (Maximum likelihood estimate: $0.51$) than negation (MLE: $0.48$) and modals ($0.47$), but lower than for conditional antecedents ($0.56$, see model \#1 in Table\ \ref{t:models}).
 	%
-	This result differs from that of \cite{sieker_projective_2022}, who observed that the CC of German attitude predicates projects more from under negation than the other operators.
+	This differs from \posscite{sieker_projective_2022} observation that the CC of German clause-embedding predicates projects more across negation than the other operators, suggesting potential cross-linguistic variation in the projection of clausal complements across entailment-canceling operators.
+
+	\textbf{(ii)} There is by-predicate variation in the effect of operator on projection, as illustrated in Figure~\ref{fig:figure1}, showing mean projection ratings for the 20 clause-embedding predicates by operator, with predicates ordered by their mean rating across operators (\emph{be annoyed} has the highest overall mean).
 	%
-	\textbf{(ii)} There is by-predicate variation in the effect of operator on projection.
-This is illustrated in Figure~\ref{fig:figure1}, which shows mean projection ratings for the 20 attitude predicates by operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
-	%
-	For instance, the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
-The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
+	For instance, the CC of \emph{be annoyed} projects more across negation and questions than conditionals and modals, and the CC of \emph{discover} projects less across negation than conditionals and questions, but more across negation than modals.
+The CC of \emph{know} projects less across negation than questions, but more across negation than modals, while the difference between negation and conditionals is not significant.
 	%
@@
 	%
+
+	Our finding of operator/predicate-interactions concurs with \citet{smith_projection_2011}, but we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals.
 	%
-	This result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals.
-This differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
-	%
-	Given that the methods and the set of projective contents of our experiment were very similar to theirs, this difference may suggest cross-linguistic variation in projection variability (cf.
+	Further, our finding differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
+Given that the methods and the set of projective contents of our experiment were very similar to theirs, this may suggest cross-linguistic variation in projection variability (cf.
 \citealt{tonhauser_projection_2020}).
 	%
-	On the other hand, in line with \citet{sieker_projective_2022}, our results are contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+	Additionally, our results are contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably across four operators, and the CC of \emph{discover,}  considered semi-factive, does not project more from under negation than the other three operators.
 The pattern observed for {\em know} does not fit into either category.
-	%
-
-	%
 
@@
 
-\noindent {\bf Discussion.}
-	Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
+\paragraph{Discussion.}
+	Our results---that various entailment-canceling operators affect projection differently and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
 	%
-	The first is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
-In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
-While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out.
+	First, they do not predict variation associated with different types of entailment-cancelling operators.
+In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency and the CC is accommodated locally.
+While it is conceivable for operator meanings to systematically affect the possibility of local accommodation, this has not been spelled out.
 	%
-	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
-A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
-In a full talk, we would like to discuss how this analysis might be extended to capture the gradient projection observed in our experiment.
+	The second reason is that many analyses only focus on (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
+This limits their predictive power for other predicates.
 	%
-	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
-Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
+	\posscite{schlenker_triggering_2021} may be an exception, which predicts the potential of projection for CCs that are contextually entailed.
+In the full paper, we discuss how Schlenker's analysis may be extended to capture the gradient projection observed in our experiment.
+	%
+	Finally, no current analysis makes sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is presupposed/entailed or not).
+Consequently, precluding predictions about the by-predicate variation in operator effects.
 
-% \noindent {\bf Discussion:  Empirical implications and lexical properties.}
-	%
-	Our results further question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates (see also \citealt{beaver_have_2010}).
-	Future research appealing to these categories must clarify their definition. Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
+	Our results further question \posscite{karttunen_observations_1971} proposed distinction between factive and semi-factive predicates (see also \citealt{beaver_have_2010,sieker_projective_2022}).
+	Future research appealing to these categories must clarify their definition. 
+	% Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
 	% Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
 	%
-	Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
+	Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{\lq inform\rq} and \emph{\lq acknowledge\rq}, for instance, are at least as projective as those of some (semi-)factive predicates.
 	%
@@
 	%
-	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting initial generalizations over lexical properties, indicated in \textbf{Figure\ 2}.
+	In a full presentation, we would discuss how more fine-grained lexical distinctions may affect projection variability: For example, the CCs of \emph{\lq think\rq} and \emph{\lq pretend\rq} (which are often used anti-veridically) project more across negation than other operators.
+The CCs of \emph{\lq reveal, confess, admit\rq} and \emph{\lq announce\rq} (communicative achievement predicates) project most from conditional antecendents.
 
@@
 
-\noindent
-
-\newpage
-{\scriptsize
-\noindent {\bf (Selected) References:}
-%
-{\bf Abrusán (2011).} Predicting the presuppositions of soft triggers. {\em Linguistics and philosophy.} ~\textbullet~
-% {\bf Beaver (2010).} Have you noticed that your belly button lint colour is related to the colour of your clothing? {\em Presuppositions and Discourse: Essays offered to Hans Kamp.}~\textbullet~
-{\bf Degen \& Tonhauser (2022).} Are there factive predicates? An empirical investigation. {\em Language.}~\textbullet~
-{\bf Djärv \& Bacovcin (2017).} Prosodic effects on factive presupposition projection. {\em Semantics and Linguistic Theory.}~\textbullet~
-% {\bf Djärv, Zehr \& Schwarz (2018).} Cognitive vs. emotive factives: An experimental differentiation. {\em Proceedings of Sinn und Bedeutung.}~\textbullet~
-{\bf Heim (1983).} On the Projection Problem for Presuppositions. {\em Proceedings of the Second West Coast Conference on Formal Linguistics.}~\textbullet~
-% {\bf Hooper \& Thompson (1973).} On the applicability of root transformations. {\em Linguistic Inquiry.} ~\textbullet~
-{\bf Karttunen (1971).} Some observations on factivity. {\em Research on Language \& Social Interaction.} ~\textbullet~
-% {\bf Mahler (2020).} The social component of projection behavior of clausal complements. {\em Linguistic Society of America.}~\textbullet~
-% {\bf van der Sandt (1992).} Presupposition projection as anaphora resolution. {\em Journal of Semantics.}~\textbullet~
-{\bf Schlenker (2021).} Triggering presuppositions. {\em Glossa: A Journal of General Linguistics.}~\textbullet~
-{\bf Sieker \& Solstad (2022).} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study. {\em Proceedings of the 23rd Amsterdam Colloquium.}~\textbullet~
-% {\bf Simons, Beaver, Roberts \& Tonhauser (2017).} The best question: Explaining the projection behavior of factives. {\em Discourse processes.}~\textbullet~
-{\bf Smith and Hall (2014).} The relationship between projection and embedding environment. {\em Proceedings of the 48th Meeting of the Chicago Linguistics Society.}~\textbullet~
-{\bf Tonhauser (2020).} Projection variability in Paraguayan Guaraní. {\em Natural Language \& Linguistic Theory.}~\textbullet~
-{\bf Tonhauser, Beaver, \& Degen (2018).} How projective is projective content? Gradience in projectivity and at-issueness. {\em Journal of Semantics.}
-% {\bf Tönnis (2021).} {\em German es-clefts in discourse. A question-based analysis involving expectedness.} Doctoral dissertation, Graz University.
-% {\bf Zimmermann (2011).} {\em The grammatical expression of focus in West Chadic: Variation and uniformity in and across languages.} Walter de Gruyter.
-
-\begin{figure}[h]
-	\vspace{-.5\baselineskip}
+\begin{table}[h]
 	\centering
-	\scalebox{.9}{\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}}\vspace{-1.8\baselineskip}
-	\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
-Embedding operator coded by letter and color:  \texttt{N} (orange): negation, \texttt{M} (black): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (blue): polar questions.}
-	\label{fig:figure1}
-\end{figure}
-
-\begin{table}[h]
-	\vspace{-1\baselineskip}\phantom{.}
-	\centering
-	\hspace{-1.3em}
-	\begin{tabular}{p{.65\linewidth} p{.31\linewidth}}
-		\scalebox{.75}{
 	\begin{tabular}{llrrrr}
@@
 	\end{tabular}
-		}
-		&
-		\parbox{\linewidth}{\caption{\small Excerpt of the output from three linear mixed effects models; \textbf{\#1} has fixed effects of operator; random effects: participant and item intercepts, \textbf{\#2-4} have fixed effect: operator, predicate, and their interaction; random effect: participant intercepts.
+	\caption{\small Excerpt of the output from three linear mixed effects models; \textbf{\#1} has fixed effects of operator; random effects: participant and item intercepts, \textbf{\#2-4} have fixed effect: operator, predicate, and their interaction; random effect: participant intercepts.
 		Models were fit with \texttt{lme4, lmertest} in \texttt{R}.
-Models \textbf{\#2--4} also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (not shown here).}\label{t:models}}
-
-	\end{tabular}
+Models \textbf{\#2--4} also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (not shown here).}\label{t:models}
 \end{table}
 
-\hspace{-2.7em}
-	\begin{minipage}{\textwidth}\centering
-	\begin{tabular}{p{.235\linewidth} p{.235\linewidth} p{.235\linewidth} p{.235\linewidth}}
-		\parbox{\linewidth}{\centering (a) \textbf{Negation high}\\
-		\includegraphics[width=.26\textwidth]{graphs/profile1.pdf}}
-		&
-		\parbox{\linewidth}{\centering (b) \textbf{Negation low}\\
-		\includegraphics[width=.26\textwidth]{graphs/profile4.pdf}}
-		&
-		\parbox{\linewidth}{\centering (c) \textbf{Conditional high}\\
-		\includegraphics[width=.26\textwidth]{graphs/profile2.pdf}}
-		&
-		\parbox{\linewidth}{\centering (d) \textbf{Modal low}\\
-		\includegraphics[width=.26\textwidth]{graphs/profile5.pdf}}
-	\end{tabular}
-\end{minipage}
+% \hspace{-2.7em}
+% 	\begin{minipage}{\textwidth}\centering
+% 	\begin{tabular}{p{.5\linewidth} p{.5\linewidth}}
+% 		\parbox{\linewidth}{\centering (a) \textbf{Negation high}\\
+% 		\includegraphics[width=.8\linewidth]{graphs/profile1.pdf}}
+% 		&
+% 		\parbox{\linewidth}{\centering (b) \textbf{Negation low}\\
+% 		\includegraphics[width=.8\linewidth]{graphs/profile4.pdf}}
+% 		\\
+% 		\parbox{\linewidth}{\centering (c) \textbf{Conditional high}\\
+% 		\includegraphics[width=.8\linewidth]{graphs/profile2.pdf}}
+% 		&
+% 		\parbox{\linewidth}{\centering (d) \textbf{Modal low}\\
+% 		\includegraphics[width=.8\linewidth]{graphs/profile5.pdf}}
+% 		\\
+% 	\end{tabular}
+% \end{minipage}
 
-\vspace{-.4\baselineskip}
-\noindent \small Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
+% \vspace{-.4\baselineskip}
+% \noindent \small Figure 2: \small Mean certainty ratings by operator (\texttt{M}: Modal, \texttt{N}: Negation, \texttt{Q}: Polar Question, \texttt{C}: Conditional antecedent) with 95\% bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).
 

### 2023-04-18 writing/XPRAG-abstract/XPRAG-abstract.tex
cut down to 1200 words

@@
 	Interpreters may infer that a speaker uttering a sentence with an embedded complement clause \ref{ex:family} is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
-	even when embedded under an entailment-canceling operator \Next[a-d]. Then we say that the CC \emph{projects}.
+	even when embedded under an entailment-canceling operator \Next[a-d]. Then we say the CC \emph{projects}.
 
@@
 	\noindent
-	Previous research has investigated projection variability across these operators only rarely, and with conflicting findings.
+	Previous research has investigated differences in projection across these operators only rarely, and with conflicting findings.
 	%
-	For English clause-embedding predicates, \citet{karttunen_observations_1971} proposed a distinction between factives (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it always projects across negation, but not always across the other operators.
+	For English clause-embedding predicates, \citet{karttunen_observations_1971} proposed distinguishing factives (e.g., \emph{be annoyed, regret}), where the CC projects across all four operators, and semi-factives (e.g., \emph{discover, realize, see, notice}) where it projects across negation, but not always across the other operators.
 	%
-	\cite{sieker_projective_2022}, investigating the projection of German clause-embedding predicates across these operators, found an overall pattern of more projection across negation than the others, but no operator/predicate interaction effect.
+	For German clause-embedding predicates,
+	\cite{sieker_projective_2022} found an overall pattern of more projection across negation than the other four operators, but no operator/predicate interactions.
 	%
 	%
-	In contrast, \citet{smith_relationship_2014} found different projectivity patterns for different types of English projective contents: That of epithets and the CC of \emph{know} were more projective under negation than conditionals, whereas that of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
+	In contrast, \citet{smith_relationship_2014} found different patterns for different English projective contents: That of epithets and the CC of \emph{know} were more projective under negation than conditionals, whereas that of non-restrictive relative clauses and \emph{win} showed the opposite pattern.
 	%
 
-	Because the studies diverge about whether they find operator/predicate interaction effects, the question arises whether they differ due to cross-linguistic variation, task differences, or different types of contents being tested.
-To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}.
-We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}),
-	and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
+	These diverging findings about whether there are operator/predicate interaction effects raise the question whether differences result from cross-linguistic variation, task differences, or different types of contents being tested.
+To address this, we conducted a series of experiments designed to assess projection across the four operators in \ref{ex:family}.
+We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factives (\emph{be annoyed, know, reveal}) semi-factives (\emph{discover, see}),
+	and 15 non-factive predicates, given recent findings that their complements also project to varying degrees (\citealt{degen_are_2022}).
 
 \paragraph{Method.}
-	Projection of the CC of the 20 clause-embedding predicates was measured in four sets of experiments: Predicates were embedded in polar questions in Exps.\ 1, under negation (Exps.\ 2), under a modal {\em perhaps} (Exps.\ 3), and in conditional antecedents (Exps.\ 4).
+	Projection of the CC of 20 clause-embedding predicates was measured in four sets of experiments: Predicates were embedded in polar questions in Exps.\ 1, under negation (Exps.\ 2), under the modal {\em \lq perhaps\rq} (Exps.\ 3), and in conditional antecedents (Exps.\ 4).
 (Each set contained three experiments using different at-issueness measures in a separate block.
-Here, we focus on the projection ratings.)
+Here, we focus on the projection-ratings.)
 	%
-	Participants read utterances like in \ref{ex:family} and rated whether the speaker (who was named) was certain of the CC (e.g.: \emph{\lq Is [the speaker] certain that Julian dances salsa?\rq}), on a slider marked \emph{`no'} (coded as \texttt{0}) at one end and \emph{`yes'} (coded as \texttt{1}) on the other.
-Each participant saw all 20 clause-embedding predicates (each with a unique content from a set of 20) under one operator.
-We analyze data from $2,682$ self-reported native speakers of American English, recruited on Prolific or Amazon's MT platform.
+	Participants read utterances like in \ref{ex:family} and rated whether the speaker (who was named) was certain of the CC (e.g.: \emph{\lq Is [the speaker] certain that Julian dances salsa?\rq}), on a slider marked \emph{`no'} (coded as \texttt{0}) at one end and \emph{`yes'} (coded as \texttt{1}) at another.
+Each participant saw all 20 clause-embedding predicates (each paired with a unique content) under one operator.
+We analyze data from $2,682$ self-reported native speakers of American English.
 
@@
 \paragraph{Results and Analysis.}
-	Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator, with higher mean projection ratings for question-operators (Maximum likelihood estimate: $0.51$) than negation (MLE: $0.48$) and modals ($0.47$), but lower than for conditional antecedents ($0.56$, see model \#1 in Table\ \ref{t:models}).
+	Our analysis reveals two key results: \textbf{(i)} Projection varies by operator, with higher mean certainty-ratings for question-operators (Maximum likelihood estimate: $0.51$) than negation (MLE: $0.48$) and modals ($0.47$), but lower than for conditionals ($0.56$, see model \#1 in Table\ \ref{t:models}).
 	%
-	This differs from \posscite{sieker_projective_2022} observation that the CC of German clause-embedding predicates projects more across negation than the other operators, suggesting potential cross-linguistic variation in the projection of clausal complements across entailment-canceling operators.
+	This differs from \posscite{sieker_projective_2022} observation that the CC of German clause-embedding predicates projects more across negation than the other operators, suggesting potential cross-linguistic differences in projection of CCs across entailment-canceling operators.
 
-	\textbf{(ii)} There is by-predicate variation in the effect of operator on projection, as illustrated in Figure~\ref{fig:figure1}, showing mean projection ratings for the 20 clause-embedding predicates by operator, with predicates ordered by their mean rating across operators (\emph{be annoyed} has the highest overall mean).
+	\textbf{(ii)} There is by-predicate variation in the effect of operator on projection, illustrated in Figure~\ref{fig:figure1}, which shows mean projection-ratings for the 20 clause-embedding predicates by operator, with predicates ordered by their mean rating across operators (\emph{be annoyed} has the highest overall mean).
 	%
@@
 	%
-	These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).
+	These generalizations are supported by models\ \#2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).
 	%
 
-	Our finding of operator/predicate-interactions concurs with \citet{smith_projection_2011}, but we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals.
+	Our finding of operator/predicate-interactions concurs with \citet{smith_projection_2011}, but we did not replicate their result that the CC of \emph{know} projects more from negation than conditionals.
 	%
-	Further, our finding differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
-Given that the methods and the set of projective contents of our experiment were very similar to theirs, this may suggest cross-linguistic variation in projection variability (cf.
+	Our results also differ from \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
+Given that our task and materials were very similar to theirs, this may suggest cross-linguistic variation in projection variability (cf.
 \citealt{tonhauser_projection_2020}).
 	%
-	Additionally, our results are contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably across four operators, and the CC of \emph{discover,}  considered semi-factive, does not project more from under negation than the other three operators.
+	Additionally, our results contradict \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates: The CC of (factive) \emph{be annoyed} does not project invariably across four operators, and the CC of \emph{discover,}  considered semi-factive, does not project more from under negation than the other three operators.
 The pattern observed for {\em know} does not fit into either category.
@@
 \paragraph{Discussion.}
-	Our results---that various entailment-canceling operators affect projection differently and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
+	Our results---that various entailment-canceling operators affect projection differently and there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses, for several reasons.
 	%
-	First, they do not predict variation associated with different types of entailment-cancelling operators.
-In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency and the CC is accommodated locally.
-While it is conceivable for operator meanings to systematically affect the possibility of local accommodation, this has not been spelled out.
+	First, current theories do not predict variation along different types of entailment-cancelling operators.
+In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would be inconsistent, and the CC is accommodated locally.
+While it is conceivable that operator meanings systematically affect the possibility of local accommodation, this has not been spelled out.
 	%
-	The second reason is that many analyses only focus on (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
+	Second, many analyses only focus on (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
 This limits their predictive power for other predicates.
 	%
-	\posscite{schlenker_triggering_2021} may be an exception, which predicts the potential of projection for CCs that are contextually entailed.
-In the full paper, we discuss how Schlenker's analysis may be extended to capture the gradient projection observed in our experiment.
+	\citet{schlenker_triggering_2021} may be an exception, predicting the potential of projection for contextually entailed CCs.
+In the full paper, we discuss how Schlenker's analysis may be extended to capture the observed gradient patterns.
 	%
-	Finally, no current analysis makes sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is presupposed/entailed or not).
-Consequently, precluding predictions about the by-predicate variation in operator effects.
+	Finally, no analysis makes sufficiently fine-grained distinctions between clause-embedding predicates (only distinguishing whether the CC is presupposed/entailed or not), precluding predictions about the by-predicate variation in operator effects.
 
-	Our results further question \posscite{karttunen_observations_1971} proposed distinction between factive and semi-factive predicates (see also \citealt{beaver_have_2010,sieker_projective_2022}).
-	Future research appealing to these categories must clarify their definition. 
-	% Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
-	% Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
+	Our results further question the effect of categorical lexical distinctions on projection: They challenge \posscite{karttunen_observations_1971} factive/semi-factive distinction (see also \citealt{beaver_have_2010,sieker_projective_2022}), and provide support (from negation, modals, and conditionals) for \posscite{degen_are_2022} result, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{\lq inform\rq} and \emph{\lq acknowledge\rq}, for instance, are at least as projective as those of some (semi-)factive predicates.
 	%
-	Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates: The CCs of \emph{\lq inform\rq} and \emph{\lq acknowledge\rq}, for instance, are at least as projective as those of some (semi-)factive predicates.
-	%
-	% Further, the variability we found is gradient, confirming previous research that by-predicate variability is not determined by categorical lexical classes (\citealt{tonhauser_how_2018,degen_are_2022}),
-	% and extending this result to by-predicate variability in interaction with entailment-cancelling contexts.
-	%
-	In a full presentation, we would discuss how more fine-grained lexical distinctions may affect projection variability: For example, the CCs of \emph{\lq think\rq} and \emph{\lq pretend\rq} (which are often used anti-veridically) project more across negation than other operators.
-The CCs of \emph{\lq reveal, confess, admit\rq} and \emph{\lq announce\rq} (communicative achievement predicates) project most from conditional antecendents.
+	In the full version, we discuss how more fine-grained lexical distinctions may affect projection: For example, the CCs of (often used anti-veridically) \emph{\lq think\rq} and \emph{\lq pretend\rq} project more across negation than other operators.
+The CCs of \emph{\lq reveal, confess, admit\rq} and \emph{\lq announce\rq} (communicative achievement predicates) project most from conditionals.
 
@@
 	\caption{\small Excerpt of the output from three linear mixed effects models; \textbf{\#1} has fixed effects of operator; random effects: participant and item intercepts, \textbf{\#2-4} have fixed effect: operator, predicate, and their interaction; random effect: participant intercepts.
-		Models were fit with \texttt{lme4, lmertest} in \texttt{R}.
+		Models were fit with \texttt{lme4, lmertest} in \texttt{R}. Significance code \lq***\rq\ indicates that $p < 0.001$.
 		Models \textbf{\#2--4} also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (not shown here).}\label{t:models}

### 2023-07-15 writing/DGfS2023-abstract/write-up.tex
ran exp13 and started analysis

\documentclass[11pt,fleqn]{article}
\usepackage[margin=1in,top=1in,bottom=1in]{geometry}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{hyperref}
%\usepackage[dvips]{graphics}
%\usepackage[table]{xcolor}
%\usepackage{amssymb}
\usepackage{float}
%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{subcaption}

\usepackage[normalem]{ulem}

\usepackage{multicol}
\usepackage{txfonts}
\usepackage{amsfonts}
\usepackage{natbib}
\usepackage{gb4e}
\usepackage[all]{xy}
\usepackage{rotating}
\usepackage{tipa}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{url}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{array}

\def\bad{{\leavevmode\llap{*}}}
\def\marginal{{\leavevmode\llap{''}}}
\def\verymarginal{{\leavevmode\llap{''''}}}
\def\swmarginal{{\leavevmode\llap{4}}}
\def\infelic{{\leavevmode\llap{\#}}}

\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
%\definecolor{gray}{rgb}{0.36, 0.54, 0.66}

\definecolor{Pink}{RGB}{240,0,120}
\newcommand{\red}[1]{\textcolor{Pink}{#1}}
\newcommand{\jd}[1]{\textbf{\textcolor{Pink}{[jd: #1]}}}

\newcommand{\dashrule}[1][black]{%
  \color{#1}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}\xleaders\hbox{\rule{4pt}{0pt}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}}\hfill\kern0pt%
}

\setlength{\parindent}{.3in}
\setlength{\parskip}{0ex}

\newcommand{\yi}{\'{\symbol{16}}}
\newcommand{\nasi}{\~{\symbol{16}}}
\newcommand{\hina}{h\nasi na}
\newcommand{\ina}{\nasi na}

\newcommand{\foc}{$_{\mbox{\small F}}$}

\hyphenation{par-ti-ci-pa-tion}

\setlength{\bibhang}{0.5in}
\setlength{\bibsep}{0mm}
\bibpunct[:]{(}{)}{,}{a}{}{,}

\newcommand{\6}{\mbox{$[\hspace*{-.6mm}[$}} 
\newcommand{\9}{\mbox{$]\hspace*{-.6mm}]$}}
\newcommand{\sem}[2]{\6#1\9$^{#2}$}
\renewcommand{\ni}{\~{\i}}

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{#1}}
\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{#1})}

\newcolumntype{R}[2]{%
    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
    l%
    <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{90}{0em}}}% no optional argument here, please!

\title{Exp 13 write-up}

\author{Judith Tonhauser}

\newcommand{\jt}[1]{\textbf{\color{blue}JT: #1}}

\begin{document}

\maketitle

\vspace*{-1cm}

\section{Introduction}\label{s-intro}

On the basis of observed gradient mean certainty ratings for the contents of the complements (CCs) of 20 clause-embedding predicates, \citealt{degen-tonhauser-openmind,degen-tonhauser-language} suggested:

\begin{itemize}
\item[a)] projection is gradient, that is, it is not the case that content either projects or it doesn't, but rather content projects to varying degrees, reflecting either varying degrees of speaker commitment or varying degrees of certainty of the interpreter on the speaker's commitment to a content, and

\item[b)] that there is no empirical support (to date) for a binary division between factive and non-factive predicates, that is, between predicates that conventionally code that the CC is presupposed and predicates that do not code such a requirement.
\end{itemize}

 \citealt[\S6.2]{mandelkern-etal2020} argued that the observed gradience is due to the inference task used. They suggested that acceptability ratings in explicit ignorance contexts are better suited to investigating projection and, specifically, to distinguishing presuppositions (conventionally specified projective content) from other projection inferences: ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).

More specifically, \citealt{mandelkern-etal2020} argued that just because participants ``tend to draw certain inferences from clauses appearing in embedded, entailment-canceling environments, does not necessarily mean that the relevant content has projected in the sense that theorists of presupposition should care about'' (p.497). They suggest that it ``may well be that inference tasks to some degree invite subjects to make the inference in question if it is compatible with the assertion, since the content in question is right there before the subjects' eyes'' (p.498). They argued that acceptability ratings of sentences presented in explicit ignorance contexts are better suited to distinguish conventionally specified projective content from other projection inferences: if participants ``draw an inference just because it is a natural conclusion to draw for any of a variety of pragmatic reasons short of entailment or presupposition, then they will tend to relinquish that inclination when there is pragmatic pressure to do so [...]. By contrast, if the inference in question is a semantic presupposition, this will not be a possibility: beyond the (marked) rescue mechanism of local accommodation, subjects will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).

p.497: ``In methodological terms, we strongly recommend at least a two-pronged approach, with careful attention paid to results stemming from the evaluation of acceptabilty [sic] in different contexts.''

We conducted an experiment that was designed to investigate \citepos{mandelkern-etal2020} claim that gradient projection is due inference ratings and will not be observed acceptability ratings in explicit ignorance contexts. We hypothesize that acceptability ratings in explicit ignorance contexts will also yield gradient distinctions between the CCs of the 20 clause-embedding predicates investigated in \citealt{degen-tonhauser-openmind,degen-tonhauser-language}, rather than revealing a binary distinction between factive and non-factive predicates.\footnote{Preregistration of our experiment: \url{https://osf.io/5fsku}}

\section{\citealt{mandelkern-etal2020}}

\begin{itemize}

\item This paper is concerned with the question of whether ``the mechanism behind presupposition projection and filtering [is] fundamentally asymmetric or symmetric'' (p.473). To this end, they investigated presupposition filtering in conjunctions, and whether there is evidence for both left-to-right filtering and right-to-left filtering, or only the former.

\item Assumptions

\begin{itemize}

\item Local accommodation is ``marked'' (p.497).

\begin{exe}
\ex  I don't know if Sue has ever played piano. Did Sue play the piano again?
\\ Global accommodation (default): unacceptable
\\ Local accommodation (marked): Did Sue played the piano and did she play it again?
\end{exe}

\item p.478: ``Local accommodation in general encompasses cases in which presuppositions triggered in entailment-canceling environments that generally project presuppositions, like negation and antecedents of conditionals, do not wind up as inferences supported by the utterance as a whole. In assessing the filtering properties of complex sentences, it is thus crucial to distinguish failure to project due to local accommodation from failure to project due to filtering, either left-to-right or right-to-left. As we discuss below, our experimental designs accomplish this by including a simple presuppositional control condition where failure to project can only be due to local accommodation, thus providing a baseline relative to which any separate effects of filtering can be assessed.''

\item p.480:

\begin{exe}
\exi{(11)} Simple Ps:\\
If Mary stopped doing yoga, then Matthew will interview her for his story.
\end{exe}

Controls like (11) provide a baseline for projection. The only thing that could prevent projection from a sentence like (11) would be local accommodation. Therefore rates of projection out of sentences like this one will provide a baseline for projection in the task. To the extent that the critical variants with conjunctive antecedents exhibit the use of an additional mechanism to prevent projection, namely filtering, this should be reflected in relatively lower inference endorsement rates than in the simple control variants, where the only mechanism that could prevent projection is local accommodation.''

\end{itemize}

\item Materials Exp.~3

\begin{itemize}

\item Two contexts

\begin{exe}
\ex 
\begin{xlist}
\ex Support context: Jacob has been traveling a lot, and he's in France this week.
\begin{xlist}
\ex If Emily is \underline{happy} that Jacob is in France, then she will call him soon. \hfill (simple ps, not a critical condition)
\ex If Emily is \underline{happy} that Jacob is in France and he is in Paris, then she will call him soon. \hfill (ps in first conjunct)
\ex If Jacob is in Paris and Emily is \underline{happy} that he is in France, then she will call him soon. \hfill (ps in second conjunct)
\end{xlist}

\ex Explicit ignorance context: Jacob has been traveling a lot, but I'm not sure where he is this week.

\end{xlist}
\end{exe}

\item 4 ps triggers: {\em is happy, is aware, has stopped, continues V{\em ing}} (six lexicalizations each)
\\ 4 non-ps triggers: {\em was hoping, is sure, now frowns on, enjoys V{\em ing}} (same six lexicalizations as the ps variant)\footnote{It is not clear that {\em X now frowns on Ving} doesn't presuppose that X Ved in the past. 
\begin{exe}
\exi{(i)} Emily used to enjoy boardgames, \{and she also loved playing with toys when she was little / but I don't know whether she ever played with toys\}. If Emily \{has stopped / now frowns on\} playing with toys and she used to play with toy cars, then Jacob will buy her racing video games for her birthday.
\end{exe}
}

\item ``24 critical items in a total of 12 versions''
\\ 4 ps triggers x 6 lexicalizations = 24 critical items
\\ 12 versions = 3 (no conjunct v first vs second conjunct) x 2 (ps vs no ps) x 2 contexts

The factors ``ps vs no ps'' and ``context'' were between subject.

p.491: ``Second, precluding participants from seeing both presuppositional and non-presuppositional variants aimed to avoid any potential confusion or noise due to lack of attention to the particular verb seen in a given trial; it also aimed to prevent strategic effects and insights into the manipulation from arising. Similar concerns arose for the context manipulation, though note that through the fillers, detailed below, each participant did see some variety of contexts throughout the entire experiment.''

p.492: ``all participants saw at least 16 equivalents of Support contexts, and 16 equivalents of Expl-Ign contexts, regardless off which type of context their group was assigned to for the critical items.''

\end{itemize}

\item Procedure: ``How natural is the sentence in the given context?", ratings given in 7-point Likert scale, with points labeled `completely unnatural' and `completely natural'

\item Participants: 126 students (no mention of exclusion criteria)

\item Analysis (linear mixed effects model): Compare naturalness ratings of ps triggers and non-ps triggers in explicit ignorance context to naturalness ratings in support context.

``Thus comparing contexts which support the inference to contexts in which it has been made clear that the speaker is ignorant about the inference provides a way to distinguish a broad class of natural and invited pragmatic inferences from those that are really encoded as presuppositions, and thus have no choice but to project.'' (p.497).

\item Results

\begin{itemize}

\item p.492: ``the Simple-Ps condition exhibits low ratings for the Expl-Ign context, but high ratings for the Support context, establishing the effectiveness of the methodology. Next, the non-presuppositional conjunctive conditions, [...] also do not seem to exhibit any variation by context or, for that matter, conjunct order. Turning to the critical conditions, the conditions with the trigger in the second conjunct also seem comparably acceptable across contexts. Only when the trigger is introduced in the first conjunct does context seem to matter, with lower ratings in the Expl-Ign context.''

\item p.494: ``As a further point, it's worth highlighting that the overall pattern is remarkably uniform across triggers (see graph showing results by trigger in Appendix for details). While the acceptability of both Simple-Ps and Ps-First varies across triggers (with particularly high acceptability for {\em aware}, perhaps due to relative ease of local accommodation for this trigger), the two conditions show comparable acceptability levels in Expl-Ign for each of the triggers.''

\end{itemize}

\end{itemize}

\section{This experiment, and expectations}

\begin{itemize}

\item In this experiment, we will investigate the projection of the CCs of the 20 clause-embedding predicates using an acceptability rating task, similar to that of \citealt{mandelkern-etal2020}.

\item Like MEAL, we place our target stimuli (400 interrogatives from \citealt{degen-tonhauser-openmind, degen-tonhauser-language}) in an explicit ignorance context, where the speaker is explicitly not committed to the CC.

Instead of MEAL's `support' context (which entails the CC), we use `neutral' contexts, which are compatible with the CC but don't entail it. These `neutral' contexts are the low and high prior probability facts of \citealt{degen-tonhauser-openmind}. Projection of the CC is possible regardless of whether Julian is German (low prior probability fact) or Cuban (high prior probability fact).

\item Predictions presupposition analysis:

\begin{exe}
\ex
\begin{xlist}
\ex I have no idea if Julian dances salsa. Did Cole discover/say that Julian dances salsa?
\ex Julian is German. Did Cole discover/say that Julian dances salsa?
\ex Julian is Cuban. Did Cole discover/say that Julian dances salsa?
\end{xlist}
\end{exe}

\begin{exe}
\ex Explicit ignorance context:
\begin{xlist}
\ex Presupposition triggers: acceptable only under local accommodation, which is marked
\ex Non-ps triggers: acceptable because no ps is triggered
\end{xlist}
\ex Neutral context
\begin{xlist}
\ex Ps triggers: acceptable, since global accommodation is possible, no sensitivity to prior probability predicted
\ex Non-ps triggers: acceptable, since no ps is triggered, but inference comes about pragmatically, sensitivity to prior probability expected
\end{xlist}
\end{exe}

\item Given \citealt{degen-tonhauser-openmind}, we expect that acceptability ratings are also sensitive to these prior probabilities. If so, researchers who want to maintain the difference between factive and non-factive predicates will have to assume that accommodation is sensitive (for factives) and whatever the mechanism for projection is for non-factives is also sensitive.

\item \citealt{mandelkern-etal2020} lead us to expect a binary distinction in acceptability ratings in explicit ignorance contexts between expressions that conventionally code presuppositions and expressions whose content projects for pragmatic reasons: The former are expected to have lower ratings than the latter.

\item \citealt{mandelkern-etal2020} also lead us to expect a binary distinction in the effect of context on acceptability ratings between expressions that conventionally code presuppositions and expressions whose content projects for pragmatic reasons: For the former, a difference is expected, but not for the latter, or, at least, the difference is expected to be greater for the former than for the latter.

\item Given \citealt{degen-tonhauser-openmind,degen-tonhauser-language}, we expect acceptability ratings in an explicit ignorance context to  reveal gradient projection.

\item Given \citealt{brst-lang11}, we expect SCF+ expressions to be maximally unacceptable in such contexts: at least {\em too} and {\em also}, perhaps also clefts (see \citealt{abusch10}).

\item However, even if we find evidence for gradient projection from acceptability ratings in an explicit ignorance contexts, researchers like \citealt{mandelkern-etal2020} can always use variation in the ``ease of local accommodation'' (p.494) to account for projection variation.

\end{itemize}

\section{Methods}\label{s-methods}

\subsection{Participants}

We recruited 425 participants on Prolific. Due to a server error, the data of only 398 participants was recorded (ages: 19-73, median: 40.8; 187 women, 201 men, 8 non-binary, 2 preferred to not disclose). The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%. The median time spent on the task was 6:24 minutes. Participants were paid \$1.78 (corresponding to an hourly pay of \$16.6).

\subsection{Materials}

The materials consisted of two-sentence discourses that were uttered by a named speaker: As shown in the sample target stimuli in (\ref{sample}), the first sentence was always a declarative sentence and the second one an interrogative sentence.

\begin{exe}
\ex\label{sample}
\begin{xlist}
\ex I have no idea if Julian dances salsa. Did Cole discover that Julian dances salsa?
\ex Julian is German. Did Cole discover that Julian dances salsa?
\ex Julian is Cuban. Did Cole discover that Julian dances salsa?
\end{xlist}
\end{exe}

In the target stimuli, the interrogative sentences were identical to those of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}: they were constructed from 20 clause-embedding predicates paired with 20 clauses, as shown in (\ref{sample}) for the predicate {\em discover} and the clause {\em Julian dances salsa}. There were two context conditions: In the `explicit ignorance' context, illustrated in (\ref{sample}a), the declarative sentence that preceded the interrogative conveyed that the speaker is ignorant about the content of the embedded clause. In the `neutral' context, illustrated in (\ref{sample}b-c), the declarative sentence conveyed a fact about the subject of the embedded clause that is compatible with the projection of the CC. These facts are identical to the low and high prior probability facts of \citealt{degen-tonhauser-openmind}. These neutral contexts are not the same, however, as the `support' context used in \citealt{mandelkern-etal2020} or \citealt{kalomoiros-schwarz2021}, which are identical to the projection inference. In our `neutral' context, conventionally coded presuppositions need to be globally accommodated.

The experiment also includes 6 filler stimuli, which were only presented in an explicit ignorance context and were the same for each participant. They include ``again'' (prejacent), ``stop'' (pre-state content), ``continue'' (pre-state content), for comparison with \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, as well as ``too'' (prejacent), ``also'' (prejacent) and an it-cleft (existential closure over relative clause), because \citealt{abusch10} proposed that these are hard triggers. (Prosody is not controlled, however, for ``also'' and ``too'': the assumption is that the existential presupposition cannot be accommodated so these triggers, if indeed hard, should receive low acceptability ratings under any focus assignment.)

\begin{exe}
\ex Filler stimuli
\begin{xlist}
\ex I don't know if Ann plays any instrument. Does Ann play the flute, too?
\ex I don't know if Svenja plays any sport. Does Svenja also play soccer?
\ex I don't know if anyone was playing outside with the kids. Was it Jack who was playing outside with the kids?
\ex I don't know if Stephen was ever in the habit of vaping. Has Stephen recently stopped vaping?
\ex I don't know if John was ever reading ``Dune''. Has John recently continued reading ``Dune''?
\ex I don't know if William was ever interested in history. Is William interested in history again?"
\end{xlist}
\end{exe}

Finally, the experiment included 4 control stimuli, which were used to exclude participants not attending to the task. These are expected to be rated as natural/acceptable.

Each participant saw a random set of 30 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates), the same 6 control stimuli, and the same 4  filler stimuli. 12 of the CCs were presented in the explicit ignorance context, 4 with a low prior probability fact, 4 with a high prior probability fact. Within-block trial order was randomized. After completing the experiment, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

The experiment started off with four practice trials (2 natural, 2 unnatural; participants are given feedback on correct and incorrect answers).

\subsection{Data exclusion}

We excluded the data of five participants who did not self-declare to be native speakers of American English. We also excluded the data of 23 participants whose mean response to the controls was more than 2 sd below the group mean.\footnote{Only three of the four controls were used to exclude participants. While three of the four controls received high acceptability ratings, as expected (mean: .86), the fourth control, shown in (i), only received a mean rating of 0.5. 
\begin{exe}
\exi{(i)} Hendrick was looking to buy a car. Was Hendrick's car expensive?
\end{exe}
} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-binary, 2 preferred to not disclose).

\newpage

\section{Results}

\subsection{Is projection binary or gradient?}

Is projection binary, as suggested in \citealt{mandelkern-etal2020}, or gradient, as suggested in \citealt{degen-tonhauser-openmind,degen-tonhauser-language}?

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
\caption{Mean naturalness ratings in explicit ignorance contexts, by expression. Error bars indicate 95\% bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants' ratings. Fillers are given in gray, factive predicates in red.}\label{fig:}
\end{figure}

\begin{itemize}

\item \citealt{brst-lang11} classified the prejacent of {\em too} and {\em also} as +SCF content, that is, content that cannot be (globally or locally) accommodated. As expected, these expression/content pairs receive among the lowest acceptability ratings in explicit ignorance contexts.

\item \citealt{brst-lang11} classified the pre-state contents of {\em stop} and {\em continue} as -SCF content, that is, content that can be accommodated. \citealt{tbd-variability} observed that the pre-state content of {\em stop} is projective, but not at ceiling. Against this background, the very low mean acceptability ratings, especially for {\em continue}, are unexpected. They suggest that the pre-state contents resist local accommodation (if ps is triggered) or are strongly projective.

\item \citealt{abusch10} assumed that {\em it-}clefts are hard triggers, that is, unacceptable in explicit ignorance contexts. Our results are more in line with the results of \citealt{smith-hall-cls}, who also did not observe strong projection for the cleft.

\item Contrary to what \citealt{mandelkern-etal2020} suggested, there is no binary distinction between factive and non-factive predicates: The mean acceptability ratings for several factive predicates ({\em see, discover, reveal}) are at least as high as that of non-factive predicates.  The mean acceptability rating of {\em know} is not at floor (notice the bimodal responses!). The mean rating of {\em be annoyed} is higher than that of {\em too} and {\em also}.

\end{itemize}

\newpage

\subsection{Difference between naturalness ratings in explicit ignorance and neutral contexts?}

Similar to Mandelkern et al?s 2020 analysis: They claim that there is a difference in the naturalness ratings in support and explicit ignorance contexts between presuppositions and non-presuppositions.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
\caption{Mean naturalness ratings, by context and expression. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:}
\end{figure}

%Here, I will plot, for each of the 20 predicates (target stimuli), the mean naturalness rating in the two contexts (explicit ignorance, neutral). 
%
%Expectations:
%    - Mandelkern et al 2020: presupposition triggers (factives) have lower naturalness ratings in explicit ignorance context than neutral context than non-factives, for which no difference is expected.
%    - Degen & Tonhauser 2022: the differences between the two contexts are gradient across the 20 predicates, without a clear distinction between factive and non-factive predicates.
%
%Models (fit only to target data):
%
%A. naturalness rating ~ factivity (RL: factive) * context (RL: explIgnorance), by-item (combination of predicate and CC) and by-participant intercepts, factivity slope for participant, and context slopes for item and participant
%
%Expectations based on Mandelkern et al 2020:
%- significant main effect of factivity, positive coefficient: nonfactive predicates better in explIgnorance context than factive ones
%- significant main effect of context, positive coefficient: factive predicates better in neutral context than explIgnorance 
%- interaction: positive coefficient: higher naturalness ratings for nonfactive predicates in neutral context than factive ones in explIgnorance context
%
%?%B. naturalness ~  expression (RL: predicate with smallest naturalness difference between the two contexts) * context (RL: explIgnorance), by-item and by-participant intercepts, and context slopes for item and participant. 
%
%Idea in choosing RL: when there is no difference, this is not a presupposition trigger.
%
%Expectations based on Degen & Tonhauser 2022:
%        - main effect of expression: 
%            - positive coefficient: naturalness ratings for the expression in explIgnorance context is higher than for RL expression (suggests projection of CC is easier to suppress)
%            - negative coefficient: naturalness ratings for the expression in neutral context is lower than for RL expression (suggests projection of CC is harder to suppress)
%        - main effect of context (will ideally be not significant at RL, to interpret interaction)
%        - interaction:
%            - positive coefficient: naturalness rating for the expression in neutral context is higher than for RL expression in explIgnorance context (uninteresting)
%            - negative coefficient: naturalness rating for the expression in neutral context is lower than for RL expression in explIgnorance context (uninteresting)
%
%Comparison of models A and B with AIC to identify whether the factivity-based model A or the predicate-based model B capture more variance in the data (see de Marneffe, Simons & Tonhauser 2019 SuB, Degen & Tonhauser 2022).
%
%Third analysis: Do the naturalness ratings in explicit ignorance contexts and the ?certain that? ratings get at the same underlying theoretical construct (namely speaker commitment)?
%
%    - Plot the mean naturalness ratings of the CCs of the 20 clause-embedding predicates against their mean certainty ratings.
%    - Calculate Spearman rank correlation.
%    - If the two measures get at the same underlying theoretical construct, then we expect to find a negative correlation between the two measures such that the higher the mean naturalness rating is (that is, the more the CC does not project), the lower the mean ?certain that? rating.
%
\section{Discussion}\label{s4}

\bibliographystyle{cslipubs-natbib}
\bibliography{/Users/tonhauser.1/Documents/bibliography}

\newpage

\section*{Supplemental materials}

\appendix

\setcounter{page}{1}
%\renewcommand{\thetable}{A\arabic{table}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\section{20 clauses}\label{a-clauses}

The contents of the following 20 clauses, which realized the complements of the 20 clause-embedding predicates, were investigated in Exps.~1-3:

\begin{enumerate}[leftmargin=3ex,itemsep=-2pt]

\begin{multicols}{2}

\item Mary is pregnant.
\item Josie went on vacation to France.
\item Emma studied on Saturday morning.
\item Olivia sleeps until noon.
\item Sophia got a tattoo.
\item Mia drank 2 cocktails last night.
\item Isabella ate a steak on Sunday.
\item  Emily bought a car yesterday.
\item  Grace visited her sister.
\item Zoe calculated the tip.

\columnbreak

\item  Danny ate the last cupcake.
\item  Frank got a cat.
\item  Jackson ran 10 miles.
\item  Jayden rented a car.
\item  Tony had a drink last night.
\item  Josh learned to ride a bike yesterday.
\item  Owen shoveled snow last winter.
\item  Julian dances salsa.
\item  Jon walks to work.
\item  Charley speaks Spanish.

\end{multicols}

\end{enumerate}

\section{Participant information and data exclusion criteria}\label{a-participants}

This supplement provides information on the participants of the 12 experiments and the criteria by which participants' data were excluded. The first three columns of Table \ref{t:exclusion} show, for each of the 12 experiments how many participants were recruited, their age range and mean age, and their self-reported gender; no gender data was collected in Exp.~1q. The next three columns provide information on the number of participants whose data were excluded based on the following criteria:

\begin{itemize}[itemsep=-2pt]

\item `multiple': Due to an experimental glitch, some participants participated more than once in Exp.~1q. Since no information was available on which one was their first take, those participants' data was removed.

\item `language': Participants' data were excluded if they did not self-identify as native speakers of American English.

\item `controls': Participants' data were excluded if their mean rating on the 6 main clause control items in the projection block was more than 2 sd above the group mean (in Exps.~1q, 2q, and 3q) or more than 2 sd below the group mean (in the remaining experiments). Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments).

\item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli. To identify such participants, we first identified participants whose mean variance on the target stimuli was more than 2 sd below the group mean variance and then manually inspecting their response patterns. The data of participants who used the full scale was not excluded.  
\end{itemize}

The remaining columns of Table \ref{t:exclusion} provide information on the remaining participants, that is, the participants' data that entered into the analysis. Participants took around 9-11 minutes to complete the various experiments. Participants were paid more in Exps.~1c, 2c, and 3c than the remaining experiments because the target stimuli in those experiments were longer (as they consisted of conditionals). More women than men were recruited in many of the experiments because the experiments were run at a time when Prolific went viral on TikTok, resulting in a large number of young women registering for the service (around July 24, 2021; see \url{https://blog.prolific.co/we-recently-went-viral-on-tiktok-heres-what-we-learned/}, \\ last accessed February 4, 2022).

\end{document}

### 2023-07-15 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -1,0 +1,76 @@
+# 13_explicitIgnorance
+# analysis
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load required packages
+require(tidyverse)
+library(lme4)
+
+# load helper functions
+source('../../../helpers.R')
+
+# load cleaned data
+d = read_tsv("../data/cd.tsv")
+
+names(d)
+table(d$context)
+
+length(unique(d$participantID)) #370 participants
+
+# predict acceptability rating in comparison to least acceptable (continue)
+
+# target data: explicit ignorance context
+t = d %>%
+  filter(context == "explicitIgnorance")
+
+# merge the two controls into one
+table(t$expression)
+t = t %>%
+  mutate(expression = recode(expression, "controlGood1" = "controls", "controlGood2" = "controls"))
+
+# set reference level
+# create item as combination of predicate and complement clause
+# cd$item = as.factor(paste(cd$verb,cd$content))
+
+# 
+str(t$expression)
+t$expression <- as.factor(t$expression)
+t$expression <- relevel(t$expression, ref = "continue")
+names(t)
+m = lmer(response ~ expression + (1|participantID), data = t, REML=F)
+summary(m)
+
+# main analysis of interest
+
+# the model reported in the paper
+m = lmer(response ~ cai * cblock_ai + (1+cai|workerid) + (0+cai|content) + (1+cai|short_trigger), data=t_nomc, REML=F)
+summary(m.mr.1)
+
+# get p-values via likelihood ratio tests
+m.mr.0a = lmer(projective ~ cai + cblock_ai + (1+cai|workerid) + (0+cai|content) + (1+cai|short_trigger), data=t_nomc, REML=F)
+summary(m.mr.0a)
+
+m.mr.0b = lmer(projective ~ cai + cai : cblock_ai + (1+cai|workerid) + (0+cai|content) + (1+cai|short_trigger), data=t_nomc, REML=F)
+summary(m.mr.0b)
+
+m.mr.0c = lmer(projective ~ cblock_ai + cai : cblock_ai + (1+cai|workerid) + (0+cai|content) + (1+cai|short_trigger), data=t_nomc, REML=F)
+summary(m.mr.0c)
+
+anova(m.mr.0a,m.mr.1) #p-value for interaction
+anova(m.mr.0b,m.mr.1) #p-value for block
[…]

### 2023-07-15 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -1,0 +1,272 @@
+# 13_explicitIgnorance
+# graphs
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load required packages
+require(tidyverse)
+library(ggrepel)
+library(dichromat)
+library(forcats)
+library(ggrepel)
+
+# color-blind-friendly palette
+cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+theme_set(theme_bw())
+
+# load helper functions
+source('../../helpers.R')
+
+# load cleaned data
+d = read_tsv("../data/cd.tsv")
+
+names(d)
+table(d$context)
+
+length(unique(d$participantID)) #370 participants
+
+# plot of mean naturalness ratings in explicit ignorance context ----
+
+# target data: explicit ignorance context
+t = d %>%
+  filter(context == "explicitIgnorance")
+
+# merge the two controls into one
+table(t$expression)
+t = t %>%
+  mutate(expression = recode(expression, "controlGood1" = "controls", "controlGood2" = "controls"))
+
+# calculate mean naturalness rating by expression, including the controls
+nat.means = t %>%
+  group_by(expression) %>%
+  summarize(Mean = mean(response), CILow = ci.low(response), CIHigh = ci.high(response)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, expression = fct_reorder(as.factor(expression),Mean))
+nat.means
+levels(nat.means$expression)
+
+# color code factives and ps triggers
+factives <- c("know", "discover", "be annoyed", "reveal", "see")
+triggers <- c("too", "also", "cleft", "again", "stop", "continue")
+
+nat.means$ps = ifelse(nat.means$expression %in% triggers, "trigger", 
+                      ifelse(nat.means$expression %in% factives, "factive", "other"))
+
+table(nat.means$ps, nat.means$expression)
+
+text.color <- ifelse(nat.means$expression[order(nat.means$Mean)] %in% factives, '#D55E00', 
+                     ifelse(nat.means$expression[order(nat.means$Mean)] %in% triggers, "gray", "black"))
+text.color
+
+t$expression = factor(t$expression, levels = nat.means$expression[order(nat.means$Mean)], ordered = TRUE)
+
[…]

### 2023-07-15 results/main/13explicitIgnorance/rscripts/preprocessing.R

@@ -1,0 +1,477 @@
+# 13_explicitIgnorance
+# preprocessing
+
+# server issue: doesn't collect all the data that were paid for
+# first round: paid 400, collected only 375
+# second round: paid 25, collected only 23
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+source('../../helpers.R')
+
+# load required packages for pre-processing data
+library(tidyverse)
+library(readr)
+
+theme_set(theme_bw())
+
+# read in the raw data
+d1 = read_tsv("../data/combined1.tsv")
+d2 = read_tsv("../data/combined2.tsv")
+# parsing warnings are due to the rows with demographics info and final screens
+
+d = rbind(d1,d2)
+
+# remove rows with info about instructions and final screens
+d = d %>% drop_na(task)
+
+# replace participant_id by random number
+length(unique(d$participant_id)) #398
+d$participantID <- match(d$participant_id, unique(sort(d$participant_id)))
+
+# how many participants?
+length(unique(d$participantID)) #398
+
+# from Prolific
+# £12.66/hr
+# median time: 00:06:24
+
+# select relevant columns
+d = d %>%
+  select(c(stimulus, response, trial_index, task, participantID))
+
+# unpack demographics info
+dg <- d %>%
+  filter(task == "demographics") %>%
+  select(c(participantID,response))
+#view(dg)
+
+table(dg$response)
+
+# age
+dg$age = as.numeric(gsub("\\D", "", dg$response))
+table(dg$age)
+
+# gender
+dg$gender = case_when(grepl("female", dg$response) ~ "female",
+                      grepl("male", dg$response) ~ "male",
+                      grepl("non-binary", dg$response) ~ "non-binary",
+                      TRUE ~ "preferNoToSay")
+table(dg$gender)
+
+# female          male    non-binary preferNoToSay 
+#187           201             8             2
+
+# language
+dg$language = case_when(grepl("language\"\":\"\"yes", dg$response) ~ "English",
+                      TRUE ~ "notSpeakerOfEnglish")
+table(dg$language)
+
+# American English
+dg$amE = case_when(grepl("amE\"\":\"\"yes", dg$response) ~ "AmE",
+                        TRUE ~ "notAmE")
+table(dg$amE)
+
+# comments
+dg$comments = gsub(".*comments", "", dg$response)
+table(dg$comments)
+                      
+
[…]

### 2023-07-16 writing/DGfS2023-abstract/write-up.tex
updated exp13 analysis and write-up

@@
 \begin{itemize}
-\item[a)] projection is gradient, that is, it is not the case that content either projects or it doesn't, but rather content projects to varying degrees, reflecting either varying degrees of speaker commitment or varying degrees of certainty of the interpreter on the speaker's commitment to a content, and
+\item[a)] projection is gradient, that is, it is not the case that content either projects or it doesn't, but rather content projects to varying degrees, reflecting varying degrees of (of certainty of the interpreter about) speaker commitment, and
 
@@
 \\ 4 ps triggers x 6 lexicalizations = 24 critical items
-\\ 12 versions = 3 (no conjunct v first vs second conjunct) x 2 (ps vs no ps) x 2 contexts
+\\ 12 versions = 3 (no conjunct vs first vs second conjunct) x 2 (ps vs no ps) x 2 contexts
 
@@
 
-``Thus comparing contexts which support the inference to contexts in which it has been made clear that the speaker is ignorant about the inference provides a way to distinguish a broad class of natural and invited pragmatic inferences from those that are really encoded as presuppositions, and thus have no choice but to project.'' (p.497).
+``Thus comparing contexts which support the inference to contexts in which it has been made clear that the speaker is ignorant about the inference provides a way to distinguish a broad class of natural and invited pragmatic inferences from those that are really encoded as presuppositions, and thus {\bf have no choice but to project}.'' (p.497, bold-face added).
 
@@
 
-\item In this experiment, we will investigate the projection of the CCs of the 20 clause-embedding predicates using an acceptability rating task, similar to that of \citealt{mandelkern-etal2020}.
+\item In this experiment, we will investigate the projection of the CCs of the 20 clause-embedding predicates using an acceptability rating task that is similar to that of \citealt{mandelkern-etal2020}.
 
 \item Like MEAL, we place our target stimuli (400 interrogatives from \citealt{degen-tonhauser-openmind, degen-tonhauser-language}) in an explicit ignorance context, where the speaker is explicitly not committed to the CC.
-
-Instead of MEAL's `support' context (which entails the CC), we use `neutral' contexts, which are compatible with the CC but don't entail it.
-These `neutral' contexts are the low and high prior probability facts of \citealt{degen-tonhauser-openmind}.
-Projection of the CC is possible regardless of whether Julian is German (low prior probability fact) or Cuban (high prior probability fact).
-
-\item Predictions presupposition analysis:
 
@@
 
+\item Instead of MEAL's `support' context (which entails the CC), we use `neutral' contexts, which are compatible with the CC but don't entail it.
+(\citealt{brst-lang11} called such contexts neutral, contrasting them with positive, which is like their support context.) These `neutral' contexts are given by the low and high prior probability facts of \citealt{degen-tonhauser-openmind}: Regardless of whether Julian is German (low prior probability fact) or Cuban (high prior probability fact), he may or may not dance salsa.
+
+Using these two types of `neutral' context allows us to identify whether acceptability ratings are also sensitive to the content's prior probability.
+It also gives us a comparison between contexts where the speaker is explicitly ignorant about the CC vs not explicitly ignorant but possibly ignorant.
+
+\item Predictions presupposition analysis:
+
 \begin{exe}
@@
 \begin{xlist}
-\ex Ps triggers: acceptable, since global accommodation is possible, no sensitivity to prior probability predicted
+\ex Ps triggers: acceptable, since global accommodation is possible, and ``default'' and ``unmarked'', no sensitivity to prior probability predicted
 \ex Non-ps triggers: acceptable, since no ps is triggered, but inference comes about pragmatically, sensitivity to prior probability expected
@@
 
-\item Given \citealt{brst-lang11}, we expect SCF+ expressions to be maximally unacceptable in such contexts: at least {\em too} and {\em also}, perhaps also clefts (see \citealt{abusch10}).
+\item Given \citealt{brst-lang11}, we expect SCF+ expressions, namely {\em too} and {\em also}, to be maximally unacceptable in such contexts.
+
+\item \citealt{abusch10} leads us to expect that it-clefts receive low acceptability ratings since it is a hard trigger, like {\em too}.
 
@@
 These facts are identical to the low and high prior probability facts of \citealt{degen-tonhauser-openmind}.
-These neutral contexts are not the same, however, as the `support' context used in \citealt{mandelkern-etal2020} or \citealt{kalomoiros-schwarz2021}, which are identical to the projection inference.
-In our `neutral' context, conventionally coded presuppositions need to be globally accommodated.
 
-The experiment also includes 6 filler stimuli, which were only presented in an explicit ignorance context and were the same for each participant.
+The experiment also included 6 filler stimuli, which were only presented in an explicit ignorance context and were the same for each participant.
 They include ``again'' (prejacent), ``stop'' (pre-state content), ``continue'' (pre-state content), for comparison with \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, as well as ``too'' (prejacent), ``also'' (prejacent) and an it-cleft (existential closure over relative clause), because \citealt{abusch10} proposed that these are hard triggers.
-(Prosody is not controlled, however, for ``also'' and ``too'': the assumption is that the existential presupposition cannot be accommodated so these triggers, if indeed hard, should receive low acceptability ratings under any focus assignment.)
+(Prosody was not controlled: the assumption is that the  presuppositions of {\em too, also} and the it-cleft cannot be accommodated so these triggers, if indeed hard, should receive low acceptability ratings under any focus assignment.)
 
@@
 
-Finally, the experiment included 4 control stimuli, which were used to exclude participants not attending to the task. These are expected to be rated as natural/acceptable.
+Finally, the experiment included 4 control stimuli, which were used to exclude participants not attending to the task. These were expected to be rated as natural/acceptable.
 
-Each participant saw a random set of 30 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates), the same 6 control stimuli, and the same 4  filler stimuli.
+Each participant saw a random set of 30 stimuli: Each set contained one target utterance for each of the 20 CCs (each paired with a unique clause-embedding predicates), the same 6 control stimuli, and the same 4  filler stimuli.
 12 of the CCs were presented in the explicit ignorance context, 4 with a low prior probability fact, 4 with a high prior probability fact.
@@
 
-The experiment started off with four practice trials (2 natural, 2 unnatural; participants are given feedback on correct and incorrect answers).
+The experiment started off with four practice trials (2 natural, 2 unnatural; participants were given feedback on correct and incorrect answers).
 
@@
 
-\newpage
+\section{Results}
 
-\section{Results}
+Data points:
+
+\begin{itemize}
+
+\item Each predicate and each CC received 370 ratings.
+
+\item Each of the 60 predicate/context combinations received at least 59 ratings (median: 80, mean: 123).
+
+\item Each predicate received at least 200 ratings in the explicit ignorance context (median: 224.5, mean: 222).
+
+\item Each predicate received at least 59 ratings in each of the neutral contexts (median: 75, mean: 74).
+
+\end{itemize}
 
@@
 \includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
-\caption{Mean naturalness ratings in explicit ignorance contexts, by expression.
+\caption{Mean naturalness ratings in explicit ignorance contexts by expression.
 Error bars indicate 95\% bootstrapped confidence intervals.
 Violin plots indicate the kernel probability density of the individual participants' ratings.
-Fillers are given in gray, factive predicates in red.}\label{fig:}
+Fillers are given in gray, factive predicates in red.}\label{fig:acc-by-expression}
 \end{figure}
@@
 
-\newpage
-
 \subsection{Difference between naturalness ratings in explicit ignorance and neutral contexts?}
 
-Similar to Mandelkern et al?s 2020 analysis: They claim that there is a difference in the naturalness ratings in support and explicit ignorance contexts between presuppositions and non-presuppositions.
+\citealt{mandelkern-etal2020} did not merely compare acceptability ratings of presuppositions and non-presuppositions in explicit ignorance contexts, but they rather claimed that there is a difference in the acceptability ratings in the two contexts between presuppositions and non-presuppositions.
+Specifically, while there is no difference for non-presuppositions, the ratings for presuppositions are lower in the explicit ignorance context than in the support context.
+
+For comparison with their way of presenting the results, Fig.~\ref{fig:acc-by-context2} plots the mean acceptability ratings by expression and context (distinguishing explicit ignorance from neutral), with the expressions sorted by the difference between the mean rating in the two contexts: The mean rating in the neutral context is higher than the mean rating in the explicit ignorance context for {\em be annoyed}; the opposite is the case for {\em confirm}.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
-\caption{Mean naturalness ratings, by context and expression. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:}
+\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context2-and-predicate}
+\caption{Mean naturalness ratings by expression and context. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:acc-by-context2}
 \end{figure}
+
+\begin{itemize}
+
+\item For {\em be annoyed} and {\em know}, the mean rating in the explicit ignorance context is lower than the mean rating in the neutral context, suggesting that the interrogatives are judged to be less neutral when the speaker is explicitly ignorant about the truth of the CC.
+
+For {\em see}, the mean rating in the two contexts does not appear to be different (but model suggests neutral is already higher), and the pattern seems to go in the opposite direction for {\em discover} and {\em reveal}.
+
+Thus, on this way of displaying the data, too, factive predicates do not show a homogenous pattern that distinguishes them from non-factive predicates.
+
+\item Non-factive predicates also do not show a homogenous pattern: For some predicates, the ratings in the two contexts do not appear to be different (e.g., {\em acknowledge, inform}), while ratings in the explicit ignorance context seem to be higher than in the neutral context for other predicates (e.g., {\em say, confirm}).
+
+\item When ratings are higher in explicit ignorance than neutral contexts, this suggests that a natural use of these predicates is for the speaker to inquire about the CC, given their explicit ignorance.
+
+\end{itemize}
+
+\subsection{The effect of the prior probability of the context on naturalness ratings}
+
+Fig.~\ref{fig:acc-by-context} plots the mean acceptability ratings by expression and context (distinguishing explicit ignorance from factL and factH), with the expressions sorted on the mean acceptability rating in the explicit ignorance context, as in Fig.~\ref{fig:acc-by-expression}.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
+\caption{Mean naturalness ratings by expression and context. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:acc-by-context}
+\end{figure}
+
+\begin{itemize}
+
+\item For all predicates except {\em pretend} (and {\em admit}?), the acceptability ratings are higher in factH contexts than in factL contexts.
+This observation suggests that acceptability ratings are sensitive to the prior probability of the CC.
+This is true regardless of whether ratings in explicit ignorance contexts are higher or lower or similar to those in neutral contexts.
+
+This observation suggests that it is generally more acceptable for a speaker to ask about some AH's relation to the CC in a context in which the speaker assigns a higher probability to the CC than in a context where the speaker assigns a lower probability to the CC.
+
+\item {\em be annoyed} and {\em know} show distinct patterns: For {\em be annoyed}, ratings are higher in the factH context than in the factL context, and higher in the factL context than in the explicit ignorance context.
+For {\em know}, the latter two ratings do not differ.
+So even these two predicates don't show the same pattern in this task.
+
+\begin{itemize}
+
+\item Assuming that both predicates are factive and trigger a ps doesn't capture why they show these different patterns.
+
+\item {\em be annoyed:} According to \citealt{karttunen2016}, this predicate codes that the AH believes the CC.
+Belief of the speaker in the CC arises as long as there is no contextual evidence/information that would justify why the speaker wouldn't share the belief.
+In the explicit ignorance context, the speaker has gone on record that they are ignorant about the CC, that is, do not share in the AH's belief.
+Given the assumption that {\em be annoyed} only codes AH belief in the CC, the very low acceptability ratings in explicit ignorance contexts are not accounted for.
+
+\item An explanation may have to do with at-issueness: The explicit ignorance context makes the question ?CC at-issue, but  the interrogative makes the CC not-at-issue, resulting in lower acceptability ratings.
+
+\item The same explanation may also work for the fillers with {\em stop} and {\em continue}: The explicit ignorance context makes the pre-state content at-issue, but the interrogative makes it not-at-issue.
+
+\item The CC of {\em know} is still quite not-at-issue, but not as much as that of {\em be annoyed}. {\bf need to check this in our other data}
+
+\end{itemize}
+
+\item For predicates other than {\em be annoyed} and {\em know}, acceptability ratings in explicit ignorance contexts are roughly comparable to those in factH contexts.
+That is, asking whether an AH has a relation to the CC is just as acceptable in a context in which the speaker is explicitly ignorant of the CC and in a context in which the speaker has a belief relative to which the CC has a high probability.
+
+It is when the speaker has a belief relative to which the CC has a low probability that the acceptability ratings are lowest (except with {\em pretend}).
+
+\end{itemize}
 

### 2023-07-16 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -20,7 +20,7 @@ table(d$context)
 
 length(unique(d$participantID)) #370 participants
 
-# predict acceptability rating in comparison to least acceptable (continue)
+# predict acceptability rating in comparison to least acceptable (continue) ----
 
 # target data: explicit ignorance context
 t = d %>%
@@ -35,13 +35,42 @@ t = t %>%
 # create item as combination of predicate and complement clause
 # cd$item = as.factor(paste(cd$verb,cd$content))
 
-# 
+# model
 str(t$expression)
 t$expression <- as.factor(t$expression)
 t$expression <- relevel(t$expression, ref = "continue")
 names(t)
 m = lmer(response ~ expression + (1|participantID), data = t, REML=F)
 summary(m)
+
+# predict acceptability rating from predicate and context ----
+
+# target data: predicates
+t = d %>%
+  filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4") %>%
+  filter(expression != "also" & expression != "too" & expression != "again" & expression != "cleft" &
+           expression != "stop" & expression != "continue") %>%
+  mutate(expression = recode(expression, "controlGood1" = "controls", "controlGood2" = "controls")) %>%
+  mutate(context2 = recode(context, "factL" = "neutral", "factH" = "neutral"))
+
+# set reference levels
+# create item as combination of predicate and complement clause
+# cd$item = as.factor(paste(cd$verb,cd$content))
+
+# reference level for expression is "see" as there is smallest difference 
+str(t$expression)
+t$expression <- as.factor(t$expression)
+t$expression <- relevel(t$expression, ref = "see")
+
+# reference level for context is "explicitIgnorance"
+t$context2 <- as.factor(t$context2)
+t$context2 <- relevel(t$context2, ref = "explicitIgnorance")
+
+# model
+m = lmer(response ~ expression * context2 + (1|participantID), data = t, REML=F)
+summary(m)
+
+#######
 
 # main analysis of interest
 

### 2023-07-16 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -201,6 +201,21 @@ ggplot(nat.means, aes(x=expression, y=Mean, group = context, fill = context)) +
   theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) 
 ggsave("../graphs/naturalness-by-context-and-predicate.pdf",height=4,width=7)
 
+ggplot(nat.means, aes(x=context, y=Mean, fill = context)) +
+  geom_bar(stat="identity", color = "black", position=position_dodge(.9)) +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1,color="black", position=position_dodge(.9)) +
+  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
+  scale_fill_manual(values=c('gray40',"#56B4E9",'#E69F00'), name = "Context") + 
+  guides(fill=FALSE) +
+  theme(text = element_text(size=12), axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) +
+  theme(legend.position="top") +
+  facet_wrap(. ~ expression) +
+  ylab("Mean naturalness rating") +
+  xlab("Context") +
+  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) 
+ggsave("../graphs/naturalness-by-context-and-predicate.pdf",height=7,width=7)
+
+
 
 # plot of mean naturalness ratings against mean certainty ratings
 

### 2023-07-16 results/main/13explicitIgnorance/rscripts/preprocessing.R

@@ -474,4 +474,58 @@ d %>%
 
 write_tsv(d, file="../data/cd.tsv")
 
+# how many data points per predicate/CC/context combination?
+names(d)
+
+tmp = d %>%
+  filter(cc != "noCC") %>%
+  group_by(expression,cc,context) %>% 
+  tally
+tmp
+nrow(tmp) # 1185 (if everything had been chosen, should be 1200 = 3 contexts x 400 predicate/cc combinations)
+min(tmp$n) #1
+mean(tmp$n) #6.2
+max(tmp$n) #23
+#view(tmp)
+
+# how many data points per predicate/context combination?
+tmp = d %>%
+  filter(cc != "noCC") %>%
+  group_by(expression,context) %>% 
+  tally
+tmp
+nrow(tmp) # 60 combinations (3 contexts x 20 predicates)
+sd(tmp$n)
+min(tmp$n) #59
+mean(tmp$n) #123
+median(tmp$n) #80
+max(tmp$n) #234
+
+# how many data points for each predicate in the explicit ignorance context?
+tmp = d %>%
+  filter(context == "explicitIgnorance") %>%
+  filter(cc != "noCC") %>%
+  group_by(expression,context) %>% 
+  tally
+tmp
+nrow(tmp) # 20 combinations (1 contexts x 20 predicates)
+sd(tmp$n) #8.7
+min(tmp$n) #200
+mean(tmp$n) #222
+median(tmp$n) #224.5
+max(tmp$n) #234
+
+# how many data points for each predicate in the two neutral contexts?
+tmp = d %>%
+  filter(context != "explicitIgnorance") %>%
+  filter(cc != "noCC") %>%
+  group_by(expression,context) %>% 
+  tally
+tmp
+nrow(tmp) # 40 combinations (2 contexts x 20 predicates)
+sd(tmp$n) #8
+min(tmp$n) #59
+mean(tmp$n) #74
+median(tmp$n) #75
+max(tmp$n) #87
 

### 2023-08-01 writing/DGfS2023-abstract/abstract.tex
updated DGfS abstract

\documentclass[12pt,fleqn]{article}
\usepackage[margin=1in,top=1in,bottom=1in]{geometry}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{hyperref}
%\usepackage[dvips]{graphics}
%\usepackage{amssymb}
\usepackage{float}
%\usepackage[usenames]{xcolor}
%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{subcaption}

\usepackage[normalem]{ulem}

\usepackage{multicol}
\usepackage{txfonts}
\usepackage{amsfonts}
\usepackage{natbib}
\usepackage{gb4e}
\usepackage[all]{xy}
\usepackage{rotating}
\usepackage{tipa}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{url}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{array}

\def\bad{{\leavevmode\llap{*}}}
\def\marginal{{\leavevmode\llap{''}}}
\def\verymarginal{{\leavevmode\llap{''''}}}
\def\swmarginal{{\leavevmode\llap{4}}}
\def\infelic{{\leavevmode\llap{\#}}}

\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
%\definecolor{gray}{rgb}{0.36, 0.54, 0.66}

\definecolor{Pink}{RGB}{240,0,120}
\newcommand{\red}[1]{\textcolor{Pink}{#1}}
\newcommand{\jd}[1]{\textbf{\textcolor{Pink}{[jd: #1]}}}

\newcommand{\dashrule}[1][black]{%
  \color{#1}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}\xleaders\hbox{\rule{4pt}{0pt}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}}\hfill\kern0pt%
}

\setlength{\parindent}{2ex}
\setlength{\parskip}{.3ex}

\newcommand{\yi}{\'{\symbol{16}}}
\newcommand{\nasi}{\~{\symbol{16}}}
\newcommand{\hina}{h\nasi na}
\newcommand{\ina}{\nasi na}

\newcommand{\foc}{$_{\mbox{\small F}}$}

\hyphenation{par-ti-ci-pa-tion}

\setlength{\bibhang}{0.5in}
\setlength{\bibsep}{0mm}
\bibpunct[:]{(}{)}{,}{a}{}{,}

\newcommand{\6}{\mbox{$[\hspace*{-.6mm}[$}} 
\newcommand{\9}{\mbox{$]\hspace*{-.6mm}]$}}
\newcommand{\sem}[2]{\6#1\9$^{#2}$}
\renewcommand{\ni}{\~{\i}}

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{#1}}
\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{#1})}

\newcolumntype{R}[2]{%
    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
    l%
    <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{90}{0em}}}% no optional argument here, please!

\newcommand{\jt}[1]{\textbf{\color{blue}JT: #1}}

\begin{document}

%Abstracts should fit two pages (letter size or A4 paper, 2.54cm or 1 inch margins on all sides, 12 point font, Times New Roman), with an additional third page used exclusively for the following elements: references (obligatory), large figures or tables,

\begin{center}
{\bf I have no idea if factivity is categorical. Did Mandelkern et al.\ 2020 discover that it is?}
\end{center}

\vspace*{-.5cm}

\noindent
This talk investigates the long-standing assumption that attitude predicates divide into factive and nonfactive ones (\citealt{kiparsky-kiparsky70,karttunen71b,heim92}, i.a.). This assumption was recently challenged by empirical investigations (e.g., \citealt*{demarneffe-etal-sub23,degen-tonhauser-language}), but \citealt*{mandelkern-etal2020} argued that the inference rating tasks used in these investigations are unsuitable to detect a categorical factivity distinction. This talk presents the results of an experiment that investigates factive and nonfactive predicates using the naturalness rating task advocated for in \citealt{mandelkern-etal2020}. The results again fail to support  a categorical factivity distinction, in line with \citealt{degen-tonhauser-language}.

\noindent
{\bf Introduction.} Attitude predicates have long been divided into factive and nonfactive ones based on whether the content of their clausal complement (CC) is presupposed (see references above): {\em know} is factive because interpreters infer that a speaker who utters (\ref{first}) with {\em know} believes that Julian dancing salsa is part of the common ground of the interlocutors (the CC ``projects''), whereas this inference does not arise with the variant of (\ref{first}) with nonfactive {\em think} (the CC ``does not project'').

\vspace*{-.2cm}
\begin{exe}
\ex\label{first} Ed: {\em ``Does Cole know/think that Julian dances salsa?''}
\end{exe}
\vspace*{-.2cm}

This classification of predicates was recently challenged based on empirical investigations (e.g., \citealt{demarneffe-etal-sub23,degen-tonhauser-language}). These investigations, which asked participants to draw inferences about the CC from examples like (\ref{first}), confirmed the long-standing intuition that the CCs of factive predicates are, generally, highly projective, but they also found i) that there is variability in how projective the CCs of factive predicates are, ii) that the CCs of factive predicates are not categorically more projective than those of nonfactive predicates, and iii) that the CCs of some factive predicates are less projective than that of nonfactive ones. These results were taken to suggest that ``there is little empirical support [...] for the assumed categorical distinction between factive and nonfactive predicates'' (\citealt[552]{degen-tonhauser-language}).

\citealt{mandelkern-etal2020} challenged this conclusion, arguing that gradient projection is an artifact of the inference task used, writing that ``inference tasks to some degree invite subjects to make the inference in question'' (p.498) and that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497). They suggested that naturalness ratings of utterances with attitude predicates in explicit ignorance contexts (EICs), as in (\ref{second}), are more suitable to distinguishing the semantic presuppositions of factive predicates (hypothesized to be unnatural in EICs) from inferences that arise with nonfactive predicates ``for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497).

\vspace*{-.2cm}
\begin{exe}
\ex\label{second} Ed: \hspace*{-.2cm} {\em ``I have no idea if Julian dances salsa.}  \hspace*{-.1cm} {\em Does Cole know/think that Julian dances salsa?''}
\end{exe}
\vspace*{-.2cm}

\noindent
{\bf Methods.} To investigate \citepos{mandelkern-etal2020} claim that a categorical factivity distinction emerges from naturalness ratings in EICs, we collected such ratings for 20 (non)factive predicates.

\noindent
\underline{Participants.} Data from 370 participants (ages: 19-80, mean: 40.7; 175 women, 185 men, 8 nonbinary, 2 undisclosed) were analyzed. Other data were excluded based on preregistered criteria.

\noindent
\underline{Materials.} Participants read two-sentence discourses consisting of an interrogative followed by a declarative, like (\ref{second}). In the target stimuli, the interrogatives combined the 20 factive and nonfactive predicates with the 20 complement clauses of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}. The preceding declaratives implemented a three-level context condition: In the `explicit ignorance' context (\ref{second}), the declarative sentence conveyed the speaker's ignorance about the CC: If the CC is presupposed, global accommodation (taken to be default in \citealt{heim83,vds92}) is not possible and the discourse is expected to be unnatural; if the CC is not presupposed, the discourse is expected to be natural. In the other two context levels, the declarative denoted a fact relative to which the CC had a `low prior probability' (\ref{sample}a) or a `high prior probability' (\ref{sample}b); these contexts were normed in \citealt{degen-tonhauser-openmind}. Presupposition analyses (e.g., \citealt{heim83,vds92}) invariably predict global accommodation for presupposed CCs, that is, high naturalness ratings.

\vspace*{-.2cm}
\begin{exe}
\ex\label{sample}
\begin{xlist}
\ex Ed: {\em ``Julian is German. Did Cole discover that Julian dances salsa?''} \hspace*{-.2cm} \hfill [low prior prob.]
\ex Ed: {\em ``Julian is Cuban. Did Cole discover that Julian dances salsa?''} \hfill [high prior prob.]
\end{xlist}
\end{exe}
\vspace*{-.2cm}

In addition to the 1,200 target stimuli (20 predicates $\times$ 20 CCs $\times$ 3 contexts), the materials included six fillers: Four with hard triggers ({\em too, also, again},  {\em it-}clefts), expected to be unnatural in EICs (\citealt{simons01,abusch10}), and two with {\em stop} and {\em continue}, which \citealt{simons01} considers soft triggers (natural in explicit ignorance contexts), contrary to \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}. The fillers were only presented in EICs.

\noindent
\underline{Procedure.} Participants rated the naturalness of ``[speaker's] question in this context'' on a slider from ``totally unnatural'' (coded 0) to ``totally natural'' (coded 1) for 30 stimuli: one target stimulus for each of the 20 predicates (each paired with a unique CC; 12 in EICs, and 4 each in low and high prior probability contexts), the six filler stimuli, and four acceptable control stimuli (used to identify participants not attending to the task). Trial order was randomized for each participant.
%\begin{exe}
%\ex Filler stimuli
%\begin{xlist}
%\ex I don't know if Ann plays any instrument. Does Ann play the flute, too?
%\ex I don't know if Svenja plays any sport. Does Svenja also play soccer?
%\ex I don't know if anyone was playing outside with the kids. Was it Jack who was playing outside with the kids?
%\ex I don't know if Stephen was ever in the habit of vaping. Has Stephen recently stopped vaping?
%\ex I don't know if John was ever reading ``Dune''. Has John recently continued reading ``Dune''?
%\ex I don't know if William was ever interested in history. Is William interested in history again?"
%\end{xlist}
%\end{exe}

%Finally, the experiment included 4 control stimuli, which were used to exclude participants not attending to the task. These were expected to be rated as natural/acceptable.

%Each participant saw a random set of 30 stimuli: Each set contained one target utterance for each of the 20 CCs (each paired with a unique clause-embedding predicates), the same 6 control stimuli, and the same 4  filler stimuli. 12 of the CCs were presented in the explicit ignorance context, 4 with a low prior probability fact, 4 with a high prior probability fact. Within-block trial order was randomized. After completing the experiment, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

%The experiment started off with four practice trials (2 natural, 2 unnatural; participants were given feedback on correct and incorrect answers).

\noindent
{\bf Results.} Contrary to \citepos{mandelkern-etal2020} claim, factives do not pattern alike, as shown in Fig.~\ref{fig:acc-by-expression}, which plots mean naturalness ratings in EICs by expression (colors: \color{orange}factives\color{black}, \color{gray}nonfactives\color{black}, hard triggers, \color{magenta}soft triggers\color{black}). Whereas the mean for {\em be annoyed} is at the level of the soft trigger {\em stop}, that of {\em know} is higher, and the means of {\em see, discover}, and {\em reveal} are as high as those of some nonfactive predicates. (These observations are confirmed by a linear mixed effects model, shown in the talk.) Thus, in line with \citealt{degen-tonhauser-language}, there is no empirical evidence for a categorical distinction between factive and non-factive predicates. (The talk also discusses that the results do not align with assumptions about hard and soft triggers; \citealt{simons01,abusch10}.)

Contrary to the predictions of presupposition analyses, the naturalness ratings of the purportedly factive predicates are sensitive to the low vs.\ high prior probability context manipulation. As shown in Fig.~\ref{fig:acc-by-context}, which plots mean naturalness ratings by context and expression, the mean naturalness ratings are higher in the high prior probability context than in the low one. This context-sensitivity, which is also observed for the nonfactive predicates (except {\em pretend}; model details provided in talk), replicates \citepos{degen-tonhauser-openmind} results, obtained using an inference rating task.

\noindent
{\bf Methodological and theoretical implications.} The recent challenge to the categorical factivity distinction by \citealt{degen-tonhauser-language} is not addressed by collecting naturalness ratings in EICs instead of inference ratings. And while naturalness ratings in EICs may provide insight into the meanings of attitude predicates, the interpretation of the results depends on the presumed linking function (as discussed in the talk; see, e.g., \citealt{sprouse2018}). The very low mean naturalness rating for {\em be annoyed} in EICs could, for instance, be due not to a semantic presupposition (see also \citealt{karttunen2016}) but because the CC is not-at-issue, contrary to what is suggested by the EIC.

Contemporary presupposition analyses (e.g., \citealt{heim83,vds92,abrusan2011,karttunen2016,simons-etal2017}) predict neither the heterogeneity of the naturalness ratings of factive predicates nor their sensitivity to the prior probability of the CC. These results, which mirror those based on inference ratings, suggest that projection analyses must incorporate more fine-grained distinctions between attitude predicates and the systematic effect of prior probabilities.

\newpage

\begin{figure}[h!]
\centering
\caption{\small{Mean naturalness rating in explicit ignorance context by expression (\color{orange}factive\color{black}, \color{gray}nonfactive\color{black}, \color{magenta} soft trigger\color{black}, \color{black}hard trigger\color{black}). Error bars indicate 95\% bootstrapped CIs. Violin plots indicate kernel probability density of participants' ratings.}}\label{fig:acc-by-expression}
\includegraphics[width=.9\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
\end{figure}

\vspace*{-1cm}

\begin{figure}[h!]
\centering
\caption{\small{Mean naturalness rating by context and predicate (\fcolorbox{black}{orange}{factive}, \fcolorbox{black}{lightgray}{nonfactive}); predicates ordered as in Fig.~\ref{fig:acc-by-expression}. Error bars indicate 95\% bootstrapped CIs. Violin plots indicate kernel probability density of participants' ratings}.}\label{fig:acc-by-context}
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/ORDER-by-LANGUAGE-naturalness-by-context-and-predicate}
\end{figure}

\vspace*{-.7cm}

\renewcommand\bibsection{\subsubsection*{\refname}}
\begin{tiny}
\bibliographystyle{cslipubs-natbib}
\bibliography{/Users/tonhauser.1/Documents/bibliography}
\end{tiny}

\end{document}

### 2023-08-01 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -31,111 +31,64 @@ length(unique(d$participantID)) #370 participants
 # plot of mean naturalness ratings in explicit ignorance context ----
 
 # target data: explicit ignorance context
+# merge the two controls into one, but exclude them
+table(d$expression)
 t = d %>%
-  filter(context == "explicitIgnorance")
+  filter(context == "explicitIgnorance") %>%
+  mutate(expression = recode(expression, "controlGood1" = "controls", "controlGood2" = "controls")) %>%
+  filter(expression != "controls")
 
-# merge the two controls into one
-table(t$expression)
-t = t %>%
-  mutate(expression = recode(expression, "controlGood1" = "controls", "controlGood2" = "controls"))
-
-# calculate mean naturalness rating by expression, including the controls
-nat.means = t %>%
+# calculate mean naturalness rating by expression, including the fillers, in explicit ignorance context
+nat.meansEIC = t %>%
   group_by(expression) %>%
   summarize(Mean = mean(response), CILow = ci.low(response), CIHigh = ci.high(response)) %>%
   mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, expression = fct_reorder(as.factor(expression),Mean))
-nat.means
-levels(nat.means$expression)
+nat.meansEIC
+levels(nat.meansEIC$expression)
 
-# color code factives and ps triggers
+# color code the expressions
 factives <- c("know", "discover", "be annoyed", "reveal", "see")
-triggers <- c("too", "also", "cleft", "again", "stop", "continue")
+hard.triggers <- c("too", "also","cleft","again")
+soft.triggers <- c("stop", "continue")
 
-nat.means$ps = ifelse(nat.means$expression %in% triggers, "trigger", 
-                      ifelse(nat.means$expression %in% factives, "factive", "other"))
+nat.meansEIC$ps = ifelse(nat.meansEIC$expression %in% soft.triggers, "softTrigger", 
+               ifelse(nat.meansEIC$expression %in% hard.triggers, "hardTrigger",
+                      ifelse(nat.meansEIC$expression %in% factives, "factive", "other")))
 
-table(nat.means$ps, nat.means$expression)
+t$ps = ifelse(t$expression %in% soft.triggers, "softTrigger", 
+              ifelse(t$expression %in% hard.triggers, "hardTrigger",
[…]

### 2023-08-04 writing/DGfS2023-abstract/abstract.tex
updated glossa psycholx analysis and paper

@@
 
-\citealt{mandelkern-etal2020} challenged this conclusion, arguing that gradient projection is an artifact of the inference task used, writing that ``inference tasks to some degree invite subjects to make the inference in question'' (p.498) and that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).
-They suggested that naturalness ratings of utterances with attitude predicates in explicit ignorance contexts (EICs), as in (\ref{second}), are more suitable to distinguishing the semantic presuppositions of factive predicates (hypothesized to be unnatural in EICs) from inferences that arise with nonfactive predicates ``for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497).
+\citealt{mandelkern-etal2020} challenged this conclusion, arguing that gradient projection is an artefact of the inference task used, writing that ``inference tasks to some degree invite subjects to make the inference in question'' (p.498) and that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).
+They suggested that naturalness ratings of utterances with attitude predicates in explicit ignorance contexts (EICs), as in (\ref{second}), are more suitable to distinguish the semantic presuppositions of factive predicates (hypothesized to be unnatural in EICs) from inferences that arise with nonfactive predicates ``for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497).
 
@@
 \noindent
-\underline{Materials.} Participants read two-sentence discourses consisting of an interrogative followed by a declarative, like (\ref{second}).
+\underline{Materials.} Participants read two-sentence discourses consisting of a declarative followed by an interrogative, like (\ref{second}).
 In the target stimuli, the interrogatives combined the 20 factive and nonfactive predicates with the 20 complement clauses of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}.

### 2023-08-17 writing/DGfS2023-abstract/abstract.tex
updated DGfS abstract

@@
 
-\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
-%\definecolor{gray}{rgb}{0.36, 0.54, 0.66}
+\definecolor{green}{RGB}{0,158,115}
+\definecolor{orange}{RGB}{213,94,0}
 
-\definecolor{Pink}{RGB}{240,0,120}
 \newcommand{\red}[1]{\textcolor{Pink}{#1}}
@@
 \citealt{mandelkern-etal2020} challenged this conclusion, arguing that gradient projection is an artefact of the inference task used, writing that ``inference tasks to some degree invite subjects to make the inference in question'' (p.498) and that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).
-They suggested that naturalness ratings of utterances with attitude predicates in explicit ignorance contexts (EICs), as in (\ref{second}), are more suitable to distinguish the semantic presuppositions of factive predicates (hypothesized to be unnatural in EICs) from inferences that arise with nonfactive predicates ``for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497).
+They suggested that naturalness ratings of utterances with attitude predicates in explicit ignorance contexts (EICs), as in (\ref{second}), are more suitable to distinguish the semantic presuppositions of factive predicates (hypothesized to be unnatural in EICs) from inferences that may arise with nonfactive predicates ``for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497).
 
@@
 \noindent
-\underline{Participants.} Data from 370 participants (ages: 19-80, mean: 40.7; 175 women, 185 men, 8 nonbinary, 2 undisclosed) were analyzed. Other data were excluded based on preregistered criteria.
+\underline{Participants.} 398 participants were recruited on Prolific. The data of 370 entered into the analysis.
 
@@
 \underline{Materials.} Participants read two-sentence discourses consisting of a declarative followed by an interrogative, like (\ref{second}).
-In the target stimuli, the interrogatives combined the 20 factive and nonfactive predicates with the 20 complement clauses of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}.
+In the target stimuli, the interrogatives combined the 20 (non)factive predicates with the 20 complement clauses of \citealt{degen-tonhauser-openmind,degen-tonhauser-language}.
 The preceding declaratives implemented a three-level context condition: In the `explicit ignorance' context (\ref{second}), the declarative sentence conveyed the speaker's ignorance about the CC: If the CC is presupposed, global accommodation (taken to be default in \citealt{heim83,vds92}) is not possible and the discourse is expected to be unnatural; if the CC is not presupposed, the discourse is expected to be natural.
@@
 
-In addition to the 1,200 target stimuli (20 predicates $\times$ 20 CCs $\times$ 3 contexts), the materials included six fillers: Four with hard triggers ({\em too, also, again},  {\em it-}clefts), expected to be unnatural in EICs (\citealt{simons01,abusch10}), and two with {\em stop} and {\em continue}, which \citealt{simons01} considers soft triggers (natural in explicit ignorance contexts), contrary to \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}.
+In addition to the 1,200 target stimuli (20 predicates $\times$ 20 CCs $\times$ 3 contexts), the materials included six fillers: Four with hard triggers ({\em too, also, again},  {\em it-}clefts), expected to be unnatural in EICs (\citealt{simons01,abusch10}), and two with {\em stop} and {\em continue}, which \citealt{simons01} considers soft triggers (natural in EICs), contrary to \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, who take them to be unnatural in EICs.
 The fillers were only presented in EICs.
@@
 \noindent
-{\bf Results.} Contrary to \citepos{mandelkern-etal2020} claim, factives do not pattern alike, as shown in Fig.~\ref{fig:acc-by-expression}, which plots mean naturalness ratings in EICs by expression (colors: \color{orange}factives\color{black}, \color{gray}nonfactives\color{black}, hard triggers, \color{magenta}soft triggers\color{black}).
+{\bf Results.} Contrary to \citepos{mandelkern-etal2020} claim, factives do not pattern alike, as shown in Fig.~\ref{fig:acc-by-expression}, which shows mean naturalness ratings in EICs by expression (colors: \color{orange}factives\color{black}, \color{gray}nonfactives\color{black}, hard triggers, \color{green}soft triggers\color{black}).
 Whereas the mean for {\em be annoyed} is at the level of the soft trigger {\em stop}, that of {\em know} is higher, and the means of {\em see, discover}, and {\em reveal} are as high as those of some nonfactive predicates.
-(These observations are confirmed by a linear mixed effects model, shown in the talk.) Thus, in line with \citealt{degen-tonhauser-language}, there is no empirical evidence for a categorical distinction between factive and non-factive predicates.
+(These observations are confirmed by a linear mixed effects model.) Thus, in line with \citealt{degen-tonhauser-language}, there is no empirical evidence for a categorical distinction between factive and non-factive predicates.
 (The talk also discusses that the results do not align with assumptions about hard and soft triggers; \citealt{simons01,abusch10}.)
@@
 Contrary to the predictions of presupposition analyses, the naturalness ratings of the purportedly factive predicates are sensitive to the low vs.\ high prior probability context manipulation.
-As shown in Fig.~\ref{fig:acc-by-context}, which plots mean naturalness ratings by context and expression, the mean naturalness ratings are higher in the high prior probability context than in the low one.
-This context-sensitivity, which is also observed for the nonfactive predicates (except {\em pretend}; model details provided in talk), replicates \citepos{degen-tonhauser-openmind} results, obtained using an inference rating task.
+As shown in Fig.~\ref{fig:acc-by-context}, which shows mean naturalness ratings by context and expression, the mean naturalness ratings are higher in the high prior probability context than in the low one.
+This context-sensitivity, which is also observed for the nonfactive predicates (except {\em pretend}), replicates \citepos{degen-tonhauser-openmind} results, obtained using an inference rating task.
 
 \noindent
-{\bf Methodological and theoretical implications.} The recent challenge to the categorical factivity distinction by \citealt{degen-tonhauser-language} is not addressed by collecting naturalness ratings in EICs instead of inference ratings.
-And while naturalness ratings in EICs may provide insight into the meanings of attitude predicates, the interpretation of the results depends on the presumed linking function (as discussed in the talk; see, e.g., \citealt{sprouse2018}).
-The very low mean naturalness rating for {\em be annoyed} in EICs could, for instance, be due not to a semantic presupposition (see also \citealt{karttunen2016}) but because the CC is not-at-issue, contrary to what is suggested by the EIC.
+{\bf Methodological and theoretical implications.} The recent challenge to the categorical factivity distinction by \citealt{degen-tonhauser-language} is not dispelled by collecting naturalness ratings in EICs instead of inference ratings.
+And while naturalness ratings in EICs may provide insight into the meanings of attitude predicates, the interpretation of the results depends on the presumed linking function (see, e.g., \citealt{sprouse2018}).
+Instead of the ratings directly reflecting whether the CC is presupposed, the very low mean rating for {\em be annoyed} in EICs could be due to its CC being not-at-issue while the EIC suggests that it is at-issue; likewise, reading {\em know} with a contrastive pitch accent in EICs may have resulted in about half of the participants giving a high rating.
 
-Contemporary presupposition analyses (e.g., \citealt{heim83,vds92,abrusan2011,karttunen2016,simons-etal2017}) predict neither the heterogeneity of the naturalness ratings of factive predicates nor their sensitivity to the prior probability of the CC.
-These results, which mirror those based on inference ratings, suggest that projection analyses must incorporate more fine-grained distinctions between attitude predicates and the systematic effect of prior probabilities.
+Contemporary presupposition analyses (e.g., \citealt{heim83,vds92,abrusan2011,karttunen2016,simons-etal2017}) predict neither the heterogeneity of the naturalness ratings of factive predicates nor their sensitivity to the prior probability of the CC (though see \citealt{schlenker2021} for the latter).
+These results, which mirror those based on inference ratings, suggest that projection analyses must incorporate more fine-grained distinctions between attitude predicates and the systematic effect of prior probabilities on projection.
 
@@
 \centering
-\caption{\small{Mean naturalness rating in explicit ignorance context by expression (\color{orange}factive\color{black}, \color{gray}nonfactive\color{black}, \color{magenta} soft trigger\color{black}, \color{black}hard trigger\color{black}).
+\caption{\small{Mean naturalness rating in explicit ignorance context by expression (\color{orange}factive\color{black}, \color{gray}nonfactive\color{black}, \color{green} soft trigger\color{black}, \color{black}hard trigger\color{black}).
 Error bars indicate 95\% bootstrapped CIs.

### 2023-08-17 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -28,6 +28,7 @@ table(d$context)
 
 length(unique(d$participantID)) #370 participants
 
+
 # Fig 1: plot of mean naturalness ratings in explicit ignorance context ----
 
 # target data: explicit ignorance context
@@ -62,9 +63,9 @@ t$ps = ifelse(t$expression %in% soft.triggers, "softTrigger",
 
 table(nat.meansEIC$ps, nat.meansEIC$expression)
 
-text.color <- ifelse(nat.meansEIC$expression[order(nat.meansEIC$Mean)] %in% factives, '#D55E00', 
-                     ifelse(nat.meansEIC$expression[order(nat.meansEIC$Mean)] %in% hard.triggers, "black", 
-                            ifelse(nat.meansEIC$expression[order(nat.meansEIC$Mean)] %in% soft.triggers, "#CC79A7",
+text.color <- ifelse(nat.meansEIC$expression[order(nat.meansEIC$Mean)] %in% factives, '#D55E00',
+                     ifelse(nat.meansEIC$expression[order(nat.meansEIC$Mean)] %in% hard.triggers, "black",
+                            ifelse(nat.meansEIC$expression[order(nat.meansEIC$Mean)] %in% soft.triggers, "#009E73",
                             "gray80")))
 text.color
 
@@ -75,13 +76,14 @@ ggplot(nat.meansEIC, aes(x=expression, y=Mean)) +
   geom_violin(data=t[t$context == "explicitIgnorance",],aes(x=expression, y=response),
                scale="width",color="gray80", fill = "gray80") +
   geom_point(aes(group = ps, fill = ps), shape=21,stroke=.5,size=3, color="black") +
-  scale_fill_manual(values=c('#D55E00','black','gray80','#CC79A7')) + 
+  scale_fill_manual(values=c('#D55E00','black','gray80','#009E73')) + 
   geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1,color="black") +
   scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0), labels = c("0",".2",".4",".6",".8","1")) +
   guides(fill=FALSE) +
   theme(legend.position="top") +
   ylab("Mean naturalness rating \n in explicit ignorance context") +
   xlab("Expression") +  
+  theme_dark() +
   theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = text.color)) 
[…]

### 2023-08-31 results/main/meta-analyses/new/1_projectivity/rscripts/graphs.R
SuB presentation updates

@@ -1,183 +1,202 @@
+##### GRAPHS FOR PROJECTION RATINGS #####
 
-### summary statistics and graphs for projection data ------
-
-# set wd to script dir
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-# load data
-data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
-
-# load helper functions
-source('../../../../helpers.R')
-
-# libraries for manipulating dataframes, and plotting
-library(dplyr)
-library(forcats)
-library(ggplot2)
-
-# projectivity by operator -----
-proj_means_op = data %>% group_by(op) %>%
-  summarize(Mean = mean(projective), CILow = ci.low(projective), 
-            CIHigh = ci.high(projective)) %>%
-  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
-         op = fct_reorder(as.factor(op),Mean)) %>% ungroup()
-
-proj_means_op %>% mutate(op = fct_reorder(op, Mean, 
-                                          .fun = mean)) %>% 
-  ggplot(aes(x = op, y=Mean)) +
-  # coord_cartesian(ylim = c(0,1)) +
-  geom_point(aes(shape = op), size = 1, color = "lightblue") +
-  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
-  geom_line(color = "lightblue") + 
-  labs(title = "Mean projectivity by operator")+
-  theme_bw() +
-  scale_color_brewer(palette = "PRGn")
-
-ggsave("../graphs/proj-by-op.pdf",height=4,width=5)
+# 1 libraries, helpers, settings
+# 2 projectivity by operator
+# 3 projectivity by predicate
+  ## 3a projectivity by predicate w distribution of participants' means
+  ## 3b projectivity by predicate w distribution of observations
+# 4 projectivity by predicate and operator
+  ## 4b projectivity means with distributions for observations
+  ## 4c projectivity means with distributions for paritipant means
+# 5 verb profiles
 
 
+# 1 Libraries,  helpers, settings -------------------------------------------
 
-# projectivity by verb -----
-proj_means_v = data %>% group_by(verb) %>%
[…]

### 2023-08-31 writing/DGfS2023-abstract/write-up.tex

@@
 
-\subsection{The effect of the prior probability of the context on naturalness ratings}
+\subsection{The effect of the prior probability of the CC on naturalness ratings}
 
@@
 
-%Here, I will plot, for each of the 20 predicates (target stimuli), the mean naturalness rating in the two contexts (explicit ignorance, neutral). 
+\newpage
+
+Here the predicates are ordered by the mean certainty rating (Language paper, Exp.1a).
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/ORDER-by-LANGUAGE-naturalness-by-context-and-predicate}
+\caption{Mean naturalness ratings by expression and context. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:acc-by-context}
+\end{figure}
+
 %
@@
 %
-%Third analysis: Do the naturalness ratings in explicit ignorance contexts and the ?certain that? ratings get at the same underlying theoretical construct (namely speaker commitment)?
+
+\subsection{Measuring speaker belief in the CC}
+
+Do the naturalness ratings in explicit ignorance contexts and the inference ratings get at the same underlying theoretical construct, namely speaker belief in the CC?
+
+Spearman rank correlation: -.62
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.7\textwidth]{../../results/main/13explicitIgnorance/graphs/mean-acceptability-against-mean-certainty}
+\caption{Mean naturalness ratings by mean certainty ratings. Error bars indicate 95\% bootstrapped confidence intervals.}\label{fig:acc-by-certain}
+\end{figure}
+
 %

### 2023-09-14 talks/XPrag-poster/xprag-poster.tex
added poster draft

\documentclass[final]{beamer}
%% Possible paper sizes: a0, a0b, a1, a2, a3, a4.
%% Possible orientations: portrait, landscape
%% Font sizes can be changed using the scale option.
\usepackage[size=a0,orientation=portrait]{beamerposter}

\usepackage{euler}
\usetheme{gemini}
% \usecolortheme{seagull}

% ====================
% Packages
% ====================

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgfplots}

% ====================
% Lengths
% ====================

% If you have N columns, choose \sepwidth and \colwidth such that
% (N+1)*\sepwidth + N*\colwidth = \paperwidth
\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.005\paperwidth}
\setlength{\colwidth}{0.48\paperwidth}

\newlength{\vboxsep}
\setlength{\vboxsep}{-.6\baselineskip}
\newlength{\mboxpreadjust}
\setlength{\mboxpreadjust}{-.4\baselineskip}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

\input{lisa-extra.tex}

%% Reference Sources
% \addbibresource{refs.bib}

\title{Projection variability of clausal complements across different operators}

\author{Lisa Hofmann\inst{1} \and Marie-Catherine de Marneffe\inst{2} \and Judith Tonhauser\inst{1}}

\institute[shortinst]{\inst{1} University of Stuttgart \samelineand \inst{2} UC Lovain}

\date{\today}

\begin{document}

\begin{frame}[t]

	\vspace{-2cm}
	\begin{columns}[t]
		\separatorcolumn

		\begin{column}{.74\colwidth}
			\begin{upshotbox}{Does the projection of content differ across entailment-canceling environments?}

				\begin{itemize}
					\item[\color{highlight}{$\blacktriangleright$}] \textbf{Yes!} Projection differs by entailment-cancelling \ophighlight{operator}

					\item[\color{highlight}{$\blacktriangleright$}] By-operator effects differ by predicate (\textbf{\ophighlight{operator}/\predhighlight{predicate} interaction})

					\item[\color{highlight}{$\blacktriangleright$}] Current theories of \cchighlight{projective content} do not predict our results
				\end{itemize}

			\end{upshotbox}
			\vspace{\vboxsep}
			\begin{normalbox}{Projection of clausal complements}
				Do you infer that Rachel is committed to the truth of the \textit{content of the complement} (CC), that \cchighlight{Julian dances salsa}?

				\vspace{-.3\baselineskip}
				\ex. \a. Rachel: \emph{\lq Does Cole \predhighlight{know} that \cchighlight{Julian dances salsa}?\rq}\\
					\cmark\ \  Yes, CC projects out of the question
					\b. Rachel: \emph{\lq Does Cole \predhighlight{think} that \cchighlight{Julian dances salsa}?\rq}\\
					\xmark\ \ No, CC does not project
					\z.
				\z.

				\vspace{-.5\baselineskip}
				\resizebox{\linewidth}{!}{
					\begin{minipage}{1.7\linewidth}
						\citet{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_observations_1971,karttunen_conventional_1979}, and many more

					\end{minipage}
				}
			\end{normalbox}

			\vspace{\vboxsep}
			\begin{normalbox}{Entailment-cancelling operators}
				\textbf{Family-of-sentences test:}\newline No mention of differences in projection between different \ophighlight{operators}

				\vspace{-.5\baselineskip}
				\ex. \a. Polar question:\\
					\textit{\ophighlight{Does} Cole \predhighlight{know} that \cchighlight{Julian dances salsa}?}
					\b. Negation:\\
					\textit{Cole \ophighlight{doesn't} \predhighlight{know} that \cchighlight{Julian dances salsa}.}
					\b. Epistemic modal:\\
					\textit{\ophighlight{Perhaps} Cole \predhighlight{knows} that \cchighlight{Julian dances salsa}.}
					\b. Conditional antecedents:\\
					\textit{\ophighlight{If} Cole \predhighlight{knows} that \cchighlight{Julian dances salsa}, Logan will be joyful.}
					\z.
				\z.

				\vspace{-.5\baselineskip}
				\resizebox{\linewidth}{!}{
					\begin{minipage}{1.7\linewidth}
						(e.g. \citealt{chierchia_meaning_1990,coppock_invitation_2020})

					\end{minipage}
				}

			\end{normalbox}

			\vspace{\vboxsep}
			\begin{normalbox}{Hints at by-operator variation}
				\textbf{Factive vs. semi-factive predicates} (\citealt{karttunen_observations_1971})
				\vspace{-.2\baselineskip}
				\begin{itemize}
					\item Factives (\textit{be annoyed, regret, \dots}): CC projects across all four operators
					\item Semi-factives (\textit{discover, realize, see, notice, \dots}):\newline CC projects across negation, but not always for the other operators
				\end{itemize}

				\textbf{Experiment with English projective contents} (\citealt{smith_relationship_2014})
				\vspace{-.2\baselineskip}
				\begin{itemize}
					\item Projective content of epithets (e.g. \textit{idiot}) and the CC of \textit{know}:\\ more projective under negation than conditionals
					\item Opposite pattern for appositive relative clauses and \textit{win}
				\end{itemize}

				\textbf{Experiment with German clause-embedding predicates}\\ (\citealt{sieker_projective_2022})
				\vspace{-.2\baselineskip}
				\begin{itemize}
					\item Higher projection ratings w/ negation than other three operators
					\item No by-predicate variation, no evidence for factive/semi-factive distinction
				\end{itemize}

				\begin{figure}[h]
					\vspace{-.2\baselineskip}
					\centering
					\includegraphics[width=.7\linewidth]{sieker-solstad.png}
					\vspace{-.4\baselineskip}
					\caption{\citealt{sieker_projective_2022}, p. 286: Projection-ratings by embedding operator, for purported factive and semi-factive predicates}
					% \label{fig:figure1}
				\end{figure}
				\vspace{-.8\baselineskip}
			\end{normalbox}

			\vspace{\vboxsep}
			\begin{normalbox}{Certain-that task for projection inferences}

				\begin{figure}[h]
					\centering
					\includegraphics[width=.6\linewidth]{task-1n-proj.eps}
					% \caption{Caption here}
					\label{fig:task}
				\end{figure}

				\resizebox{\linewidth}{!}{
					\begin{minipage}{1.7\linewidth}
						\citet{tonhauser_prosodic_2016,djarv_prosodic_2017,tonhauser_how_2018,de_marneffe_commitmentbank_2019,mahler_social_2020,degen_are_2022,sieker_projective_2022}

					\end{minipage}
				}

			\end{normalbox}

			\vspace{\vboxsep}
			\begin{normalbox}{Variation among clause-embedding predicates}
				20 \predhighlight{predicates} that have shown projection variability in PQs\newline (\citealt{degen_are_2022})

				\begin{figure}[h]
					\centering
					\includegraphics[width=.95\linewidth]{degen-tonhauser.png}
					\caption{\citealt{degen_are_2022}, p. 562: Mean certainty ratings by predicate}
					\label{fig:task}
				\end{figure}
				\vspace{-.8\baselineskip}
			\end{normalbox}

			\vspace{\vboxsep}
			\begin{normalbox}{Materials}
				To assess the effect of \ophighlight{operator} and \predhighlight{predicate} on \cchighlight{\textbf{projection}}:
				\vspace{.2\baselineskip}

					\textbf{4 experiments} (roughly $750$ participants each)
					\vspace{-.3\baselineskip}
					\begin{itemize}
						\item One per \ophighlight{operator}: polar questions, negation, modal \textit{perhaps}, conditional
					\end{itemize}

					\vspace{-.2\baselineskip}
					Participants saw:
					\vspace{-.3\baselineskip}
					\begin{itemize}
						\item \textbf{20 clause embedding \predhighlight{predicates}}

						\begin{itemize}
							\item  Crossed with 20 CCs ($20 \times 20 = 400$ combinations)
						\end{itemize}

						\item (6 controls for exclusion)

					\end{itemize}
				\vspace{-.2\baselineskip}
				{\small (Experiments also used different at-issueness measures in separate block, not analyzed here)}

			\end{normalbox}

		\end{column}

		\separatorcolumn

		\begin{column}{1.25\colwidth}
			\begin{normalbox}{Main effect of embedding operator \& By-predicate variation in the effect of operator}
				% \vspace{-\baselineskip}
				\hspace{-.2cm}\begin{tabular}{p{.70\linewidth} p{.3\linewidth}}
					\textcolor{highlight}{\large \Raleway By-operator variation aggregating across predicates (Figure 3)}
					\begin{itemize}
						\item Conditional > Question > Negation, Modal
							\vspace{\mboxpreadjust}
							\begin{modelbox}{Model \#1: Linear mixed effect regression}
								\footnotesize
								response: \textbf{certainty ratings}; fixed effect: \ophighlight{\textbf{operator}} (base level: Question);
								random intercepts: participants, items; \newline
								MLEs: question (intercept) $0.51$, conditional $+0.05$, modal $-0.04$, negation $-0.03$; with all $p < 0.001$
							\end{modelbox}

						\item But small differences, as in Sieker \& Solstad’s (2022) study
						\item Sieker \& Solstad’s results for German: Negation > Question, Conditional, Modal

					\end{itemize}

				\textcolor{highlight}{\large \Raleway Effect of operator differs by predicate (Figure 4), e.g.}
				\begin{itemize}
					\item CC of \predhighlight{be annoyed}: Negation, Conditional > Question, Modal
						\vspace{\mboxpreadjust}
						\begin{modelbox}{Model \#2: Linear mixed effect regression}
							\footnotesize
							response: \textbf{certainty ratings}; fixed effects: \ophighlight{operator}, \predhighlight{predicate}, and interaction (base level: \textbf{be annoyed} / negation); random intercepts: participant; \newline
							MLEs: negation (intercept) $0.87$, conditional $-0.12$, modal $-0.16$; with $p < 0.001$; question $+0.02$ (n.s.)
						\end{modelbox}

					\item CC of \predhighlight{know}: Question > Negation, Conditional > Modal
						\vspace{\mboxpreadjust}
						\begin{modelbox}{Model \#3: Linear mixed effect regression}
							\footnotesize
							response: \textbf{certainty ratings}; fixed effects: \ophighlight{operator}, \predhighlight{predicate}, and interaction (base level: \textbf{know} / negation); random intercepts: participant; \newline
							MLEs: negation (intercept) $0.79$, modal $-0.14$, question $+0.08$; with $p < 0.001$; , conditional $+/- 0$, (n.s.)
						\end{modelbox}

					\item CC of \predhighlight{discover}: Modal > Negation > Conditional, Question
						\vspace{\mboxpreadjust}
						\begin{modelbox}{Model \#4: Linear mixed effect regression}
							\footnotesize
							response: \textbf{certainty ratings}; fixed effects: \ophighlight{operator}, \predhighlight{predicate}, and interaction (base level: \textbf{discover} / negation); random intercepts: participant; \newline
							MLEs: negation (intercept) $0.68$, conditional $+0.11$, modal $-0.06$, question $+0.10$; with $p < 0.001$
						\end{modelbox}

				\end{itemize}

					&

					\vspace{-1.25\baselineskip}
					\begin{figure}[h]
						% \hspace{-1cm}
						\centering
						\includegraphics[width=.9\linewidth]{projective-op.pdf}
						\vspace{-\baselineskip}
						\caption{Mean certainty ratings by operator}
						\label{fig:figure2}
					\end{figure}
				\end{tabular}
				\vspace{-\baselineskip}
				\begin{figure}[h]
					\centering
					\includegraphics[width=\linewidth]{projective-pred-op.pdf}
					\vspace{-2\baselineskip}
					\caption{Mean certainty ratings by predicate, grouped by operator}
					\label{fig:figure3}
				\end{figure}
				\vspace{-.8\baselineskip}
			\end{normalbox}

			\vspace{\vboxsep}	
			\begin{normalbox}{Discussion --- By predicate variation in the effect of operator}
				\begin{itemize}
					\item Concurs with \citet{smith_relationship_2014}, who found content/operator interactions for English projective contents
					\item Differs from \citet{sieker_projective_2022}, who found no predicate/operator interaction for CCs of German clause-embedding predicates
				\end{itemize}

				\textbf{No evidence for factive vs. semi-factive distinction (\citealt{karttunen_observations_1971})}
				\begin{itemize}
					\item CC of purported factive \textit{be annoyed} does not invariably project across operators
					\item CC of purported semi-factives (\textit{discover, see}) do not project more across negation than other operators
				\end{itemize}

				\textbf{Provides support (from negation, modals, conditionals) for Degen \& Tonhauser’s (2022) result:}
				\begin{itemize}
					\item Projection does not categorically differentiate between (semi-)factive/-factive predicates 
				\end{itemize}

			\end{normalbox}

			\vspace{\vboxsep}
			\begin{normalbox}{Do existing theories of projection predict our results?}
				\textbf{Dynamic accounts of projection: Lexical factivity \& dynamic operators} \hfill (\citealt{heim_projection_1983,van_der_sandt_presupposition_1992})

				Distinguish factive and non-factive predicates:
				\begin{itemize}
					\item factive predicates (\textit{be annoyed, regret, \dots}): CC conventionally required to be contextually entailed in common ground

					\item non-factive predicates (\textit{believe, say, …}): no such requirement

				\end{itemize}

				Factive content projects globally, unless not admitted by common ground

				\textbf{Entailments \& discourse structure} \hfill (\citealt{abrusan_predicting_2011,simons_best_2017})

				\textbf{Contextual entailments \& Epistemic preconditions} \hfill (\citealt{schlenker_triggering_2021})

			\end{normalbox}

			\vspace{\vboxsep}
			\begin{upshotbox}{Theoretical implications}
				\begin{itemize}
					\item Previous work: projection analyses need to consider the effect of \textbf{lexical meaning} (e.g. \citealt{kiparsky_fact_1970,karttunen_observations_1971}, et. seq.), \textbf{world knowledge} (\citealt{de_marneffe_did_2012,degen_prior_2021}), and \textbf{discourse structure} (e.g. \citealt{simons_best_2017,tonhauser_how_2018})

					\item Add to that the effect of various \ophighlight{entailment-cancelling operators}

					\item An analysis of projection should be able to address \ophighlight{operator} / \predhighlight{predicate} interaction effects. None of the extant projection analyses capture our data.

				\end{itemize}
			\end{upshotbox}

			\vspace{\vboxsep}
			\begin{normalbox}{References}
				\tiny

				\textbf{Márta Abrusán}. Predicting the presuppositions of soft triggers. \textit{Linguistics and philosophy}, 34:491–535, 2011. \quad \textbullet \quad
				%
				\textbf{Gennaro Chierchia and Sally McConnell-Ginet.} \textit{Meaning and grammar: An introduction to semantics.} 1990. \quad \textbullet \quad
				%
				\textbf{Elizabeth Coppock and Lucas Champollion.} \textit{Invitation to formal semantics.} online publication, in progress, 2020. \quad \textbullet \quad
				%
				\textbf{Marie de Marneffe, Mandy Simons, and Judith Tonhauser.} The CommitmentBank: Investigating projection in naturally occurring discourse. \textit{Sinn und Bedeutung}, 23:107–124, 2019. \quad \textbullet \quad
				%
				\textbf{Marie-Catherine de Marneffe, Christopher D. Manning, and Christopher Potts.} Did It Happen? The Pragmatic Complexity of Veridicality Assessment. \textit{Computational Linguistics}, 38(2):301–333, February 2012. \quad \textbullet \quad
				%
				\textbf{Judith Degen and Judith Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 5:59–70, 2021. \quad \textbullet \quad
				%
				\textbf{Judith Degen and Judith Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 98(3):552–591, 2022. \quad \textbullet \quad
				%
				\textbf{Kajsa Djärv and Hezekiah Akiva Bacovcin.} Prosodic effects on factive presupposition projection. \textit{Semantics and Linguistic Theory}, 27:116–133, 2017. \quad \textbullet \quad
				%
				\textbf{Gottlob Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, pages 25–50, 1892. \quad \textbullet \quad
				%
				\textbf{Irene Heim.} On the projection problem for presuppositions. \textit{Formal semantics–the essential readings}, pages 249–260, 1983. \quad \textbullet \quad
				%
				\textbf{Lauri Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 4(1):55–69, 1971. \quad \textbullet \quad
				%
				\textbf{Lauri Karttunen and Stanley Peters.} Conventional implicature. In Choon-Kyu Oh and David A. Dinneen, editors, \textit{Presuppositions, Syntax and Semantics} Vol.11, pages 1–56. New York: Academic Press, 1979. \quad \textbullet \quad
				%
				\textbf{Paul Kiparsky and Carol Kiparsky.} Fact. In Manfred Bierwisch and Karl Erich Heidolph, editors, \textit{Progress in Linguistics}, pages 143–173. The Hague: Mouton, 1970. \quad \textbullet \quad
				%
				\textbf{Taylor Mahler.} The social component of projection behavior of clausal complements. \textit{Linguistic Society of America}, 5:777–791, 2020. \quad \textbullet \quad
				%
				\textbf{Philippe Schlenker.} Triggering presuppositions. \textit{Glossa: a journal of general linguistics}, 6(1), 2021. \quad \textbullet \quad
				%
				\textbf{Judith Sieker and Torgrim Solstad.} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study. \textit{In Proceedings of the 23rd Amsterdam Colloquium}, 2022. \quad \textbullet \quad
				%
				\textbf{Mandy Simons, David Beaver, Craige Roberts, and Judith Tonhauser.} The best question: Explaining the projection behavior of factives. \textit{Discourse processes}, 54(3):187–206, 2017. \quad \textbullet \quad
				%
				\textbf{E. Allyn Smith and Kathleen Currie Hall.} The relationship between projection and embedding environment. \textit{In Proceedings of the 48th Meeting of the Chicago Linguistics Society}. 2014. \quad \textbullet \quad
				%
				\textbf{Peter F. Strawson.} On referring. \textit{Mind}, 59(235):320–344, 1950. \quad \textbullet \quad
				%
				\textbf{Judith Tonhauser.} Prosodic cues to presupposition projection. In \textit{Semantics and Linguistic Theory}, volume 26, pages 934–960, 2016. \quad \textbullet \quad
				%
				\textbf{Judith Tonhauser, David I. Beaver, and Judith Degen.} How projective is projective content? Gradience in projectivity and at-issueness. \textit{Journal of Semantics}, 35(3):495–542, 2018. \quad \textbullet \quad
				%
				\textbf{Rob A. van der Sandt.} Presupposition projection as anaphora resolution. \textit{Journal of semantics}, 9(4):333–377, 1992.

			\end{normalbox}

			\begin{normalbox}{References (not visible later)}
				\bibliographystyle{plainnat}
				\bibliography{projection}
			\end{normalbox}

		\end{column}
		\separatorcolumn
	\end{columns}

\end{frame}
\end{document}

### 2023-09-15 talks/XPrag-poster/xprag-poster.tex
XPRAG poster draft

@@
-\documentclass[final]{beamer}
+\documentclass[final, table]{beamer}
 %% Possible paper sizes: a0, a0b, a1, a2, a3, a4.
@@
 \usepackage{euler}
+% \usepackage[table]{xcolor}
 \usetheme{gemini}
@@
 		\begin{column}{1.25\colwidth}
-			\begin{normalbox}{Main effect of embedding operator \& By-predicate variation in the effect of operator}
+			\begin{normalbox}{\phantom.\hfill Effects of operator \& predicate on projection}
 				% \vspace{-\baselineskip}
 				\hspace{-.2cm}\begin{tabular}{p{.70\linewidth} p{.3\linewidth}}
-					\textcolor{highlight}{\large \Raleway By-operator variation aggregating across predicates (Figure 3)}
-					\begin{itemize}
+					\textcolor{highlight}{\large \Raleway \bfseries \underline{By-operator variation aggregating across predicates} (Figure 3)}
+					\vspace{-.15\baselineskip}
+					\begin{itemize} \small
 						\item Conditional > Question > Negation, Modal
@@
 							\end{modelbox}
+							\vspace{\mboxpreadjust}
 
@@
 
-				\textcolor{highlight}{\large \Raleway Effect of operator differs by predicate (Figure 4), e.g.}
-				\begin{itemize}
+				\vspace{-.1\baselineskip}
+				\textcolor{highlight}{\large \Raleway \bfseries\underline{Effect of operator differs by predicate} (Figure 4), e.g.}
+				\vspace{-.15\baselineskip}
+				\begin{itemize}\small
 					\item CC of \predhighlight{be annoyed}: Negation, Conditional > Question, Modal
@@
 							\footnotesize
-							response: \textbf{certainty ratings}; fixed effects: \ophighlight{operator}, \predhighlight{predicate}, and interaction (base level: \textbf{be annoyed} / negation); random intercepts: participant; \newline
-							MLEs: negation (intercept) $0.87$, conditional $-0.12$, modal $-0.16$; with $p < 0.001$; question $+0.02$ (n.s.)
+							response: \textbf{certainty ratings}; fixed effects: \ophighlight{operator}, \predhighlight{predicate}, and interaction (base lvl: \textbf{be annoyed} / negation); random intercepts: participant;
+							MLEs: negation (intercept) $0.87$, conditional $-0.12$, modal $-0.16$; ($p < 0.001$); question $+0.02$ (n.s.)
 						\end{modelbox}
@@
 
-					\vspace{-1.25\baselineskip}
+					\vspace{-1.5\baselineskip}
 						\begin{figure}[h]
-						% \hspace{-1cm}
+							\resizebox{.85\linewidth}{!}{
 							\centering
-						\includegraphics[width=.9\linewidth]{projective-op.pdf}
-						\vspace{-\baselineskip}
+							\begin{tikzpicture}
+							    \node at (0,0) {\includegraphics[width=.9\linewidth]{projective-op.pdf}};
+							    % \draw[red,ultra thick,rounded corners] (7.5,5.3) rectangle (9.4,6.2);
+							    \draw[red, ultra thick] (-3.37,2.3) -- (-3.37,1) -- node[below] {***} (-.66,1) -- (-.66,1.5);
+							    \draw[red, ultra thick] (-.66,1.5) -- (-.66,.2) -- node[below] {***} (2.07,.2) -- (2.07,.7);
+							    \draw[red, ultra thick] (2.07,.8) -- (2.07,.12) -- node[below] {\footnotesize n.s.} (4.78,.12) -- (4.78,.62);
+							\end{tikzpicture}}
+							\vspace{-1.4\baselineskip}
 							\caption{Mean certainty ratings by operator}
@@
 						\end{figure}
+
 				\end{tabular}
-				\vspace{-\baselineskip}
+				\vspace{-1.6\baselineskip}
 				\begin{figure}[h]
@@
 					\includegraphics[width=\linewidth]{projective-pred-op.pdf}
-					\vspace{-2\baselineskip}
+					\vspace{-2.2\baselineskip}
 					\caption{Mean certainty ratings by predicate, grouped by operator}
@@
 			\vspace{\vboxsep}	
-			\begin{normalbox}{Discussion --- By predicate variation in the effect of operator}
+			\begin{normalbox}{\phantom.\hfill Discussion --- By-predicate variation in the effect of operator}
 				\begin{itemize}
 					\item Concurs with \citet{smith_relationship_2014}, who found content/operator interactions for English projective contents
-					\item Differs from \citet{sieker_projective_2022}, who found no predicate/operator interaction for CCs of German clause-embedding predicates
+					\item Differs from \citet{sieker_projective_2022}: found no predicate/operator interaction for CCs of German clause-embedding predicates
 				\end{itemize}
@@
 				\textbf{No evidence for factive vs. semi-factive distinction (\citealt{karttunen_observations_1971})}
+				\vspace{\mboxpreadjust}
 				\begin{itemize}
@@
 				\textbf{Provides support (from negation, modals, conditionals) for Degen \& Tonhauser’s (2022) result:}
+				\vspace{\mboxpreadjust}
 				\begin{itemize}
@@
 			\vspace{\vboxsep}
-			\begin{normalbox}{Do existing theories of projection predict our results?}
-				\textbf{Dynamic accounts of projection: Lexical factivity \& dynamic operators} \hfill (\citealt{heim_projection_1983,van_der_sandt_presupposition_1992})
+			\begin{normalbox}{\phantom.\hfill Existing theories of projection do not predict our results}
+				\centering
+				\begin{theorybox}{Dynamic accounts of projection: Lexical triggering + dynamic semantics \hfill (\citealt{heim_projection_1983,van_der_sandt_presupposition_1992})}
+					\small Distinguish factive and non-factive predicates:
+					\vspace{-.3\baselineskip}
+					\begin{itemize}
+						\item \underline{factive} predicates (\textit{be annoyed, regret, \dots}): CC conventionally required to be contextually entailed in common ground
+						\item \underline{non-factive} predicates (\textit{believe, say, …}): no such requirement
+					\end{itemize}
+					\vspace{-.2\baselineskip}
+					Factive content projects globally, unless not admitted by common ground
+				\end{theorybox}
 
-				Distinguish factive and non-factive predicates:
+				\begin{theorybox}{Lexical entailments + discourse-based triggering \hfill (\citealt{abrusan_predicting_2011,simons_best_2017})}
+					\small Distinguish veridical predicates (CC is entailed) from non-veridical ones:
+					\vspace{-.3\baselineskip}
 					\begin{itemize}
-					\item factive predicates (\textit{be annoyed, regret, \dots}): CC conventionally required to be contextually entailed in common ground
-
-					\item non-factive predicates (\textit{believe, say, …}): no such requirement
-
+						\item \underline{veridical} predicates (\textit{be right, demonstrate, \dots}): entailed CC projects if not at-issue
+						\item \underline{non-veridical} predicates (\textit{believe, say, \dots}): no predictions / CC projects if required by discourse coherence
 					\end{itemize}
 
-				Factive content projects globally, unless not admitted by common ground
+				\end{theorybox}
 
-				\textbf{Entailments \& discourse structure} \hfill (\citealt{abrusan_predicting_2011,simons_best_2017})
+				\begin{theorybox}{Contextual entailments + triggering based on cognitive inertness \hfill (\citealt{schlenker_triggering_2021})}
+					\small Potential of projection for contents that are contextually entailed (given a context $C$ and the utterance $U$), including inferences from:
+					\vspace{-.3\baselineskip}
+					\begin{itemize}
+						\item Lexically veridical predicates
+						\item \lq Distributed veridicality\rq\ contexts (\citealt{roberts_i_2019}) \hfill (\textit{Cole \{was not wrong, can’t believe\} that Julian dances salsa.})
+						\item Other sources of contextual inference \hfill ((Cole is honest + knowledgeable.) \textit{Cole said that Julian dances salsa.})
+					\end{itemize}
+					\vspace{-.2\baselineskip}
+					Contextually entailed CC projects if it is an epistemic precondition of $U$ in $C$ (it is typically/likely already known).
 
-				\textbf{Contextual entailments \& Epistemic preconditions} \hfill (\citealt{schlenker_triggering_2021})
+				\end{theorybox}
+
+				\resizebox{\linewidth}{!}{
+						\begin{tabular}{l c c c}
+							\textbf{Our data} 
+								& \multicolumn{3}{c}{\textbf{Predictions}}\\ \toprule
+							%
+								& \parbox{.3\linewidth}{\centering Lexical triggering + \par dynamic semantics}
+									& \parbox{.3\linewidth}{\centering Lexical entailments +\par discourse-based triggering}
+										& \parbox{.3\linewidth}{\centering Contextual entailments +\par triggering based on cognitive inertness} \\ \midrule
+							%
+							\parbox{.3\linewidth}{\raggedright\textbf{Superadditive predicate/operator interaction}\\[2.5\baselineskip]}
+								& \parbox{.29\linewidth}{\centering \cellcolor{red!25} Meaning of entailment-canceling operators (invariably) encodes interaction with conventional content of embedded factives}
+									& \parbox{.29\linewidth}{\centering \cellcolor{red!25} No systematic predictions for how veridicality or at-issueness interact with the meaning of entailment-canceling operators}
+										& \parbox{.29\linewidth}{\centering\cellcolor{green!25} May be extended to our data by making explicit how combinations of operator + predicate are associated with contextual inferences}\\
+							%
+							\parbox{.3\linewidth}{\raggedright\textbf{Projection variability for all predicates}\par {\small Projection for some non-factive (/non-veridical) predicates as high as for some factive (/veridical) ones (see also D\&T'22)}\\[.5\baselineskip]}
+								& \parbox{.29\linewidth}{\centering \cellcolor{red!25} No predictions for non-factive predicates}
+									& \parbox{.29\linewidth}{\centering \cellcolor{red!25} No systematic predictions for non-veridical predicates}
+										& \parbox{.29\linewidth}{\centering \cellcolor{green!25} Makes predictions about CCs of all clause-embedding predicates} \\
+							%
+							\parbox{.3\linewidth}{\raggedright\textbf{Out-of-the-blue contexts}\par {\small Projection variability in the out-of-the-blue contexts used in experiment (see also D\&T'22)}\\[1.7\baselineskip]}
+								& \parbox{.29\linewidth}{\centering \cellcolor{red!25} Consistent projection of factive CCs}
+									& \parbox{.3\linewidth}{\centering \cellcolor{green!25} Veridical predicates: analyses may be extended by assuming that the CCs of veridical predicates differ in at-issueness in out-of-the-blue contexts}
+										& \parbox{.29\linewidth}{\centering \cellcolor{red!25} “Out-of-the-blue” contexts do not warrant assumption of contextual \underline{entailment}: No projection expected} \\
+						\end{tabular}
+				}
 
@@
 
-					\item An analysis of projection should be able to address \ophighlight{operator} / \predhighlight{predicate} interaction effects. None of the extant projection analyses capture our data.
+					\item An analysis of projection needs to be able to address \ophighlight{operator} / \predhighlight{predicate} interaction effects.
 
@@
 			\vspace{\vboxsep}
-			\begin{normalbox}{References}
+			\begin{normalbox}{\phantom.\hfill References}
 				\tiny
 
-				\textbf{Márta Abrusán}. Predicting the presuppositions of soft triggers. \textit{Linguistics and philosophy}, 34:491–535, 2011. \quad \textbullet \quad
+				\textbf{Márta Abrusán}. Predicting the presuppositions of soft triggers. \textit{Linguistics and philosophy}, 34:491–535, 2011.\quad\textbullet\quad
 				%
-				\textbf{Gennaro Chierchia and Sally McConnell-Ginet.} \textit{Meaning and grammar: An introduction to semantics.} 1990. \quad \textbullet \quad
+				\textbf{Gennaro Chierchia and Sally McConnell-Ginet.} \textit{Meaning and grammar: An introduction to semantics.} 1990.\quad\textbullet\quad
 				%
-				\textbf{Elizabeth Coppock and Lucas Champollion.} \textit{Invitation to formal semantics.} online publication, in progress, 2020. \quad \textbullet \quad
+				\textbf{Elizabeth Coppock and Lucas Champollion.} \textit{Invitation to formal semantics.} online publication, in progress, 2020.\quad\textbullet\quad
 				%
 				\textbf{Marie de Marneffe, Mandy Simons, and Judith Tonhauser.} The CommitmentBank: Investigating projection in naturally occurring discourse.
-\textit{Sinn und Bedeutung}, 23:107–124, 2019.
-\quad \textbullet \quad
+\textit{Sinn und Bedeutung}, 23:107–124, 2019.\quad\textbullet\quad
 				%
@@
 The Pragmatic Complexity of Veridicality Assessment.
-\textit{Computational Linguistics}, 38(2):301–333, February 2012.
-\quad \textbullet \quad
+\textit{Computational Linguistics}, 38(2):301–333, February 2012.\quad\textbullet\quad
 				%
-				\textbf{Judith Degen and Judith Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 5:59–70, 2021. \quad \textbullet \quad
+				\textbf{Judith Degen and Judith Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 5:59–70, 2021.\quad\textbullet\quad
 				%
-				\textbf{Judith Degen and Judith Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 98(3):552–591, 2022. \quad \textbullet \quad
+				\textbf{Judith Degen and Judith Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 98(3):552–591, 2022.\quad\textbullet\quad
 				%
-				\textbf{Kajsa Djärv and Hezekiah Akiva Bacovcin.} Prosodic effects on factive presupposition projection. \textit{Semantics and Linguistic Theory}, 27:116–133, 2017. \quad \textbullet \quad
+				\textbf{Kajsa Djärv and Hezekiah Akiva Bacovcin.} Prosodic effects on factive presupposition projection. \textit{Semantics and Linguistic Theory}, 27:116–133, 2017.\quad\textbullet\quad
 				%
-				\textbf{Gottlob Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, pages 25–50, 1892. \quad \textbullet \quad
+				\textbf{Tom Roberts.} I can’t believe it’s not lexical: Deriving distributed veridicality.
+Semantics and Linguistic Theory, \textit{Semantics and Linguistic Theory}, 29:665–685, 2019.\quad\textbullet\quad
 				%
-				\textbf{Irene Heim.} On the projection problem for presuppositions. \textit{Formal semantics–the essential readings}, pages 249–260, 1983. \quad \textbullet \quad
+				\textbf{Gottlob Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, pages 25–50, 1892.\quad\textbullet\quad
 				%
-				\textbf{Lauri Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 4(1):55–69, 1971. \quad \textbullet \quad
+				\textbf{Irene Heim.} On the projection problem for presuppositions. \textit{Formal semantics–the essential readings}, pages 249–260, 1983.\quad\textbullet\quad
+				%
+				\textbf{Lauri Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 4(1):55–69, 1971.\quad\textbullet\quad
 				%
@@
 Dinneen, editors, \textit{Presuppositions, Syntax and Semantics} Vol.11, pages 1–56.
-New York: Academic Press, 1979.
-\quad \textbullet \quad
+New York: Academic Press, 1979.\quad\textbullet\quad
 				%
@@
 In Manfred Bierwisch and Karl Erich Heidolph, editors, \textit{Progress in Linguistics}, pages 143–173.
-The Hague: Mouton, 1970.
-\quad \textbullet \quad
+The Hague: Mouton, 1970.\quad\textbullet\quad
 				%
-				\textbf{Taylor Mahler.} The social component of projection behavior of clausal complements. \textit{Linguistic Society of America}, 5:777–791, 2020. \quad \textbullet \quad
+				\textbf{Taylor Mahler.} The social component of projection behavior of clausal complements. \textit{Linguistic Society of America}, 5:777–791, 2020.\quad\textbullet\quad
 				%
-				\textbf{Philippe Schlenker.} Triggering presuppositions. \textit{Glossa: a journal of general linguistics}, 6(1), 2021. \quad \textbullet \quad
+				\textbf{Philippe Schlenker.} Triggering presuppositions. \textit{Glossa: a journal of general linguistics}, 6(1), 2021.\quad\textbullet\quad
 				%
 				\textbf{Judith Sieker and Torgrim Solstad.} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study.
-\textit{In Proceedings of the 23rd Amsterdam Colloquium}, 2022.
-\quad \textbullet \quad
+\textit{In Proceedings of the 23rd Amsterdam Colloquium}, 2022.\quad\textbullet\quad
 				%
 				\textbf{Mandy Simons, David Beaver, Craige Roberts, and Judith Tonhauser.} The best question: Explaining the projection behavior of factives.
-\textit{Discourse processes}, 54(3):187–206, 2017.
-\quad \textbullet \quad
+\textit{Discourse processes}, 54(3):187–206, 2017.\quad\textbullet\quad
 				%
@@
 \textit{In Proceedings of the 48th Meeting of the Chicago Linguistics Society}.
-2014.
-\quad \textbullet \quad
+2014.\quad\textbullet\quad
 				%
-				\textbf{Peter F. Strawson.} On referring. \textit{Mind}, 59(235):320–344, 1950. \quad \textbullet \quad
+				\textbf{Peter F. Strawson.} On referring. \textit{Mind}, 59(235):320–344, 1950.\quad\textbullet\quad
 				%
-				\textbf{Judith Tonhauser.} Prosodic cues to presupposition projection. In \textit{Semantics and Linguistic Theory}, volume 26, pages 934–960, 2016. \quad \textbullet \quad
+				\textbf{Judith Tonhauser.} Prosodic cues to presupposition projection. In \textit{Semantics and Linguistic Theory}, volume 26, pages 934–960, 2016.\quad\textbullet\quad
 				%
@@
 Gradience in projectivity and at-issueness.
-\textit{Journal of Semantics}, 35(3):495–542, 2018.
-\quad \textbullet \quad
+\textit{Journal of Semantics}, 35(3):495–542, 2018.\quad\textbullet\quad
 				%
@@
 
+			\vspace{20\baselineskip}
 			\begin{normalbox}{References (not visible later)}

### 2023-09-18 talks/XPrag-poster/xprag-poster.tex
abbreviated references

@@
 				\centering
-				\begin{theorybox}{Dynamic accounts of projection: Lexical triggering + dynamic semantics \hfill (\citealt{heim_projection_1983,van_der_sandt_presupposition_1992})}
+				\begin{theorybox}{Dynamic accounts of projection: Lexical triggering + dynamic semantics \hfill (\citealt{heim_presupposition_1992,van_der_sandt_presupposition_1992})}
 					\small Distinguish factive and non-factive predicates:
@@
 
-				\textbf{Márta Abrusán}. Predicting the presuppositions of soft triggers. \textit{Linguistics and philosophy}, 34:491–535, 2011.\quad\textbullet\quad
+				\textbf{Márta Abrusán}. Predicting the presuppositions of soft triggers. \textit{L\&P}, 2011.\quad\textbullet\quad
 				%
@@
 				%
-				\textbf{Elizabeth Coppock and Lucas Champollion.} \textit{Invitation to formal semantics.} online publication, in progress, 2020.\quad\textbullet\quad
+				\textbf{Elizabeth Coppock and Lucas Champollion.} \textit{Invitation to formal semantics.} 2020.\quad\textbullet\quad
 				%
-				\textbf{Marie de Marneffe, Mandy Simons, and Judith Tonhauser.} The CommitmentBank: Investigating projection in naturally occurring discourse.
-\textit{Sinn und Bedeutung}, 23:107–124, 2019.\quad\textbullet\quad
+				\textbf{Marie de Marneffe, Mandy Simons, and Judith Tonhauser.} The CommitmentBank: Investigating projection in naturally occurring discourse. \textit{SuB}, 2019.\quad\textbullet\quad
 				%
@@
 The Pragmatic Complexity of Veridicality Assessment.
-\textit{Computational Linguistics}, 38(2):301–333, February 2012.\quad\textbullet\quad
+\textit{Comp.
+Ling.}, February 2012.\quad\textbullet\quad
 				%
-				\textbf{Judith Degen and Judith Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 5:59–70, 2021.\quad\textbullet\quad
+				\textbf{Judith Degen and Judith Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 2021.\quad\textbullet\quad
 				%
-				\textbf{Judith Degen and Judith Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 98(3):552–591, 2022.\quad\textbullet\quad
+				\textbf{Judith Degen and Judith Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 2022.\quad\textbullet\quad
 				%
-				\textbf{Kajsa Djärv and Hezekiah Akiva Bacovcin.} Prosodic effects on factive presupposition projection. \textit{Semantics and Linguistic Theory}, 27:116–133, 2017.\quad\textbullet\quad
+				\textbf{Kajsa Djärv and Hezekiah Akiva Bacovcin.} Prosodic effects on factive presupposition projection. \textit{SALT}, 2017.\quad\textbullet\quad
 				%
-				\textbf{Tom Roberts.} I can’t believe it’s not lexical: Deriving distributed veridicality.
-Semantics and Linguistic Theory, \textit{Semantics and Linguistic Theory}, 29:665–685, 2019.\quad\textbullet\quad
+				\textbf{Tom Roberts.} I can’t believe it’s not lexical: Deriving distributed veridicality. Semantics and Linguistic Theory, \textit{SALT}, 2019.\quad\textbullet\quad
 				%
-				\textbf{Gottlob Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, pages 25–50, 1892.\quad\textbullet\quad
+				\textbf{Gottlob Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, 1892.\quad\textbullet\quad
 				%
-				\textbf{Irene Heim.} On the projection problem for presuppositions. \textit{Formal semantics–the essential readings}, pages 249–260, 1983.\quad\textbullet\quad
+				\textbf{Irene Heim.} Presupposition projection and the semantics of attitude verbs. \textit{JoS},1992.\quad\textbullet\quad
 				%
-				\textbf{Lauri Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 4(1):55–69, 1971.\quad\textbullet\quad
+				\textbf{Lauri Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 1971.\quad\textbullet\quad
 				%
-				\textbf{Lauri Karttunen and Stanley Peters.} Conventional implicature.
-In Choon-Kyu Oh and David A.
-Dinneen, editors, \textit{Presuppositions, Syntax and Semantics} Vol.11, pages 1–56.
-New York: Academic Press, 1979.\quad\textbullet\quad
+				\textbf{Lauri Karttunen and Stanley Peters.} Conventional implicature. In Choon-Kyu Oh and David A. Dinneen, eds., \textit{Presuppositions, Syntax and Semantics}, 1979.\quad\textbullet\quad
 				%
-				\textbf{Paul Kiparsky and Carol Kiparsky.} Fact.
-In Manfred Bierwisch and Karl Erich Heidolph, editors, \textit{Progress in Linguistics}, pages 143–173.
-The Hague: Mouton, 1970.\quad\textbullet\quad
+				\textbf{Paul Kiparsky and Carol Kiparsky.} Fact. In Manfred Bierwisch and Karl Erich Heidolph, eds., \textit{Progress in Linguistics}, 1970.\quad\textbullet\quad
 				%
-				\textbf{Taylor Mahler.} The social component of projection behavior of clausal complements. \textit{Linguistic Society of America}, 5:777–791, 2020.\quad\textbullet\quad
+				\textbf{Taylor Mahler.} The social component of projection behavior of clausal complements. \textit{LSA}, 2020.\quad\textbullet\quad
 				%
-				\textbf{Philippe Schlenker.} Triggering presuppositions. \textit{Glossa: a journal of general linguistics}, 6(1), 2021.\quad\textbullet\quad
+				\textbf{Philippe Schlenker.} Triggering presuppositions. \textit{Glossa}, 2021.\quad\textbullet\quad
 				%
-				\textbf{Judith Sieker and Torgrim Solstad.} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study.
-\textit{In Proceedings of the 23rd Amsterdam Colloquium}, 2022.\quad\textbullet\quad
+				\textbf{Judith Sieker and Torgrim Solstad.} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study. \textit{AC 23}, 2022.\quad\textbullet\quad
 				%
 				\textbf{Mandy Simons, David Beaver, Craige Roberts, and Judith Tonhauser.} The best question: Explaining the projection behavior of factives.
-\textit{Discourse processes}, 54(3):187–206, 2017.\quad\textbullet\quad
+\textit{Discourse processes}, 2017.\quad\textbullet\quad
 				%
-				\textbf{E.
-Allyn Smith and Kathleen Currie Hall.} The relationship between projection and embedding environment.
-\textit{In Proceedings of the 48th Meeting of the Chicago Linguistics Society}.
-2014.\quad\textbullet\quad
+				\textbf{E. Allyn Smith and Kathleen Currie Hall.} The relationship between projection and embedding environment. \textit{CLS 48}. 2014.\quad\textbullet\quad
 				%
-				\textbf{Peter F. Strawson.} On referring. \textit{Mind}, 59(235):320–344, 1950.\quad\textbullet\quad
+				\textbf{Peter F. Strawson.} On referring. \textit{Mind}, 1950.\quad\textbullet\quad
 				%
-				\textbf{Judith Tonhauser.} Prosodic cues to presupposition projection. In \textit{Semantics and Linguistic Theory}, volume 26, pages 934–960, 2016.\quad\textbullet\quad
+				\textbf{Judith Tonhauser.} Prosodic cues to presupposition projection. In \textit{SALT}, 2016.\quad\textbullet\quad
 				%
-				\textbf{Judith Tonhauser, David I.
-Beaver, and Judith Degen.} How projective is projective content?
-Gradience in projectivity and at-issueness.
-\textit{Journal of Semantics}, 35(3):495–542, 2018.\quad\textbullet\quad
+				\textbf{Judith Tonhauser, David I. Beaver, and Judith Degen.} How projective is projective content? Gradience in projectivity and at-issueness. \textit{JoS}, 2018.\quad\textbullet\quad
 				%
-				\textbf{Rob A. van der Sandt.} Presupposition projection as anaphora resolution. \textit{Journal of semantics}, 9(4):333–377, 1992.
+				\textbf{Rob A. van der Sandt.} Presupposition projection as anaphora resolution. \textit{JoS}, 1992.
 

### 2023-09-18 talks/XPrag-poster/xprag-poster.tex
abbreviated references further (names)

@@
 
-				\textbf{Márta Abrusán}. Predicting the presuppositions of soft triggers. \textit{L\&P}, 2011.\quad\textbullet\quad
+				\textbf{M. Abrusán}. Predicting the presuppositions of soft triggers. \textit{L\&P}, 2011.\quad\textbullet\quad
 				%
-				\textbf{Gennaro Chierchia and Sally McConnell-Ginet.} \textit{Meaning and grammar: An introduction to semantics.} 1990.\quad\textbullet\quad
+				\textbf{G. Chierchia and S. McConnell-Ginet.} \textit{Meaning and grammar: An introduction to semantics.} 1990.\quad\textbullet\quad
 				%
-				\textbf{Elizabeth Coppock and Lucas Champollion.} \textit{Invitation to formal semantics.} 2020.\quad\textbullet\quad
+				\textbf{E. Coppock and L. Champollion.} \textit{Invitation to formal semantics.} 2020.\quad\textbullet\quad
 				%
-				\textbf{Marie de Marneffe, Mandy Simons, and Judith Tonhauser.} The CommitmentBank: Investigating projection in naturally occurring discourse. \textit{SuB}, 2019.\quad\textbullet\quad
+				\textbf{M.-C. de Marneffe, M. Simons, and J. Tonhauser.} The CommitmentBank: Investigating projection in naturally occurring discourse. \textit{SuB}, 2019.\quad\textbullet\quad
 				%
-				\textbf{Marie-Catherine de Marneffe, Christopher D.
-Manning, and Christopher Potts.} Did It Happen?
-The Pragmatic Complexity of Veridicality Assessment.
-\textit{Comp.
-Ling.}, February 2012.\quad\textbullet\quad
+				\textbf{M.-C. de Marneffe, C.D. Manning, and C. Potts.} Did It Happen? The Pragmatic Complexity of Veridicality Assessment. \textit{Comp. Ling.}, February 2012.\quad\textbullet\quad
 				%
-				\textbf{Judith Degen and Judith Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 2021.\quad\textbullet\quad
+				\textbf{J. Degen and J. Tonhauser.} Prior beliefs modulate projection. \textit{Open Mind}, 2021.\quad\textbullet\quad
 				%
-				\textbf{Judith Degen and Judith Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 2022.\quad\textbullet\quad
+				\textbf{J. Degen and J. Tonhauser.} Are there factive predicates? An empirical investigation. \textit{Language}, 2022.\quad\textbullet\quad
 				%
-				\textbf{Kajsa Djärv and Hezekiah Akiva Bacovcin.} Prosodic effects on factive presupposition projection. \textit{SALT}, 2017.\quad\textbullet\quad
+				\textbf{K. Djärv and H.A. Bacovcin.} Prosodic effects on factive presupposition projection. \textit{SALT}, 2017.\quad\textbullet\quad
 				%
-				\textbf{Tom Roberts.} I can’t believe it’s not lexical: Deriving distributed veridicality. Semantics and Linguistic Theory, \textit{SALT}, 2019.\quad\textbullet\quad
+				\textbf{T. Roberts.} I can’t believe it’s not lexical: Deriving distributed veridicality. Semantics and Linguistic Theory, \textit{SALT}, 2019.\quad\textbullet\quad
 				%
-				\textbf{Gottlob Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, 1892.\quad\textbullet\quad
+				\textbf{G. Frege.} Über Sinn und Bedeutung. \textit{Zeitschrift für Philosophie und philosophische Kritik}, 1892.\quad\textbullet\quad
 				%
-				\textbf{Irene Heim.} Presupposition projection and the semantics of attitude verbs. \textit{JoS},1992.\quad\textbullet\quad
+				\textbf{I. Heim.} Presupposition projection and the semantics of attitude verbs. \textit{JoS},1992.\quad\textbullet\quad
 				%
-				\textbf{Lauri Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 1971.\quad\textbullet\quad
+				\textbf{L. Karttunen.} Some observations on factivity. \textit{Research on Language \& Social Interaction}, 1971.\quad\textbullet\quad
 				%
-				\textbf{Lauri Karttunen and Stanley Peters.} Conventional implicature. In Choon-Kyu Oh and David A. Dinneen, eds., \textit{Presuppositions, Syntax and Semantics}, 1979.\quad\textbullet\quad
+				\textbf{L. Karttunen and S. Peters.} Conventional implicature. In Choon-Kyu Oh and David A. Dinneen, eds., \textit{Presuppositions, Syntax and Semantics}, 1979.\quad\textbullet\quad
 				%
-				\textbf{Paul Kiparsky and Carol Kiparsky.} Fact. In Manfred Bierwisch and Karl Erich Heidolph, eds., \textit{Progress in Linguistics}, 1970.\quad\textbullet\quad
+				\textbf{P. Kiparsky and C. Kiparsky.} Fact. In Manfred Bierwisch and Karl Erich Heidolph, eds., \textit{Progress in Linguistics}, 1970.\quad\textbullet\quad
 				%
-				\textbf{Taylor Mahler.} The social component of projection behavior of clausal complements. \textit{LSA}, 2020.\quad\textbullet\quad
+				\textbf{T. Mahler.} The social component of projection behavior of clausal complements. \textit{LSA}, 2020.\quad\textbullet\quad
 				%
-				\textbf{Philippe Schlenker.} Triggering presuppositions. \textit{Glossa}, 2021.\quad\textbullet\quad
+				\textbf{P. Schlenker.} Triggering presuppositions. \textit{Glossa}, 2021.\quad\textbullet\quad
 				%
-				\textbf{Judith Sieker and Torgrim Solstad.} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study. \textit{AC 23}, 2022.\quad\textbullet\quad
+				\textbf{J. Sieker and T. Solstad.} Projective variability of (semi) factive verbs in family of sentence contexts: A rating study. \textit{AC 23}, 2022.\quad\textbullet\quad
 				%
-				\textbf{Mandy Simons, David Beaver, Craige Roberts, and Judith Tonhauser.} The best question: Explaining the projection behavior of factives.
-\textit{Discourse processes}, 2017.\quad\textbullet\quad
+				\textbf{M. Simons, D. Beaver, C. Roberts, and J. Tonhauser.} The best question: Explaining the projection behavior of factives. \textit{Discourse processes}, 2017.\quad\textbullet\quad
 				%
-				\textbf{E. Allyn Smith and Kathleen Currie Hall.} The relationship between projection and embedding environment. \textit{CLS 48}. 2014.\quad\textbullet\quad
+				\textbf{E.A. Smith and K.C. Hall.} The relationship between projection and embedding environment. \textit{CLS 48}. 2014.\quad\textbullet\quad
 				%
-				\textbf{Peter F. Strawson.} On referring. \textit{Mind}, 1950.\quad\textbullet\quad
+				\textbf{P.F. Strawson.} On referring. \textit{Mind}, 1950.\quad\textbullet\quad
 				%
-				\textbf{Judith Tonhauser.} Prosodic cues to presupposition projection. In \textit{SALT}, 2016.\quad\textbullet\quad
+				\textbf{J. Tonhauser.} Prosodic cues to presupposition projection. In \textit{SALT}, 2016.\quad\textbullet\quad
 				%
-				\textbf{Judith Tonhauser, David I. Beaver, and Judith Degen.} How projective is projective content? Gradience in projectivity and at-issueness. \textit{JoS}, 2018.\quad\textbullet\quad
+				\textbf{J. Tonhauser, D.I. Beaver, and J. Degen.} How projective is projective content? Gradience in projectivity and at-issueness. \textit{JoS}, 2018.\quad\textbullet\quad
 				%
-				\textbf{Rob A. van der Sandt.} Presupposition projection as anaphora resolution. \textit{JoS}, 1992.
+				\textbf{R.A. v.d.Sandt.} Presupposition projection as anaphora resolution. \textit{JoS}, 1992.
 

### 2023-09-18 talks/XPrag-poster/xprag-poster.tex
poster print

@@
-\documentclass[final, table]{beamer}
+\documentclass[final, table, cmyk]{beamer}
 %% Possible paper sizes: a0, a0b, a1, a2, a3, a4.
@@
 
-\institute[shortinst]{\inst{1} University of Stuttgart \samelineand \inst{2} UC Lovain}
+\institute[shortinst]{\inst{1} University of Stuttgart \samelineand \inst{2} UC Louvain}
 
@@
 
-					\item[\color{highlight}{$\blacktriangleright$}] Current theories of \cchighlight{projective content} do not predict our results
+					\item[\color{highlight}{$\blacktriangleright$}] Current theories of projective content do not predict our results
 				\end{itemize}
@@
 			\begin{normalbox}{Projection of clausal complements}
-				Do you infer that Rachel is committed to the truth of the \textit{content of the complement} (CC), that \cchighlight{Julian dances salsa}?
+				Do you infer that Rachel is committed to the truth of the \textit{content of the complement} (CC), that \underline{\em Julian dances salsa}?
 
 				\vspace{-.3\baselineskip}
-				\ex. \a. Rachel: \emph{\lq Does Cole \predhighlight{know} that \cchighlight{Julian dances salsa}?\rq}\\
+				\ex. \a. Rachel: \emph{\lq Does Cole \predhighlight{know} that \underline{Julian dances salsa}?\rq}\\
 					\cmark\ \  Yes, CC projects out of the question
-					\b. Rachel: \emph{\lq Does Cole \predhighlight{think} that \cchighlight{Julian dances salsa}?\rq}\\
+					\b. Rachel: \emph{\lq Does Cole \predhighlight{think} that \underline{Julian dances salsa}?\rq}\\
 					\xmark\ \ No, CC does not project
@@
 
-				\vspace{-.5\baselineskip}
+				\vspace{-.7\baselineskip}
 				\resizebox{\linewidth}{!}{
@@
 
-				\vspace{-.5\baselineskip}
-				\ex. \a. Polar question:\\
-					\textit{\ophighlight{Does} Cole \predhighlight{know} that \cchighlight{Julian dances salsa}?}
-					\b. Negation:\\
-					\textit{Cole \ophighlight{doesn't} \predhighlight{know} that \cchighlight{Julian dances salsa}.}
-					\b. Epistemic modal:\\
-					\textit{\ophighlight{Perhaps} Cole \predhighlight{knows} that \cchighlight{Julian dances salsa}.}
-					\b. Conditional antecedents:\\
-					\textit{\ophighlight{If} Cole \predhighlight{knows} that \cchighlight{Julian dances salsa}, Logan will be joyful.}
-					\z.
+				% \vspace{-.5\baselineskip}
+				\ex. \phantom.\vspace{-2.8\baselineskip}\newline
+					\begin{questionbox}{i. Polar question:}
+						\hfill \textit{\ophighlight{Does} Cole know that Julian dances salsa?}
+					\end{questionbox}
+					\vspace{-.8\baselineskip}
+					\begin{negbox}{ii. Negation:} \hfill \textit{Cole \ophighlight{doesn't} know that Julian dances salsa.}
+					\end{negbox}
+					\vspace{-.8\baselineskip}
+					\begin{modalbox}{iii. Epistemic modal:}
+						\hfill \textit{\ophighlight{Perhaps} Cole knows that Julian dances salsa.}
+					\end{modalbox}
+					\vspace{-.8\baselineskip}
+					\begin{condbox}{iv. Conditional antecedents:}\newline
+						\phantom.\hfill\textit{\ophighlight{If} Cole knows that Julian dances salsa, Logan will be joyful.}
+					\end{condbox}
 				\z.
@@
 			\begin{normalbox}{Hints at by-operator variation}
-				\textbf{Factive vs. semi-factive predicates} (\citealt{karttunen_observations_1971})
-				\vspace{-.2\baselineskip}
+				% \hspace{-.02\linewidth}
+					\begin{minipage}{1.08\linewidth}
 						\begin{itemize}
-					\item Factives (\textit{be annoyed, regret, \dots}): CC projects across all four operators
-					\item Semi-factives (\textit{discover, realize, see, notice, \dots}):\newline CC projects across negation, but not always for the other operators
+							\item \citet{karttunen_observations_1971} proposes \textbf{factive vs semi-factive} distinction
+							\item Smith \& Hall (\citeyear{smith_relationship_2014}): Experiment with \textbf{English projective contents}
+							\begin{itemize}
+								\item Effect of operator differs by projective content
 							\end{itemize}
-
-				\textbf{Experiment with English projective contents} (\citealt{smith_relationship_2014})
-				\vspace{-.2\baselineskip}
+							\item Sieker \& Solstad (\citeyear{sieker_projective_2022}): Exp. with \textbf{German clause-embedding predicates}
 							\begin{itemize}
-					\item Projective content of epithets (e.g. \textit{idiot}) and the CC of \textit{know}:\\ more projective under negation than conditionals
-					\item Opposite pattern for appositive relative clauses and \textit{win}
-				\end{itemize}
-
-				\textbf{Experiment with German clause-embedding predicates}\\ (\citealt{sieker_projective_2022})
-				\vspace{-.2\baselineskip}
-				\begin{itemize}
-					\item Higher projection ratings w/ negation than other three operators
 								\item No by-predicate variation, no evidence for factive/semi-factive distinction
 							\end{itemize}
+						\end{itemize}
+					\end{minipage}
+
+				% \smallskip
+
+				\begin{center}
+					\scalebox{.85}{
+					\begin{tabular}{l l c c c c}
+					 	&& \nhl{\textcolor{neg}{\bf Neg}}  & \chl{\textcolor{cond}{\bf Cond}} &  \qhl{\textcolor{question}{\bf PQ}} & \mhl{\textcolor{modal}{\bf Mod}}  \\ \toprule
+
+					 	&factives \\
+					 	& \small (\textit{be annoyed, regret, \dots}) & 
+					 		\multirow{-2}{*}{\cmark}	&
+					 		\multirow{-2}{*}{\cmark}	&
+					 		\multirow{-2}{*}{\cmark}	&
+					 		\multirow{-2}{*}{\cmark}	\\
+
+					 	 & semi-factives\\
+					 	\multirow{-4}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-\baselineskip}\phantom.\newline \citet{karttunen_observations_1971}}} &
+					 	\small (\textit{discover, realize, see, notice, \dots}) & 
+					 		\multirow{-2}{*}{\cmark}	&
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-\baselineskip}\phantom.\newline not\\ always}}	&
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-\baselineskip}\phantom.\newline not\\ always}}	&
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-\baselineskip}\phantom.\newline not\\ always}}	\\ \midrule
+
+						& epithets (e.g. \textit{idiot}),
+							&&& \cellcolor{gray!20} & \cellcolor{gray!20} \\
+						& CC of \textit{know}&
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf more}\\ projection}} &
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf less}\\ projection}} &
+					 		\cellcolor{gray!20}
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline N/A}} &
+					 		\cellcolor{gray!20}
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline N/A}} \\
+
+					 	& Appositive RCs,
+							&&& \cellcolor{gray!20} & \cellcolor{gray!20} \\
+
+						\multirow{-4}{*}{\parbox{3cm}{\footnotesize\centering \vspace{-\baselineskip}\phantom.\newline Smith \& Hall (\citeyear{smith_relationship_2014})}} &
+
+						prep. content of \textit{win} &
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf less}\\ projection}} &
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf more}\\ projection}} &
+					 		\cellcolor{gray!20}
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline N/A}} &
+					 		\cellcolor{gray!20}
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline N/A}} \\ \midrule
+
+					 	&CCs of German \lq factives\rq
+							&&&&\\
+
+						\multirow{-2}{*}{\parbox{3cm}{\footnotesize\centering \vspace{-\baselineskip}\phantom.\newline Sieker \& Solstad (\citeyear{sieker_projective_2022})}} &
+
+						\& \lq semi-factives\rq&
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf more}\\ projection}} &
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf less}\\ projection}} &
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf less}\\ projection}} &
+					 		\multirow{-2}{*}{\parbox{2.5cm}{\footnotesize\centering \vspace{-.5\baselineskip}\phantom.\newline {\footnotesize\bf less}\\ projection}} \\ \bottomrule
+					\end{tabular}}
+				\end{center}
+
+				% \vspace{-\baselineskip}
 
@@
 					\vspace{-.4\baselineskip}
-					\caption{\citealt{sieker_projective_2022}, p. 286: Projection-ratings by embedding operator, for purported factive and semi-factive predicates}
+					\caption{Sieker \& Solstad \citeyear{sieker_projective_2022}, p. 286: Projection-ratings by embedding operator, for purported factive and semi-factive predicates}
 					% \label{fig:figure1}
@@
 			\vspace{\vboxsep}
-			\begin{normalbox}{Certain-that task for projection inferences}
+			\begin{normalbox}{\lq Certain-that\rq\ task for projection inferences}
 
@@
 					\includegraphics[width=.95\linewidth]{degen-tonhauser.png}
+					\vspace{-.8\baselineskip}
 					\caption{\citealt{degen_are_2022}, p. 562: Mean certainty ratings by predicate}
@@
 			\begin{normalbox}{Materials}
-				To assess the effect of \ophighlight{operator} and \predhighlight{predicate} on \cchighlight{\textbf{projection}}:
+				To assess the effect of \ophighlight{operator} and \predhighlight{predicate} on \textbf{projection}:
 				\vspace{.2\baselineskip}
@@
 					\begin{itemize}
-						\item One per \ophighlight{operator}: polar questions, negation, modal \textit{perhaps}, conditional
+						\item One per operator: \textcolor{question}{\bf polar questions}, \textcolor{neg}{\bf negation}, \textcolor{modal}{\bf modal \textit{perhaps}}, \textcolor{cond}{\bf conditional}
 					\end{itemize}
@@
 					\begin{itemize} \small
-						\item Conditional > Question > Negation, Modal
+						\item {\bf \textcolor{cond}{\bf Conditional} > \textcolor{question}{\bf Question} > \textcolor{neg}{\bf Negation}, \textcolor{modal}{\bf Modal}}
 							\vspace{\mboxpreadjust}
@@
 								\footnotesize
-								response: \textbf{certainty ratings}; fixed effect: \ophighlight{\textbf{operator}} (base level: Question);
+								response: \textbf{certainty ratings}; fixed effect: \ophighlight{operator} (base level: Question);
 								random intercepts: participants, items; \newline
@@
 				\begin{itemize}\small
-					\item CC of \predhighlight{be annoyed}: Negation, Conditional > Question, Modal
+					\item CC of \predhighlight{be annoyed}: {\bf \textcolor{neg}{\bf Negation}, \textcolor{cond}{\bf Conditional} > \textcolor{question}{\bf Question}, \textcolor{modal}{\bf Modal}}
 						\vspace{\mboxpreadjust}
@@
 
-					\item CC of \predhighlight{know}: Question > Negation, Conditional > Modal
+					\item CC of \predhighlight{know}: {\bf \textcolor{question}{\bf Question} > \textcolor{neg}{\bf Negation}, \textcolor{cond}{\bf Conditional} > \textcolor{modal}{\bf Modal}}
 						\vspace{\mboxpreadjust}
@@
 
-					\item CC of \predhighlight{discover}: Modal > Negation > Conditional, Question
+					\item CC of \predhighlight{discover}: {\bf \textcolor{modal}{\bf Modal} > \textcolor{neg}{\bf Negation} > \textcolor{cond}{\bf Conditional}, \textcolor{question}{\bf Question}}
 						\vspace{\mboxpreadjust}
@@
 							    % \draw[red,ultra thick,rounded corners] (7.5,5.3) rectangle (9.4,6.2);
-							    \draw[red, ultra thick] (-3.37,2.3) -- (-3.37,1) -- node[below] {***} (-.66,1) -- (-.66,1.5);
-							    \draw[red, ultra thick] (-.66,1.5) -- (-.66,.2) -- node[below] {***} (2.07,.2) -- (2.07,.7);
-							    \draw[red, ultra thick] (2.07,.8) -- (2.07,.12) -- node[below] {\footnotesize n.s.} (4.78,.12) -- (4.78,.62);
+							    \draw[highlight, ultra thick] (-3.37,2.3) -- (-3.37,1) -- node[below] {***} (-.66,1) -- (-.66,1.5);
+							    \draw[highlight, ultra thick] (-.66,1.5) -- (-.66,.2) -- node[below] {***} (2.07,.2) -- (2.07,.7);
+							    \draw[highlight, ultra thick] (2.07,.8) -- (2.07,.12) -- node[below] {\footnotesize n.s.} (4.78,.12) -- (4.78,.62);
 							\end{tikzpicture}}
@@
 				\end{tabular}
-				\vspace{-1.6\baselineskip}
+				\vspace{-1.24\baselineskip}
 				\begin{figure}[h]
@@
 					\includegraphics[width=\linewidth]{projective-pred-op.pdf}
-					\vspace{-2.2\baselineskip}
-					\caption{Mean certainty ratings by predicate, grouped by operator}
+					\vspace{-1.7\baselineskip}
+					\caption{Mean certainty ratings by predicate with 95\% bootstrapped confidence intervals, grouped by operator.
+Entailment-cancelling operator coded by letter and color:  \textcolor{neg}{\texttt{N}: negation}, \textcolor{modal}{\texttt{M}: modals}, \textcolor{cond}{\texttt{C}: conditional antecedents}, \textcolor{question}{\texttt{Q}: polar questions}.
+					}
 					\label{fig:figure3}
@@
 				\begin{itemize}
-					\item Concurs with \citet{smith_relationship_2014}, who found content/operator interactions for English projective contents
-					\item Differs from \citet{sieker_projective_2022}: found no predicate/operator interaction for CCs of German clause-embedding predicates
+					\item Concurs with Smith \& Hall (\citeyear{smith_relationship_2014}), who found content/operator interactions for English projective contents
+					\item Differs from Sieker \& Solstad (\citeyear{sieker_projective_2022}): found no predicate/operator interaction for CCs of German clause-embedding predicates
 				\end{itemize}
@@
 
+				\smallskip
 				\resizebox{\linewidth}{!}{
 						\begin{tabular}{l c c c}
-							\textbf{Our data} 
 								& \multicolumn{3}{c}{\textbf{Predictions}}\\ \toprule
 							%
+							\textbf{Our data} 
 								& \parbox{.3\linewidth}{\centering Lexical triggering + \par dynamic semantics}
@@
 							\parbox{.3\linewidth}{\raggedright\textbf{Superadditive predicate/operator interaction}\\[2.5\baselineskip]}
-								& \parbox{.29\linewidth}{\centering \cellcolor{red!25} Meaning of entailment-canceling operators (invariably) encodes interaction with conventional content of embedded factives}
-									& \parbox{.29\linewidth}{\centering \cellcolor{red!25} No systematic predictions for how veridicality or at-issueness interact with the meaning of entailment-canceling operators}
-										& \parbox{.29\linewidth}{\centering\cellcolor{green!25} May be extended to our data by making explicit how combinations of operator + predicate are associated with contextual inferences}\\
+								& \parbox{.29\linewidth}{\centering \cellcolor{red!15} Meaning of entailment-canceling operators (invariably) encodes interaction with conventional content of embedded factives}
+									& \parbox{.29\linewidth}{\centering \cellcolor{red!15} No systematic predictions for how veridicality or at-issueness interact with the meaning of entailment-canceling operators}
+										& \parbox{.29\linewidth}{\centering\cellcolor{green!15} May be extended to our data by making explicit how combinations of operator + predicate are associated with contextual inferences}\\
 							%
 							\parbox{.3\linewidth}{\raggedright\textbf{Projection variability for all predicates}\par {\small Projection for some non-factive (/non-veridical) predicates as high as for some factive (/veridical) ones (see also D\&T'22)}\\[.5\baselineskip]}
-								& \parbox{.29\linewidth}{\centering \cellcolor{red!25} No predictions for non-factive predicates}
-									& \parbox{.29\linewidth}{\centering \cellcolor{red!25} No systematic predictions for non-veridical predicates}
-										& \parbox{.29\linewidth}{\centering \cellcolor{green!25} Makes predictions about CCs of all clause-embedding predicates} \\
+								& \parbox{.29\linewidth}{\centering \cellcolor{red!15} No predictions for non-factive predicates}
+									& \parbox{.29\linewidth}{\centering \cellcolor{red!15} No systematic predictions for non-veridical predicates}
+										& \parbox{.29\linewidth}{\centering \cellcolor{green!15} Makes predictions about CCs of all clause-embedding predicates} \\
 							%
 							\parbox{.3\linewidth}{\raggedright\textbf{Out-of-the-blue contexts}\par {\small Projection variability in the out-of-the-blue contexts used in experiment (see also D\&T'22)}\\[1.7\baselineskip]}
-								& \parbox{.29\linewidth}{\centering \cellcolor{red!25} Consistent projection of factive CCs}
-									& \parbox{.3\linewidth}{\centering \cellcolor{green!25} Veridical predicates: analyses may be extended by assuming that the CCs of veridical predicates differ in at-issueness in out-of-the-blue contexts}
-										& \parbox{.29\linewidth}{\centering \cellcolor{red!25} “Out-of-the-blue” contexts do not warrant assumption of contextual \underline{entailment}: No projection expected} \\
+								& \parbox{.29\linewidth}{\centering \cellcolor{red!15} Consistent projection of factive CCs}
+									& \parbox{.3\linewidth}{\centering \cellcolor{green!15} Veridical predicates: analyses may be extended by assuming that the CCs of veridical predicates differ in at-issueness in out-of-the-blue contexts}
+										& \parbox{.29\linewidth}{\centering \cellcolor{red!15} “Out-of-the-blue” contexts do not warrant assumption of contextual \underline{entailment}: No projection expected} \\
 						\end{tabular}

### 2023-09-18 results/main/meta-analyses/new/1_projectivity/rscripts/graphs.R

@@ -117,8 +117,8 @@
       mutate(op = fct_reorder(op, projective, .fun = mean)) %>% ungroup()
   
     # plot ----
-    pomeans %>% 
-      ggplot(aes(x=fct_reorder(verb, projective), y=projective, group = op, color = op)) +
+    pmeans %>% 
+      ggplot(aes(x=fct_reorder(verb, Mean), y=Mean, group = op, color = op)) +
       geom_point(aes(shape = op), size = 4) + 
       geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
       geom_line() +
@@ -159,7 +159,7 @@
       theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))
     ggsave(f="../graphs/projectivity-pred-op-participant.pdf",height=6,width=8)
     
-  ## 4c projectivity means with distributions for paricipant means ----
+  ## 4c projectivity means with distributions for participant means ----
     # plot
     ggplot(pomeans, aes(x=fct_reorder(verb, projective), y=projective)) +
       data %>% mutate(verb = fct_reorder(verb, projective, .fun = mean)) %>%

### 2023-09-21 talks/XPrag-poster/xprag-poster.tex
poster print

@@
 					\begin{itemize} \small
-						\item {\bf \textcolor{cond}{\bf Conditional} > \textcolor{question}{\bf Question} > \textcolor{neg}{\bf Negation}, \textcolor{modal}{\bf Modal}}
+						\item \chl{\textcolor{cond}{\bf Conditional}} > \qhl{\textcolor{question}{\bf Question}} > \nhl{\textcolor{neg}{\bf Negation}}, \mhl{\textcolor{modal}{\bf Modal}}
 							\vspace{\mboxpreadjust}
@@
 				\begin{itemize}\small
-					\item CC of \predhighlight{be annoyed}: {\bf \textcolor{neg}{\bf Negation}, \textcolor{cond}{\bf Conditional} > \textcolor{question}{\bf Question}, \textcolor{modal}{\bf Modal}}
+					\item CC of \predhighlight{be annoyed}: \qhl{\textcolor{question}{\bf Question}}, \nhl{\textcolor{neg}{\bf Negation}} > \chl{\textcolor{cond}{\bf Conditional}}, \mhl{\textcolor{modal}{\bf Modal}}
 						\vspace{\mboxpreadjust}
@@
 
-					\item CC of \predhighlight{know}: {\bf \textcolor{question}{\bf Question} > \textcolor{neg}{\bf Negation}, \textcolor{cond}{\bf Conditional} > \textcolor{modal}{\bf Modal}}
+					\item CC of \predhighlight{know}: \qhl{\textcolor{question}{\bf Question}} > \nhl{\textcolor{neg}{\bf Negation}}, \chl{\textcolor{cond}{\bf Conditional}} > \mhl{\textcolor{modal}{\bf Modal}}
 						\vspace{\mboxpreadjust}
@@
 
-					\item CC of \predhighlight{discover}: {\bf \textcolor{modal}{\bf Modal} > \textcolor{neg}{\bf Negation} > \textcolor{cond}{\bf Conditional}, \textcolor{question}{\bf Question}}
+					\item CC of \predhighlight{discover}: \chl{\textcolor{cond}{\bf Conditional}}, \qhl{\textcolor{question}{\bf Question}} > \nhl{\textcolor{neg}{\bf Negation}} > \mhl{\textcolor{modal}{\bf Modal}}
 						\vspace{\mboxpreadjust}
@@
 							response: \textbf{certainty ratings}; fixed effects: \ophighlight{operator}, \predhighlight{predicate}, and interaction (base level: \textbf{discover} / negation); random intercepts: participant; \newline
-							MLEs: negation (intercept) $0.68$, conditional $+0.11$, modal $-0.06$, question $+0.10$; with $p < 0.001$
+							MLEs: negation (intercept) $0.68$, conditional $+0.11$, question $+0.10$, modal $-0.06$; with $p < 0.001$
 						\end{modelbox}

### 2023-09-21 results/main/meta-analyses/new/1_projectivity/rscripts/graphs.R

@@ -63,20 +63,20 @@
   
   # means and confidence intervals for projectivity rating by predicate ----
   pmeans = data %>% group_by(verb) %>%
-    summarize(projective = mean(projective), CILow = ci.low(projective), 
+    summarize(Mean = mean(projective), CILow = ci.low(projective), 
               CIHigh = ci.high(projective)) %>%
-    mutate(YMin = projective - CILow, YMax = projective + CIHigh, 
-           verb = fct_reorder(as.factor(verb), projective)) %>% ungroup()
+    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
+           verb = fct_reorder(as.factor(verb), Mean), projective = Mean) %>% ungroup()
   
   ## 3a projectivity by predicate w distribution of participants' means ----
     # get the mean of participants' projectivity ratings by verb
     subjmeans = data %>%
       group_by(verb,workerid) %>%
-      summarize(projective = mean(projective))
+      summarize(Mean = mean(projective))
     subjmeans$verb <- factor(subjmeans$verb, levels = unique(levels(pmeans$verb)))
   
     # plot 
-    ggplot(pmeans, aes(x=verb, y=projective)) +
+    ggplot(pmeans, aes(x=verb, y=Mean)) +
       geom_violin(data=subjmeans,scale="width",color="gray80") +
       geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.3,color="black") +
       geom_point(size=0.5,color="black") +
@@ -109,15 +109,15 @@
   ## 4a means compare between operators ----
     # projectivity by verb and operator
     pomeans = data %>% group_by(verb, op) %>%
-      summarize(projective = mean(projective), CILow = ci.low(projective), 
+      summarize(Mean = mean(projective), CILow = ci.low(projective), 
                 CIHigh = ci.high(projective)) %>%
-      mutate(YMin = projective - CILow, YMax = projective + CIHigh, 
-             verb = fct_reorder(as.factor(verb),projective)) %>%
-      mutate(verb = fct_reorder(verb, projective, .fun = mean)) %>%
-      mutate(op = fct_reorder(op, projective, .fun = mean)) %>% ungroup()
+      mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
+             verb = fct_reorder(as.factor(verb),Mean)) %>%
[…]

### 2023-11-05 writing/naturalness-paper/paper.tex
updated L-and-P paper and scripts ls -l
q

q
:q

q

\documentclass[11pt,fleqn]{article}
\usepackage[margin=1in,top=1in,bottom=1in]{geometry}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
%\usepackage[dvips]{graphics}
%\usepackage[table]{xcolor}
%\usepackage{amssymb}
\usepackage{float}
%\usepackage{subfig}
\usepackage{booktabs}
\usepackage{subcaption}

\usepackage[normalem]{ulem}

\usepackage{multicol}
\usepackage{txfonts}
\usepackage{amsfonts}
\usepackage{natbib}
\usepackage{gb4e}
\usepackage[all]{xy}
\usepackage{rotating}
\usepackage{tipa}
\usepackage{multirow}
\usepackage{authblk}
\usepackage{url}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{adjustbox}
\usepackage{array}

\def\bad{{\leavevmode\llap{*}}}
\def\marginal{{\leavevmode\llap{?}}}
\def\verymarginal{{\leavevmode\llap{??}}}
\def\swmarginal{{\leavevmode\llap{4}}}
\def\infelic{{\leavevmode\llap{\#}}}

\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
%\definecolor{gray}{rgb}{0.36, 0.54, 0.66}

\definecolor{Pink}{RGB}{240,0,120}
\newcommand{\red}[1]{\textcolor{Pink}{#1}}
\newcommand{\jd}[1]{\textbf{\textcolor{Pink}{[jd: #1]}}}
\definecolor{green}{RGB}{0,158,115}
\definecolor{orange}{RGB}{213,94,0}

\newcommand{\dashrule}[1][black]{%
  \color{#1}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}\xleaders\hbox{\rule{4pt}{0pt}\rule[\dimexpr.5ex-.2pt]{4pt}{.4pt}}\hfill\kern0pt%
}

\setlength{\parindent}{.3in}
\setlength{\parskip}{0ex}

\newcommand{\yi}{\'{\symbol{16}}}
\newcommand{\nasi}{\~{\symbol{16}}}
\newcommand{\hina}{h\nasi na}
\newcommand{\ina}{\nasi na}

\newcommand{\foc}{$_{\mbox{\small F}}$}

\hyphenation{par-ti-ci-pa-tion}

\setlength{\bibhang}{0.5in}
\setlength{\bibsep}{0mm}
\bibpunct[:]{(}{)}{,}{a}{}{,}

\newcommand{\6}{\mbox{$[\hspace*{-.6mm}[$}} 
\newcommand{\9}{\mbox{$]\hspace*{-.6mm}]$}}
\newcommand{\sem}[2]{\6#1\9$^{#2}$}
\renewcommand{\ni}{\~{\i}}

\newcommand{\citepos}[1]{\citeauthor{#1}'s \citeyear{#1}}
\newcommand{\citeposs}[1]{\citeauthor{#1}'s}
\newcommand{\citetpos}[1]{\citeauthor{#1}'s (\citeyear{#1})}

\newcolumntype{R}[2]{%
    >{\adjustbox{angle=#1,lap=\width-(#2)}\bgroup}%
    l%
    <{\egroup}%
}
\newcommand*\rot{\multicolumn{1}{R{90}{0em}}}% no optional argument here, please!

\title{I don't know if projection is binary. \\ Did \citealt{mandelkern-etal2020} discover that it is?}

\author[$\circ$]{Judith Tonhauser}
\author[$\bullet$]{Judith Degen}
\affil[$\circ$]{University of Stuttgart}
\affil[$\bullet$]{Stanford University}

\renewcommand\Authands{ and }

\newcommand{\jt}[1]{\textbf{\color{blue}JT: #1}}

\begin{document}

\maketitle

\begin{abstract}

Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{frege1892,ccmg90}). While projection is often characterized as binary (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and proposed that naturalness ratings in explicit ignorance contexts are more suitable to measure binary projection and, thereby, to distinguish presupposition triggers from nontriggers. This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} can distinguish factive predicates (presumed presupposition triggers) from nonfactive ones (nontriggers). While the results do not support \citepos{mandelkern-etal2020} hypothesis and are rather in line with results based on inference rating measures (e.g., \citealt{degen-tonhauser-language}), the results suggest that naturalness ratings in explicit ignorance contexts are sensitive to felicity conditions of anaphoric projective content.

\end{abstract}

\bigskip

\noindent
{\bf Keywords:} Presuppositions, binary vs.\ gradient projection, inference ratings, naturalness ratings in explicit ignorance contexts.

\bigskip

\noindent
{\bf Acknowledgments:}

\noindent
For helpful comments on this project we thank Craige Roberts, Gregory Scontras, and Mandy Simons.

\newpage

\section{Introduction}\label{s1}

Presuppositions are defined as content that typically projects out of entailment-canceling environments, like the scope of negation or a polar interrogative (e.g., \citealt{frege1892,ccmg90}).\footnote{Conventional implicatures also project out of these environments. They differ from presuppositions in that they contribute novel information. For discussion see, e.g., \citealt{ccmg90,potts05}.} Thus, a content like the content of the clausal complement (CC) of {\em know} in (\ref{first}a) is taken to be a presupposition because a speaker who utters (\ref{first}a) is typically taken to believe the CC, that Julian dances salsa, even though the clausal complement occurs in a polar interrogative. By contrast, the CC of {\em think} in (\ref{first}b) is not assumed to be a presupposition because it does not typically project. Accordingly, {\em know} is analyzed as a presupposition trigger, but {\em think} is not.

\begin{exe}
\ex\label{first} 
\begin{xlist}
\ex Does Cole know that Julian dances salsa?
\ex Does Cole think that Julian dances salsa?
\end{xlist}
\end{exe}

To investigate whether a particular content is a presupposition, empirical research has long relied on inference rating measures. Such measures involve presenting participants with an utterance (like that in (\ref{first}a)) and contents (like the content that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content. Many theoretical works on projection assume that projection is a binary property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition). There is, however, mounting empirical evidence (based on investigations that used inference rating measures) that projection is not a binary property of  content but rather a gradient one, with  content being more or less projective. \citealt{karttunen71b}, for instance, already suggested that the contents of the clausal complements (CCs) of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}). Specifically, he suggested that a speaker who utters (\ref{kart}a) believes that the addressee has not told the truth, whereas a speaker who utters (\ref{kart}b) ``is not sure about the truth of the complement'' (p.63).

\begin{exe}
\ex\label{kart}
\begin{xlist}
\ex Did you regret that you had not told the truth?
\ex Did you discover that you had not told the truth? \hfill (\citealt[63]{karttunen71b})
\end{xlist}
\end{exe}
More recently, experimental investigations have also observed projection variability between presuppositions (e.g., \citealt{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}). \citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CCs of {\em know, discover,} and {\em reveal} exhibit projection variability: The CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}. These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''. Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}. In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a binary distinction in the projection of the CCs of factive and nonfactive predicates. This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content. A sample illustration of these results is given in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{figures/Language-Figure2}
\caption{Mean certainty rating (measuring projection) of main clause content and the CCs of the 20 predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category). Error bars indicate 95\% bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the individual participants' ratings.}\label{fig:dt1a}
\end{figure}

Recently,  \citealt[\S6.2]{mandelkern-etal2020} challenged the hypothesis that projection is gradient by arguing that inference rating measures are not suitable to detect binary projection. Such measures are not suitable, they argued, because they do not distinguish between, on the one hand, contents that project because they are presuppositions and, on the other hand, contents that project because they are ``a natural conclusion to draw for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497). Inference rating measures, they suggested, might invite participants to draw projection inferences even when not lexically required because ``the content in question is right there before the [participants'] eyes" (p.498). They concluded that ``we should think twice before embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).

\citealt[\S6.2]{mandelkern-etal2020} suggested that naturalness ratings of utterances in explicit ignorance contexts are better suited to distinguish presuppositions from nonpresuppositional projection inferences.\footnote{\citealt{mandelkern-etal2020} refer to the measure as an acceptability rating measure, but since participants were asked to rate the naturalness of utterances, we refer to it as a naturalness rating measure.} On this measure, participants are presented with utterances in which a speaker states that they are ignorant about a content, as in (\ref{mtrig}a), and then follows up with an utterance in which the content occurs embedded in an entailment-canceling environment, such as the antecedent of a conditional in (\ref{mtrig}c). \citealt{mandelkern-etal2020} assumed that if the content is a presupposition (such as the pre-state content of {\em stop} in (\ref{mtrig}c)), the utterance is rated as unnatural in the explicit ignorance context because presuppositions are contents that the speaker believes to be true. That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).  If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497).\footnote{Presuppositions won't be relinquished, modulo local accommodation} Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).

\begin{exe}
\ex\label{mtrig} \citealt[490f.]{mandelkern-etal2020}
\begin{xlist}
\ex Explicit ignorance context: \\ Mary always was involved in a lot of sports, but I don't know whether she ever did any yoga.
\ex Support context: \\ Mary always was involved in a lot of sports, and she used to do yoga, too.
\ex Sentence with presupposition / nonpresupposition: \\ If Mary \{has stopped / now frowns on\} doing yoga, then Matthew will interview her for his story.
\end{xlist}
\end{exe}
\citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, only provided limited empirical evidence for the hypothesis that projection is a binary property of content. This is primarily because the goal of this work (including Exp.~3) was not to investigate this hypothesis, but rather whether presupposition projection in conjunctions is symmetric. The evidence provided in this work is also limited, however, because only four presupposition triggers and nontriggers each were investigated: The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i). It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past.

\begin{exe}
\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
\end{exe}
} In the control items of Exp.~3, these eight expressions were realized in (single-clause) antecedents of conditionals, as shown in (\ref{mtrig}c). Naturalness ratings were elicited in explicit ignorance and support contexts.

\citealt{mandelkern-etal2020} found (as shown in their Fig.~3) that the naturalness ratings for the nontriggers in the control items were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale). The presupposition triggers, on the other hand, had lower mean naturalness ratings in the explicit ignorance context (mean around 3) than in the support context (mean just above 5). Mandelkern and his colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492). They proposed (p.497):

\begin{quote}

[\ldots] comparing contexts which support the inference to contexts in which it has been made clear that the speaker is ignorant about the inference provides a way to distinguish a broad class of natural and invited pragmatic inferences from those that are really encoded as presuppositions, and thus have no choice but to project. [...]  In methodological terms, we strongly recommend at least a two-pronged approach, with careful attention paid to results stemming from the evaluation of acceptabilty [sic] in different contexts.

\end{quote}

However, an inspection of the naturalness ratings for the eight (non)triggers provided in Fig.~8 in their Appendix 3 raises doubt about whether the measure indeed provides evidence for a binary distinction between presuppositions and nonpresuppositions. First, considering the mean naturalness ratings in the explicit ignorance contexts, we find ratings for the presuppositions range from just above 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}). As the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}), it is not clear whether naturalness ratings in explicit ignorance contexts provide evidence for a binary distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}. Second, the effect of context (explicit ignorance vs.\ support) also does not appear to be consistent for presuppositions or nonpresuppositions. Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}. In sum, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a binary projection either.

There also appears to be disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts. Whereas Mandelkern and his colleagues assumed that all presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only hard (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural. For instance,  {\em stop} and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic}c) with `??' instead of `$\#$'.} This disagreement in the literature necessitates further investigation.

\begin{exe}
\ex\label{eic2}
\begin{xlist}
\ex I have no idea whether Jane ever smoked, but she hasn't stopped smoking. \hfill (\citealt[443]{simons01})
\ex I have no idea whether John ended up participating in the
Road Race yesterday. But if he won it, then he has more victories than anyone else in history. \hfill (\citealt[39]{abusch10})
\end{xlist}
\ex\label{eic3}
\begin{xlist}
\ex\infelic I don't know if Jane ever rented ``Manhattan'' before, but perhaps she's renting it again. \hfill (\citealt[443]{simons01})

\ex \infelic I have no idea whether anyone read that letter. But if it is John
who read it, let's ask him to be discreet about the content. \hfill (\citealt[40]{abusch10})

\ex \infelic I have no idea whether John read that proposal. But if Bill read it too, let's ask them to confer and simply give us a yes-no response. \hfill (\citealt[40]{abusch10})
\end{xlist}
\end{exe}

This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} provides evidence that projection is a binary property of content rather than a gradient one.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated. The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft, to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.

One change we made to the design of \citepos{mandelkern-etal2020} Exp.~3 was that our experiment implemented a three-level context condition instead of a two-level one. Specifically, we replaced \citepos{mandelkern-etal2020} support context (which entailed the relevant context, see (\ref{mtrig}b)) with two contexts that were compatible with the relevant content but did not entail it. The two contexts differed in the prior probability of the relevant content: In the `lower prior probability' context, illustrated in (\ref{prior}a), the  relevant content (in (\ref{prior}c), that Julian dances salsa) had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context, illustrated in (\ref{prior}b). These two contexts were normed in \citealt{degen-tonhauser-openmind}.

\begin{exe}
\ex\label{prior}
\begin{xlist}
\ex Julian is German. \hfill [lower prior probability]
\ex Julian is Cuban. \hfill [higher prior probability]
\ex Did Cole discover that Julian dances salsa?
\end{xlist}
\end{exe}
This change was implemented to allow for comparison to the result of \citealt{degen-tonhauser-openmind}, namely that the projection of a content is sensitive to the content's prior probability. Specifically, this work found that the higher the prior probability of a content, the stronger the projection inference. Under standard analyses of presuppositions, like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in such contexts. Accordingly, such analyses predict that presuppositions are rated as natural in both lower and higher probability contexts, just as in \citepos{mandelkern-etal2020} support contexts.

%A challenge in the identification of presuppositions is that they only {\em typically} project: It has long been know that presuppositions need not project in any utterance in which the trigging expression is realized (e.g., \citealt{} REFERENCES). For example, in the naturally occurring example in (\ref{acc}) the presupposition of {\em know} (in the framed box) does not project but is (analyzed as) locally accommodated in the scope of the epistemic modal adverb {\em perhaps}. 
%
%
%\begin{exe}
%\ex\label{acc} Whatever the reason, she knows she can't just bounce back, and she decides to leave. For others, the decision is based purely on fear. Perhaps she \fbox{knows} that if she were to stay another day, her life or the lives of her children would be gone. \hfill (\citealt[85]{beaver-belly})
%\end{exe}
%
%
%In view of this potential for presuppositions to not project, investigations into the question of which utterance contents are presuppositions (and which ones are not) must

\section{Experiment}\label{s2}

To investigate whether the measure proposed in \citealt{mandelkern-etal2020} provides support for a binary distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive clause-embedding predicate), and rated the naturalness of the interrogative in the context of the declarative.

\subsection{Methods}\label{s-methods}

\subsubsection{Participants}

We recruited 425 participants on Prolific. Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-binary, 2 preferred to not disclose). The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%. The median time spent on the task was 6:24 minutes. Participants were paid \$1.78 (corresponding to an hourly pay of \$16.6).

\subsubsection{Materials}

Participants read two-sentence discourses consisting of a declarative followed by an interrogative, as shown in (\ref{sample}). In the target stimuli, the interrogatives combined the 20 (non)factive predicates of \citealt{degen-tonhauser-language} (see Fig.~\ref{fig:dt1a}) with the 20 complement clauses of \citealt{degen-tonhauser-language}, for a total of 400 interrogatives. The preceding declaratives implemented a three-level context condition: In the `explicit ignorance' context (\ref{sample}a), the declarative sentence conveyed the speaker's ignorance about the CC. In the `lower prior probability' context (\ref{sample}b), the CC had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context (\ref{sample}c). See Supplement \ref{a:clauses} for the full set of 20 complement clauses and the two contexts for each complement clause.

\begin{exe}
\ex\label{sample}
\begin{xlist}
\ex Explicit ignorance context: \\ I have no idea if Julian dances salsa. Does Cole discover that Julian dances salsa?
\ex Lower prior probability context: \\ Julian is German. Did Cole discover that Julian dances salsa?
\ex Higher prior probability context: \\ Julian is Cuban. Did Cole discover that Julian dances salsa?'
\end{xlist}
\end{exe}

In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included six control stimuli, shown in (\ref{filler}). The interrogatives of the control stimuli featured six expressions typically analyzed as presupposition triggers, namely {\em stop, continue, again, too, also}, and an {\em it-}cleft. The preceding declaratives conveyed the speaker's explicit ignorance about the presuppositions associated with these expressions.\footnote{As the stimuli were presented in writing, prosody was not controlled for. Since the projective contents of {\em too, also}, and the {\em it-}cleft cannot be accommodated, the filler stimuli with these expressions should receive low naturalness in explicit ignorance context under any focus assignment.} Three of these expressions, namely {\em stop, continue,} and {\em again}, were target expressions in the relevant experiments in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, who assumed these expressions to be unnatural in explicit ignorance contexts. This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is acceptable in explicit ignorance contexts. The other three expressions, namely {\em too, also,} and the {\em it-}cleft, were assumed to be hard triggers in \citealt{simons01} and \citealt{abusch10}, that is, unnatural in explicit ignorance contexts. The control stimuli were included for comparison to the target stimuli, to the claims of \citealt{simons01} and \citealt{abusch10}, and to the results of \citealt{mandelkern-etal2020}.

\begin{exe}
\ex\label{filler} 
\begin{xlist}
\ex I don't know if Stephen was ever in the habit of vaping. Has Stephen recently stopped vaping?
\ex I don't know if John was ever reading ``Dune''. Has John recently continued reading ``Dune''?
\ex I don't know if William was ever interested in history. Is William interested in history again?"
\ex I don't know if Ann plays any instrument. Does Ann play the flute, too?
\ex I don't know if Svenja plays any sport. Does Svenja also play soccer?
\ex I don't know if anyone was playing outside with the kids. Was it Jack who was playing outside with the kids?

\end{xlist}
\end{exe}

The experiment also included 4 filler stimuli that were used to exclude participants not attending to the task (see Supplement \ref{a:fillerPractice}). The filler stimuli were expected to receive high naturalness ratings.

A random set of 30 stimuli was created for each participant. Each set contained 20 target stimuli in which each of the 20 complement clauses was paired with a unique clause-embedding predicate. Twelve of the target stimuli were presented in the explicit ignorance context, and the other eight in a low or a higher prior probability context (four each). Each participants' set of 30 stimuli contained the same six control stimuli and the same four filler stimuli. Each of the 30 stimuli were presented as utterances by a unique named speaker. Trial order was randomized for each participant.

\subsubsection{Procedure}

Participants were instructed to rate ``how natural the question sounds in the context of the statement''. As shown in Figure \ref{f:trials}, they were asked give their rating on a slider from `totally unnatural' (coded as 0) to `totally natural' (coded as 1). The experiment began with four practice trials to familiarize participants with the task (see Supplement \ref{a:fillerPractice} for details). After responding to the 30 trials, participants filled out a short optional demographic survey. To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.

\begin{figure}[h]
\centering
%\begin{subfigure}{0.49\textwidth}
\fbox{\includegraphics[width=.7\textwidth]{figures/trial-eic}}
%\caption{Trial with explicit ignorance context}
%\end{subfigure}
%\\[.1cm]
%\begin{subfigure}{0.49\textwidth}
%\fbox{\includegraphics[width=\textwidth]{figures/trial-low}}
%\caption{Trial with lower prior probability context}
%\end{subfigure} \hfill \begin{subfigure}{0.49\textwidth}
%\fbox{\includegraphics[width=\textwidth]{figures/trial-high}}
%\caption{Trial with higher prior probability context}
%\end{subfigure}
\caption{Sample experiment trial with explicit ignorance context.}\label{f:trials}
\end{figure}

\subsubsection{Data exclusion}

We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the attention checks was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four controls was not used to exclude participants because it received a mean rating of only .5. See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-binary, 2 preferred to not disclose). Each of the 20 predicates received at least 200 ratings in the explicit ignorance context (mean: 222 ratings), at least 59 ratings in the lower prior probability context (mean: 74), and at least 61 ratings in the higher prior probability context (mean: 74). The six control stimuli received 370 ratings each.

\subsection{Results}

We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for a binary difference between presuppositions and nonpresuppositions (\S\ref{s:analysis1}). We then consider whether a comparison of naturalness ratings in the contexts provides such evidence (\S\ref{s:analysis2}).

\subsubsection{Naturalness ratings in explicit ignorance contexts}\label{s:analysis1}

Recall that \citealt{simons01} and \citealt{abusch10} assumed that hard presuppositions are unnatural in explicit ignorance contexts, whereas \citealt{mandelkern-etal2020} assumed that all presuppositions are unnatural in such contexts. Figure \ref{fig:acc-by-expression} shows the mean naturalness ratings in the explicit ignorance context by expression (in distinct colors: \color{orange}factive predicates\color{black}, \color{green}nonfactive predicates\color{black},  controls). As shown, purported presuppositions, that is, the controls and the factive predicates, do not exhibit a unified pattern: The mean naturalness ratings of four controls ({\em continue, too, also, again}) is at floor, those of {\em stop} and {\em be annoyed} are a bit higher, those of the {\em it-}cleft and {\em know} are higher again,\footnote{The distribution of the ratings for the CC of {\em know} is bimodal. We discuss this in \S\ref{s3}.} and those of {\em see, discover}, and {\em reveal} are just as high as the means of some nonfactive predicates. These results do not provide empirical support for a binary distinction between presuppositions and nonpresuppositions, contrary to what \citealt{mandelkern-etal2020} hypothesized. Furthermore, these results do not provide empirical support for for a class of factive predicates that is categorically distinct from nonfactive predicates, in line with the result of \citealt{degen-tonhauser-language} based on inference rating measures.

\begin{figure}[h!]
\centering
\includegraphics[width=1\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
\caption{Mean naturalness rating in explicit ignorance context by expression (\color{orange}factive\color{black}, \color{green}nonfactive\color{green}, \color{black}filler\color{black}). Error bars indicate 95\% bootstrapped confidence intervals. Overlaid violin plots indicate the kernel probability density of the participants' ratings.}\label{fig:acc-by-expression}
\end{figure}

These observations were confirmed by a posthoc pairwise comparison of the naturalness ratings of the contents associated with the six controls and 20 target expressions using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}). The input to the pairwise comparison came from a Bayesian mixed-effects beta regression model with weakly informative priors using the `brms' package (\citealt{buerkner2017}). The model predicted naturalness ratings from a fixed effect of expression (with treatment coding and `continue' as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause). We thus obtain, for each predicate, a 95\% credible interval for the mean rating for that predicate. This analysis provided us with with posterior distributions of estimated marginal means of pairwise differences. We assume that two expressions differ in their mean naturalness rating when the posterior distribution of their difference does not include 0. See Supplement \ref{a:analysis1} for the full model output.

As shown in Table \ref{t:pairwise}, the analysis suggests difference between the naturalness ratings of the factive predicates in the explicit ignorance context: the ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}. Furthermore, the naturalness ratings for the CCs of the five factive predicates are indistinguishable from different sets of expressions: The naturalness ratings for {\em be annoyed} do not differ from those of {\em stop} (but are higher than those the prejacents of the controls {\em continue, too}, and {\em also}), the naturalness ratings for {\em know} do not differ from those of the {\em it-}cleft, the naturalness ratings for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, acknowledge}, and {\em admit}, and those of {\em reveal} do not differ from those of the nonfactives {\em prove} and {\em announce}. These results confirm that naturalness ratings in explicit ignorance contexts do not provide empirical support for a binary distinction between factive and nonfactive predicates, or between presuppositions and nonpresuppositions.

\begin{table}[!h]
\begin{center}
\begin{tabular}{r ccccccccccccccccccccccccc}
\toprule
 &  \rot{continue} & \rot{too} & \rot{also} & \rot{again} & \rot{stop} & \rot{be annoyed} & \rot{cleft} & \rot{know} & \rot{demonstrate} & \rot{pretend} & \rot{inform} & \rot{confess} & \rot{see} & \rot{acknowledge} & \rot{discover} & \rot{admit} & \rot{hear} & \rot{prove} & \rot{reveal} & \rot{announce} & \rot{be right} & \rot{establish} & \rot{confirm} & \rot{suggest} & \rot{think} \\
 \midrule
 too & & & & & & & & & & & & & & & & & & \\
also & & & & & & & & & & & & & & & & & & \\
again  & & & & & & & & & & & & & & & & & & \\
stop  & & & & & & & & & & & & & & & & & & \\
be annoyed & & & & & & & & & & & & & & & & & & \\
cleft & & & & & & & & & & & & & & & & & & \\
know & & & & & & & & & & & & & & & & & & \\
demonstrate & & & & & & & & & & & & & & & & & & \\
pretend & & & & & & & & & & & & & & & & & & \\
inform & & & & & & & & & & & & & & & & & & \\
confess & & & & & & & & & & & & & & & & & & \\
see & & & & & & & & & & & & & & & & & & \\
acknowledge & & & & & & & & & & & & & & & & & & \\
discover & & & & & & & & & & & & & & & & & & \\
admit & & & & & & & & & & & & & & & & & & \\
hear & & & & & & & & & & & & & & & & & & \\
prove & & & & & & & & & & & & & & & & & & \\
reveal & & & & & & & & & & & & & & & & & & \\
announce & & & & & & & & & & & & & & & & & & \\
be right & & & & & & & & & & & & & & & & & & \\
establish & & & & & & & & & & & & & & & & & & \\
confirm & & & & & & & & & & & & & & & & & & \\
suggest & & & & & & & & & & & & & & & & & & \\
say  & & & & & & & & & & & & & & & & & & \\
\bottomrule
\end{tabular}
\caption{Posterior distributions of estimated marginal means of pairwise differences `***' indicates significance at .001, `**' at .01, `*' at .05, `.' marginal significance at .1, and \emph{n.s} indicates no significant difference in means.}\label{t:pairwise}
\end{center}
\end{table}

Contrary to what is assumed in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, presuppositions are not invariably unnatural in explicit ignorance contexts. Rather, in line with what was assumed in \citealt{simons01} and \citealt{abusch10}, some presuppositions seem to be rated as more unnatural in such contexts than others. On the assumption that naturalness ratings in explicit ignorance contexts differentiate between undefeasible/hard and defeasible/soft triggers, one might take the results of our experiment to suggest that the pre-state contents of {\em continue} and {\em again}, and the existence requirements of {\em too} and {\em also} are not defeasible. The CCs of {\em see, discover}, and {\em reveal}, on the other hand, might be taken to be defeasible. Contrary to what \citealt{abusch10} assumed, the existence requirement of the {\em it-}cleft is not as clearly undefeasible as that of other projective contents (\citealt{smith-hall11} also observed that the {\em it-}cleft did not pattern like a hard trigger in their experiment.) The results of our experiment align with our observation about possible by-content variation in \citepos{mandelkern-etal2020} Exp.~3. Recall from above that in their Exp.~3, the pre-state of {\em continue} was judged as (numerically) less natural in explicit ignorance contexts than the CC of the cognitive predicate {\em be aware}. This result is in line with the result of our experiment that the pre-state of {\em continue} is judged as less natural than the CC of the cognitive predicate {\em know}.

\subsubsection{Comparison of naturalness ratings across contexts}\label{s:analysis2}

We next consider whether presuppositions and nonpresuppositions differ with respect to the ratings in contexts in which the speaker is explicitly ignorant of the content and in contexts that are compatible with the content. Recall that \citealt{mandelkern-etal2020} expected presuppositions to be sensitive to their context manipulation (explicit ignorance vs.\ support) but not nonpresuppositions. Fig.~\ref{fig:acc-by-context} shows mean naturalness ratings for the CCs of the 20 \fcolorbox{black}{orange}{factive} and \fcolorbox{black}{green}{nonfactive} predicates in the three contexts featured in our experiment, with the predicates ordered by their mean naturalness ratings in the explicit ignorance context, as in Fig.~\ref{fig:acc-by-expression}.

As shown, the effect of context is not uniform for the factive predicates: For {\em be annoyed} and {\em know}, the mean naturalness rating is lower in the explicit ignorance context than in the higher prior probability context, the two ratings are equally high for {\em see, discover}, and {\em reveal}. This latter pattern is also observed for many of the nonfactive predicates, for instance {\em confess, hear, announce}, and {\em be right}. Thus, naturalness ratings in explicit ignorance vs.\ higher prior probability contexts do not provide support for a binary distinction between the CCs of factive and nonfactive predicates. A comparison between the explicit ignorance and lower prior probability contexts also suggests variation between the factive predicates: Whereas for {\em be annoyed} the mean naturalness ratings are lower in the explicit ignorance context than in the lower prior probability context, they are equally high for {\em know}, but lower in the lower prior probability context than the explicit ignorance context for {\em see, discover}, and {\em reveal}. This latter pattern is observed for all of the nonfactive predicates, except possibly {\em pretend}. Thus, naturalness ratings in explicit ignorance vs.\ lower prior probability contexts also do not provide support for a binary distinction between the CCs of factive and nonfactive predicates.

Finally, we observe that the prior probability of content modulates its naturalness ratings for all predicates, except {\em demonstrate} and {\em pretend}): Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability. This result is in line with the result of \citealt{degen-tonhauser-openmind} based on an inference rating task.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
\caption{\small{Mean naturalness rating by context and predicate (\fcolorbox{black}{orange}{factive}, \fcolorbox{black}{green}{nonfactive}); predicates ordered as in Fig.~\ref{fig:acc-by-expression}. Error bars indicate 95\% bootstrapped CIs. Violin plots indicate kernel probability density of participants' ratings}.}\label{fig:acc-by-context}
\end{figure}

These results were confirmed by a model...{\bf BETA MODELS}

\newpage

bla

\newpage

\section{Discussion}\label{s3}

The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides empirical support for the hypothesis that projection is a binary property of content and distinguishes presuppositions from nonpresuppositions. Neither the by-content comparison of naturalness ratings in explicit ignorance contexts nor the by-context comparison of naturalness ratings provided empirical support for a binary distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions. In sum, the experiment did not find support for \citepos{mandelkern-etal2020} claim nor for the hypothesis that projection is a binary property of content. Instead, the experiment provided further support for the hypothesis of \citealt{tbd-variability} that projection is a gradient property of content and the result of \citealt{degen-tonhauser-language} that there is no well-defined class of factive predicates. In this section, we discuss theoretical and methodological implications of these results.

\subsection{What are presuppositions?}

Contemporary literature still frequently treats presuppositions as if they were a homogenous class, for instance by characterizing them as contents that typically project or as felicity requirements. Research on projective content has long recognized that the set of contents that have been called presuppositions is heterogeneous. Two dimensions of variation that were already mentioned in \S\ref{s1}, namely defeasibility and strength of projection, fly in the face of this homogeneity assumption.  A third dimension of variation is whether the presupposition is associated with what \citealt{brst-lang11} referred to as a `strong contextual felicity' constraint: Some presumed presuppositions are not associated with such a constraint, which means that they are judged to be acceptable in a context that is neutral with respect to the content (like the CC of {\em know}, which is informative in such a context), while others are associated with such a constraint, that is, judged to be unacceptable in a context that does not entail or satisfy the content (like the existence requirement of pronouns or {\em too}). A fourth dimension of variation is at-issue, that is, whether the content can address the Question Under Discussion: As shown in \citealt{tbd-variability}, the CCs of factive predicates vary in this regard, as the CC of {\em discover} is more at-issue than those of  {\em be annoyed} and {\em know}.

Given that the set of contents called presuppositions is heterogeneous, it is not entirely surprising that these contents do not pattern alike in explicit ignorance contexts. For instance, a content that is associated with a strong contextual felicity constraint, like the existence requirement of {\em too}, might be expected to be judged as unnatural in a context in which the speaker is explicitly ignorant about the content. On the other hand, a content that is not associated with a strong contextual felicity constraint and that can be at-issue, like the CC of {\em discover}, might be expected to be judged as somewhat natural in a context in which the speaker is explicitly ignorant about the content, as the content can be interpreted as what the speaker is inquiring about. Our experiment suggests that both of these expectations were borne out by the data.

\begin{exe}
\ex
\begin{xlist}
\ex If Julian dances salsa, then Cole knows that he does.
\ex If Julian dances salsa, then 
\end{xlist}
\end{exe}

Against this background, it does not seem fruitful to us to

\begin{itemize}

\item When \citealt{mandelkern-etal2020} talk about presuppositions, it's not even clear what they mean: +SCF or -SCF? They only investigated -SCF/+OLE expression but assumed an analysis for them that was developed for +SCF triggers.

\end{itemize}

{\bf need to discuss in the results what naturalness in EIC appears to be sensitive to, and also bimodal for know and low responses for be annoyed}

\subsection{Is projection a binary or a gradient property of content?}

\begin{itemize}

\item We need to identify which level we're talking about

\begin{itemize}

\item Speaker belief:

Option 1: Speaker belief is binary, gradience is introduced by interpretation.

Option 2: Speaker belief is gradient.

\item Lexical representation:

Option 1: Representation codes projection

Option 2: Representation does not code projection but projection is possible in particular utterances

\end{itemize}

\item Typically, researchers ask this because they wonder about the linguistic representation we want to assume in our models. Simple picture:

triggers introduce context requirement and hence content typically projects

nontriggers do not introduce this and hence exhibit projection to varying degrees

\item More complex picture, already represented by contemporary analyses:

+SCF: introduce anaphoric requirement/felicity requirement

soft triggers: lexical meaning typically gives rise to projection inference

nontriggers: lexical meaning may give rise to projection inference

\item To address this question, we have to distinguish whether we're talking about content associated with a particular expression (aggregated projection over multiple utterances) versus utterance content

\item The characterization of presuppositions is at the level of content, but has to involve `typically' to capture that a presupposition in some utterances does not project.

\item We also have to distinguish the speaker and the hearer: The speaker may be binary or gradient, and even if binary, interpretation may vary because of different assumptions about speaker, different integration of other factors.

\item Projection inferences, sources of gradience

\begin{tabular}{l c c c c} \\ 
             & expression & utterance & speaker & interpreter \\\hline
class 1 &  binary & binary & binary & binary \\
class 2 &  binary & & & \\ 
class 3 & & & & \\

\end{tabular}

\item What does gradient projection mean?

\begin{itemize}

\item[{[H1]}] Projection is binary at the level of the expression: An expression either triggers a presupposition or not. The observed gradience at the utterance level is due to utterance-level factors that modulate projection.

\item[{[H2]}] Projection is not binary at the level of the expression: The lexical representations of (classes of) expressions give rise to projection inferences of varying strength. The observed gradience at the utterance level is due to the lexical variation in projection strength as well as utterance-level factors that modulate projection.

\end{itemize}

\item Contemporary presupposition analyses (e.g., \citealt{heim83,vds92,abrusan2011,karttunen2016,simons-etal2017}) predict neither the heterogeneity of the naturalness ratings of factive predicates nor their sensitivity to the prior probability of the CC (though see \citealt{schlenker2021} for the latter). These results, which mirror those based on inference ratings, suggest that projection analyses must incorporate more fine-grained distinctions between attitude predicates and the systematic effect of prior probabilities on projection.

\end{itemize}

\subsection{Which measures are suitable to investigate projection?}

\begin{itemize}

\item And while naturalness ratings in EICs may provide insight into the meanings of attitude predicates, the interpretation of the results depends on the presumed linking function (see, e.g., \citealt{sprouse2018}). Instead of the ratings directly reflecting whether the CC is presupposed, the very low mean rating for {\em be annoyed} in EICs could be due to its CC being not-at-issue while the EIC suggests that it is at-issue; likewise, reading {\em know} with a contrastive pitch accent in EICs may have resulted in about half of the participants giving a high rating.

\item This result suggests that interrogatives are particularly acceptable when the speaker is either ignorant about the CC or the CC has a higher prior probability: to inquire about the CC, or to inquire about the matrix clause.

\item We also observe that the mean naturalness rating is lower in the lower prior probability context than in the higher prior probability context, for all of the predicates except for {\em pretend}. This suggests that the prior probability of the CC modulates the naturalness of the interrogative regardless of whether the embedding predicate is factive, where the observed context-sensitivity is unexpected under standard analyses (e.g., \citealt{heim83,vds92}) or nonfactive (where it is expected). This result mirrors that of \citealt{degen-tonhauser-openmind}, who found in an inference rating task the strength of projection inferences to the CCs of factive and nonfactive predicates is modulated by the prior probability of the CC. One could, of course, propose to make the global accommodation of presuppositions to not be default but to be sensitive to the prior probability of content. But the fact of the matter is that the context-sensitivity is observed not just for factive but also for nonfactive predicates, thereby motivating a common projection mechanism.

\item The talk also discusses that the results do not align with assumptions about hard and soft triggers; \citealt{simons01,abusch10}.

\end{itemize}

\subsection{What does an empirically adequate analysis of projection need to capture?}

\section{Conclusions}\label{s4}

% end document here for word count
%\end{document}

\bibliographystyle{../cslipubs-natbib}
%\bibliographystyle{/Users/tonhauser.1/Library/Latex/cslipubs-natbib}
\bibliography{/Users/tonhauser.1/Documents/bibliography}

\newpage

\section*{Supplemental materials}

\appendix

\setcounter{page}{1}
%\renewcommand{\thetable}{A\arabic{table}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\section{Experiment stimuli}\label{a:clauses}

The twenty predicates used in the experiment are the same as in \citealt{degen-tonhauser-openmind,degen-tonhauser-language}:

\begin{exe}
\ex\label{predicates}
\begin{xlist}
\ex Factive: be annoyed, discover, know, reveal, see
\ex Non-factive: acknowledge, admit, announce, be right, confess, confirm, establish, hear, inform, pretend, prove, say, suggest, think
\end{xlist}
\end{exe}

Eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense. The direct object of {\em inform} was realized by the proper name {\em Sam}. Each clause-embedding predicate was paired with a unique subject proper name. The speaker of the target stimuli was realized by a randomly sampled unique proper name.

The following list shows the 20 clauses that realized the complements of the predicates in the target and filler stimuli, together with their lower and higher probability facts, respectively, that realized the preceding declarative sentences in the filler stimuli.

\begin{enumerate}[leftmargin=4ex,itemsep=-2pt]
\item Mary is pregnant. Facts: Mary is a middle school student / Mary is taking a prenatal yoga class
\item Josie went on vacation to France. Facts:  Josie doesn't have a passport / Josie loves France 
\item Emma studied on Saturday morning. Facts: Emma is in first grade / Emma is in law school 
\item Olivia sleeps until noon. Facts: Olivia has two small children / Olivia works the third shift
\item Sophia got a tattoo. Facts: Sophia is a high end fashion model / Sophia is a hipster
\item Mia drank 2 cocktails last night. Facts: Mia is a nun / Mia is a college student
\item Isabella ate a steak on Sunday. Facts: Isabella is a vegetarian / Isabella is from Argentina
\item Emily bought a car yesterday. Facts: Emily never has any money / Emily has been saving for a year
\item Grace visited her sister. Facts: Grace hates her sister / Grace loves her sister
\item Zoe calculated the tip. Facts: Zoe is 5 years old / Zoe is a math major
\item Danny ate the last cupcake. Facts: Danny is a diabetic / Danny loves cake
\item Frank got a cat. Facts: Frank is allergic to cats / Frank has always wanted a pet
\item Jackson ran 10 miles. Facts: Jackson is obese / Jackson is training for a marathon
\item Jayden rented a car. Facts: Jayden doesn't have a driver's license / Jayden's car is in the shop
\item Tony had a drink last night. Facts: Tony has been sober for 20 years / Tony really likes to party with his friends
\item Josh learned to ride a bike yesterday. Facts: Josh is a 75-year old man / Josh is a 5-year old boy
\item Owen shoveled snow last winter. Facts: Owen lives in New Orleans / Owen lives in Chicago
\item Julian dances salsa. Facts: Julian is German / Julian is Cuban
\item Jon walks to work. Facts: Jon lives 10 miles away from work / Jon lives 2 blocks away from work
\item Charley speaks Spanish. Facts: Charley lives in Korea / Charley lives in Mexico
\end{enumerate}

\section{Control and practice stimuli}\label{a:fillerPractice}

The following list shows the four control stimuli where the interrogative was expected to receive high naturalness ratings in the context of the preceding declarative sentence. The values in parentheses indicate the mean naturalness rating for each of the four controls. As shown, the third control did not receive the expected high naturalness mean ratings (probably because participants were unwilling to accommodate that Hendrick has a car in a context in which Hendrick was looking to buy a car). This control was therefore not used to exclude participants' data.

\begin{enumerate}[leftmargin=4ex,itemsep=-2pt]

\item I don't know if Samantha has a new hat. Does Samantha have a new hat? (.89)

\item  I don't know if this pizza has mushrooms on it. Does this pizza have mushrooms on it? (.87)

\item Hendrick was looking to buy a car. Was Hendrick's car expensive? (.5)

\item Mary visited her aunt yesterday. Is Mary's aunt sick? (.91)

\end{enumerate}

\noindent
The following list shows the four practice stimuli in the order in which they were presented to the participants. Participants were able to advance to the experiment only if they gave a naturalness rating higher than .6 for the first and third stimulus, and a naturalness rating lower than .4 for the second and fourth stimulus.

\begin{enumerate}[leftmargin=4ex,itemsep=-2pt]

\item I have no idea where Natalie is from. Is Natalie from the USA?

\item I don't have any sisters. Have you met my sister yet?

\item I am going on vacation to Ireland. Does Fritz realize that Joe is going with me?

\item I have no idea if Anna has any dogs. Is Samuel glad that Anna fed her dogs?

\end{enumerate}

\section{Model output: Pairwise comparison of expressions}\label{a:analysis1}

\begin{sidewaysfigure}[h!]
\centering
\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/comparisons-in-EIC}
\caption{.}\label{fig:comparisons1}
\end{sidewaysfigure}

\section{Model output: Pairwise comparison of contexts}\label{a:analysis2}

\begin{figure}[h!]
\centering
\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/context-comparisons}
\caption{.}\label{fig:comparisons2}
\end{figure}

\end{document}

### 2023-11-05 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -7,44 +7,534 @@ setwd(this.dir)
 
 # load required packages
 library(tidyverse)
-library(lme4)
+library(tidybayes)
+library(dplyr)
+library(dichromat)
+library(forcats)
+library(ggrepel)
+library(brms)
+library(knitr)
 library(emmeans)
+library(lme4)
+library(lmerTest)
+library(padr)
+library(performance)
+library(MuMIn)
 
 # load cleaned data
 d = read_tsv("../data/cd.tsv")
 
-names(d)
-table(d$context)
+# analysis 1: pairwise comparison of ratings in explicit ignorance context ----
 
-length(unique(d$participantID)) #370 participants
-
-# predict naturalness rating from expression in explicit ignorance context ----
-
-# target data: explicit ignorance context, no controls
-t = d %>%
+# target data and set reference levels
+t = d %>% 
   filter(context == "explicitIgnorance") %>%
-  filter(expression != "controlGood1" & expression != "controlGood2")
+  filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4")
+  
+# set reference levels
+t = t %>%
+  mutate(expression = fct_relevel(expression, "continue"))
 
-# sort the expressions by mean naturalness rating (for ease of reading multiple comparisons)
-means = t %>%
-  group_by(expression) %>%
-  summarize(mean = mean(response)) %>%
-  mutate(expression = fct_reorder(as.factor(expression),mean)) 
-means
+table(t$expression)
+table(t$context)
+
+# first, because response assumes values of 0 and 1, which beta regression cannot handle, transform: (Smithson & Verkuilen 2006)
+# y'' = (y' ?? (n ??? 1) + 0.5)/n
+# note: first rescaling of y'=(y-a)/(b-a) not necessary because highest and lowest value are 0 and 1 already
+t$betaresponse = (t$response*(nrow(t)-1) + .5)/nrow(t)
+
+prior = get_prior(betaresponse ~ expression + (1|participantID) + (1|cc),family = Beta(),data=t)
+prior
+
+betamodel = bf(betaresponse ~ expression + (1|participantID) + (1|cc),
+               phi ~ expression + (1|participantID) + (1|cc), # beta distribution's precision 
+               family = Beta())
+
+m.b = brm(formula = betamodel,
+          family=Beta(),
+          data=t, 
[…]

### 2023-11-05 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -31,7 +31,6 @@ length(unique(d$participantID)) #370 participants
 # Fig 1: plot of mean naturalness ratings in explicit ignorance context ----
 
 # target data: explicit ignorance context
-# merge the two controls into one, but exclude them
 table(d$expression)
 t = d %>%
   filter(context == "explicitIgnorance") %>%
@@ -48,24 +47,28 @@ levels(nat.meansEIC$expression)
 
 # color code the expressions
 factives <- c("know", "discover", "be annoyed", "reveal", "see")
-hard.triggers <- c("too", "also","cleft","again")
-soft.triggers <- c("stop", "continue")
+fillers <- c("too", "also","cleft","again","stop", "continue")
+#hard.triggers <- c("too", "also","cleft","again")
+#soft.triggers <- c("stop", "continue")
 
-nat.meansEIC$ps = ifelse(nat.meansEIC$expression %in% soft.triggers, "softTrigger", 
-               ifelse(nat.meansEIC$expression %in% hard.triggers, "hardTrigger",
-                      ifelse(nat.meansEIC$expression %in% factives, "factive", "other")))
+# nat.meansEIC$ps = ifelse(nat.meansEIC$expression %in% soft.triggers, "softTrigger", 
+#                ifelse(nat.meansEIC$expression %in% hard.triggers, "hardTrigger",
+#                       ifelse(nat.meansEIC$expression %in% factives, "factive", "other")))
+# 
+# t$ps = ifelse(t$expression %in% soft.triggers, "softTrigger", 
+#               ifelse(t$expression %in% hard.triggers, "hardTrigger",
+#                      ifelse(t$expression %in% factives, "factive", "other")))
 
-t$ps = ifelse(t$expression %in% soft.triggers, "softTrigger", 
-              ifelse(t$expression %in% hard.triggers, "hardTrigger",
-                     ifelse(t$expression %in% factives, "factive", "other")))
+nat.meansEIC$ps = ifelse(nat.meansEIC$expression %in% fillers, "filler",
+                                ifelse(nat.meansEIC$expression %in% factives, "factive", "other"))
 
+t$ps = ifelse(t$expression %in% fillers, "filler",
+                     ifelse(t$expression %in% factives, "factive", "other"))
 
 table(nat.meansEIC$ps, nat.meansEIC$expression)
 
[…]

### 2023-11-05 results/main/13explicitIgnorance/rscripts/preprocessing.R

@@ -477,6 +477,7 @@ write_tsv(d, file="../data/cd.tsv")
 # how many data points per predicate/CC/context combination?
 names(d)
 
+# exclude the fillers and the controls
 tmp = d %>%
   filter(cc != "noCC") %>%
   group_by(expression,cc,context) %>% 
@@ -515,17 +516,31 @@ mean(tmp$n) #222
 median(tmp$n) #224.5
 max(tmp$n) #234
 
+table(d$context)
 # how many data points for each predicate in the two neutral contexts?
 tmp = d %>%
-  filter(context != "explicitIgnorance") %>%
+  filter(context == "factL") %>%
   filter(cc != "noCC") %>%
   group_by(expression,context) %>% 
   tally
 tmp
-nrow(tmp) # 40 combinations (2 contexts x 20 predicates)
-sd(tmp$n) #8
+nrow(tmp) # 20 (1 context x 20 predicates)
+sd(tmp$n) #8.25
 min(tmp$n) #59
 mean(tmp$n) #74
 median(tmp$n) #75
 max(tmp$n) #87
+
+tmp = d %>%
+  filter(context == "factH") %>%
+  filter(cc != "noCC") %>%
+  group_by(expression,context) %>% 
+  tally
+tmp
+nrow(tmp) # 20 (1 context x 20 predicates)
+sd(tmp$n) #8
+min(tmp$n) #61
+mean(tmp$n) #74
+median(tmp$n) #75
+max(tmp$n) #86
 

### 2023-11-16 writing/naturalness-paper/paper.tex
updated naturalness paper

@@
 \usepackage{subcaption}
+\usepackage{colortbl}
 
@@
 
-\title{I don't know if projection is binary. \\ Did \citealt{mandelkern-etal2020} discover that it is?}
+\newcommand*\rots{\multicolumn{1}{R{90}{.7em}}}% no optional argument here, please!
+
+% positive coefficients/difference
+\definecolor{yellow1}{RGB}{102,102,0} 
+\definecolor{yellow2}{RGB}{153,153,0} 
+\definecolor{yellow3}{RGB}{204,204,0} 
+\definecolor{yellow4}{RGB}{255,255,204}
+%\definecolor{green1}{RGB}{0, 158, 115} % >.95
+%\definecolor{green2}{RGB}{55, 185, 141} % .85-.95
+%\definecolor{green3}{RGB}{88, 214, 167} % .75-.85
+%\definecolor{green4}{RGB}{119, 242, 194} % <.75
+
+% negative coefficients/difference
+\definecolor{purple1}{RGB}{150,57,109}
+\definecolor{purple2}{RGB}{188,78,139}
+\definecolor{purple3}{RGB}{215,149,186}
+\definecolor{purple4}{RGB}{236,206,223}
+
+\title{I don't know if projection is categorical. \\ Did \citealt{mandelkern-etal2020} discover that it is?}
 
@@
 
-Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{frege1892,ccmg90}).
-While projection is often characterized as binary (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
-\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and proposed that naturalness ratings in explicit ignorance contexts are more suitable to measure binary projection and, thereby, to distinguish presupposition triggers from nontriggers.
+Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{ccmg90}).
+While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
+\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that naturalness ratings in explicit ignorance contexts are more suitable to measure categorical projection and, thereby, to distinguish presuppositions from nonpresuppositions.
 This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} can distinguish factive predicates (presumed presupposition triggers) from nonfactive ones (nontriggers).
-While the results do not support \citepos{mandelkern-etal2020} hypothesis and are rather in line with results based on inference rating measures (e.g., \citealt{degen-tonhauser-language}), the results suggest that naturalness ratings in explicit ignorance contexts are sensitive to felicity conditions of anaphoric projective content.
+The results do not support \citepos{mandelkern-etal2020} claim as they do not provide evidence for a categorical distinction between presuppositions and nonpresuppositions.
+{\bf FIX: Furthermore, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to a variety of semantic and pragmatic properties of utterances in addition to the felicity conditions of anaphoric projective content.}
 
@@
 \noindent
-{\bf Keywords:} Presuppositions, binary vs.\ gradient projection, inference ratings, naturalness ratings in explicit ignorance contexts.
+{\bf Keywords:} Presuppositions, categorical vs.\ gradient projection, inference ratings, naturalness ratings in explicit ignorance contexts.
 
@@
 \noindent
-For helpful comments on this project we thank Craige Roberts, Gregory Scontras, and Mandy Simons.
+For helpful comments on this project we thank Craige Roberts, Gregory Scontras, Mandy Simons, and the audience at the syntax/semantics discussion group at the University of Stuttgart.
 
@@
 
-Presuppositions are defined as content that typically projects out of entailment-canceling environments, like the scope of negation or a polar interrogative (e.g., \citealt{frege1892,ccmg90}).\footnote{Conventional implicatures also project out of these environments.
+Presuppositions are often characterized as content that typically projects out of entailment-canceling environments, like the scope of negation or a polar interrogative (e.g., \citealt{ccmg90}).\footnote{Conventional implicatures also project out of these environments.
 They differ from presuppositions in that they contribute novel information.
@@
 
-To investigate whether a particular content is a presupposition, empirical research has long relied on inference rating measures.
-Such measures involve presenting participants with an utterance (like that in (\ref{first}a)) and contents (like the content that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content.
-Many theoretical works on projection assume that projection is a binary property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition).
-There is, however, mounting empirical evidence (based on investigations that used inference rating measures) that projection is not a binary property of  content but rather a gradient one, with  content being more or less projective.
-\citealt{karttunen71b}, for instance, already suggested that the contents of the clausal complements (CCs) of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}).
+Many theoretical works assume that projection is a categorical property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition).
+There is, however, mounting empirical evidence that projection is not a categorical property of  content but rather a gradient one, with content being more or less projective.
+The measure used to establish this evidence are inference rating measures: Such measures involve presenting participants with an utterance (like that in (\ref{first}a)) and contents (like the content that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content.
+\citealt{karttunen71b}, for instance, suggested that the CCs of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}).
 Specifically, he suggested that a speaker who utters (\ref{kart}a) believes that the addressee has not told the truth, whereas a speaker who utters (\ref{kart}b) ``is not sure about the truth of the complement'' (p.63).
@@
 \end{exe}
-More recently, experimental investigations have also observed projection variability between presuppositions (e.g., \citealt{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
+More recently, experimental investigations have also observed projection variability between presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
 \citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CCs of {\em know, discover,} and {\em reveal} exhibit projection variability: The CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}.
-These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''.
+These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a categorical, categorical one''.
 Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}.
-In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a binary distinction in the projection of the CCs of factive and nonfactive predicates.
+In a series of experiments and analyses of existing datasets, this work not only found projection variability between the CCs of factive predicates, but also no empirical support for a categorical distinction in the projection of the CCs of factive and nonfactive predicates.
 This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content.
@@
 \centering
-\includegraphics[width=.8\textwidth]{figures/Language-Figure2}
-\caption{Mean certainty rating (measuring projection) of main clause content and the CCs of the 20 predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category).
+\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/mean-certainty-by-predicateType}
+\caption{Mean certainty rating (measuring projection) of main clause (`MC') content and the CCs of the 20 \color{orange}factive \color{black} and \color{green}nonfactive \color{black} predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category).
 Error bars indicate 95\% bootstrapped confidence intervals.
@@
 
-Recently,  \citealt[\S6.2]{mandelkern-etal2020} challenged the hypothesis that projection is gradient by arguing that inference rating measures are not suitable to detect binary projection.
+Recently,  \citealt*[\S6.2]{mandelkern-etal2020} challenged the hypothesis that projection is gradient by arguing that inference rating measures are not suitable to detect categorical projection.
 Such measures are not suitable, they argued, because they do not distinguish between, on the one hand, contents that project because they are presuppositions and, on the other hand, contents that project because they are ``a natural conclusion to draw for any of a variety of pragmatic reasons short of entailment or presupposition'' (p.497).
@@
 That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).
-If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497).\footnote{Presuppositions won't be relinquished, modulo local accommodation} Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).
+If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497).
+Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).
 
@@
 \end{exe}
-\citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, only provided limited empirical evidence for the hypothesis that projection is a binary property of content.
+\citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, only provided limited empirical evidence for the hypothesis that projection is a categorical property of content.
 This is primarily because the goal of this work (including Exp.~3) was not to investigate this hypothesis, but rather whether presupposition projection in conjunctions is symmetric.
-The evidence provided in this work is also limited, however, because only four presupposition triggers and nontriggers each were investigated: The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i).
+The evidence provided in this work is also limited, however, because only four presupposition triggers and four nontriggers were investigated: The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.
+%\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i).
 It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past.
-
-\begin{exe}
-\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
-\end{exe}
-} In the control items of Exp.~3, these eight expressions were realized in (single-clause) antecedents of conditionals, as shown in (\ref{mtrig}c).
+%
+%\begin{exe}
+%\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
+%\end{exe}
+%} 
+In the control items of Exp.~3, these eight expressions were realized in (single-clause) antecedents of conditionals, as shown in (\ref{mtrig}c).
 Naturalness ratings were elicited in explicit ignorance and support contexts.
@@
 \citealt{mandelkern-etal2020} found (as shown in their Fig.~3) that the naturalness ratings for the nontriggers in the control items were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale).
-The presupposition triggers, on the other hand, had lower mean naturalness ratings in the explicit ignorance context (mean around 3) than in the support context (mean just above 5).
+The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean just above 5).
 Mandelkern and his colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492).
@@
 
-However, an inspection of the naturalness ratings for the eight (non)triggers provided in Fig.~8 in their Appendix 3 raises doubt about whether the measure indeed provides evidence for a binary distinction between presuppositions and nonpresuppositions.
+However, an inspection of the naturalness ratings for the eight (non)triggers provided in Fig.~8 in their Appendix 3 raises doubt about whether the measure indeed provides evidence for a categorical distinction between presuppositions and nonpresuppositions.
 First, considering the mean naturalness ratings in the explicit ignorance contexts, we find ratings for the presuppositions range from just above 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}).
-As the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}), it is not clear whether naturalness ratings in explicit ignorance contexts provide evidence for a binary distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}.
+As the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}), it is not clear that naturalness ratings in explicit ignorance contexts provide evidence for a categorical distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}.
 Second, the effect of context (explicit ignorance vs.\ support) also does not appear to be consistent for presuppositions or nonpresuppositions.
 Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}.
-In sum, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a binary projection either.
+Thus, it is not clear that comparisons of ratings in support and explicit ignorance contexts provides empirical support for a categorical projection either.
 
-There also appears to be disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts.
-Whereas Mandelkern and his colleagues assumed that all presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only hard (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural.
-For instance,  {\em stop} and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic}c) with `??' instead of `$\#$'.} This disagreement in the literature necessitates further investigation.
+Adding to this, there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts.
+Whereas Mandelkern and his colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural.
+For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.} This disagreement about the naturalness of presuppositions in explicit ignorance contexts necessitates further investigation.
 
@@
 \ex I have no idea whether Jane ever smoked, but she hasn't stopped smoking. \hfill (\citealt[443]{simons01})
+\ex Context: ``two people [\ldots] know that Henry is searching for Jane, but who don't themselves know where Jane is:'' \\ If Henry discovers that Jane is in New York, there'll be trouble.
+\hfill (\citealt[434]{simons01})
 \ex I have no idea whether John ended up participating in the
@@
 
-This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} provides evidence that projection is a binary property of content rather than a gradient one.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated.
-The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft, to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.
+This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} provides evidence that projection is a categorical property of content rather than a gradient one.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated.
+The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.
 
-One change we made to the design of \citepos{mandelkern-etal2020} Exp.~3 was that our experiment implemented a three-level context condition instead of a two-level one.
-Specifically, we replaced \citepos{mandelkern-etal2020} support context (which entailed the relevant context, see (\ref{mtrig}b)) with two contexts that were compatible with the relevant content but did not entail it.
-The two contexts differed in the prior probability of the relevant content: In the `lower prior probability' context, illustrated in (\ref{prior}a), the  relevant content (in (\ref{prior}c), that Julian dances salsa) had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context, illustrated in (\ref{prior}b).
-These two contexts were normed in \citealt{degen-tonhauser-openmind}.
+One change made to the design of \citepos{mandelkern-etal2020} Exp.~3 was that the experiment implemented a three-level context condition instead of a two-level one.\footnote{Another change was that participants gave their naturalness ratings on a slider with endpoints labeled `totally unnatural' and `totally natural' rather than on a 7-point Likert scale with endpoints labeled `completely unnatural' and `completely natural'.}  Specifically, \citepos{mandelkern-etal2020} support context (which entailed the relevant context, see (\ref{mtrig}b)) was replaced with two contexts that were compatible with the relevant content but did not entail it.
+The two contexts differed in the prior probability of the  content: In the `lower prior probability' context, illustrated in (\ref{prior}a), the  relevant content (in (\ref{prior}c), that Julian dances salsa) had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context, illustrated in (\ref{prior}b).
+The two contexts for each of the 20 complement clauses that were combined with the 20 predicates were normed in \citealt{degen-tonhauser-openmind}.
 
@@
 \end{exe}
-This change was implemented to allow for comparison to the result of \citealt{degen-tonhauser-openmind}, namely that the projection of a content is sensitive to the content's prior probability.
+This change in the context condition was implemented to allow for comparison to the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability.
 Specifically, this work found that the higher the prior probability of a content, the stronger the projection inference.
-Under standard analyses of presuppositions, like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in such contexts.
+Under standard analyses of presuppositions, like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in contexts that are compatible with the presupposition, like the lower and higher prior probability contexts.
 Accordingly, such analyses predict that presuppositions are rated as natural in both lower and higher probability contexts, just as in \citepos{mandelkern-etal2020} support contexts.
-
-%A challenge in the identification of presuppositions is that they only {\em typically} project: It has long been know that presuppositions need not project in any utterance in which the trigging expression is realized (e.g., \citealt{} REFERENCES).
-For example, in the naturally occurring example in (\ref{acc}) the presupposition of {\em know} (in the framed box) does not project but is (analyzed as) locally accommodated in the scope of the epistemic modal adverb {\em perhaps}.
-%
-%
-%\begin{exe}
-%\ex\label{acc} Whatever the reason, she knows she can't just bounce back, and she decides to leave.
-For others, the decision is based purely on fear.
-Perhaps she \fbox{knows} that if she were to stay another day, her life or the lives of her children would be gone.
-\hfill (\citealt[85]{beaver-belly})
-%\end{exe}
-%
-%
-%In view of this potential for presuppositions to not project, investigations into the question of which utterance contents are presuppositions (and which ones are not) must
 
@@
 
-To investigate whether the measure proposed in \citealt{mandelkern-etal2020} provides support for a binary distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive clause-embedding predicate), and rated the naturalness of the interrogative in the context of the declarative.
+To investigate whether the measure proposed in \citealt{mandelkern-etal2020} provides support for a categorical distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive clause-embedding predicate), and rated the naturalness of the interrogative in the context of the declarative.
 
@@
 We recruited 425 participants on Prolific.
-Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-binary, 2 preferred to not disclose).
+Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-categorical, 2 preferred to not disclose).
 The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%.
 The median time spent on the task was 6:24 minutes.
-Participants were paid \$1.78 (corresponding to an hourly pay of \$16.6).
+Participants were paid \$1.78, corresponding to an hourly pay of \$16.6.
 
@@
 
-In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included six control stimuli, shown in (\ref{filler}).
+In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included the six control stimuli shown in (\ref{filler}).
 The interrogatives of the control stimuli featured six expressions typically analyzed as presupposition triggers, namely {\em stop, continue, again, too, also}, and an {\em it-}cleft.
-The preceding declaratives conveyed the speaker's explicit ignorance about the presuppositions associated with these expressions.\footnote{As the stimuli were presented in writing, prosody was not controlled for.
-Since the projective contents of {\em too, also}, and the {\em it-}cleft cannot be accommodated, the filler stimuli with these expressions should receive low naturalness in explicit ignorance context under any focus assignment.} Three of these expressions, namely {\em stop, continue,} and {\em again}, were target expressions in the relevant experiments in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, who assumed these expressions to be unnatural in explicit ignorance contexts.
-This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is acceptable in explicit ignorance contexts.
-The other three expressions, namely {\em too, also,} and the {\em it-}cleft, were assumed to be hard triggers in \citealt{simons01} and \citealt{abusch10}, that is, unnatural in explicit ignorance contexts.
+The preceding declaratives conveyed the speaker's explicit ignorance about the presuppositions associated with these expressions.\footnote{The contents investigated for {\em too} and {\em also} in (\ref{filler}d) and (\ref{filler}e) are that Ann plays an instrument other than the flute and that Svenja plays a sport other than soccer.
+This interpretation arises if {\em Ann} and {\em Svenja} associate with {\em too} and {\em also}, respectively.
+While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts.
+This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is natural in explicit ignorance contexts.
+Three expressions were assumed to be unnatural in explicit ignorance contexts in \citealt{simons01} ({\em again}) and \citealt{abusch10} ({\em too}, {\em it-}cleft).
 The control stimuli were included for comparison to the target stimuli, to the claims of \citealt{simons01} and \citealt{abusch10}, and to the results of \citealt{mandelkern-etal2020}.
@@
 The experiment began with four practice trials to familiarize participants with the task (see Supplement \ref{a:fillerPractice} for details).
-After responding to the 30 trials, participants filled out a short optional demographic survey.
+After rating the 30 trials, participants filled out a short optional demographic survey.
 To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.
@@
 %\end{subfigure}
-\caption{Sample experiment trial with explicit ignorance context.}\label{f:trials}
+\caption{Sample trial with explicit ignorance context.}\label{f:trials}
 \end{figure}
@@
 
-We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the attention checks was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four controls was not used to exclude participants because it received a mean rating of only .5.
-See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-binary, 2 preferred to not disclose).
+We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the  filler stimuli was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four filler stimuli was not used to exclude participants because it received a mean rating of only .5.
+See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-categorical, 2 preferred to not disclose).
 Each of the 20 predicates received at least 200 ratings in the explicit ignorance context (mean: 222 ratings), at least 59 ratings in the lower prior probability context (mean: 74), and at least 61 ratings in the higher prior probability context (mean: 74).
@@
 
-We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for a binary difference between presuppositions and nonpresuppositions (\S\ref{s:analysis1}).
-We then consider whether a comparison of naturalness ratings in the contexts provides such evidence (\S\ref{s:analysis2}).
+We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for a categorical difference between presuppositions and nonpresuppositions (\S\ref{s:analysis1}).
+We then consider whether a comparison of naturalness ratings in the three contexts provides such evidence (\S\ref{s:analysis2}).
 
@@
 
-Recall that \citealt{simons01} and \citealt{abusch10} assumed that hard presuppositions are unnatural in explicit ignorance contexts, whereas \citealt{mandelkern-etal2020} assumed that all presuppositions are unnatural in such contexts.
-Figure \ref{fig:acc-by-expression} shows the mean naturalness ratings in the explicit ignorance context by expression (in distinct colors: \color{orange}factive predicates\color{black}, \color{green}nonfactive predicates\color{black},  controls).
-As shown, purported presuppositions, that is, the controls and the factive predicates, do not exhibit a unified pattern: The mean naturalness ratings of four controls ({\em continue, too, also, again}) is at floor, those of {\em stop} and {\em be annoyed} are a bit higher, those of the {\em it-}cleft and {\em know} are higher again,\footnote{The distribution of the ratings for the CC of {\em know} is bimodal.
-We discuss this in \S\ref{s3}.} and those of {\em see, discover}, and {\em reveal} are just as high as the means of some nonfactive predicates.
-These results do not provide empirical support for a binary distinction between presuppositions and nonpresuppositions, contrary to what \citealt{mandelkern-etal2020} hypothesized.
-Furthermore, these results do not provide empirical support for for a class of factive predicates that is categorically distinct from nonfactive predicates, in line with the result of \citealt{degen-tonhauser-language} based on inference rating measures.
+Recall that \citealt{simons01} and \citealt{abusch10} assumed that only hard presuppositions are unnatural in explicit ignorance contexts, whereas \citealt{mandelkern-etal2020} assumed that all presuppositions are unnatural in such contexts.
+Figure \ref{fig:acc-by-expression} shows mean naturalness ratings in the explicit ignorance context by expression (in distinct colors: \color{orange}factive predicates\color{black}, \color{green}nonfactive predicates\color{black},  controls).
+As shown, the purported presupposition triggers, that is, the controls and the factive predicates, do not exhibit a unified pattern: The mean naturalness ratings of four controls ({\em continue, too, also, again}) are at floor, those of {\em stop} and {\em be annoyed} are a bit higher, those of the {\em it-}cleft and {\em know} are higher again, and those of {\em see, discover}, and {\em reveal} are just as high as those of some nonfactive predicates.
+These results do not provide empirical support for a categorical distinction between presuppositions and nonpresuppositions, contrary to what \citealt{mandelkern-etal2020} hypothesized.
+These results also do not provide support for a class of factive predicates that is categorically distinct from nonfactive predicates, in line with the result of \citealt{degen-tonhauser-language} based on inference rating measures.
 
@@
 \centering
-\includegraphics[width=1\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
+\includegraphics[width=.9\textwidth]{../../results/main/13explicitIgnorance/graphs/explicit-ignorance-naturalness-by-predicate}
 \caption{Mean naturalness rating in explicit ignorance context by expression (\color{orange}factive\color{black}, \color{green}nonfactive\color{green}, \color{black}filler\color{black}).
@@
 
-These observations were confirmed by a posthoc pairwise comparison of the naturalness ratings of the contents associated with the six controls and 20 target expressions using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}).
-The input to the pairwise comparison came from a Bayesian mixed-effects beta regression model with weakly informative priors using the `brms' package (\citealt{buerkner2017}).
+These observations were confirmed by a posthoc pairwise comparison of the naturalness ratings of the contents associated with the six controls and 20 target expressions in the explicit ignorance context using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}).
+The input to the pairwise comparison was a Bayesian mixed-effects beta regression model with weakly informative priors that was fitted using the `brms' package (\citealt{buerkner2017}).
 The model predicted naturalness ratings from a fixed effect of expression (with treatment coding and `continue' as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause).
-We thus obtain, for each predicate, a 95\% credible interval for the mean rating for that predicate.
-This analysis provided us with with posterior distributions of estimated marginal means of pairwise differences.
-We assume that two expressions differ in their mean naturalness rating when the posterior distribution of their difference does not include 0.
-See Supplement \ref{a:analysis1} for the full model output.
+The pairwise comparison provided us with posterior distributions of estimated marginal means of pairwise differences between the contents associated with each of the 26 expression.
+We assume that two contents differ in their naturalness in the explicit ignorance context when the posterior distribution of their difference does not include 0.
 
-As shown in Table \ref{t:pairwise}, the analysis suggests difference between the naturalness ratings of the factive predicates in the explicit ignorance context: the ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.
-Furthermore, the naturalness ratings for the CCs of the five factive predicates are indistinguishable from different sets of expressions: The naturalness ratings for {\em be annoyed} do not differ from those of {\em stop} (but are higher than those the prejacents of the controls {\em continue, too}, and {\em also}), the naturalness ratings for {\em know} do not differ from those of the {\em it-}cleft, the naturalness ratings for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, acknowledge}, and {\em admit}, and those of {\em reveal} do not differ from those of the nonfactives {\em prove} and {\em announce}.
-These results confirm that naturalness ratings in explicit ignorance contexts do not provide empirical support for a binary distinction between factive and nonfactive predicates, or between presuppositions and nonpresuppositions.
+As shown in Table \ref{t:pairwise},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.} the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates, such that the ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.
+There is also variation between the factive predicates and the presuppositional controls: The CCs of each of the five purportedly factive are more natural in explicit ignorance contexts than the presuppositions contributed by {\em too} and {\em also}.
+Furthermore, the naturalness ratings for the CCs of the five factive predicates are indistinguishable from different sets of expressions: The naturalness ratings for {\em be annoyed} do not differ from those of {\em stop} (but are higher than those the prejacents of the controls {\em continue, too}, and {\em also}), the naturalness ratings for {\em know} do not differ from those of the {\em it-}cleft, the naturalness ratings for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}.
+These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for a categorical distinction between factive and nonfactive predicates, or between presuppositions and nonpresuppositions.
+
+%\begin{table}[!h]
+%\addtolength{\tabcolsep}{-.1em}
+%\centering
+%\begin{tabular}{r ccccccccccccccccccccccccc}
+%\toprule
+% &  \rot{continue\color{black}} & \rot{too\color{black}} & \rot{also\color{black}} & \rot{again\color{black}} & \rot{stop\color{black}} & \rot{\color{orange}be annoyed\color{black}} & \rot{cleft\color{black}} & \rot{\color{orange}know\color{black}} & \rot{\color{green}demonstrate\color{black}} & \rot{\color{green}pretend\color{black}} & \rot{\color{green}inform\color{black}} & \rot{\color{green}confess\color{black}} & \rot{\color{orange}see\color{black}} & \rot{\color{green}acknowledge\color{black}} & \rot{\color{orange}discover\color{black}} & \rot{\color{green}admit\color{black}} & \rot{\color{green}hear\color{black}} & \rot{\color{green}prove\color{black}} & \rot{\color{orange}reveal\color{black}} & \rot{\color{green}announce\color{black}} & \rot{\color{green}be right\color{black}} & \rot{\color{green}establish\color{black}} & \rot{\color{green}confirm\color{black}} & \rot{\color{green}suggest\color{black}} & \rot{\color{green}think\color{black}} \\
+% \midrule
+%too & -- & \cellcolor{lightgray} & & & & & & & & & & & & & & & & \\
+%also & -- & -- & \cellcolor{lightgray} & & & & & & & & & & & & & & & \\
+%again  & -- & -- & -- & \cellcolor{lightgray} & & & & & & & & & & & & & & \\
+%stop  & $\ast$ & $\ast$  & $\ast$ & $\ast$ & \cellcolor{lightgray} & & & & & & & & & & & & & \\
+%\color{orange} be annoyed\color{black} & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} & & & & & & & & & & & & \\
+%cleft &  $\ast$ & $\ast$ & $\ast$& $\ast$ & $\ast$ & $\ast$ & \cellcolor{lightgray} & & & & & & & & & & & \\
+%\color{orange} know\color{black}  & $\ast$ & $\ast$ & $\ast$&  $\ast$& $\ast$ & $\ast$ & -- & \cellcolor{lightgray} & & & & & & & & & & \\
+%\color{green} demonstrate\color{black}  &  $\ast$   &$\ast$  & $\ast$ & $\ast$ &  $\ast$& $\ast$ & $\ast$ & $\ast$ & \cellcolor{lightgray} & & & & &  & & & & \\
+%\color{green} pretend\color{black}  & $\ast$  &$\ast$  & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & -- & \cellcolor{lightgray} & & & &  & & & & \\
+%\color{green} inform\color{black}  & $\ast$  &$\ast$  & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray}& & &  & & & & \\
+%\color{green} confess\color{black}  &  $\ast$ &$\ast$  & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} & &  & & & & \\
+%\color{orange} see\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- &--  &-- & \cellcolor{lightgray} &  & & & & \\
+%\color{green} acknowledge\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & -- &--  &--  &--  &--  & \cellcolor{lightgray} & & & & \\
+%\color{orange} discover\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ &$\ast$  &$\ast$  &--  & -- &--  & -- &--  & -- & \cellcolor{lightgray} & & & \\
+%\color{green} admit\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ &$\ast$  &$\ast$  & $\ast$ &$\ast$  &-- & -- & -- & -- & \cellcolor{lightgray} & & \\
+%\color{green} hear\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  & $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  &-- & \cellcolor{lightgray} & \\
+%\color{green} prove\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  & -- &-- & $\ast$ &$\ast$  & -- &-- & \cellcolor{lightgray} \\
+%\color{orange} reveal\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ &$\ast$  &$\ast$  &$\ast$  &$\ast$  & $\ast$ &$\ast$  &$\ast$  & $\ast$ & -- & -- & -- & \cellcolor{lightgray}\\
+%\color{green} announce\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$&$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  & -- &$\ast$  &$\ast$  & \cellcolor{lightgray} \\
+%\color{green} be right\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & $\ast$ & -- & -- & \cellcolor{lightgray}\\
+%\color{green} establish\color{black}  &  $\ast$ & $\ast$ & $\ast$& $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} \\
+%\color{green} confirm\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$&$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  & $\ast$ &$\ast$  & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & \cellcolor{lightgray}\\
+%\color{green} suggest\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  &$\ast$  &$\ast$  &-- & \cellcolor{lightgray}\\
+%\color{green} think\color{black}   &  $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} \\
+%\color{green} say\color{black}   &  $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- \\
+%\bottomrule
+%\end{tabular}
+%\caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating.
+`*' indicates a difference in means, that is, the posterior distribution does not include 0, `--' indicates no difference in means, that is, the posterior distribution includes 0.}\label{t:pairwise}
+%\end{table}
 
 \begin{table}[!h]
-\begin{center}
-\begin{tabular}{r ccccccccccccccccccccccccc}
-\toprule
- &  \rot{continue} & \rot{too} & \rot{also} & \rot{again} & \rot{stop} & \rot{be annoyed} & \rot{cleft} & \rot{know} & \rot{demonstrate} & \rot{pretend} & \rot{inform} & \rot{confess} & \rot{see} & \rot{acknowledge} & \rot{discover} & \rot{admit} & \rot{hear} & \rot{prove} & \rot{reveal} & \rot{announce} & \rot{be right} & \rot{establish} & \rot{confirm} & \rot{suggest} & \rot{think} \\
- \midrule
- too & & & & & & & & & & & & & & & & & & \\
-also & & & & & & & & & & & & & & & & & & \\
-again  & & & & & & & & & & & & & & & & & & \\
-stop  & & & & & & & & & & & & & & & & & & \\
-be annoyed & & & & & & & & & & & & & & & & & & \\
-cleft & & & & & & & & & & & & & & & & & & \\
-know & & & & & & & & & & & & & & & & & & \\
-demonstrate & & & & & & & & & & & & & & & & & & \\
-pretend & & & & & & & & & & & & & & & & & & \\
-inform & & & & & & & & & & & & & & & & & & \\
-confess & & & & & & & & & & & & & & & & & & \\
-see & & & & & & & & & & & & & & & & & & \\
-acknowledge & & & & & & & & & & & & & & & & & & \\
-discover & & & & & & & & & & & & & & & & & & \\
-admit & & & & & & & & & & & & & & & & & & \\
-hear & & & & & & & & & & & & & & & & & & \\
-prove & & & & & & & & & & & & & & & & & & \\
-reveal & & & & & & & & & & & & & & & & & & \\
-announce & & & & & & & & & & & & & & & & & & \\
-be right & & & & & & & & & & & & & & & & & & \\
-establish & & & & & & & & & & & & & & & & & & \\
-confirm & & & & & & & & & & & & & & & & & & \\
-suggest & & & & & & & & & & & & & & & & & & \\
-say  & & & & & & & & & & & & & & & & & & \\
-\bottomrule
+\addtolength{\tabcolsep}{-.19em}
+\centering
+\begin{tabular}{r | cccccccccccccccccccccccccc}
+& \rots{continue} & \rots{too} & \rots{also} & \rots{again} & \rots{stop} & \rots{\color{orange}be annoyed\color{black}} & \rots{cleft} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
+\hline
+\input{../../results/main/13explicitIgnorance/models/latex-tables/table1}
+\hline
 \end{tabular}
-\caption{Posterior distributions of estimated marginal means of pairwise differences `***' indicates significance at .001, `**' at .01, `*' at .05, `.' marginal significance at .1, and \emph{n.s} indicates no significant difference in means.}\label{t:pairwise}
-\end{center}
+\caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating in the same context.
+Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
+Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
+Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
+\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
+}\label{t:pairwise}
 \end{table}
 
-Contrary to what is assumed in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021}, presuppositions are not invariably unnatural in explicit ignorance contexts.
+Contrary to what is assumed in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021,kalomoiros-schwarz-JoS}, presuppositions are not invariably unnatural in explicit ignorance contexts.
 Rather, in line with what was assumed in \citealt{simons01} and \citealt{abusch10}, some presuppositions seem to be rated as more unnatural in such contexts than others.
-On the assumption that naturalness ratings in explicit ignorance contexts differentiate between undefeasible/hard and defeasible/soft triggers, one might take the results of our experiment to suggest that the pre-state contents of {\em continue} and {\em again}, and the existence requirements of {\em too} and {\em also} are not defeasible.
+On the assumption that naturalness ratings in explicit ignorance contexts differentiate between undefeasible/hard presuppositions and defeasible/soft presuppositions, one might take the results of our experiment to suggest that the pre-state contents of {\em continue} and {\em again}, and the existence requirements of {\em too} and {\em also} are not defeasible.
 The CCs of {\em see, discover}, and {\em reveal}, on the other hand, might be taken to be defeasible.
-Contrary to what \citealt{abusch10} assumed, the existence requirement of the {\em it-}cleft is not as clearly undefeasible as that of other projective contents (\citealt{smith-hall11} also observed that the {\em it-}cleft did not pattern like a hard trigger in their experiment.) The results of our experiment align with our observation about possible by-content variation in \citepos{mandelkern-etal2020} Exp.~3.
+Contrary to what \citealt{abusch10} assumed, the existence requirement of the {\em it-}cleft is not as clearly undefeasible as that of other projective contents (\citealt{smith-hall11} also observed that the {\em it-}cleft did not pattern like a hard trigger in their experiment).
+Finally, the results of our experiment align with our observation about possible by-content variation in \citepos{mandelkern-etal2020} Exp.~3.
 Recall from above that in their Exp.~3, the pre-state of {\em continue} was judged as (numerically) less natural in explicit ignorance contexts than the CC of the cognitive predicate {\em be aware}.
 This result is in line with the result of our experiment that the pre-state of {\em continue} is judged as less natural than the CC of the cognitive predicate {\em know}.
+
+In our experiment, the CC of {\em be annoyed} received lower naturalness ratings than the CC of {\em know}.
+This result leads to the undesirable conclusion that the CC of {\em be annoyed} is more plausibly analyzed as a presupposition (perhaps even a hard one) than the CC of {\em know}, regardless of whether naturalness ratings in explicit ignorance contexts are taken to distinguish hard and soft presuppositions (\citealt{simons01,abusch10}) or presuppositions and nonpresuppositions (\citealt{mandelkern-etal2020,kalomoiros-schwarz2021,kalomoiros-schwarz-JoS}).
+This conclusion is undesirable because the presupposition of emotive predicates like {\em be annoyed} is that the attitude holder believes the CC to be true (e.g., \citealt{heim92,karttunen2016}), not that the speaker believes the CC to be true, as with {\em know}.
+In (\ref{emo}), for instance, the speaker does not believe that Sally did not leave a tip.
+\citealt{karttunen2016} suggested that the inference to speaker belief is a generalized conversational implicature that comes about in contexts that support the ``default assumption'' that the beliefs of the speaker and the attitude holder are aligned (p.712).
+
+\begin{exe}
+\ex\label{emo} Sally misremembered not having left a tip and regretted it. \hfill (\citealt[712]{karttunen2016})
+\end{exe}
+Given this status of the CC of {\em be annoyed}, it is surprising that this content received lower mean naturalness ratings than any of the other purportedly factive predicates.
+After all, in a context in which the speaker is explicitly ignorant of the CC, the default assumption that the speaker's and the attitude holder's beliefs are aligned does not arise, and so the interrogative with {\em be annoyed} should be judged natural in the explicit ignorance context, as it only presupposes that the attitude holder, not the speaker, believes the CC.
+
+One hypothesis for why the CC of {\em be annoyed} received such low ratings is that naturalness ratings in explicit ignorance context are sensitive to at-issueness.
+On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether the CC is true (?CC; e.g., {\em Does Julian dances salsa?}).
+The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?CC is true.
+If so, the naturalness of the interrogative is modulated by whether the QUD ?CC  together with the interpretation of the interrogative is a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is, if the interrogative can be interpreted as inquiring about the CC.
+As shown in \citealt{tbd-variability}, projective content differs in whether it can be at-issue.
+The CCs of emotive predicates, including {\em be annoyed}, are particularly resistant to being interpreted as at-issue.
+Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with the QUD ?CC (e.g., {\em Does Julian dance salsa?}).
+This incongruence may be why the CC of {\em be annoyed} received comparatively low naturalness ratings in an explicit ignorance context.
+
+For the CC of {\em know}, Fig.~\ref{fig:acc-by-expression} shows a bimodal pattern for the naturalness ratings, with about half of the participants rating it as fairly unnatural in the explicit ignorance context and about half as fairly natural.
+For the first group of participants, one could assume either that they reacted to the contradiction between the speaker's explicit ignorance statement and the inference arising from the interrogative that the speaker believes the CC to be true, or one could assume an analysis along the same lines as for {\em be annoyed} above.
+For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read both utterances with the rise-fall-rise melodies indicated by the ToBI transcription of the example,\footnote{Explain transcription, what is TOBI} that is, with rising pitch on each of the bold-faced expressions:
+
+\begin{exe}
+\ex\label{pros} \gll {\bf I} don't {\bf know} whether Julian dances {\bf salsa}. Does {\bf Cole} know that Julian dances {\bf salsa}? 
+\\ L*+H {} {L* H-} {} {} {} {L* H-H\%} {} L*+H {} {} {} {} {\hspace*{.2cm} L-H\%} \\ \glt 
+\end{exe}
+Analyses of the rise-fall-rise contour generally agree that it conveys that the speaker is uncertain about something and that their utterance does not fully answer the QUD (though the analyses differ in other regards; see, e.g., \citealt{ward-hirschberg85,buering97,buering03,wagner-etal2013}).
+In the context of the utterance of the declarative in (\ref{pros}), which conveys that the speaker takes the QUD to be whether Julian dances salsa, the rise-fall-rise contour that the interrogative utterance is realized with conveys that this utterance does not fully answer the QUD, which in turn requires that the speaker is uncertain about the CC.
+Under this interpretation, utterances like (\ref{pros}) may be rated as natural.
+
+In sum, naturalness ratings in explicit ignorance contexts do not support \citepos{mandelkern-etal2020} hypothesis that projection is a categorical property of content, or for a categorical distinction between presuppositions and nonpresupposition.
+Instead, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to whether the content is defeasible, to at-issueness, and to prosody.
 
@@
 
-As shown, the effect of context is not uniform for the factive predicates: For {\em be annoyed} and {\em know}, the mean naturalness rating is lower in the explicit ignorance context than in the higher prior probability context, the two ratings are equally high for {\em see, discover}, and {\em reveal}.
-This latter pattern is also observed for many of the nonfactive predicates, for instance {\em confess, hear, announce}, and {\em be right}.
-Thus, naturalness ratings in explicit ignorance vs.\ higher prior probability contexts do not provide support for a binary distinction between the CCs of factive and nonfactive predicates.
+As shown, the effect of context is not uniform for the factive predicates.
+For {\em be annoyed} and {\em know} the mean naturalness rating is quite a bit lower in the explicit ignorance context than in the higher prior probability context, whereas the difference is much less pronounced for {\em see} and {\em discover}, and not observed for {\em reveal}.
+This latter pattern is also observed for several nonfactive predicates, including {\em confess, hear, announce}, and {\em be right}.
+Thus, naturalness ratings in explicit ignorance vs.\ higher prior probability contexts do not provide support for a categorical distinction between the CCs of factive and nonfactive predicates.
 A comparison between the explicit ignorance and lower prior probability contexts also suggests variation between the factive predicates: Whereas for {\em be annoyed} the mean naturalness ratings are lower in the explicit ignorance context than in the lower prior probability context, they are equally high for {\em know}, but lower in the lower prior probability context than the explicit ignorance context for {\em see, discover}, and {\em reveal}.
-This latter pattern is observed for all of the nonfactive predicates, except possibly {\em pretend}.
-Thus, naturalness ratings in explicit ignorance vs.\ lower prior probability contexts also do not provide support for a binary distinction between the CCs of factive and nonfactive predicates.
+This latter pattern is observed for all of the nonfactive predicates.
+Thus, naturalness ratings in explicit ignorance vs.\ lower prior probability contexts also do not provide support for a categorical distinction between the CCs of factive and nonfactive predicates.
 
-Finally, we observe that the prior probability of content modulates its naturalness ratings for all predicates, except {\em demonstrate} and {\em pretend}): Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability.
+Finally, we observe that the prior probability of content modulates its naturalness ratings with all predicates, except {\em demonstrate} and {\em pretend}): Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability.
 This result is in line with the result of \citealt{degen-tonhauser-openmind} based on an inference rating task.
@@
 
-These results were confirmed by a model...{\bf BETA MODELS}
+These observations were confirmed by posthoc pairwise comparisons of the naturalness ratings in the three context for each of the 20 predicates using the `emmeans' package (\citealt{emmeans}).
+The input to the pairwise comparisons were 20 Bayesian mixed-effects beta regression models with weakly informative priors that were fitted using the `brms' package (\citealt{buerkner2017}).
+The models predicted, for each predicate, the naturalness ratings from a fixed effect of context (with treatment coding and `explicit ignorance' as reference level) and included random by-item intercepts.
+The pairwise comparisons provided us with posterior distributions of estimated marginal means of pairwise differences between the three contexts for each predicate.
+We assume that the naturalness ratings in two contexts differ when the posterior distribution of their difference does not include 0.
+See Supplement \ref{a:analysis1} for the full model output.
 
-\newpage
+ As shown in Table \ref{t:pairwise2},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.} the model confirms that the effect of context is not uniform for the five factive predicates: The difference between the naturalness ratings in the higher prior probability context and the explicit ignorance context is larger for {\em be annoyed} than for {\em know}, which in turn is larger than for {\em see} and {\em discover}.
+For {\em reveal} there is no difference, just like for many nonfactive predicates, such as {\em admit, hear, announce} and {\em be right}.
+The comparison of the lower prior probability context and the explicit ignorance context reveals a positive difference for {\em be annoyed}, no difference for {\em know} and a negative difference for {\em see, discover}, and {\em reveal}.
+Finally, there is a positive difference between the higher and lower prior probability context for all 20 predicates except for {\em demonstrate} and {\em pretend}.
 
-bla
+\begin{table}[!h]
+\addtolength{\tabcolsep}{-.19em}
+\centering
+\begin{tabular}{r  cccccccccccccccccccc}
+& \rots{\color{orange}be annoyed\color{black}} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
+\hline
+\input{../../results/main/13explicitIgnorance/models/latex-tables/table2}
+\hline
+\end{tabular}
+\caption{Posterior distribution of estimated marginal means of pairwise differences between the three contexts for each expression, with expressions ordered by mean naturalness rating in explicit ignorance context.
+Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
+Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
+Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
+\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
+}\label{t:pairwise2}
+\end{table}
 
-\newpage
+In contrast to \citepos{mandelkern-etal2020} Exp.~3, our experiment did not collect naturalness ratings in a support context that entails the relevant content but in higher and lower prior probability contexts that are compatible with the CC.
+As mentioned above, contemporary presupposition analyses (e.g., \citealt{heim83,vds92}) assume that presuppositions are by default globally accommodated when they are not entailed by or satisfied in the context, and they can be locally accommodated (e.g., under the polar question operator) to avoid contradiction, uninformativity or problems with binding.
+Such analyses lead us to expect that presuppositions are rated as natural in higher and lower prior probability contexts, as these contexts do not contradict the CCs, make the CCs uninformative, or lead to problems with binding.
+Contrary to this expectation, the CCs of the five purportedly factive predicates received lower ratings in the lower prior probability context than in the higher prior probability context, where they received ratings that are roughly comparable to those of the factive predicates investigated in \citealt{mandelkern-etal2020}.\footnote{In the support context, the CCs of {\em be aware} and {\em be happy} received a mean naturalness rating of about 4.4 and 4.7 out of 7 in \citepos{mandelkern-etal2020} Exp.~3 (based on Fig.~8 in their Appendix).
+In our higher prior probability context, the mean naturalness ratings for the CCs of the five factive predicates ranged from .64 ({\em reveal}) to .78 ({\em know}) on a scale from 0 to 1.} This result suggests that naturalness ratings are sensitive to the prior probability of content, such that an interrogative with a clause-embedding predicate is judged as more natural in a context in which the CC has a higher prior probability than in a context in which it has a lower prior probability.
+For contemporary analyses of presuppositions, the result means that local accommodation must be sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
 
-\section{Discussion}\label{s3}
+\section{General discussion}\label{s3}
 
-The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides empirical support for the hypothesis that projection is a binary property of content and distinguishes presuppositions from nonpresuppositions.
-Neither the by-content comparison of naturalness ratings in explicit ignorance contexts nor the by-context comparison of naturalness ratings provided empirical support for a binary distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
-In sum, the experiment did not find support for \citepos{mandelkern-etal2020} claim nor for the hypothesis that projection is a binary property of content.
-Instead, the experiment provided further support for the hypothesis of \citealt{tbd-variability} that projection is a gradient property of content and the result of \citealt{degen-tonhauser-language} that there is no well-defined class of factive predicates.
+The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides support for a categorical distinction between presuppositions and nonpresuppositions.
+Neither the by-content comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the by-context comparison in \S\ref{s:analysis2} provided empirical support for a categorical distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
+Rather, the results of the experiment provided further support for the result of \citealt{degen-tonhauser-language} that factive predicates are not a natural class that is categorically different from nonfactive predicates.
 In this section, we discuss theoretical and methodological implications of these results.
@@
 
-Contemporary literature still frequently treats presuppositions as if they were a homogenous class, for instance by characterizing them as contents that typically project or as felicity requirements.
-Research on projective content has long recognized that the set of contents that have been called presuppositions is heterogeneous.
-Two dimensions of variation that were already mentioned in \S\ref{s1}, namely defeasibility and strength of projection, fly in the face of this homogeneity assumption.
-A third dimension of variation is whether the presupposition is associated with what \citealt{brst-lang11} referred to as a `strong contextual felicity' constraint: Some presumed presuppositions are not associated with such a constraint, which means that they are judged to be acceptable in a context that is neutral with respect to the content (like the CC of {\em know}, which is informative in such a context), while others are associated with such a constraint, that is, judged to be unacceptable in a context that does not entail or satisfy the content (like the existence requirement of pronouns or {\em too}).
-A fourth dimension of variation is at-issue, that is, whether the content can address the Question Under Discussion: As shown in \citealt{tbd-variability}, the CCs of factive predicates vary in this regard, as the CC of {\em discover} is more at-issue than those of  {\em be annoyed} and {\em know}.
+Contemporary literature still frequently treats presuppositions as if they were a homogenous class, by characterizing them as contents that typically project out of entailment-canceling environments or by analyzing them as conventionally-coded felicity requirements (e.g., \citealt{mandelkern-etal2020}).
+However, research on projective content has long recognized that the set of contents that have been called presuppositions is heterogeneous.
+One dimension of variation that was already mentioned in \S\ref{s1} is projection.
+Experimental investigations in different languages suggest that contents traditionally called presuppositions exhibit variation in the strength of the inference to the content (e.g., \citealt{xue-onea11} on German, \citealt{tonhauser-guarani-variability} on Paraguayan Guaran\'i, and \citealt{demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language} on English).
+In English, for instance, the CC of {\em discover} is less projective than the CC of {\em know}.
+These results challenge the traditional characterization of presuppositions on the basis of projection.
 
-Given that the set of contents called presuppositions is heterogeneous, it is not entirely surprising that these contents do not pattern alike in explicit ignorance contexts.
-For instance, a content that is associated with a strong contextual felicity constraint, like the existence requirement of {\em too}, might be expected to be judged as unnatural in a context in which the speaker is explicitly ignorant about the content.
-On the other hand, a content that is not associated with a strong contextual felicity constraint and that can be at-issue, like the CC of {\em discover}, might be expected to be judged as somewhat natural in a context in which the speaker is explicitly ignorant about the content, as the content can be interpreted as what the speaker is inquiring about.
-Our experiment suggests that both of these expectations were borne out by the data.
+A second dimension of variation is whether the content is associated with what \citealt{brst-lang11} referred to as a `strong contextual felicity' constraint.
+Some presumed presuppositions are not associated with such a constraint, which means that they are judged to be acceptable in a context that is neutral with respect to the content.
+This is the case for the CC of {\em know}, illustrated in (\ref{scf}a), which is informative in such a context.
+Other presumed presuppositions are associated with such a constraint, that is, they are judged to be unacceptable in a context that does not entail or satisfy the content.
+This is the case for the existence requirement of {\em too}, illustrated in (\ref{scf}b).
 
 \begin{exe}
-\ex
+\ex\label{scf} \citealt[78, 80]{brst-lang11}
 \begin{xlist}
-\ex If Julian dances salsa, then Cole knows that he does.
-\ex If Julian dances salsa, then 
+\ex Context: A girl backs out of a driveway and hits Susi's car.
+A woman comes running out of the house, apologizes that her daughter hit Susi's car, and says: \\ She knows that she has to use her glasses to drive.
+
+\ex Context: Malena is eating her lunch, a hamburger, on the bus going into town. A woman who she doesn't know sits down next to her and says:
+\\ \infelic Our bus driver is eating empanadas, too.
+
 \end{xlist}
 \end{exe}
+This variation suggests that analyses of presuppositions as conventionally-coded felicity requirements generalize an analysis that is empirically justified for contents associated with a strong contextual felicity constraint to contents that are not associated with such a constraint.
 
-Against this background, it does not seem fruitful to us to
+A third dimension of variation is at-issue, that is, whether the content can address the question under discussion: As shown in \citealt{tbd-variability}, contents traditionally called presuppositions vary in this regard: For instance, the pre-state content of {\em stop} and the CC of {\em discover} are more at-issue than the CCs of {\em know} and {\em be annoyed}.
+And, finally, there is the dimension investigated in our experiment, which was referred to as defeasibility by \citealt{simons01} and \citealt{abusch10}.
+While the variation observed in our experiment does not line up perfectly with the variation assumed in these works, the results of our experiment suggest that contents traditionally called presuppositions exhibit variation in how natural they are in explicit ignorance contexts.
+For instance,  the pre-state content of {\em continue} was rated as less natural than the pre-state content of {\em stop}, which was rated as less natural than the existential content of the {\em it-}cleft, which was rated as less natural than the CC of {\em see}, which was rated as less natural than the CC of {\em reveal}.
+
+These dimensions of variation show that the set of contents traditionally referred to as presuppositions does not constitute a natural class.
+It is in response to such variation that researchers have developed analyses of subsets of these contents that do not rely on a conventional specification of felicity conditions or assume uniform projection (e.g., \citealt{abrusan2011,abrusan2013,abrusan2016,abusch02,abusch10,romoli2015}).
+Consequently, research on projective content can no longer take for granted a characterization of ``presuppositions''  based on projection or an analysis as conventionally-specified felicity conditions.
+
+\subsection{Are there factive predicates?}
+
+As discussed in \citealt{degen-tonhauser-language}, factive predicates are variably defined as predicates whose CC is presupposed (e.g., \citealt{kiparsky-kiparsky70}) or as predicates whose CC is presupposed and entailed (e.g., \citealt{gazdar79a,schlenker10,abrusan2011}).
+\citealt{degen-tonhauser-language} did not, however, find empirical support for a natural class of `factive' predicates that is categorically distinct from nonfactive ones.
+First, based on the results of two experiments designed to investigate the projection of the CCs of (non)factive predicates as well as the analyses of three existing projection datasets for 45, 78, and 517 (non)factive predicates, respectively,\footnote{These datasets were the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality dataset (\citealt{white-rawlins-nels2018}).} this work suggested that projection does not provide support for a natural class of factive predicates that is categorically distinct from nonfactive ones.
+Second, based on four experiments designed to investigate the entailment of the CCs of (non)factive predicates as well as analyses of entailment in two of the aforementioned datasets, this work suggested that considering entailment in addition to presuppositionality also does not provide support for a natural class of factive predicates that is categorically distinct from nonfactive ones.
+
+\citealt{degen-tonhauser-language} used a variety of inference rating measures to investigate the projection of the CCs of (non)factive predicates: In some measures, participants read constructed examples, in others naturally occurring ones; in some, examples were presented with a context, in others without a context; in some, the examples had diverse content, in others, they had minimal lexical content; across the measures, the clause-embedding predicates were embedded under different entailment-canceling operators, participants were asked to respond to different questions, and participants gave their response on different scales and sliders.
+Despite this variation, the results invariably suggested that factive predicates are not categorically distinct from nonfactive ones.
+
+The experiment presented in this paper presents support for the claim that there is no natural class of factive predicates based on a different measure, namely naturalness ratings in explicit ignorance contexts.
+The CCs of the five purportedly factive predicates exhibited variation in how natural they were judged to be when the speaker was explicitly ignorant of the CC, and there were no differences between the CCs of {\em see, discover} and {\em reveal} and several nonfactive predicates.
+Thus, neither data based on inference rating measures nor data based on naturalness ratings in explicit ignorance contexts provide support for a natural class of factive predicates.
+As discussed in \citealt{degen-tonhauser-language}, ``the search for a class of factive predicates might continue'' (p.586), but we believe that it is more fruitful to investigate the more fine-grained distinctions between clause-embedding predicates that can lead to a better understanding of their use in discourse.
+We offer some suggestions in \S\ref{s:disc3} based on the naturalness ratings we collected.
+
+\subsection{Is projection a categorical or a gradient property of content?}
+
+\citealt[497]{mandelkern-etal2020} urged us to ``think twice before embracing a notion of [\ldots] projection that is gradient based on results from inference tasks alone'', advocating instead for a view of projection that is categorical.
+In this section, we return to the question of whether projection is a categorical or a gradient property of content.
+In addressing this question, it is important to distinguish between the projection of {\em content} and the projection of {\em utterance content}.
+The latter concerns the strength of the projection inference that a particular interpreter draws from a particular utterance of a sentence in which a content is contributed in the scope of an entailment-canceling operator.
+The empirical evidence pointed to in \S\ref{s1} shows that different interpreters draw projection inferences of varying strength to, for instance, the CC that Julian dances salsa in utterances of {\em Did Cole discover that Julian dances salsa?} In other words, it is uncontroversial that projection is a gradient property {\em utterance content}, not a categorical one.
+It is gradient because the projection of utterance content is sensitive to several factors on which  interpreters differ, including interpreters' prior beliefs about the speaker or about the utterance content, interpreters' perception and interpretation of the prosodic realization of the utterances, and interpreters' assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Also: noise, inattentiveness, but replication of variation exclude this.}
+
+What is an open question, however, is whether projection is a gradient or a categorical property of {\em content}, where a content is, for instance, the CC of {\em discover} or the pre-state content of {\em stop}.
+This question amounts to asking whether the observed projection variability (that is attributable at least to some extent to interpreters) is derived from underlying representations that are categorical or underlying representations that are not categorical.
+Two aspects of the underlying representations we consider here are speaker belief and lexical entries.
+A speaker who wants to communicate a content could have a belief in the content that is categorical (that is, they either believe that is is true or they are uncertain) or or that is gradient (in which case they believe the truth to some degree).
+Likewise, the expressions associated with the content (like {\em discover} or {\em stop}) could have lexical entries that either license a projection inference or not (categorical) or their lexical entries are such that whatever is specified may license a projection inference depending on other factors in the uttered sentence (gradient).
+This means that there are four hypotheses for any given content:
+
+\begin{table}[h]
+\centering
+\begin{tabular}{c | c c}
+Hypothesis & speaker & lexicon  \\ \hline\hline
+1 & categorical & categorical  \\ 
+2 & categorical & {\bf gradient}  \\ 
+3 & {\bf gradient} & categorical \\ 
+4 & {\bf gradient} & {\bf gradient}  \\ 
+\hline
+\end{tabular}
+\caption{Hypotheses about speaker belief and lexical entries for any given content.}\label{t:hyp}
+\end{table}
+
+Some of this hypothesis space has already been advocated for in the literature.
+Analyses like \citealt{heim83,vds92} and \citealt{potts05} align with hypotheses 1 and 3, as they assume that some expression (triggers of presuppositions or conventional implicatures) are conventionally specified to trigger projection.
+While these analyses do not specify whether speaker belief is categorical or gradient, one could assume, on the former view (hypothesis 1), that a speaker who believes that a content is true chooses an expression that triggers projection, whereas a speaker who is uncertain about a content chooses an expression that does not trigger projection.
+On the latter view (hypothesis 3), a speaker believes content to be true to some degree and chooses an expression that either triggers projection or not, depending on what comes closest to their degree of belief.
+On this view, interpreters might draw projection inferences of different strengths due to uncertainty about the degree of the speaker's belief.
+
+Analyses like those of \citealt{abrusan2011,abrusan2016,abusch02,abusch10,romoli2015} and \citealt{brst-salt10,simons-etal2017} could be taken to align with hypotheses 2 and 4.
+On some of these analyses, some expressions specify alternatives as part of their lexical entries and this specification of alternatives may, depending on other factors, give lead to a projection inference (e.g., \citealt{abusch02,abusch10,romoli2015}).
+On other analyses, projection inferences arise from entailments in particular utterances (e.g., \citealt{abrusan2011,abrusan2016,brst-salt10,simons-etal2017}).
+Crucially, while these analyses differ in many ways, they all share the assumption that the lexical entries of expressions do not categorically trigger (or not trigger) projection.
+Rather, whether a projection inference is licensed depends on a combination of the lexical entries and other utterance factors.
+
+{\bf MOVE EARLIER: Any of these analyses must be evaluated against the observed projection variability, of which there are two types, namely variability within a given content (e.g., the CC of {\em discover}) as well as variability between different contents (e.g., comparing the CC of {\em discover} and the CC of {\em know}).}
+
+While projection variability is solely due to utterance interpretation on hypothesis 1, it is due to both utterance interpretation and to speaker belief being gradient on hypothesis 3. 
+On these analyses, then, the observed projection variability...
+
+\citealt[498f.]{tbd-variability} suggested that projection variability could be due to speaker belief being gradient, as a consequence of which an interpreter's  projection inference is gradient (hypotheses 1, 3, 5, 7).
+This position still allows for lexical representations being categorical or gradient, and utterance interpretation introducing variability or not.
+A second position entertained in \citealt[498f.]{tbd-variability} is that speaker belief is categorical and that projection variability ``arises from the [interpreter's] uncertainty about whether the speaker is committed'', that is, as part of the utterance interpretation (hypotheses 2, 6).
+
+p.518: the projectivity of projective content is also influenced by the expression that the projective content is associated with
+
+Recall that \citealt[498]{tbd-variability} proposed that projection is a gradient property of content rather than a categorical one after observing variation in the projection of content in their experiments.
+Above, we suggested that the results of \citealt{degen-tonhauser-language}, some of which are shown in Fig.~\ref{fig:dt1a}, provide further support for this proposal.
+\citealt{mandelkern-etal2020}, on the other hand, warned us against ``embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).
+Contrary to what \citealt{mandelkern-etal2020} led us to expect, the results of our experiment do not provide empirical support for their claim that (presupposition) projection is categorical.
+
+What would constitute empirical evidence that projection is categorical?
+One type of evidence would be the results of behavioral experiments with a broad range of contents that suggest a nonarbitrary division of contents into ones that typically project and ones that typically don't.
+Such results does not currently exist, as noted in \citealt{degen-tonhauser-language} and this paper.
+Another type of evidence could be provided by computational models that evaluate whether the data collected in behavioral experiments is better captured by analyses under which projection is categorical or analyses under which it is gradient.
 
 \begin{itemize}
-
-\item When \citealt{mandelkern-etal2020} talk about presuppositions, it's not even clear what they mean: +SCF or -SCF?
-They only investigated -SCF/+OLE expression but assumed an analysis for them that was developed for +SCF triggers.
-
-\end{itemize}
-
-{\bf need to discuss in the results what naturalness in EIC appears to be sensitive to, and also bimodal for know and low responses for be annoyed}
-
-\subsection{Is projection a binary or a gradient property of content?}
-
-\begin{itemize}
-
-\item We need to identify which level we're talking about
-
-\begin{itemize}
-
-\item Speaker belief:
-
-Option 1: Speaker belief is binary, gradience is introduced by interpretation.
-
-Option 2: Speaker belief is gradient.
-
-\item Lexical representation:
-
-Option 1: Representation codes projection
-
-Option 2: Representation does not code projection but projection is possible in particular utterances
-
-\end{itemize}
 
@@
 
-\item To address this question, we have to distinguish whether we're talking about content associated with a particular expression (aggregated projection over multiple utterances) versus utterance content
-
-\item The characterization of presuppositions is at the level of content, but has to involve `typically' to capture that a presupposition in some utterances does not project.
-
-\item We also have to distinguish the speaker and the hearer: The speaker may be binary or gradient, and even if binary, interpretation may vary because of different assumptions about speaker, different integration of other factors.
-
-\item Projection inferences, sources of gradience
-
-\begin{tabular}{l c c c c} \\ 
-             & expression & utterance & speaker & interpreter \\\hline
-class 1 &  binary & binary & binary & binary \\
-class 2 &  binary & & & \\ 
-class 3 & & & & \\
-
-\end{tabular}
-
-\item What does gradient projection mean?
-
-\begin{itemize}
-
-\item[{[H1]}] Projection is binary at the level of the expression: An expression either triggers a presupposition or not.
-The observed gradience at the utterance level is due to utterance-level factors that modulate projection.
-
-\item[{[H2]}] Projection is not binary at the level of the expression: The lexical representations of (classes of) expressions give rise to projection inferences of varying strength.
-The observed gradience at the utterance level is due to the lexical variation in projection strength as well as utterance-level factors that modulate projection.
-
-\end{itemize}
-
 \item Contemporary presupposition analyses (e.g., \citealt{heim83,vds92,abrusan2011,karttunen2016,simons-etal2017}) predict neither the heterogeneity of the naturalness ratings of factive predicates nor their sensitivity to the prior probability of the CC (though see \citealt{schlenker2021} for the latter).
@@
 
-\subsection{Which measures are suitable to investigate projection?}
+\subsection{Future avenues in the investigation and theoretical analysis of projective content}\label{s:disc3}
+
+What are naturalness ratings sensitive to?
+
+- strong contextual felicity constraint
+- at-issueness
+- prior probability of content
 
@@
 
-\item And while naturalness ratings in EICs may provide insight into the meanings of attitude predicates, the interpretation of the results depends on the presumed linking function (see, e.g., \citealt{sprouse2018}).
-Instead of the ratings directly reflecting whether the CC is presupposed, the very low mean rating for {\em be annoyed} in EICs could be due to its CC being not-at-issue while the EIC suggests that it is at-issue; likewise, reading {\em know} with a contrastive pitch accent in EICs may have resulted in about half of the participants giving a high rating.
+\item Lots of interesting properties of projective content to investigate
+
+\item What do the naturalness ratings tell us about the lexical meanings of these predicates?
 
@@
 
-\item The talk also discusses that the results do not align with assumptions about hard and soft triggers; \citealt{simons01,abusch10}.
-
 \end{itemize}
 
-\subsection{What does an empirically adequate analysis of projection need to capture?}
+measure doesn't just track anaphoric presuppositions but possibly also mismatches in at-issueness
+
+does the measure track at-issueness or projection?
 
@@
 
+{\bf JT PROPOSES TO LEAVE OUT THIS AND THE NEXT FIGURE}
+
+Fig.~\ref{fig:comparisons1} provides a graphical representation of the results of the pairwise comparisons between the ratings for each expression in the explicit ignorance context.
+The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.
+
 \begin{sidewaysfigure}[h!]
@@
 
+Fig.~\ref{fig:comparisons2} provides a graphical representation of the results of the pairwise comparisons of the ratings in the three contexts for each of the 20 predicates.
+The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.
+
 \begin{figure}[h!]

### 2023-11-16 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -20,6 +20,7 @@ library(lmerTest)
 library(padr)
 library(performance)
 library(MuMIn)
+library(xtable)
 
 # load cleaned data
 d = read_tsv("../data/cd.tsv")
@@ -31,12 +32,10 @@ t = d %>%
   filter(context == "explicitIgnorance") %>%
   filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4")
   
-# set reference levels
+# set reference level
 t = t %>%
   mutate(expression = fct_relevel(expression, "continue"))
-
-table(t$expression)
-table(t$context)
+levels(t$expression)
 
 # first, because response assumes values of 0 and 1, which beta regression cannot handle, transform: (Smithson & Verkuilen 2006)
 # y'' = (y' ?? (n ??? 1) + 0.5)/n
@@ -71,10 +70,38 @@ pairwise <- m.b %>%
   mean_hdi() %>%
   # create new column of "first" element in contrast
   mutate(first = gsub(" -.*", "", contrast)) %>%
+  mutate(second = gsub(".* -", "", contrast)) %>%
   # sort by mean value
   mutate(contrast = fct_reorder(as.factor(contrast),.value))
 pairwise
+
+saveRDS(pairwise,file="../models/pairwise1.csv")
+pairwise = readRDS(file="../models/pairwise1.csv")
+pairwise
+names(pairwise)
+
+# select relevant columns for printing
+pairwise_reduced = pairwise %>%
+  select(c(contrast, .value, .lower, .upper))
+
+#### table of model output in Appendix ----
+
+tableApp1 = print(xtable(pairwise_reduced),
+               #only.contents = T,
+               include.rownames=FALSE,
+               include.colnames=TRUE,
+               tabular.environment="longtable",
+               floating=FALSE,
+               hline.after = NULL,
+               latex.environments=NULL,
+               booktabs=TRUE,
+               sanitize.text.function = function(x){x},
+               comment = F
+)
+
+write(tableApp1, "../models/latex-tables/tableApp1.tex")
        
+#### plot of model output in Appendix ----
 theme_set(theme_bw())   
 ggplot(pairwise, aes(x = .value, y = contrast)) +
   geom_point() +
@@ -82,7 +109,190 @@ ggplot(pairwise, aes(x = .value, y = contrast)) +
   geom_vline(xintercept = 0, color = "red") +
[…]

### 2023-11-16 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -28,7 +28,71 @@ table(d$context)
 
 length(unique(d$participantID)) #370 participants
 
-# Fig 1: plot of mean naturalness ratings in explicit ignorance context ----
+# Fig 1: plot of mean certainty ratings from Exp 1a of Degen & Tonhauser 2022 -----
+# import data from repo
+cd <- read_csv("https://raw.githubusercontent.com/judith-tonhauser/projective-probability/master/results/5-projectivity-no-fact/data/cd.csv")
+summary(cd)
+
+# mean projectivity by predicate, including the main clause controls
+means = cd %>%
+  group_by(verb) %>%
+  summarize(Mean = mean(response), CILow = ci.low(response), CIHigh = ci.high(response)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, verb = fct_reorder(as.factor(verb),Mean))
+means
+
+# define colors for the predicates
+cols = data.frame(V=levels(means$verb))
+
+cols$VeridicalityGroup = as.factor(
+  ifelse(cols$V %in% c("know", "discover", "reveal", "see", "be annoyed"), "F", 
+         ifelse(cols$V %in% c("MC"),"MC","NF")))
+
+levels(cols$V)
+cols$V <- factor(cols$V, levels = cols[order(as.character(means$verb)),]$V, ordered = TRUE)
+
+cols$Colors =  ifelse(cols$VeridicalityGroup == "F", "#D55E00",
+                      ifelse(cols$VeridicalityGroup == "NF", "#009E73",'black'))
+
+
+cols$V <- factor(cols$V, levels = cols[order(as.character(means$verb)),]$V, ordered = TRUE)
+levels(cols$V)
+
+means$VeridicalityGroup = factor(x=
+                                   ifelse(means$verb %in% c("know", "discover", "reveal", "see", "be annoyed"), "F", 
+                                          ifelse(means$verb  %in% c("MC"),"MC","NF")),levels=rev(c("F","NF","MC")))
+
+subjmeans = cd %>%
+  group_by(verb,workerid) %>%
+  summarize(Mean = mean(response)) 
+subjmeans$verb <- factor(subjmeans$verb, levels = unique(levels(means$verb)))
+subjmeans$VeridicalityGroup = factor(x=
+                                   ifelse(subjmeans$verb %in% c("know", "discover", "reveal", "see", "be annoyed"), "F", 
[…]

### 2023-11-16 writing/naturalness-paper/paper.tex
updated naturalness paper

@@
 \citealt[497]{mandelkern-etal2020} urged us to ``think twice before embracing a notion of [\ldots] projection that is gradient based on results from inference tasks alone'', advocating instead for a view of projection that is categorical.
-In this section, we return to the question of whether projection is a categorical or a gradient property of content.
-In addressing this question, it is important to distinguish between the projection of {\em content} and the projection of {\em utterance content}.
-The latter concerns the strength of the projection inference that a particular interpreter draws from a particular utterance of a sentence in which a content is contributed in the scope of an entailment-canceling operator.
-The empirical evidence pointed to in \S\ref{s1} shows that different interpreters draw projection inferences of varying strength to, for instance, the CC that Julian dances salsa in utterances of {\em Did Cole discover that Julian dances salsa?} In other words, it is uncontroversial that projection is a gradient property {\em utterance content}, not a categorical one.
-It is gradient because the projection of utterance content is sensitive to several factors on which  interpreters differ, including interpreters' prior beliefs about the speaker or about the utterance content, interpreters' perception and interpretation of the prosodic realization of the utterances, and interpreters' assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Also: noise, inattentiveness, but replication of variation exclude this.}
+In this section, we turn to the question of whether projection is a categorical or a gradient property of content.
+Before addressing the question, we first consider what it is (not) about.
 
-What is an open question, however, is whether projection is a gradient or a categorical property of {\em content}, where a content is, for instance, the CC of {\em discover} or the pre-state content of {\em stop}.
-This question amounts to asking whether the observed projection variability (that is attributable at least to some extent to interpreters) is derived from underlying representations that are categorical or underlying representations that are not categorical.
+What the question is not about is the projection of {\em utterance content}.
+What we mean by that is the strength of the projection inference that a particular interpreter draws from a particular utterance of a sentence in which a content is contributed in the scope of an entailment-canceling operator.
+The empirical evidence pointed to in \S\ref{s1} shows that different interpreters draw projection inferences of varying strength.
+For instance, as shown in Fig.~\ref{fig:dt1a-JULIAN}, the 266 participants of \citepos{degen-tonhauser-language} Exp.~1a drew projection inferences of varying strength for the CC that Julian dances salsa from utterances of the interrogative {\em Did Scott discover that Julian dances salsa?} It is uncontroversial that this utterance content variability is gradient, not categorical.
+It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Also: noise, inattentiveness, but replication of variation exclude this.}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/mean-certainty-by-predicateType-JULIAN}
+\caption{Participants' certainty ratings (measuring projection) of the CC that Julian dances salsa for the 20 \color{orange}factive \color{black} and \color{green}nonfactive \color{black} predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category).}\label{fig:dt1a-JULIAN}
+\end{figure}
+
+What the question about whether projection is gradient or categorical is about is {\em content}, where a content is, for instance, the CC of {\em discover}, the CC of {\em inform}, or the pre-state content of {\em stop}.
+Asking whether projection is gradient or categorical at this level of abstraction (that is, abstracting away from variation attributable to interpreter variability) amounts to asking whether the observed by-content variability (for instance that the CC of {\em discover} is less projective than the CC of {\em inform}; see Fig.~\ref{fig:dt1a}) is derived from underlying representations that are categorical or underlying representations that are not categorical.
 Two aspects of the underlying representations we consider here are speaker belief and lexical entries.
-A speaker who wants to communicate a content could have a belief in the content that is categorical (that is, they either believe that is is true or they are uncertain) or or that is gradient (in which case they believe the truth to some degree).
-Likewise, the expressions associated with the content (like {\em discover} or {\em stop}) could have lexical entries that either license a projection inference or not (categorical) or their lexical entries are such that whatever is specified may license a projection inference depending on other factors in the uttered sentence (gradient).
+A speaker who wants to communicate a content might have a belief in the content that is categorical (that is, they either believe that is is true or they are uncertain) or one that is gradient (in which case they believe that the content is true to some degree).\footnote{\citealt[498f.]{tbd-variability} already considered that speaker belief could be gradient or categorical.} Likewise, the expressions associated with the content (like {\em discover} or {\em stop}) could have lexical entries that either license a projection inference or not (categorical) or their lexical entries could be such that whatever is specified may license a projection inference depending on other factors in the uttered sentence (gradient).
 This means that there are four hypotheses for any given content:
@@
 On the latter view (hypothesis 3), a speaker believes content to be true to some degree and chooses an expression that either triggers projection or not, depending on what comes closest to their degree of belief.
-On this view, interpreters might draw projection inferences of different strengths due to uncertainty about the degree of the speaker's belief.
+On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.
 
 Analyses like those of \citealt{abrusan2011,abrusan2016,abusch02,abusch10,romoli2015} and \citealt{brst-salt10,simons-etal2017} could be taken to align with hypotheses 2 and 4.
-On some of these analyses, some expressions specify alternatives as part of their lexical entries and this specification of alternatives may, depending on other factors, give lead to a projection inference (e.g., \citealt{abusch02,abusch10,romoli2015}).
-On other analyses, projection inferences arise from entailments in particular utterances (e.g., \citealt{abrusan2011,abrusan2016,brst-salt10,simons-etal2017}).
+On some of these analyses, some expressions specify alternatives as part of their lexical entries and this specification of alternatives can, depending on other factors, give lead to a projection inference (e.g., \citealt{abusch02,abusch10,romoli2015}).
+On other analyses, entailed content may project in particular utterances  (e.g., \citealt{abrusan2011,abrusan2016,brst-salt10,simons-etal2017}).
 Crucially, while these analyses differ in many ways, they all share the assumption that the lexical entries of expressions do not categorically trigger (or not trigger) projection.
 Rather, whether a projection inference is licensed depends on a combination of the lexical entries and other utterance factors.
+These analyses, too, don't specify whether speaker belief is categorical or gradient.
 
-{\bf MOVE EARLIER: Any of these analyses must be evaluated against the observed projection variability, of which there are two types, namely variability within a given content (e.g., the CC of {\em discover}) as well as variability between different contents (e.g., comparing the CC of {\em discover} and the CC of {\em know}).}
+To evaluate which of the four hypotheses in Table \ref{t:hyp} is appropriate for any given content, we have to consider the observed by-content variability.
+An extreme position is to assume for every projective content that speaker belief is categorical and triggered by conventional specification.
+To capture the observed by-content variability, one would then have to assume that it arises during utterance interpretation and affects different expression/content pairs differently.
+That does not seem plausible.
+A more plausible position seems to us to be one on which each of the four hypotheses is correct but for different sets of expression/content pairs.
+For instance, hypothesis 1 might be appropriate  the existence requirement of pronouns and {\em too} (expression/content pairs associated with a strong contextual felicity requirement), hypothesis 2 for the pre-state of {\em stop}, and hypothesis 4 for the CC of {\em inform}.
+The observed by-content variability would then be attributable to a combination of factors, with expression/content pairs differing on the factors and their relative weights.
 
-While projection variability is solely due to utterance interpretation on hypothesis 1, it is due to both utterance interpretation and to speaker belief being gradient on hypothesis 3. 
-On these analyses, then, the observed projection variability...
-
-\citealt[498f.]{tbd-variability} suggested that projection variability could be due to speaker belief being gradient, as a consequence of which an interpreter's  projection inference is gradient (hypotheses 1, 3, 5, 7).
-This position still allows for lexical representations being categorical or gradient, and utterance interpretation introducing variability or not.
-A second position entertained in \citealt[498f.]{tbd-variability} is that speaker belief is categorical and that projection variability ``arises from the [interpreter's] uncertainty about whether the speaker is committed'', that is, as part of the utterance interpretation (hypotheses 2, 6).
-
-p.518: the projectivity of projective content is also influenced by the expression that the projective content is associated with
-
-Recall that \citealt[498]{tbd-variability} proposed that projection is a gradient property of content rather than a categorical one after observing variation in the projection of content in their experiments.
-Above, we suggested that the results of \citealt{degen-tonhauser-language}, some of which are shown in Fig.~\ref{fig:dt1a}, provide further support for this proposal.
-\citealt{mandelkern-etal2020}, on the other hand, warned us against ``embracing a notion of presupposition projection that is gradient based on results from inference tasks alone'' (p.497).
-Contrary to what \citealt{mandelkern-etal2020} led us to expect, the results of our experiment do not provide empirical support for their claim that (presupposition) projection is categorical.
-
-What would constitute empirical evidence that projection is categorical?
-One type of evidence would be the results of behavioral experiments with a broad range of contents that suggest a nonarbitrary division of contents into ones that typically project and ones that typically don't.
-Such results does not currently exist, as noted in \citealt{degen-tonhauser-language} and this paper.
-Another type of evidence could be provided by computational models that evaluate whether the data collected in behavioral experiments is better captured by analyses under which projection is categorical or analyses under which it is gradient.
-
-\begin{itemize}
-
-\item Typically, researchers ask this because they wonder about the linguistic representation we want to assume in our models. Simple picture:
-
-triggers introduce context requirement and hence content typically projects
-
-nontriggers do not introduce this and hence exhibit projection to varying degrees
-
-\item More complex picture, already represented by contemporary analyses:
-
-+SCF: introduce anaphoric requirement/felicity requirement
-
-soft triggers: lexical meaning typically gives rise to projection inference
-
-nontriggers: lexical meaning may give rise to projection inference
-
-\item Contemporary presupposition analyses (e.g., \citealt{heim83,vds92,abrusan2011,karttunen2016,simons-etal2017}) predict neither the heterogeneity of the naturalness ratings of factive predicates nor their sensitivity to the prior probability of the CC (though see \citealt{schlenker2021} for the latter).
-These results, which mirror those based on inference ratings, suggest that projection analyses must incorporate more fine-grained distinctions between attitude predicates and the systematic effect of prior probabilities on projection.
-
-\end{itemize}
+There is, to date, no projection analysis that can predict the observed by-content variability (for discussions, see \citealt{tbd-variability,degen-tonhauser-language}).
+What seems clear, given the observed by-content variability, is that the question of whether projection is a gradient or a categorical property of content does not have a single answer for all projective content.
+What the source of the variability is must be decided in future research that carefully compares...
 

### 2023-11-16 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -70,7 +70,7 @@ subjmeans$VeridicalityGroup = factor(x=
                                           ifelse(subjmeans$verb  %in% c("MC"),"MC","NF")),levels=rev(c("F","NF","MC")))
 
 levels(subjmeans$verb)
-view(subjmeans)
+#view(subjmeans)
 
 # version of Figure 2 Language paper
 # plot of means, 95% CIs and participants' ratings 
@@ -218,6 +218,74 @@ ggplot(nat.means, aes(x=context, y=Mean)) +
   theme(strip.background = element_rect(fill="white")) +
   theme(strip.text = element_text(color = "black")) 
 ggsave("../graphs/naturalness-by-context-and-predicate.pdf",height=4,width=9)
+
+# Fig 4: plot of mean certainty ratings for "Julian dances salsa" from Exp 1a of Degen & Tonhauser 2022 -----
+# import data from repo
+cd <- read_csv("https://raw.githubusercontent.com/judith-tonhauser/projective-probability/master/results/5-projectivity-no-fact/data/cd.csv")
+summary(cd)
+table(cd$content)
+table(cd$verb)
+
+# mean projectivity by predicate, including the main clause controls
+means = cd %>%
+  filter(verb != "MC") %>%
+  group_by(verb) %>%
+  summarize(Mean = mean(response), CILow = ci.low(response), CIHigh = ci.high(response)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, verb = fct_reorder(as.factor(verb),Mean))
+means
+
+# define colors for the predicates
+cols = data.frame(V=levels(means$verb))
+
+cols$VeridicalityGroup = as.factor(
+  ifelse(cols$V %in% c("know", "discover", "reveal", "see", "be annoyed"), "F", 
+         ifelse(cols$V %in% c("MC"),"MC","NF")))
+
+levels(cols$V)
+cols$V <- factor(cols$V, levels = cols[order(as.character(means$verb)),]$V, ordered = TRUE)
+
+cols$Colors =  ifelse(cols$VeridicalityGroup == "F", "#D55E00",
+                      ifelse(cols$VeridicalityGroup == "NF", "#009E73",'black'))
+
+
+cols$V <- factor(cols$V, levels = cols[order(as.character(means$verb)),]$V, ordered = TRUE)
+levels(cols$V)
+
+means$VeridicalityGroup = factor(x=
[…]

### 2023-11-17 writing/naturalness-paper/paper.tex
updated naturalness paper

@@
 This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} can distinguish factive predicates (presumed presupposition triggers) from nonfactive ones (nontriggers).
-The results do not support \citepos{mandelkern-etal2020} claim as they do not provide evidence for a categorical distinction between presuppositions and nonpresuppositions.
-{\bf FIX: Furthermore, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to a variety of semantic and pragmatic properties of utterances in addition to the felicity conditions of anaphoric projective content.}
+The results do not support \citepos{mandelkern-etal2020} claim and rather align with \citepos{degen-tonhauser-language} result that there is no evidence for a categorical distinction between factive and nonfactive predicates.
 
@@
 They differ from presuppositions in that they contribute novel information.
-For discussion see, e.g., \citealt{ccmg90,potts05}.} Thus, a content like the content of the clausal complement (CC) of {\em know} in (\ref{first}a) is taken to be a presupposition because a speaker who utters (\ref{first}a) is typically taken to believe the CC, that Julian dances salsa, even though the clausal complement occurs in a polar interrogative.
+For discussion see, e.g., \citealt{ccmg90} and \citealt{potts05}.} Thus, a content like the content of the clausal complement (CC) of {\em know} in (\ref{first}a) is taken to be a presupposition because a speaker who utters (\ref{first}a) is typically taken to believe the CC, that Julian dances salsa, even though the CC occurs in a polar interrogative.
 By contrast, the CC of {\em think} in (\ref{first}b) is not assumed to be a presupposition because it does not typically project.
@@
 There is, however, mounting empirical evidence that projection is not a categorical property of  content but rather a gradient one, with content being more or less projective.
-The measure used to establish this evidence are inference rating measures: Such measures involve presenting participants with an utterance (like that in (\ref{first}a)) and contents (like the content that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content.
+The measure used to establish this evidence are inference rating measures: Such measures involve presenting participants with an utterance (like (\ref{first}a)) and contents (like that Julian dances salsa), and asking for ratings that are taken to reveal the existence or strength of the inference to the content.
 \citealt{karttunen71b}, for instance, suggested that the CCs of {\em regret} and {\em discover}, which he took to be presuppositions, exhibit projection variability in examples like (\ref{kart}).
@@
 \end{exe}
-More recently, experimental investigations have also observed projection variability between presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
-\citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CCs of {\em know, discover,} and {\em reveal} exhibit projection variability: The CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}.
-These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a categorical, categorical one''.
+More recently, experimental investigations have also observed projection variability among presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
+\citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}.
+These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''.
 Further support for this proposal comes from  investigations of both factive and nonfactive predicates in \citealt{degen-tonhauser-language}.
@@
 This is primarily because the goal of this work (including Exp.~3) was not to investigate this hypothesis, but rather whether presupposition projection in conjunctions is symmetric.
-The evidence provided in this work is also limited, however, because only four presupposition triggers and four nontriggers were investigated: The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.
+The evidence provided in this work is also limited, however, because only four presupposition triggers and four nontriggers were investigated.
+The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.
 %\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i).
@@
 \citealt{mandelkern-etal2020} found (as shown in their Fig.~3) that the naturalness ratings for the nontriggers in the control items were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale).
-The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean just above 5).
+The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean at about 5.3).
 Mandelkern and his colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492).
@@
 Whereas Mandelkern and his colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural.
-For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.} This disagreement about the naturalness of presuppositions in explicit ignorance contexts necessitates further investigation.
+For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}
 
@@
 
-This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} provides evidence that projection is a categorical property of content rather than a gradient one.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated.
+This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim that the measure they proposed provides evidence that projection is categorical and distinguishes presuppositions from nonpresuppositions.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx} {\bf DO THIS AFTER FINALIZING SCRIPTS}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated.
 The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.
@@
 This interpretation arises if {\em Ann} and {\em Svenja} associate with {\em too} and {\em also}, respectively.
-While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts.
+While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements, which evoke the question of whether Ann plays an instrument and whether Svenja plays any sport, respectively.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts.
 This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is natural in explicit ignorance contexts.
@@
 
-We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the  filler stimuli was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four filler stimuli was not used to exclude participants because it received a mean rating of only .5.
+We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the filler stimuli (expected to be natural) was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four filler stimuli was not used to exclude participants because it received a mean rating of only .5.
 See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-categorical, 2 preferred to not disclose).
@@
 
-\subsection{Results}
+\subsection{Results and discussion}
 
@@
 Figure \ref{fig:acc-by-expression} shows mean naturalness ratings in the explicit ignorance context by expression (in distinct colors: \color{orange}factive predicates\color{black}, \color{green}nonfactive predicates\color{black},  controls).
-As shown, the purported presupposition triggers, that is, the controls and the factive predicates, do not exhibit a unified pattern: The mean naturalness ratings of four controls ({\em continue, too, also, again}) are at floor, those of {\em stop} and {\em be annoyed} are a bit higher, those of the {\em it-}cleft and {\em know} are higher again, and those of {\em see, discover}, and {\em reveal} are just as high as those of some nonfactive predicates.
-These results do not provide empirical support for a categorical distinction between presuppositions and nonpresuppositions, contrary to what \citealt{mandelkern-etal2020} hypothesized.
-These results also do not provide support for a class of factive predicates that is categorically distinct from nonfactive predicates, in line with the result of \citealt{degen-tonhauser-language} based on inference rating measures.
+As shown, the purported presupposition triggers, that is, the controls and the factive predicates, do not exhibit a unified pattern: The mean naturalness ratings of four controls ({\em continue, too, also, again}) are at floor, those of {\em stop} and {\em be annoyed} are slightly higher, those of the {\em it-}cleft and {\em know} are higher again, but lower than those of {\em see} and {\em discover}, and, finally, those of {\em reveal} are highest.
+We also see that the naturalness ratings of {\em see, discover} and {\em reveal} are just as high as those of some nonfactive predicates.
+These results do not provide empirical support for a categorical distinction between presuppositions and nonpresuppositions, contrary to what \citealt{mandelkern-etal2020} claimed.
+These results also do not provide support for a class of factive predicates that is categorically distinct from nonfactive ones, in line with the result of \citealt{degen-tonhauser-language} based on inference rating measures.
 
@@
 The model predicted naturalness ratings from a fixed effect of expression (with treatment coding and `continue' as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause).
-The pairwise comparison provided us with posterior distributions of estimated marginal means of pairwise differences between the contents associated with each of the 26 expression.
+{\bf FIX: The pairwise comparison provided us with posterior distributions of estimated marginal means of pairwise differences between the contents associated with each of the 26 expression.
 We assume that two contents differ in their naturalness in the explicit ignorance context when the posterior distribution of their difference does not include 0.
+}
 
-As shown in Table \ref{t:pairwise},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.} the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates, such that the ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.
-There is also variation between the factive predicates and the presuppositional controls: The CCs of each of the five purportedly factive are more natural in explicit ignorance contexts than the presuppositions contributed by {\em too} and {\em also}.
+As shown in Table \ref{t:pairwise},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO} {\bf ADD THIS AFTER FINALIZING SCRIPTS}.} the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.
+There is also variation between the factive predicates and the presuppositional controls: The CCs of each of the five purportedly factive predicates are more natural in explicit ignorance contexts than the presuppositions contributed by {\em too} and {\em also}.
 Furthermore, the naturalness ratings for the CCs of the five factive predicates are indistinguishable from different sets of expressions: The naturalness ratings for {\em be annoyed} do not differ from those of {\em stop} (but are higher than those the prejacents of the controls {\em continue, too}, and {\em also}), the naturalness ratings for {\em know} do not differ from those of the {\em it-}cleft, the naturalness ratings for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}.
@@
 Finally, the results of our experiment align with our observation about possible by-content variation in \citepos{mandelkern-etal2020} Exp.~3.
-Recall from above that in their Exp.~3, the pre-state of {\em continue} was judged as (numerically) less natural in explicit ignorance contexts than the CC of the cognitive predicate {\em be aware}.
-This result is in line with the result of our experiment that the pre-state of {\em continue} is judged as less natural than the CC of the cognitive predicate {\em know}.
+Recall from above that in their Exp.~3, the pre-state of {\em continue} was rated as (at least numerically) less natural in explicit ignorance contexts than the CC of the cognitive predicate {\em be aware}.
+This is in line with the result of our experiment that the pre-state of {\em continue} is rated less natural than the CC of the cognitive predicate {\em know}.
 
 In our experiment, the CC of {\em be annoyed} received lower naturalness ratings than the CC of {\em know}.
-This result leads to the undesirable conclusion that the CC of {\em be annoyed} is more plausibly analyzed as a presupposition (perhaps even a hard one) than the CC of {\em know}, regardless of whether naturalness ratings in explicit ignorance contexts are taken to distinguish hard and soft presuppositions (\citealt{simons01,abusch10}) or presuppositions and nonpresuppositions (\citealt{mandelkern-etal2020,kalomoiros-schwarz2021,kalomoiros-schwarz-JoS}).
+This result, if taken at face value, leads to the undesirable conclusion that the CC of {\em be annoyed} is more plausibly analyzed as a presupposition (perhaps even a hard one) than the CC of {\em know}.
 This conclusion is undesirable because the presupposition of emotive predicates like {\em be annoyed} is that the attitude holder believes the CC to be true (e.g., \citealt{heim92,karttunen2016}), not that the speaker believes the CC to be true, as with {\em know}.
@@
 \end{exe}
-Given this status of the CC of {\em be annoyed}, it is surprising that this content received lower mean naturalness ratings than any of the other purportedly factive predicates.
-After all, in a context in which the speaker is explicitly ignorant of the CC, the default assumption that the speaker's and the attitude holder's beliefs are aligned does not arise, and so the interrogative with {\em be annoyed} should be judged natural in the explicit ignorance context, as it only presupposes that the attitude holder, not the speaker, believes the CC.
 
-One hypothesis for why the CC of {\em be annoyed} received such low ratings is that naturalness ratings in explicit ignorance context are sensitive to at-issueness.
+Given this status of the CC of {\em be annoyed}, it is surprising that this content received lower mean naturalness ratings in the explicit ignorance context than any of the other purportedly factive predicates.
+After all, in a context in which the speaker is explicitly ignorant of the CC, the default assumption that the speaker's and the attitude holder's beliefs are aligned does not arise, and so the interrogative with {\em be annoyed} should be rated as natural in the explicit ignorance context, as it only presupposes that the attitude holder, not the speaker, believes the CC.
+
+One hypothesis for why the CC of {\em be annoyed} received such low ratings is that naturalness ratings in explicit ignorance context are sensitive not just to whether the presupposition is defeasible (which is not the case for {\em too, also} and {\em again}) but also to at-issueness.
 On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether the CC is true (?CC; e.g., {\em Does Julian dances salsa?}).
-The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?CC is true.
-If so, the naturalness of the interrogative is modulated by whether the QUD ?CC  together with the interpretation of the interrogative is a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is, if the interrogative can be interpreted as inquiring about the CC.
+The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?CC.
+If so, the naturalness of the interrogative is modulated by whether the QUD ?CC  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about the CC.
 As shown in \citealt{tbd-variability}, projective content differs in whether it can be at-issue.
@@
 Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with the QUD ?CC (e.g., {\em Does Julian dance salsa?}).
-This incongruence may be why the CC of {\em be annoyed} received comparatively low naturalness ratings in an explicit ignorance context.
+This incongruence may be why the CC of {\em be annoyed} received comparatively low naturalness ratings in the explicit ignorance context.
 
 For the CC of {\em know}, Fig.~\ref{fig:acc-by-expression} shows a bimodal pattern for the naturalness ratings, with about half of the participants rating it as fairly unnatural in the explicit ignorance context and about half as fairly natural.
-For the first group of participants, one could assume either that they reacted to the contradiction between the speaker's explicit ignorance statement and the inference arising from the interrogative that the speaker believes the CC to be true, or one could assume an analysis along the same lines as for {\em be annoyed} above.
-For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read both utterances with the rise-fall-rise melodies indicated by the ToBI transcription of the example,\footnote{Explain transcription, what is TOBI} that is, with rising pitch on each of the bold-faced expressions:
+For the first group of participants, one could assume either that they reacted to the contradiction between the speaker's explicit ignorance statement and the inference arising from the interrogative that the speaker believes the CC to be true, or one could assume an analysis along the same lines as for {\em be annoyed} above as the CC of {\em know} is also highly not-at-issue (\citealt{tbd-variability}).
+For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read both utterances with the rise-fall-rise melodies indicated by the ToBI transcription of the example,\footnote{ToBI refers to the Tones and Break Indices annotation system of \citealt{beckman-ayers97}.
+In this system, L*+H identifies a pitch accent whose low target is aligned with the stressed syllable and which is followed by a trailing high target.
+A L* pitch accent may also be followed by a high intermediate phrase tone (H-).
+An intonational phrases may end with a low intermediate phrase tone and a high intonational phrase tone (L-H\%) or a high intermediate and intonational phrase tone (H-H\%).} that is, with rising pitch on each of the bold-faced expressions:
 
@@
 Analyses of the rise-fall-rise contour generally agree that it conveys that the speaker is uncertain about something and that their utterance does not fully answer the QUD (though the analyses differ in other regards; see, e.g., \citealt{ward-hirschberg85,buering97,buering03,wagner-etal2013}).
-In the context of the utterance of the declarative in (\ref{pros}), which conveys that the speaker takes the QUD to be whether Julian dances salsa, the rise-fall-rise contour that the interrogative utterance is realized with conveys that this utterance does not fully answer the QUD, which in turn requires that the speaker is uncertain about the CC.
-Under this interpretation, utterances like (\ref{pros}) may be rated as natural.
+In the context of the utterance of the declarative in (\ref{pros}), which conveys that the speaker takes the QUD to be whether Julian dances salsa, the rise-fall-rise realization of the interrogative conveys that this utterance does not fully answer the QUD.
+This interpretation of the interrogative is possible only if the speaker is uncertain about the CC, in which case the interrogative does not contradict the speaker's preceding explicit ignorance statement.
 
-In sum, naturalness ratings in explicit ignorance contexts do not support \citepos{mandelkern-etal2020} hypothesis that projection is a categorical property of content, or for a categorical distinction between presuppositions and nonpresupposition.
-Instead, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to whether the content is defeasible, to at-issueness, and to prosody.
+In sum, naturalness ratings in explicit ignorance contexts do not support \citepos{mandelkern-etal2020} claim of a categorical distinction between presuppositions and nonpresupposition.
+Instead, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to whether the content is defeasible, to at-issueness, and to prosodically contributed meanings.
 
@@
 For {\em be annoyed} and {\em know} the mean naturalness rating is quite a bit lower in the explicit ignorance context than in the higher prior probability context, whereas the difference is much less pronounced for {\em see} and {\em discover}, and not observed for {\em reveal}.
-This latter pattern is also observed for several nonfactive predicates, including {\em confess, hear, announce}, and {\em be right}.
+This last pattern is also observed for several nonfactive predicates, including {\em confess, hear, announce}, and {\em be right}.
 Thus, naturalness ratings in explicit ignorance vs.\ higher prior probability contexts do not provide support for a categorical distinction between the CCs of factive and nonfactive predicates.
-A comparison between the explicit ignorance and lower prior probability contexts also suggests variation between the factive predicates: Whereas for {\em be annoyed} the mean naturalness ratings are lower in the explicit ignorance context than in the lower prior probability context, they are equally high for {\em know}, but lower in the lower prior probability context than the explicit ignorance context for {\em see, discover}, and {\em reveal}.
-This latter pattern is observed for all of the nonfactive predicates.
+A comparison between the explicit ignorance and lower prior probability contexts also suggests differences between the factive predicates: Whereas for {\em be annoyed} the mean naturalness ratings are lower in the explicit ignorance context than in the lower prior probability context, they are equally high for {\em know}, but lower in the lower prior probability context than the explicit ignorance context for {\em see, discover}, and {\em reveal}.
+This last pattern is observed for all of the nonfactive predicates.
 Thus, naturalness ratings in explicit ignorance vs.\ lower prior probability contexts also do not provide support for a categorical distinction between the CCs of factive and nonfactive predicates.
 
-Finally, we observe that the prior probability of content modulates its naturalness ratings with all predicates, except {\em demonstrate} and {\em pretend}): Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability.
+Finally, we observe that the prior probability of content modulates its naturalness ratings with all predicates, except {\em demonstrate} and {\em pretend}: Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability.
 This result is in line with the result of \citealt{degen-tonhauser-openmind} based on an inference rating task.
@@
 
-These observations were confirmed by posthoc pairwise comparisons of the naturalness ratings in the three context for each of the 20 predicates using the `emmeans' package (\citealt{emmeans}).
+These observations were confirmed by posthoc pairwise comparisons of the naturalness ratings in the three contexts for each of the 20 predicates, again using the `emmeans' package (\citealt{emmeans}).
 The input to the pairwise comparisons were 20 Bayesian mixed-effects beta regression models with weakly informative priors that were fitted using the `brms' package (\citealt{buerkner2017}).
 The models predicted, for each predicate, the naturalness ratings from a fixed effect of context (with treatment coding and `explicit ignorance' as reference level) and included random by-item intercepts.
-The pairwise comparisons provided us with posterior distributions of estimated marginal means of pairwise differences between the three contexts for each predicate.
-We assume that the naturalness ratings in two contexts differ when the posterior distribution of their difference does not include 0.
-See Supplement \ref{a:analysis1} for the full model output.
+{\bf FIX: The pairwise comparisons provided us with posterior distributions of estimated marginal means of pairwise differences between the three contexts for each predicate.
+We assume that the naturalness ratings in two contexts differ when the posterior distribution of their difference does not include 0.}
 
- As shown in Table \ref{t:pairwise2},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.} the model confirms that the effect of context is not uniform for the five factive predicates: The difference between the naturalness ratings in the higher prior probability context and the explicit ignorance context is larger for {\em be annoyed} than for {\em know}, which in turn is larger than for {\em see} and {\em discover}.
+ As shown in Table \ref{t:pairwise2},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.{\bf ADD AFTER FINALIZING SCRIPTS}} the model confirms that the effect of context is not uniform for the five factive predicates: The difference between the naturalness ratings in the higher prior probability context and the explicit ignorance context is larger for {\em be annoyed} than for {\em know}, which in turn is larger than for {\em see} and {\em discover}.
 For {\em reveal} there is no difference, just like for many nonfactive predicates, such as {\em admit, hear, announce} and {\em be right}.
@@
 \caption{Posterior distribution of estimated marginal means of pairwise differences between the three contexts for each expression, with expressions ordered by mean naturalness rating in explicit ignorance context.
+`EIC' stands for explicit ignorance context.
 Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
@@
 Such analyses lead us to expect that presuppositions are rated as natural in higher and lower prior probability contexts, as these contexts do not contradict the CCs, make the CCs uninformative, or lead to problems with binding.
-Contrary to this expectation, the CCs of the five purportedly factive predicates received lower ratings in the lower prior probability context than in the higher prior probability context, where they received ratings that are roughly comparable to those of the factive predicates investigated in \citealt{mandelkern-etal2020}.\footnote{In the support context, the CCs of {\em be aware} and {\em be happy} received a mean naturalness rating of about 4.4 and 4.7 out of 7 in \citepos{mandelkern-etal2020} Exp.~3 (based on Fig.~8 in their Appendix).
-In our higher prior probability context, the mean naturalness ratings for the CCs of the five factive predicates ranged from .64 ({\em reveal}) to .78 ({\em know}) on a scale from 0 to 1.} This result suggests that naturalness ratings are sensitive to the prior probability of content, such that an interrogative with a clause-embedding predicate is judged as more natural in a context in which the CC has a higher prior probability than in a context in which it has a lower prior probability.
-For contemporary analyses of presuppositions, the result means that local accommodation must be sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
+Contrary to this expectation, the CCs of the five purportedly factive predicates received lower ratings in the lower prior probability context than in the higher prior probability context.\footnote{In the higher prior probability context, our five purportedly factive predicates received ratings roughly comparable to the ratings of the factive predicates investigated in \citealt{mandelkern-etal2020} in their support context.
+Specifically, in our higher prior probability context, the mean naturalness ratings for the CCs of the five factive predicates ranged from .64 ({\em reveal}) to .78 ({\em know}) on a scale from 0 to 1.
+In \citepos{mandelkern-etal2020} support context, the CCs of {\em be aware} and {\em be happy} received a mean naturalness rating of about 4.4 and 4.7 out of 7 (based on Fig.~8 in their Appendix).} This result suggests that naturalness ratings are sensitive to the prior probability of content, such that an interrogative with a clause-embedding predicate is judged as more natural in a context in which the CC has a higher prior probability than in a context in which it has a lower prior probability.
+
+This result has several methodological and theoretical implications.
+First, as the result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure, it suggests that inference rating measures and naturalness rating measures are sensitive to some of the same factors.
+Second, naturalness ratings in explicit ignorance and other contexts are sensitive not to whether content is presupposed or not (as claimed in \citealt{mandelkern-etal2020}) but to a whole slew of factors, including whether content is defeasible, at-issueness, prosody, and the prior probability of content.
+Finally, for analyses of presuppositions that assume that presuppositions can be locally accommodated, the result means that local accommodation must be sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
+
+{\bf What do the results tell us about the lexical semantics of the different predicates?}
 
@@
 The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides support for a categorical distinction between presuppositions and nonpresuppositions.
-Neither the by-content comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the by-context comparison in \S\ref{s:analysis2} provided empirical support for a categorical distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
-Rather, the results of the experiment provided further support for the result of \citealt{degen-tonhauser-language} that factive predicates are not a natural class that is categorically different from nonfactive predicates.
-In this section, we discuss theoretical and methodological implications of these results.
+Neither the by-content comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the by-context comparisons in \S\ref{s:analysis2} provided empirical support for a categorical distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
+Rather, the results aligned with that of \citealt{degen-tonhauser-language} that factive predicates are not a natural class that is categorically different from nonfactive predicates.
+In this section, we discuss theoretical implications of these results.
 
@@
 
-Contemporary literature still frequently treats presuppositions as if they were a homogenous class, by characterizing them as contents that typically project out of entailment-canceling environments or by analyzing them as conventionally-coded felicity requirements (e.g., \citealt{mandelkern-etal2020}).
+Contemporary literature like \citealt{mandelkern-etal2020} still frequently treats presuppositions as if they were a homogenous class, by characterizing them as backgrounded content that typically projects out of entailment-canceling environments or by analyzing them as conventionally-coded felicity requirements.
 However, research on projective content has long recognized that the set of contents that have been called presuppositions is heterogeneous.
 One dimension of variation that was already mentioned in \S\ref{s1} is projection.
-Experimental investigations in different languages suggest that contents traditionally called presuppositions exhibit variation in the strength of the inference to the content (e.g., \citealt{xue-onea11} on German, \citealt{tonhauser-guarani-variability} on Paraguayan Guaran\'i, and \citealt{demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language} on English).
+Experimental investigations suggest that contents traditionally called presuppositions exhibit variation in the strength of the inference to the content (e.g., \citealt{xue-onea11} on German, \citealt{tonhauser-guarani-variability} on Paraguayan Guaran\'i, and \citealt{demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language} on English).
 In English, for instance, the CC of {\em discover} is less projective than the CC of {\em know}.
-These results challenge the traditional characterization of presuppositions on the basis of projection.
+These results challenge the traditional characterization of presuppositions as contents that ``typically project''.
 
@@
 \end{exe}
-This variation suggests that analyses of presuppositions as conventionally-coded felicity requirements generalize an analysis that is empirically justified for contents associated with a strong contextual felicity constraint to contents that are not associated with such a constraint.
+This variation suggests that analyses of presuppositions as conventionally-coded felicity requirements (e.g., \citealt{heim83,vds92}) generalize an analysis that is empirically justified for contents associated with a strong contextual felicity constraint to contents that are not associated with such a constraint.
 
-A third dimension of variation is at-issue, that is, whether the content can address the question under discussion: As shown in \citealt{tbd-variability}, contents traditionally called presuppositions vary in this regard: For instance, the pre-state content of {\em stop} and the CC of {\em discover} are more at-issue than the CCs of {\em know} and {\em be annoyed}.
-And, finally, there is the dimension investigated in our experiment, which was referred to as defeasibility by \citealt{simons01} and \citealt{abusch10}.
-While the variation observed in our experiment does not line up perfectly with the variation assumed in these works, the results of our experiment suggest that contents traditionally called presuppositions exhibit variation in how natural they are in explicit ignorance contexts.
+A third dimension of variation is at-issue, that is, whether the content can address the question under discussion.
+For instance, as shown in \citealt{tbd-variability}, the pre-state content of {\em stop} and the CC of {\em discover} are more at-issue than the CCs of {\em know} and {\em be annoyed}.
+This variability challenges a unified characterization of presuppositions as backgrounded content.
+Finally, there is the dimension investigated in our experiment, referred to as defeasibility in \citealt{simons01} and \citealt{abusch10}.
+The results of our experiment suggest that contents traditionally called presuppositions exhibit variation in how natural (or: defeasible) they are in explicit ignorance contexts.
 For instance,  the pre-state content of {\em continue} was rated as less natural than the pre-state content of {\em stop}, which was rated as less natural than the existential content of the {\em it-}cleft, which was rated as less natural than the CC of {\em see}, which was rated as less natural than the CC of {\em reveal}.
+No unified characterizations is forthcoming from this measure either.
 
 These dimensions of variation show that the set of contents traditionally referred to as presuppositions does not constitute a natural class.
-It is in response to such variation that researchers have developed analyses of subsets of these contents that do not rely on a conventional specification of felicity conditions or assume uniform projection (e.g., \citealt{abrusan2011,abrusan2013,abrusan2016,abusch02,abusch10,romoli2015}).
-Consequently, research on projective content can no longer take for granted a characterization of ``presuppositions''  based on projection or an analysis as conventionally-specified felicity conditions.
+It is in response to such variation that researchers have developed analyses of subsets of these contents that capture projection variability and do not rely on a conventional specification of felicity conditions (e.g., \citealt{abrusan2011,abrusan2013,abrusan2016,abusch02,abusch10,romoli2015}).
+A fruitful path forward for contemporary research on projective content is to engage with this variation rather than to keep taking for granted a characterization of ``presuppositions''  based on projection or an analysis as conventionally-specified felicity conditions.
 
@@
 As discussed in \citealt{degen-tonhauser-language}, factive predicates are variably defined as predicates whose CC is presupposed (e.g., \citealt{kiparsky-kiparsky70}) or as predicates whose CC is presupposed and entailed (e.g., \citealt{gazdar79a,schlenker10,abrusan2011}).
-\citealt{degen-tonhauser-language} did not, however, find empirical support for a natural class of `factive' predicates that is categorically distinct from nonfactive ones.
-First, based on the results of two experiments designed to investigate the projection of the CCs of (non)factive predicates as well as the analyses of three existing projection datasets for 45, 78, and 517 (non)factive predicates, respectively,\footnote{These datasets were the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality dataset (\citealt{white-rawlins-nels2018}).} this work suggested that projection does not provide support for a natural class of factive predicates that is categorically distinct from nonfactive ones.
-Second, based on four experiments designed to investigate the entailment of the CCs of (non)factive predicates as well as analyses of entailment in two of the aforementioned datasets, this work suggested that considering entailment in addition to presuppositionality also does not provide support for a natural class of factive predicates that is categorically distinct from nonfactive ones.
+\citealt{degen-tonhauser-language} did not find empirical support for a natural class of factive predicates that is categorically distinct from nonfactive ones: First, based on the results of two experiments designed to investigate the projection of the CCs of (non)factive predicates as well as the analyses of three existing projection datasets with 45, 78, and 517 (non)factive predicates, respectively,\footnote{These datasets were the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality dataset (\citealt{white-rawlins-nels2018}).} this work suggested that projection does not categorically distinguish factive and nonfactive predicates.
+Second, based on four experiments designed to investigate the entailment of the CCs of (non)factive predicates as well as analyses of entailment in two of the aforementioned datasets, this work suggested that considering entailment in addition to projection also does not provide support for a natural class of factive predicates that is categorically distinct from nonfactive ones.
 
-\citealt{degen-tonhauser-language} used a variety of inference rating measures to investigate the projection of the CCs of (non)factive predicates: In some measures, participants read constructed examples, in others naturally occurring ones; in some, examples were presented with a context, in others without a context; in some, the examples had diverse content, in others, they had minimal lexical content; across the measures, the clause-embedding predicates were embedded under different entailment-canceling operators, participants were asked to respond to different questions, and participants gave their response on different scales and sliders.
+\citealt{degen-tonhauser-language} used a variety of inference rating measures to investigate projection: In some measures, participants read constructed examples, in others naturally occurring ones; in some, examples were presented with a context, in others without a context; in some, the examples had diverse content, in others, they had minimal lexical content.
+Across the measures, the clause-embedding predicates were embedded under different entailment-canceling operators, participants were asked to respond to different questions, and participants gave their response on different scales and sliders.
 Despite this variation, the results invariably suggested that factive predicates are not categorically distinct from nonfactive ones.
 
-The experiment presented in this paper presents support for the claim that there is no natural class of factive predicates based on a different measure, namely naturalness ratings in explicit ignorance contexts.
+The experiment presented in this paper presents supports \citepos{degen-tonhauser-language} result based on a different measure, namely naturalness ratings in explicit ignorance contexts.
 The CCs of the five purportedly factive predicates exhibited variation in how natural they were judged to be when the speaker was explicitly ignorant of the CC, and there were no differences between the CCs of {\em see, discover} and {\em reveal} and several nonfactive predicates.
@@
 As discussed in \citealt{degen-tonhauser-language}, ``the search for a class of factive predicates might continue'' (p.586), but we believe that it is more fruitful to investigate the more fine-grained distinctions between clause-embedding predicates that can lead to a better understanding of their use in discourse.
-We offer some suggestions in \S\ref{s:disc3} based on the naturalness ratings we collected.
 
@@
 
-\citealt[497]{mandelkern-etal2020} urged us to ``think twice before embracing a notion of [\ldots] projection that is gradient based on results from inference tasks alone'', advocating instead for a view of projection that is categorical.
+\citealt[497]{mandelkern-etal2020} urged us to ``think twice before embracing a notion of [...] projection that is gradient based on results from inference tasks alone'', advocating instead for a view of projection that is categorical.
 In this section, we turn to the question of whether projection is a categorical or a gradient property of content.
@@
 The empirical evidence pointed to in \S\ref{s1} shows that different interpreters draw projection inferences of varying strength.
-For instance, as shown in Fig.~\ref{fig:dt1a-JULIAN}, the 266 participants of \citepos{degen-tonhauser-language} Exp.~1a drew projection inferences of varying strength for the CC that Julian dances salsa from utterances of the interrogative {\em Did Scott discover that Julian dances salsa?} It is uncontroversial that this utterance content variability is gradient, not categorical.
-It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Also: noise, inattentiveness, but replication of variation exclude this.}
+For instance, as shown in Fig.~\ref{fig:dt1a-JULIAN}, the participants of \citepos{degen-tonhauser-language} Exp.~1a drew projection inferences of varying strength for the CC that Julian dances salsa from utterances of the interrogative {\em Did Scott discover that Julian dances salsa?} It is uncontroversial that this utterance content variability is gradient, not categorical.
+It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Variation can also be introduced by lso: noise, inattentiveness, but replication of variation exclude this.
+{\bf FIX THIS}}
 
@@
 \includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/mean-certainty-by-predicateType-JULIAN}
-\caption{Participants' certainty ratings (measuring projection) of the CC that Julian dances salsa for the 20 \color{orange}factive \color{black} and \color{green}nonfactive \color{black} predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language} (adapted by collapsing the nonfactive predicates into a single category).}\label{fig:dt1a-JULIAN}
+\caption{Participants' certainty ratings (measuring projection) of the CC that Julian dances salsa for the 20 \color{orange}factive \color{black} and \color{green}nonfactive \color{black} predicates investigated in Exp.~1a of \citealt{degen-tonhauser-language}.}\label{fig:dt1a-JULIAN}
 \end{figure}
@@
 What the question about whether projection is gradient or categorical is about is {\em content}, where a content is, for instance, the CC of {\em discover}, the CC of {\em inform}, or the pre-state content of {\em stop}.
-Asking whether projection is gradient or categorical at this level of abstraction (that is, abstracting away from variation attributable to interpreter variability) amounts to asking whether the observed by-content variability (for instance that the CC of {\em discover} is less projective than the CC of {\em inform}; see Fig.~\ref{fig:dt1a}) is derived from underlying representations that are categorical or underlying representations that are not categorical.
-Two aspects of the underlying representations we consider here are speaker belief and lexical entries.
-A speaker who wants to communicate a content might have a belief in the content that is categorical (that is, they either believe that is is true or they are uncertain) or one that is gradient (in which case they believe that the content is true to some degree).\footnote{\citealt[498f.]{tbd-variability} already considered that speaker belief could be gradient or categorical.} Likewise, the expressions associated with the content (like {\em discover} or {\em stop}) could have lexical entries that either license a projection inference or not (categorical) or their lexical entries could be such that whatever is specified may license a projection inference depending on other factors in the uttered sentence (gradient).
-This means that there are four hypotheses for any given content:
+The question, thus, is about whether projection is gradient or categorical when we abstract away from variation that is attributable to interpreter variability and consider the observed by-content variation (for instance that the CC of {\em discover} is less projective than the CC of {\em know}; see Fig.~\ref{fig:dt1a}): Is this by-content variation derived from underlying representations that either categorically trigger projection inferences or not, or from underlying representations that license projection inferences in combination with other factors in the uttered sentence?\footnote{Another potential source of projection variability is speaker belief.
+\citealt[498f.]{tbd-variability} pointed out that speaker belief could be categorical (that is, they either believe that is is true or they are uncertain) or gradient (in which case they believe that the content is true to some degree).
+On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.}
 
-\begin{table}[h]
-\centering
-\begin{tabular}{c | c c}
-Hypothesis & speaker & lexicon  \\ \hline\hline
-1 & categorical & categorical  \\ 
-2 & categorical & {\bf gradient}  \\ 
-3 & {\bf gradient} & categorical \\ 
-4 & {\bf gradient} & {\bf gradient}  \\ 
-\hline
-\end{tabular}
-\caption{Hypotheses about speaker belief and lexical entries for any given content.}\label{t:hyp}
-\end{table}
+Both positions have already been advocated for in the literature.
+Categorical underlying representations are assumed in analyses of presuppositions like \citealt{heim83} and \citealt{vds92}, as they analyze presupposition triggers as being conventionally specified to trigger projection, whereas nontriggers do not trigger projection.
+To account for examples in which presuppositions do not project, such analyses rely on the mechanism of local accommodation.
+As discussed in detail in \citealt{tbd-variability}, it is not clear how local accommodation can capture the observed by-content variation in a principled way.
+Such analyses would need to assume that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}.
+Another problem such analyses face is that the CC of nonfactive predicates is also projective, sometimes even more so than the CC of factive predicates.
+Thus, as also discussed in \citealt{degen-tonhauser-language}, it is not clear which expressions would be analyzed as triggering projection, and which ones not.
 
-Some of this hypothesis space has already been advocated for in the literature.
-Analyses like \citealt{heim83,vds92} and \citealt{potts05} align with hypotheses 1 and 3, as they assume that some expression (triggers of presuppositions or conventional implicatures) are conventionally specified to trigger projection.
-While these analyses do not specify whether speaker belief is categorical or gradient, one could assume, on the former view (hypothesis 1), that a speaker who believes that a content is true chooses an expression that triggers projection, whereas a speaker who is uncertain about a content chooses an expression that does not trigger projection.
-On the latter view (hypothesis 3), a speaker believes content to be true to some degree and chooses an expression that either triggers projection or not, depending on what comes closest to their degree of belief.
-On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.
+Other projection analyses do not assume underlying representations on which particular expressions are conventionally specified to trigger projection.
+For instance, in the analyses of \citealt{abusch02,abusch10} and \citealt{romoli2015}, expressions may specify alternatives as part of their lexical entries and this specification can, depending on other factors in the uttered sentence, lead to a projection inference.
+In the analyses of  \citealt{abrusan2011,abrusan2016} and \citealt{brst-salt10,simons-etal2017} it is entailed content that may project, depending on the utterance context.
+While it is not clear either that these analyses predict the observed by-content projection variability or can be extended to the CCs of nonfactive predicates (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion), what is important for our discussion here, is that such analyses do not assume categorical underlying representations.
 
-Analyses like those of \citealt{abrusan2011,abrusan2016,abusch02,abusch10,romoli2015} and \citealt{brst-salt10,simons-etal2017} could be taken to align with hypotheses 2 and 4.
-On some of these analyses, some expressions specify alternatives as part of their lexical entries and this specification of alternatives can, depending on other factors, give lead to a projection inference (e.g., \citealt{abusch02,abusch10,romoli2015}).
-On other analyses, entailed content may project in particular utterances  (e.g., \citealt{abrusan2011,abrusan2016,brst-salt10,simons-etal2017}).
-Crucially, while these analyses differ in many ways, they all share the assumption that the lexical entries of expressions do not categorically trigger (or not trigger) projection.
-Rather, whether a projection inference is licensed depends on a combination of the lexical entries and other utterance factors.
-These analyses, too, don't specify whether speaker belief is categorical or gradient.
-
-To evaluate which of the four hypotheses in Table \ref{t:hyp} is appropriate for any given content, we have to consider the observed by-content variability.
-An extreme position is to assume for every projective content that speaker belief is categorical and triggered by conventional specification.
-To capture the observed by-content variability, one would then have to assume that it arises during utterance interpretation and affects different expression/content pairs differently.
-That does not seem plausible.
-A more plausible position seems to us to be one on which each of the four hypotheses is correct but for different sets of expression/content pairs.
-For instance, hypothesis 1 might be appropriate  the existence requirement of pronouns and {\em too} (expression/content pairs associated with a strong contextual felicity requirement), hypothesis 2 for the pre-state of {\em stop}, and hypothesis 4 for the CC of {\em inform}.
-The observed by-content variability would then be attributable to a combination of factors, with expression/content pairs differing on the factors and their relative weights.
-
-There is, to date, no projection analysis that can predict the observed by-content variability (for discussions, see \citealt{tbd-variability,degen-tonhauser-language}).
-What seems clear, given the observed by-content variability, is that the question of whether projection is a gradient or a categorical property of content does not have a single answer for all projective content.
-What the source of the variability is must be decided in future research that carefully compares...
-
-\subsection{Future avenues in the investigation and theoretical analysis of projective content}\label{s:disc3}
-
-What are naturalness ratings sensitive to?
-
-- strong contextual felicity constraint
-- at-issueness
-- prior probability of content
-
-\begin{itemize}
-
-\item Lots of interesting properties of projective content to investigate
-
-\item What do the naturalness ratings tell us about the lexical meanings of these predicates?
-
-\item This result suggests that interrogatives are particularly acceptable when the speaker is either ignorant about the CC or the CC has a higher prior probability: to inquire about the CC, or to inquire about the matrix clause.
-
-\item We also observe that the mean naturalness rating is lower in the lower prior probability context than in the higher prior probability context, for all of the predicates except for {\em pretend}.
-This suggests that the prior probability of the CC modulates the naturalness of the interrogative regardless of whether the embedding predicate is factive, where the observed context-sensitivity is unexpected under standard analyses (e.g., \citealt{heim83,vds92}) or nonfactive (where it is expected).
-This result mirrors that of \citealt{degen-tonhauser-openmind}, who found in an inference rating task the strength of projection inferences to the CCs of factive and nonfactive predicates is modulated by the prior probability of the CC.
-One could, of course, propose to make the global accommodation of presuppositions to not be default but to be sensitive to the prior probability of content.
-But the fact of the matter is that the context-sensitivity is observed not just for factive but also for nonfactive predicates, thereby motivating a common projection mechanism.
-
-\end{itemize}
-
-measure doesn't just track anaphoric presuppositions but possibly also mismatches in at-issueness
-
-does the measure track at-issueness or projection?
+So, where does this leave us with respect to the question of whether the projection of content is categorical or gradient?
+The challenge is clear: Proponents of either position must confront the observed by-content projection variability and develop analyses that capture the variability in a principled way.
+As none of the contemporary analyses of projection are able to do so, the verdict is still out.
 
 \section{Conclusions}\label{s4}
+
+For many decades the term `presuppositions' has served researchers well to refer to content with properties distinct from, for instance, entailments, conventional implicatures and conversational implicatures.
+However, over the past two decades, both empirical and theoretical research has pointed out that the set of contents typically referred to as presuppositions is not as homogenous as perhaps initially assumed, resulting in theoretical analyses of particular subsets and an understanding of multiple dimensions of empirical variation among this set of contents.
+Against this background, the claim in \citealt{mandelkern-etal2020} that there is an empirical measure that categorically distinguishes presuppositions from nonpresuppositions sparked hope that there might be, after all, a way of preserving the term `presupposition' as a well-defined notion in the way it has been traditionally used.
+Unfortunately, the results of our experiment did not support this claim, instead providing further evidence for variation among the set of contents traditionally referred to as presuppositions.
+We suggest retiring the term `presupposition' or, at the least, confining its use to a particular subtype of projective content, such as those associated with a strong contextual felicity constraint.
 

### 2023-11-17 results/main/13explicitIgnorance/models/analysis1/table1.tex
updated naturalness paper

 continue & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  too & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  also & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  again & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  stop & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  \color{orange}be annoyed\color{black} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  cleft & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  \color{orange}know\color{black} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  \color{green}demonstrate\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}pretend\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}inform\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}confess\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{orange}see\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}acknowledge\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{orange}discover\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}admit\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}hear\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}prove\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{orange}reveal\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}announce\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\ 
  \color{green}be right\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}establish\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\ 
  \color{green}confirm\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} \\ 
  \color{green}suggest\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} \\ 
  \color{green}think\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} \\ 
  \color{green}say\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} \\

### 2023-11-17 results/main/13explicitIgnorance/models/analysis2/table2.tex
updated naturalness paper

 higher prior - EIC & \cellcolor{yellow1} & \cellcolor{yellow2} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} \\ 
  lower prior - EIC & \cellcolor{yellow2} & \cellcolor{white} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} \\ 
  higher - lower prior & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\

### 2023-11-17 writing/naturalness-paper/paper.tex
updated naturalness paper

@@
 \hline
-\input{../../results/main/13explicitIgnorance/models/latex-tables/table1}
+\input{../../results/main/13explicitIgnorance/models/analysis1/table1}
 \hline
@@
 \hline
-\input{../../results/main/13explicitIgnorance/models/latex-tables/table2}
+\input{../../results/main/13explicitIgnorance/models/analysis2/table2}
 \hline
@@
 This result has several methodological and theoretical implications.
-First, as the result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure, it suggests that inference rating measures and naturalness rating measures are sensitive to some of the same factors.
-Second, naturalness ratings in explicit ignorance and other contexts are sensitive not to whether content is presupposed or not (as claimed in \citealt{mandelkern-etal2020}) but to a whole slew of factors, including whether content is defeasible, at-issueness, prosody, and the prior probability of content.
-Finally, for analyses of presuppositions that assume that presuppositions can be locally accommodated, the result means that local accommodation must be sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
+First, as the result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure, it suggests that inference rating and naturalness rating measures are sensitive to some of the same factors.
+Second, naturalness ratings in explicit ignorance and other contexts are sensitive to a whole slew of factors, including the (implicit) prosody of the utterance and the defeasibility, at-issueness, and the prior probability of content.
+Finally, for analyses of presuppositions that local accommodation, the result means that local accommodation is sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
 
-{\bf What do the results tell us about the lexical semantics of the different predicates?}
+%One thing that is striking about the naturalness ratings is that for many  predicates the ratings are quite high in both the explicit ignorance and the higher prior probability context (e.g., {\em inform, see, discover, hear, reveal, be right, think, say}).
+This suggests that interrogatives with these predicates are natural both in contexts in which the speaker is uncertain about the truth of the CC and in contexts in which the speaker believes the CC to have a high probability.
+This result, which crosscuts the factive/nonfactive distinction, highlights the flexibility with which these predicates are used in discourse,
 
@@
 For instance, as shown in Fig.~\ref{fig:dt1a-JULIAN}, the participants of \citepos{degen-tonhauser-language} Exp.~1a drew projection inferences of varying strength for the CC that Julian dances salsa from utterances of the interrogative {\em Did Scott discover that Julian dances salsa?} It is uncontroversial that this utterance content variability is gradient, not categorical.
-It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).\footnote{Variation can also be introduced by lso: noise, inattentiveness, but replication of variation exclude this.
-{\bf FIX THIS}}
+It is gradient because the projection of utterance content is sensitive to several factors on which interpreters differ, including their prior beliefs about the speaker or about the utterance content, their perception and interpretation of the prosodic realization of the utterances, and their assumptions about the question under discussion addressed by the utterance (e.g., \citealt{djaerv-bacovcin2020,degen-tonhauser-openmind,mahler2020,mahler-thesis,tonhauser-salt26,tbd-variability,tonhauser-etal-sub23}).
 
@@
 The question, thus, is about whether projection is gradient or categorical when we abstract away from variation that is attributable to interpreter variability and consider the observed by-content variation (for instance that the CC of {\em discover} is less projective than the CC of {\em know}; see Fig.~\ref{fig:dt1a}): Is this by-content variation derived from underlying representations that either categorically trigger projection inferences or not, or from underlying representations that license projection inferences in combination with other factors in the uttered sentence?\footnote{Another potential source of projection variability is speaker belief.
-\citealt[498f.]{tbd-variability} pointed out that speaker belief could be categorical (that is, they either believe that is is true or they are uncertain) or gradient (in which case they believe that the content is true to some degree).
+\citealt[498f.]{tbd-variability} pointed out that speaker belief could be categorical (the speaker either believe that the content is true or they are uncertain) or gradient (the speaker believes that the content is true to some degree).
 On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.}
@@
 The challenge is clear: Proponents of either position must confront the observed by-content projection variability and develop analyses that capture the variability in a principled way.
-As none of the contemporary analyses of projection are able to do so, the verdict is still out.
+As none of the contemporary projection analyses are able to do so, one might say that the verdict is still out.
+The empirical evidence, however, seems to favor the hypothesis that projection is gradient, as there is, to date, no empirical evidence that projection is categorical, and several pieces of evidence that it is gradient.
 
@@
 
-\section{Model output: Pairwise comparison of expressions}\label{a:analysis1}
-
-{\bf JT PROPOSES TO LEAVE OUT THIS AND THE NEXT FIGURE}
-
-Fig.~\ref{fig:comparisons1} provides a graphical representation of the results of the pairwise comparisons between the ratings for each expression in the explicit ignorance context.
+%\section{Model output: Pairwise comparison of expressions}\label{a:analysis1}
+%
+%{\bf JT PROPOSES TO LEAVE OUT THIS AND THE NEXT FIGURE}
+%
+%Fig.~\ref{fig:comparisons1} provides a graphical representation of the results of the pairwise comparisons between the ratings for each expression in the explicit ignorance context.
 The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.
-
-\begin{sidewaysfigure}[h!]
-\centering
-\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/comparisons-in-EIC}
-\caption{.}\label{fig:comparisons1}
-\end{sidewaysfigure}
-
-\section{Model output: Pairwise comparison of contexts}\label{a:analysis2}
-
-Fig.~\ref{fig:comparisons2} provides a graphical representation of the results of the pairwise comparisons of the ratings in the three contexts for each of the 20 predicates.
+%
+%\begin{sidewaysfigure}[h!]
+%\centering
+%\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/comparisons-in-EIC}
+%\caption{.}\label{fig:comparisons1}
+%\end{sidewaysfigure}
+%
+%\section{Model output: Pairwise comparison of contexts}\label{a:analysis2}
+%
+%Fig.~\ref{fig:comparisons2} provides a graphical representation of the results of the pairwise comparisons of the ratings in the three contexts for each of the 20 predicates.
 The full model output in table form is available here: \url{LINK TO TABLE IN REPO}.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/context-comparisons}
-\caption{.}\label{fig:comparisons2}
-\end{figure}
+%
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=.8\textwidth]{../../results/main/13explicitIgnorance/graphs/context-comparisons}
+%\caption{.}\label{fig:comparisons2}
+%\end{figure}
 

### 2023-11-17 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -57,10 +57,10 @@ m.b = brm(formula = betamodel,
 
 summary(m.b)
 
-saveRDS(m.b,file="../models/beta-model-mixed1.rds")
+saveRDS(m.b,file="../models/analysis1/beta-model-mixed1.rds")
 
 # to load model
-m.b = readRDS(file="../models/beta-model-mixed1.rds")
+m.b = readRDS(file="../models/analysis1/beta-model-mixed1.rds")
 
 # draws of posterior distributions of estimated marginal means of pairwise differences
 pairwise <- m.b %>%
@@ -75,16 +75,16 @@ pairwise <- m.b %>%
   mutate(contrast = fct_reorder(as.factor(contrast),.value))
 pairwise
 
-saveRDS(pairwise,file="../models/pairwise1.csv")
-pairwise = readRDS(file="../models/pairwise1.csv")
+write_csv(pairwise,file="../models/analysis1/pairwise1.csv")
+pairwise = read_csv(file="../models/analysis1/pairwise1.csv")
 pairwise
-names(pairwise)
 
 # select relevant columns for printing
 pairwise_reduced = pairwise %>%
   select(c(contrast, .value, .lower, .upper))
+pairwise_reduced
 
-#### table of model output in Appendix ----
+#### full model output  ----
 
 tableApp1 = print(xtable(pairwise_reduced),
                #only.contents = T,
@@ -99,19 +99,19 @@ tableApp1 = print(xtable(pairwise_reduced),
                comment = F
 )
 
-write(tableApp1, "../models/latex-tables/tableApp1.tex")
+write(tableApp1, "../models/analysis1/fullModelOutput/analysis1.tex")
        
-#### plot of model output in Appendix ----
-theme_set(theme_bw())   
-ggplot(pairwise, aes(x = .value, y = contrast)) +
-  geom_point() +
-  geom_linerange(aes(xmin = .lower, xmax = .upper)) +
-  geom_vline(xintercept = 0, color = "red") +
-  labs(x = "Average marginal effect of expression", y = NULL) +
-  facet_wrap(. ~ first, scales = "free_y", ncol = 5, drop = TRUE)
-ggsave("../graphs/comparisons-in-EIC.pdf",height=13,width=20)
+# #### plot of model output in Appendix
+# theme_set(theme_bw())   
+# ggplot(pairwise, aes(x = .value, y = contrast)) +
+#   geom_point() +
[…]

### 2023-11-17 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -92,7 +92,7 @@ ggplot(means, aes(x=verb, y=Mean, fill=VeridicalityGroup)) +
   theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) 
 ggsave("../graphs/mean-certainty-by-predicateType.pdf",height=4.5,width=7)
 
-# Fig 2: plot of mean naturalness ratings in explicit ignorance context ----
+# Fig 3: plot of mean naturalness ratings in explicit ignorance context ----
 
 # target data: explicit ignorance context
 table(d$expression)
@@ -143,7 +143,7 @@ ggplot(nat.meansEIC, aes(x=expression, y=Mean)) +
   theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = text.color)) 
 ggsave("../graphs/explicit-ignorance-naturalness-by-predicate.pdf",height=4,width=7)
 
-# Fig 3: plot of mean naturalness ratings in by context ----
+# Fig 4: plot of mean naturalness ratings in by context ----
 # for 20 clause-embedding predicates only
 
 # calculate mean naturalness rating by predicate and context
@@ -186,7 +186,7 @@ t$expression = factor(t$expression, levels=tmp$expression[order(tmp$expression)]
 levels(nat.means$expression)
 levels(t$expression)
 
-# order the contexts: EI, low, high
+# order the contexts: EIC, low, high
 levels(nat.means$context)
 nat.means$context = factor(nat.means$context, levels = c("explicitIgnorance", "factL", "factH"))
 levels(t$context)
@@ -219,7 +219,7 @@ ggplot(nat.means, aes(x=context, y=Mean)) +
   theme(strip.text = element_text(color = "black")) 
 ggsave("../graphs/naturalness-by-context-and-predicate.pdf",height=4,width=9)
 
-# Fig 4: plot of mean certainty ratings for "Julian dances salsa" from Exp 1a of Degen & Tonhauser 2022 -----
+# Fig 5: plot of mean certainty ratings for "Julian dances salsa" from Exp 1a of Degen & Tonhauser 2022 -----
 # import data from repo
 cd <- read_csv("https://raw.githubusercontent.com/judith-tonhauser/projective-probability/master/results/5-projectivity-no-fact/data/cd.csv")
 summary(cd)
@@ -287,7 +287,9 @@ ggplot(subjmeans, aes(x=verb, y=Mean)) +
[…]

### 2023-11-17 results/main/13explicitIgnorance/rscripts/preprocessing.R

@@ -251,49 +251,6 @@ trialCount = d %>%
 trialCount
 # participants who had to redo practice trials have different trial numbers!
 
-# # create separate data for practice trials
-# table(d$task)
-# practice = d %>%
-#   filter(task == "practice") %>%
-#   select(c(participantID, response, trial_index))
-# practice
-# # participants who did the 4 practice trials without errors should have alternating
-# # values above 60, below 40, above 60, below 40
-# 
-# # remove practice trials from data
-# d = d %>%
-#   filter(task != "practice")
-# view(d)
-# 
-# # how many target trials per participant?
-# trialCount = d %>% 
-#   select(trial_index, participantID) %>% 
-#   unique() %>% 
-#   group_by(participantID) %>% 
-#   summarize(count=n())
-# trialCount # now everybody has 30 trials
-# 
-# # make a trial number
-# table(d$trial_index, d$participantID) 
-# table(d$participantID,d$trial_index) 
-# # participants who had to repeat practice trials have other trial numbers
-# # specifically, the more often they had to repeat the practice trials
-# # the later their target trials start
-# 
-# d$trial = d$trial_index
-# table(d$trial, d$participantID)
-# table(d$participantID,d$trial)
-# # for participants who don't have a trial 10, subtract 1
-# # for participant 3 subtract 2 for trials above 11
-# d[d$participantID == 3 & d$trial > 11,]$trial = d[d$participantID == 3 & d$trial > 11,]$trial - 2
-# # everybody: trials above 9 subtract 9
-# d[d$trial > 9,]$trial = d[d$trial > 9,]$trial - 9
-# # now everybody has 30 trials between 1 and 30
-# 
-# # remove original trial_index
-# d = d %>%
-#   select(-c(trial_index))
-
 ### exclude non-English speakers and non-American English speakers
 # exclude non-English speakers
 length(which(is.na(d$language))) #no missing responses

### 2023-11-27 results/main/13explicitIgnorance/models/analysis1/table1.tex
updated naturalness paper analysis and paper

@@
- continue & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  too & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  also & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  again & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  stop & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  \color{orange}be annoyed\color{black} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  cleft & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  \color{orange}know\color{black} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
-  \color{green}demonstrate\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}pretend\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}inform\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}confess\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{orange}see\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}acknowledge\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{orange}discover\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}admit\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}hear\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}prove\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{orange}reveal\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}announce\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\
-  \color{green}be right\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
-  \color{green}establish\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\
-  \color{green}confirm\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} \\
-  \color{green}suggest\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} \\
-  \color{green}think\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} \\
-  \color{green}say\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} \\
+ {\bf continue} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  {\bf too} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  {\bf also} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  {\bf again} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  {\bf stop} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  \color{orange}{\bf be annoyed}\color{black} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  {\bf cleft} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  \color{orange}{\bf know}\color{black} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} \\
+  \color{green}{\bf demonstrate}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf pretend}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf inform}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf confess}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{orange}{\bf see}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf acknowledge}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{orange}{\bf discover}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf admit}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf hear}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf prove}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{orange}{\bf reveal}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf announce}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\
+  \color{green}{\bf be right}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} \\
+  \color{green}{\bf establish}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\
+  \color{green}{\bf confirm}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} \\
+  \color{green}{\bf suggest}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} \\
+  \color{green}{\bf think}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} \\
+  \color{green}{\bf say}\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} \\

### 2023-11-27 writing/naturalness-paper/paper.tex
updated naturalness paper
updated naturalness paper analysis and paper

@@
 % positive coefficients/difference
-\definecolor{yellow1}{RGB}{102,102,0} 
-\definecolor{yellow2}{RGB}{153,153,0} 
-\definecolor{yellow3}{RGB}{204,204,0} 
-\definecolor{yellow4}{RGB}{255,255,204}
+\definecolor{purple1}{RGB}{178,24,43}
+\definecolor{purple2}{RGB}{239,138,98} 
+\definecolor{purple3}{RGB}{253,219,199} 
+%\definecolor{yellow4}{RGB}{255,255,204}
 %\definecolor{green1}{RGB}{0, 158, 115} % >.95
@@
 % negative coefficients/difference
-\definecolor{purple1}{RGB}{150,57,109}
-\definecolor{purple2}{RGB}{188,78,139}
-\definecolor{purple3}{RGB}{215,149,186}
-\definecolor{purple4}{RGB}{236,206,223}
+\definecolor{yellow1}{RGB}{33,102,172}
+\definecolor{yellow2}{RGB}{103,169,207}
+\definecolor{yellow3}{RGB}{209,229,240}
+%\definecolor{purple4}{RGB}{236,206,223}
 
@@
 While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
-\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that naturalness ratings in explicit ignorance contexts are more suitable to measure categorical projection and, thereby, to distinguish presuppositions from nonpresuppositions.
-This paper presents the results of an experiment designed to investigate whether the measure proposed by \citealt{mandelkern-etal2020} can distinguish factive predicates (presumed presupposition triggers) from nonfactive ones (nontriggers).
+\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and categorically distinguishes presuppositions from nonpresuppositions.
+This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim for factive predicates (presumed presupposition triggers) and nonfactive ones (nontriggers).
 The results do not support \citepos{mandelkern-etal2020} claim and rather align with \citepos{degen-tonhauser-language} result that there is no evidence for a categorical distinction between factive and nonfactive predicates.
@@
 \end{exe}
-\citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, only provided limited empirical evidence for the hypothesis that projection is a categorical property of content.
-This is primarily because the goal of this work (including Exp.~3) was not to investigate this hypothesis, but rather whether presupposition projection in conjunctions is symmetric.
-The evidence provided in this work is also limited, however, because only four presupposition triggers and four nontriggers were investigated.
-The presupposition triggers were the change-of-state verbs {\em stop} and {\em continue}, and the two clause-embedding predicates {\em be aware} and {\em be happy}; the nontriggers were {\em now frown on, enjoy, hope} and {\em be sure}.
-%\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i).
-It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past.
-%
-%\begin{exe}
-%\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
-%\end{exe}
-%} 
-In the control items of Exp.~3, these eight expressions were realized in (single-clause) antecedents of conditionals, as shown in (\ref{mtrig}c).
-Naturalness ratings were elicited in explicit ignorance and support contexts.
 
-\citealt{mandelkern-etal2020} found (as shown in their Fig.~3) that the naturalness ratings for the nontriggers in the control items were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale).
+In \citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, items like (\ref{mtrig}) were control items.
+Mandelkern and colleagues found that the naturalness ratings for nontriggers were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale).
 The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean at about 5.3).
-Mandelkern and his colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492).
+Mandelkern and colleagues took this to ``establish[] the effectiveness of the methodology'' (p.492).
 They proposed (p.497):
@@
 
-However, an inspection of the naturalness ratings for the eight (non)triggers provided in Fig.~8 in their Appendix 3 raises doubt about whether the measure indeed provides evidence for a categorical distinction between presuppositions and nonpresuppositions.
-First, considering the mean naturalness ratings in the explicit ignorance contexts, we find ratings for the presuppositions range from just above 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}).
-As the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}), it is not clear that naturalness ratings in explicit ignorance contexts provide evidence for a categorical distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}.
-Second, the effect of context (explicit ignorance vs.\ support) also does not appear to be consistent for presuppositions or nonpresuppositions.
+However, the empirical evidence in \citealt{mandelkern-etal2020} in support of the claim projection is categorical is quite weak.
+This is primarily because the goal of their work was not to investigate this claim, but rather whether presupposition projection in conjunctions is symmetric.
+It is also weak, however, because the data provided in Fig.~8 in their Appendix 3 do not appear to support a categorical distinction between the four presupposition triggers ({\em be aware, be happy, stop, continue}) and the four nontriggers ({\em now frown on, enjoy, hope, be sure}) that were investigated.
+First, considering just the mean naturalness ratings in the explicit ignorance context, it appears that the ratings for the presuppositions and nonpresuppositions overlap: The ratings for the presuppositions ranged from 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}), and the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}).
+It is therefore not clear that naturalness ratings in explicit ignorance contexts provide evidence for a categorical distinction between the four presuppositions and the four nonpresuppositions investigated in \citealt{mandelkern-etal2020}.
+Second, there is also no consistent effect of context.
 Whereas the difference between the mean naturalness ratings in the explicit ignorance and support contexts was quite large for the (presuppositional) pre-state of {\em continue}, the difference was much less pronounced for the (presuppositional) CC of {\em be aware}, where it was roughly comparable to that of the (nonpresuppositional) pre-state of {\em frown on}.
@@
 
-Adding to this, there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts.
-Whereas Mandelkern and his colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural.
+%\footnote{While their Exps.~1 and 2 investigated {\em frown on}, Exp.~3 investigated {\em now frown on}, as shown in (i).
+It is possible that the pre-state content contributed by this expression is projective: (i), for instance, appears to give rise to the projective content that Mary did not frown on doing yoga in the past.
+%
+%\begin{exe}
+%\exi{(i)} If Mary now frowns on doing yoga, then Matthew will interview her for his story. \hfill (\citealt[491]{mandelkern-etal2020})
+%\end{exe}
+%}
+
+Another issue with \citepos{mandelkern-etal2020} claim is that there is disagreement in the literature about the naturalness of presuppositions in explicit ignorance contexts.
+Whereas Mandelkern and colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural.
 For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}
@@
 
-One change made to the design of \citepos{mandelkern-etal2020} Exp.~3 was that the experiment implemented a three-level context condition instead of a two-level one.\footnote{Another change was that participants gave their naturalness ratings on a slider with endpoints labeled `totally unnatural' and `totally natural' rather than on a 7-point Likert scale with endpoints labeled `completely unnatural' and `completely natural'.}  Specifically, \citepos{mandelkern-etal2020} support context (which entailed the relevant context, see (\ref{mtrig}b)) was replaced with two contexts that were compatible with the relevant content but did not entail it.
-The two contexts differed in the prior probability of the  content: In the `lower prior probability' context, illustrated in (\ref{prior}a), the  relevant content (in (\ref{prior}c), that Julian dances salsa) had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context, illustrated in (\ref{prior}b).
-The two contexts for each of the 20 complement clauses that were combined with the 20 predicates were normed in \citealt{degen-tonhauser-openmind}.
+We slightly changed the design of \citepos{mandelkern-etal2020} Exp.~3 to allow for comparison with the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability.
+Specifically, what \citealt{degen-tonhauser-openmind} found was that the higher the prior probability of a content, the stronger the projection inference.
+For instance, a speaker's commitment to the CC of (\ref{prior}c), that Julian dances salsa, was stronger in the context in (\ref{prior}b), where the CC has a higher prior probability, than in the context in (\ref{prior}a), where the CC has a lower prior probability.
 
@@
 \end{exe}
-This change in the context condition was implemented to allow for comparison to the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability.
-Specifically, this work found that the higher the prior probability of a content, the stronger the projection inference.
-Under standard analyses of presuppositions, like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in contexts that are compatible with the presupposition, like the lower and higher prior probability contexts.
-Accordingly, such analyses predict that presuppositions are rated as natural in both lower and higher probability contexts, just as in \citepos{mandelkern-etal2020} support contexts.
+Our experiment replaced \citepos{mandelkern-etal2020} support context that entailed the relevant context with two support contexts that differed in the prior probability of the CC, namely a lower and a higher prior probability context.\footnote{Another change was that participants gave their naturalness ratings on a slider with endpoints labeled `totally unnatural' and `totally natural' rather than on a 7-point Likert scale with endpoints labeled `completely unnatural' and `completely natural'.}  Under analyses of presuppositions like \citealt{heim83} or \citealt{vds92}, presuppositions are by default globally accommodated in contexts that are compatible with the presupposition, like the lower and higher prior probability contexts.
+Accordingly, such analyses predict that presuppositions are rated as natural in both lower and higher probability contexts, just as in \citepos{mandelkern-etal2020} entailing support context.
 
@@
 
-To investigate whether the measure proposed in \citealt{mandelkern-etal2020} provides support for a categorical distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive clause-embedding predicate), and rated the naturalness of the interrogative in the context of the declarative.
+To investigate \citepos{mandelkern-etal2020} claim that their naturalness rating measure provides support for a categorical distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive predicate), and rated the naturalness of the interrogative in the context of the declarative.
 
@@
 In the `lower prior probability' context (\ref{sample}b), the CC had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context (\ref{sample}c).
+The two contexts for each of the 20 complement clauses that were combined with the 20 predicates were normed in \citealt{degen-tonhauser-openmind}.
 See Supplement \ref{a:clauses} for the full set of 20 complement clauses and the two contexts for each complement clause.
@@
 
-Recall that \citealt{simons01} and \citealt{abusch10} assumed that only hard presuppositions are unnatural in explicit ignorance contexts, whereas \citealt{mandelkern-etal2020} assumed that all presuppositions are unnatural in such contexts.
+\paragraph{Results.} Recall that \citealt{simons01} and \citealt{abusch10} assumed that only hard presuppositions are unnatural in explicit ignorance contexts, whereas \citealt{mandelkern-etal2020} assumed that all presuppositions are unnatural in such contexts.
 Figure \ref{fig:acc-by-expression} shows mean naturalness ratings in the explicit ignorance context by expression (in distinct colors: \color{orange}factive predicates\color{black}, \color{green}nonfactive predicates\color{black},  controls).
-As shown, the purported presupposition triggers, that is, the controls and the factive predicates, do not exhibit a unified pattern: The mean naturalness ratings of four controls ({\em continue, too, also, again}) are at floor, those of {\em stop} and {\em be annoyed} are slightly higher, those of the {\em it-}cleft and {\em know} are higher again, but lower than those of {\em see} and {\em discover}, and, finally, those of {\em reveal} are highest.
+As shown, the purported presupposition triggers, that is, the controls and the factive predicates, do not exhibit a unified pattern: The means of four controls ({\em continue, too, also, again}) are at floor, those of {\em stop} and {\em be annoyed} are slightly higher, those of the {\em it-}cleft and {\em know} are higher again, but lower than those of {\em see} and {\em discover}, and, finally, those of {\em reveal} are highest.
 We also see that the naturalness ratings of {\em see, discover} and {\em reveal} are just as high as those of some nonfactive predicates.
@@
 The model predicted naturalness ratings from a fixed effect of expression (with treatment coding and `continue' as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause).
-{\bf FIX: The pairwise comparison provided us with posterior distributions of estimated marginal means of pairwise differences between the contents associated with each of the 26 expression.
-We assume that two contents differ in their naturalness in the explicit ignorance context when the posterior distribution of their difference does not include 0.
-}
+The output of the pairwise comparison are 95\% highest density intervals of estimated marginal mean differences between each of the 26 contents under investigation.
+We assume that two contents differ in their naturalness in the explicit ignorance context if the 95\% highest density interval does not include 0.
 
@@
 
-%\begin{table}[!h]
-%\addtolength{\tabcolsep}{-.1em}
-%\centering
-%\begin{tabular}{r ccccccccccccccccccccccccc}
-%\toprule
-% &  \rot{continue\color{black}} & \rot{too\color{black}} & \rot{also\color{black}} & \rot{again\color{black}} & \rot{stop\color{black}} & \rot{\color{orange}be annoyed\color{black}} & \rot{cleft\color{black}} & \rot{\color{orange}know\color{black}} & \rot{\color{green}demonstrate\color{black}} & \rot{\color{green}pretend\color{black}} & \rot{\color{green}inform\color{black}} & \rot{\color{green}confess\color{black}} & \rot{\color{orange}see\color{black}} & \rot{\color{green}acknowledge\color{black}} & \rot{\color{orange}discover\color{black}} & \rot{\color{green}admit\color{black}} & \rot{\color{green}hear\color{black}} & \rot{\color{green}prove\color{black}} & \rot{\color{orange}reveal\color{black}} & \rot{\color{green}announce\color{black}} & \rot{\color{green}be right\color{black}} & \rot{\color{green}establish\color{black}} & \rot{\color{green}confirm\color{black}} & \rot{\color{green}suggest\color{black}} & \rot{\color{green}think\color{black}} \\
-% \midrule
-%too & -- & \cellcolor{lightgray} & & & & & & & & & & & & & & & & \\
-%also & -- & -- & \cellcolor{lightgray} & & & & & & & & & & & & & & & \\
-%again  & -- & -- & -- & \cellcolor{lightgray} & & & & & & & & & & & & & & \\
-%stop  & $\ast$ & $\ast$  & $\ast$ & $\ast$ & \cellcolor{lightgray} & & & & & & & & & & & & & \\
-%\color{orange} be annoyed\color{black} & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} & & & & & & & & & & & & \\
-%cleft &  $\ast$ & $\ast$ & $\ast$& $\ast$ & $\ast$ & $\ast$ & \cellcolor{lightgray} & & & & & & & & & & & \\
-%\color{orange} know\color{black}  & $\ast$ & $\ast$ & $\ast$&  $\ast$& $\ast$ & $\ast$ & -- & \cellcolor{lightgray} & & & & & & & & & & \\
-%\color{green} demonstrate\color{black}  &  $\ast$   &$\ast$  & $\ast$ & $\ast$ &  $\ast$& $\ast$ & $\ast$ & $\ast$ & \cellcolor{lightgray} & & & & &  & & & & \\
-%\color{green} pretend\color{black}  & $\ast$  &$\ast$  & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & -- & \cellcolor{lightgray} & & & &  & & & & \\
-%\color{green} inform\color{black}  & $\ast$  &$\ast$  & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray}& & &  & & & & \\
-%\color{green} confess\color{black}  &  $\ast$ &$\ast$  & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} & &  & & & & \\
-%\color{orange} see\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- &--  &-- & \cellcolor{lightgray} &  & & & & \\
-%\color{green} acknowledge\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & -- &--  &--  &--  &--  & \cellcolor{lightgray} & & & & \\
-%\color{orange} discover\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ &$\ast$  &$\ast$  &--  & -- &--  & -- &--  & -- & \cellcolor{lightgray} & & & \\
-%\color{green} admit\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ &$\ast$  &$\ast$  & $\ast$ &$\ast$  &-- & -- & -- & -- & \cellcolor{lightgray} & & \\
-%\color{green} hear\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  & $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  &-- & \cellcolor{lightgray} & \\
-%\color{green} prove\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  & -- &-- & $\ast$ &$\ast$  & -- &-- & \cellcolor{lightgray} \\
-%\color{orange} reveal\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ &$\ast$  &$\ast$  &$\ast$  &$\ast$  & $\ast$ &$\ast$  &$\ast$  & $\ast$ & -- & -- & -- & \cellcolor{lightgray}\\
-%\color{green} announce\color{black}  & $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$&$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  & -- &$\ast$  &$\ast$  & \cellcolor{lightgray} \\
-%\color{green} be right\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & $\ast$ & -- & -- & \cellcolor{lightgray}\\
-%\color{green} establish\color{black}  &  $\ast$ & $\ast$ & $\ast$& $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} \\
-%\color{green} confirm\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$&$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  &$\ast$  & $\ast$ &$\ast$  & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & \cellcolor{lightgray}\\
-%\color{green} suggest\color{black}  &  $\ast$ & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ &$\ast$  &$\ast$  &$\ast$  &-- & \cellcolor{lightgray}\\
-%\color{green} think\color{black}   &  $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- & \cellcolor{lightgray} \\
-%\color{green} say\color{black}   &  $\ast$  & $\ast$ & $\ast$&  $\ast$&  $\ast$& $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & $\ast$ & -- & -- \\
-%\bottomrule
-%\end{tabular}
-%\caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating.
-`*' indicates a difference in means, that is, the posterior distribution does not include 0, `--' indicates no difference in means, that is, the posterior distribution includes 0.}\label{t:pairwise}
-%\end{table}
-
 \begin{table}[!h]
@@
 \begin{tabular}{r | cccccccccccccccccccccccccc}
-& \rots{continue} & \rots{too} & \rots{also} & \rots{again} & \rots{stop} & \rots{\color{orange}be annoyed\color{black}} & \rots{cleft} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
+& \rots{\bf continue} & \rots{\bf too} & \rots{\bf also} & \rots{\bf again} & \rots{\bf stop} & \rots{\color{orange}{\bf be annoyed}\color{black}} & \rots{\bf cleft} & \rots{\color{orange}{\bf know}\color{black}} & \rots{\color{green}{\bf demonstrate}\color{black}} & \rots{\color{green}{\bf pretend}\color{black}} & \rots{\color{green}{\bf inform}\color{black}} & \rots{\color{green}{\bf confess}\color{black}} & \rots{\color{orange}{\bf see}\color{black}} & \rots{\color{green}{\bf acknowledge}\color{black}} & \rots{\color{orange}{\bf discover}\color{black}} & \rots{\color{green}{\bf admit}\color{black}} & \rots{\color{green}{\bf hear}\color{black}} & \rots{\color{green}{\bf prove}\color{black}} & \rots{\color{orange}{\bf reveal}\color{black}} & \rots{\color{green}{\bf announce}\color{black}} & \rots{\color{green}{\bf be right}\color{black}} & \rots{\color{green}{\bf establish}\color{black}} & \rots{\color{green}{\bf confirm}\color{black}} & \rots{\color{green}{\bf suggest}\color{black}} & \rots{\color{green}{\bf think}\color{black}} & \rots{\color{green}{\bf say}\color{black}} \\
 \hline
@@
 \caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating in the same context.
-Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
-Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
-Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
-\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
+Color coding indicates whether the difference between the naturalness ratings for contents of the expressions in the first column and those of the expressions in the top row were positive or negative, and the size of the difference; a white cell means that there was no difference: \\
+Positive difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
+\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} \\
+Negative difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} 
 }\label{t:pairwise}
@@
 
-Contrary to what is assumed in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021,kalomoiros-schwarz-JoS}, presuppositions are not invariably unnatural in explicit ignorance contexts.
+\paragraph{Discussion.} Contrary to what is assumed in \citealt{mandelkern-etal2020} and \citealt{kalomoiros-schwarz2021,kalomoiros-schwarz-JoS}, presuppositions are not invariably unnatural in explicit ignorance contexts.
 Rather, in line with what was assumed in \citealt{simons01} and \citealt{abusch10}, some presuppositions seem to be rated as more unnatural in such contexts than others.
@@
 The CCs of {\em see, discover}, and {\em reveal}, on the other hand, might be taken to be defeasible.
-Contrary to what \citealt{abusch10} assumed, the existence requirement of the {\em it-}cleft is not as clearly undefeasible as that of other projective contents (\citealt{smith-hall11} also observed that the {\em it-}cleft did not pattern like a hard trigger in their experiment).
-Finally, the results of our experiment align with our observation about possible by-content variation in \citepos{mandelkern-etal2020} Exp.~3.
-Recall from above that in their Exp.~3, the pre-state of {\em continue} was rated as (at least numerically) less natural in explicit ignorance contexts than the CC of the cognitive predicate {\em be aware}.
+Contrary to what \citealt{abusch10} assumed, the existence requirement of the {\em it-}cleft is not as clearly undefeasible as that of other projective contents.\footnote{\citealt{smith-hall11} also observed that the {\em it-}cleft did not pattern like a hard trigger in their experiment.
+\citealt{jayez-etal2015} found that French {\em aussi} `too', {\em regretter} `regret' and clefts are not entirely unnatural in explicit ignorance contexts.} Finally, the results of our experiment align with our observation about possible by-content variation in \citepos{mandelkern-etal2020} Exp.~3.
+Recall that in their Exp.~3, the pre-state of {\em continue} was rated as (at least numerically) less natural in explicit ignorance contexts than the CC of the cognitive predicate {\em be aware}.
 This is in line with the result of our experiment that the pre-state of {\em continue} is rated less natural than the CC of the cognitive predicate {\em know}.
@@
 For the first group of participants, one could assume either that they reacted to the contradiction between the speaker's explicit ignorance statement and the inference arising from the interrogative that the speaker believes the CC to be true, or one could assume an analysis along the same lines as for {\em be annoyed} above as the CC of {\em know} is also highly not-at-issue (\citealt{tbd-variability}).
-For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read both utterances with the rise-fall-rise melodies indicated by the ToBI transcription of the example,\footnote{ToBI refers to the Tones and Break Indices annotation system of \citealt{beckman-ayers97}.
+For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read the utterances with a prosody that supports an interpretation on which the CC does not project out of the interrogative utterance.
+Two possible prosodies are indicated by the ToBI transcriptions of in (\ref{pros});\footnote{ToBI refers to the Tones and Break Indices annotation system of \citealt{beckman-ayers97}.
 In this system, L*+H identifies a pitch accent whose low target is aligned with the stressed syllable and which is followed by a trailing high target.
-A L* pitch accent may also be followed by a high intermediate phrase tone (H-).
-An intonational phrases may end with a low intermediate phrase tone and a high intonational phrase tone (L-H\%) or a high intermediate and intonational phrase tone (H-H\%).} that is, with rising pitch on each of the bold-faced expressions:
+A L* pitch accent may also be followed by a low (L-) or high (H-) intermediate phrase tone.
+An intonational phrases may end with a low intermediate phrase tone and a high intonational phrase tone (L-H\%) or a high intermediate and intonational phrase tone (H-H\%).} in both, the interrogative is realized with the so-called rise-fall-rise contour.
 
 \begin{exe}
-\ex\label{pros} \gll {\bf I} don't {\bf know} whether Julian dances {\bf salsa}. Does {\bf Cole} know that Julian dances {\bf salsa}? 
+\ex\label{pros} 
+\begin{xlist}
+\ex \gll I don't {\bf know} if Julian dances {\bf salsa}. Does {\bf Cole} know that Julian dances {\bf salsa}? 
+\\ {} {} {L* H-} {} {} {} {L* L-H\%} {} L*+H {} {} {} {} {\hspace*{.2cm} L-H\%} \\ \glt 
+\ex \gll {\bf I} don't {\bf know} if Julian dances {\bf salsa}. Does {\bf Cole} know that Julian dances {\bf salsa}? 
 \\ L*+H {} {L* H-} {} {} {} {L* H-H\%} {} L*+H {} {} {} {} {\hspace*{.2cm} L-H\%} \\ \glt 
+\end{xlist}
 \end{exe}
 Analyses of the rise-fall-rise contour generally agree that it conveys that the speaker is uncertain about something and that their utterance does not fully answer the QUD (though the analyses differ in other regards; see, e.g., \citealt{ward-hirschberg85,buering97,buering03,wagner-etal2013}).
-In the context of the utterance of the declarative in (\ref{pros}), which conveys that the speaker takes the QUD to be whether Julian dances salsa, the rise-fall-rise realization of the interrogative conveys that this utterance does not fully answer the QUD.
+In the context of the utterance of the declarative, which conveys that the speaker takes the QUD to be whether Julian dances salsa, the rise-fall-rise realization of the interrogative conveys that this utterance does not fully answer the QUD.
 This interpretation of the interrogative is possible only if the speaker is uncertain about the CC, in which case the interrogative does not contradict the speaker's preceding explicit ignorance statement.
@@
 
-We next consider whether presuppositions and nonpresuppositions differ with respect to the ratings in contexts in which the speaker is explicitly ignorant of the content and in contexts that are compatible with the content.
-Recall that \citealt{mandelkern-etal2020} expected presuppositions to be sensitive to their context manipulation (explicit ignorance vs.\ support) but not nonpresuppositions.
-Fig.~\ref{fig:acc-by-context} shows mean naturalness ratings for the CCs of the 20 \fcolorbox{black}{orange}{factive} and \fcolorbox{black}{green}{nonfactive} predicates in the three contexts featured in our experiment, with the predicates ordered by their mean naturalness ratings in the explicit ignorance context, as in Fig.~\ref{fig:acc-by-expression}.
+\paragraph{Results.} We next consider whether presuppositions and nonpresuppositions differ with respect to the ratings in explicit ignorance and support contexts.
+Recall that \citealt{mandelkern-etal2020} claimed that presuppositions are sensitive to the context manipulation, but not nonpresuppositions.
+Fig.~\ref{fig:acc-by-context} shows mean naturalness ratings for the CCs of the 20 \fcolorbox{black}{orange}{factive} and \fcolorbox{black}{green}{nonfactive} predicates in the three contexts featured in our experiment (with predicates ordered by their mean rating in the explicit ignorance context, as in Fig.~\ref{fig:acc-by-expression}).
 
@@
 Thus, naturalness ratings in explicit ignorance vs.\ higher prior probability contexts do not provide support for a categorical distinction between the CCs of factive and nonfactive predicates.
-A comparison between the explicit ignorance and lower prior probability contexts also suggests differences between the factive predicates: Whereas for {\em be annoyed} the mean naturalness ratings are lower in the explicit ignorance context than in the lower prior probability context, they are equally high for {\em know}, but lower in the lower prior probability context than the explicit ignorance context for {\em see, discover}, and {\em reveal}.
+A comparison between the explicit ignorance and lower prior probability contexts also suggests differences between the factive predicates: Whereas for {\em be annoyed} the mean naturalness rating is lower in the explicit ignorance context than in the lower prior probability context, the two means are equally high for {\em know}, but the rating in the lower prior probability context is lower than that in the explicit ignorance context for {\em see, discover}, and {\em reveal}.
 This last pattern is observed for all of the nonfactive predicates.
@@
 Finally, we observe that the prior probability of content modulates its naturalness ratings with all predicates, except {\em demonstrate} and {\em pretend}: Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability.
-This result is in line with the result of \citealt{degen-tonhauser-openmind} based on an inference rating task.
+This result is in line with the result of \citealt{degen-tonhauser-openmind} that projection modulated by prior probability.
 
@@
 \centering
-\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate}
-\caption{\small{Mean naturalness rating by context and predicate (\fcolorbox{black}{orange}{factive}, \fcolorbox{black}{green}{nonfactive}); predicates ordered as in Fig.~\ref{fig:acc-by-expression}.
+\includegraphics[width=\textwidth]{../../results/main/13explicitIgnorance/graphs/naturalness-by-context-and-predicate-with-stats}
+\caption{Mean naturalness rating by predicate (\fcolorbox{black}{orange}{factive}, \fcolorbox{black}{green}{nonfactive}) and context, with predicates ordered as in Fig.~\ref{fig:acc-by-expression}.
 Error bars indicate 95\% bootstrapped CIs.
-Violin plots indicate kernel probability density of participants' ratings}.}\label{fig:acc-by-context}
+Violin plots indicate kernel probability density of participants' ratings.
+At the bottom of each facet, a line spanning two contexts  indicates that the ratings in the two contexts differ according to the pairwise comparison.
+The linetype indicates  whether the difference is larger than 1.5 (solid line: ---), below 1.5 but above 0.5 (dashed line: -- --), or below 0.5 and above 0 (dotted line: \raisebox{1mm}{\ldots}).
+}\label{fig:acc-by-context}
 \end{figure}
@@
 The models predicted, for each predicate, the naturalness ratings from a fixed effect of context (with treatment coding and `explicit ignorance' as reference level) and included random by-item intercepts.
-{\bf FIX: The pairwise comparisons provided us with posterior distributions of estimated marginal means of pairwise differences between the three contexts for each predicate.
-We assume that the naturalness ratings in two contexts differ when the posterior distribution of their difference does not include 0.}
+The output of the pairwise comparison are 95\% highest density intervals of estimated marginal mean differences between each of the three contexts for each of the 20 contents.
+We assume that the naturalness ratings between two contexts differ if the 95\% highest density interval does not include 0.\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.{\bf ADD AFTER FINALIZING SCRIPTS}} As shown in Fig.~\ref{fig:acc-by-context}, naturalness ratings in explicit ignorance vs.\ lower or higher prior probability contexts do not provide support for a categorical distinction between presuppositions and nonpresuppositions.
 
- As shown in Table \ref{t:pairwise2},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.{\bf ADD AFTER FINALIZING SCRIPTS}} the model confirms that the effect of context is not uniform for the five factive predicates: The difference between the naturalness ratings in the higher prior probability context and the explicit ignorance context is larger for {\em be annoyed} than for {\em know}, which in turn is larger than for {\em see} and {\em discover}.
+% As shown in Table \ref{t:pairwise2},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO}.{\bf ADD AFTER FINALIZING SCRIPTS}} the model confirms that the effect of context is not uniform for the five factive predicates: The difference between the naturalness ratings in the higher prior probability context and the explicit ignorance context is larger for {\em be annoyed} than for {\em know}, which in turn is larger than for {\em see} and {\em discover}.
 For {\em reveal} there is no difference, just like for many nonfactive predicates, such as {\em admit, hear, announce} and {\em be right}.
@@
 Finally, there is a positive difference between the higher and lower prior probability context for all 20 predicates except for {\em demonstrate} and {\em pretend}.
-
-\begin{table}[!h]
-\addtolength{\tabcolsep}{-.19em}
-\centering
-\begin{tabular}{r  cccccccccccccccccccc}
-& \rots{\color{orange}be annoyed\color{black}} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
-\hline
-\input{../../results/main/13explicitIgnorance/models/analysis2/table2}
-\hline
-\end{tabular}
-\caption{Posterior distribution of estimated marginal means of pairwise differences between the three contexts for each expression, with expressions ordered by mean naturalness rating in explicit ignorance context.
+% 
+%\begin{table}[!h]
+%\addtolength{\tabcolsep}{-.19em}
+%\centering
+%\begin{tabular}{r  cccccccccccccccccccc}
+%& \rots{\color{orange}be annoyed\color{black}} & \rots{\color{orange}know\color{black}} & \rots{\color{green}demonstrate\color{black}} & \rots{\color{green}pretend\color{black}} & \rots{\color{green}inform\color{black}} & \rots{\color{green}confess\color{black}} & \rots{\color{orange}see\color{black}} & \rots{\color{green}acknowledge\color{black}} & \rots{\color{orange}discover\color{black}} & \rots{\color{green}admit\color{black}} & \rots{\color{green}hear\color{black}} & \rots{\color{green}prove\color{black}} & \rots{\color{orange}reveal\color{black}} & \rots{\color{green}announce\color{black}} & \rots{\color{green}be right\color{black}} & \rots{\color{green}establish\color{black}} & \rots{\color{green}confirm\color{black}} & \rots{\color{green}suggest\color{black}} & \rots{\color{green}think\color{black}} & \rots{\color{green}say\color{black}} \\
+%\hline
+%\input{../../results/main/13explicitIgnorance/models/analysis2/table2}
+%\hline
+%\end{tabular}
+%\caption{Posterior distribution of estimated marginal means of pairwise differences between the three contexts for each expression, with expressions ordered by mean naturalness rating in explicit ignorance context.
 `EIC' stands for explicit ignorance context.
 Color coding indicates whether the difference was positive or negative, and the size of the difference; a white cell means that there was no difference: \\
-Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
-Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
-\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
-}\label{t:pairwise2}
-\end{table}
+%Positive difference: \colorbox{yellow1}{\makebox[4em][c]{$>=-1.5$}}, \colorbox{yellow2}{\makebox[4em][c]{($-1.5, -0.5]$}}, \colorbox{yellow3}{\makebox[4em][c]{($-0.5, 0]$}} \\
+%Negative difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
+%\colorbox{purple3}{\makebox[4em][c]{($0.5, 0]$}} 
+%}\label{t:pairwise2}
+%\end{table}
 
-In contrast to \citepos{mandelkern-etal2020} Exp.~3, our experiment did not collect naturalness ratings in a support context that entails the relevant content but in higher and lower prior probability contexts that are compatible with the CC.
+\paragraph{Discussion.} In contrast to \citepos{mandelkern-etal2020} Exp.~3, our experiment did not collect naturalness ratings in a support context that entails the relevant content but in higher and lower prior probability contexts that are compatible with the CC.
 As mentioned above, contemporary presupposition analyses (e.g., \citealt{heim83,vds92}) assume that presuppositions are by default globally accommodated when they are not entailed by or satisfied in the context, and they can be locally accommodated (e.g., under the polar question operator) to avoid contradiction, uninformativity or problems with binding.
@@
 The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides support for a categorical distinction between presuppositions and nonpresuppositions.
-Neither the by-content comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the by-context comparisons in \S\ref{s:analysis2} provided empirical support for a categorical distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
-Rather, the results aligned with that of \citealt{degen-tonhauser-language} that factive predicates are not a natural class that is categorically different from nonfactive predicates.
-In this section, we discuss theoretical implications of these results.
+Neither the comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the comparison of naturalness ratings in explicit ignorance and support contexts in \S\ref{s:analysis2} provided empirical support for a categorical distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
+Rather, the results align with that of \citealt{degen-tonhauser-language}, namely that factive predicates are not a natural class that is categorically different from nonfactive predicates.
+In this section, we discuss theoretical implications of the results.
 
@@
 \end{exe}
-This variation suggests that analyses of presuppositions as conventionally-coded felicity requirements (e.g., \citealt{heim83,vds92}) generalize an analysis that is empirically justified for contents associated with a strong contextual felicity constraint to contents that are not associated with such a constraint.
+This variation suggests that analyses of presuppositions as conventionally-coded felicity requirements (e.g., \citealt{heim83,vds92}) generalize an analysis that is empirically justified for contents associated with a strong contextual felicity constraint to contents that can be informative.
 
-A third dimension of variation is at-issue, that is, whether the content can address the question under discussion.
+A third dimension of variation is at-issueness, that is, whether the content can address the question under discussion.
 For instance, as shown in \citealt{tbd-variability}, the pre-state content of {\em stop} and the CC of {\em discover} are more at-issue than the CCs of {\em know} and {\em be annoyed}.
@@
 Categorical underlying representations are assumed in analyses of presuppositions like \citealt{heim83} and \citealt{vds92}, as they analyze presupposition triggers as being conventionally specified to trigger projection, whereas nontriggers do not trigger projection.
-To account for examples in which presuppositions do not project, such analyses rely on the mechanism of local accommodation.
-As discussed in detail in \citealt{tbd-variability}, it is not clear how local accommodation can capture the observed by-content variation in a principled way.
-Such analyses would need to assume that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}.
+To account for examples in which presuppositions do not project, such analyses rely on local accommodation.
+As discussed in detail in \citealt{tbd-variability}, it is not clear that local accommodation can capture the observed by-content variation in a principled way, as analyses would need to assume that the CC of {\em discover} is more amenable to local accommodation than the CC of {\em know}.
 Another problem such analyses face is that the CC of nonfactive predicates is also projective, sometimes even more so than the CC of factive predicates.
@@
 In the analyses of  \citealt{abrusan2011,abrusan2016} and \citealt{brst-salt10,simons-etal2017} it is entailed content that may project, depending on the utterance context.
-While it is not clear either that these analyses predict the observed by-content projection variability or can be extended to the CCs of nonfactive predicates (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion), what is important for our discussion here, is that such analyses do not assume categorical underlying representations.
+While it is not clear that these analyses predict the observed by-content projection variability or can be extended to the CCs of nonfactive predicates (see \citealt{tbd-variability} and \citealt{degen-tonhauser-language} for discussion), what is important for our discussion here, is that such analyses do not assume categorical underlying representations.
 
-So, where does this leave us with respect to the question of whether the projection of content is categorical or gradient?
+Where does this leave us with the question of whether the projection of content is categorical or gradient?
 The challenge is clear: Proponents of either position must confront the observed by-content projection variability and develop analyses that capture the variability in a principled way.
 As none of the contemporary projection analyses are able to do so, one might say that the verdict is still out.
-The empirical evidence, however, seems to favor the hypothesis that projection is gradient, as there is, to date, no empirical evidence that projection is categorical, and several pieces of evidence that it is gradient.
+The empirical evidence, however, seems to favor the hypothesis that projection is gradient, as there is, to date, no empirical evidence that projection is categorical, and empirical evidence from several different measures that it is gradient.
 
@@
 
-For many decades the term `presuppositions' has served researchers well to refer to content with properties distinct from, for instance, entailments, conventional implicatures and conversational implicatures.
+For many decades the term `presupposition' has served researchers well to refer to content with properties distinct from, for instance, entailments, conventional implicatures and conversational implicatures.
 However, over the past two decades, both empirical and theoretical research has pointed out that the set of contents typically referred to as presuppositions is not as homogenous as perhaps initially assumed, resulting in theoretical analyses of particular subsets and an understanding of multiple dimensions of empirical variation among this set of contents.
@@
 %\bibliographystyle{/Users/tonhauser.1/Library/Latex/cslipubs-natbib}
-\bibliography{/Users/tonhauser.1/Documents/bibliography}
+\bibliography{bibliography}
 
@@
 
-The twenty predicates used in the experiment are the same as in \citealt{degen-tonhauser-openmind,degen-tonhauser-language}:
+The twenty clause-embedding predicates used in the experiment are the same as in \citealt{degen-tonhauser-openmind,degen-tonhauser-language}:
 
@@
 \end{exe}
-
 Eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.
@@
 
-The following list shows the 20 clauses that realized the complements of the predicates in the target and filler stimuli, together with their lower and higher probability facts, respectively, that realized the preceding declarative sentences in the filler stimuli.
+The following list shows the 20 clauses that realized the complements of the predicates in the target stimuli, together with their lower and higher probability facts, respectively, that realized the preceding declarative sentences in the two support contexts.
 
@@
 
-\section{Control and practice stimuli}\label{a:fillerPractice}
+\section{Filler and practice stimuli}\label{a:fillerPractice}
 
-The following list shows the four control stimuli where the interrogative was expected to receive high naturalness ratings in the context of the preceding declarative sentence.
-The values in parentheses indicate the mean naturalness rating for each of the four controls.
-As shown, the third control did not receive the expected high naturalness mean ratings (probably because participants were unwilling to accommodate that Hendrick has a car in a context in which Hendrick was looking to buy a car).
-This control was therefore not used to exclude participants' data.
+The following list shows the four filler stimuli where the interrogative was expected to receive high naturalness ratings in the context of the preceding declarative sentence.
+The values in parentheses indicate the mean naturalness rating that each of the four filler stimuli received.
+As shown, the third filler stimulus did not receive the expected high naturalness mean ratings, probably because participants were unwilling to accommodate that Hendrick has a car in a context in which Hendrick was looking to buy a car.
+This filler stimulus was therefore not used to exclude participants' data.
 

### 2023-11-27 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -215,26 +215,33 @@ tableData = tableData %>%
 # replace NA with gray cells and expressions with color coded versions
 tableData = tableData %>% mutate(across(everything(), ~replace_na(.x, "\\cellcolor{black}")))
 tableData = tableData %>%
-  mutate(expression = recode(expression,"be.annoyed" = "\\color{orange}be annoyed\\color{black}",
-                             "know" = "\\color{orange}know\\color{black}",
-                             "demonstrate" = "\\color{green}demonstrate\\color{black}",
-                             "pretend" = "\\color{green}pretend\\color{black}",
-                             "inform" = "\\color{green}inform\\color{black}",
-                             "confess" = "\\color{green}confess\\color{black}",
-                             "see" = "\\color{orange}see\\color{black}",
-                             "acknowledge" = "\\color{green}acknowledge\\color{black}",
-                             "discover" = "\\color{orange}discover\\color{black}",
-                             "admit" = "\\color{green}admit\\color{black}",
-                             "hear" = "\\color{green}hear\\color{black}",
-                             "prove" = "\\color{green}prove\\color{black}",
-                             "reveal" = "\\color{orange}reveal\\color{black}",
-                             "announce" = "\\color{green}announce\\color{black}",
-                             "be.right" = "\\color{green}be right\\color{black}",
-                             "establish" = "\\color{green}establish\\color{black}",
-                             "confirm" = "\\color{green}confirm\\color{black}",
-                             "suggest" = "\\color{green}suggest\\color{black}",
-                             "think" = "\\color{green}think\\color{black}",
-                             "say" = "\\color{green}say\\color{black}"                             
+  mutate(expression = recode(expression,
+                             "continue" = "{\\bf continue}",
+                             "too" = "{\\bf too}",
+                             "also" = "{\\bf also}",
+                             "again" = "{\\bf again}",
+                             "stop" = "{\\bf stop}",
+                             "cleft" = "{\\bf cleft}",
[…]

### 2023-11-27 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -143,8 +143,86 @@ ggplot(nat.meansEIC, aes(x=expression, y=Mean)) +
   theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = text.color)) 
 ggsave("../graphs/explicit-ignorance-naturalness-by-predicate.pdf",height=4,width=7)
 
-# Fig 4: plot of mean naturalness ratings in by context ----
+
+# Fig 4.1: plot of mean naturalness ratings in by context ----
 # for 20 clause-embedding predicates only
+
+# calculate mean naturalness rating by predicate and context
+table(d$expression)
+table(d$context)  #explicit ignorance / factL / factH
+
+nat.means = d %>%
+  filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4") %>%
+  filter(expression != "also" & expression != "too" & expression != "again" & expression != "cleft" &
+           expression != "stop" & expression != "continue") %>%
+  group_by(expression,context) %>%
+  summarize(Mean = mean(response), CILow = ci.low(response), CIHigh = ci.high(response)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh) %>%
+  ungroup %>%
+  select(-c(CILow,CIHigh)) %>%
+  mutate(context = as.factor(context))
+nat.means
+table(nat.means$context)
+nat.means$expression <- as.factor(nat.means$expression)
+levels(nat.means$expression)
+
+t = d %>%
+  filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4") %>%
+  filter(expression != "also" & expression != "too" & expression != "again" & expression != "cleft" &
+           expression != "stop" & expression != "continue") %>%
+  mutate(context = as.factor(context))
+levels(t$context)
+
+# order predicates by mean naturalness rating in EIC
+tmp <- t %>%
+  filter(context == "explicitIgnorance") %>%
+  group_by(expression) %>%
+  summarize(Mean = mean(response)) %>%
+  mutate(expression = fct_reorder(as.factor(expression),Mean))
+tmp
+levels(tmp$expression)
+
[…]

### 2023-11-29 writing/naturalness-paper/paper.tex
updated naturalness paper and analysis scripts

@@
 \author[$\bullet$]{Judith Degen}
-\affil[$\circ$]{University of Stuttgart}
-\affil[$\bullet$]{Stanford University}
+\affil[$\circ$]{University of Stuttgart, Department of Linguistics, Stuttgart, Germany, judith.tonhauser@ling.uni-stuttgart.de (corresponding author)}
+\affil[$\bullet$]{Stanford University, Department of Linguistics, Stanford, USA, jdegen@stanford.edu}
 
@@
 \maketitle
+\thispagestyle{empty}
 
@@
 While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
-\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and categorically distinguishes presuppositions from nonpresuppositions.
+\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and distinguishes presuppositions from nonpresuppositions.
 This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim for factive predicates (presumed presupposition triggers) and nonfactive ones (nontriggers).
@@
 
+\bigskip
+
+\noindent
+{\bf Funding:} The experiment was funded by Judith Tonhauser's research funds.
+
+\bigskip
+
+\noindent
+{\bf Competing interests:} The authors have no competing interests to declare.
+
+\bigskip
+
+\noindent
+{\bf Ethics approval and informed consent:} The experiment was approved by the ethics review committee of the University of Stuttgart. Informed consent was obtained from the participants.
+
+\bigskip
+
+\noindent
+{\bf Availability of materials, data, and code:} See footnote \ref{f:github}.
+
+\bigskip
+
+\noindent
+{\bf Author contributions:} Both authors contributed to the conception and design of the experiment.
+Material preparation, data collection and analysis were performed by both authors.
+The first draft of the manuscript was written by Judith Tonhauser.
+Both authors commented on previous versions of the manuscript.
+Both authors read and approved the final manuscript.
+
 \newpage
+
+\clearpage
+\pagenumbering{arabic}
+
+\begin{center}
+{\LARGE I don't know if projection is categorical. \\[.2cm] Did \citealt{mandelkern-etal2020} discover that it is?}	
+\end{center}
+
+\begin{abstract}
+
+Presuppositions are taken to typically project out of entailment-canceling environments like the scope of negation (e.g., \citealt{ccmg90}).
+While projection is often characterized as categorical (that is, content either typically projects or not), there is mounting empirical evidence that projection is gradient, with content being more or less projective (e.g., \citealt{karttunen71b,xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
+\citealt{mandelkern-etal2020} critically evaluated the inference rating measures on which this empirical evidence is based and claimed that a different measure, namely naturalness ratings in explicit ignorance contexts, provides support for categorical projection and distinguishes presuppositions from nonpresuppositions.
+This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim for factive predicates (presumed presupposition triggers) and nonfactive ones (nontriggers).
+The results do not support \citepos{mandelkern-etal2020} claim and rather align with \citepos{degen-tonhauser-language} result that there is no evidence for a categorical distinction between factive and nonfactive predicates.
+
+\end{abstract}
 
@@
 
-Many theoretical works assume that projection is a categorical property of utterance content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition).
+Many theoretical works assume that projection is a categorical property of  content, that is, content either typically projects (in which case it is a presupposition) or it does not (in which case it is not a presupposition).
 There is, however, mounting empirical evidence that projection is not a categorical property of  content but rather a gradient one, with content being more or less projective.
@@
 \end{exe}
-More recently, experimental investigations have also observed projection variability among presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
-\citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CC of {\em know} is more projective than that of {\em discover}, and the CC of {\em discover} is more projective than that of {\em reveal}.
+More recently, experimental investigations also observed projection variability among presuppositions (e.g., \citealt*{xue-onea11,demarneffe-etal-sub23,tbd-variability,degen-tonhauser-language}).
+\citealt{tbd-variability}, for instance, observed in their Exps.~1 that the CC of {\em know} is more projective than that of {\em discover}, which is more projective than that of {\em reveal}.
 These results led \citealt[498]{tbd-variability} to propose that ``projectivity [\ldots is] a gradient property of content, rather than a binary, categorical one''.
@@
 This is because the CCs of both factive and nonfactive predicates were projective to varying degrees compared to nonprojective main clause content.
-A sample illustration of these results is given in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).
+A sample illustration of these results is in Fig.~\ref{fig:dt1a} from their Exp.~1a, which shows the mean projection ratings of 5 factive predicates (in \color{orange}orange\color{black}) and 15 nonfactive predicates (in \color{green}green\color{black}).\footnote{The data from \citepos{degen-tonhauser-language} Exp.~1a were accessed here: \url{https://github.com/judith-tonhauser/projective-probability/tree/master/results/5-projectivity-no-fact}.
+The script that was used to create Fig.~\ref{fig:dt1a} can be found in the GitHub repository linked in footnote \ref{f:github}.}
 
@@
 That is, in such constellations,  participants ``will not have an alternative to interpreting the presupposition at the utterance level, and in turn seeing the utterance as infelicitous and the speaker as incoherent'' (p.497).
-If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'' as in the explicit ignorance contexts (p.497).
-Both presuppositions and nonpresuppositions were expected to be judged as natural in the support contexts, which entail the relevant content, illustrated in (\ref{mtrig}b).
+If, on the other hand, the content is not a presupposition (such as the pre-state content of {\em now frown on} in (\ref{mtrig}c)), the utterance is assumed to be natural: Even if participants might, in a different context, draw a projection inference ``for any of a variety of pragmatic reasons short of entailment or presupposition, [\ldots] they will tend to relinquish that inclination when there is pragmatic pressure to do so'', as in the explicit ignorance contexts (p.497).
+Both presupposition triggers and nontriggers were expected to be judged as natural in the support context, which entails the relevant content, illustrated in (\ref{mtrig}b).
 
@@
 In \citepos{mandelkern-etal2020} Exp.~3, which used the proposed measure, items like (\ref{mtrig}) were control items.
-Mandelkern and colleagues found that the naturalness ratings for nontriggers were relatively high in both contexts (with means just below 5, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'' on their 7-point Likert scale).
+Mandelkern and colleagues found that the naturalness ratings for nontriggers were relatively high in both contexts (with means just below 5 on their 7-point Likert scale, where 1 meant ``completely unnatural'' and 7 meant ``completely natural'').
 The presupposition triggers, on the other hand, had a lower mean naturalness rating in the explicit ignorance context (mean around 3) than in the support context (mean at about 5.3).
@@
 This is primarily because the goal of their work was not to investigate this claim, but rather whether presupposition projection in conjunctions is symmetric.
-It is also weak, however, because the data provided in Fig.~8 in their Appendix 3 do not appear to support a categorical distinction between the four presupposition triggers ({\em be aware, be happy, stop, continue}) and the four nontriggers ({\em now frown on, enjoy, hope, be sure}) that were investigated.
+It is also weak, however, because only four presupposition triggers ({\em be aware, be happy, stop, continue}) and four nontriggers ({\em now frown on, enjoy, hope, be sure}) were investigated, and the results (provided in Fig.~8 in their Appendix 3) do not appear to support a categorical distinction between the four triggers and the four nontriggers.
 First, considering just the mean naturalness ratings in the explicit ignorance context, it appears that the ratings for the presuppositions and nonpresuppositions overlap: The ratings for the presuppositions ranged from 1 (for the pre-state of {\em continue}) to about 3.5 (for the CC of {\em be aware}), and the ratings for the nonpresuppositions range from just below 3 (for the CC of {\em hope}) to 4.5 (for the CC of {\em be aware}).
@@
 Whereas Mandelkern and colleagues assumed that presuppositions are unnatural in explicit ignorance contexts, \citealt{simons01} and \citealt{abusch10} assumed that only `hard' (nondefeasible) presuppositions are, but that `soft' (defeasible) ones are natural.
-For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}
+For instance,  the examples with {\em stop}, {\em discover}, and {\em win} in (\ref{eic2}) were assumed to be natural in explicit ignorance contexts, and the examples with {\em again}, the {\em it-}cleft and {\em too} in (\ref{eic3}) unnatural.%\footnote{\citealt{abusch10} marked (\ref{eic3}b) and (\ref{eic3}c) with `??' instead of `$\#$'.}
 
@@
 
-\ex \infelic I have no idea whether anyone read that letter. But if it is John
+\ex \verymarginal I have no idea whether anyone read that letter. But if it is John
 who read it, let's ask him to be discreet about the content. \hfill (\citealt[40]{abusch10})
 
-\ex \infelic I have no idea whether John read that proposal. But if Bill read it too, let's ask them to confer and simply give us a yes-no response. \hfill (\citealt[40]{abusch10})
+\ex \verymarginal I have no idea whether John read that proposal. But if Bill read it too, let's ask them to confer and simply give us a yes-no response. \hfill (\citealt[40]{abusch10})
 \end{xlist}
@@
 
-This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim that the measure they proposed provides evidence that projection is categorical and distinguishes presuppositions from nonpresuppositions.\footnote{Link to experiment, materials, and analysis scripts: \url{xxx} {\bf DO THIS AFTER FINALIZING SCRIPTS}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated.
+This paper presents the results of an experiment designed to investigate \citepos{mandelkern-etal2020} claim that the measure they proposed provides evidence that projection is categorical and distinguishes presuppositions from nonpresuppositions.\footnote{\label{f:github}The experiment was approved by the ethics review committee of the [name of university suppressed].
+The experiment, materials, data, and analysis scripts can be accessed here:  \url{xxx} {\bf DO THIS AFTER FINALIZING SCRIPTS}}  To allow for comparison with the results of \citealt{degen-tonhauser-language}, the CCs of the same 20 (non)factive clause-embedding predicates (see Fig.~\ref{fig:dt1a}) were investigated.
 The experiment also included control stimuli with {\em stop, continue, again, too, also}, and an {\em it-}cleft to allow for comparison with the results of \citealt{mandelkern-etal2020} and the assumptions made in \citealt{simons01} and \citealt{abusch10}.
 
-We slightly changed the design of \citepos{mandelkern-etal2020} Exp.~3 to allow for comparison with the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability.
+The design of \citepos{mandelkern-etal2020} Exp.~3 was slightly changed to allow for comparison with the result of \citealt{degen-tonhauser-openmind} that the projection of content is sensitive to the content's prior probability.
 Specifically, what \citealt{degen-tonhauser-openmind} found was that the higher the prior probability of a content, the stronger the projection inference.
@@
 
-To investigate \citepos{mandelkern-etal2020} claim that their naturalness rating measure provides support for a categorical distinction between presuppositions and nonpresuppositions, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a (non)factive predicate), and rated the naturalness of the interrogative in the context of the declarative.
+To investigate \citepos{mandelkern-etal2020} claim that their naturalness rating measure provides evidence for categorical projection, participants read two-sentence discourses consisting of a declarative (which provided the context) and an interrogative (which realized a factive or nonfactive predicate), and rated the naturalness of the interrogative in the context of the declarative.
 
@@
 We recruited 425 participants on Prolific.
-Due to a server error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-categorical, 2 preferred to not disclose).
+Due to a programming error, the data of only 398 participants was recorded (ages: 19-73, mean: 40.8; 187 women, 201 men, 8 non-categorical, 2 preferred to not disclose).
 The recruited participants were required to live in the USA, to speak English as their first language, to have completed at least 100 tasks, and to have an approval rating of at least 99\%.
@@
 In the `lower prior probability' context (\ref{sample}b), the CC had a comparatively lower prior probability, whereas it had a comparatively higher prior probability in the `higher prior probability' context (\ref{sample}c).
-The two contexts for each of the 20 complement clauses that were combined with the 20 predicates were normed in \citealt{degen-tonhauser-openmind}.
+The two contexts for each of the 20 complement clauses were normed in \citealt{degen-tonhauser-openmind}.
 See Supplement \ref{a:clauses} for the full set of 20 complement clauses and the two contexts for each complement clause.
@@
 
-In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included the six control stimuli shown in (\ref{filler}).
+In addition to the 1,200 target stimuli (400 interrogatives $\times$ 3 contexts), the materials also included the six control stimuli in (\ref{filler}).
 The interrogatives of the control stimuli featured six expressions typically analyzed as presupposition triggers, namely {\em stop, continue, again, too, also}, and an {\em it-}cleft.
 The preceding declaratives conveyed the speaker's explicit ignorance about the presuppositions associated with these expressions.\footnote{The contents investigated for {\em too} and {\em also} in (\ref{filler}d) and (\ref{filler}e) are that Ann plays an instrument other than the flute and that Svenja plays a sport other than soccer.
-This interpretation arises if {\em Ann} and {\em Svenja} associate with {\em too} and {\em also}, respectively.
-While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements, which evoke the question of whether Ann plays an instrument and whether Svenja plays any sport, respectively.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts.
+This interpretation arises if {\em too} and {\em also} associate with {\em the flute} and {\em soccer}, respectively.
+While prosody was not controlled for, this focus association is made plausible by the preceding explicit ignorance statements, which evoke the question of whether Ann plays any instrument and whether Svenja plays any sport, respectively.} Two of these expressions, namely {\em stop} and {\em continue}, were target expressions in \citepos{mandelkern-etal2020} Exp.~3, where they were assumed to be unnatural in explicit ignorance contexts.
 This is in contrast to \citealt{simons01}, who took {\em stop} to be a soft trigger that is natural in explicit ignorance contexts.
@@
 Twelve of the target stimuli were presented in the explicit ignorance context, and the other eight in a low or a higher prior probability context (four each).
-Each participants' set of 30 stimuli contained the same six control stimuli and the same four filler stimuli.
+Each participants' set also contained the same six control stimuli and the same four filler stimuli.
 Each of the 30 stimuli were presented as utterances by a unique named speaker.
@@
 
-Participants were instructed to rate ``how natural the question sounds in the context of the statement''.
+Participants were instructed to rate how natural the question sounds in the context of the statement.
 As shown in Figure \ref{f:trials}, they were asked give their rating on a slider from `totally unnatural' (coded as 0) to `totally natural' (coded as 1).
@@
 
-We excluded the data of five participants who did not self-identify as native speakers of American English as well as the data of 23 participants whose mean response to the filler stimuli (expected to be natural) was more than 2 sd below the group mean.\footnote{Contrary to what was preregistered, one of the four filler stimuli was not used to exclude participants because it received a mean rating of only .5.
-See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 male, 8 non-categorical, 2 preferred to not disclose).
-Each of the 20 predicates received at least 200 ratings in the explicit ignorance context (mean: 222 ratings), at least 59 ratings in the lower prior probability context (mean: 74), and at least 61 ratings in the higher prior probability context (mean: 74).
-The six control stimuli received 370 ratings each.
+We excluded the data of 5 participants who did not self-identify as native speakers of American English and that of 23 participants whose mean response to the fillers (expected to be natural) was more than 2 sd below the group mean.\footnote{Contrary to what was planned, one of the four filler stimuli was not used to exclude participants' data because it received a mean rating of only .5.
+See Supplement \ref{a:fillerPractice} for details.} The data of 370 participants entered into the analysis (ages: 19-80, mean: 40.7; 175 women, 185 men, 8 nonbinary, 2 did not disclose).
+The 20 predicates each received at least 200 ratings in the explicit ignorance context (mean: 222 ratings), at least 59 ratings in the lower prior probability context (mean: 74), and at least 61 ratings in the higher prior probability context (mean: 74).
+The six controls received 370 ratings each.
 
@@
 
-We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for a categorical difference between presuppositions and nonpresuppositions (\S\ref{s:analysis1}).
+We first address the question of whether naturalness ratings in explicit ignorance contexts provide empirical evidence for categorical projection (\S\ref{s:analysis1}).
 We then consider whether a comparison of naturalness ratings in the three contexts provides such evidence (\S\ref{s:analysis2}).
@@
 The output of the pairwise comparison are 95\% highest density intervals of estimated marginal mean differences between each of the 26 contents under investigation.
-We assume that two contents differ in their naturalness in the explicit ignorance context if the 95\% highest density interval does not include 0.
+We assume that two contents differ in their naturalness in the explicit ignorance context if this 95\% highest density interval does not include 0.
 
-As shown in Table \ref{t:pairwise},\footnote{The full model output is available here: \url{LINK TO TABLE IN REPO} {\bf ADD THIS AFTER FINALIZING SCRIPTS}.} the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.
+As shown in Table \ref{t:pairwise},\footnote{Even though the naturalness ratings collected from the participants range from 0 to 1, the estimated marginal mean differences represented in this table range from -2.7 (difference between the pre-state content of {\em continue} and the CC of {\em say} to 2.7 (difference between the CC of {\em say} and the prejacent of {\em too}).
+These large estimated marginal mean differences are possible because they are based on the estimates and 95\% credible intervals of the  contents associated with the 26 expressions, where the estimates ranged from -1.32 (for the intercept, which was {\em continue}) to 2.69 (for {\em say}).
+The full model output is available here: \url{LINK TO TABLE IN REPO} {\bf ADD THIS AFTER FINALIZING SCRIPTS}.} the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.
 There is also variation between the factive predicates and the presuppositional controls: The CCs of each of the five purportedly factive predicates are more natural in explicit ignorance contexts than the presuppositions contributed by {\em too} and {\em also}.
 Furthermore, the naturalness ratings for the CCs of the five factive predicates are indistinguishable from different sets of expressions: The naturalness ratings for {\em be annoyed} do not differ from those of {\em stop} (but are higher than those the prejacents of the controls {\em continue, too}, and {\em also}), the naturalness ratings for {\em know} do not differ from those of the {\em it-}cleft, the naturalness ratings for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}.
-These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for a categorical distinction between factive and nonfactive predicates, or between presuppositions and nonpresuppositions.
+These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for categorical projection or for a categorical distinction between factive and nonfactive predicates.
 
@@
 \end{tabular}
-\caption{Posterior distributions of estimated marginal means of pairwise differences in the explicit ignorance context, with expressions ordered by mean naturalness rating in the same context.
-Color coding indicates whether the difference between the naturalness ratings for contents of the expressions in the first column and those of the expressions in the top row were positive or negative, and the size of the difference; a white cell means that there was no difference: \\
+\caption{Estimated marginal mean differences between the 26 target and control contents in the explicit ignorance context by expression, with expressions ordered by mean naturalness rating in that context.
+The color coding indicates whether the difference between the naturalness ratings for contents of the expressions in the first column and those of the expressions in the top row was positive or negative, as well as the size of the difference.
+A white cell means that there was no difference.\\
 Positive difference: \colorbox{purple1}{\makebox[4em][c]{$>=1.5$}}, \colorbox{purple2}{\makebox[4em][c]{($1.5,0.5]$}}, 
@@
 Given this status of the CC of {\em be annoyed}, it is surprising that this content received lower mean naturalness ratings in the explicit ignorance context than any of the other purportedly factive predicates.
-After all, in a context in which the speaker is explicitly ignorant of the CC, the default assumption that the speaker's and the attitude holder's beliefs are aligned does not arise, and so the interrogative with {\em be annoyed} should be rated as natural in the explicit ignorance context, as it only presupposes that the attitude holder, not the speaker, believes the CC.
+After all, in a context in which the speaker is explicitly ignorant of the CC, the default assumption that the speaker's and the attitude holder's beliefs are aligned does not hold, and so the interrogative with {\em be annoyed} should be rated as natural in the explicit ignorance context, as it only presupposes that the attitude holder, not the speaker, believes the CC.
 
 One hypothesis for why the CC of {\em be annoyed} received such low ratings is that naturalness ratings in explicit ignorance context are sensitive not just to whether the presupposition is defeasible (which is not the case for {\em too, also} and {\em again}) but also to at-issueness.
-On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether the CC is true (?CC; e.g., {\em Does Julian dances salsa?}).
-The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?CC.
-If so, the naturalness of the interrogative is modulated by whether the QUD ?CC  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about the CC.
+On this hypothesis, the explicit ignorance statement (e.g., {\em I don't know whether Julian dances salsa}) does not merely convey the speaker's ignorance about the content $c$ to be investigated, but also identifies that the speaker takes the question under discussion (QUD) to be whether $c$ is true (?$c$ e.g., {\em Does Julian dances salsa?}).
+The speaker's immediately following interrogative utterance is interpreted with respect to the QUD ?$c$.
+If so, the naturalness of the interrogative is modulated by whether the QUD ?$c$  and the interpretation of the interrogative constitute a felicitous strategy of inquiry (\citealt[32f.]{roberts12}), that is,  by whether the interrogative can be interpreted as inquiring about $c$.
 As shown in \citealt{tbd-variability}, projective content differs in whether it can be at-issue.
 The CCs of emotive predicates, including {\em be annoyed}, are particularly resistant to being interpreted as at-issue.
-Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with the QUD ?CC (e.g., {\em Does Julian dance salsa?}).
+Thus, an interrogative with {\em be annoyed} (e.g., {\em Is Cole annoyed that Julian dances salsa?}) is unlikely to be interpreted as congruent with a QUD about the CC (e.g., {\em Does Julian dance salsa?}).
 This incongruence may be why the CC of {\em be annoyed} received comparatively low naturalness ratings in the explicit ignorance context.
 
-For the CC of {\em know}, Fig.~\ref{fig:acc-by-expression} shows a bimodal pattern for the naturalness ratings, with about half of the participants rating it as fairly unnatural in the explicit ignorance context and about half as fairly natural.
+For the CC of {\em know}, Fig.~\ref{fig:acc-by-expression} shows a bimodal pattern for the naturalness ratings, with about half of the participants rating it as relatively unnatural in the explicit ignorance context and about half as relatively natural.
 For the first group of participants, one could assume either that they reacted to the contradiction between the speaker's explicit ignorance statement and the inference arising from the interrogative that the speaker believes the CC to be true, or one could assume an analysis along the same lines as for {\em be annoyed} above as the CC of {\em know} is also highly not-at-issue (\citealt{tbd-variability}).
-For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read the utterances with a prosody that supports an interpretation on which the CC does not project out of the interrogative utterance.
-Two possible prosodies are indicated by the ToBI transcriptions of in (\ref{pros});\footnote{ToBI refers to the Tones and Break Indices annotation system of \citealt{beckman-ayers97}.
+For the second group of participants, those who judged speakers' utterances like (\ref{pros}) as fairly natural, one hypothesis is that they read the utterances with a (possibly implicit) prosody that supports an interpretation on which the CC does not project out of the interrogative utterance.
+Two possible prosodies are indicated by the ToBI transcriptions of in (\ref{pros}); in both, the interrogative is realized with the so-called rise-fall-rise contour.\footnote{ToBI refers to the Tones and Break Indices annotation system of \citealt{beckman-ayers97}.
 In this system, L*+H identifies a pitch accent whose low target is aligned with the stressed syllable and which is followed by a trailing high target.
 A L* pitch accent may also be followed by a low (L-) or high (H-) intermediate phrase tone.
-An intonational phrases may end with a low intermediate phrase tone and a high intonational phrase tone (L-H\%) or a high intermediate and intonational phrase tone (H-H\%).} in both, the interrogative is realized with the so-called rise-fall-rise contour.
+An intonational phrases may end with a low intermediate phrase tone and a high intonational phrase tone (L-H\%) or a high intermediate and intonational phrase tone (H-H\%).}
 
@@
 In the context of the utterance of the declarative, which conveys that the speaker takes the QUD to be whether Julian dances salsa, the rise-fall-rise realization of the interrogative conveys that this utterance does not fully answer the QUD.
-This interpretation of the interrogative is possible only if the speaker is uncertain about the CC, in which case the interrogative does not contradict the speaker's preceding explicit ignorance statement.
+This interpretation of the interrogative is possible only if the speaker is uncertain about the CC, in which case the interrogative does not contradict the speaker's preceding explicit ignorance statement, leading some participants to give a relatively high naturalness rating.
 
-In sum, naturalness ratings in explicit ignorance contexts do not support \citepos{mandelkern-etal2020} claim of a categorical distinction between presuppositions and nonpresupposition.
+In sum, naturalness ratings in explicit ignorance contexts do not support \citepos{mandelkern-etal2020} claim that projection is categorical, or that their measure distinguishes presuppositions from nonpresuppositions.
 Instead, the results suggest that naturalness ratings in explicit ignorance contexts may be sensitive to whether the content is defeasible, to at-issueness, and to prosodically contributed meanings.
@@
 Finally, we observe that the prior probability of content modulates its naturalness ratings with all predicates, except {\em demonstrate} and {\em pretend}: Content is judged as more natural in a context in which it has a higher prior probability than in a context in which it has a lower prior probability.
-This result is in line with the result of \citealt{degen-tonhauser-openmind} that projection modulated by prior probability.
+This result is in line with the result of \citealt{degen-tonhauser-openmind} that projection is modulated by prior probability.
 
@@
 Violin plots indicate kernel probability density of participants' ratings.
-At the bottom of each facet, a line spanning two contexts  indicates that the ratings in the two contexts differ according to the pairwise comparison.
-The linetype indicates  whether the difference is larger than 1.5 (solid line: ---), below 1.5 but above 0.5 (dashed line: -- --), or below 0.5 and above 0 (dotted line: \raisebox{1mm}{\ldots}).
+Below each facet, a line spanning two contexts  indicates that the ratings differ according to the pairwise comparison of contexts.
+The linetype indicates  whether the difference is at least 1.5 (solid line: ---), below 1.5 but at least 0.5 (dashed line: -- --), or below 0.5 but above 0 (dotted line: \raisebox{1mm}{\ldots}).
 }\label{fig:acc-by-context}
@@
 
-\paragraph{Discussion.} In contrast to \citepos{mandelkern-etal2020} Exp.~3, our experiment did not collect naturalness ratings in a support context that entails the relevant content but in higher and lower prior probability contexts that are compatible with the CC.
-As mentioned above, contemporary presupposition analyses (e.g., \citealt{heim83,vds92}) assume that presuppositions are by default globally accommodated when they are not entailed by or satisfied in the context, and they can be locally accommodated (e.g., under the polar question operator) to avoid contradiction, uninformativity or problems with binding.
-Such analyses lead us to expect that presuppositions are rated as natural in higher and lower prior probability contexts, as these contexts do not contradict the CCs, make the CCs uninformative, or lead to problems with binding.
+\paragraph{Discussion.} As mentioned above, contemporary presupposition analyses (e.g., \citealt{heim83,vds92}) assume that presuppositions are by default globally accommodated when they are not entailed by or satisfied in the context, and they can be locally accommodated (e.g., under the polar question operator) to avoid contradiction, uninformativity or problems with binding.
+Such analyses led us to expect that presuppositions are rated as natural in higher and lower prior probability contexts, as these contexts do not contradict the CCs, make the CCs uninformative, or lead to problems with binding.
 Contrary to this expectation, the CCs of the five purportedly factive predicates received lower ratings in the lower prior probability context than in the higher prior probability context.\footnote{In the higher prior probability context, our five purportedly factive predicates received ratings roughly comparable to the ratings of the factive predicates investigated in \citealt{mandelkern-etal2020} in their support context.
 Specifically, in our higher prior probability context, the mean naturalness ratings for the CCs of the five factive predicates ranged from .64 ({\em reveal}) to .78 ({\em know}) on a scale from 0 to 1.
-In \citepos{mandelkern-etal2020} support context, the CCs of {\em be aware} and {\em be happy} received a mean naturalness rating of about 4.4 and 4.7 out of 7 (based on Fig.~8 in their Appendix).} This result suggests that naturalness ratings are sensitive to the prior probability of content, such that an interrogative with a clause-embedding predicate is judged as more natural in a context in which the CC has a higher prior probability than in a context in which it has a lower prior probability.
+In \citepos{mandelkern-etal2020} support context, the CCs of {\em be aware} and {\em be happy} received mean naturalness ratings of about 4.4 and 4.7 out of 7 (based on Fig.~8 in their Appendix).} This result suggests that naturalness ratings are sensitive to the prior probability of content, such that an interrogative with a clause-embedding predicate is judged as more natural in a context in which the CC has a higher prior probability than in a context in which it has a lower prior probability.
 
 This result has several methodological and theoretical implications.
-First, as the result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure, it suggests that inference rating and naturalness rating measures are sensitive to some of the same factors.
-Second, naturalness ratings in explicit ignorance and other contexts are sensitive to a whole slew of factors, including the (implicit) prosody of the utterance and the defeasibility, at-issueness, and the prior probability of content.
-Finally, for analyses of presuppositions that local accommodation, the result means that local accommodation is sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
+First, this result aligns with the result of \citealt{degen-tonhauser-openmind} based on an inference rating measure.
+It therefore suggests that inference rating and naturalness rating measures are sensitive to some of the same factors.
+Second, naturalness ratings in explicit ignorance and support contexts are sensitive to a whole slew of factors, including the (implicit) prosody of the utterance and the defeasibility, at-issueness, and the prior probability of content.
+The assumption that naturalness ratings in such contexts distinguish presuppositions from nonpresuppositions cannot be maintained.
+Finally, for analyses of presuppositions that assume a mechanism of local accommodation, the result means that local accommodation is sensitive to the prior probability of content, not just to whether the content is contradictory or uninformative in the context.
 
@@
 
-The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides support for a categorical distinction between presuppositions and nonpresuppositions.
-Neither the comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the comparison of naturalness ratings in explicit ignorance and support contexts in \S\ref{s:analysis2} provided empirical support for a categorical distinction between the CCs of factive and nonfactive predicates, or, more generally, between presuppositions and nonpresuppositions.
+The experiment was designed to investigate \citepos{mandelkern-etal2020} claim that the measure used in their Exp.~3 provides support for categorical  projection and a categorical distinction between presuppositions and nonpresuppositions.
+Neither the comparison of naturalness ratings in explicit ignorance contexts  in \S\ref{s:analysis1} nor the comparison of naturalness ratings in explicit ignorance and support contexts in \S\ref{s:analysis2} provided support for this claim.
 Rather, the results align with that of \citealt{degen-tonhauser-language}, namely that factive predicates are not a natural class that is categorically different from nonfactive predicates.
@@
 Contemporary literature like \citealt{mandelkern-etal2020} still frequently treats presuppositions as if they were a homogenous class, by characterizing them as backgrounded content that typically projects out of entailment-canceling environments or by analyzing them as conventionally-coded felicity requirements.
-However, research on projective content has long recognized that the set of contents that have been called presuppositions is heterogeneous.
+However, research on projective content has long recognized that the set of contents that have been called presuppositions is heterogeneous along several dimensions.
 One dimension of variation that was already mentioned in \S\ref{s1} is projection.
@@
 
-A second dimension of variation is whether the content is associated with what \citealt{brst-lang11} referred to as a `strong contextual felicity' constraint.
+A second dimension of variation is whether the content is associated with what \citealt*{brst-lang11} referred to as a `strong contextual felicity' constraint.
 Some presumed presuppositions are not associated with such a constraint, which means that they are judged to be acceptable in a context that is neutral with respect to the content.
@@
 \begin{xlist}
-\ex Context: A girl backs out of a driveway and hits Susi's car.
+\ex A girl backs out of a driveway and hits Susi's car.
 A woman comes running out of the house, apologizes that her daughter hit Susi's car, and says: \\ She knows that she has to use her glasses to drive.
 
-\ex Context: Malena is eating her lunch, a hamburger, on the bus going into town. A woman who she doesn't know sits down next to her and says:
+\ex Malena is eating her lunch, a hamburger, on the bus going into town. A woman who she doesn't know sits down next to her and says:
 \\ \infelic Our bus driver is eating empanadas, too.
@@
 For instance,  the pre-state content of {\em continue} was rated as less natural than the pre-state content of {\em stop}, which was rated as less natural than the existential content of the {\em it-}cleft, which was rated as less natural than the CC of {\em see}, which was rated as less natural than the CC of {\em reveal}.
-No unified characterizations is forthcoming from this measure either.
+No unified characterizations of `presupposition' is forthcoming from this measure either.
 
-These dimensions of variation show that the set of contents traditionally referred to as presuppositions does not constitute a natural class.
+These dimensions of variation strongly suggest that the set of contents traditionally referred to as presuppositions does not constitute a natural class.
 It is in response to such variation that researchers have developed analyses of subsets of these contents that capture projection variability and do not rely on a conventional specification of felicity conditions (e.g., \citealt{abrusan2011,abrusan2013,abrusan2016,abusch02,abusch10,romoli2015}).
-A fruitful path forward for contemporary research on projective content is to engage with this variation rather than to keep taking for granted a characterization of ``presuppositions''  based on projection or an analysis as conventionally-specified felicity conditions.
+A fruitful path forward for contemporary research on projective content is to engage with this variation rather than to keep taking for granted a characterization of `presupposition'  based on projection or an analysis as conventionally-specified felicity conditions.
 
@@
 As discussed in \citealt{degen-tonhauser-language}, factive predicates are variably defined as predicates whose CC is presupposed (e.g., \citealt{kiparsky-kiparsky70}) or as predicates whose CC is presupposed and entailed (e.g., \citealt{gazdar79a,schlenker10,abrusan2011}).
-\citealt{degen-tonhauser-language} did not find empirical support for a natural class of factive predicates that is categorically distinct from nonfactive ones: First, based on the results of two experiments designed to investigate the projection of the CCs of (non)factive predicates as well as the analyses of three existing projection datasets with 45, 78, and 517 (non)factive predicates, respectively,\footnote{These datasets were the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality dataset (\citealt{white-rawlins-nels2018}).} this work suggested that projection does not categorically distinguish factive and nonfactive predicates.
+\citealt{degen-tonhauser-language} did not find empirical support for a natural class of factive predicates that is categorically distinct from nonfactive ones: First, based on the results of two experiments designed to investigate the projection of the CCs of (non)factive predicates as well as the analyses of three existing projection datasets with up to 517 (non)factive predicates,\footnote{These datasets were the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality dataset (\citealt{white-rawlins-nels2018}).} this work suggested that projection does not categorically distinguish factive and nonfactive predicates.
 Second, based on four experiments designed to investigate the entailment of the CCs of (non)factive predicates as well as analyses of entailment in two of the aforementioned datasets, this work suggested that considering entailment in addition to projection also does not provide support for a natural class of factive predicates that is categorically distinct from nonfactive ones.
@@
 What the question about whether projection is gradient or categorical is about is {\em content}, where a content is, for instance, the CC of {\em discover}, the CC of {\em inform}, or the pre-state content of {\em stop}.
-The question, thus, is about whether projection is gradient or categorical when we abstract away from variation that is attributable to interpreter variability and consider the observed by-content variation (for instance that the CC of {\em discover} is less projective than the CC of {\em know}; see Fig.~\ref{fig:dt1a}): Is this by-content variation derived from underlying representations that either categorically trigger projection inferences or not, or from underlying representations that license projection inferences in combination with other factors in the uttered sentence?\footnote{Another potential source of projection variability is speaker belief.
+The question, thus, is about whether projection is gradient or categorical when we abstract away from variation that is attributable to interpreter variability\footnote{Another potential source of projection variability is speaker belief.
 \citealt[498f.]{tbd-variability} pointed out that speaker belief could be categorical (the speaker either believe that the content is true or they are uncertain) or gradient (the speaker believes that the content is true to some degree).
-On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.}
+On either view, interpreters might draw projection inferences of different strengths due to uncertainty about (the degree of) the speaker's belief.}  and consider the observed by-content variation (for instance that the CC of {\em discover} is less projective than the CC of {\em know}; see Fig.~\ref{fig:dt1a}).
+The question, then, is whether this by-content variation derives from underlying categorical representations (where an expression either categorically triggers a projection inferences, or it does not), or from underlying representations that license projection inferences in combination with other factors in the uttered sentence?
 
@@
 Unfortunately, the results of our experiment did not support this claim, instead providing further evidence for variation among the set of contents traditionally referred to as presuppositions.
-We suggest retiring the term `presupposition' or, at the least, confining its use to a particular subtype of projective content, such as those associated with a strong contextual felicity constraint.
+We suggest retiring the term `presupposition' or, at least, confining its use to a particular subtype of projective content, such as content associated with a strong contextual felicity constraint.
 

### 2023-11-29 results/main/13explicitIgnorance/rscripts/analysis.R

@@ -1,5 +1,8 @@
 # 13_explicitIgnorance
-# analysis
+# analysis.R
+# analysis 1: section 2.2.1
+# analysis 2: section 2.2.2
+# this script also creates the visual outputs of the analyses reported in the paper
 
 # set working directory to directory of script
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
@@ -79,6 +82,11 @@ write_csv(pairwise,file="../models/analysis1/pairwise1.csv")
 pairwise = read_csv(file="../models/analysis1/pairwise1.csv")
 pairwise
 
+# which differences are observed?
+summary(pairwise$.value)
+#Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+#-2.6943 -1.0429 -0.2822 -0.2895  0.3865  2.7232 
+
 # select relevant columns for printing
 pairwise_reduced = pairwise %>%
   select(c(contrast, .value, .lower, .upper))
@@ -111,7 +119,7 @@ write(tableApp1, "../models/analysis1/fullModelOutput/analysis1.tex")
 #   facet_wrap(. ~ first, scales = "free_y", ncol = 5, drop = TRUE)
 # ggsave("../graphs/comparisons-in-EIC.pdf",height=13,width=20)
 
-#### Table 1 ----
+#### create latex input for Table 1 ----
 
 # select needed columns
 tableInput = pairwise %>%
@@ -371,7 +379,7 @@ analysis2 = print(xtable(values),
 
 write(analysis2, "../models/analysis2/fullModelOutput/analysis2.tex")
 
-# create data for presenting results as part of Figure 4
+#### create data for presenting results as part of Figure 4 ----
 
 # there are three lines in each facet (that is, for each expression)
 # one line for each context (x = context)
@@ -449,76 +457,77 @@ contrasts = contrasts %>%
 write_csv(contrasts, file="../data/contrasts.csv")
 
 
-# create Table 2 
+# the remainder of this code is to create Table 2 (which has been removed from the paper because
+# the results of the statistical comparison have been integrated to Fig 4)
 
-# what is the distribution of the values?
-ggplot(values, aes(x=value)) +
-  geom_histogram() 
-    
-    
-diff2 = data.frame(predicate = character(), contrast = character(), significant = character())
-diff2
-
-for (p in predicates) {
-  for (i in 1:nrow(get(paste("pairwise.",p,sep="")))) {
-    print(i)
[…]

### 2023-11-29 results/main/13explicitIgnorance/rscripts/graphs.R

@@ -72,7 +72,7 @@ subjmeans$VeridicalityGroup = factor(x=
 levels(subjmeans$verb)
 #view(subjmeans)
 
-# version of Figure 2 Language paper
+# version of Figure 2 Degen & Tonhauser 2022 Language paper
 # plot of means, 95% CIs and participants' ratings 
 ggplot(means, aes(x=verb, y=Mean, fill=VeridicalityGroup)) +
   geom_violin(data=subjmeans,scale="width",linewidth = 0, alpha = .3) +
@@ -144,83 +144,7 @@ ggplot(nat.meansEIC, aes(x=expression, y=Mean)) +
 ggsave("../graphs/explicit-ignorance-naturalness-by-predicate.pdf",height=4,width=7)
 
 
-# Fig 4.1: plot of mean naturalness ratings in by context ----
-# for 20 clause-embedding predicates only
-
-# calculate mean naturalness rating by predicate and context
-table(d$expression)
-table(d$context)  #explicit ignorance / factL / factH
-
-nat.means = d %>%
-  filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4") %>%
-  filter(expression != "also" & expression != "too" & expression != "again" & expression != "cleft" &
-           expression != "stop" & expression != "continue") %>%
-  group_by(expression,context) %>%
-  summarize(Mean = mean(response), CILow = ci.low(response), CIHigh = ci.high(response)) %>%
-  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh) %>%
-  ungroup %>%
-  select(-c(CILow,CIHigh)) %>%
-  mutate(context = as.factor(context))
-nat.means
-table(nat.means$context)
-nat.means$expression <- as.factor(nat.means$expression)
-levels(nat.means$expression)
-
-t = d %>%
-  filter(expression != "practice" & expression != "controlGood1" & expression != "controlGood2" & expression != "controlGood3" & expression != "controlGood4") %>%
-  filter(expression != "also" & expression != "too" & expression != "again" & expression != "cleft" &
-           expression != "stop" & expression != "continue") %>%
-  mutate(context = as.factor(context))
-levels(t$context)
-
-# order predicates by mean naturalness rating in EIC
-tmp <- t %>%
-  filter(context == "explicitIgnorance") %>%
[…]

### 2023-11-29 results/main/13explicitIgnorance/rscripts/preprocessing.R

@@ -1,7 +1,7 @@
 # 13_explicitIgnorance
 # preprocessing
 
-# server issue: doesn't collect all the data that were paid for
+# programming issue: not all the data that were paid for are collected
 # first round: paid 400, collected only 375
 # second round: paid 25, collected only 23
 
@@ -20,8 +20,9 @@ theme_set(theme_bw())
 # read in the raw data
 d1 = read_tsv("../data/combined1.tsv")
 d2 = read_tsv("../data/combined2.tsv")
-# parsing warnings are due to the rows with demographics info and final screens
+# parsing warnings are due to the rows with demographics info and final screens (ignore)
 
+# bind the data
 d = rbind(d1,d2)
 
 # remove rows with info about instructions and final screens
@@ -34,7 +35,7 @@ d$participantID <- match(d$participant_id, unique(sort(d$participant_id)))
 # how many participants?
 length(unique(d$participantID)) #398
 
-# from Prolific
+# info from Prolific
 # £12.66/hr
 # median time: 00:06:24
 
@@ -298,8 +299,8 @@ ggplot(c.means, aes(x=expression,y=Mean)) +
   ylab("Mean response")
 
 # controlGood3 didn't work: here the mean naturalness rating is way lower
-# makes sense (in hindsight: question presupposes that he has a car, but the statement says
-# that he was looking to buy one)
+# makes sense in hindsight: question presupposes that he has a car, but the statement says
+# that he was looking to buy one
 
 # "item_id" : "controlGood1",
 # "utterance" : "Does Samantha have a new hat?",

### 2023-12-13 writing/SuB-paper/Template_SuB-Latex.tex
started SuB paper draft

%!TEX TS-program = pdflatex
\documentclass[a4paper,12pt,twoside]{article}

\usepackage{mathptmx}
\usepackage{parskip}
\usepackage[top=1.5in,left=1in,right=1in,bottom=1in]{geometry}
%\setlength{\parindent}{0pt}
%\setlength{\parskip}{\the\baselineskip}
\pagestyle{empty}

\makeatletter
\renewcommand\@makefntext[1]{%
  \parindent 1em\noindent
  \hbox{\@makefnmark}#1}
\renewcommand\large{\@setfontsize\large{14pt}{18}} % For standardizing with the Word template: normally \large is 14.4pt
\makeatother
\usepackage{titlesec}
\titleformat{\subsection}{\normalfont\normalsize}{\thesubsection.}{1ex}{}
\titleformat{\subsubsection}{\normalfont\normalsize}{\thesubsubsection.}{1ex}{}
\titleformat{\section}{\normalfont\normalsize\bfseries}{\thesection.}{1ex}{}
\titlespacing*{\section}{0pt}{*0}{0pt}
\titlespacing*{\subsection}{0pt}{\the\baselineskip}{0pt}

\usepackage{natbib} 
\bibpunct{(}{)}{;}{a}{,}{,}

\setlength{\bibsep}{0pt} \relax
\setcitestyle{notesep={: },yysep={, }} \relax

\usepackage{linguex,url}
\def\refdash{} % To suppress the dash in (2-a) etc. Defining both as different versions of linguex had different names for this command.
\def\firstrefdash{}

\usepackage{titleps}
\usepackage{lipsum}
\newpagestyle{mystyle}{
\sethead[\thepage][\Author][]{}{\Title}{\thepage}
}
\pagestyle{mystyle}
\author{First Author \and Second Author}
\title{The Title}
\makeatletter
\newcommand\Author{First Author---Second Author}
\let\Title\@title
\makeatother
%header with author and title

\pagenumbering{gobble} 
%suppresses page numbering in header

\begin{document}
%%\maketitle
\setlength{\Extopsep}{0pt}
\thispagestyle{empty}

{\large \textbf{Title}}\footnote{We would like to thank\ldots }\\
Firstname LASTNAME --- \textit{Affiliation}\\
Firstname LASTNAME --- \textit{Affiliation}\\

\textbf{Abstract.} This is the abstract.

\textbf{Keywords:} key, words, hello.

\section{Introduction}

This is where the introduction starts.

\subsection{Subsection}

This is what a subsection looks like. Here is an example, namely example \ref{Example1}.  I made it with \texttt{linguex.sty}.  If you want to make your examples with \texttt{expex} or \texttt{gb4e} or whatever, you can.

\ex. \label{Example1}I am an example \hfill (Hello!)\\ Maybe I need two lines.

\ex. Me, too.
	\a. \label{AnotherRef}I am \ref{AnotherRef}.
	\b. I am not.

\exg. De~temps~en~temps je m'ennuie de l'anglais.\\
occasionally I \textsc{me}.bore of the.English\\
`Occasionally I get bored of English.'

\ex.	
\ag. Maintenant, par exemple.\\
	now by example\\
	`Now, for example.'
\bg. I ara tamb\'{e}.\\
and now also \\
`And now, as well.'

Here are some examples of references in the main text: \citet{Montague73, HeimKratzer98}.  I could also cite them parenthetically \citep{Montague73, HeimKratzer98}.  I did this with \texttt{natbib.sty}.  Please use \texttt{natbib.sty}. Here are some examples of references in the main text: \citet{Dowty78}, \citet{HeimKratzer98}.  If you want to cite \citeauthor{kz84}'s \citeyearpar{kz84} paper possessively you can. Please note that \texttt{natbib.sty} truncates ``to appear'' \citep[as in][]{Montague73a} in citations; you must use a hack if you want to cite \nocite{Montague73a} Montague (to appear).

I could also cite them parenthetically Here are some examples of references in the main text: \citet{Montague73, HeimKratzer98}. 
\bibliographystyle{sub-chicago}
\bibliography{SuB}

\newpage
\subsection{Towards the end}

For other tasks, you can use pretty much any other package you like, so long as you don't alter the margins, spacing between paragraphs, font, font size, etc.  If you have any doubts, please contact:

\newpage

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.

\end{document}

### 2023-12-13 writing/SuB-paper/projection-operators-SuB.tex
started SuB paper draft

%!TEX TS-program = pdflatex
\documentclass[a4paper,12pt,twoside]{article}

\usepackage{mathptmx}
\usepackage{parskip}
\usepackage[top=1.5in,left=1in,right=1in,bottom=1in]{geometry}
%\setlength{\parindent}{0pt}
%\setlength{\parskip}{\the\baselineskip}
\pagestyle{empty}

\makeatletter
\renewcommand\@makefntext[1]{%
  \parindent 1em\noindent
  \hbox{\@makefnmark}#1}
\renewcommand\large{\@setfontsize\large{14pt}{18}} % For standardizing with the Word template: normally \large is 14.4pt
\makeatother
\usepackage{titlesec}
\titleformat{\subsection}{\normalfont\normalsize}{\thesubsection.}{1ex}{}
\titleformat{\subsubsection}{\normalfont\normalsize}{\thesubsubsection.}{1ex}{}
\titleformat{\section}{\normalfont\normalsize\bfseries}{\thesection.}{1ex}{}
% \titlespacing*{\section}{0pt}{*0}{0pt}
% \titlespacing*{\subsection}{0pt}{\the\baselineskip}{0pt}

\usepackage{natbib} 
\bibpunct{(}{)}{;}{a}{,}{,}

\setlength{\bibsep}{0pt} \relax
\setcitestyle{notesep={: },yysep={, }} \relax

\usepackage{linguex,url}
\def\refdash{} % To suppress the dash in (2-a) etc. Defining both as different versions of linguex had different names for this command.
\def\firstrefdash{}

\usepackage{titleps}
\usepackage{lipsum}
\newpagestyle{mystyle}{
\sethead[\thepage][\Author][]{}{\Title}{\thepage}
}
\pagestyle{mystyle}
\author{Hofmann \and de Marneffe  \and Tonhauser}
\title{Projection variability of clausal complements across different operators}
\makeatletter
\newcommand\Author{Hofmann, de Marneffe, and Tonhauser}
\let\Title\@title
\makeatother
%header with author and title

\pagenumbering{gobble} 
%suppresses page numbering in header

\input{lisa-extra}

\begin{document}
%%\maketitle
\setlength{\Extopsep}{0pt}
\thispagestyle{empty}

{\large \textbf{Projection variability of clausal complements across different operators}}\footnote{We would like to thank\ldots Taylor Mahler; Reviewers at SALT, SuB, XPrag; Audience at Syntax \& Semantics colloquium at University of Stuttgart}\\
Lisa HOFMANN --- \textit{University of Stuttgart}\\
Marie-Catherine DE MARNEFFE --- \textit{UC Louvain}\\
Judith TONHAUSER --- \textit{University of Stuttgart}\\

\textbf{Abstract.}
	We present experimental evidence that the projection of clausal complements \textbf{(i)} varies between entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend the result from \citet{degen_are_2022}, that projection ratings in polar questions are gradient in ways that cannot be predicted by categorical lexical classes, to contexts with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
	%
	The observed variability is not captured by existing theoretical accounts of projection
	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
	%
	Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.

\textbf{Keywords:} projection, commitment, discourse interpretation.

\section{Introduction}
	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \textit{projects} (\citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}, and many more).

	\ex. \label{ex:family}
		\a. \label{ex:neg}
			{\bf Negation:} \hfill
			\emph{\lq Cole didn't discover that Julian dances salsa.\rq}
		\b. \label{ex:q}
			{\bf Polar Question:} \hfill
			\emph{\lq Did Cole discover that Julian dances salsa?\rq}
		\c. \label{ex:mod}
			{\bf Modal:} \hfill
			\emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
		\d. \label{ex:cond}
			{\bf Conditional:} \hfill
			\emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
		\z.
	\z.

	We investigate the question whether the projection of embedded propositional content across entailment-cancelling operators differs by operator, and whether by-operator differences vary between embedding predicates.
	%
	Textbook diagnostics for projection often assume that, if a content projects across one of the operators in \Last, it uniformly projects across the others (\citealt{chierchia_meaning_1990}). Accordingly, current research examines projection variability between these operators only rarely, and with conflicting findings.

	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \textit{be annoyed, regret, reveal, know}), where the CC projects across all four operators, and semi-factives (e.g., \textit{discover, realize, see, notice, find out}) where it always projects from under negation, but not always from under polar questions, modals, or conditionals.

	\citet{smith_relationship_2014} investigated by-operator variation, comparing negation and conditionals, for various types of projective content. They found that the projective content of epithets (e.g. \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals, whereas that of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and that of clefts showed no difference. Their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance. Since other factors besides speaker commitment modulate discourse (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.

	\posscite{sieker_projective_2022} study on German clause-embedding predicates compared the projection of purported factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{enthüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') across the four contexts in \Last. They replicated Smith and Hall's result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know'). However, this finding was part of an overall pattern of higher projection ratings with negation than other operators. They found no interaction with predicate type, and therefore no evidence for the distinction of factives vs. semi-factives. This is based on an analysis which aggregated the ratings for the four purported factive and semi-factive predicates, under each of the entailment-cancelling operators.

	Studies in \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} observed by-predicate variation in polar questions, where Karttunen's classification predicts more projection for factives than semi-factives.
	%
	\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive). But here, it is also not obvious how exactly this task relates to projection.
	%
	\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC. The observed differences between individual clause-embedding predicates do not align with theoretically assumed lexical classes. While Degen \& Tonhauser demonstrate this for the distinction between purported factive and non-factive predicates, the results also did not match the expectations from Karttunen's factive vs. semi-factive classification.
	%
	Using speaker-certainty ratings, \citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}. \citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}. These findings suggest that individually examining projection ratings for various predicates, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate variation in operator effects on projection.

	While \citet{smith_relationship_2014} found by-predicate variation in the effect of the operator on projection, \citet{sieker_projective_2022} did not.
	%
	These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis. To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}. We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}), and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).

\section{Assessing projection across entailment-canceling operators}

	To assess the effect of entailment-cancelling operators and clause-embedding predicates on the projection of the embedded content, we collected projectivity judgments for contents embedded under 20 clause-embedding predicates in four sets of experiments. The predicates were embedded under negation in Exps.~1, under polar questions (Exps.~2), under the epistemic possibility modal {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4). (Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block. We focus on the projection ratings here.)

		In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).

		\begin{itemize}
			\item say more about task
		\end{itemize}

		Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.

	\subsection{Experimental method}

		\subsubsection{Participants}
			We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform. (roughly 750 participants in each experiment)

		\subsubsection{Materials}
			Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.

			\begin{itemize}
				\item 20 clause-embedding predicates that have shown projection variability in question contexts (Degen \& Tonhauser, 2022)
				\item Crossed w/ 20 CCs: 20 x 20 = 400 combinations
				\item One experiment per operator
				\item six main-clause controls for data exclusion
			\end{itemize}

		\subsubsection{Procedure}

			On each trial, participants read an utterance, which was presented as an utterance of a speaker with a randomly chosen name, as well as the corresponding projection question, and then gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other. This is illustrated in Fig.~\ref{fig:trial}.

			Following \citet{tonhauser_how_2018}, we took a `yes' response to indicate that participants interpret the utterance in a way that the presented speaker (e.g., Christopher in Fig.~\ref{fig:trial}) is committed to the embedded content, which reflects projection. A `no' response was taken to indicate that the content does not project.

			\begin{figure}[ht]
				\centering
				\includegraphics[width = .7\linewidth]{task-1n-proj.eps}
				\caption{A sample trial in Exp.~1. In the corresponding trials in Exps.~2-4, participants were presented with an utterance involving a different entailment-cancelling operator.}
				\label{fig:trial}
			\end{figure}

	\subsection{Results and discussion}
		Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator and \textbf{(ii)} there is by-predicate variation in the effect of operator on projection.

		\subsubsection{By-operator variation}

			We find an effect of operator when aggregating across predicates. Mean projection ratings were higher under question-embeddings than under negation and modals, but lower than in conditional antecedents. This is shown in Fig.~\ref{fig:op-ratings}, and the generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.

			\begin{figure}[ht]
				\centering
				\includegraphics[width = .35\linewidth]{projective-op}
				\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals.}
				\label{fig:op-ratings}
			\end{figure}

			\begin{table}[ht]
					\centering
					\hspace{-1.3em}
					\begin{tabular}{llrrrr}
						Model & & Estimate & Std. Error & t-value\\
						\midrule
						\#1 & Intercept: \emph{question} & 0.51 & 0.01 & 44.78 & ***\\
						& operator: conditional & 0.05 & 0.01 & 5.30 & ***\\
						& operator: modal & -0.04 & 0.01 & -4.45 & ***\\
						& operator: negation & -0.03 & 0.01 & -4.67 & ***\\
						\bottomrule
					\end{tabular}

					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fit with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
				\end{table}

			This result differs from that of \cite{sieker_projective_2022}, who observed that the CC of German attitude predicates projects more from under negation than the other operators.

			\begin{itemize}
				\item but like them, we observe small differences 
				\item overall, they have much higher projection ratings
				\item but they chose a different set of predicates, we can understand these different results when taking into account by-predicate variation in the effect of operator
			\end{itemize}

		\subsubsection{By-predicate variation in the effect of operator}

			We observe differences across the predicates in by-operator variation. This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
			%
			For instance, whereas the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals. The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
			%
			These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).

			\begin{figure}[ht]
				\centering
				\includegraphics[width = \linewidth]{projective-pred-op}
				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
				\label{fig:op-pred-ratings}
			\end{figure}

			\begin{table}[ht]
				\centering
				\begin{tabular}{llrrrr}
					Model & & Estimate & Std. Error & t-value\\
					\midrule
					\#2 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01 & 79.86 & ***\\
					& operator: conditional & -0.12 & 0.02  & -7.36 & ***\\
					& operator: modal & -0.16 & 0.02  & -10.01 & ***\\
					& operator: question & 0.02 & 0.01 & 1.72 & n.s.\\
					\midrule
					\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01 & 62.70 & ***\\
					& operator: conditional & 0.11 & 0.02 & 7.11 & ***\\
					& operator: modal & -0.06 & 0.02 & -3.63 & ***\\
					& operator: question & 0.10 & 0.01 & 7.08 & ***\\
					\midrule
					\#4 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 & 72.97 & ***\\
					& operator: conditional & 0.00 & 0.02 & -0.06 & n.s.\\
					& operator: modal & -0.14 & 0.02 & -9.18 & ***\\
					& operator: question & 0.08 & 0.01 & 5.67 & ***\\
					\bottomrule
				\end{tabular}
				\caption{\small Excerpt of the output from three linear mixed effects models, with fixed effects: operator, predicate, and their interaction; random effect: participant intercepts.
				Models were fit with \texttt{lme4, lmertest} in \texttt{R}. Models \textbf{\#2--4} also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (not shown here).\label{t:models}}
			\end{table}

			This result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals. This differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
			%
			Given that the methods and the set of projective contents of our experiment were very similar to theirs, this difference may suggest cross-linguistic variation in projection variability (cf. \citealt{tonhauser_projection_2020}).

		\subsubsection{Gradience and lexical classes}

			Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions. 
			E.g., the CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.

			Similarly, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates, in line with \citet{sieker_projective_2022} (see also \citealt{beaver_have_2010}): Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
			%
			Further, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators. The pattern observed for {\em know} does not fit into either category.

		\subsubsection{Converging evidence for operator / predicate interactions}
			MegaVeridicality dataset (\citealt{white_role_2018}): 517 predicates in three sentence types

			\ex. \a. Somebody didn’t know that a particular thing happened. (Did that thing happen?)
				\b. If somebody knows that a particular thing happened, did that thing happen?
				\b. If somebody didn’t know that a particular thing happened, did that thing happen?
				\z.
			\z.

			\begin{figure}[ht]
				\centering
				\includegraphics[width = \linewidth]{mega-veridicality}
				\caption{Mean projection ratings by embedding context and predicate.}
				\label{fig:figure1}
			\end{figure}

		\subsubsection{Interim conclusion}

			Results:

			\begin{enumerate}
				\item Effect of operator on projection

				\item Effect of operator on projection differs by predicate

				\item By-predicate variation on operator effects are gradient and cannot be predicted by categorical lexical classes

			\end{enumerate}

			Methodological implications:
			\begin{itemize}
				\item projection ratings do not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.

				\item Additionally, claims about projection variability must be relativized to the entailment-canceling operator.

				\item but we can still use FoS test in teaching, JT made this claim, but I don't understand the argument... help?

			\end{itemize}

			Theoretical implications: Accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability. The following section argues that our results provide evidence for a triggering algorithm based on contextual rather than lexical or conceptual information (e.g., in the sense of \citealt{abrusan_predicting_2011,simons_best_2017}), and that the input to this triggering algorithm should be contextual rather than lexical inferences (e.g., in the sense of \citealt{schlenker_triggering_2021}).

\section{Theoretical implications}
	Our results---that projection is modulated by entailment-canceling operators, that there is by-predicate variation in the effect of operator on projection, and that this variation is gradient in ways that cannot be predicted by categorical lexical classes---are not captured by contemporary projection analyses, for several reasons.

	Here, we discuss (i) dynamic approaches, which assume that projection is determined lexically for factive predicates (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) approaches which assume that projection is determined by a contextual triggering algorithm that operates over the lexical entailments of an utterance (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) approaches with a conceptual triggering algorithm that operates over contextual entailments of an utterance (\citealt{schlenker_triggering_2021}). While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).

	- effect of different entailment-cancelling operators, not captured (heim vs santd, if projects projects; abtrusan operators?; schlenker can be imagined) in all systems it is conceivable, but no systemaitc itneraction spelled out

	- rely on lexical classes, factives, veriudical, no predictions of non-veridicals

	- not capture how the operator differentially affects different predicates (contextual entailment may provied an approach, but also needs to be spelled out yet)

	The first is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators. While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out systematically.

	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}). A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed. In a full talk, we would like to discuss how this analysis might be extended to capture the gradient projection observed in our experiment.

	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed). Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.

	\subsection{Factivity: Lexical triggering}

		In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.

	\subsection{Entailments and contextual triggering}

	\subsection{Contextual entailment and conceptual triggering}

	\subsection{Conclusion: Contextual entailment and contextual triggering}
		\begin{itemize}
			\item contextual entailments (but weaker notion of entailment)

			\begin{itemize}

				\item cant believe its not lexical roberts

				\item pragmatics of full embedding context; Does Cole think that Julian dances salsa?

				\item why would that be the case?

			\end{itemize}

			\item contextual triggering

			\begin{itemize}

				\item what is the difference between know and prove (in terms of conceptual precondition, causal model?) there is no conceptually inherent relation between the two, the difference is lexical

				\item QUD influences

				\item what is the relationship between lexical differences and QUD?

			\end{itemize}

		\end{itemize}

\section{Lexical differences}

	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting initial generalizations over lexical properties, indicated in \textbf{Figure\ 2}.

	So, can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
	%
	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality associated with these predicates.
	%
	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf `Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
	%
	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This {\bf `Conditional high'} pattern (c) may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
	%
	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern (d). The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.

	%%%%
	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns. \textbf{Fig.~\ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, identifying some groups of predicates that show similar by-operator variation. We highlight four \lq projection-profiles\rq here: \emph{pretend} and \emph{think} are the only predicates that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}). \emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}). \emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}). Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).

\section{Conclusion}

\newpage
\bibliographystyle{sub-chicago}
\bibliography{projection}

\end{document}

### 2024-02-05 writing/SuB-paper/projection-operators-SuB.tex
sub paper edits before overleaf

@@
 
-{\large \textbf{Projection variability of clausal complements across different operators}}\footnote{We would like to thank\ldots Taylor Mahler; Reviewers at SALT, SuB, XPrag; Audience at Syntax \& Semantics colloquium at University of Stuttgart}\\
+{\large \textbf{Projection variability of clausal complements across different operators}}\footnote{We would like to thank\ldots }\\
 Lisa HOFMANN --- \textit{University of Stuttgart}\\
@@
 \textbf{Abstract.}
-	We present experimental evidence that the projection of clausal complements \textbf{(i)} varies between entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend the result from \citet{degen_are_2022}, that projection ratings in polar questions are gradient in ways that cannot be predicted by categorical lexical classes, to contexts with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
+	We present experimental evidence that the projection of clausal complements \textbf{(i)} varies among entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend the result from \citet{degen_are_2022}, that projection ratings in polar questions are gradient in ways that cannot be predicted by categorical lexical classes, to contexts with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
 	%
@@
 \section{Introduction}
-	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
-	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \textit{projects} (\citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}, and many more).
+	Language users may infer that a speaker who uses a clause-embedding predicate, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,}) even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}, in which case we say that it \textit{projects} (e.g., \citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}).
 
@@
 
-	We investigate the question whether the projection of embedded propositional content across entailment-cancelling operators differs by operator, and whether by-operator differences vary between embedding predicates.
+	We investigate whether the projection of embedded propositional content from under entailment-cancelling operators differs by operator and whether by-operator differences vary between embedding predicates.
 	%
-	Textbook diagnostics for projection often assume that, if a content projects across one of the operators in \Last, it uniformly projects across the others (\citealt{chierchia_meaning_1990}).
-Accordingly, current research examines projection variability between these operators only rarely, and with conflicting findings.
+	Textbook diagnostics for projection often assume that if a content projects across one of the operators in \Last, it uniformly projects across the others (e.g., \citealt{chierchia_meaning_1990}).
+Accordingly, current research examines projection variability between these operators only rarely and with conflicting findings.
 
@@
 	\citet{smith_relationship_2014} investigated by-operator variation, comparing negation and conditionals, for various types of projective content.
-They found that the projective content of epithets (e.g.
-\textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals, whereas that of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and that of clefts showed no difference.
-Their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance.
+They found that the expressive content of epithets (e.g.
+\textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals, whereas appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference.
+However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance.
 Since other factors besides speaker commitment modulate discourse (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.
@@
 However, this finding was part of an overall pattern of higher projection ratings with negation than other operators.
-They found no interaction with predicate type, and therefore no evidence for the distinction of factives vs.
+	%
+	Importantly, their analysis did not explore the impact of operators on each predicate individually.
+Instead, it grouped ratings for the four purported factive and semi-factive predicates, respectively.
+Based on this analysis, they found no interaction of operator with predicate type, and thus no evidence for the distinction of factives vs.
 semi-factives.
-This is based on an analysis which aggregated the ratings for the four purported factive and semi-factive predicates, under each of the entailment-cancelling operators.
 
@@
 	\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-But here, it is also not obvious how exactly this task relates to projection.
+But here, it is also not obvious how exactly this task relates to projection.\footnote{Say why}
+
+	\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured speaker commitments towards a broad range of projective contents more directly, by collecting judgments of how certain participants take a speaker to be of the contents under investigation.
 	%
-	\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
-The observed differences between individual clause-embedding predicates do not align with theoretically assumed lexical classes.
-While Degen \& Tonhauser demonstrate this for the distinction between purported factive and non-factive predicates, the results also did not match the expectations from Karttunen's factive vs.
-semi-factive classification.
+	The observed differences between CCs of various clause-embedding predicates do not align with theoretically assumed lexical classes.
+While \citet{degen_are_2022} demonstrate this for the distinction between purported factive and non-factive predicates, the results did not match the expectations from Karttunen's factive vs.
+semi-factive classification either.
 	%
-	Using speaker-certainty ratings, \citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}.
+	\citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}.
 \citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}.
-These findings suggest that individually examining projection ratings for various predicates, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate variation in operator effects on projection.
+These findings suggest that examining projection ratings for various predicates individually, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate projection variation.
 
@@
 To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}.
-We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}), and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
+We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive and semi-factive predicates, and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
 
@@
 	To assess the effect of entailment-cancelling operators and clause-embedding predicates on the projection of the embedded content, we collected projectivity judgments for contents embedded under 20 clause-embedding predicates in four sets of experiments.
-The predicates were embedded under negation in Exps.~1, under polar questions (Exps.~2), under the epistemic possibility modal {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
-(Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block.
-We focus on the projection ratings here.)
+The predicates were embedded under negation in Experiments~1, under polar questions (Exps.~2), under the epistemic possibility modal {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
+Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block.
+We focus on the projection ratings here.
 
-		In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
+		Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} Participants were presented with utterances like those in \Next and judged whether the speaker (who was named) was certain of the CC (e.g.: Is Christopher certain that Julian dances salsa?).
 
-		\begin{itemize}
-			\item say more about task
-		\end{itemize}
+		\ex. \label{ex:certain-that}
+			\a. Christopher: \emph{\lq Cole didn't discover that Julian dances salsa.\rq}
+			\b. Christopher: \emph{\lq Did Cole discover that Julian dances salsa?\rq}
+			\c. Christopher: \emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
+			\d. Christopher: \emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
+			\z.
+			Projection question: Is Christopher certain that Julian dances salsa?
+		\z.
 
+		Following \citet{tonhauser_prosodic_2016,djarv_prosodic_2017,tonhauser_how_2018,lorson_influence_2018,mahler_does_2019}, we took a `yes' response to indicate that participants interpret the utterance in a way that the presented speaker (e.g., Christopher in \Last) is committed to the embedded content, which reflects projection.
+A `no' response was taken to indicate that the content does not project.
+		%
 		Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.
@@
 		\subsubsection{Participants}
-			We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform. (roughly 750 participants in each experiment)
+			In each of the 12 experiments, 250-300 inhabitants of the USA participated via Prolific or Amazon's MT platform. 
+			%
+			We analyze the data from $2,682$ self-reported native speakers of American English, after excluding participants who failed one of the following two attention-checks: their mean rating on six control items is more than 2 sd different from the group mean, or they always roughly selected the same point on the scale for target stimuli (identified by low variance, and then manually inspecting their response patterns).
 
-		\subsubsection{Materials}
-			Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
+		\subsubsection{Materials and procedure }
+			We tested 20 clause-embedding predicates \Next which have shown projection variability in question contexts (\citealt{degen_are_2022}).
+They comprise three purported factive predicates \Next[a], two purported semi-factives \Next[b], nine that have been characterized as optionally factive in \citealt{kiparsky_fact_1970} \Next[c], and six non-factives which include two veridical predicates \Next[d], and four non-veridical ones \Next[e].
 
-			\begin{itemize}
-				\item 20 clause-embedding predicates that have shown projection variability in question contexts (Degen \& Tonhauser, 2022)
-				\item Crossed w/ 20 CCs: 20 x 20 = 400 combinations
-				\item One experiment per operator
-				\item six main-clause controls for data exclusion
-			\end{itemize}
+			\ex. 20 clause-embedding predicates \hfill \citealt{degen_are_2022}
+				\a. canonically factive:
+					\hfill \emph{be annoyed, know, reveal}
+				\b. semi-factive:
+					\hfill \emph{discover, see}
+				\b. optionally factive: \newline
+					\phantom.\hfill \emph{acknowledge admit, announce, confess, confirm, establish, hear, inform, prove}
+				\b. Non-factive veridical:
+					\hfill \emph{be right, demonstrate}
+				\b. Non-veridical:
+					\hfill \emph{pretend, say, suggest, think}
+				\z.
+			\z.
 
-		\subsubsection{Procedure}
+			Each participant saw all 20 predicates, randomly paired with a unique embedded clause from a set of 20 contents, and a subject from a set of 20 proper names.
+The predicates associated with projective content were embedded under (i) negation, (ii) polar questions, (iii) the epistemic possibility modal \textit{perhaps}, and (iv) in conditional antecedents in the four different sets of experiments.
 
-			On each trial, participants read an utterance, which was presented as an utterance of a speaker with a randomly chosen name, as well as the corresponding projection question, and then gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
-This is illustrated in Fig.~\ref{fig:trial}.
-
-			Following \citet{tonhauser_how_2018}, we took a `yes' response to indicate that participants interpret the utterance in a way that the presented speaker (e.g., Christopher in Fig.~\ref{fig:trial}) is committed to the embedded content, which reflects projection.
-A `no' response was taken to indicate that the content does not project.
+			On each trial, participants read an utterance and the corresponding projection question.
+They gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
+This is illustrated in Figure~\ref{fig:trial}.
+Following, e.g., \citet{tonhauser_how_2018}, we take high ratings of speaker certainty to reflect speaker greater speaker commitment towards the CC.
 
@@
 
+			Each experiment also included six control stimuli, which were main clauses with appositive relative clauses, which were also presented as utterances made by a named speaker.
+Here, the main clause content was targeted by the projection question.
+In the experiments with polar questions, the controls were expected to receive low certainty ratings, whereas in the experiments with declaratives, they were expected to receive high ratings.
+
+			Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else.
+Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates) and the same 6 (experiment-specific) control stimuli.
+Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
+Block order and within-block trial order was randomized.
+After completing the experiment, participants filled out a short optional demographic survey.
+
 	\subsection{Results and discussion}
-		Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator and \textbf{(ii)} there is by-predicate variation in the effect of operator on projection.
+		Our data reveals three key results: \textbf{(i)} There is projection variability by operator; \textbf{(ii)} There is by-predicate variation in the effect of operator on projection; and \textbf{(iii)} We find further support (from entailment-cancelling operators beyond polar questions) for \posscite{degen_are_2022} result that by-predicate projection variation is gradient in a way that cannot be predicted by categorical lexical classes.
 
 		\subsubsection{By-operator variation}
-
-			We find an effect of operator when aggregating across predicates.
-Mean projection ratings were higher under question-embeddings than under negation and modals, but lower than in conditional antecedents.
-This is shown in Fig.~\ref{fig:op-ratings}, and the generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.
+			When aggregating across predicates, we found an effect of \texttt{operator}.
+The distribution of projection ratings, as well as means and $95\%$ bootstrapped confidence intervals by operator are shown in Figure~\ref{fig:op-ratings}.
 
@@
 				\centering
-				\includegraphics[width = .35\linewidth]{projective-op}
-				\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals.}
+				\includegraphics[scale = .8]{operator-graph-1}
+				\caption{Certainty ratings by operator with means and $95\%$ bootstrapped confidence intervals.}
 				\label{fig:op-ratings}
 			\end{figure}
+
+			Mean projection ratings were higher for CCs in question-embeddings than under negation and modals, but lower than in conditional antecedents.
+These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.
 
@@
 
-					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fit with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
+					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fitted with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
 				\end{table}
 
-			This result differs from that of \cite{sieker_projective_2022}, who observed that the CC of German attitude predicates projects more from under negation than the other operators.
+			Like \citet{sieker_projective_2022}, we observe a small main effect of entailment-cancelling operator on projection.
+However, our result, that the CC of clause-embedding predicates projects most from conditional antecedents and least from under negation and modals, differs from their finding that the CC of the German predicates they tested projects more from under negation than the other operators.
+Further, the projection ratings they report are overall quite a bit higher than the ones reported here (their overall mean is at around $0.855$, while ours is at around $0.501$).
+			%
+			We can understand these differences based on an effect of \texttt{predicate}, established in the upcoming subsections, and a different set of predicates tested in the two studies.
+While Sieker \& Solstad tested only predicates that are characterized as factive or semi-factive in the literature, yielding relatively high projection ratings, we also included ones that are usually characterized as non-factive, lowering the overall mean projection ratings.
+The difference in the direction of the effect of the operator between the two studies can also be tied to a different set of predicates tested, because the effect of operator on projection differs by predicate.
 
-			\begin{itemize}
-				\item but like them, we observe small differences 
-				\item overall, they have much higher projection ratings
-				\item but they chose a different set of predicates, we can understand these different results when taking into account by-predicate variation in the effect of operator
-			\end{itemize}
+		\subsubsection{Gradient by-predicate variation}
+
+			Taking into account the effect of \texttt{predicate} on projection, we find gradient projection variation, like \citet{tonhauser_how_2018} and \citet{degen_are_2022}.
+			%
+			This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 predicates grouped by embedding operator.
+Here, predicates on the x-axis are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
+
+			\begin{figure}[ht]
+				\centering
+				\includegraphics[width = \linewidth]{predicate-operator-graph-1}
+				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
+Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
+				\label{fig:op-pred-ratings}
+			\end{figure}
+
+			Our results therefore provide further support from negation, modals, and conditionals for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions.
+The CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
 
@@
 
-			We observe differences across the predicates in by-operator variation.
-This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
+			We observe differences between predicates in by-operator variation, i.e., a operator/predicate interaction effect, which is also illustrated in Figure~\ref{fig:op-pred-ratings}.
 			%
@@
 			These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).
-
-			\begin{figure}[ht]
-				\centering
-				\includegraphics[width = \linewidth]{projective-pred-op}
-				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
-Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
-				\label{fig:op-pred-ratings}
-			\end{figure}
 
@@
 
-			This result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals.
-This differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
+			Our result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals (we found no difference here).
+Our results differ from those of \citet{sieker_projective_2022}, who found no difference between the two groups of predicates they assumed in the effect of operator.
 			%
-			Given that the methods and the set of projective contents of our experiment were very similar to theirs, this difference may suggest cross-linguistic variation in projection variability (cf.
-\citealt{tonhauser_projection_2020}).
-
-		\subsubsection{Gradience and lexical classes}
-
-			Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions.
-			E.g., the CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
-
-			Similarly, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates, in line with \citet{sieker_projective_2022} (see also \citealt{beaver_have_2010}): Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
+			However in line with Sieker \& Solstad, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates (see also \citealt{beaver_have_2010}).
+Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates.
 			%
-			Further, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+			Further, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
 The pattern observed for {\em know} does not fit into either category.
@@
 		\subsubsection{Converging evidence for operator / predicate interactions}
-			MegaVeridicality dataset (\citealt{white_role_2018}): 517 predicates in three sentence types
 
-			\ex. \a. Somebody didn’t know that a particular thing happened. (Did that thing happen?)
+			Our experimental result that that the effect of entailment-cancelling operator differs by predicate finds further support in the data from \posscite{white_role_2018} MegaVeridicality dataset.
+			%
+			White \& Rawlins assess projection inferences associated with 517 clause-embedding predicates by presenting them in 50 \lq low-content\rq\ syntactic frames, like the ones in \Next, illustrated here with \textit{know}.
+The frames in \Next present predicates associated with projective contents embedded under negation \Next[a], in conditional antecedents \Next[b], or both \Next[c], in combination with a projection question (\textit{Did that thing happen?}).
+
+			\ex. \a. Somebody didn’t know that a particular thing happened. Did that thing happen?
 				\b. If somebody knows that a particular thing happened, did that thing happen?
@@
 
+			In a series of experiments, participants presented with stimuli like \Last responded with \lq yes\rq, \lq no\rq, or  \lq maybe\rq (coded as \texttt{1, -1,} or \texttt{0}, respectively).
+Here, we present data from their data set, for the frames in \Last, and 16 of their predicates.
+These include ones also used in our experiment as purported factives (\textit{be annoyed, know, reveal}) and semi-factives (\textit{discover, see}), and eleven further predicates commonly characterized as factive or semi-factive (\textit{amuse, find out, forget, learn, love, notice, realize, recognize, regret, remember, understand}).
+
+			The mean projection ratings by embedding context and predicate are presented in Figure~\ref{fig:figure4}, which shows that effect of embedding context on projection differs by predicate.
+
+			For instance,\footnote{Give some examples of differences here, but first double check if lines are labeled correctly. Also angle labels on x-axis.}
+
+			% the predicates \textit{see} and \textit{learn} have the same mean ratings for both negative and conditional contexts.
+However, the effect of negative conditional contexts does not simply reflect the additive result of both effects, but
+
 			\begin{figure}[ht]
@@
 				\caption{Mean projection ratings by embedding context and predicate.}
-				\label{fig:figure1}
+				\label{fig:figure4}
 			\end{figure}
+
+			It is worth noting that the task assesses global projection only in case of \Last[a], but not \Last[b+c], where the question \textit{did it happen?} is embedded in the conditional consequent.
+It assesses whether the projective inference holds in the conditional local context.
+Nevertheless, the data provides supporting evidence for by-predicate variation in the effect of operators on projection.
+Crucially, the by-predicate variation in the effect of negation in the global context is different from the by-predicate variation in the effect of negation in the conditional context.
 
@@
 
-			Results:
-
-			\begin{enumerate}
-				\item Effect of operator on projection
-
-				\item Effect of operator on projection differs by predicate
-
-				\item By-predicate variation on operator effects are gradient and cannot be predicted by categorical lexical classes
-
-			\end{enumerate}
-
-			Methodological implications:
-			\begin{itemize}
-				\item projection ratings do not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.
-
-				\item Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
-
-				\item but we can still use FoS test in teaching, JT made this claim, but I don't understand the argument... help?
-
-			\end{itemize}
-
-			Theoretical implications: Accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability.
-The following section argues that our results provide evidence for a triggering algorithm based on contextual rather than lexical or conceptual information (e.g., in the sense of \citealt{abrusan_predicting_2011,simons_best_2017}), and that the input to this triggering algorithm should be contextual rather than lexical inferences (e.g., in the sense of \citealt{schlenker_triggering_2021}).
+			Based on our data, we conclude that embedded propositional content projects differently from under different entailment-cancelling operators.
+A methodological implication of this finding is that claims about projection variability must be relativized to the entailment-canceling operator in future research.
+			%
+			Further, the observed by-predicate variation does not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.
+			%
+			Finally, the effect of operator on projection differs by predicate.
+The theoretical implications of our findings are discussed in the following section.
+Importantly, accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability.
 
 \section{Theoretical implications}
-	Our results---that projection is modulated by entailment-canceling operators, that there is by-predicate variation in the effect of operator on projection, and that this variation is gradient in ways that cannot be predicted by categorical lexical classes---are not captured by contemporary projection analyses, for several reasons.
+	Our results---that projection is modulated by entailment-canceling operators, that there is by-predicate variation in the effect of operator on projection, and that this variation is gradient in ways that cannot be predicted by categorical lexical classes---are not captured by contemporary projection analyses.
 
-	Here, we discuss (i) dynamic approaches, which assume that projection is determined lexically for factive predicates (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) approaches which assume that projection is determined by a contextual triggering algorithm that operates over the lexical entailments of an utterance (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) approaches with a conceptual triggering algorithm that operates over contextual entailments of an utterance (\citealt{schlenker_triggering_2021}).
+		Here, we discuss this for (i) dynamic approaches, which assume that projection is determined lexically for factive predicates (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) approaches which assume that projection is determined by a triggering algorithm based on information structure that operates over the literal entailments of an utterance (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) approaches with a triggering algorithm based on world-knowledge that operates over contextual entailments of an utterance (\citealt{schlenker_triggering_2021}).
 While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).
 
-	- effect of different entailment-cancelling operators, not captured (heim vs santd, if projects projects; abtrusan operators?; schlenker can be imagined) in all systems it is conceivable, but no systemaitc itneraction spelled out
+	\subsection{Lexical triggering in dynamic semantics}
 
-	- rely on lexical classes, factives, veriudical, no predictions of non-veridicals
+		Dynamic semantic accounts of projection (e.g., \citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}) assume lexical triggering: there is a class of (semi-)factive predicates, which lexically presuppose their CCs.
+Presupposed CCs project to the global context, except when that would produce an inconsistency, in which case they are accommodated to the local context of the operator.
 
-	- not capture how the operator differentially affects different predicates (contextual entailment may provied an approach, but also needs to be spelled out yet)
+		\paragraph{By-operator variation.} In our study, the stimulus utterances were presented out of the blue, leaving it to the participants to imagine a suitable discourse context.
+Therefore, none of the CCs should be interpreted as necessarily inconsistent with contextual information.
+		%
+		Consequently, dynamic approaches do not lead us to expect differential effects of entailment-cancelling operators on projection in out-of-the blue contexts.
+Instead, negation and conditional antecedents are given a semantics that derives their behavior as presuppositional \lq holes\rq\ (in the sense of \citealt{karttunen_observations_1971}).
+For instance in \citealt{heim_projection_1983}, presuppositional expressions in negative prejacents and conditional antecedents are evaluated relative to the global context set.
+Therefore, presuppositional expressions in these contexts are predicted to behave like unembedded ones, with no expected differences.
+While the approach does not explicitly address epistemic possibility modals or polar questions, we might expect that they would be treated as presuppositional holes along the same lines.
+		%
+		Differential effects of operators on projection could conceivably be addressed in such a system by assuming that the meaning of entailment-cancelling operators systematically affects the possibility of local accommodation (e.g., by assuming that accommodation is more likely under modals than conditional antecedents).
+However, no such effects have been spelled out.
 
-	The first is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
-While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out systematically.
+		\paragraph{Gradient by-predicate variation.} The assumption of categorical lexical triggering does not address the projection variation among purported factive predicates (see also discussion in \citealt{degen_are_2022}).
+It also cannot address the fact that the CCs of some non-factive predicates (e.g., \emph{inform, acknowledge}) project just as much, or more, than some (semi-)factive predicates (e.g., \emph{reveal, discover, see}).
+That is because, it only makes predictions for the projection of the CC of (semi-)factive predicates, whose CCs are analyzed as presuppositions.
+It therefore does not address the data for the non-veridical predicates in our study.
+		%
+		Further, the categorical lexical distinctions are not sufficiently fine-grained to make predictions about the gradient by-predicate variation in the effect of operator on projection.
+		%
+		This interaction effect can only be addressed by taking into account more fine-grained lexical distinctions than assuming that some predicates are presuppositional, while others are not.
 
-	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
-A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
-In a full talk, we would like to discuss how this analysis might be extended to capture the gradient projection observed in our experiment.
+	\subsection{Triggering based on information structure} % (fold)
+		The information-structural triggering approaches in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} assume that projection is not determined lexically, but by a triggering mechanism, which operates over the entailments of an expression, and determines which of them projects.
+They assume entailments project that do not contribute to the main point invoked by the utterance.
 
-	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
-Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
+		% Instead of assuming that the possibility of projection is determined lexically, the accounts in \cite{abrusan_predicting_2011,simons_best_2017} and \citet{schlenker_triggering_2021} assume a triggering mechanism, which operates over a set of inferences associated with an utterance, and determines which of them projects.
+The assumed triggering mechanisms differ on two dimensions: (i) how the set of inferences is determined that constitutes the input to the triggering mechanism, and (ii) the algorithm by which these inferences are derived as projective or not.
 
-	\subsection{Factivity: Lexical triggering}
+		For instance, \citet{simons_best_2017} operationalize the main point of an utterance in terms of at-issueness relative to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}).
+		%
+		In the QUD-based approach to information-structure in discourse, each utterance addresses a question, which is determined by its alternative-structure, which, can be indicated by focus-marking.
+For example, in a modalized utterance with \emph{be annoyed}, the CC can be focused \Next[b], which leads to the alternative structure in \Next[c], allowing the utterance to address a question like \Next[a].
 
-		In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
+		\ex. \a. What was Cole annoyed about today?
+			\b. Perhaps Cole was annoyed [$_F$ that Julian dances salsa].
+			\b. $\{p \mid \exists q(p = \lambda w.$\emph{Cole was annoyed that} $q$ \emph{in} $w)\}$
+			\z.
+		\z.
 
-	\subsection{Entailments and contextual triggering}
+		In contrast, focus on the subject \Next[b] leads to the alternative structure in \Next[c], allowing the utterance to address a question like \Next[a].
 
-	\subsection{Contextual entailment and conceptual triggering}
+		\ex. \a. Who was annoyed today?
+			\b. Perhaps [$_F$ Cole] was annoyed that Julian dances salsa.
+			\b. $\{p \mid \exists x(p = \lambda w.x$ \emph{was annoyed in} $w$ \emph{that Julian dances salsa}$)\}$
+			\z.
+		\z.
 
-	\subsection{Conclusion: Contextual entailment and contextual triggering}
-		\begin{itemize}
-			\item contextual entailments (but weaker notion of entailment)
+		The account assumes that an inference projects if it is entailed by the union of the alternatives (see also \citealt{abusch_presupposition_2010}).
+therefore predicting projection of the CC in \Last, but not \LLast.
+That is, under the assumption that $x$ \emph{was annoyed that} $p$ entails $p$.
 
-			\begin{itemize}
+		\paragraph{Gradient by-predicate variation.} The triggering mechanisms assumed in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017}, predict a potential of projection for inferences that are entailed by the literal content of the modal prejacent $p$ in \Last and \LLast, and similarly for propositions embedded in negative contexts, conditional antecedents and polar questions.
+These accounts, therefore, predict a potential of projection for CCs that are entailed by the prejacent, due to the lexical entailments of veridical predicates, or due to entailments that are contributed compositionally by the literal content (see e.g., \citealt{roberts_i_2019}).
+However, they do not make systematic predictions for CCs, such as in \Last, that are not entailed.
+As a result, these accounts cannot address the fact that we see some amount of projection for all of the tested predicates.
 
-				\item cant believe its not lexical roberts
+		Because the main point of an utterance is often determined in the discourse context (\citealt{roberts_information_1996,roberts_information_2012}, \emph{et seq.}), these approaches incorporate discourse effects on projection.
+		%
+		% \citet{abrusan_predicting_2011} also makes predictions for utterances in out-of-the-blue contexts: Utterances are associated with a default main point that is grammatically determined by the time at which the matrix predicate is interpreted.
+		%
+		The account in \citealt{simons_best_2017} could be extended to address our gradient data in out-of-the blue contexts, by assuming (i) that embedded contents that are not at-issue can project even if they are not entailed, and (ii) that linguistic expressions differ in how likely the contributed contents are interpreted as at-issue (as suggested in \citealt{tonhauser_how_2018}).
+		%
+		In our data, for instance, the CC of \emph{discover} \Next[a] has received relatively high projection ratings, while that of \emph{confirm} \Next[b] has received medium to low ratings.
+To explain this difference, the information-structure based triggering approach would need to assume that the CC of \emph{discover} is less likely to be at issue than that of \emph{confirm}.
 
-				\item pragmatics of full embedding context; Does Cole think that Julian dances salsa?
+		\ex. \a. Perhaps Cole discovered that Julian dances salsa.
+			\b. Perhaps Cole confirmed that Julian dances salsa.
+			\z.
+		\z.
 
-				\item why would that be the case?
+		While \citet{tonhauser_how_2018} experimental evidence that projection ratings negatively correlate with at-issueness ratings in out-of-the-blue polar questions.
+\citet{sieker_projective_2022} argue that their data also shows the same correlation for modal, conditional antecedent, and negation contexts for the German clause-embedding predicates they investigated.
+To address the gradient by-predicate variation theoretically, this approach would need to explicitly spell out how various predicates are associated with different likelihoods that their CCs is interpreted as at-issue.
 
-			\end{itemize}
+		\paragraph{Operator/predicate interaction.}
+			Since we found an operator/predicate interaction effect on projection, a further prediction of this type of approach would be that at-issueness ratings would show an operator-predicate interaction in this experimental setup as well.
+To address the interaction effect theoretically, an information structural triggering approach would need to make explicit predictions about how the meaning of entailment-cancelling operators affects at-issueness differentially for different predicates.
+This is certainly conceivable, when we take default assumptions about how likely embedded contents are interpreted at issue to be associated with the entire embedding context, including clause-embedding predicates, but also entailment-cancelling operators.
+However, this has not been spelled out explicitly and needs to be investigated in more detail.
 
-			\item contextual triggering
+	\subsection{Triggering based on world knowledge}
 
-			\begin{itemize}
+			The account in \citealt{schlenker_triggering_2021} assumes a triggering mechanism which operates on local contextual entailments of an expression, and determines which one of them projects based on probabilistic world knowledge about the truth of the invoked contents.
+For instance, in a negated utterance with \emph{discover} \Next[a], the CC expressed by \Next[c] has the potential to project if it is entailed by the negative prejacent \Next[b] along with the information that is available in its local context.
+Based on \citealt{heim_projection_1983} and \citealt{schlenker_local_2009}, negation is treated as a presuppositional hole, and the local context under negation is taken to be the global context.
 
-				\item what is the difference between know and prove (in terms of conceptual precondition, causal model?) there is no conceptually inherent relation between the two, the difference is lexical
+			\ex. \a. Cole didn't discover that Julian dances salsa.
+				\b. Cole discovered that Julian dances salsa.
+				\b. Julian dances salsa.
+			\z.
 
-				\item QUD influences
+			The triggering mechanism states that an inference $p$ of some propositional expression $E$ projects if $p$ is an epistemic precondition for the truth of $E$ in its local context $c^\prime$.
+			%
+			$p$ is an epistemic precondition of $E$ in $c^\prime$ if, usually, when one acquires the belief in $c^\prime$ that $E$, one already knows that $p$.
+			%
+			For \Last[a], the CC expressed by \Last[c] can be characterizes as an epistemic precondition for the negative prejacent \Last[b] based on the \lq subjective conditional probability\rq\ that a generic epistemic agent would already believe \Last[c] when learning that \Last[a] is true given the local context.
+Specifically, it is taken to be an epistemic precondition, if that probability reaches a contextually given threshold.
 
-				\item what is the relationship between lexical differences and QUD?
+			\paragraph{Gradient by-predicate variation.} This account predicts the potential of projection for CCs that are entailed by the prejacent in combination with contextual information.
+It can therefore incorporate effects of contextual information on projection, such as the reliability of the source of a speech report (\citealt{rieh_credibility_2010,de_marneffe_did_2012}).
+It also predicts the potential to of projection for CCs that are not entailed based on the literal content.
+In order to extend an account along these lines to address the out-of-the-blue utterances in our study, we might assume that these out-of-the-blue utterances are interpreted relative to an imagined plausible discourse context (c.f., discussion in \citealt{simons_best_2017}).
 
-			\end{itemize}
+			Due to the probabilistic nature and context-sensitivity of the triggering mechanism, the approach can make gradient predictions.
+However, the subjective conditional probabilities associated with expressions and their contextual inferences are taken as given in the account.
+In order to make concrete predictions about the gradient by-predicate variation, this account would need to make explicit how the relevant probabilities are derived based on different predicates.
+For instance, to explain the contrast that \emph{discover} receives higher projection ratings than \emph{confirm}, this account would assume that the situations in which one learns that a sentence like \emph{$x$ discovered that $p$} is true, are such that one already knew that $p$ is true more often than situations in which one learns the truth of a sentence like \emph{$x$ confirmed that $p$}.
 
-		\end{itemize}
+			\paragraph{By-operator variation.} The account does not incorporate differential effects of entailment-cancelling operators on projection.
+Although subjective conditional probabilities are relativized to a local context, the local context in negative contexts and conditional antecedents is assumed to the global context (based on \citealt{heim_projection_1983,schlenker_local_2009}).
+Therefore, no differential effect of operator is expected.
+For instance, the conditional probability for $p$ given $E$ vs that of $p$ given $(\emph{not } E)$ or that of $p$ given $(\emph{if } p, q)$ should not differ.
+
+			In order to address the operator-predicate interaction effect, an approach based on epistemic preconditions would need to make explicit how the meaning of entailment-cancelling operators affects the conditional probabilities differentially for different predicates.
+			%
+			The conditional probability that some embedded content is antecedently known would need to be computed relative to a global context, and take into account the entire embedding context, including the combination of entailment-cancelling operator and clause embedding predicate.
+
+	\subsection{Discussion: Contextual inferences and probabilistic triggering}
+
+		Our data showed gradient projection variation, based on a main effect of predicate, and an operator-predicate interaction.
+We found these effects for the CCs of clause-embedding predicates under four entailment-cancelling operators in out-of-the-blue contexts.
+The fact that we see some amount of projection for all predicates (even non-veridical ones), can only be addressed by an account that assumes the potential of projection for contextual inferences, rather than only literal ones.
+
+		The main effect of predicate and the operator-predicate interaction are both gradient effects, and cannot be explained by a categorical distinction between presuppositional and non-pre\-suppo\-si\-tion\-al predicates.
+The effects could rather be addressed in a probabilistic account, where various embedding contexts (i.e., combinations of entailment-cancelling operator and predicate) are associated with different probabilities that the embedded content projects.
+
+		Under the assumption of information-structural triggering, this is the probability that the embedded content is not at issue in the (possibly imagined) discourse context.
+Under a world-knowledge-based triggering account, the projection probability can be associated with the probability that the embedded content is already known in the discourse context.
+These two notions are, of course, related.
+If some proposition is already known, raising an issue about whether $p$ is the case will be uninformative.
+Therefore, it is unclear whether these two approaches would make different predictions for our data.
+Both types of accounts would need to be extended to make concrete predictions for the contribution of the combinations of entailment-cancelling operators and predicates.
 
 \section{Lexical differences}
+	Can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
+	This is a pressing question for future research, to which our data offer some tentative answers.
+We can find some initial generalizations over lexical properties, indicated in \textbf{Figure~\ref{fig:patterns}}, which gives the mean certainty ratings for the four operators by predicate, identifying four groups of predicates that show similar by-operator variation.
 
-	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting initial generalizations over lexical properties, indicated in \textbf{Figure\ 2}.
+	\begin{figure}[ht]
+		\centering
+		% \hspace{-.5em}\resizebox{\linewidth}{!}{
+		\begin{tabular}{p{.205\linewidth} p{.18\linewidth} p{.18\linewidth} p{.18\linewidth} p{.18\linewidth}}
+			\footnotesize \hspace{2.1em} (a) Negation high &
+			\footnotesize \ (b) Negation low &
+			\footnotesize (c) Conditional high &
+			\footnotesize \ (d) Modal low
+			% & \footnotesize \  (e) Discover
+			\\
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.225 \textwidth, valign=T]{negation-high-1.pdf}
+			&
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.2\textwidth, valign=T]{negation-low-1.pdf}
+			&
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.2\textwidth, valign=T]{conditional-high-1.pdf}
+			&
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.2\textwidth, valign=T]{modal-low-1.pdf}
+			% &
+			% \vspace{-.6\baselineskip}
+			% \includegraphics[width=.2\textwidth, valign=T]{discover-1.pdf}
+			\\
+		\end{tabular}
+		% }
 
-	So, can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
-	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
+		\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).}
+		\label{fig:patterns}
+	\end{figure}
+
 	%
-	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality associated with these predicates.
-	%
-	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf `Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
-	%
-	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This {\bf `Conditional high'} pattern (c) may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
-	%
-	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern (d).
+	The non-veridical predicates \emph{pretend} and \emph{think} exhibit the `Negation high' pattern, shown in panel (a) of Figure~\ref{fig:patterns}.
+These are the only predicates that are most projective under negation compared to all other operators.
+This could be related to
+	assumptions that non-veridical doxastics like \emph{think} can often be interpreted in comparison to a stronger veridical alternative (e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
+We tentatively hypothesize that this can lead to an inference that the CC is false in upward-monotone contexts, but not under negation.
+A similar alternative seems to be salient for \emph{pretend}, but not the non-veridical communicative predicates \emph{say} or \emph{suggest}.
+
+	The inferential predicates \emph{confirm, demonstrate}, and \emph{prove} exhibit a `Negation low' pattern, shown in panel (b).
+These are most projective under modals and conditionals and least projective under negation.
+Here, we may hypothesize that a negative statement involving these predicates is often only taken as relevant (in a Gricean sense) when it is antecedently known that there was an attempt to confirm/demonstrate/prove the truth of the CC.
+This may result in an inference that this attempt failed and result in lower projection ratings under negation.
+
+	For \emph{admit, announce, confess}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This `Conditional high' pattern (c) may suggest that the discourse effect of a conditional interacts with these change-of-state communication predicates.
+
+	Finally, the predicates \emph{hear, inform,} and \emph{know} exhibit a `Modal low' pattern (d).
 The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
-
-	%%%%
-	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
-\textbf{Fig.~\ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, identifying some groups of predicates that show similar by-operator variation.
-We highlight four \lq projection-profiles\rq here: \emph{pretend} and \emph{think} are the only predicates that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
-\emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}).
-\emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}).
-Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).
 
@@
 
-\newpage
+% \newpage
 \bibliographystyle{sub-chicago}

### 2024-02-05 results/main/meta-analyses/new/1_projectivity/projection-operators-summary.Rmd
added rmd overview file

---
title: "Projection operators data analysis summary"
output:
  pdf_document:
    toc: true
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage{euler}
   - \usepackage[T1]{fontenc}
   - \usepackage[margin = 1in]{geometry}
   # - \usepackage{fontspec}
   # - \defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
   # - \setmonofont[Scale=MatchLowercase]{Andale Mono}
mainfont: SourceSansPro
date: "2024-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/', cache = TRUE)
library(tidyverse)

# summaries
source('../../../helpers.R')

# graphing
# Color blind friendly palette (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)):
cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", "#CC79A7", "#000000")
cbbRotate <- c("#0072B2", "#CC79A7", "#000000", "#D55E00", "#F0E442", "#E69F00", "#56B4E9", "#009E73")
graphcolor <- "#0072B2"
theme_set(theme_bw())

# analysis
library(rstan)
options(mc.cores=parallel::detectCores())
rstan_options(auto_write=TRUE)
library(brms)
```

```{r load-data, include=FALSE}
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
data <- read.csv("../data_combined.csv", header = TRUE, sep = ",")
data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), operator = as.factor(op), participant = as.factor(workerid), item = as.factor(content), .keep = "none")
levels(data$operator) <- c("conditional", "modal", "negation", "question")
```

## The dataset

The dataset consists of $57160$ observations from $2682$ participants (recruited on Amazon MTurk and Prolific). We are interested in how projection ratings (`projection`) are affected by `predicate` and `operator` .

```{r dataset}
str(data)
```

The levels of the independent variables are:

```{r independent-variables}
levels(data$operator)
levels(data$predicate)
```

\pagebreak

## Graphs and summaries

### Summary for overall projection ratings

```{r projection-means, echo=FALSE}
data %>% summarize(Mean = mean(projection), CILow = ci.low(projection), 
              CIHigh = ci.high(projection)) %>%
    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh) %>% knitr::kable()
```

```{r projection-plot, echo=FALSE, fig.dim = c(4,2)}
data %>% ggplot(aes(x = projection)) +
  geom_area(stat = "bin", binwidth = .01, fill = graphcolor) +
  xlab("Certainty rating")
```

### Projection by operator

```{r operator-means, echo=FALSE}
operator_means <- data %>% group_by(operator) %>% 
    summarize(Mean = mean(projection), CILow = ci.low(projection), 
              CIHigh = ci.high(projection)) %>%
    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
           operator = fct_reorder(as.factor(operator), Mean)) %>% ungroup()
operator_means %>% knitr::kable()
operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
```

```{r operator-graph, echo=FALSE, fig.dim = c(3, 4.5), warning=FALSE, message=FALSE}
operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), 
                                y = projection, label = round(projection, digits = 2))) +
    geom_violin(data = data, scale="width", fill = graphcolor, 
                alpha = .4, color = graphcolor) +
    geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "black") +
    geom_point(size=0.5, color = "black") +
    geom_text(hjust = 0, nudge_x = -0.08, nudge_y = -0.06) + 
    scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Certainty rating") +
    theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
    xlab("Operator")
    # + labs(title = "Projection by operator")
```

### Projection by predicate

```{r predicate-means, echo=FALSE}
predicate_means <- data %>% group_by(predicate) %>% 
    summarize(Mean = mean(projection), CILow = ci.low(projection), 
              CIHigh = ci.high(projection)) %>%
    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
           predicate = fct_reorder(as.factor(predicate), Mean)) %>% ungroup()
predicate_means %>% knitr::kable()
predicate_means <- predicate_means %>% mutate(projection = Mean, .keep = "unused")

data$VeridicalityGroup = as.factor(
  ifelse(data$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))

predicate_means$VeridicalityGroup = as.factor(
  ifelse(predicate_means$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
```

```{r color-coding, echo=FALSE}
colors <- c("darkorchid", "dodgerblue")
pred_order <- levels(fct_reorder(predicate_means$predicate, predicate_means$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")
```

```{r predicate-graph, echo=FALSE, fig.dim = c(7, 4.5)}
predicate_means %>% ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = data, scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = "Predicate") +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="bottom", text = element_text(size=12)
            )
```

### Projection by predicate and operator

```{r predicate-operator-means, echo=FALSE, warning=FALSE, message = FALSE}
predicate_operator_means <- data %>% group_by(predicate, operator) %>%
      summarize(Mean = mean(projection), CILow = ci.low(projection), 
                CIHigh = ci.high(projection)) %>%
      mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
             predicate = fct_reorder(as.factor(predicate),Mean)) %>%
      mutate(predicate = fct_reorder(predicate, Mean, .fun = mean)) %>%
      mutate(operator = fct_reorder(operator, Mean, .fun = mean)) %>% ungroup()
predicate_operator_means %>% knitr::kable()
predicate_operator_means <- predicate_operator_means %>% mutate(projection = Mean, .keep = "unused")

predicate_operator_means$VeridicalityGroup = as.factor(
  ifelse(predicate_operator_means$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
```

```{r predicate-operator-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 4.5)}
pred_order <- levels(fct_reorder(
  predicate_operator_means$predicate, 
  predicate_operator_means$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

predicate_operator_means %>%
  ggplot(aes(x = fct_reorder(predicate, projection), y = projection, group = operator, color = operator)) +
      geom_point(aes(shape = operator), size = 4) + 
      geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
      geom_line() +
      xlab("Predicate") +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1), 
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Mean certainty rating") +
      scale_shape_manual(values = c("M", "N", "C", "Q")) +
      scale_colour_manual(values=cbbPalette) +
      theme_bw() +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            legend.position="none", text = element_text(size=12)
            )
```

```{r predicate-operator-graph-violin, echo=FALSE, fig.dim = c(10, 5)}
predicate_operator_means %>%
  ggplot(aes(x = fct_reorder(predicate, projection), y = projection)) +
    geom_violin(data = data, scale="width", fill = graphcolor, alpha = .4) +
    geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.5, color="black") +
    geom_point(size=0.5, color="black") +
    xlab("Predicate") +
    scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Mean certainty rating") +
    facet_wrap(~operator) +
    # theme(text = element_text(size=12), axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = -0.1, color=cols$Colors)) +
    theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))
```

### Projection by predicate, negation only
```{r negation-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.2)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "negation")$predicate, 
  subset(predicate_operator_means, operator == "negation")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "negation") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "negation"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="none", text = element_text(size=12)
            )

```

### Projection by predicate, question only
```{r question-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.2)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "question")$predicate, 
  subset(predicate_operator_means, operator == "question")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "question") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "question"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="none", text = element_text(size=12)
            )
```

### Projection by predicate, modal only
```{r modal-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.2)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "modal")$predicate, 
  subset(predicate_operator_means, operator == "modal")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "modal") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "modal"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="none", text = element_text(size=12)
            )
```

### Projection by predicate, conditional only
```{r conditional-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.75)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "conditional")$predicate, 
  subset(predicate_operator_means, operator == "conditional")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "conditional") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "conditional"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="bottom", text = element_text(size=12)
            )
```

## Analysis

**Response:** projection ratings (rescale for beta regression). Scaling to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$, excluding boundaries using method used in Degen & Tonhauser (2022), from Smithson & Verkuilen (2006), for proportional data. $$y' = \frac{y \cdot (n - 1) + 0.5}{n}$$

```{r dependent-variable}
data$betaresponse <- (data$projection*(nrow(data)-1) + .5)/nrow(data)
data$betaresponse <- as.numeric(data$betaresponse)
summary(data$betaresponse)
str(data)
```

Fixed effect: operator. Using deviation coding to compare each level of operator to the overall mean.

```{r fixed-effect}
data <- data %>% mutate(operator = fct_reorder(operator, projection, .fun = "mean"))
contrasts(data$operator) <- contr.sum(4)
colnames(contrasts(data$operator)) <- c("modal", "negation", "question")
contrasts(data$operator)
```

## Predicate profiles

```{r predicate-profiles, echo=FALSE, fig.dim = c(10, 10)}
predicate_operator_means %>%
  ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), y = projection, color = fct_reorder(operator, projection, .fun = "mean"), shape = fct_reorder(operator, projection, .fun = "mean"))) +
  geom_violin(data = data, scale="width", alpha = .4) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.5, color="black", alpha = .6) +
  geom_point(size = 4) +
  facet_wrap(vars(fct_reorder(predicate, projection, .fun = "mean"))) +
  xlab("Operator") +
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  scale_colour_manual(values=cbbPalette) +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Certainty rating") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

Clustering predicates based on the differences between means for the operators.

```{r profiles}
# cross-tabulate means for predicate / operator
pomeans <- data.frame(matrix(nrow = length(levels(predicate_operator_means$predicate)),
                             ncol = length(levels(predicate_operator_means$operator))),
                      row.names = levels(predicate_operator_means$predicate))

colnames(pomeans) <- levels(predicate_operator_means$operator)

for (i in 1:length(levels(predicate_operator_means$predicate))) {
  current_pred = levels(predicate_operator_means$predicate)[i]
  for(j in 1:length(levels(predicate_operator_means$operator))) {
    current_op = levels(predicate_operator_means$operator)[j]
    pomeans[current_pred, current_op] <- subset(predicate_operator_means,
                                                predicate == current_pred &
                                                  operator == current_op)$projection[1]
  }
}

# get differences between operator means for each predicate
pomeansdiffs <- pomeans %>% mutate(mn = modal - negation, nc = negation - conditional, 
                                   cq = conditional - question, .keep = "none")

# plot hierarchical clusters based on differences
plot(hclust(dist(pomeansdiffs), method = "complete"))
```

### Negation high

```{r negation-high, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>%  mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>%
  subset(predicate == "think" | predicate == "pretend") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 4, y = 0.3, label = "think", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 3.85, y = 0.1, label = "pretend", size = 2.8, color = graphcolor) +
  scale_colour_manual(values=cbbRotate) +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Mean certainty rating") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Negation low

```{r negation-low, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "confirm" | predicate == "prove" | predicate == "demonstrate"| predicate == "establish") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 3.4, y = 0.57, label = "demonstrate", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 3.9, y = 0.28, label = "confirm", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 2, y = 0.17, label = "prove", size = 2.8, color = "black") +
  scale_colour_manual(values=cbbRotate) +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Conditional high

```{r conditional-high, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "announce" | predicate == "confess" | predicate == "admit" | predicate == "reveal") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 4, y = 0.79, label = "reveal", size = 2.8, color = "#D55E00") +
  annotate(geom = "text", x = 3.7, y = 0.48, label = "announce", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 1, y = 0.64, label = "admit", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 2, y = 0.44, label = "confess", size = 2.8, color = "black") +
  scale_colour_manual(values=cbbRotate) +
  xlab("Operator") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Modal low

```{r modal-low, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "know" | predicate == "inform" | predicate == "hear" | predicate == "be_annoyed") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 1, y = 0.45, label = "hear", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 3, y = 0.92, label = "know", size = 2.8, color = "black") +
  annotate(geom = "text", x = 4, y = 0.68, label = "inform", size = 2.8, color = "#CC79A7") +
  scale_colour_manual(values=cbbRotate) +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Question high

```{r discover, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "see" | predicate == "acknowledge" | predicate == "discover") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  scale_colour_manual(values=cbbRotate) +
  annotate(geom = "text", x = 3.5, y = 0.63, label = "acknowledge", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 1.2, y = 0.73, label = "discover", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 3, y = 0.87, label = "see", size = 2.8, color = "black") +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Suggest

```{r suggest, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "be_right" | predicate == "suggest") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  scale_colour_manual(values=cbbRotate) +
  annotate(geom = "text", x = 3.5, y = 0.1, label = "be right", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 3.9, y = 0.35, label = "suggest", size = 2.8, color = "#CC79A7") +
  xlab("Operator") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

## R session info

```{r session-info, echo=FALSE, message=FALSE}
sessionInfo()
```

### 2024-02-05 writing/SuB-paper/projection-operators-SuB.tex
WiP SuB paper, sec 2 graphs

--- 2023-12-13 writing/SuB-paper/projection-operators-SuB.tex
@@
 %\setlength{\parskip}{\the\baselineskip}
-\pagestyle{empty}
+
+% JT: I've commented this out so that we can see page numbers while writin
+%\pagestyle{empty}
 
@@
 \usepackage{natbib} 
-\bibpunct{(}{)}{;}{a}{,}{,}
+%\bibpunct{(}{)}{;}{a}{,}{,}
+\bibpunct{(}{)}{,}{a}{,}{,}
 
@@
 \author{Hofmann \and de Marneffe  \and Tonhauser}
-\title{Projection variability of clausal complements across different operators}
+%\title{Projection variability of clausal complements across different operators}
+\title{Projection variability: Is the family of sentences really a family?}
 \makeatletter
@@
 
-\pagenumbering{gobble} 
+% JT commented out whle editing
+%\pagenumbering{gobble} 
 %suppresses page numbering in header
 
-\input{lisa-extra}
+%% other packages I added (LH)
+% misc formatting
+	\usepackage{booktabs}
+    \usepackage[export]{adjustbox}
+    \usepackage{rotating}
+
+%% bibliography
+	\newcommand{\posscite}[1]{\citeauthor{#1}'s (\citeyear{#1})}
+
+ \newcommand{\poscite}[1]{\citeauthor{#1}'s \citeyear{#1}}
+
+% drawing
+	\usepackage{graphicx}
+	\usepackage{xcolor}
 
@@
 
-{\large \textbf{Projection variability of clausal complements across different operators}}\footnote{We would like to thank\ldots Taylor Mahler; Reviewers at SALT, SuB, XPrag; Audience at Syntax \& Semantics colloquium at University of Stuttgart}\\
+%{\large \textbf{Projection variability of clausal complements across different operators}}
+{\large \textbf{Projection variability: Is the family of sentences really a family?}}\footnote{We thank Taylor Mahler for assistance in collecting the data presented here as well as valuable comments.
+We gratefully acknowledge the National Science foundation grant \#1452674 (to Marie-Catherine de Marneffe, Craige Roberts, and Judith Tonhauser), which provided financial support for this project.}\\
 Lisa HOFMANN --- \textit{University of Stuttgart}\\
-Marie-Catherine DE MARNEFFE --- \textit{UC Louvain}\\
+Marie-Catherine DE MARNEFFE --- \textit{UCLouvain}\\
 Judith TONHAUSER --- \textit{University of Stuttgart}\\
 
-\textbf{Abstract.}
-	We present experimental evidence that the projection of clausal complements \textbf{(i)} varies between entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend the result from \citet{degen_are_2022}, that projection ratings in polar questions are gradient in ways that cannot be predicted by categorical lexical classes, to contexts with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
-	%
-	The observed variability is not captured by existing theoretical accounts of projection
-	(e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
-	%
+\textbf{Abstract.} We present experimental evidence that the projection of the content of clausal complements \textbf{(i)} varies across entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from nonfactive predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
+The observed variability is not captured by existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
 Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.
 
-\textbf{Keywords:} projection, commitment, discourse interpretation.
+\textbf{Keywords:} Projection variability, entailment-canceling operators, (non)factive predicates.
 
-\section{Introduction}
-	Language users may infer that a speaker who utters an attitude ascription, as in \ref{ex:family}, is committed to the content of the complement (CC, here: \emph{Julian dances salsa,})
-	even when it occurs under an entailment-canceling operator, like negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditionals \ref{ex:cond}, in which case we say that it \textit{projects} (\citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}, and many more).
+\section{Introduction} \label{sec:intro}
+
+    Textbooks typically present the family-of-sentences diagnostic for projection without addressing by-operator variation (e.g., \citealt{chierchia_meaning_1990}).
+For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from its variants in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
+
+    % In formal semantics/pragmatics, the so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}).
+For instance, to diagnose the content of the clausal complement of \emph{discover} as projective content, a researcher would typically observe that the content that Julian dances salsa follows not just from \ref{ex:simple}, but also from the family-of-sentences variants of \ref{ex:simple} in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
+
+    % (e.g., \citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}).
+
+    \ex.\label{ex:simple} Cole discovered that Julian dances salsa.
 
@@
 			{\bf Negation:} \hfill
-			\emph{\lq Cole didn't discover that Julian dances salsa.\rq}
+			\emph{Cole didn't discover that Julian dances salsa.}
 		\b. \label{ex:q}
 			{\bf Polar Question:} \hfill
-			\emph{\lq Did Cole discover that Julian dances salsa?\rq}
+			\emph{Did Cole discover that Julian dances salsa?}
 		\c. \label{ex:mod}
 			{\bf Modal:} \hfill
-			\emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
+			\emph{Perhaps Cole discovered that Julian dances salsa.}
 		\d. \label{ex:cond}
 			{\bf Conditional:} \hfill
-			\emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
+			\emph{If Cole discovered that Julian dances salsa, Logan will be joyful.}
 		\z.
@@
 
-	We investigate the question whether the projection of embedded propositional content across entailment-cancelling operators differs by operator, and whether by-operator differences vary between embedding predicates.
+    However, prior research is suggestive of variation in the effect of different entailment-cancelling operators on projection.
+For instance, \citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \textit{regret}) and semi-factives (e.g., \textit{discover}).
+    Based on the examples in \Next, he suggested that the CC of true factives projects across all four operators, but that of semi-factives does not always project from under polar questions, modals, or conditionals.\footnote{See \citealt{beaver_have_2010} for naturally occurring examples in which the CCs of factive predicates do not project from under the four operators, thereby calling into question the distinction between factive and semi-factive predicates.
+For experimental research on the distinction see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
+We return to this literature in section \ref{s:analysis}.}
+
+    \ex. \citealt{karttunen_observations_1971}: (22, 24--26)
+        \a. John didn't \{regret/discover\} that he had not told the truth.
+        \b. Did you \{regret/discover\} that you had not told the truth?
+        \b. If I \{regret/discover\} later that I have not told the truth, I will confess it to everyone.
+        \b. It is possible that I \{regret/discover\} later that I have not told the truth.
+        \z.
+    \z.
+
+    There have been two experimental investigations of by-operator projection variation.
+First, the experiment reported in \citet{smith_relationship_2014} investigated projection from under negation and antecents of conditionals for various types of English projective contents.
+They found that the expressive content of epithets (e.g., \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals.
+In contrast, the content of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference.
+%However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance.
+Since other factors besides speaker commitment modulate discourse (un)expectedness (see e.g., \citealt{zimmermann_grammatical_2011,tonnis_german_2021}), it is unclear that they (only) measured projection.
+
+    Second, \citet{sieker_projective_2022}  compared the projection of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') between the four operators in \ref{ex:family}.
+The results showed that the CCs of the predicates investigated projected more from under negation than the other three operators.
+contrary to what \citealt{karttunen_observations_1971} suggested, a comparison of factive and semifactive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of nonfactive predicates.
+
+    %They replicated \poscite{smith_relationship_2014} result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know').
+However, this finding was part of an overall pattern of higher projection ratings with negation than other operators.
     %
-	Textbook diagnostics for projection often assume that, if a content projects across one of the operators in \Last, it uniformly projects across the others (\citealt{chierchia_meaning_1990}).
-Accordingly, current research examines projection variability between these operators only rarely, and with conflicting findings.
+    %Grouping ratings for the four factive and semi-factive predicates, respectively, they found no interaction of operator with predicate type, and thus no evidence for the factive/semi-factive distinction (iii), or an operator/predicate interaction (ii).
+However, their analysis did not explore the impact of operators on each predicate individually.
 
-	\citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \textit{be annoyed, regret, reveal, know}), where the CC projects across all four operators, and semi-factives (e.g., \textit{discover, realize, see, notice, find out}) where it always projects from under negation, but not always from under polar questions, modals, or conditionals.
+    This paper reports on the results of a set of experiments that were designed to compare projection from under the four entailment-canceling operators in \ref{ex:family}.
+Our experiments extend the empirical scope of prior research by investigating projection for a broader set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}:
 
-	\citet{smith_relationship_2014} investigated by-operator variation, comparing negation and conditionals, for various types of projective content.
-They found that the projective content of epithets (e.g.
-\textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals, whereas that of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and that of clefts showed no difference.
-Their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance.
-Since other factors besides speaker commitment modulate discourse (un)expectedness \citep[see e.g.][]{zimmermann_grammatical_2011,tonnis_german_2021}, it is unclear that they (only) measured projection.
+    \ex. \label{ex:predicates}
+    		\a. (Semi)factive predicates: {\em be annoyed, know, reveal, discover, see}
+      \b. Nonfactive predicates: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, be right, demonstrate, pretend, say, suggest, think}
 
-	\posscite{sieker_projective_2022} study on German clause-embedding predicates compared the projection of purported factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{enthüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') across the four contexts in \Last.
-They replicated Smith and Hall's result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know').
-However, this finding was part of an overall pattern of higher projection ratings with negation than other operators.
-They found no interaction with predicate type, and therefore no evidence for the distinction of factives vs.
-semi-factives.
-This is based on an analysis which aggregated the ratings for the four purported factive and semi-factive predicates, under each of the entailment-cancelling operators.
+    As shown, the  20 predicates include both (semi)factive and nonfactive predicates.
+The inclusion of the latter set of predicates is motivated by the results of the empirical investigations in \citealt{degen_are_2022}, which suggested that the contents of complements of nonfactive predicates is also projective and that projection does not categorically distinguish factive and nonfactive predicates.
+As \poscite{degen_are_2022} result was established based on sentences where the predicates were embedded under polar questions, our experiments, which feature all four entailment-canceling operators, serve to replicate and extend this result of \citealt{degen_are_2022}.
 
-	Studies in \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} observed by-predicate variation in polar questions, where Karttunen's classification predicts more projection for factives than semi-factives.
+    %    The factive/semi-factive distinction (iii) is further called into question by \citealt{beaver_have_2010}, who reports many naturally occurring examples of cognitive factive predicates, showing that their CC does not necessarily project (e.g., English \emph{be/become aware, know, discover, notice, find out, realize, accept the fact}; French \emph{trouver} \lq discover\rq).
+
+    %Experimental findings from \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} can be evaluated with respect to whether they support Karttunen's distinction (iii), with mixed results.
+These studies observed by-predicate variation in polar questions, where more projection is expected for factives than semi-factives.
 	%
-	\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-But here, it is also not obvious how exactly this task relates to projection.
+	%\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for English \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
+This could be interpreted as limited support for (iii).
+But here, it is also not obvious how exactly this task relates to projection.\footnote{Say why}
+
+	%\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured speaker commitments toward the CC of a broad range of English clause-embedding predicates more directly, by collecting judgments of how certain participants take a speaker to be of the contents under investigation.
 	%
-	\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured projection of the CC of a broad range of attitude predicates more directly, collecting ratings about speaker certainty about the CC.
-The observed differences between individual clause-embedding predicates do not align with theoretically assumed lexical classes.
-While Degen \& Tonhauser demonstrate this for the distinction between purported factive and non-factive predicates, the results also did not match the expectations from Karttunen's factive vs.
-semi-factive classification.
+	%The observed differences between the various predicates do not align with theoretically assumed lexical classes.
+While \citet{degen_are_2022} demonstrate this for the distinction between purported factive and non-factive predicates, the results did not match the expectations from Karttunen's factive vs.
+semi-factive classification either.
 	%
-	Using speaker-certainty ratings, \citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}.
+	%\citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}.
 \citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}.
-These findings suggest that individually examining projection ratings for various predicates, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate variation in operator effects on projection.
+These findings also suggest that examining projection ratings for various predicates individually, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate projection variation.
 
-	While \citet{smith_relationship_2014} found by-predicate variation in the effect of the operator on projection, \citet{sieker_projective_2022} did not.
+	%Both \citet{smith_relationship_2014} \citet{sieker_projective_2022} found that projection is affected differently by entailment-cancelling operators.
+However, while Smith \& Hall found variation in the effect of the operator on projection between various projective contents, Sieker \& Solstad did not.
 	%
-	These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis.
-To address this, we conducted a series of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family}.
-We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' diagnostic; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English clause-embedding predicates, including factive (\emph{be annoyed, know, reveal}) and semi-factive predicates (\emph{discover, see}), and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
+	%These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis.
+Finally, much previous research calls into question the hypothesis that by-predicate differences in the effect of operator on projection pattern according to Kattunen's factive/semi-factive distinction (\citealt{beaver_have_2010,tonhauser_how_2018,sieker_projective_2022,degen_are_2022}).
 
-\section{Assessing projection across entailment-canceling operators}
+  %Karttunen's distinction gives rise to three hypotheses about the projection associated with clause-embedding predicates, which we explore in this paper: \textbf{(i) by-operator variation:} entailment-cancelling operators differentially affect projection; \textbf{(ii) operator/predicate interaction:} the effect of operator differs between predicates; \textbf{(iii) factive/semi-factive distinction:} this difference can be addressed by characterizing lexical classes of (semi-)factive predicates.
 
-	To assess the effect of entailment-cancelling operators and clause-embedding predicates on the projection of the embedded content, we collected projectivity judgments for contents embedded under 20 clause-embedding predicates in four sets of experiments.
-The predicates were embedded under negation in Exps.~1, under polar questions (Exps.~2), under the epistemic possibility modal {\em perhaps} (Exps.~3), and in conditional antecedents (Exps.~4).
-(Each set of experiments contained three experiments differing in an at-issueness measure used in a separate block.
-We focus on the projection ratings here.)
+    %To address the three hypotheses, we conducted a set of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family} for various clause-embedding predicates.
+We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' task; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English predicates, including factives and semi-factives, and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
+Our findings support (i) by-operator variation, and an operator/predicate interaction (ii), but not Karttunen's factive/semi-factive distinction (iii).
 
-		In each experiment, participants read utterances like those in \ref{ex:family} and judged whether the speaker (who was named) was certain of the CC (e.g.: Is [the speaker] certain that Julian dances salsa?).
+\section{Experiments}
 
-		\begin{itemize}
-			\item say more about task
-		\end{itemize}
+	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.\footnote{The data and analysis scripts can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.
+{\bf LETS BE MORE SPECIFIC HERE ABOUT WHERE THINGS RELATED TO THIS PAPER ARE}} The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4.
+Each set of experiments consisted of three experiments that differed in the at-issueness measure used in a separate block.
+In this paper, we limit our attention to the projection ratings collected in the 12 experiments.
 
-		Based on Karttunen's generalization, we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.
+	Projection was measured in each experiment with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} This diagnostic was also used in \citealt{sieker_projective_2022}.
+Under this diagnostic, participants are presented with utterances like those in \Next.
+They are asked to rate whether the (named) speaker is certain of the CC.
 
-	\subsection{Experimental method}
+	\ex. \label{ex:certain-that}
+		\a. Christopher: \emph{\lq Cole didn't discover that Julian dances salsa.\rq}
+		\b. Christopher: \emph{\lq Did Cole discover that Julian dances salsa?\rq}
+		\c. Christopher: \emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
+		\d. Christopher: \emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
+		\z.
+		Projection question: Is Christopher certain that Julian dances salsa?
+	\z.
+
+    We assume, following \citealt{tonhauser_how_2018} and \citealt{degen_are_2022}, that judgments of speaker certainty about some propositional content reflect speaker commitment to that content, that is, projection.
+If a participant interprets an utterance like \Last[a--d] in a way that the speaker (e.g., Christopher) is certain of the CC, it is assumed to project out of the entailment-canceling environment.
+If a participant does not take the speaker to be certain of the CC, this is taken to indicate that it does not project.
+
+    %This setup allows us to test the three hypotheses outlined in Section \ref{sec:intro}.
+First, a main effect of operator on speaker certainty judgments for embedded contents would provide evidence for by-operator variation (i).
+Second, if there is a operator/predicate interaction (ii), we would expect the effect of operator to be different (in magnitude or direction) for different predicates.
+Finally, based on the factive/semi-factive distinction (iii), we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.
+
+	\subsection{Methods}
 
 		\subsubsection{Participants}
-			We analyze the data from 2,682 self-reported native speakers of American English recruited on Prolific or Amazon's MT platform. (roughly 750 participants in each experiment)
+			We recruited 250-300 participants for each of the 12 experiments.
+Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs.
+Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
+See Supplement \ref{app:d-participants} for further information about the participants.
 
 		\subsubsection{Materials}
-			Each participant saw all 20 attitude predicates (each paired with a unique content from a set of 20 contents) under one operator.
 
-			\begin{itemize}
-				\item 20 clause-embedding predicates that have shown projection variability in question contexts (Degen \& Tonhauser, 2022)
-				\item Crossed w/ 20 CCs: 20 x 20 = 400 combinations
-				\item One experiment per operator
-				\item six main-clause controls for data exclusion
-			\end{itemize}
+            The target sentences consisted of 400 combinations of the 20 clause-embedding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}).
+As mentioned above, the predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal \emph{perhaps} in Exps.~3, and in conditional antecedents in Exps.~4, for a total of 400 target stimuli in each of the four sets of experiments.
+
+			%We tested the 20 clause-embedding predicates in \Next, which have shown projection variability in polar questions (\citealt{degen_are_2022}).
+They comprise three factive predicates \Next[a], two semi-factives \Next[b], nine that have been characterized as optionally factive in \citealt{kiparsky_fact_1970} \Next[c], and six non-factives which include two veridical predicates \Next[d], and four non-veridical ones \Next[e].
+
+	%		\ex. 20 clause-embedding predicates \hfill \citealt{degen_are_2022}
+			%	\a. canonically factive:					\emph{be annoyed, know, reveal}
+				%\b. semi-factive: 		\emph{discover, see}
+%				\b. optionally factive: \newline \emph{acknowledge admit, announce, confess, confirm, establish, hear, inform, prove}
+%				\b. Non-factive veridical:
+%					\emph{be right, demonstrate}
+%				\b. Non-factive non-veridical:
+%				\emph{pretend, say, suggest, think}
+%				\z.
+%			\z.
+
+To assess whether participants were attending to the task, each experiment included six control stimuli. For details on the six control stimuli see Supplement \ref{app:c-control}.
+
+Each participant saw a random set of twenty-six stimuli: Each set contained one target stimulus for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same six control stimuli.\footnote{Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
+Block order was randomized.
+As mentioned above, we focus on the projection ratings here.} Trial order was randomized.
+
+  %which were main clauses with appositive relative clauses, which were also presented as utterances made by a named speaker.
+Here, the main clause content was targeted by the projection question.
+In the experiments with polar questions, the controls were expected to receive low certainty ratings, whereas in the experiments with declaratives, they were expected to receive high ratings.
+
+  %Each participant saw all 20 predicates, randomly paired with a unique embedded clause from a set of 20 contents, and a subject from a set of 20 proper names.
+%The predicates associated with projective content were embedded under (i) negation, (ii) polar questions, (iii) the epistemic possibility modal \textit{perhaps}, and (iv) in conditional antecedents in the four different sets of experiments.
 
@@
 
-			On each trial, participants read an utterance, which was presented as an utterance of a speaker with a randomly chosen name, as well as the corresponding projection question, and then gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
-This is illustrated in Fig.~\ref{fig:trial}.
+			Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else.
+            %
+            On each trial, participants read an utterance and responded to the `certain that'   question.
+They gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
+A sample trial from Exps.~1 is illustrated in Figure~\ref{fig:trial}.
+Following, e.g., \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things:
+            %
+            On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection.
+This assumes that projection (and the relevant notion of speaker commitment) can itself be gradient.
+            %
+            Conversely, higher ratings could also reflect the probability that a participant associates with a projective interpretation for the utterance.
+Under this interpretation, commitment may be interpreted categorically, and the gradience is a result of the uncertainty about possible interpretations (see also discussion in \citealt{grove_factivity_2023}, which suggests the latter kind of interpretation).
 
-			Following \citet{tonhauser_how_2018}, we took a `yes' response to indicate that participants interpret the utterance in a way that the presented speaker (e.g., Christopher in Fig.~\ref{fig:trial}) is committed to the embedded content, which reflects projection.
-A `no' response was taken to indicate that the content does not project.
+            %  we take high ratings of speaker certainty to reflect speaker greater speaker commitment towards the CC, that is, greater projection.
 
@@
 				\centering
-				\includegraphics[width = .7\linewidth]{task-1n-proj.eps}
-				\caption{A sample trial in Exp.~1. In the corresponding trials in Exps.~2-4, participants were presented with an utterance involving a different entailment-cancelling operator.}
+				\fbox{\includegraphics[width = .7\linewidth]{task-1n-proj.eps}}
+				\caption{A sample trial in Exps.~1. In the corresponding trials in Exps.~2-4, participants were presented with an utterance with a different entailment-cancelling operator.}
 				\label{fig:trial}
@@
 
+             After completing the experiment, participants filled out a short optional demographic survey.
+To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.
+
+        \subsubsection{Data exclusion}
+
+        Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}.
+The data from 2,682 participants entered into the analysis.
+
+            %We analyze the data from $2,682$ self-reported native speakers of American English, after excluding participants who failed one of the following two attention-checks: their mean rating on six control items is more than 2 sd different from the group mean, or they always roughly selected the same point on the scale for target stimuli (identified by low variance, and then manually inspecting their response patterns).
+
     \subsection{Results and discussion}
-		Our analysis reveals two key results: \textbf{(i)} There is projection variability by operator and \textbf{(ii)} there is by-predicate variation in the effect of operator on projection.
 
-		\subsubsection{By-operator variation}
+    We first address the question of whether there is by-operator variation (section \ref{s:by-op}) and then the question of whether there is by-predicate variation in the observed by-operator variation (section \ref{s:by-pred}).
 
-			We find an effect of operator when aggregating across predicates.
-Mean projection ratings were higher under question-embeddings than under negation and modals, but lower than in conditional antecedents.
-This is shown in Fig.~\ref{fig:op-ratings}, and the generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.
+		%Our data reveals three key results: \textbf{(i)} There is projection variability by operator; \textbf{(ii)} There is by-predicate variation in the effect of operator on projection; and \textbf{(iii)} We find further support (from entailment-cancelling operators beyond polar questions) for \posscite{degen_are_2022} result that by-predicate projection variation is gradient in a way that cannot be predicted by categorical lexical classes.
+
+	\subsubsection{By-operator variation}\label{s:by-op}
+
+        Fig.~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates.
+As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relative more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.
+
+			%When aggregating across predicates, we found an effect of \texttt{operator}.
+The distribution of projection ratings, as well as means and $95\%$ bootstrapped confidence intervals by operator are shown in Figure~\ref{fig:op-ratings}.
 
@@
 				\centering
-				\includegraphics[width = .35\linewidth]{projective-op}
-				\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals.}
+				\includegraphics[scale = .8]{operator-graph-1}
+				\caption{Mean certainty ratings by operator.
+Error bars indicate $95\%$ bootstrapped confidence intervals.
+Violin plots indicate the kernel probability density of the individual participants’ ratings.}
 				\label{fig:op-ratings}
 			\end{figure}
+
+			%Mean projection ratings were higher for CCs in question-embeddings than under negation and modals, but lower than in conditional antecedents.
+
+   These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.
 
@@
 
-					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fit with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
+					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fitted with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
 				\end{table}
 
-			This result differs from that of \cite{sieker_projective_2022}, who observed that the CC of German attitude predicates projects more from under negation than the other operators.
+			Like \citet{sieker_projective_2022}, we observe a small main effect of entailment-cancelling operator on projection.
+However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.
+            We note, however, that the differences in mean certainty ratings between the four entailment-canceling operators are very small.
+This suggests that, when abstracting away from individual projective contents, projection from under the four entailment-canceling operators is very similar.
+In other words, when abstracting away from individual projective contents, the family of sentences really are a family.
 
-			\begin{itemize}
-				\item but like them, we observe small differences 
-				\item overall, they have much higher projection ratings
-				\item but they chose a different set of predicates, we can understand these different results when taking into account by-predicate variation in the effect of operator
-			\end{itemize}
+		\subsubsection{By-predicate variation in the effect of entailment-canceling operator}\label{s:by-pred}
 
-		\subsubsection{By-predicate variation in the effect of operator}
+  Fig.~\ref{fig:op-pred-ratings} shows mean certainty ratings by predicate and entailment-canceling operator (aggregating over the 20 CCs that each predicate was paired with).
+As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
+The observed variation does not align with widely assumed predicate classes.
+For instance, the five (semi)factive predicates exhibit five different patterns:
+i) the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modal {\em perhaps}; ii) the CC of \emph{know} projects most from under questions, less from questions and negation, and least from under the modal {\em perhaps}; iii) the CC of \emph{see} projects most from under questions and conditionals, less from under negation, and least from under the modal \emph{perhaps}; iv) the CC of  \emph{discover} projects more from under conditionals and questions than negation and modal {\em perhaps}; and v) the CC of \emph{reveal} projects most from conditionals, less from questions, and least from negation and modal \emph{perhaps}.
+Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
+We also observe by-operator projection variation for nonfactive predicates.
+Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}.
+Other nonfactive predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.
 
-			We observe differences across the predicates in by-operator variation.
-This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 attitude predicates by embedding operator; predicates are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
-			%
-			For instance, whereas the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modals, and the CC of \emph{discover} projects less from negation than conditionals and questions, and more from under negation than modals.
+\begin{figure}[ht]
+				\centering
+				\includegraphics[width = \linewidth]{predicate-operator-graph-1}
+				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Predicates are ordered by mean certainty rating pooled over the four operators.
+Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
+				\label{fig:op-pred-ratings}
+			\end{figure}
+
+The observation that there is by-operator variation for the CC of each of the 20 clause-embedding predicates is confirmed by the statistical analysis. We fitted...
+
   The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
@@
 
-			This result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals.
-This differs from the result of \citet{sieker_projective_2022}, who found no interaction between predicates and operators.
+Our results align with those of \citealt{smith_relationship_2014}, who also observed by-expression by-operator variation.
+However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find an effect here.
+We hypothesize that this difference is due to the difference in projection diagnostic used.
+
+			%Our result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals (we found no difference here).
+   Our results differ from those of \citealt{sieker_projective_2022}: While that work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
+As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic.
+Other factors that varied between our experiments are the language under investigated (German vs.\ English), the clause-embedding predicates investigated, and the CCs that the predicates were paired with.
+Future research will need to establish which of these factors are implicated in the observed differences.
+
+%   , who found no difference between the two groups of predicates they assumed in the effect of operator.
 			%
-			Given that the methods and the set of projective contents of our experiment were very similar to theirs, this difference may suggest cross-linguistic variation in projection variability (cf.
-\citealt{tonhauser_projection_2020}).
-
-		\subsubsection{Gradience and lexical classes}
-
-			Finally, our results provide further support (from negation, modals, and conditionals) for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions.
-			E.g., the CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
-
-			Similarly, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates, in line with \citet{sieker_projective_2022} (see also \citealt{beaver_have_2010}): Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals.
+			%However in line with Sieker \& Solstad, our results question \posscite{karttunen_observations_1971} proposed difference between factive and semi-factive predicates (see also \citealt{beaver_have_2010}).
+Although our data replicate the result from \citet{tonhauser_how_2018} that, in polar questions, the CC of semi-factive \emph{discover} is less projective than that of \emph{know}, the same does not hold in conditionals, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates.
 			%
-			Further, contrary to what would be expected based on \posscite{karttunen_observations_1971} distinction between factive and semi-factive predicates, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
+			%Further, the CC of (factive) \emph{be annoyed} does not project invariably from all four operators, and the CC of \emph{discover,}  which is considered semi-factive, does not project more from under negation than the other three operators.
 The pattern observed for {\em know} does not fit into either category.
 
-		\subsubsection{Converging evidence for operator / predicate interactions}
-			MegaVeridicality dataset (\citealt{white_role_2018}): 517 predicates in three sentence types
+	\subsubsection{Factive vs.\ nonfactive predicates}
 
-			\ex. \a. Somebody didn’t know that a particular thing happened. (Did that thing happen?)
+Factive predicates are defined as predicates that presuppose the CC, while the CC of nonfactive predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entailment-canceling operators, this definition leads one to expect that factive predicates are distinguished from nonfactive ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of nonfactive ones.
+This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates}, which were embedded in polar questions.
+Contrary to expectation,  \citealt{degen_are_2022} found variation among the CCs of the five factive predicates investigated and that the CCs of some nonfactive predicates were as projective, or even more projective, than the CCs of some factive predicates.
+In short, projection of the CC from under polar questions did not categorically distinguish factive from nonfactive predicates.
+
+The results of the set of experiments reported on in this paper replicate this result.
+As shown in Fig.~\ref{fig:op-pred-ratings}, there is variation between the CCs of the five factive predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from nonfactive predicates.
+Furthermore, the results of  this set of experiments suggest that projection from under the three other entailment-canceling operators also does not categorically distinguish factive and nonfactive predicates.
+This result strengthens the conclusion of \citealt{degen_are_2022} that there is, to date no empirical evidence for a coherent class of factive predicates.
+
+IF THERE IS ROOM I WOULD LIKE TO INCLUDE 4 VERSIONS OF FIG 3 WHERE THE 20 PREDICATES ARE ORDERED BY OPERATOR, TO MAKE THE POINT OF THIS SECTION CLEARER.
+THAT WOULD ALSO BE GOOD FOR THE METHODOLOGICAL IMPLICATION SECTION BELOW.
+
+		Modal:\newline
+		\includegraphics[width = \linewidth]{modal-predicate-graph-1.pdf}
+        Negation:\newline
+        \includegraphics[width = \linewidth]{negation-predicate-graph-1.pdf}
+        Polar questions:\newline
+        \includegraphics[width = \linewidth]{question-predicate-graph-1.pdf}
+        Conditionals:\newline
+        \includegraphics[width = \linewidth]{conditional-predicate-graph-1.pdf}
+		%\caption{Mean certainty ratings by predicate, for each operator.
+Error bars indicate $95\%$ bootstrapped confidence intervals.
+Violin plots indicate kernel probability densities of individual observations.}
+
+%			Taking into account the effect of \texttt{predicate} on projection, we find gradient projection variation, like \citet{tonhauser_how_2018} and \citet{degen_are_2022}.
+			%
+			%This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 predicates grouped by embedding operator.
+Here, predicates on the x-axis are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
+
+			%Our results therefore provide further support from negation, modals, and conditionals for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions.
+The CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
+
+		\subsubsection{Summary}
+
+  The results of our experiments suggest that there is little by-operator variation when aggregating over clause-embedding predicates, but that the CCs of different clause embedding predicates exhibit by-operator projection variation.
+			%Based on our data, we conclude that embedded propositional content projects differently from under different entailment-cancelling operators. 
+    Crucially, the observed by-operator variation does not align with prior claims about differences between factive and semi-factive predicates in by-operator variation (e.g., \citealt{karttunen_observations_1971}).
+Finally, the results of our experiments provide further support for the results of \citealt{degen_are_2022}, who did not find empirical support for a class of factive predicates based on the projection of the CC from under polar questions.
+Our experiments suggest that projection of the CC from under negation, the antecedent of conditionals and epistemic modals also do not provide empirical support for a natural class of factive predicates.
+
+    Before discussing the methodological and theoretical implications of these results in section \ref{s:general}, we provide converging evidence from a different dataset in section \ref{s:converging}.
+
+%   does not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.
+			%
+			%Finally, the effect of operator on projection differs by predicate.
+The theoretical implications of our findings are discussed in the following section.
+Importantly, accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability.
+
+		\section{Converging evidence for the by-operator variation}\label{s:converging}
+
+This section provides converging evidence for the by-operator variation based on the MegaVeridicality dataset (\citealt{white_role_2018}).
+This dataset contains projection ratings for the CCs of 517 English clause-embedding predicates.
+The stimuli that participants rated consisted of combinations of these predicates with what White and Rawlins (2018) referred to as ‘low content arguments’, as shown in \Next for \emph{know}.
+The predicates were embedded under negation in stimuli like \Next[a], in the antecedent of a conditional and a polar question in stimuli like \Next[b], and under negation, in the antecedent of a conditional, and in a question in stimuli like \Next[c].
+To assess projection, participants were asked to respond to the question \emph{Did that thing happen?} for stimuli like \Next[a] and to respond to the question posed by stimuli like \Next[b] and \Next[c].
+The response options were ‘yes’, ‘maybe or maybe not’, and ‘no’.
+
+			%Our experimental result that that the effect of entailment-cancelling operator differs by predicate finds further support in the data from \posscite{white_role_2018} MegaVeridicality dataset.
+			%
+			%White \& Rawlins assess projection inferences associated with 517 clause-embedding predicates by presenting them in 50 \lq low-content\rq\ syntactic frames, like the ones in \Next, illustrated here with \textit{know}.
+The frames in \Next present predicates associated with projective contents embedded under negation \Next[a], in conditional antecedents \Next[b], or both \Next[c], in combination with a projection question (\textit{Did that thing happen?}).
+
+			\ex. \a. Somebody didn’t know that a particular thing happened. Did that thing happen?
 				\b. If somebody knows that a particular thing happened, did that thing happen?
@@
 
+To investigate by-operator projection variability, we recoded `yes' responses as 1, `no' responses as -1, and `maybe or maybe not' responses as 0.
+We then calculated the mean projection rating for 16 predicates under the three types of operator combinations shown in \Last.
+The 16 predicates include ones also used in our experiment as purported factives (\textit{be annoyed, know, reveal}) and semi-factives (\textit{discover, see}), and eleven further predicates commonly characterized as factive or semi-factive (\textit{amuse, find out, forget, learn, love, notice, realize, recognize, regret, remember, understand}).
+
+			%In a series of experiments, participants presented with stimuli like \Last responded with \lq yes\rq, \lq no\rq, or  \lq maybe\rq (coded as \texttt{1, -1,} or \texttt{0}, respectively).
+Here, we present data from their data set, for the frames in \Last, and 16 of their predicates.
+These include ones also used in our experiment as purported factives (\textit{be annoyed, know, reveal}) and semi-factives (\textit{discover, see}), and eleven further predicates commonly characterized as factive or semi-factive (\textit{amuse, find out, forget, learn, love, notice, realize, recognize, regret, remember, understand}).
+
+			The mean projection ratings by embedding context and predicate are presented in Figure~\ref{fig:figure4}, which shows that effect of embedding context on projection differs by predicate.
+
+			For instance,\footnote{Give some examples of differences here, but first double check if lines are labeled correctly. Also angle labels on x-axis.}
+
+			% the predicates \textit{see} and \textit{learn} have the same mean ratings for both negative and conditional contexts.
+However, the effect of negative conditional contexts does not simply reflect the additive result of both effects, but
+
 			\begin{figure}[ht]
@@
 				\caption{Mean projection ratings by embedding context and predicate.}
-				\label{fig:figure1}
+				\label{fig:figure4}
 			\end{figure}
 
-		\subsubsection{Interim conclusion}
+			It is worth noting that the task assesses global projection only in case of \Last[a], but not \Last[b+c], where the question \textit{did it happen?} is embedded in the conditional consequent.
+It assesses whether the projective inference holds in the conditional local context.
+Nevertheless, the data provides supporting evidence for by-predicate variation in the effect of operators on projection.
+Crucially, the by-predicate variation in the effect of negation in the global context is different from the by-predicate variation in the effect of negation in the conditional context.
 
-			Results:
+\section{General discussion}\label{s:general}
 
-			\begin{enumerate}
-				\item Effect of operator on projection
+In this section, we point out methodological implications of our results (section \ref{s:method-impl}), discuss whether contemporary projection analyses can capture the observed variation (section \ref{s:analysis}), and speculate about lexical differences between the clause-embedding predicates that might predict the by-operator variation observed (section \ref{s:lex}).
 
-				\item Effect of operator on projection differs by predicate
+\subsection{Methodological implications}\label{s:method-impl}
 
-				\item By-predicate variation on operator effects are gradient and cannot be predicted by categorical lexical classes
+A methodological implication of this latter result is that future research must relativize claims about projection variation to the entailment-canceling operator.
+
+\subsection{Theoretical implications}\label{s:analysis}
+
+In this section, we consider whether contemporary analyses can capture the observed by-predicate projection variation and the interactions between predicate and entailment-canceling operator in projection ratings.
+We consider three types of analyses: (i) analyses that assume that the CCs of factive predicates are lexically specified as presupposed  (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) analyses on which backgrounded entailments are projective (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) an analysis on which contextual entailments are projective
+%with a triggering algorithm based on world-knowledge that operates over contextual entailments of an utterance 
+(\citealt{schlenker_triggering_2021}).
+
+%While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).
+
+	%Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses.
+
+\subsubsection{Analyses on which presuppositions are lexically specified}
+
+On the analyses in \citealt{heim_projection_1983,heim_presupposition_1992} and \citealt{van_der_sandt_presupposition_1992}, the CCs of factive predicates are lexically specified as presupposed, in contrast to the CCs of nonfactive predicates.
+Presupposed CCs project to the global context, except when that would produce an inconsistency, in which case they are accommodated to the local context of the operator.
+
+\paragraph{By-predicate variation.} Analyses on which some clause-embedding predicates are factive and others are nonfactive do not predict the observed by-predicate variation, for several reasons.
+One reason is that there is variation between the factive predicates: For instance, the CC of \emph{be annoyed} is more projective than that of \emph{discover}, which in turn is more projective than that of \emph{reveal}.
+This variation is not predicted because the CCs of factive predicates are invariably specified as presupposed.
+The variation is also not predicted because, in our experiment, the stimuli were presented in minimal contexts that were not inconsistent with the CCs, thereby rendering local accommodation unlikely.
+A second reason is that the CCs of some nonfactive predicates (e.g., \emph{inform, acknowledge})  are just as projective as those of some factive predicates, or even more projective.
+Analyses like those in \citealt{heim_projection_1983,heim_presupposition_1992} and \citealt{van_der_sandt_presupposition_1992} fail to make predictions about the projection of the CCs of nonfactive predicates.
+Thus, as discussed in detail in \citealt{degen_are_2022}, the factive/nonfactive distinction is not sufficiently fine-grained to adequately predict the observed by-predicate projection variation.
+
+%The assumption of categorical lexical triggering does not address the projection variation among purported factive predicates (see also discussion in \citealt{degen_are_2022}).
+It also cannot address the fact that the CCs of some non-factive predicates (e.g., \emph{inform, acknowledge}) project just as much, or more, than some (semi-)factive predicates (e.g., \emph{reveal, discover, see}).
+That is because, it only makes predictions for the projection of the CC of (semi-)factive predicates, whose CCs are analyzed as presuppositions.
+It therefore does not address the data for the non-veridical predicates in our study.
+		%
+%		Further, the categorical lexical distinctions are not sufficiently fine-grained to make predictions about the gradient by-predicate variation in the effect of operator on projection.
+		%
+%		This interaction effect can only be addressed by taking into account more fine-grained lexical distinctions than assuming that some predicates are presuppositional, while others are not.
+
+	\paragraph{By-operator variation.} 
+        %In our experiment, the stimulus utterances were presented out of the blue, leaving it to the participants to imagine a suitable discourse context.
+Therefore, none of the CCs should be interpreted as necessarily inconsistent with contextual information.
+		%
+		The analyses in \citealt{heim_projection_1983,heim_presupposition_1992} and \citealt{van_der_sandt_presupposition_1992} do not lead us to expect differential effects of entailment-cancelling operators on projection.
+This is because negation and conditional antecedents are given a semantics that derives their behavior as presuppositional \lq holes\rq\ (in the sense of \citealt{karttunen_observations_1971}).
+For instance in \citealt{heim_projection_1983}, presupposition triggers under negation or in conditional antecedents are evaluated relative to the global context set.
+Therefore, presupposition triggers under these operators are predicted to behave like unembedded ones, with no expected differences.
+While the analyses does not explicitly address epistemic possibility modals or polar questions, we might expect that they would be treated as presuppositional holes along the same lines, therefore also not predicting the observed projection variation.
+		%
+		%Differential effects of operators on projection could conceivably be addressed in such a system by assuming that the meaning of entailment-cancelling operators systematically affects the possibility of local accommodation (e.g., by assuming that accommodation is more likely under modals than conditional antecedents).
+However, no such effects have been spelled out.
+
+	\subsubsection{Analyses on which backgrounded entailments are projective}
+
+		\citealt{abrusan_predicting_2011,abrusan_presupposition_2016} and \citealt{simons_best_2017} do not assume that presuppositions are lexically specified but rather that those entailments that are backgrounded with respect to the discourse context in which the utterance is made may project.
+%assume that projection is not determined lexically, but by a triggering mechanism, which operates over the entailments of an expression, and determines which of them projects.
+They assume entailments project that do not contribute to the main point invoked by the utterance.
+		% Instead of assuming that the possibility of projection is determined lexically, the accounts in \cite{abrusan_predicting_2011,simons_best_2017} and \citet{schlenker_triggering_2021} assume a triggering mechanism, which operates over a set of inferences associated with an utterance, and determines which of them projects.
+The assumed triggering mechanisms differ on two dimensions: (i) how the set of inferences is determined that constitutes the input to the triggering mechanism, and (ii) the algorithm by which these inferences are derived as projective or not.
+		In \citealt{abrusan_predicting_2011} and \citealt{abrusan_presupposition_2016}, an entailment is backgrounded if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}).
+In \citealt{simons_best_2017}, an entailment is backgrounded if it is not at-issueness relative to the QUD and not conventionally marked as at-issue.
+		%
+		%In such analyses, every utterance addresses a question, which is determined by its alternative-structure, which, can be indicated by focus-marking.
+For example, in a modalized utterance with \emph{be annoyed}, the CC can be focused \Next[b], which leads to the alternative structure in \Next[c], allowing the utterance to address a question like \Next[a].
+
+		%\ex. \a. What was Cole annoyed about today?
+		%	\b. Perhaps Cole was annoyed [$_F$ that Julian dances salsa].
+		%	\b. $\{p \mid \exists q(p = \lambda w.$\emph{Cole was annoyed that} $q$ \emph{in} $w)\}$
+		%	\z.
+		%\z.
+
+		%In contrast, focus on the subject \Next[b] leads to the alternative structure in \Next[c], allowing the utterance to address a question like \Next[a].
+
+		%\ex. \a. Who was annoyed today?
+		%	\b. Perhaps [$_F$ Cole] was annoyed that Julian dances salsa.
+		%	\b. $\{p \mid \exists x(p = \lambda w.x$ \emph{was annoyed in} $w$ \emph{that Julian dances salsa}$)\}$
+		%	\z.
+		%\z.
+
+		%The account assumes that an inference projects if it is entailed by the union of the alternatives (see also \citealt{abusch_presupposition_2010}).
+therefore predicting projection of the CC in \Last, but not \LLast.
+That is, under the assumption that $x$ \emph{was annoyed that} $p$ entails $p$.
+
+		\paragraph{By-predicate variation.} Like the analyses discussed in the previous section, the analyses in \citealt{abrusan_predicting_2011,abrusan_presupposition_2016} and \citealt{simons_best_2017} do not predict the observed by-predicate variation.
+First, if the CCs of factive predicates are entailed, they are invariably predicted to project, and so the observed differences are not predicted.
+Second, the analyses do not make predictions for nonentailed CCs, and hence they do not predict that the CCs of some nonfactive predicates project at least as much or even more than those of some factive predicates.
+
+%  The triggering mechanisms assumed in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017}, predict a potential of projection for inferences that are entailed by the literal content of the modal prejacent $p$ in \Last and \LLast, and similarly for propositions embedded under negation, and in conditional antecedents or polar questions.
+These accounts, therefore, predict a potential of projection for CCs that are entailed by the prejacent, due to the lexical entailments of veridical predicates, or due to entailments that are contributed compositionally by the literal content (see e.g., \citealt{roberts_i_2019}).
+However, they do not make systematic predictions for CCs, such as in \Last, that are not entailed.
+As a result, these accounts cannot address the fact that we see some amount of projection for all of the tested predicates.
+
+		%Because the main point of an utterance is often determined in the discourse context (\citealt{roberts_information_1996,roberts_information_2012}, \emph{et seq.}), these approaches incorporate discourse effects on projection.
+		%
+		% \citet{abrusan_predicting_2011} also makes predictions for utterances in out-of-the-blue contexts: Utterances are associated with a default main point that is grammatically determined by the time at which the matrix predicate is interpreted.
+		%
+		%The account in \citealt{simons_best_2017} could be extended to address our gradient data in out-of-the blue contexts, by assuming (i) that embedded contents that are not at-issue can project even if they are not entailed, and (ii) that linguistic expressions differ in how likely the contributed contents are interpreted as at-issue (as suggested in \citealt{tonhauser_how_2018}).
+		%
+		%In our data, for instance, the CC of \emph{discover} \Next[a] has received relatively high projection ratings, while that of \emph{confirm} \Next[b] has received medium to low ratings.
+To explain this difference, the information-structure based triggering approach would need to assume that the CC of \emph{discover} is less likely to be at issue than that of \emph{confirm}.
+
+	%	\ex. \a. Perhaps Cole discovered that Julian dances salsa.
+	%		\b. Perhaps Cole confirmed that Julian dances salsa.
+%			\z.
+%		\z.
+
+		%While \citet{tonhauser_how_2018} experimental evidence that projection ratings negatively correlate with at-issueness ratings in out-of-the-blue polar questions.
+\citet{sieker_projective_2022} argue that their data also shows the same correlation for modals, conditional antecedents, and negation for the German clause-embedding predicates they investigated.
+To address the gradient by-predicate variation theoretically, this approach would need to explicitly spell out how various predicates are associated with different likelihoods that their CCs is interpreted as at-issue.
+
+		\paragraph{By-operator variation.} Neither of the analyses make predictions about by-operator variation.
+			%Since we found an operator/predicate interaction effect on projection, a further prediction of this type of approach would be that at-issueness ratings would show an operator-predicate interaction in this experimental setup as well.
+To address the interaction effect theoretically, an information structural triggering approach would need to make explicit predictions about how the meaning of entailment-cancelling operators affects at-issueness differentially for different predicates.
+This is certainly conceivable, when we take default assumptions about how likely embedded contents are interpreted at issue to be associated with the entire embedding context, including clause-embedding predicates, but also entailment-cancelling operators.
+However, this has not been spelled out explicitly and needs to be investigated in more detail.
+
+\subsubsection{Triggering based on contextual entailment}
+
+The triggering mechanism assumed in \citealt{schlenker_triggering_2021} operates on local contextual entailments of an expression, and determines which one of them projects based on probabilistic world knowledge about the truth of the invoked contents.
+For instance, the CC of the utterance with \emph{discover} in \Next[a], that Julian dances salsa, has the potential to project if it is entailed by the sentence in the scope of negation (namely \Next[b]) together with information that is available in its local context.
+Negation is treated as a presuppositional hole, and the local context under negation is taken to be the global context (as in  \citealt{heim_projection_1983} and \citealt{schlenker_local_2009}).
+
+			\ex. \a. Cole didn't discover that Julian dances salsa.
+				\b. Cole discovered that Julian dances salsa.
+			\z.
+
+			The triggering mechanism states that an inference $p$ of some propositional expression $E$ pro\-jects if $p$ is an epistemic precondition for the truth of $E$ in its local context $c^\prime$.
+So, the CC of \Last[a] projects if the CC is an epistemic precondition for the truth of \Last[b].
+			%
+			$p$ is an epistemic precondition of $E$ in $c^\prime$ if, usually, when one acquires the belief in $c^\prime$ that $E$, one already knows that $p$. 
+			%
+			The CC of \Last[a] can be taken to be an epistemic precondition for \Last[b] based on the \lq subjective conditional probability\rq\ that a generic epistemic agent would already believe that Julian dances salsa when learning that \Last[b] is true.
+Specifically, the CC is taken to be an epistemic precondition if that probability reaches a contextually given threshold.
+
+			\paragraph{By-predicate variation.} \citealt{schlenker_triggering_2021} predicts   
+   the potential of projection for CCs that are contextually entailed by the variants not embedded under an entailment-canceling operator.
+Thus, in contrast to the analyses reviewed above, this analysis can, in principle, predict the projection of non-entailed CCs, namely when the CCs are contextually entailed.
+While this is a step in the right direction, it is not clear that the analysis can predict the projection of non-entailed CCs.
+Consider \Next, the variant of \Last[a] with {\em inform}:
+
+   \ex. \a. Cole didn't inform Sam that Julian dances salsa.
+       \b. Cole informed Sam that Julian dances salsa
+
+On \poscite{schlenker_triggering_2021} analysis, the CC projects in \Last[a] if, usually, when one acquires the belief that \Last[b], one already knows that Julian dances salsa.
+While it is certainly possible for Julian dancing salsa to be already known when one acquires the belief that \Last[b], it is not clear that this is ``usually'' the case.
+(Another question, of course, is what ``usually'' means.)
+
+It is also not clear that this analysis can predict the observed by-predicate variation. To illustrate, consider the schemata in \Next:
+
+\ex. \a. Cole didn't PREDICATE that Julian dances salsa.
+    \b. Cole PREDICATE that Julian dances salsa.
+
+Given that the non-entailed CC of a given predicate projects in \Last[a] only if one ``usually'' already knows that the CC is true when one acquires the belief that \Last[b], \poscite{schlenker_triggering_2021} analysis seems to predict that predicates fall into two groups: Those whose CCs one ``usually'' already knows to be true when one acquires the belief that \Last[b], and those for which this is not the case.
+If so, this analysis cannot predict the observed by-predicate projection variation.
+
+%   In order to extend an account along these lines to address the out-of-the-blue utterances in our experiment, we might assume that these utterances are interpreted relative to an imagined plausible discourse context (for discussion, see \citealt{simons_best_2017}).
+
+   %It can therefore incorporate effects of contextual information on projection, such as the reliability of the source of a speech report (\citealt{rieh_credibility_2010,de_marneffe_did_2012}).
+
+%Due to the probabilistic nature and context-sensitivity of the triggering mechanism, the approach can make gradient predictions.
+However, the subjective conditional probabilities associated with expressions and their contextual inferences are taken as given in the account.
+In order to make concrete predictions about the gradient by-predicate variation, this account would need to make explicit how the relevant probabilities are derived based on different predicates.
+For instance, to explain the contrast that \emph{discover} receives higher projection ratings than \emph{confirm}, this account would assume that the situations in which one learns that a sentence like \emph{$x$ discovered that $p$} is true, are such that one already knew that $p$ is true more often than situations in which one learns the truth of a sentence like \emph{$x$ confirmed that $p$}.
+
+\paragraph{By-operator variation.} \citealt{schlenker_triggering_2021} does not incorporate differential effects of entailment-cancelling operators on projection.
+Although subjective conditional probabilities are relativized to a local context, the local context under negation and in conditional antecedents is assumed to the global context (based on \citealt{heim_projection_1983,schlenker_local_2009}).
+Therefore, no differential effect of operator is predicted, contrary to what is observed.
+For instance, the conditional probability for $p$ given $E$ vs.\ that of $p$ given $(\emph{not } E)$ or that of $p$ given $(\emph{if } p, q)$ should not differ.
+
+%			In order to address the operator-predicate interaction effect, an approach based on epistemic preconditions would need to make explicit how the meaning of entailment-cancelling operators affects the conditional probabilities differentially for different predicates.
+			%
+%			The conditional probability that some embedded content is antecedently known would need to be computed relative to a global context, and take into account the entire embedding context, including the combination of entailment-cancelling operator and clause embedding predicate.
+
+	\subsubsection{Discussion: Contextual inferences and probabilistic triggering}
+
+		Our data showed gradient projection variation, based on a main effect of predicate, and an operator-predicate interaction.
+We found these effects for the CCs of clause-embedding predicates under four entailment-cancelling operators in out-of-the-blue contexts.
+The fact that we see some amount of projection for all predicates (even non-veridical ones), can only be addressed by an account that assumes the potential of projection for contextual inferences, rather than only literal ones.
+
+		The main effect of predicate and the operator-predicate interaction are both gradient effects, and cannot be explained by a categorical distinction between presuppositional and non-pre\-suppo\-si\-tion\-al predicates.
+The effects could rather be addressed in a probabilistic account, where various embedding contexts (i.e., combinations of entailment-cancelling operator and predicate) are associated with different probabilities that the embedded content projects.
+
+		Under the assumption of information-structural triggering, this is the probability that the embedded content is not at issue in the (possibly imagined) discourse context.
+Under a world-knowledge-based triggering account, the projection probability can be associated with the probability that the embedded content is already known in the discourse context.
+These two notions are, of course, related.
+If some proposition is already known, raising an issue about whether $p$ is the case will be uninformative.
+Therefore, it is unclear whether these two approaches would make different predictions for our data.
+Both types of accounts would need to be extended to make concrete predictions for the contribution of the combinations of entailment-cancelling operators and predicates.
+
+\subsection{Lexical patterns}\label{s:lex}
+
+Can the observed interactions between predicate and entailment-canceling operator be predicted from lexical semantic/pragmatic properties of the predicates?
+This is a pressing question for future research, to which our data offer some tentative answers.
+
+We can find some initial generalizations over lexical properties, indicated in Figure~\ref{fig:patterns}, which gives the mean certainty ratings for the four operators by predicate, identifying four groups of predicates that show similar by-operator variation.
+
+	\begin{figure}[ht]
+		\centering
+		% \hspace{-.5em}\resizebox{\linewidth}{!}{
+		\begin{tabular}{p{.205\linewidth} p{.18\linewidth} p{.18\linewidth} p{.18\linewidth} p{.18\linewidth}}
+			\footnotesize \hspace{2.1em} (a) Negation high &
+			\footnotesize \ (b) Negation low &
+			\footnotesize (c) Conditional high &
+			\footnotesize \ (d) Modal low
+			% & \footnotesize \  (e) Discover
+			\\
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.225 \textwidth, valign=T]{negation-high-1.pdf}
+			&
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.2\textwidth, valign=T]{negation-low-1.pdf}
+			&
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.2\textwidth, valign=T]{conditional-high-1.pdf}
+			&
+			\vspace{-.6\baselineskip}
+			\includegraphics[width=.2\textwidth, valign=T]{modal-low-1.pdf}
+			% &
+			% \vspace{-.6\baselineskip}
+			% \includegraphics[width=.2\textwidth, valign=T]{discover-1.pdf}
+			\\
+		\end{tabular}
+		% }
+
+		\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).}
+		\label{fig:patterns}
+	\end{figure}
+
+	%
+	The non-veridical predicates \emph{pretend} and \emph{think} exhibit the `Negation high' pattern, shown in panel (a) of Figure~\ref{fig:patterns}.
+These are the only predicates that are most projective under negation compared to all other operators.
+This could be related to
+	assumptions that non-veridical doxastics like \emph{think} can often be interpreted in comparison to a stronger veridical alternative (e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
+We tentatively hypothesize that this can lead to an inference that the CC is false in upward-monotone contexts, but not under negation.
+A similar alternative seems to be salient for \emph{pretend}, but not the non-veridical communicative predicates \emph{say} or \emph{suggest}.
+
+	The inferential predicates \emph{confirm, demonstrate}, and \emph{prove} exhibit a `Negation low' pattern, shown in panel (b).
+These are most projective under modals and conditionals and least projective under negation.
+Here, we may hypothesize that a negative statement involving these predicates is often only taken as relevant (in a Gricean sense) when it is antecedently known that there was an attempt to confirm/demonstrate/prove the truth of the CC.
+This may result in an inference that this attempt failed and result in lower projection ratings under negation.
+
+	For \emph{admit, announce, confess}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This `Conditional high' pattern (c) may suggest that the discourse effect of a conditional interacts with these change-of-state communication predicates.
+
+	Finally, the predicates \emph{hear, inform,} and \emph{know} exhibit a `Modal low' pattern (d).
+The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
+
+\section{Conclusion}
+
+\bibliographystyle{sub-chicago}
+\bibliography{projection}
+
+\newpage
+\appendix
+
+\section{20 clauses}\label{app:a-clauses}
+    The contents of the following 20 clauses, which realized the complements of the 20 clause-embedding predicates, were investigated in Exps.~1-4:
+
+    %\begin{multicols}{2}
+    \begin{enumerate}%[leftmargin=3ex,itemsep=-2pt]
+        \item Mary is pregnant.
+        \item Josie went on vacation to France.
+        \item Emma studied on Saturday morning.
+        \item Olivia sleeps until noon.
+        \item Sophia got a tattoo.
+        \item Mia drank 2 cocktails last night.
+        \item Isabella ate a steak on Sunday.
+        \item  Emily bought a car yesterday.
+        \item  Grace visited her sister.
+        \item Zoe calculated the tip.
+
+    %\columnbreak
+
+        \item  Danny ate the last cupcake.
+        \item  Frank got a cat.
+        \item  Jackson ran 10 miles.
+        \item  Jayden rented a car.
+        \item  Tony had a drink last night.
+        \item  Josh learned to ride a bike yesterday.
+        \item  Owen shoveled snow last winter.
+        \item  Julian dances salsa.
+        \item  Jon walks to work.
+        \item  Charley speaks Spanish.
+
+    \end{enumerate}
+    %\end{multicols}
+
+\section{Consequents for conditional target stimuli in Exps.~4} \label{app:b-consequents}
+    We created 20 consequent clauses for the three experiments in which the 20 clauses were embedded in the antecedent of a conditional (Exps.~4).
+Each of the 20 clauses was paired with a unique consequent clause, as shown in the list below.
+To minimize the variability of the effect of the contents of these consequent clauses on the projection of the contents of the 20 complement clauses, the consequent clauses all consist of a uniquely named subject and an adjectival predication in the future tense ({\em will be}), and the adjectives all denote an emotion.
+We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner_norms_2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
+
+    \begin{enumerate}%[leftmargin=3ex,itemsep=-2pt]
+
+    \item \ldots that Mary is pregnant, Esther will be mad.
+    \item \ldots that Josie went on vacation to France, Arnold will be frustrated.
+    \item \ldots that Emma studied on Saturday morning, Liam will be proud.
+    \item \ldots that Olivia sleeps until noon, Elijah will be embarrassed.
+    \item \ldots that Sophia got a tattoo, Ariel will be giddy.
+    \item \ldots that Mia drank 2 cocktails last night, Mariela will be worried.
+    \item \ldots that Isabella ate a steak on Sunday, Liz will be delighted.
+    \item  \ldots that Emily bought a car yesterday, Kate will be excited.
+    \item \ldots that  Grace visited her sister, Henry will be surprised.
+    \item \ldots that Zoe calculated the tip, Alex will be astonished.
+    \item  \ldots that Danny ate the last cupcake, Harper will be disgusted.
+    \item  \ldots that Frank got a cat, Lucas will be grouchy.
+    \item  \ldots that Jackson ran 10 miles, Kayla will be cheerful.
+    \item  \ldots that Jayden rented a car, Brittany will be furious.
+    \item  \ldots that Tony had a drink last night, Victoria will be ashamed.
+    \item  \ldots that Josh learned to ride a bike yesterday, Mason will be envious.
+    \item  \ldots that Owen shoveled snow last winter, Bianca will be jealous.
+    \item  \ldots that Julian dances salsa, Logan will be joyful.
+    \item  \ldots that Jon walks to work, Caleb will be suspicious.
+    \item  \ldots that Charley speaks Spanish, Jay will be happy.
 
@@
 
-			Methodological implications:
-			\begin{itemize}
-				\item projection ratings do not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.
+\section{Control stimuli in Exps.~1-3}\label{app:c-control}
 
-				\item Additionally, claims about projection variability must be relativized to the entailment-canceling operator.
+    The control stimuli in Exps.~1-3 were the contents of main clauses.
+In Exps.~1q, 2q and 3q, the control stimuli consisted of the polar questions in (1).
+The non-restrictive relative clauses (NRRCs), given in parentheses in (1), were included in Exps.~2q and 3q, where at-issueness was measured with an assent diagnostic.
+The control stimuli here consisted of two clauses (like the target stimuli), to allow the relevant speaker to assent with one of two clauses.
 
-				\item but we can still use FoS test in teaching, JT made this claim, but I don't understand the argument... help?
+    \ex.[(1)] Sentences for control stimuli in in question embedding experiments (Exps.~1q, 2q and 3q)
+      \a. Do these muffins (, which are really delicious,) have blueberries in them?
+      \b. Does this pizza (, which I just made from scratch,) have mushrooms on it? 
+      \b. Was Jack (, who is my long-time neighbor,) playing outside with the kids? 
+      \b. Does Ann (, who is a local performer,) dance ballet?
+      \b. Were John's kids (, who are very well-behaved,) in the garage?
+      \b. Does Samantha (, who is really into fashion,) have a new hat?
+      \z.
+    \z.
+
+    We expected participants to give low responses on the `certain that' diagnostic for the control stimuli in (1), indicating that the speaker is not certain of the main clause content, because main clause content is hypothesized to not project out of polar questions.
+These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for Exps.~1q, 2q, and 3q.
+We also expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (1), indicating that the main clause content is at-issue.
+These expectations were borne out for Exps.~1q and 2q, as shown in the fourth column of Table~\ref{t-controls}, but the ratings were higher than expected for Exp.~3q.
+See below for discussion.
+
+    In the remaining experiments, the control stimuli consisted of the positive declarative variants of (1) given in (2).
+The NRRCs in parentheses were realized in Exps.~2 and 3 for the reason explained above, as well as in Exp.~1c, to make the control stimuli more similar to the target stimuli (which also consisted of two clauses, namely the antecedent and the consequent).
+
+    \ex.[(2)]  Sentences for control stimuli in negation, modal and conditional embeddings
+      \a. These muffins (, which are really delicious,) have blueberries in them.
+      \b. This pizza (, which I just made from scratch,) has mushrooms on it. 
+      \b. Jack (, who is my long-time neighbor,) was playing outside with the kids. 
+      \b. Ann (, who is a local performer,) dances ballet.
+      \b. John's kids (, who are very well-behaved,) were in the garage.
+      \b. Samantha (, who is really into fashion,) has a new hat.
+      \z.
+    \z.
+
+    We expected participants to give high responses on the `certain that' diagnostic for the control stimuli in (2), indicating that the speaker is certain of the main clause content, because speakers are hypothesized to be committed to the asserted main clause content.
+These expectations were borne out, as shown in the third column of Table~\ref{t-controls} for the `n', `m' and `c' variants of Exps.~1, 2, and 3.
+We expected participants to give low responses on the at-issueness diagnostics for the control stimuli in (2), indicating that the main clause content is at-issue.
+These expectations were borne out for the `n', `m', and `c' variants of Exps.~1, as shown in the fourth column of Table~\ref{t-controls}, but not for the `n', `m', and `c' variants of Exps.~2 and 3.
+See below for discussion.
+
+    \begin{table}[h!]
+        \centering
+        \begin{tabular}{r r r r l }
+            & &  \multicolumn{2}{c}{Mean ratings} &  \\ 
+            Exp. & Control stimuli & Certainty & Not-at-issueness & At-issueness measure \\ 
+            \hline
+            1q & (1) & .14 & .05  & asking whether $c$ \\
+            1n & (2) &  .95 & .04 & sure that $c$\\
+            1m & (2) & .96 & .03 & sure that $c$\\
+            1c & (2) with NRRC & .94  & .08 & sure that $c$\\
+            \hline
+            2q & (1) with NRRC & .18 & .07 & {\em yes}, $c$\\
+            2n & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
+            2m & (2) with NRRC& .96 & .25 & {\em yes, that's true}, $c$\\
+            2c & (2) with NRRC& .96 & .22 & {\em yes, that's true}, $c$\\
+            \hline
+            3q & (1) with NRRC & .17 &  .28 & {\em yes}, but $\neg c'$\\
+            3n & (2) with NRRC & .94 & .44 & {\em yes, that's true}, but $\neg c'$\\
+            3m & (2) with NRRC & .93 & .50 & {\em yes, that's true}, but $\neg c'$\\
+            3c & (2) with NRRC & .93 & .53 & {\em yes, that's true}, but $\neg c'$\\
+            \hline
+        \end{tabular}
+        \caption{Mean certainty and at-issueness ratings for control stimuli, for self-declared American English participants}\label{t-controls}
+    \end{table}
+
+    As mentioned above, the mean at-issueness ratings were higher than expected in the `n', `m', and `c' variants of Exps.~2 and all four of the Exps.~3.
+We discuss these exceptions in more detail here because they allow us to further understand the various measures for at-issueness investigated in this paper.
+The example in (3a) illustrates the version of the assent diagnostic applied in Exps.~2n, 2m, and 2c: the assent particle {\em yes} is followed by the clause whose content is diagnosed, here, the content of the main clause.
+The related affirmation diagnostic applied in Exp.~2q (where the mean at-issueness rating was as low as expected) is illustrated in (3b).
+
+    \ex.[(3)]
+        \a. Sample control stimulus in Exps.~2n, 2m, and 2c
+            \a.[A:] These muffins, which are really delicious, have blueberries in them.
+            \b.[B:] Yes, that's true, they have blueberries in them.
+            \z.
+        Question to participants: Does A's response to B sound good?
+        \b. Sample control stimulus in Exp.~2q
+            \a.[A:] Do these muffins, which are really delicious, have blueberries in them?
+            \b.[B:] Yes, they have blueberries in them.
+            \z.
+        Question to participants: Does A's response to B sound good?
+        \z.
+    \z.
+
+    As shown in Table \ref{t-controls},  the group mean on the control stimuli is numerically higher (at .22 or .25) than for Exp.~2q (.07).
+We hypothesize that a possible explanation for this difference is that participants take A to assert both the content of the main clause and of the NRRC in (3a) but is only asking about the main clause content in (3b).
+Whereas B's affirmation in (3b) specifies the one content that is affirmed, B's assent in (3a) specifies only one of the two contents that A asserted, seemingly leaving out a specification of the second content, that of the NRRC.
+Participants may judge B's response in (3b) to be less acceptable than in (3a) because of this missing content specification.
+This hypothesis is consistent with the results of \posscite{syrett_experimental_2015} Exp.~3, which suggests that content of a sentence-medial NRRCs can be the target of a direct denial, though the main clause content is preferred as the target of such a denial.
+That the group means on the control stimuli in our Exps.~2n, 2m, and 2c are still relatively low may be due to the fact that the one content that was specified is the at-issue main clause content.
+One would expect lower acceptability ratings in a version of this diagnostic in which the one content that was specified is the NRRC.
+
+    The mean at-issueness ratings for the control stimuli were also higher than expected in Exps.~3.
+The examples in (4) illustrate the versions of the affirmation and assent diagnostics used in these experiments:
+
+    \ex.[(4)]
+        \a. Sample control stimulus in Exp.~3q
+            \a.[A:] Do these muffins, which are really delicious, have blueberries in them?
+            \b.[B:] Yes, but they aren't really delicious.
+            \z.
+        Question to participants: Does A's response to B sound good?
+        \b. Sample control stimulus in Exps.~3n, 3m, and 3c
+            \a.[A:] These muffins, which are really delicious, have blueberries in them.
+            \b.[B:] Yes, that's true, they but they aren't really delicious.
+            \z.
+        Question to participants: Does A's response to B sound good?
+        \z.
+    \z.
+
+    The mean at-issueness rating for the control stimuli in Exp.~3q was comparatively higher (at .28) than for Exp.~1q (.05) or Exp.~2q (.07).
+We hypothesize that that this difference (especially to Exp.~2q) is due the content of the NRRC being directly denied, even though it was presented as backgrounded content in A's question (as shown in (2), the content of none of the other NRRCs was a matter of personal taste).
+
+    The mean at-issueness ratings for the control stimuli in Exps.~3n, 3m, and 3c were numerically even higher than for Exp.~3q and, in fact, the highest across all 12 experiments (at .44, .50, and .53, respectively).
+There are two factors that could be implicated in the difference between Exp.~3q and Exps.~3n, 3m, and 3c: first, the NRRC is included in a polar question in the former but a declarative assertion in the latter; second, B utters an affirmation {\em yes} in the former but an assent {\em yes, that's true} in the latter.
+For instance, participants might judge B's direct denial of the content of the NRRC as less acceptable when A uttered a declarative assertion and B assented with that assertion using {\em yes, that's true} than when A uttered a polar question and B responded in the affirmative with {\em yes}.
+Both factors must be considered in future research on at-issueness measures.
+
+\section{Participant information and data exclusion criteria}\label{app:d-participants}
+
+    This supplement provides information on the participants of the 12 experiments and the criteria by which participants' data were excluded.
+The first three columns of Table \ref{t:exclusion} show, for each of the 12 experiments how many participants were recruited, their age range and mean age, and their self-reported gender; no gender data was collected in Exp.~1q.
+The next three columns provide information on the number of participants whose data were excluded based on the following criteria:
+
+    \begin{itemize}%[itemsep=-2pt]
+
+        \item `multiple': Due to an experimental glitch, some participants participated more than once in Exp.~1q.
+Since no information was available on which one was their first take, those participants' data was removed.
+
+        \item `language': Participants' data were excluded if they did not self-identify as native speakers of American English.
+
+        \item `controls': Participants' data were excluded if their mean rating on the 6 main clause control items in the projection block was more than 2 sd above the group mean (in Exps.~1q, 2q, and 3q) or more than 2 sd below the group mean (in the remaining experiments).
+Participants' data were also excluded if their mean rating on the 6 main clause control items in the at-issueness block was more than 2 sd above the group mean (across all experiments).
+
+        \item `variance': Participants' data were excluded if they always selected roughly the same point on the response scale for the target stimuli.
+To identify such participants, we first identified participants whose mean variance on the target stimuli was more than 2 sd below the group mean variance and then manually inspecting their response patterns.
+The data of participants who used the full scale was not excluded.
 
@@
 
-			Theoretical implications: Accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability.
-The following section argues that our results provide evidence for a triggering algorithm based on contextual rather than lexical or conceptual information (e.g., in the sense of \citealt{abrusan_predicting_2011,simons_best_2017}), and that the input to this triggering algorithm should be contextual rather than lexical inferences (e.g., in the sense of \citealt{schlenker_triggering_2021}).
+    The remaining columns of Table \ref{t:exclusion} provide information on the remaining participants, that is, the participants' data that entered into the analysis.
+Participants took around 9-11 minutes to complete the various experiments.
+Participants were paid more in Exps.~1c, 2c, and 3c than the remaining experiments because the target stimuli in those experiments were longer (as they consisted of conditionals).
+More women than men were recruited in many of the experiments because the experiments were run at a time when Prolific went viral on TikTok, resulting in a large number of young women registering for the service (around July 24, 2021; see \url{https://blog.prolific.co/we-recently-went-viral-on-tiktok-heres-what-we-learned/}, \\ last accessed February 4, 2022).
 
-\section{Theoretical implications}
-	Our results---that projection is modulated by entailment-canceling operators, that there is by-predicate variation in the effect of operator on projection, and that this variation is gradient in ways that cannot be predicted by categorical lexical classes---are not captured by contemporary projection analyses, for several reasons.
-
-	Here, we discuss (i) dynamic approaches, which assume that projection is determined lexically for factive predicates (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) approaches which assume that projection is determined by a contextual triggering algorithm that operates over the lexical entailments of an utterance (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) approaches with a conceptual triggering algorithm that operates over contextual entailments of an utterance (\citealt{schlenker_triggering_2021}).
-While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).
-
-	- effect of different entailment-cancelling operators, not captured (heim vs santd, if projects projects; abtrusan operators?; schlenker can be imagined) in all systems it is conceivable, but no systemaitc itneraction spelled out
-
-	- rely on lexical classes, factives, veriudical, no predictions of non-veridicals
-
-	- not capture how the operator differentially affects different predicates (contextual entailment may provied an approach, but also needs to be spelled out yet)
-
-	The first is that contemporary analyses do not lead us to expect interactions with different types of entailment-cancelling operators.
-While it is conceivable for the meaning of the operator to systematically interact with the possibility of local accommodation, no such interaction has been spelled out systematically.
-
-	The second reason is that many contemporary analyses do not make predictions for the projection of the CC of many of the 20 predicates, as they are limited to (semi-)factive predicates, whose CCs are analyzed as presuppositions (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), or entailed CCs that project unless at-issue with respect to the Question Under Discussion (e.g., \citealt{abrusan_predicting_2011,simons_best_2017}).
-A possible exception is the analysis of \citealt{schlenker_triggering_2021}, which predicts the potential for projection for CCs that are contextually entailed.
-In a full talk, we would like to discuss how this analysis might be extended to capture the gradient projection observed in our experiment.
-
-	The third reason is that contemporary projection analyses do not make sufficiently fine-grained distinctions between different clause-embedding predicates (but only between whether the CC is a presupposition or entailed).
-Consequently, they do not make predictions about the by-predicate variation in the effect of operator on projection.
-
-	\subsection{Factivity: Lexical triggering}
-
-		In \citet{heim_projection_1983}, for instance, the CC of (semi-)factive predicates projects to the global context, except when that would produce an inconsistency, in which case the CC is accommodated to the local context of the operator.
-
-	\subsection{Entailments and contextual triggering}
-
-	\subsection{Contextual entailment and conceptual triggering}
-
-	\subsection{Conclusion: Contextual entailment and contextual triggering}
-		\begin{itemize}
-			\item contextual entailments (but weaker notion of entailment)
-
-			\begin{itemize}
-
-				\item cant believe its not lexical roberts
-
-				\item pragmatics of full embedding context; Does Cole think that Julian dances salsa?
-
-				\item why would that be the case?
-
-			\end{itemize}
-
-			\item contextual triggering
-
-			\begin{itemize}
-
-				\item what is the difference between know and prove (in terms of conceptual precondition, causal model?) there is no conceptually inherent relation between the two, the difference is lexical
-
-				\item QUD influences
-
-				\item what is the relationship between lexical differences and QUD?
-
-			\end{itemize}
-
-		\end{itemize}
-
-\section{Lexical differences}
-
-	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting initial generalizations over lexical properties, indicated in \textbf{Figure\ 2}.
-
-	So, can the observed interaction between predicate and operator in mean projection ratings be predicted from lexical semantic/pragmatic properties of the predicates, and, if so, how?
-	This is a pressing question for future research, to which our data offer some tentative answers. We identify four major patterns.
-	%
-	The predicates \emph{pretend} and \emph{think} exhibit the {\bf `Negation high'} pattern, shown in panel (a) of \textbf{Figure~2}: We tentatively hypothesize that negation (but not the other operators) interacts with the semantic or pragmatic antiveridicality associated with these predicates.
-	%
-	The inferential predicates \emph{prove, confirm}, and \emph{establish} exhibit a {\bf `Negation low'} pattern, shown in panel (b): Here, we tentatively hypothesize that the veridical meaning component interacts with negation (but not the other operators), to result in lower projection ratings under negation.
-	%
-	For \emph{announce, confess, admit}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This {\bf `Conditional high'} pattern (c) may suggest that the discourse effect of a conditional interacts with the change-of-state communication predicates.
-	%
-	Finally, the predicates  \emph{inform, know}, and \emph{be annoyed} exhibit a {\bf `Modal low'} pattern (d).
-The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
-
-	%%%%
-	In spite of a lack of categorical distinctions about the projection behavior of our verbs, we can find some interesting patterns.
-\textbf{Fig.~\ref{fig:figure2}} gives the mean certainty ratings for the four operators by predicate, identifying some groups of predicates that show similar by-operator variation.
-We highlight four \lq projection-profiles\rq here: \emph{pretend} and \emph{think} are the only predicates that are most projective under negation compared to all other operators (\texttt{N} $>$ \texttt{M, Q, C}).
-\emph{annouce, confess, admit,} and \emph{reveal} are most projective under conditionals, while there is also a tendency that there is more projection from questions compared to modals and negation, a difference that may not be robust for \emph{announce} (\texttt{C} $>$ \texttt{Q} \textcolor{gray!40}{$>_?$} \texttt{M, N}).
-\emph{prove, confirm,} and \emph{establish} are more projective under modals and conditionals than under questions and negation (\texttt{M, C} $>$ \texttt{Q, N}).
-Finally, \emph{inform} and \emph{know} are most projective under questions, and least projective under modals (\texttt{Q} $>$ \texttt{N, Q} $>$ \texttt{M}).
-
-\section{Conclusion}
-
-\newpage
-\bibliographystyle{sub-chicago}
-\bibliography{projection}
+    \begin{sidewaystable}[h!]
+      \centering
+      \begin{tabular}{l | r r r | r r r r | r r r | r }
+      &  \multicolumn{3}{c|}{Recruited participants} & \multicolumn{4}{c|}{Exclusion criteria} & \multicolumn{3}{c|}{Remaining participants} &  \\ 
+      Exp. & total & ages (mean) & gender (f/m/o/u) & multiple & language & controls & variance & total & ages (mean) & gender (f/m/o/u) & payment   \\ 
+      \hline
+      1q & 300  & 19-74 (38.2)  & --  & 5  & 7  & 35  & 0  & 242   & 21-74 (39.2) &  -- & \$1.70\\
+      1n & 300  & 18-74 (33.2)  &  150/145/5/0 & 0  & 8 & 17 & 1 & 274  & 18-74 (33.3) & 141/128/5/0 & \$1.70\\
+      1m & 300  & 18-74 (32.7) & 150/141/7/2  &  0 & 0  & 19 & 0   & 281   & 18-74 (32.7) & 144/129/7/1 & \$1.70 \\
+      1c &  300 & 18-58 (25.9)  & 249/45/6/0 & 0 & 6  & 26 & 2  & 266  &  18-58 (24.8) & 235/25/6/0  & \$2.15 \\
+      2q & 250  &  18-58 (25.5) &  201/43/6/0  & 0 & 4  & 24  & 1 & 220  & 18-58 (24.8)  & 187/28/5/0  & \$1.70 \\
+      2n & 250  &  18-69 (33.2)  &  127/114/6/1 & 1  & 4  & 29 & 0& 215  &  18-69 (33.1) & 113/95/6/1    & \$1.70\\
+      2m & 251  & 18-74 (31.7)  & 132/113/6/0  & 0  & 4  & 27 &  0 & 220  & 18-70 (31.9) & 116/98/6/0 & \$1.70\\
+      2c &  250 &  18-56 (24.5)  & 212/30/8/0  & 0  & 0  & 26 & 0  & 224  & 18-56 (24.4) & 195/24/5/0 & \$2.15\\
+      3q & 250  &  18-66 (32.4) &  140/102/7/1 &  0 & 4  & 20  & 0  &  225 & 18-66 (32.6) & 125/93//7/0  & \$1.70 \\
+      3n & 250  &  18-70 (24.6) & 114/31/5/0  & 0  & 5  &  13 & 4  & 228  & 18-70 (24.3) &  198/25/5/0 &\$1.70 \\
+      3m & 250  & 18-63 (25.5)  & 205/40/5/0 &  0 & 3  & 14 &  0 & 233  &  18-63 (24.8) & 197/31/5/0 &\$1.70 \\
+      3c & 250  & 18-59  (27.5) & 182/64/4/0  & 0  & 3  & 17 & 0 & 230  &  18-59 (26.7)  & 177/49/4/0  & \$2.15\\
+      \end{tabular}
+      \caption{Recruited participants, excluded data, and remaining participants in Exps.~1, 2 and 3.
+Gender distinctions were `f' = female, `m' = male, `o' = other, and `u' = undeclared.}\label{t:exclusion}
+    \end{sidewaystable}
 

### 2024-02-14 writing/SuB-paper/projection-operators-SuB.tex
updated analysis scripts

@@
 
-    Textbooks typically present the family-of-sentences diagnostic for projection without addressing by-operator variation (e.g., \citealt{chierchia_meaning_1990}).
-For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from its variants in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
-
-    % In formal semantics/pragmatics, the so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}).
-For instance, to diagnose the content of the clausal complement of \emph{discover} as projective content, a researcher would typically observe that the content that Julian dances salsa follows not just from \ref{ex:simple}, but also from the family-of-sentences variants of \ref{ex:simple} in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
-
-    % (e.g., \citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}).
+    Textbooks usually present the family-of-sentences diagnostic for projection without discussing the possibility of by-operator variation (e.g., \citealt{chierchia_meaning_1990}).
+For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, the epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.
 
@@
 
-    However, prior research is suggestive of variation in the effect of different entailment-cancelling operators on projection.
-For instance, \citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \textit{regret}) and semi-factives (e.g., \textit{discover}).
-    Based on the examples in \Next, he suggested that the CC of true factives projects across all four operators, but that of semi-factives does not always project from under polar questions, modals, or conditionals.\footnote{See \citealt{beaver_have_2010} for naturally occurring examples in which the CCs of factive predicates do not project from under the four operators, thereby calling into question the distinction between factive and semi-factive predicates.
-For experimental research on the distinction see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
-We return to this literature in section \ref{s:analysis}.}
+     However, prior research is suggestive that different entailment-cancelling operators may affect projection differently.
+For instance, \citet{karttunen_observations_1971} proposed distinguishing between English factive predicates (e.g., \textit{regret}) and semi-factives (e.g., \textit{discover}).
+Based on the examples in \Next, he suggested that the content of the complement (CC) of true factives consistently projects across the four operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%
+    %
+    \footnote{The factive/semi-factive distinction is called into question by reports of many naturally occurring examples where the CCs of factive predicates do not project from under the four operators (see \citealt{beaver_have_2010,de_marneffe_did_2012,demarneffe-etal-sub23}).
+For experimental research on the distinction, see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
+We return to this literature in Section \ref{s:analysis}.
+    %See \citealt{beaver_have_2010} for naturally occurring examples in which the CCs of factive predicates do not project from under the four operators, thereby calling into question the distinction between factive and semi-factive predicates.
+\mcdm{add CB too?
+We have quite a lot of naturally occurring examples for that too} For experimental research on the distinction see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
+We return to this literature in section \ref{s:analysis}.
+    }
 
@@
     There have been two experimental investigations of by-operator projection variation.
-First, the experiment reported in \citet{smith_relationship_2014} investigated projection from under negation and antecents of conditionals for various types of English projective contents.
+First, \citet{smith_relationship_2014} investigated projection from under negation and conditional antecedents for various types of English projective contents.
 They found that the expressive content of epithets (e.g., \textit{idiot}) and the CC of \textit{know} was more projective under negation than conditionals.
@@
 The results showed that the CCs of the predicates investigated projected more from under negation than the other three operators.
-contrary to what \citealt{karttunen_observations_1971} suggested, a comparison of factive and semifactive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of nonfactive predicates.
+Contrary to what \citet{karttunen_observations_1971} suggested, a comparison of the factive and
+    % \mcdm{need to make a pass to make sure semi-factive is written consistantly, above it is with hyphen}
+    semi-factive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of semi-factive predicates.
 
@@
 
-    This paper reports on the results of a set of experiments that were designed to compare projection from under the four entailment-canceling operators in \ref{ex:family}.
-Our experiments extend the empirical scope of prior research by investigating projection for a broader set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}:
+    This paper presents the results of a set of experiments designed to compare projection from under the four entailment-canceling operators in \ref{ex:family} in English.
+Our experiments extend the empirical scope of prior research on by-operator projection variation by investigating a larger set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}:
+
+    Our research expands on previous studies by exploring a broader range of content, specifically the complements of the 20 English clause-embedding predicates listed in (4).
 

### 2024-02-14 writing/SuB-paper/supplement.tex
updated analysis scripts

--- 2022-10-28 writing/investigating-GPP.tex
@@
 
-\title{Does at-issueness predict projection? Further investigations of the Gradient Projection Principle}
-
-%\thanks{For helpful comments on the research presented here, we thank David Beaver, Cleo Condoravdi, Kai von Fintel, Lauri Karttunen, Mandy Simons, Greg Scontras, the anonymous reviewers for {\em Semantics and Linguistic Theory} 2018, as well as the audiences at the MIT Linguistics colloquium, the 2018 Annual Meeting of XPRAG.de and at the University of T\"ubingen.
-We gratefully acknowledge financial support for this research from {\em National Science Foundation} grant BCS-1452674 (JT) and the Targeted Investment for Excellence Initiative at The Ohio State University (JT).
-IGOR Tuebingen, SALT TALK}}
-
-\author{Author(s)}
-
-%\author[$\circ$]{Judith Tonhauser}
-%\author[$\bullet$]{Judith Degen}
-%\affil[$\circ$]{The Ohio State University / University of Stuttgart}
-%\affil[$\bullet$]{Stanford University}
-%
-%\renewcommand\Authands{ and }
-
-\newcommand{\jt}[1]{\textbf{\color{blue}JT: #1}}
+\title{Is the family of sentences really a family?}
 
 \begin{document}
-
-%\tableofcontents
-%\newpage
-
-\maketitle
-
-\vspace*{-1cm}
-
-\begin{abstract}
-
-\citealt{tbd-variability} hypothesized that at-issueness is one factor that modulates the projection of content (see also \citealt{brst-salt10,brst-ar}).
-Specifically, according to their Gradient Projection Principle, if content $C$ is expressed by a constituent embedded under an entailment-canceling operator, then $C$ projects to the extent that it is not at-issue.
-Their work provided evidence for the GPP from xx contents that had been described as projective in the literature, at-issueness was measured in two different ways, only one consistently brought evidence in support (question, asking whether, are you sure?
-mixed results 2b).
-This paper provides further investigate the GPP by investigating a) more varied items, b) with more types of embedding, and c) other measures of at-issueness.
-These experiments provide additional support for the GPP, and they also show: comparison of projection across entailment-canceling operators, comparison of at-issueness diagnostics.
-Suggests that the diagnostics may not all measure the same (\citealt{snider2017}).
-
-\end{abstract}
-
-\newpage
-
-\tableofcontents
-
-\newpage
-
-\section{Introduction}\label{s1}
-
-\begin{exe}
-\ex Context: A and B live near a dog that regularly barks for extended periods of time. It is barking right now. A says:  \\ Ugh. The damn dog is barking again.
-\end{exe}
-
-Contents:
-
-\begin{itemize}
-\item there is a unique salient dog (backgrounded, ps)
-\item the dog has barked before (backgrounded, ps)
-\item the speaker is annoyed at the dog/barking (new, could be what the speaker wants to contribute)
-\item the dog is barking (``at-issue'', ordinary entailment)
-\end{itemize}
-
-First dimension: Role in discourse context: some anchor to context, some convey relevant backgrounded information, some are what the speaker wants to communicate.
-
-Second dimension: Compositional semantics: some interact, some don't.
-
-Third dimension: Role in subsequent discourse: some are what carries the conversation on, some do not.
-
-One of the contents is privileged in that it is the content that is put up for negotiation.
-It is often new information, though not necessarily.
-It is what is ``linguistically relevant'': it interacts with operators and is accessible for anaphoric reference.
-
-\begin{exe}
-\ex\label{gpp} {\bf Gradient Projection Principle} \hfill (\citealt[400]{tbd-variability}) \\ If content $C$ is expressed by a constituent embedded under an entailment-canceling operator, then $C$ projects to the extent that it is not at-issue.
-\end{exe}
-
-We already know that the GPP is stated too strongly: prior influences projection (\citealt{degen-tonhauser-openmind}) {\bf independently of at-issueness (unpublished)}.
-At-issueness is one factor among many.
-
-\begin{exe}
-\ex\label{rqs} {\bf Main research question} \\
-Are not-at-issueness and projection positively correlated, as predicted by the Gradient Projection Principle?
-\end{exe}
-
-\begin{exe}
-\ex\label{rqs} {\bf Ancillary research questions}
-\begin{xlist}
-
-\ex Is the at-issueness of contents embedded under different entailment-canceling operators positively correlated?
-
-\ex Is the at-issueness of contents positively correlated under different measures of at-issueness?
-\end{xlist}
-\end{exe}
-
-Previous investigations of/claims about the relationship between projection and at-issueness:
-
-\begin{itemize}
-
-\item \citealt{brst-salt10,brst-ar}
-
-\item \citealt{xue-onea11}
-
-\item \citealt{tbd-variability}
-
-\item \citealt{koev2018}
-
-\end{itemize}
-
-Limited how?
-(only presuppositions + CCs of factive predicates: \citealt{degen-tonhauser-factive} / few measures of at-issueness / some supporting results, some not / some embedded, other matrix clause)
-
-\begin{exe}
-\ex\label{pred} 20 clause-embedding predicates
-
-\begin{xlist}
-
-\ex canonically factive: {\em be annoyed, discover, know, reveal, see}
-
-\ex nonfactive:
-
-\begin{xlist}
-
-\ex nonveridical nonfactive: {\em pretend, suggest, say, think}
-
-\ex veridical nonfactive: {\em be right, demonstrate}
-
-\end{xlist}
-
-\ex optionally factive: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove}
-
-\end{xlist}
-
-\end{exe}
-
-\subsection{At-issueness diagnostics}\label{s:ai-diagnostics}
-
-In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q, illustrated in (\ref{ai-exps1}a), and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, illustrated in (\ref{ai-exps1}b).
-The `asking whether' diagnostic requires that the content to be diagnosed is realized by an expression in a polar question; this diagnostic was also used in \citealt{tbd-variability} Exps.~1.
-A speaker (Christopher in (\ref{ai-exps1}a) asks a polar question and the participant is asked whether the speaker is asking about the content to be diagnosed (here, that Julian dances salsa).
-If a participant takes the speaker to be asking of the content, the content is assumed to be at-issue; if a participant does not take the speaker to be asking about the content, the content is assumed to be not-at-issue.
-The `sure that' diagnostic requires that the content to be diagnosed is realized by an expression in declarative sentence; this diagnostic was also used in \citealt{tbd-variability} Exps.~2.
-A speaker utters the declarative sentence and the addressee (here, Sandy) asks {\em Are you sure?}.
-The speaker response with {\em Yes, I am sure that} followed by the expression that denotes the content to be diagnosed.
-Participants are asked whether the speaker answered the addressee's question.
-If a participant takes the speaker to have answered the question, the content is assumed to be at-issue; if a participant does not take the speaker to have answered the question, the content is assumed to be not-at-issue.
-{\bf move all this into section 1}
-
-\begin{exe}
-\ex\label{ai-exps1} At-issueness measures in Exps.~1
-\begin{xlist}
-\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
-\\ Question to participants: Is Christopher asking whether Julian dances salsa?
-
-\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
-\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
-\\ Sandy: Are you sure?
-\\ Christopher: Yes, I'm sure that Julian dances salsa.
-\\ Question to participants: Did Christopher answer Sandy's question?
-
-\end{xlist}
-\end{exe}
-
-In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed.
-Participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
-The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps2}a)) or assents (for declaratives, as shown in (\ref{ai-exps2}b).
-The affirmation/assent is followed by the content to be diagnosed, here that Julian dances salsa.
-
-\begin{exe}
-\ex\label{ai-exps2} At-issueness measures in Exps.~2
-\begin{xlist}
-\ex {\em yes}, $c$
-\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
-\\ Sandy: Yes, Julian dances salsa.
-\\ Question to participants: Does Sandy's response to Christoper sound good?
-
-\ex {\em yes, that's true}, $c$
-\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
-\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
-\\ Sandy: Yes, that's true, Julian dances salsa.
-\\ Question to participants: Does Sandy's response to Christopher sound good?
-
-\end{xlist}
-\end{exe}
-Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
-If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
-If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
-
-In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content.
-As in the other assent diagnostic, participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
-The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps3}a)) or assents (for declaratives, as shown in (\ref{ai-exps3}b).
-The affirmation/assent is followed by the denial of another salient content, here, that Cole (didn't) discover it.
-
-\begin{exe}
-\ex\label{ai-exps3} At-issueness measures in Exps.~3
-\begin{xlist}
-\ex {\em yes}, $c$
-\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
-\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
-\\ Question to participants: Does Sandy's response to Christoper sound good?
-
-\ex {\em yes, that's true}, $c$
-\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
-\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
-\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
-\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
-\\ Question to participants: Does Sandy's response to Christopher sound good?
-
-\end{xlist}
-\end{exe}
-As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
-If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
-If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
-
-\subsection{Formal analyses of at-issueness}
-
-- Esipova, Maria. 2018. QUD-addressing appositives don?t have to be clause-final. Snippets 33. 7?8. https://doi.org/10.7358/snip-2018-033-esip.
-
-- Dj\"arv, Kajsa. 2019. Factive and assertive attitude reports. Philadelphia, PA: Uni- versity of Pennsylvania dissertation. https://repository.upenn.edu/edissertations/ 3645.
-
-backgrounded content
-
-\begin{exe}
-\ex Formal analyses of at-issueness
-
-\begin{xlist}
-\ex Dimension-based \\ At-issue content interacts with operators
-
-\ex Question-based \\ At-issue content addresses the question under discussion
-
-\ex Anaphora-based \\ At-issue content introduces a discourse referent for subsequent reference
-
-\end{xlist}
-\end{exe}
-
-\begin{itemize}
-
-\item \citealt{potts05}
-
-\begin{itemize}
-
-\item p.23/Figure 2.1: at-issue entailments vs.
-backgrounded content (conventional presuppositions, CIs, conversationally-triggered presuppositions) vs.
-conversational implicatures (not backgrounded!!) {\bf 3-way distinction}
-
-\item p.24: But the supplementary relative who lived in a working-class suburb of Boston plays a secondary role relative to the information conveyed by the main clause.
-The issue is not where the grandmother lived, but rather the fact that the speaker summered with her as a child.
-
-\item p.30: the utterance might be followed in the discourse by ?Hey, she?ll do both!?
-Or it might be preceded by an agreement that if Mary does one, then she does the other.
-The maxims of quality and quantity would then conspire to en- sure that (2.35b) disappeared.
-Thus, the conversational-implicature dimension is a negotiable part of denotations.
-
-\item p.31: a CI is never relativized to the beliefs of an entity other than the speaker.
-But at-issue content certainly is; in Sue wrongly believes Conner got promoted, the at-issue proposition that Conner got promoted is asserted to hold only in Sue?s belief worlds.
-this embedded proposition is not speaker-oriented, and hence not classifiable as a CI contri- bution,
-
-\item p.32:  CIs are independent of the at-issue con- tent.
-In contrast, the fundamental goal of almost all presupposition logics is to create a dependency between the presuppositions and the at-issue entailments
-
-\item p.34: Neither at- issue content nor CI content should be presupposed.
-
-\item p.43: everything about the definition says that CIs contribute new information. In this sense, they are like at-issue content.
-
-\item {\bf can we diagnose not-at-issue content independent of a particular type of not-at-issue content?
-that is, is there a diagnostic for not-at-issueness that applies to all such content, regardless of its specific type?}
-
-\item p.51: discourse-level phenomena. It seems that none are sensitive to the distinction between these two types of content. (at-issue vs CI): VP ellipsis and do it/that pronominalization
-
-\item p.53: Overall, they are an indication that we distinguish the at-issue and CI dimensions in the composition. From the point of view of a discourse as a whole, they are identical.
-
-\item p.54: A similar, and less fraught, test involves discourse referents.
-Because noth- ing can scope over supplementary CI content, it is quite good at establishing discourse referents.
-...
-Here again, presupposed material patterns differently.
-It alone cannot establish a discourse referent, at least not without a significant amount of contextual priming and some creative inferences on the part of the hearer;
-
-\item p.54: Speaker-oriented adverbs, for instance, suggest a new test. Exam- ples (3.16)?(3.17) show that their content can be queried just as at-issue content can.
-
-\item p.54: The upshot of these examples is that CI content is model-theoretically the same as at-issue content.
-...
-Thus, we are led to the con- clusion that the distinction exists only at the level of the meaning language.
-{\bf potts assumes that ai and nai content are not distinct on a discourse level!}
-
-\item p.70: in the CI logic employed here, translation is a nontrivial step, as it is the locus of the at-issue/CI distinction.
-If we interpreted natural language expres- sions directly, we would be left with only two options, both undesirable: we could locate the distinction in the natural language syntax or in the models.
-The syntactic view is initially unpromising given the semantic nature of the defini- tion of CIs, and it receives sustained criticism in chapter 6.
-The model-theoretic option is to locate the differences in the models.
-But section 3.3 shows that in- tersentential phenomena treat the two classes of meaning as model-theoretically identical.
-Moreover, it is hard to see what the distinctions could be.
-
-\end{itemize}
-
-\item \citealt{potts2007-compass}
-
-\begin{itemize}
-
-\item p.666: ``typically the content that speakers offer as primary and also the content that they are most expecting to have to negotiate with their interlocutors before it is accepted into the common ground''
-
-\end{itemize}
-
-\item \citealt{koev2013,koev2018,koev2021}
-
-\begin{itemize}
-
-\item \citealt{koev2018}
-
-\begin{itemize}
-
-\item p.1: At-issue content is that part of the sentence meaning that is intuitively felt to express the ?main point?
-when the sentence is uttered in a given context.
-Potts (2005; 2007) was among the first to use the term ?at-issue content?
-in the relevant sense, characterizing it as content that ?carr[ies] the main themes of a discourse?
-(Potts, 2005, 7), and that speakers are ?most expect- ing to have to negotiate with their interlocutors before it is accepted into the common ground?
-(Potts, 2007, 666).
-The first property, i.e.
-the ability to contribute to the discourse topic, is often expressed by saying that at-issue content is what is relevant to the question under discussion.
-The second property, i.e.
-the ability to represent what is being negotiated among interlocu- tors, is typically captured by saying that at-issue content constitutes a proposal to update the common ground.
-
-\item p.2: There exist several empirical tests for at-issueness, e.g.
-question/answer pairs and accessibility to direct denials or other discourse continuations.
-Exist- ing accounts have employed such tests rather indiscriminately or without elaborating on how these tests relate to a given theoretical construal.
-In order to be able to judge a given theory on its own merits, in each case I will consider only those diagnostics that transparently follow from what counts as at issue in the relevant sense.
-
-\item p.3: q-at-issueness:  A proposition p is Q-atissue relative to a question under discussion Q and a context c iff ? p is relevant to Q in c and
-? p is appropriately conventionally marked relative to Q.
-
-\item p.4: Above, Q-at-issueness was diagnosed in question/answer pairs.
-This is consistent with the question-based definition in 2, provided that an overt question is assumed to manifest the current question under discussion.
-
-\item p.5: several authors identify at-issue content with the proposal introduced by a declarative utterance, i.e.
-with its asserted content (AnderBois et al., 2015; Farkas \& Bruce, 2010; Koev, 2013; Murray, 2014; Potts, 2005).
-{\bf potts 2005 does not assume that CIs are not asserted!}
-
-\item p.5: A proposition p is P-at-issue in a context c iff ? p is a proposal in c and ? p has not been accepted or rejected in c.
-
-\item p.5: to test a proposition for proposalhood, we need a lin- guistic tool that tracks what is being negotiated at any given point.
-One widely used diagnostic to this effect is the assent/dissent test, which probes the amenability of semantic content to direct responses
-
-\item p.6: This test brings up the question of what counts as a direct response.
-One peculiarity about the responses in 10?11 is that they are anaphoric to parts of the target utterance through propositional anaphors like that or no.7 This is important because nonanaphoric denials behave differently: They can felicitously target nominal appositives (see also Hunter \& Asher, 2016).
-
-\item p.6: One possible reaction to the contrast in 10?11 v.
-12?13 is to say that the assent/dissent test tracks anaphoric potential rather than P-at-issueness.
-This is reasonable, although it has also been claimed that P-at-issue content obligatorily establishes a propositional discourse referent (Murray, 2014).
-
-\item p.6: I will thus follow the bulk of the literature in assuming that such responses target the assertion and will employ them as the main diagnostic for P-at-issueness, leaving the precise link between anaphoricity and at-issueness to further research.
-
-\item p.7: A third notion of at-issueness is situated within theories which view discourse as structured by coherence relations, also called ?rhetorical?
-or ?discourse?
-relations (Asher \& Lascarides, 2003; Hobbs, 1979, 1985; Kehler, 2002; Mann \& Thompson, 1988).
-
-\item p.8:  A proposition expressed by a node ?? of a discourse tree D is C-atissue iff ?? is on the right frontier of D.
-
-\item p.8: A proposition is C-atissue iff a freshly uttered segment can attach to it by some appropriate coherence relation.
-
-\item p.9: Coherence-based accounts thus view the assent/dissent test as a special case of discourse attachment, i.e.
-one in which the response is connected to the target by an Agreement/Denial relation.
-
-\item p.9: Q-at-issueness is primarily a backward-looking notion or relates to existing discourse. By contrast, P-at-issueness and C-at-issueness are forward-looking or relate to what is to come.
-
-\item p.10: I first compare Q-at-issueness and P-at-issueness, arguing that neither property entails the other.
-
-\item p.10: Koev argues that 26 shows that the appositive is not P-at-issue but that 27 shows that it is Q-at-issue: ``appositive relative clauses are often taken into consideration when answering the question under discussion.
-Here, the crucial information that Mary experienced firsthand the terrible consequences of cancer comes from the appositive; the main clause alone does not answer the question, at least not without loss of information.''
-
-(26) A: Mary took care of her husband, who had prostate cancer, for almost a year. B: ? No, he didn't.
-
-(27) Q: Why is Mary fundraising for the upcoming Walk For Cancer?
-A: She took care of her husband, who had prostate cancer, for almost a year.
-
-\item p.10: Relevance implicatures. They are not P-at-issue (as diagnosed by dissent) but Q-at-issue (as diagnosed by Q/A pairs).
-
-\item p.10f: slifting constructions. Can be P-at-issue (dissented with) but don't answer Q (diagnosed by Q/A pairs)
-
-\item p.11:  Simons et al. (2010) and Beaver et al. (2017) claim that projectivity is perfectly correlated with (the lack of) Q-at-issueness {\bf NOT TRUE}.
-
-\item p.12: The Projection Principle then appears to be problematic in both directions: Projecting content can be Q-at issue, and nonprojecting content can be not Q-at issue.
-{\bf Need to engage with Koev 2018 as somebody who doesn't believe in the GPP}
-
-\item p.12: The second observation is that the Projection Principle is stated in terms of Q-at-issueness {\bf NO!} and other notions of at-issueness may lead to different predictions.
-The question of which notion of at-issueness (if any) makes the right predictions about projection is yet to be systematically investigated.
-
-\item endnote 1: I also limit my attention to declarative sentences. The question of how or whether at-issueness interacts with other sentence types will have to await another study.
-
-\item endnote 5: The assent/dissent test is known under different names in the literature and has been used to distinguish asserted content from implications triggered by constructions as different as presupposition markers (Chierchia \& McConnell-Ginet, 2000; Karttunen \& Peters, 1979; Shanon, 1976; Strawson, 1950), epistemic modals (Lyons, 1977; Papafragou, 2006; von Fintel \& Gillies, 2007), evidentials (Faller, 2002; Matthewson et al., 2007; Murray, 2014), and appositives (AnderBois et al., 2015; Koev, 2013; Syrett \& Koev, 2015; Tonhauser, 2012).
-
-\item endnote 16: Appositive content typically projects, so it could be argued that it is not ?intended? as being at issue (see Simons et al., 2010).
-However, given its significance in the above context and the fact that conventional marking of at-issueness is often overruled, the information that the director fired Jon would very likely be taken into consideration when answering the question under discussion.
-
-\end{itemize}
-
-\end{itemize}
-
-\item \citealt{amaral-etal2007}
-
-\item \citealt{gutzmann}
-
-\item \citealt{korotkova2020}
-
-\item \citealt{murray2014}
-
-\item \citealt{anderbois-etal2015}
-
-\item \citealt{snider-thesis}
-
-\item \citealt{snider2017} (superseded for the most part by 2018 paper)
-
-\begin{itemize}
-
-\item In this paper, I argue that there is no tight link between at-issueness and anaphoric availability?
-that these two properties of propositions are distinct.
-I present data which demonstrates that a proposition?s at-issue status alone is not sufficient to determine whether it will be available for anaphoric reference (including but not limited to direct assent/dissent), as well as data which shows that a proposition?s being available for anaphoric reference cannot be used to diagnose its at-issue status.
-
-\item  Tonhauser asserts that they are all indeed testing for at-issueness: ?they are all useful to diagnose (not-)at-issueness with at least one kind of content? (p. 252). {\bf I was wrong}
-
-\item p.6: The question/answer tests return the same results, because their use of anaphors (response particles) is only incidental.
-In contrast, the direct dissent tests rely on explicit anaphors by design; an anaphor-less equivalent to the direct dissent test in (11) ceases to be the same diagnostic.
-
-\end{itemize}
-
-\item \citealt{snider2018}
-
-\begin{itemize}
-
-\item p.374: at-issueness and the anaphoric potential of a proposition are distinct notions; one class of tests commonly used to diagnose at-issue status in fact diagnoses only anaphoric potential
-
-\item NRRCs: anaphoric potential varies, especially with position, but never good to address question under discussion
-
-\item at-issue content: certain content conveyed by an utterance has a distinguished status as the main point of the utterance, while other content does not.
-Potts 2005 discusses ?at-issue entailments?
-as distinct from presuppositions and conventional implicatures, but does not describe how to identify content as at-issue.
-
-\item p.375: Properties of at-issue content (from \citealt{tonhauser-sula6})
-
-At-issue content can be directly assented or dissented with {\bf third dimension}
-
-At-issue content addresses the question under discussion {\bf first dimension}
-
-At-issue content determines the relevant set of alternatives {\bf second dimension, compositionality}
-
-Some content may overlap on all three dimensions: e.g., that Fritz is happy in {\em Fritz is happy}.
-But sentence-final appositives can be assented/dissented with, but they do not address the QUD or determine the relevant set of alternatives.
-And that the speaker is annoyed may address the QUD, but does not determine the relevant set of alternatives.
-
-\item p.376f: Snider argues against \citealt{syrett-koev2015} claim that appositives can address the QUD.
-``Why questions are crucially different from these other questions in that they seek explanations.
-Because they seek explanations, such questions have no single comprehensive answer: any answer given is defeasible2 or can be considered insufficiently precise, and additional contextual information can always contribute to an explanation (see discussion in Bjorndahl \& Snider, 2016).
-The appositive might contribute information which bolsters an explanation, supporting the matrix clause?s answer to the QUD, but that alone does not constitute addressing the QUD.
-If the matrix clause does not address the QUD, then an appositive alone cannot be taken to address the QUD, even in a why question, as in (7)''
-
-Syrett \& Koev 2015:587 further points to an example where an appositive appears to ?address one part of a coordinated QUD?, inspired by an example from Koev 2013.
-But the concept of a ?coordinated QUD?
-is novel: Roberts 1996 describes no such structure, instead describing sequences of ?relevant sub-questions to some super-question?
-as an ?enumeration, suggesting a plan for how to attack the super-question?, and says that one ?can only address such sub-questions one at a time?
-(5).
-The response in (8A) thus addresses the first sub-question?as it must, because the second requires an answer to the first?and not both at once.
-
-\item p.378: The crucial difference between the classes of diagnostics illustrated above is that the first is anaphora-based, where the other two are QUD-based, as they test what can address or establish a Question Under Discussion.
-
-Snider assumes that the two QUD diagnostics (addressing QUD, establishing QUD) are QUD-based, i.e., tap into the same property?
-But establishing a QUD is about compositionality, and addressing a QUD is about the role of content in discourse.
-
-\item p.380: Alternatively, this observation might lead one to move away from the Simons et al.
-2010 definition entirely.
-Hunter \& Asher 2016 proposes a discourse based account using SDRT (Asher, 1993; Asher \& Lascarides, 2003) in which at-issueness and anaphoric potential are indeed closely linked?which isn?t at odds with the data discussed here, because they use an entirely different notion of at-issueness.
-On this account, ?content is at-issue at a certain point of the discourse just in case it is on the [Right Frontier (RF)] of the discourse at that time.
-Content that is AI becomes NAI when knocked off the RF; content that is NAI can become AI if targeted through discourse subordination?
-and thus returned to the Right Frontier (Hunter \& Asher, 2016:1036).
-And, because discourse referents (propositional and otherwise) are only accessible for anaphora resolution in SDRT if they are on the Right Frontier, propositional content which is available for anaphoric reference is necessarily at-issue on this account.
-
-\item p.380: However one chooses to define or model at-issueness, it is important to recognize the difference between the at-issue status of a proposition and its anaphoric potential.
-The two are at least conceptually distinct: at-issueness is about what the main point of an utterance is, while anaphoric potential is about what propositions are available for subsequent anaphoric reference.
-
-\end{itemize}
-
-\item \citealt{syrett-koev2015}
-
-\item \citealt{tonhauser-sula6}
-
-\item \citealt{brst-salt10}
-
-\item \citealt{tbd-variability}
-
-\item \citealt{brst-ar}
-
-\end{itemize}
-
-\subsection{Diagnosing at-issueness}
-
-\begin{itemize}
-
-\item Diagnostics for at-issueness are either suitable for polar questions or for declaratives.
-
-\begin{itemize}
-
-\item Polar questions: at-issue content partitions the context set
-
-`asking whether'
-
-`yes, $c$'
-
-`yes, but not-c''
-
-\item Declaratives: at-issue content is the target of anaphora
-
-`sure that'
-
-`yes, that's true, $c$'
-
-`yes, that's true, but not-c''
-
-\end{itemize}
-
-\end{itemize}
-
-\subsection{Next steps}
-
-projection of contents embedded under different entailment-canceling operators is positively correlated, as assumed in the semantics/pragmatics literature (e.g., REFERENCES).
-Prior experimental literature: \citealt{smith-hall11,smith-hall-cls,demarneffe-etal-sub23}.
-
-\newpage
-
-\section{Experiments 1-3: Methods}\label{s2}
-
-Exps.~1-3 were designed to investigate the three research questions in (\ref{rqs}), repeated here for convenience:\footnote{\label{f-github}The experiments, data, and R code for generating the figures and analyses of the experiments reported in this paper are available at \url{https://github.com/judith-tonhauser/projective-probability}.
-All experiments were conducted with IRB approval.
-Exps.~1b, 2b, and 3b were preregistered: \url{https://osf.io/cxq47}.}
-
-\begin{exe}
-\ex\label{rqs} Research questions
-\begin{xlist}
-\ex Does the projection of content differ across entailment-canceling environments?
-\ex Are the results of different measures of at-issueness positively correlated?
-\ex Is the projection of content positively correlated with its not-at-issueness, as predicted by \citepos{tbd-variability} Gradient Projection Principle?
-\end{xlist}
-\end{exe}
-To address these questions, Exps.~1-3 investigated the projection and at-issueness of same 20 contents.
-Each of the experiments consisted of two blocks: a projection block, in which the projection of the contents was measured, and an at-issueness block, in which the at-issueness of the contents was measured.
-To address the research question in (\ref{rqs}b), different measures of at-issueness were implemented in the at-issueness blocks of Exps.~1-3, respectively.
-To address the research question in (\ref{rqs}a), each of Exps.~1-3 consisted of four experiments (for a total of 12 experiments) where the contents were realized by expressions in a polar question, in the scope of sentential negation, in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.
-The research question (\ref{rqs}c) is addressed by comparing the projection and at-issueness ratings of each of the 12 experiments.
-For an overview of the 12 experiments see Table \ref{t:overview}, where the name of each experiment identifies whether belongs to the set of Exps.~1, 2, or 3 (which differ in at-issueness measures) and whether the embedding-environment is a polar question (`q'), negation (`n'), the epistemic possibility modal (`m'), or the antecedent of a conditional (`c').
-
-Projection was measured across all 12 experiments with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser-salt26,djaerv-bacovcin-salt27,stevens-etal2017,lorson2018,tbd-variability,mahler-nels,mahler2020,demarneffe-etal-sub23}).\footnote{For other diagnostics for projection see, e.g.\ \citealt{smith-hall11,xue-onea11}, and \citealt{brst-lang11}; see also the discussion in \citealt{tbd-variability}.}  On this diagnostic, participants are presented with an utterance, like one of Daniel's utterances in (\ref{proj-stim}), in which the content to be diagnosed for projection (here, that Julian dances salsa) is realized by an expression that occurs in one of the four entailment-canceling environments, that is, in a polar question (\ref{proj-stim}a), in the scope of negation (\ref{proj-stim}b), in the scope of the epistemic possibility modal {\em perhaps}, or in the antecedent of a conditional.
-
-\begin{exe}
-\ex\label{proj-stim} `certain that' diagnostic for projection
-\begin{xlist}
-\ex Daniel: Did Cole discover that Julian dances salsa? \hfill [Exps.~1q, 2q, 3q]
-\ex Daniel: Cole didn't discover that Julian dances salsa. \hfill [Exps.~1n, 2n, 3n]
-\ex Daniel: Perhaps Cole discovered that Julian dances salsa. \hfill [Exps.~1m, 2m, 3m]
-\ex Daniel: If Cole discovered that Julian dances salsa, Logan will be joyful. \hfill [Exps.~1c, 2c, 3c]
-\end{xlist}
-Question to participants: Is Daniel certain that Julian dances salsa?
-\end{exe}
-To measure projection, participants are asked whether the speaker is certain of the relevant content.
-In \ref{proj-stim}, for instance, participants are asked whether Daniel is certain that Julian dances salsa.
-If a participant takes the speaker to be certain of the content, the content is assumed to project out of the entailment-canceling environment; if a participant does not take the speaker to be certain of the content, the content is assumed to not project.
-
-As shown in Table \ref{t:overview}, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q (as this diagnostic can only applied to content in polar questions) and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c (as this diagnostic can only be applied to content in declarative sentences).
-As discussed in section \ref{s:ai-diagnostics}, these two measures of at-issueness differ not just with respect to the required sentence type but also in whether it is a question- or anaphora-based diagnostic.
-In Exps.~2, at-issueness was measured a version of the `affirmation diagnostic' in Exp.~2q and by a version of the `assent diagnostic' in Exps.~2n, 2m, and 2c.
-Specifically, it was measured by versions of these diagnostics in which the affirmation ({\em yes}) or assent ({\em yes, that's true}) is followed by the content to be diagnosed.
-In Exps.~3, at-issueness was measured by a different version of these diagnostics, namely by the version in which the affirmation or assent is followed by a denial of another salient content.
-
-%
-%In Exps.~1, at-issueness was measured with the `asking whether' diagnostic in Exp.~1q, illustrated in (\ref{ai-exps1}a), and with the `sure that' diagnostic in Exps.~1n, 1m, and 1c, illustrated in (\ref{ai-exps1}b).
-The `asking whether' diagnostic requires that the content to be diagnosed is realized by an expression in a polar question; this diagnostic was also used in \citealt{tbd-variability} Exps.~1.
-A speaker (Christopher in (\ref{ai-exps1}a) asks a polar question and the participant is asked whether the speaker is asking about the content to be diagnosed (here, that Julian dances salsa).
-If a participant takes the speaker to be asking of the content, the content is assumed to be at-issue; if a participant does not take the speaker to be asking about the content, the content is assumed to be not-at-issue.
-The `sure that' diagnostic requires that the content to be diagnosed is realized by an expression in declarative sentence; this diagnostic was also used in \citealt{tbd-variability} Exps.~2.
-A speaker utters the declarative sentence and the addressee (here, Sandy) asks {\em Are you sure?}.
-The speaker response with {\em Yes, I am sure that} followed by the expression that denotes the content to be diagnosed.
-Participants are asked whether the speaker answered the addressee's question.
-If a participant takes the speaker to have answered the question, the content is assumed to be at-issue; if a participant does not take the speaker to have answered the question, the content is assumed to be not-at-issue.
-{\bf move all this into section 1}
-%
-%\begin{exe}
-%\ex\label{ai-exps1} At-issueness measures in Exps.~1
-%\begin{xlist}
-%\ex `asking whether' \\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~1q]
-%\\ Question to participants: Is Christopher asking whether Julian dances salsa?
-%
-%\ex `sure that' \\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~1n]
-%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~1m]
-%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~1c]
-%\\ Sandy: Are you sure?
-%\\ Christopher: Yes, I'm sure that Julian dances salsa.
-%\\ Question to participants: Did Christopher answer Sandy's question?
-%
-%\end{xlist}
-%\end{exe}
-%
-%In Exps.~2, at-issueness was measured with an assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the content to be diagnosed.
-Participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
-The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps2}a)) or assents (for declaratives, as shown in (\ref{ai-exps2}b).
-The affirmation/assent is followed by the content to be diagnosed, here that Julian dances salsa.
-%
-%\begin{exe}
-%\ex\label{ai-exps2} At-issueness measures in Exps.~2
-%\begin{xlist}
-%\ex {\em yes}, $c$
-%\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~2q]
-%\\ Sandy: Yes, Julian dances salsa.
-%\\ Question to participants: Does Sandy's response to Christoper sound good?
-%
-%\ex {\em yes, that's true}, $c$
-%\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~2n]
-%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~2m]
-%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~2c]
-%\\ Sandy: Yes, that's true, Julian dances salsa.
-%\\ Question to participants: Does Sandy's response to Christopher sound good?
-%
-%\end{xlist}
-%\end{exe}
-%Participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
-If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
-If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
-%
-%In Exps.~3, at-issueness was measured with another assent diagnostic, specifically, with the version of the assent diagnostic in which the assent is followed with the denial of another, salient content.
-As in the other assent diagnostic, participants are presented with a dialogue in which the first speaker's utterance gives rise to the content to be diagnosed (here, that Julian dances salsa).
-The second speaker responds in the affirmative (for polar interrogatives, as shown in (\ref{ai-exps3}a)) or assents (for declaratives, as shown in (\ref{ai-exps3}b).
-The affirmation/assent is followed by the denial of another salient content, here, that Cole (didn't) discover it.
-%
-%\begin{exe}
-%\ex\label{ai-exps3} At-issueness measures in Exps.~3
-%\begin{xlist}
-%\ex {\em yes}, $c$
-%\\ Christopher: Did Cole discover that Julian dances salsa? \hfill [Exp.~3q]
-%\\ Sandy: Yes, but Cole didn't discover it. {\bf or that?}
-%\\ Question to participants: Does Sandy's response to Christoper sound good?
-%
-%\ex {\em yes, that's true}, $c$
-%\\ Christopher: Cole didn't discover that Julian dances salsa.  \hfill [Exp.~3n]
-%\\ \hspace*{2cm} Perhaps Cole discovered that Julian dances salsa.  \hfill [Exp.~3m]
-%\\ \hspace*{2cm} If Cole discovered that Julian dances salsa, Logan will be joyful.  \hfill [Exp.~3c]
-%\\ Sandy: Yes, that's true, but Cole discovered / didn't discover it. {\bf or that?}
-%\\ Question to participants: Does Sandy's response to Christopher sound good?
-%
-%\end{xlist}
-%\end{exe}
-%As with the other assent diagnostic, participants are asked whether the second speaker's response is acceptable in response to the first speaker's utterance.
-If a participant judges the response to be acceptable, the relevant content is taken to be at-issue.
-If a participant judges the response to be unacceptable, the relevant content is taken to be not at-issue.
-
-\begin{table}[h!]
-\centering
-\begin{tabular}{r | r | r | r}
-
-{\bf Exp.} & {\bf Embedding} & {\bf Projection measure} & {\bf At-issueness measure} \\ 
-\hline
-1q & question & certain that $c$ & asking whether $c$ \\
-1n & negation & certain that $c$ & sure that $c$ \\
-1m & modal & certain that $c$ & sure that $c$ \\
-1c & conditional & certain that $c$ & sure that $c$ \\
-\hline
-2q & question & certain that $c$ & {\em yes}, $c$ \\ 
-2n & negation & certain that $c$& {\em yes, that's true}, $c$ \\ 
-2m & modal & certain that $c$& {\em yes, that's true}, $c$ \\ 
-2c & conditional & certain that $c$& {\em yes, that's true}, $c$ \\ 
-\hline
-3q & question & certain that $c$& {\em yes}, but $\neg c'$ \\ 
-3n & negation & certain that $c$& {\em yes, that's true}, but $\neg c'$ \\ 
-3m & modal & certain that $c$& {\em yes, that's true}, but $\neg c'$ \\ 
-3c & conditional & certain that $c$& {\em yes, that's true}, but $\neg c'$ \\ 
-\hline
-\end{tabular}
-\caption{Overview of Exps.~1-3, which differ in the entailment-canceling embedding (q = polar question, n = negation, m = epistemic possibility modal, c = antecedent of conditional) and measures for not-at-issueness.
-$c$ is the content under investigation (in the target trials: the content of the clausal complement) and $c'$ is a different, salient content (in the target trials: the main clause content).}\label{t:overview}
-\end{table}
-
-\subsection{Participants}
-
-We recruited 250-300 participants for each of Exps.~1-3.
-Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs.
-Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
-Information on the participants and the exclusion criteria is provided in Supplement \ref{a-participants}.
-
-\subsection{Materials and procedure}
-
-The projection and at-issueness of the contents of 20 clauses (such as {\em Julian dances salsa}) were measured in separate blocks in Exps.~1-3.
-For the full set of clauses, see Supplement \ref{a-clauses}.
-
-In the projection blocks of Exps.~1-3, target stimuli consisted of a polar question sentence, a negated sentence, a sentence with {\em perhaps}, or a conditional sentence that was uttered by a named speaker, as shown in Fig.~\ref{fig:projection-trials}.
-Each sentence was formed by realizing the 20 clauses as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates}.\footnote{The proper names that realized the speakers, the subjects of the clause-embedding predicates, the subjects of the complement clauses, and the subjects of the consequents  were all unique.
-The indirect object of {\em inform} was realized by the proper name {\em Sam}.
-The predicates differed in the tense in which they were realized.
-In polar interrogatives as well as in negated and modalized declaratives, eventive predicates, like {\em discover} and {\em hear}, were realized in the past tense and stative predicates, like {\em know} and {\em be annoyed}, were realized in the present tense.
-In the antecedents of conditionals, all predicates were realized in the present tense.
-For the consequents of the conditional sentences see Supplement \ref{a-target}.} Projection was measured using the `certain that' diagnostic \cite{djaerv-bacovcin-salt27,tbd-variability, lorson2018, mahler2020}: participants were asked to rate whether the speaker was certain of the CC, taking into consideration the fact.
-They gave their responses on a slider marked `no' at one end (coded as 0) and `yes' at the other (coded as 1).
-Greater speaker commitment to the CC should result in higher slider ratings.
-
-%Following \citealt{tbd-variability}, the `certain that' diagnostic was implemented in Exps.~1-3 with a gradient response scale: participants gave their responses on a slider marked `no' at one end and `yes' at the other.
-We assume that the closer to `yes' a participant's response is, the more the speaker is certain of the content, that is, the more projective the content is.
-In other words, we assume that gradient certainty ratings reflect gradience in the speaker's commitment to the truth of the content.\footnote{Strictly speaking,  gradient certainty ratings reflect gradience in the degree to which participants \emph{perceive} the speaker to be committed to the truth of the content.
-That is, as in any experiment, the quantity of interest, in this case speaker commitment, is only indirectly measured.} As discussed in \citealt{tbd-variability}, a second interpretation of gradient certainty ratings is that they reflect a participant's degree of belief in the speaker's (binary) commitment to the truth of the content.
-While we adopt the first interpretation in our discussion, we remain agnostic about the interpretation of gradient certainty ratings.
-
-\begin{figure}[h!]
-\centering
-\fbox{\begin{minipage}{15cm}{\em acknowledge, admit, announce, be annoyed, be right, confess, confirm, demonstrate, discover, establish, hear, inform, know, pretend, prove, reveal, say, see, suggest, think.
-}\end{minipage}}
-\caption{20 clause-embedding predicates in Exps.~1-3.}\label{fig:predicates}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/2q-proj}}
-\caption{Exps.~1q, 2q, and 3q.}\label{fig-exp1q-projection}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/1n-proj}} 
-\caption{Exps.~1n, 2n, and 3n.}\label{fig-exp1n-projection}
- \end{subfigure}
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/1m-proj}}
-\caption{Exps.~1m, 2m, and 3m}\label{fig-exp1m-projection}
- \end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=4.4cm,width=7.6cm]{figures/1c-proj}} 
-\caption{Exps.~1c, 2c, and 3c}\label{fig-exp1c-projection}
-\end{subfigure}
-\caption{Target trials in projection blocks of Exps.~1, 2, and 3 for the complement {\em Julian dances salsa}.}\label{fig:projection-trials}
-\end{figure}
-
-In the at-issueness blocks of Exps.~1-3, the 20 clauses were also realized as the complements of the 20 clause-embedding predicates in Fig.~\ref{fig:predicates} and the target stimuli realized either a polar question sentence (in Exps.~1q, 2q, and 3q), a negated sentence (in Exps.~1n, 2n, and 3n), a sentence with {\em perhaps} (in Exps.~1m, 2m, 3m), or a conditional sentence (in Exps.~1c, 2c, 3c).
-Sample trials of the at-issueness measures implemented in the 12 experiments are provided in Fig.~\ref{fig:ai-exps1} for Exps.~1, in Fig.~\ref{fig:ai-exps2} for Exps.~2, and in Fig.~\ref{fig:ai-exps3} for Exps.~3.
-Across all these measures, participants gave their responses on a slider marked `no' at one end (coded as 1) and `yes' at the other (coded as 0).
-Greater not-at-issueness of the CC should result in higher slider ratings.
-
-\begin{figure}[h!]
-\centering
-
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1q-ai}}
-\caption{Exp.~1q.}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1n-ai}} 
-\caption{Exp.~1n.}
- \end{subfigure}
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1m-ai}}
-\caption{Exp.~1m}
- \end{subfigure}%
- \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/1c-ai}}
-\caption{Exp.~1c}
- \end{subfigure}
-
-\caption{Target trials in at-issueness blocks of Exps.~1 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps1}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2q-ai}}
-\caption{Exp.~2q.}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2n-ai}} 
-\caption{Exp.~2n.}
- \end{subfigure}
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2m-ai}}
-\caption{Exp.~2m}
- \end{subfigure}%
- \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/2c-ai}}
-\caption{Exp.~2c}
- \end{subfigure}
-
-\caption{Target trials in at-issueness blocks of Exps.~2 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps2}
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3q-ai}}
-\caption{Exp.~3q.}
-\end{subfigure}%
-\begin{subfigure}[t]{0.5\textwidth}
-\centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3n-ai}} 
-\caption{Exp.~3n.}
- \end{subfigure}
-\begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3m-ai}}
-\caption{Exp.~3m}
- \end{subfigure}%
- \begin{subfigure}[t]{0.5\textwidth}
-        \centering
-\fbox{\includegraphics[height=5.4cm,width=7.6cm]{figures/3c-ai}}
-\caption{Exp.~3c}
- \end{subfigure}
-
-\caption{Target trials in at-issueness blocks of Exps.~3 for the complement {\em Julian dances salsa}.}\label{fig:ai-exps3}
-\end{figure}
-
-To assess whether participants were attending to the task, each experiment also included six control stimuli, which were also utterances made by a named speaker.
-Depending on the experiment, the content of these stimuli was expected to receive either high or low certainty ratings, and either high or low ratings on the at-issueness measure; see Supplement \ref{a-control} for details and the control stimuli.
-
-Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else.
-Each participant saw a random set of 26 stimuli: each set contained one target utterance for each of the 20 contents (each paired with a unique clause-embedding predicates) and the same 6 (experiment-specific) control stimuli.
-Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
-Block order and within-block trial order was randomized.
-After completing the experiment, participants filled out a short optional demographic survey.
-To encourage truthful responses, participants were told that they would be paid no matter what answers they gave in the survey.
-
-\subsection{Data exclusion}
-
-We excluded data from participants who took any experiment more than once and who did not self-declare to be native speakers of American English.
-We also excluded data from participants based on their ratings on the main clause controls and other criteria given in Supplement \ref{a-participants}.
-In each experiment, the data from between 215-281 remaining participants were analyzed.
-Information on the participants whose data entered into the analysis (total number, ages, gender) can be found in Supplement \ref{a-participants}.
-
-\newpage
-
-\section{Projection}
-
-Compare projectivity ratings across entailment-cancelling operators. 
-\begin{itemize}
-  \item is there a difference between projectivity across contexts? 
-  \item is there a difference between verbs in terms of how much embedding context makes a difference (semi-factive v factive)?
-  \item previous point that non-factive (i.e. veridical) predicates can also project can be made here for other embedding operators as well
-\end{itemize}
-
-\section{At-issueness}
-
-At-issueness diagnostics (kind of) work for questions:
-\begin{itemize}
-  \item asking whether diagnostic (exp. 1q)
-  \item Q-A diagnostic (exp. 2q)
-\end{itemize}
-
-\section{GPP}
-
-Replicating results from \cite{tbd-variability}: Correlation between NAI \& projection for PolQs found for:
-\begin{itemize}
-  \item asking whether diagnostic (exp. 1q)
-  \item Q-A diagnostic (exp. 2q)
-\end{itemize}
-
-\noindent Both of these also show some interesting lexical exceptions:
-\begin{itemize}
-  \item non-veridical predicates
-  \item some specific kinds of reportatives
-  \item how can we systematically identify exceptions based on data analysis?
-  \item are these the same kinds of exceptions as in the glossa paper?
-\end{itemize}
-
-\noindent Further analyses:
-\begin{itemize}
-  \item look at how many unique participants
-
-  \item negative correlation between Q/A and assent might be compound of a) embedding (negation) and b) assent, because we don?t see negative correlation, but no correlation, with other embeddings (modal, conditional)
-
-  \item  compare variance for both projection and at-issueness, under the different embeddings and diagnostics: very little variance for assent diagnostic/negation embedding
-
-  \item have a look at the predicates that are exceptional to the GPP in Exps1 and 2, how do they behave across the other experiments? Exp3: looks like they are also their own little group
-
-  \item  when we compare negation, modal, conditional with assent diagnostic, we can see what effect embedding has
-
-  \item next experiments: no embedding of the 20 predicates, with assent diagnostic, to compare to prior literature who didn?t use embedding with the assent diagnostic; do assent diagnostic without ?that?s true?
-anaphor to engage with Snider?s assumption that ?yes?
-is not really anaphoric in the assent diagnostic and hence also not in the Q/A diagnostic
-
-\end{itemize}
-
-\section{General discussion}\label{s4}
-
-\section{Conclusions}\label{s5}
-
-% end document here for word count
-%\end{document}
-
-\bibliographystyle{cslipubs-natbib}
-\bibliography{bibliography}
-
-\newpage
 
@@
   \end{exe}
+
   As shown in Table \ref{t-controls},  the group mean on the control stimuli is numerically higher (at .22 or .25) than for Exp.~2q (.07).
@@
   \end{exe}
+
   The mean at-issueness rating for the control stimuli in Exp.~3q was comparatively higher (at .28) than for Exp.~1q (.05) or Exp.~2q (.07).

### 2024-02-14 writing/investigating-GPP.tex
updated analysis scripts

@@
 
-Prijection comparison hasn't been done, except in CommitmentBank paper (right?, perhaps A'dam colloquium presentation?).
+Projection comparison hasn't been done, except in CommitmentBank paper (right?, perhaps A'dam colloquium presentation?).
 

### 2024-02-14 results/main/meta-analyses/0_data prep/preprocessing.R

@@ -1,0 +1,93 @@
+
+
+### data prep for projection data
+
+# set wd to script dir
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load data from exps into list
+df_list <- list(data_1q = read.csv("../../1_projaiQ/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_1n = read.csv("../../2_projaiN/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_1m = read.csv("../../3_projaiM/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_1c = read.csv("../../4_projaiC/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_2q = read.csv("../../5_projaiQ/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_2n = read.csv("../../6_projaiN/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_2m = read.csv("../../7_projaiM/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_2c = read.csv("../../8_projaiC/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_3q = read.csv("../../9_projaiQ/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_3n = read.csv("../../10_projaiN/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_3m = read.csv("../../11_projaiM/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=","),
+                data_3c = read.csv("../../12_projaiC/data/data_preprocessed.csv", 
+                                   header=TRUE, sep=",")
+)
+
+# libraries for manipulating dataframes, strings, and lists
+require(tidyverse)
+library(stringr)
+library(rlist)
+
+# spread responses over separate columns for projectivity and at-issueness
+# add label for which position at block appeared in
+df_list <- lapply(df_list, function(df) {
+  df = df %>%
+    mutate(ai_block = ifelse(question_type == "ai", 
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/1_summary-graphs.R

@@ -1,0 +1,247 @@
+##### Graphs for certainty ratings #####
+
+# 1 load libraries, data, etc --------------------------------------------------
+
+library(tidyverse)
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# summary + graphing
+library(ggplot2)
+source('../../../../helpers.R')
+cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
+                "#CC79A7", "#000000")
+graphcolor <- "#0072B2"
+pinkk <- "#CC79A7"
+pred_colors <- c(pinkk, "black")
+theme_set(theme_bw())
+
+# LOAD DATA
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
+data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), 
+                        operator = as.factor(op), participant = as.factor(workerid), 
+                        item = as.factor(content), .keep = "none")
+levels(data$operator) <- c("conditional", "modal", "negation", "question")
+str(data)
+
+
+# 2 projection by operator (Figure 2) ------------------------------------------
+
+# projection means and 95% CIs by operator 
+operator_means <- data %>% group_by(operator) %>% 
+  summarize(Mean = mean(projection), CILow = ci.low(projection), 
+            CIHigh = ci.high(projection)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
+         operator = fct_reorder(as.factor(operator), Mean)) %>% ungroup()
+operator_means
+operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
+
+# plot
+operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), 
+                              y = projection, label = round(projection, digits = 2))) +
+  geom_violin(data = data, scale="width", fill = graphcolor, 
+              alpha = .4, color = graphcolor) +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "black") +
+  geom_point(size=0.5, color = "black") +
+  geom_text(hjust = 0, nudge_x = -0.08, nudge_y = -0.06) + 
+  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
+                     labels = c("0","0.2","0.4","0.6","0.8","1"),
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R

@@ -1,0 +1,60 @@
+##### ANALYSIS BY OPERATORS #####
+
+# 1 load libraries, data, etc --------------------------------------------------
+
+library(tidyverse)
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# analysis
+library(rstan)
+options(mc.cores=parallel::detectCores())
+rstan_options(auto_write=TRUE)
+library(brms)
+library(tidybayes)
+library(emmeans)
+
+
+# LOAD DATA
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
+data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), 
+                        operator = as.factor(op), participant = as.factor(workerid), 
+                        item = as.factor(content), .keep = "none")
+levels(data$operator) <- c("conditional", "modal", "negation", "question")
+str(data)
+
+
+# 2 ANALYSIS -------------------------------------------------------------------
+# BAYESIAN MIXED EFFECTS BETA REGRESSION
+
+# Response: certainty ratings (rescale for beta regression)
+# Scaling to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
+# using method used in Degen & Tonhauser (2022), from Smithson & Verkuilen (2006), for proportional data.
+# y' = (y * (n - 1) + 0.5) / n
+data$betaresponse <- (data$projection*(nrow(data)-1) + .5)/nrow(data)
+data$betaresponse <- as.numeric(data$betaresponse)
+summary(data$betaresponse)
+
+# beta regression w response: scaled projection ratings, fixed effect: operator
+# random effects: random intercepts for item, participant, slopes for both by operator
+# also estimating these effects on beta precision parameter
+beta_formula = bf(betaresponse ~ operator + (1 | predicate) + (1 | item) + (1 | participant),
+                  phi ~ operator + (1 | predicate) + (1 | item)  + (1 | participant), # beta distribution's precision
+                  family = Beta())
+
+# dummy coding for operator fixed effect w baseline modal  
+data$operator <- relevel(data$operator, ref = "modal")
+contrasts(data$operator)
+
+m.b.operators <- brm(formula = beta_formula,
+    family = Beta(),
+    data = data,
+    cores = 4, iter = 4000, warmup = 700,
+    control = list(adapt_delta = .95,max_treedepth=15))
+
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/3_analysis-predicates.R

@@ -1,0 +1,253 @@
+##### ANALYSIS BY OPERATOR FOR EACH PREDICATE #####
+
+# 1 load libraries, data, etc --------------------------------------------------
+
+library(tidyverse)
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# analysis
+library(rstan)
+options(mc.cores=parallel::detectCores())
+rstan_options(auto_write=TRUE)
+library(brms)
+library(tidybayes)
+library(emmeans)
+
+# summary + graphing
+library(ggplot2)
+library(ggh4x)
+source('../../../../helpers.R')
+cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
+                "#CC79A7", "#000000")
+graphcolor <- "#0072B2"
+pinkk <- "#CC79A7"
+pred_colors <- c(pinkk, "black")
+theme_set(theme_bw())
+
+# LOAD DATA
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
+data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), 
+                        operator = as.factor(op), participant = as.factor(workerid), 
+                        item = as.factor(content), .keep = "none")
+levels(data$operator) <- c("conditional", "modal", "negation", "question")
+str(data)
+
+# 2 ANALYSIS -------------------------------------------------------------------
+# USING BAYESIAN MIXED EFFECTS BETA REGRESSION
+
+# Response: certainty ratings (rescale for beta regression)
+# Scaling to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
+# using method used in Degen & Tonhauser (2022), from Smithson & Verkuilen (2006), for proportional data.
+# y' = (y * (n - 1) + 0.5) / n
+data$betaresponse <- (data$projection*(nrow(data)-1) + .5)/nrow(data)
+data$betaresponse <- as.numeric(data$betaresponse)
+summary(data$betaresponse)
+
+# beta regression w response: scaled projection ratings, fixed effect: operator
+# random effects: item (embedded clause) intercepts and item slopes by operator
+# (no participant ranodm effects, bc each part saw each item only once)
+# also estimating these effects on beta precision parameter
+beta_formula = bf(betaresponse ~ operator + (1 + operator | item),
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/analysis.R

@@ -1,0 +1,92 @@
+##### DATA ANALYSIS FOR PROJECTION RATINGS #####
+
+# set wd to script dir
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load packages
+library(tidyverse)
+
+# load data
+data <- read.csv("data_combined.csv", header = TRUE, sep = ",")
+
+
+# variables and contrasts for data analysis: ----
+data <- data %>% mutate(projective = as.numeric(projective), 
+                        predicate = as.factor(verb), operator = as.factor(op), 
+                        participant = as.factor(workerid), 
+                        item = as.factor(content), .keep = "none")
+
+# contrasts
+contrasts(data$pred)
+### dummy coding with "be annoyed" as baseline
+data$pred <- relevel(data$pred, ref = "be_annoyed")
+contrasts(data$pred)
+
+## embedding operator, coded as "op"
+levels(data$op)
+contrasts(data$op)
+### op dummy coding with "n" (negation) as baseline
+data$op <- relevel(data$op, ref = "q")
+contrasts(data$op)
+
+
+# coding random effects as factors
+data$workerid <- as.factor(data$workerid)
+data$item <- as.factor(paste(data$content, data$pred))
+## content should not be a random effect, since contents are uniquely paired w
+## preds, which are independent variable
+## nmo
+
+# exploring a couple of models (linear)
+## library for linear mixed models
+library(lme4)
+library(lmerTest)
+library(knitr)
+
+# linear model w op as predictor only; q baseline
+glmm0 <- lmer(projective ~ op + (1 | workerid) + (1 | item), data = data)
+summary0 <- summary(glmm0)
+print(summary0, cor=F, dig=3)
+kable(summary0$coefficients, format = "latex", booktabs = TRUE, dig=2)
+
+### op dummy coding with "n" (negation) as baseline
+data$op <- relevel(data$op, ref = "n")
+contrasts(data$op)
+
+
+# glmm1 <- lmer(projective ~ op * pred + (1 | workerid), data=data)
+# save.image("linear-models.RData")
+load("linear-models.RData")
+
+summary1 <- summary(glmm1)
+print(summary1, cor=F, dig=3)
+kable(summary1$coefficients, format = "latex", booktabs = TRUE, dig=2)
+
+
+### dummy coding with "know" as baseline
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/by-predicate-analysis.R

@@ -1,0 +1,278 @@
+#### BY-PREDICATE ANALYSIS + graph
+
+
+
+
+
+
+#### full model output  ----
+
+# analysis2 = print(xtable(values),
+#                   #only.contents = T,
+#                   include.rownames=FALSE,
+#                   include.colnames=TRUE,
+#                   tabular.environment="longtable",
+#                   floating=FALSE,
+#                   hline.after = NULL,
+#                   latex.environments=NULL,
+#                   booktabs=TRUE,
+#                   sanitize.text.function = function(x){x},
+#                   comment = F
+# )
+
+
+
+
+#### create data for presenting results as part of Figure 4 ----
+
+# there are three lines in each facet (that is, for each expression)
+# one line for each context (x = context)
+# geom_segment(aes(x=1,xend=2,y=-.05,yend=-.05), linetype = "solid")
+# x, xend, y, yend, linetype: depend on expression and contrast
+# create data called "contrasts"
+# with columns "expression" and "context" (for binding and faceting with nat.means)
+# and with columns "x", "xend", "y", "yend", and "linetype"
+
+contrasts = data.frame(predicate = character(), contrast = character(), linetype = numeric())
+contrasts
+
+# show in table w colors
+# linetypes
+# 0 blank
+# 1 solid
+# 2 dashed
+# 3 dotted
+
+for (p in predicates) {
+  for (i in 1:nrow(get(paste("pairwise.",p,sep="")))) {
+    # print(i)
+    cntrst = get(paste("pairwise.",p,sep=""))$contrast[i]
+    lower = get(paste("pairwise.",p,sep=""))[get(paste("pairwise.",p,sep=""))$contrast == cntrst,]$.lower
+    upper = get(paste("pairwise.",p,sep=""))[get(paste("pairwise.",p,sep=""))$contrast == cntrst,]$.upper
+    value = get(paste("pairwise.",p,sep=""))[get(paste("pairwise.",p,sep=""))$contrast == cntrst,]$.value
+    l = ifelse(lower <= 0 & upper >= 0, 0,
+               ifelse(lower < 0 & upper < 0 & value <= -1.5, 1,
+                      ifelse(lower < 0 & upper < 0 & -1.5 < value & value <= -0.5, 6,
+                             ifelse(lower < 0 & upper < 0 & -.5 < value & value <= 0, 3,
+                                    ifelse(lower > 0 & upper > 0 & value >= 1.5, 1,
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/graphs.R

@@ -1,0 +1,204 @@
+
+
+# 4 projectivity by predicate and operator --------------------------------
+## means and confidence intervals for projection rating by predicate + operator
+
+
+
+
+
+
+# 5 verb profiles ---------------------------------------------------------
+## 5a projection by operator, for each verb (verb profiles)
+predicate_operator_means %>%
+  ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), y = projection, color = fct_reorder(operator, projection, .fun = "mean"), shape = fct_reorder(operator, projection, .fun = "mean"))) +
+  geom_violin(data = data, scale="width", alpha = .4) +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.5, color="black", alpha = .6) +
+  geom_point(size = 4) +
+  facet_wrap(vars(fct_reorder(predicate, projection, .fun = "mean"))) +
+  xlab("Operator") +
+  scale_shape_manual(values = c("M", "N", "Q", "C")) +
+  scale_colour_manual(values=cbbPalette) +
+  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
+                     labels = c("0","0.2","0.4","0.6","0.8","1"),
+                     name = "Certainty rating") +
+  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
+
+ggsave("../figures/predicate-profiles-1.pdf", width=10, height=10)
+
+
+
+predicate_operator_means$group = factor(
+  ifelse(predicate_operator_means$predicate %in% c("think", "pretend"), "negation-high", 
+  ifelse(predicate_operator_means$predicate  %in% 
+           c("demonstrate", "prove", "confirm"), "negation-low", 
+  ifelse(predicate_operator_means$predicate  %in%
+           c("admit", "confess", "announce", "reveal"), "modal-low",
+  ifelse(predicate_operator_means$predicate  %in% 
+           c("hear", "know", "inform"), "conditional-high",
+  ifelse(predicate_operator_means$predicate  %in% 
+           c("discover", "see", "acknowledge"), "question-high",
[…]

### 2024-02-14 results/main/meta-analyses/2_ai-ness/rscripts/graphs.R

@@ -1,0 +1,381 @@
+
+
+### summary statistics and graphs for projection data ------
+
+# set wd to script dir
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+# load data
+data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
+
+# load helper functions
+source('../../../../helpers.R')
+
+# libraries for manipulating dataframes, and plotting
+library(dplyr)
+library(forcats)
+library(ggplot2)
+
+### summary statistics and graphs for projection data ------
+
+# revert ai ratings back to original scale orientation
+data$ai <- 1-data$ai
+
+# add labels for ai-ness diagnostics ----
+data$exp = as.factor(paste(data$op, data$exp_block, ""))
+data$test = as.factor(paste(data$op, data$exp_block, ""))
+levels(data$test)
+new_levels <- c("sure?", "assent", "A/R")
+new_levels <- rep(new_levels, 3)
+new_levels <- append(new_levels, c("whether", "assent", "A/R"))
+levels(data$test) <- new_levels
+
+# mean ainess by diagnostic -----
+ai_means_test = data %>% group_by(test) %>%
+  summarize(Mean = mean(ai), CILow = ci.low(ai), 
+            CIHigh = ci.high(ai)) %>%
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
+         test = fct_reorder(as.factor(test),Mean)) %>% ungroup()
+
+ai_means_test %>% mutate(test = fct_reorder(test, Mean, 
+                                          .fun = mean)) %>% 
+  ggplot(aes(x = test, y=Mean)) +
+  coord_cartesian(ylim = c(0,1)) +
+  geom_point(aes(shape = test), size = 1, color = "lightblue") +
+  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
+  geom_line(color = "lightblue") + 
+  labs(title = "Mean ai-ness by diagnostic")+
+  theme_bw() +
+  scale_color_brewer(palette = "PRGn")
+
+ggsave("../graphs/ai-by-test.pdf")
+
+# mean ai-ness by operator -----
[…]

### 2024-02-14 results/main/meta-analyses/3_correlation/rscripts/analysis.R

@@ -1,0 +1,616 @@
+# meta-analyses
+# projection: comparing embedding under polar questions, negation, "perhaps", conditionals
+# different at-issueness diagnostics
+
+# set working directory to directory of script
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
+
+source('../../helpers.R')
+
+# load required packages
+require(tidyverse)
+library(ggrepel)
+library(lmerTest)
+
+# read in the data
+
+# at-issueness diagnostics: Q: asking whether, N/M/C: sure that
+exp1q = read.csv(file="../../1_projaiQ/data/data_preprocessed.csv") %>%
+  select(workerid,content,short_trigger,question_type,response) %>%
+  filter(short_trigger != "MC") %>%
+  spread(question_type,response) %>%
+  mutate(SubExperiment="exp1q")
+
+exp1n = read.csv(file="../../2_projaiN/data/data_preprocessed.csv") %>%
+  select(workerid,content,short_trigger,question_type,response) %>%
+  filter(short_trigger != "MC") %>%
+  spread(question_type,response) %>%
+  mutate(SubExperiment="exp1n")
+
+exp1m = read.csv(file="../../3_projaiM/data/data_preprocessed.csv") %>%
+  select(workerid,content,short_trigger,question_type,response) %>%
+  filter(short_trigger != "MC") %>%
+  spread(question_type,response) %>%
+  mutate(SubExperiment="exp1m")
+
+exp1c = read.csv(file="../../4_projaiC/data/data_preprocessed.csv") %>%
+  select(workerid,content,short_trigger,question_type,response) %>%
+  filter(short_trigger != "MC") %>%
+  spread(question_type,response) %>%
+  mutate(SubExperiment="exp1c")
+
+# at-issueness diagnostic: assent with positive continuation
+
+exp2q = read.csv(file="../../5_projaiQ/data/data_preprocessed.csv") %>%
[…]

### 2024-02-14 results/main/meta-analyses/new/0_data prep/preprocessing.R

@@ -1,93 +1,0 @@
-
-
-### data prep for projection data
-
-# set wd to script dir
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-# load data from exps into list
-df_list <- list(data_1q = read.csv("../../1_projaiQ/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_1n = read.csv("../../2_projaiN/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_1m = read.csv("../../3_projaiM/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_1c = read.csv("../../4_projaiC/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_2q = read.csv("../../5_projaiQ/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_2n = read.csv("../../6_projaiN/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_2m = read.csv("../../7_projaiM/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_2c = read.csv("../../8_projaiC/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_3q = read.csv("../../9_projaiQ/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_3n = read.csv("../../10_projaiN/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_3m = read.csv("../../11_projaiM/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=","),
-                data_3c = read.csv("../../12_projaiC/data/data_preprocessed.csv", 
-                                   header=TRUE, sep=",")
-)
-
-# libraries for manipulating dataframes, strings, and lists
-require(tidyverse)
-library(stringr)
-library(rlist)
-
-# spread responses over separate columns for projectivity and at-issueness
-# add label for which position at block appeared in
-df_list <- lapply(df_list, function(df) {
-  df = df %>%
-    mutate(ai_block = ifelse(question_type == "ai", 
[…]

### 2024-02-14 results/main/meta-analyses/new/1_projectivity/rscripts/analysis.R

@@ -1,92 +1,0 @@
-##### DATA ANALYSIS FOR PROJECTION RATINGS #####
-
-# set wd to script dir
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-# load packages
-library(tidyverse)
-
-# load data
-data <- read.csv("data_combined.csv", header = TRUE, sep = ",")
-
-
-# variables and contrasts for data analysis: ----
-data <- data %>% mutate(projective = as.numeric(projective), 
-                        predicate = as.factor(verb), operator = as.factor(op), 
-                        participant = as.factor(workerid), 
-                        item = as.factor(content), .keep = "none")
-
-# contrasts
-contrasts(data$pred)
-### dummy coding with "be annoyed" as baseline
-data$pred <- relevel(data$pred, ref = "be_annoyed")
-contrasts(data$pred)
-
-## embedding operator, coded as "op"
-levels(data$op)
-contrasts(data$op)
-### op dummy coding with "n" (negation) as baseline
-data$op <- relevel(data$op, ref = "q")
-contrasts(data$op)
-
-
-# coding random effects as factors
-data$workerid <- as.factor(data$workerid)
-data$item <- as.factor(paste(data$content, data$pred))
-## content should not be a random effect, since contents are uniquely paired w
-## preds, which are independent variable
-## nmo
-
-# exploring a couple of models (linear)
-## library for linear mixed models
-library(lme4)
-library(lmerTest)
-library(knitr)
-
-# linear model w op as predictor only; q baseline
-glmm0 <- lmer(projective ~ op + (1 | workerid) + (1 | item), data = data)
-summary0 <- summary(glmm0)
-print(summary0, cor=F, dig=3)
-kable(summary0$coefficients, format = "latex", booktabs = TRUE, dig=2)
-
-### op dummy coding with "n" (negation) as baseline
-data$op <- relevel(data$op, ref = "n")
-contrasts(data$op)
-
-
-# glmm1 <- lmer(projective ~ op * pred + (1 | workerid), data=data)
-# save.image("linear-models.RData")
-load("linear-models.RData")
-
-summary1 <- summary(glmm1)
-print(summary1, cor=F, dig=3)
-kable(summary1$coefficients, format = "latex", booktabs = TRUE, dig=2)
-
-
-### dummy coding with "know" as baseline
[…]

### 2024-02-14 results/main/meta-analyses/new/1_projectivity/rscripts/graphs.R

@@ -1,462 +1,0 @@
-##### GRAPHS FOR PROJECTION RATINGS #####
-
-# 1 libraries, helpers, settings
-# 2 projectivity by operator
-# 3 projectivity by predicate
-  ## 3a projectivity by predicate w distribution of participants' means
-  ## 3b projectivity by predicate w distribution of observations
-# 4 projectivity by predicate and operator
-  ## 4b projectivity means with distributions for observations
-  ## 4c projectivity means with distributions for paritipant means
-# 5 verb profiles
-
-
-# 1 Libraries,  helpers, settings -------------------------------------------
-
-# R SETUP / PACKAGES / DEFINITIONS
-library(tidyverse)
-library(ggplot2)
-source('../../../../helpers.R')
-# Color blind friendly palette (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)):
-cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", "#CC79A7", "#000000")
-cbbRotate <- c("#0072B2", "#CC79A7", "#000000", "#D55E00", "#F0E442", "#E69F00", "#56B4E9", "#009E73")
-graphcolor <- "#0072B2"
-theme_set(theme_bw())
-
-# LOAD DATA
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
-data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), 
-                        operator = as.factor(op), participant = as.factor(workerid), 
-                        item = as.factor(content), .keep = "none")
-levels(data$operator) <- c("conditional", "modal", "negation", "question")
-  
-str(data)
-  
-
-# 2 projection by operator ------------------------------------------------
-# projection means and 95% CIs by operator 
-operator_means <- data %>% group_by(operator) %>% 
-  summarize(Mean = mean(projection), CILow = ci.low(projection), 
-            CIHigh = ci.high(projection)) %>%
-  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
-         operator = fct_reorder(as.factor(operator), Mean)) %>% ungroup()
-operator_means
-operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
-
-# plot
[…]

### 2024-02-14 results/main/meta-analyses/new/2_ai-ness/rscripts/graphs.R

@@ -1,381 +1,0 @@
-
-
-### summary statistics and graphs for projection data ------
-
-# set wd to script dir
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-# load data
-data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
-
-# load helper functions
-source('../../../../helpers.R')
-
-# libraries for manipulating dataframes, and plotting
-library(dplyr)
-library(forcats)
-library(ggplot2)
-
-### summary statistics and graphs for projection data ------
-
-# revert ai ratings back to original scale orientation
-data$ai <- 1-data$ai
-
-# add labels for ai-ness diagnostics ----
-data$exp = as.factor(paste(data$op, data$exp_block, ""))
-data$test = as.factor(paste(data$op, data$exp_block, ""))
-levels(data$test)
-new_levels <- c("sure?", "assent", "A/R")
-new_levels <- rep(new_levels, 3)
-new_levels <- append(new_levels, c("whether", "assent", "A/R"))
-levels(data$test) <- new_levels
-
-# mean ainess by diagnostic -----
-ai_means_test = data %>% group_by(test) %>%
-  summarize(Mean = mean(ai), CILow = ci.low(ai), 
-            CIHigh = ci.high(ai)) %>%
-  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
-         test = fct_reorder(as.factor(test),Mean)) %>% ungroup()
-
-ai_means_test %>% mutate(test = fct_reorder(test, Mean, 
-                                          .fun = mean)) %>% 
-  ggplot(aes(x = test, y=Mean)) +
-  coord_cartesian(ylim = c(0,1)) +
-  geom_point(aes(shape = test), size = 1, color = "lightblue") +
-  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
-  geom_line(color = "lightblue") + 
-  labs(title = "Mean ai-ness by diagnostic")+
-  theme_bw() +
-  scale_color_brewer(palette = "PRGn")
-
-ggsave("../graphs/ai-by-test.pdf")
-
-# mean ai-ness by operator -----
[…]

### 2024-02-14 results/main/meta-analyses/old/rscripts/analysis.R

@@ -1,616 +1,0 @@
-# meta-analyses
-# projection: comparing embedding under polar questions, negation, "perhaps", conditionals
-# different at-issueness diagnostics
-
-# set working directory to directory of script
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-source('../../helpers.R')
-
-# load required packages
-require(tidyverse)
-library(ggrepel)
-library(lmerTest)
-
-# read in the data
-
-# at-issueness diagnostics: Q: asking whether, N/M/C: sure that
-exp1q = read.csv(file="../../1_projaiQ/data/data_preprocessed.csv") %>%
-  select(workerid,content,short_trigger,question_type,response) %>%
-  filter(short_trigger != "MC") %>%
-  spread(question_type,response) %>%
-  mutate(SubExperiment="exp1q")
-
-exp1n = read.csv(file="../../2_projaiN/data/data_preprocessed.csv") %>%
-  select(workerid,content,short_trigger,question_type,response) %>%
-  filter(short_trigger != "MC") %>%
-  spread(question_type,response) %>%
-  mutate(SubExperiment="exp1n")
-
-exp1m = read.csv(file="../../3_projaiM/data/data_preprocessed.csv") %>%
-  select(workerid,content,short_trigger,question_type,response) %>%
-  filter(short_trigger != "MC") %>%
-  spread(question_type,response) %>%
-  mutate(SubExperiment="exp1m")
-
-exp1c = read.csv(file="../../4_projaiC/data/data_preprocessed.csv") %>%
-  select(workerid,content,short_trigger,question_type,response) %>%
-  filter(short_trigger != "MC") %>%
-  spread(question_type,response) %>%
-  mutate(SubExperiment="exp1c")
-
-# at-issueness diagnostic: assent with positive continuation
-
-exp2q = read.csv(file="../../5_projaiQ/data/data_preprocessed.csv") %>%
[…]

### 2024-02-14 writing/SuB-paper/projection-operators-SuB.tex
tracking overleaf changes

--- 2024-02-05 writing/SuB-paper/projection-operators-SuB.tex
@@
 
+\usepackage{subcaption}
+
 % JT: I've commented this out so that we can see page numbers while writin
 %\pagestyle{empty}
+
+\newcommand{\mcdm}[1]{\textbf{\color{blue}[mcdm: #1]}}
 
@@
 
-\usepackage{natbib} 
+\usepackage[longnamesfirst]{natbib} 
 %\bibpunct{(}{)}{;}{a}{,}{,}
@@
 	\usepackage{xcolor}
+    \definecolor{op-n-color}{HTML}{56B4E9}
+    \definecolor{op-m-color}{HTML}{E69F00}
+    \definecolor{op-c-color}{HTML}{009E73}
+    \definecolor{op-q-color}{HTML}{0072B2}
+    \definecolor{pred-fcv-color}{HTML}{CC79A7}
 
@@
 
-%{\large \textbf{Projection variability of clausal complements across different operators}}
 {\large \textbf{Projection variability: Is the family of sentences really a family?}}\footnote{We thank Taylor Mahler for assistance in collecting the data presented here as well as valuable comments.
@@
 
-\textbf{Abstract.} We present experimental evidence that the projection of the content of clausal complements \textbf{(i)} varies across entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from nonfactive predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
+\textbf{Abstract.} We present experimental evidence \textbf{(i)} that the projection of the content of clausal complements  varies across entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factives predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
 The observed variability is not captured by existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
@@
 \section{Introduction} \label{sec:intro}
-
-    Textbooks typically present the family-of-sentences diagnostic for projection without addressing by-operator variation (e.g., \citealt{chierchia_meaning_1990}).
-For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from its variants in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
+    The so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective in formal semantics/pragmatics (e.g., \citealt{chierchia_meaning_1990}).
+For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.
 
@@
 For instance, to diagnose the content of the clausal complement of \emph{discover} as projective content, a researcher would typically observe that the content that Julian dances salsa follows not just from \ref{ex:simple}, but also from the family-of-sentences variants of \ref{ex:simple} in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
-
-    % (e.g., \citealt{frege_uber_1892,strawson_referring_1950,kiparsky_fact_1970,karttunen_conventional_1979}).
 
@@
 			{\bf Negation:} \hfill
-			\emph{Cole didn't discover that Julian dances salsa.}
+			Cole didn't discover that Julian dances salsa.
 		\b. \label{ex:q}
 			{\bf Polar Question:} \hfill
-			\emph{Did Cole discover that Julian dances salsa?}
+			Did Cole discover that Julian dances salsa?
 		\c. \label{ex:mod}
 			{\bf Modal:} \hfill
-			\emph{Perhaps Cole discovered that Julian dances salsa.}
+			Perhaps Cole discovered that Julian dances salsa.
 		\d. \label{ex:cond}
 			{\bf Conditional:} \hfill
-			\emph{If Cole discovered that Julian dances salsa, Logan will be joyful.}
-		\z.
-	\z.
-
-    However, prior research is suggestive of variation in the effect of different entailment-cancelling operators on projection.
-For instance, \citet{karttunen_observations_1971} proposed a distinction between English factive predicates (e.g., \textit{regret}) and semi-factives (e.g., \textit{discover}).
-    Based on the examples in \Next, he suggested that the CC of true factives projects across all four operators, but that of semi-factives does not always project from under polar questions, modals, or conditionals.\footnote{See \citealt{beaver_have_2010} for naturally occurring examples in which the CCs of factive predicates do not project from under the four operators, thereby calling into question the distinction between factive and semi-factive predicates.
-For experimental research on the distinction see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
-We return to this literature in section \ref{s:analysis}.}
+			If Cole discovered that Julian dances salsa, Logan will be joyful.
+		\z.
+	\z.
+
+    Some research, however, suggests that entailment-cancelling operators may affect projection differentially.
+For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}).
+Based on \Next, he suggested that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%
 
@@
 
-    Second, \citet{sieker_projective_2022}  compared the projection of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') between the four operators in \ref{ex:family}.
-The results showed that the CCs of the predicates investigated projected more from under negation than the other three operators.
-contrary to what \citealt{karttunen_observations_1971} suggested, a comparison of factive and semifactive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of nonfactive predicates.
+    Second, \citet{sieker_projective_2022}  compared the projection of the CCs of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') from under the four operators in \ref{ex:family}.
+The results suggest that the CCs project more from under negation than from under the other three operators.
+Contrary to what \citet{karttunen_observations_1971} suggested, a comparison of the factive and semi-factive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of semi-factive predicates.\footnote{The factive/semi-factive distinction is also called into question by naturally occurring examples where the CCs of factive predicates do not project from under the four operators (see \citealt{beaver_have_2010,de_marneffe_did_2012,de_marneffe_commitmentbank_2019}).
+For experimental research on the distinction, see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
+    }
 
@@
 
-    This paper reports on the results of a set of experiments that were designed to compare projection from under the four entailment-canceling operators in \ref{ex:family}.
-Our experiments extend the empirical scope of prior research by investigating projection for a broader set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}:
+    This paper reports on the results of a set of experiments that were designed to compare projection from under the four entailment-canceling operators in \ref{ex:family} in English.
+Our experiments extend the empirical scope of prior research on by-operator projection variation by investigating projection for a larger set of contents, namely the contents of the complements of the 20 English clause-embedding predicates in \ref{ex:predicates}, from \citet{degen_are_2022}.
 
     \ex. \label{ex:predicates}
-    		\a. (Semi)factive predicates: {\em be annoyed, know, reveal, discover, see}
-      \b. Nonfactive predicates: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, be right, demonstrate, pretend, say, suggest, think}
-
-    As shown, the  20 predicates include both (semi)factive and nonfactive predicates.
-The inclusion of the latter set of predicates is motivated by the results of the empirical investigations in \citealt{degen_are_2022}, which suggested that the contents of complements of nonfactive predicates is also projective and that projection does not categorically distinguish factive and nonfactive predicates.
-As \poscite{degen_are_2022} result was established based on sentences where the predicates were embedded under polar questions, our experiments, which feature all four entailment-canceling operators, serve to replicate and extend this result of \citealt{degen_are_2022}.
+    		\a. (Semi-)factive predicates: {\em be annoyed, know, reveal, discover, see}
+      \b. Non-factive predicates: {\em acknowledge, admit, announce, confess, confirm, establish, hear, inform, prove, be right, demonstrate, pretend, say, suggest, think}
+
+    The five predicates in \Last[a] have been characterized as factive or semi-factive.
+Our set of predicates also includes the 15 non-factives in \Last[b].
+Including non-factive predicates in investigations of projection is motivated by the empirical investigations in \citealt{de_marneffe_commitmentbank_2019} and \citealt{degen_are_2022}, which suggest that the CCs of non-factives may also project and that projection ratings do not categorically distinguish factive and non-factive predicates.
+
+    The results of our study suggest that the projection of the CCs of these 20 predicates is affected differently by the four entailment-cancelling operators in \ref{ex:family},  but not in a way that is consistent with \poscite{karttunen_observations_1971} factive/semi-factive distinction.
+The results also replicate a result of \citealt{degen_are_2022}, namely that projection from under polar questions does not categorically distinguish factive and non-factive predicates.
+Finally, the results of our experiment extend this result of \citealt{degen_are_2022} to projection from under the other three entailment-canceling operators, thereby solidifying their claim that ``research on projective content has a much broader empirical scope than previously assumed'' (p.585), as this scope includes the CCs of both factive and non-factive predicates.
 
@@
 
-\section{Experiments}
-
-	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.\footnote{The data and analysis scripts can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.
-{\bf LETS BE MORE SPECIFIC HERE ABOUT WHERE THINGS RELATED TO THIS PAPER ARE}} The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4.
-Each set of experiments consisted of three experiments that differed in the at-issueness measure used in a separate block.
-In this paper, we limit our attention to the projection ratings collected in the 12 experiments.
-
-	Projection was measured in each experiment with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} This diagnostic was also used in \citealt{sieker_projective_2022}.
-Under this diagnostic, participants are presented with utterances like those in \Next.
-They are asked to rate whether the (named) speaker is certain of the CC.
+\section{Experiments}\label{s2}
+
+	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
+    %
+    The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4.
+Each set of experiments contained three experiments, that differed in an at-issueness measure used in a separate block.
+In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}
+
+	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} This diagnostic was also used in \citealt{sieker_projective_2022}.
+Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.
 
 	\ex. \label{ex:certain-that}
-		\a. Christopher: \emph{\lq Cole didn't discover that Julian dances salsa.\rq}
-		\b. Christopher: \emph{\lq Did Cole discover that Julian dances salsa?\rq}
-		\c. Christopher: \emph{\lq Perhaps Cole discovered that Julian dances salsa.\rq}
-		\d. Christopher: \emph{\lq If Cole discovered that Julian dances salsa, Logan will be joyful.\rq}
+		\a. Christopher: ``Cole didn't discover that Julian dances salsa.''
+		\b. Christopher: ``Did Cole discover that Julian dances salsa?''
+		\c. Christopher: ``Perhaps Cole discovered that Julian dances salsa.''
+		\d. Christopher: ``If Cole discovered that Julian dances salsa, Logan will be joyful.''
 		\z.
@@
 
-    We assume, following \citealt{tonhauser_how_2018} and \citealt{degen_are_2022}, that judgments of speaker certainty about some propositional content reflect speaker commitment to that content, that is, projection.
-If a participant interprets an utterance like \Last[a--d] in a way that the speaker (e.g., Christopher) is certain of the CC, it is assumed to project out of the entailment-canceling environment.
-If a participant does not take the speaker to be certain of the CC, this is taken to indicate that it does not project.
+    We assume, following \citet{tonhauser_how_2018} and \citet{degen_are_2022}, that judgments of speaker certainty about the embedded content reflect speaker commitment to that content, that is, projection.
+If a participant interprets utterances like \Last[a--d] in a way that the speaker (here, Christopher) is certain of the CC, the CC is assumed to project.
+If a participant does not take the speaker to be certain of the CC, the CC is taken to not project.
 
@@
 			We recruited 250-300 participants for each of the 12 experiments.
-Participants for Exp.~1q were recruited on Amazon's Mechanical Turk platform; the participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs.
-Participants for the remaining experiments were recruited on Prolific; these participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
-See Supplement \ref{app:d-participants} for further information about the participants.
+Participants for one experiment were recruited on Amazon's Mechanical Turk platform.
+These participants were required to have U.S.\ IP addresses and at least 99\% of previously approved HITs.
+Participants for the remaining experiments were recruited on Prolific.
+These participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
+See Supplement \ref{app:d-participants} (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.
+
+            \begin{figure}[h!]
+				\centering
+				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
+				\caption{A sample trial in Exps.~2. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
+				\label{fig:trial}
+			\end{figure}
 
@@
 
-            The target sentences consisted of 400 combinations of the 20 clause-embedding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}).
+            The target sentences consisted of the 400 combinations of the 20 clause-embedding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}).
 As mentioned above, the predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal \emph{perhaps} in Exps.~3, and in conditional antecedents in Exps.~4, for a total of 400 target stimuli in each of the four sets of experiments.
-
-			%We tested the 20 clause-embedding predicates in \Next, which have shown projection variability in polar questions (\citealt{degen_are_2022}).
-They comprise three factive predicates \Next[a], two semi-factives \Next[b], nine that have been characterized as optionally factive in \citealt{kiparsky_fact_1970} \Next[c], and six non-factives which include two veridical predicates \Next[d], and four non-veridical ones \Next[e].
-
-	%		\ex. 20 clause-embedding predicates \hfill \citealt{degen_are_2022}
-			%	\a. canonically factive:					\emph{be annoyed, know, reveal}
-				%\b. semi-factive: 		\emph{discover, see}
-%				\b. optionally factive: \newline \emph{acknowledge admit, announce, confess, confirm, establish, hear, inform, prove}
-%				\b. Non-factive veridical:
-%					\emph{be right, demonstrate}
-%				\b. Non-factive non-veridical:
-%				\emph{pretend, say, suggest, think}
-%				\z.
-%			\z.
-
-To assess whether participants were attending to the task, each experiment included six control stimuli. For details on the six control stimuli see Supplement \ref{app:c-control}.
+To assess whether participants were attending to the task, each experiment included six control stimuli.
+For details on the six control stimuli, see Supplement \ref{app:c-control}.
 
@@
 
-			Participants were told to imagine that they are at a party and that, on walking into the kitchen, they overhear somebody say something to somebody else.
+			Participants were told to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
             %
             On each trial, participants read an utterance and responded to the `certain that'   question.
-They gave their response on a slider marked `no' (coded as 0) at one end and and `yes' (coded as 1) on the other.
-A sample trial from Exps.~1 is illustrated in Figure~\ref{fig:trial}.
-Following, e.g., \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things:
-            %
-            On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection.
-This assumes that projection (and the relevant notion of speaker commitment) can itself be gradient.
-            %
-            Conversely, higher ratings could also reflect the probability that a participant associates with a projective interpretation for the utterance.
-Under this interpretation, commitment may be interpreted categorically, and the gradience is a result of the uncertainty about possible interpretations (see also discussion in \citealt{grove_factivity_2023}, which suggests the latter kind of interpretation).
+They gave their response on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other.
+A sample trial from Exps.~1 is given in Figure~\ref{fig:trial}.
+
+            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things:
+            %
+            On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection. This assumes that speaker commitment is gradient.
+            %
+            Conversely, higher certainty ratings could reflect the probability that an interpreter takes the speaker to be committed to the CC.
+On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of the uncertainty about possible interpretations.
+%, which suggests the latter kind of interpretation).
+            In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation (for discussion, see also \citealt{grove_factivity_2023}).
 
             %  we take high ratings of speaker certainty to reflect speaker greater speaker commitment towards the CC, that is, greater projection.
-
-			\begin{figure}[ht]
-				\centering
-				\fbox{\includegraphics[width = .7\linewidth]{task-1n-proj.eps}}
-				\caption{A sample trial in Exps.~1. In the corresponding trials in Exps.~2-4, participants were presented with an utterance with a different entailment-cancelling operator.}
-				\label{fig:trial}
-			\end{figure}
 
@@
     We first address the question of whether there is by-operator variation (section \ref{s:by-op}) and then the question of whether there is by-predicate variation in the observed by-operator variation (section \ref{s:by-pred}).
+Finally, in section \ref{s:factivity}, we relate the results to the result of \citealt{degen_are_2022}, that projection from under polar questions does not categorically distinguish factive and non-factive predicates.
 
@@
 
-        Fig.~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates.
-As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relative more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.
+        \mcdm{again needs to be consistent, sometimes Figure} Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates.
+As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.
 
@@
 
-   These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.
+   These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.\mcdm{we don't know what Model \#1 is}
 
@@
 
-			Like \citet{sieker_projective_2022}, we observe a small main effect of entailment-cancelling operator on projection.
-However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.
+			Like \citet{sieker_projective_2022} for German, we observe a small main effect of entailment-cancelling operator on projection.
+However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.\mcdm{does it make sense to compare different predicates?}
             We note, however, that the differences in mean certainty ratings between the four entailment-canceling operators are very small.
@@
 
-  Fig.~\ref{fig:op-pred-ratings} shows mean certainty ratings by predicate and entailment-canceling operator (aggregating over the 20 CCs that each predicate was paired with).
-As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
-The observed variation does not align with widely assumed predicate classes.
-For instance, the five (semi)factive predicates exhibit five different patterns:
-i) the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modal {\em perhaps}; ii) the CC of \emph{know} projects most from under questions, less from questions and negation, and least from under the modal {\em perhaps}; iii) the CC of \emph{see} projects most from under questions and conditionals, less from under negation, and least from under the modal \emph{perhaps}; iv) the CC of  \emph{discover} projects more from under conditionals and questions than negation and modal {\em perhaps}; and v) the CC of \emph{reveal} projects most from conditionals, less from questions, and least from negation and modal \emph{perhaps}.
-Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
-We also observe by-operator projection variation for nonfactive predicates.
-Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}.
-Other nonfactive predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.
+    Figure~\ref{fig:op-pred-ratings} shows mean certainty ratings by entailment-canceling operator for the five (semi-)factive predicates (indicated in pink), and the 15 non-factive predicates, where predicates are ordered by their overall mean certainty ratings.
 
@@
 		\includegraphics[width = \linewidth]{predicate-operator-graph-1}
-				\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals. Predicates are ordered by mean certainty rating pooled over the four operators.
-Embedding operator coded by letter and color:  \texttt{N} (light blue): negation, \texttt{M} (orange): modals, \texttt{C} (green): conditional antecedents, \texttt{Q} (dark blue): polar questions.}
+		\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
+\textcolor{pred-fcv-color}{(Semi-)factive} and non-factive predicates are ordered by mean certainty rating pooled over the four operators.
+        Embedding operator coded by letter and color:  \textcolor{op-n-color}{\texttt{N}: negation}, \textcolor{op-m-color}{\texttt{M}: modal \emph{perhaps}}, \textcolor{op-c-color}{\texttt{C}: conditional antecedents}, \textcolor{op-q-color}{\texttt{Q}: polar questions}.}
 		\label{fig:op-pred-ratings}
@@
 
-The observation that there is by-operator variation for the CC of each of the 20 clause-embedding predicates is confirmed by the statistical analysis. We fitted...
-
-  The CC of \emph{know} projects less from under negation than questions, but more from under negation than modals, while the difference between negation and conditionals is not significant.
-			%
-			These generalizations are supported by models\ \# 2--4 in Table\ \ref{t:models}, which also each have at least $34$ highly significant interaction terms (out of $57$ possible interactions of operator and predicate).
-
-			\begin{table}[ht]
+    As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
+Further, the effect of operator differs between predicates.
+Examining the differences in certainty-ratings between the predicates for the four operators suggests that the observed variation does not align with widely assumed predicate classes.
+For instance, the five (semi-)factive predicates exhibit five different patterns:
+
+    i) the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modal {\em perhaps}; ii) the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under the modal {\em perhaps}; iii) the CC of \emph{see} projects most from under questions and conditionals, less from under negation, and least from under the modal \emph{perhaps}; iv) the CC of  \emph{discover} projects more from under conditionals and questions than negation and modal {\em perhaps}; and v) the CC of \emph{reveal} projects most from conditionals, less from questions, and least from negation and modal \emph{perhaps}.
+Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
+We also observe by-operator projection variation for non-factives predicates.
+Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}.
+Other non-factives predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.
+
+    These generalizations are confirmed by the statistical analysis.
+We fitted a Bayesian mixed-effects beta regression to the subset of the data containing the observations for each predicate, using the \texttt{brms} package (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}).
+Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect) on scaled\footnote{The beta regression model assumes a beta-distribution defined over the open unit interval $(0,1)$, excluding boundaries.
+Following \citet{degen_are_2022}, we scaled the response variable, which is distributed over the closed interval $[0,1]$ (including boundaries), using the method in \citealt{smithson_better_2006}.} projection ratings, with random by-item intercepts (where an item is a complement clause), and by-item slopes for operators.
+
+    For each predicate, the differences between operators were established by a pairwise comparison of posterior samples for the four operators using the \texttt{emmeans} package (\citealt{lenth_emmeans_2024}).
+This comparison yielded the $95\%$ high density interval (HDI) of estimated marginal mean differences for each pairwise combination of operators.
+We assume that two conditions differ when their HDI does not include $0$.
+
+    The results are indicated in Figure~\ref{fig:op-pred-analysis}, which shows shows mean certainty ratings by entailment-canceling operator for the 20 predicates, here grouped by predicate.
+As in Figure~\ref{fig:op-pred-ratings}, (semi-)factive predicates are indicated in pink, and predicates are ordered by their overall mean certainty ratings.
+Non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
+
+    \begin{figure}[t]
         \centering
-				\begin{tabular}{llrrrr}
-					Model & & Estimate & Std. Error & t-value\\
-					\midrule
-					\#2 & Intercept: \emph{\bf be annoyed}/negation & 0.87 & 0.01 & 79.86 & ***\\
-					& operator: conditional & -0.12 & 0.02  & -7.36 & ***\\
-					& operator: modal & -0.16 & 0.02  & -10.01 & ***\\
-					& operator: question & 0.02 & 0.01 & 1.72 & n.s.\\
-					\midrule
-					\#3 & Intercept: \emph{\bf discover}/negation & 0.68 & 0.01 & 62.70 & ***\\
-					& operator: conditional & 0.11 & 0.02 & 7.11 & ***\\
-					& operator: modal & -0.06 & 0.02 & -3.63 & ***\\
-					& operator: question & 0.10 & 0.01 & 7.08 & ***\\
-					\midrule
-					\#4 & Intercept: \emph{\bf know}/negation & 0.79 & 0.01 & 72.97 & ***\\
-					& operator: conditional & 0.00 & 0.02 & -0.06 & n.s.\\
-					& operator: modal & -0.14 & 0.02 & -9.18 & ***\\
-					& operator: question & 0.08 & 0.01 & 5.67 & ***\\
-					\bottomrule
-				\end{tabular}
-				\caption{\small Excerpt of the output from three linear mixed effects models, with fixed effects: operator, predicate, and their interaction; random effect: participant intercepts.
-				Models were fit with \texttt{lme4, lmertest} in \texttt{R}.
-Models \textbf{\#2--4} also had at least $34$ highly significant interaction terms of \texttt{operator} and \texttt{predicate} with $p < 0.001$ (not shown here).\label{t:models}}
-			\end{table}
-
-Our results align with those of \citealt{smith_relationship_2014}, who also observed by-expression by-operator variation.
+		\includegraphics[width = \linewidth]{analysis.pdf}
+		\caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, non-factive) and operator (\textcolor{op-m-color}{modal},  \textcolor{op-n-color}{negation}, \textcolor{op-q-color}{polar questions}, \textcolor{op-c-color}{conditional antecedents}) with 95\% bootstrapped confidence intervals.
+Violin plots indicate kernel probability density of observations.
+Below each facet, a line spanning two operators indicates a non-zero difference according to the pairwise comparison of operators.
+The linetype indicates whether the difference is
+        at least $1$ (solid line: —),
+        below $1$ but at least $0.5$ (dashed line: – –),
+        or below $0.5$ but above $0$ (dotted line: . . . )}
+		\label{fig:op-pred-analysis}
+	\end{figure}
+
+    Because we found statistically robust non-zero differences by-operator differences for each of the predicates, and these differ beween predicates, the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
+
+    As a result, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator.
 However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find an effect here.
@@
 
-			%Our result that by-operator projection variability interacts with predicate concurs with \citet{smith_relationship_2014}, while we did not reproduce their result that the CC of \emph{know} projects more from negation than conditionals (we found no difference here).
-   Our results differ from those of \citealt{sieker_projective_2022}: While that work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
+	Our results differ from those of \citealt{sieker_projective_2022}: While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
 As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic.
@@
 
-	\subsubsection{Factive vs.\ nonfactive predicates}
-
-Factive predicates are defined as predicates that presuppose the CC, while the CC of nonfactive predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entailment-canceling operators, this definition leads one to expect that factive predicates are distinguished from nonfactive ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of nonfactive ones.
-This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates}, which were embedded in polar questions.
-Contrary to expectation,  \citealt{degen_are_2022} found variation among the CCs of the five factive predicates investigated and that the CCs of some nonfactive predicates were as projective, or even more projective, than the CCs of some factive predicates.
-In short, projection of the CC from under polar questions did not categorically distinguish factive from nonfactive predicates.
-
-The results of the set of experiments reported on in this paper replicate this result.
-As shown in Fig.~\ref{fig:op-pred-ratings}, there is variation between the CCs of the five factive predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from nonfactive predicates.
-Furthermore, the results of  this set of experiments suggest that projection from under the three other entailment-canceling operators also does not categorically distinguish factive and nonfactive predicates.
-This result strengthens the conclusion of \citealt{degen_are_2022} that there is, to date no empirical evidence for a coherent class of factive predicates.
-
-IF THERE IS ROOM I WOULD LIKE TO INCLUDE 4 VERSIONS OF FIG 3 WHERE THE 20 PREDICATES ARE ORDERED BY OPERATOR, TO MAKE THE POINT OF THIS SECTION CLEARER.
-THAT WOULD ALSO BE GOOD FOR THE METHODOLOGICAL IMPLICATION SECTION BELOW.
-
-		Modal:\newline
-		\includegraphics[width = \linewidth]{modal-predicate-graph-1.pdf}
-        Negation:\newline
-        \includegraphics[width = \linewidth]{negation-predicate-graph-1.pdf}
-        Polar questions:\newline
-        \includegraphics[width = \linewidth]{question-predicate-graph-1.pdf}
-        Conditionals:\newline
-        \includegraphics[width = \linewidth]{conditional-predicate-graph-1.pdf}
-		%\caption{Mean certainty ratings by predicate, for each operator.
+	\subsubsection{Factive vs.\ non-factives predicates}\label{s:factivity}
+
+Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker10,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entail- ment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones.
+This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates} embedded in polar questions.
+Contrary to expectation,  \poscite{degen_are_2022} Exps.~1 found that the CCs of the five factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives predicates were as projective, or even more projective, than the CCs of some factive predicates.
+In short, projection of the CC from under polar questions did not categorically distinguish factive from non-factives predicates.
+Further support for this result came from the analysis of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).
+
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[t]{0.57\textwidth}
+\includegraphics[width = 1\linewidth]{question-predicate-graph-1.pdf}
+\caption{Polar question}\label{fig:op-pred-ratings-q}
+\end{subfigure}
+
+\medskip
+\begin{subfigure}[t]{0.57\textwidth}
+\includegraphics[width = 1\linewidth]{modal-predicate-graph-1.pdf}
+\caption{Modal \emph{perhaps}}\label{fig:op-pred-ratings-m}
+\end{subfigure}
+
+\medskip
+\begin{subfigure}[t]{0.57\textwidth}
+\includegraphics[width = 1\linewidth]{negation-predicate-graph-1.pdf}
+\caption{Negation}\label{fig:op-pred-ratings-n}
+\end{subfigure}
+
+\medskip
+
+\begin{subfigure}[t]{0.57\textwidth}
+ \includegraphics[width = 1\linewidth]{conditional-predicate-graph-1.pdf}
+\caption{Conditional}\label{fig:op-pred-ratings-c}
+\end{subfigure}
+\caption{Mean certainty ratings by predicate (with factives in violet) under (a) polar question, (b) modal \emph{perhaps}, (c) negation, (d) conditional.
 Error bars indicate $95\%$ bootstrapped confidence intervals.
-Violin plots indicate kernel probability densities of individual observations.}
+Violin plots indicate kernel probability densities of individual ratings.}\label{fig:op-pred-ratings2}
+\end{figure}
+
+The results of the experiments reported on in this paper replicate this result.
+As shown in Fig.~\ref{fig:op-pred-ratings-q}, there is variation between the CCs of the five factive predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from non-factives predicates.
+This result replicates the results of \poscite{degen_are_2022} Exps.~1.
+Furthermore, the results of the experiments reported on in section \ref{s2} suggest that projection from under the three other entailment-canceling operators also does not categorically distinguish factive and non-factives predicates.
+As shown in Figs.~\ref{fig:op-pred-ratings-m}-\ref{fig:op-pred-ratings-c}, there is variation in the projection of the CCs of the five factive predicates from under the other three entailment-canceling operators, and projection from under the other three entailment-canceling operators also does not categorically distinguish factive from non-factives predicates.
+These results strengthens the conclusion of \citealt{degen_are_2022} that there is, to date, no empirical evidence for a coherent class of factive predicates.
 
@@
 				\includegraphics[width = \linewidth]{mega-veridicality}
-				\caption{Mean projection ratings by embedding context and predicate.}
+				\caption{Mean projection ratings by embedding context and predicate in MegaVeridicality.}
 				\label{fig:figure4}
@@
 
-A methodological implication of this latter result is that future research must relativize claims about projection variation to the entailment-canceling operator.
+The results of the experiments reported on in section \ref{s2} suggest that there is little by-operator projection variation when we aggregate over the clause-embedding predicates and their CCs, but that there is by-operator variation that cannot be neglected when we do not aggregate.
+These results have two methodological implications.
+First, when initially investigating the projection of a content (or teaching projection to students), the family-of-sentences can indeed be treated as a family, as standard textbooks assume.
+However, it is advisable at this step already to apply the diagnostic with all four operators and attend to possible by-operator projection variation.
+
+A second methodological implication of the results of our experiments is that research on projection must consider which entailment-canceling operator to use, and report which entailment-canceling operator was used.
+This is because the results of the investigation may very well depend on the entailment-canceling operator under which the expressions and contents were embedded.
+For instance, two contents that exhibit projection variation from under polar questions might not exhibit such differences when embedded under a different entailment-canceling operator, and two contents that do not exhibit projection variation when embedded under negation, might exhibit variation when embedded under a different entailment-canceling operator.
 
@@
 In this section, we consider whether contemporary analyses can capture the observed by-predicate projection variation and the interactions between predicate and entailment-canceling operator in projection ratings.
-We consider three types of analyses: (i) analyses that assume that the CCs of factive predicates are lexically specified as presupposed  (\citealt{heim_projection_1983,heim_presupposition_1992,van_der_sandt_presupposition_1992}), (ii) analyses on which backgrounded entailments are projective (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) an analysis on which contextual entailments are projective
+%We consider three types of analyses: (i) analyses that assume that the CCs of factive predicates are lexically specified as presupposed  (\citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), (ii) analyses on which backgrounded entailments are projective (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) an analysis on which contextual entailments are presupposed
 %with a triggering algorithm based on world-knowledge that operates over contextual entailments of an utterance 
-(\citealt{schlenker_triggering_2021}).
+%(\citealt{schlenker_triggering_2021}).
 
@@
 
-\subsubsection{Analyses on which presuppositions are lexically specified}
-
-On the analyses in \citealt{heim_projection_1983,heim_presupposition_1992} and \citealt{van_der_sandt_presupposition_1992}, the CCs of factive predicates are lexically specified as presupposed, in contrast to the CCs of nonfactive predicates.
+\subsubsection{\citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992}}\label{s:heim}
+
+On the analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992}, the CCs of factive predicates are lexically specified as presupposed, in contrast to the CCs of non-factives predicates.
 Presupposed CCs project to the global context, except when that would produce an inconsistency, in which case they are accommodated to the local context of the operator.
 
-\paragraph{By-predicate variation.} Analyses on which some clause-embedding predicates are factive and others are nonfactive do not predict the observed by-predicate variation, for several reasons.
+These analyses do not predict the observed by-predicate variation, for several reasons.
 One reason is that there is variation between the factive predicates: For instance, the CC of \emph{be annoyed} is more projective than that of \emph{discover}, which in turn is more projective than that of \emph{reveal}.
-This variation is not predicted because the CCs of factive predicates are invariably specified as presupposed.
-The variation is also not predicted because, in our experiment, the stimuli were presented in minimal contexts that were not inconsistent with the CCs, thereby rendering local accommodation unlikely.
-A second reason is that the CCs of some nonfactive predicates (e.g., \emph{inform, acknowledge})  are just as projective as those of some factive predicates, or even more projective.
-Analyses like those in \citealt{heim_projection_1983,heim_presupposition_1992} and \citealt{van_der_sandt_presupposition_1992} fail to make predictions about the projection of the CCs of nonfactive predicates.
-Thus, as discussed in detail in \citealt{degen_are_2022}, the factive/nonfactive distinction is not sufficiently fine-grained to adequately predict the observed by-predicate projection variation.
+This variation is not predicted because the CCs of factive predicates are invariably specified as presupposed, with no mechanism to predict projection variation between factive predicates.
+(Recourse to local accommodation is not possible because the stimuli in our experiments were presented in minimal contexts that were not inconsistent with the CCs.)  A second reason is that the CCs of some non-factives predicates (e.g., \emph{inform, acknowledge})  are just as projective as those of some factive predicates, or even more projective.
+Analyses like those in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} fail to make predictions about the projection of the CCs of non-factives predicates.
+Thus, as discussed in detail in \citealt{degen_are_2022}, the factive/non-factive distinction is not sufficiently fine-grained to adequately predict the observed by-predicate projection variation.
 
@@
 
-	\paragraph{By-operator variation.} 
         %In our experiment, the stimulus utterances were presented out of the blue, leaving it to the participants to imagine a suitable discourse context.
@@
 		%
-		The analyses in \citealt{heim_projection_1983,heim_presupposition_1992} and \citealt{van_der_sandt_presupposition_1992} do not lead us to expect differential effects of entailment-cancelling operators on projection.
-This is because negation and conditional antecedents are given a semantics that derives their behavior as presuppositional \lq holes\rq\ (in the sense of \citealt{karttunen_observations_1971}).
-For instance in \citealt{heim_projection_1983}, presupposition triggers under negation or in conditional antecedents are evaluated relative to the global context set.
-Therefore, presupposition triggers under these operators are predicted to behave like unembedded ones, with no expected differences.
-While the analyses does not explicitly address epistemic possibility modals or polar questions, we might expect that they would be treated as presuppositional holes along the same lines, therefore also not predicting the observed projection variation.
+		The analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} also do not predict differential effects of entailment-cancelling operators on projection.
+This is because negation and conditional antecedents are given a semantics that derives their behavior as presuppositional holes, in the sense of \citealt{karttunen_observations_1971}.
+For instance in \citealt{heim_projection_1983}, presupposition triggers under negation or in conditional antecedents are evaluated relative to the global context, with the consequence that presuppositions are not expected to differentially project from under negation or conditional antecedents.
+%Therefore, presupposition triggers under these operators are predicted to behave like unembedded ones, with no expected differences.
+  While the analysis does not explicitly address epistemic possibility modals or polar questions, we might expect that they would be treated as presuppositional holes along the same lines, therefore also not predicting any by-operator projection variation, by by-predicate variation in the effect of entailment-canceling operators.
 		%
@@
 
-	\subsubsection{Analyses on which backgrounded entailments are projective}
-
-		\citealt{abrusan_predicting_2011,abrusan_presupposition_2016} and \citealt{simons_best_2017} do not assume that presuppositions are lexically specified but rather that those entailments that are backgrounded with respect to the discourse context in which the utterance is made may project.
+	\subsubsection{\citealt{abrusan_predicting_2011} and \citealt*{simons_best_2017}}\label{s:abrusan}
+
+\citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} do not assume that presuppositions are lexically specified.
+In \citealt{abrusan_predicting_2011}, a lexical entailment of an uttered sentence is a presupposition if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}) addressed by the utterance.
+For instance, the CC of B's utterance in \ref{ent}, that Phil's ballet class is canceled, is predicted to be a presupposition (and therefore to project) because it is a lexical entailment of the sentence embedded under the epistemic modal and not at-issue with respect to A's interrogative utterance.
+
+\ex.\label{ent} Adapted from \citealt[188]{simons_best_2017} \\ Context: It's early on Saturday morning. A and B are talking about their son. 
+\a.[A:] Why is Phil up already?
+\b.[B:] Perhaps he forgot that his ballet class is canceled today.
+
+In \citealt{simons_best_2017}, the CC of a clause-embedding predicate projects if it is entailed by the Current Question of the utterance (where the Current Question is the question that is congruent with the utterance).\footnote{The Current Question is defined in \citealt[194]{simons_best_2017} as follows: ``The Current Question for an utterance is a privileged subset of the focal alternative set of the uttered sentence (given a structural analysis of that sentence, including focus marking)'' which meets the conditions that ``i) the proposition expressed is a member of the Current Question and ii) the Current Question has at least one additional member.''} In \ref{ent},  the Current Question of B's utterance might be the set of propositions \{Phil forgot that his ballet class is canceled today, Phil is aware that his ballet class is canceled today\}.
+If so, the Current Question entails that Phil's ballet class is canceled today, and the CC may therefore project under \poscite{simons_best_2017} analysis.
+
 %assume that projection is not determined lexically, but by a triggering mechanism, which operates over the entailments of an expression, and determines which of them projects.
@@
 The assumed triggering mechanisms differ on two dimensions: (i) how the set of inferences is determined that constitutes the input to the triggering mechanism, and (ii) the algorithm by which these inferences are derived as projective or not.
-		In \citealt{abrusan_predicting_2011} and \citealt{abrusan_presupposition_2016}, an entailment is backgrounded if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}).
-In \citealt{simons_best_2017}, an entailment is backgrounded if it is not at-issueness relative to the QUD and not conventionally marked as at-issue.
+
 		%
@@
 
-		\paragraph{By-predicate variation.} Like the analyses discussed in the previous section, the analyses in \citealt{abrusan_predicting_2011,abrusan_presupposition_2016} and \citealt{simons_best_2017} do not predict the observed by-predicate variation.
-First, if the CCs of factive predicates are entailed, they are invariably predicted to project, and so the observed differences are not predicted.
-Second, the analyses do not make predictions for nonentailed CCs, and hence they do not predict that the CCs of some nonfactive predicates project at least as much or even more than those of some factive predicates.
+Like the analyses discussed in the previous section, the analyses in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} do not predict the observed by-predicate variation.
+First, if the CCs of factive predicates are entailed, they are invariably predicted to project, and so the observed differences between the CCs of different factive predicates are not predicted.
+Second, the analyses do not make predictions for non-entailed CCs, and hence they do not predict that the CCs of some non-factives predicates project at least as much or even more than those of some factive predicates.
+Neither of the analyses makes predictions about by-operator projection variation, nor about by-predicate variation in the effect of entailment-canceling operators.
 
@@
 
-		\paragraph{By-operator variation.} Neither of the analyses make predictions about by-operator variation.
+		%\paragraph{By-operator variation.} Neither of the analyses make predictions about by-operator variation.
 			%Since we found an operator/predicate interaction effect on projection, a further prediction of this type of approach would be that at-issueness ratings would show an operator-predicate interaction in this experimental setup as well.
@@
 
-\subsubsection{Triggering based on contextual entailment}
-
-The triggering mechanism assumed in \citealt{schlenker_triggering_2021} operates on local contextual entailments of an expression, and determines which one of them projects based on probabilistic world knowledge about the truth of the invoked contents.
+\subsubsection{\citealt{schlenker_triggering_2021}}
+
+On the analysis proposed in \citealt{schlenker_triggering_2021}, the CC of a sentence S, like \ref{knowa}, that is uttered in a context $c$ is presupposed if the CC is presupposed by the sentence S$'$under the entailment-canceling operator, that is \ref{knowb}, in the local context $c'$.
+For the CC to be presupposed in \ref{knowb}, two conditions must be met: i) S$'$ contextually entails the CC relative to $c'$; and ii) If we consider ``a generic agent'' who believes the propositions in $c'$ and who has now learned about the truth of S$'$, then the probability that this generic agent already believed the CC is above a contextual threshold $\alpha$ (\S6.2); more colloquially, condition ii) requires that the generic agent ``typically antecedently believes'' the CC (p.6) upon interpreting S$'$ in $c'$.
+Applied to \ref{know}, \poscite{schlenker_triggering_2021} analysis predicts that the CC  is presupposed if i) \ref{knowb} contextually entails that Julian dances salsa, and ii) if a generic agent would typically antecedently believe that Julian dances salsa upon interpreting \ref{knowb} in the minimal contexts we provided to our participants.
+(For \citealt{schlenker_triggering_2021}, the local context under negation $c'$ is identical to $c$.)
+
+\ex.\label{know} 
+\a.\label{knowa} Cole doesn't know that Julian dances salsa.
+\b.\label{knowb} Cole knows that Julian dances salsa.
+\z.
+
+Condition i) is met under the assumption that the CC of \ref{knowb} is an entailment.
+\citealt{schlenker_triggering_2021} also assumes the condition ii) is met: ``in many cases, one's knowledge of facts will precede one's knowledge of [Cole's] beliefs about them\ldots believing that [Julian dances salsa] is often an epistemic precondition for believing that'' Cole knows that Julian dances salsa (p.6).
+One might, however, challenge this assumption on the basis of the corpus study presented in \citealt{spenader02}, which showed that the CCs of the majority of the utterances of sentences with factive verbs (namely 81 out of 109) had to be accommodated (i.e., were not contextually entailed).
+In other words, utterance of sentences with the factive predicates investigated by \citet{spenader02}, which included \emph{know}, were ``generally used to communicate information the speaker thought was hearer-new'' (p.99).
+This result might therefore suggest that one cannot assume that a generic agent typically antecedently believes the CC of {\em know}.
+
+We assume that it is an open, empirical question which predicates are such that the probability of a generic agent antecedently believing the CC is above the contextual threshold $\alpha$ in the minimal contexts we provided our participants (and, of course, what that threshold might be).
+\citealt{schlenker_triggering_2021} appears to assume that there are two classes of predicates: those where the probability is above the threshold (including {\em know, inform}, and {\em announce}), and those where it is not (including {\em demonstrate} and {\em establish}; see p.12 and appendix I).
+Thus, one advantage of \poscite{schlenker_triggering_2021} analysis over the analyses reviewed in sections \ref{s:heim} and \ref{s:abrusan} is that it is able to predict the projection of the CCs of non-factives predicates (modulo the open questions about condition ii)).
+
+Does \poscite{schlenker_triggering_2021} analysis capture the observed by-predicate variation?
+The analysis correctly predicts that the CCs of \emph{know, inform}, and \emph{announce} are more projective than the CCs of \emph{demonstrate} and \emph{establish}, by virtue of the CCs of the former being possibly presupposed, in contrast to the CCs of the latter.
+It is not clear, however, that the analysis is able to predict the observed variation because the analysis -- even  though it does not divide predicates into factive and non-factives ones -- nevertheless suggests a binary, categorical distinction between predicates, such that the CCs of predicates with probabilities above the threshold may be presupposed (depending on the context), and those below the threshold are not.
+First, as shown in Figs.~\ref{fig:op-pred-ratings} and \ref{fig:op-pred-ratings2}, and discussed extensively in \citealt{degen_are_2022}, there is projection variation between the supposedly presupposed CCs that is not captured by an analysis that does not make more fine-grained distinctions between the meanings of clause-embedding predicates.
+Furthermore, it is not clear that the supposedly nonpresupposed CCs are not projective.
+Second, as discussed in \citealt{degen_are_2022}, the CCs of non-factives predicates like \emph{establish} and \emph{demonstrate} are projective when compared to nonprojective main clause content.
+\poscite{schlenker_triggering_2021} analysis does not make predictions for these CCs.
+In the experiments reported on in section \ref{s2}, the mean certainty rating of the CC of \emph{announce}, whose CC is assumed to possibly be presupposed, is .XX, and that of {\em demonstrate}, whose CC is not assumed to be presupposed, is .XX (both means are aggregated across entailment-canceling operators).
+It is not clear whether this particular difference in mean certainty rating motivates analyzing the CC of \emph{announce} as presupposed in contrast to that of \emph{demonstrate}.
+
+%The triggering mechanism assumed in \citealt{schlenker_triggering_2021} operates on local contextual entailments of an expression, and determines which one of them projects based on probabilistic world knowledge about the truth of the invoked contents.
 For instance, the CC of the utterance with \emph{discover} in \Next[a], that Julian dances salsa, has the potential to project if it is entailed by the sentence in the scope of negation (namely \Next[b]) together with information that is available in its local context.
@@
 
-			\ex. \a. Cole didn't discover that Julian dances salsa.
-				\b. Cole discovered that Julian dances salsa.
-			\z.
-
-			The triggering mechanism states that an inference $p$ of some propositional expression $E$ pro\-jects if $p$ is an epistemic precondition for the truth of $E$ in its local context $c^\prime$.
+	%		\ex. \a. Cole didn't discover that Julian dances salsa.
+	%			\b. Cole discovered that Julian dances salsa.
+	%		\z.
+
+			%The triggering mechanism states that an inference $p$ of some propositional expression $E$ pro\-jects if $p$ is an epistemic precondition for the truth of $E$ in its local context $c^\prime$.
 So, the CC of \Last[a] projects if the CC is an epistemic precondition for the truth of \Last[b].
 			%
-			$p$ is an epistemic precondition of $E$ in $c^\prime$ if, usually, when one acquires the belief in $c^\prime$ that $E$, one already knows that $p$. 
-			%
-			The CC of \Last[a] can be taken to be an epistemic precondition for \Last[b] based on the \lq subjective conditional probability\rq\ that a generic epistemic agent would already believe that Julian dances salsa when learning that \Last[b] is true.
+			%$p$ is an epistemic precondition of $E$ in $c^\prime$ if, usually, when one acquires the belief in $c^\prime$ that $E$, one already knows that $p$. 
+			%
+			%The CC of \Last[a] can be taken to be an epistemic precondition for \Last[b] based on the \lq subjective conditional probability\rq\ that a generic epistemic agent would already believe that Julian dances salsa when learning that \Last[b] is true.
 Specifically, the CC is taken to be an epistemic precondition if that probability reaches a contextually given threshold.
-
-			\paragraph{By-predicate variation.} \citealt{schlenker_triggering_2021} predicts   
-   the potential of projection for CCs that are contextually entailed by the variants not embedded under an entailment-canceling operator.
-Thus, in contrast to the analyses reviewed above, this analysis can, in principle, predict the projection of non-entailed CCs, namely when the CCs are contextually entailed.
-While this is a step in the right direction, it is not clear that the analysis can predict the projection of non-entailed CCs.
-Consider \Next, the variant of \Last[a] with {\em inform}:
-
-   \ex. \a. Cole didn't inform Sam that Julian dances salsa.
-       \b. Cole informed Sam that Julian dances salsa
-
-On \poscite{schlenker_triggering_2021} analysis, the CC projects in \Last[a] if, usually, when one acquires the belief that \Last[b], one already knows that Julian dances salsa.
-While it is certainly possible for Julian dancing salsa to be already known when one acquires the belief that \Last[b], it is not clear that this is ``usually'' the case.
-(Another question, of course, is what ``usually'' means.)
-
-It is also not clear that this analysis can predict the observed by-predicate variation. To illustrate, consider the schemata in \Next:
-
-\ex. \a. Cole didn't PREDICATE that Julian dances salsa.
-    \b. Cole PREDICATE that Julian dances salsa.
-
-Given that the non-entailed CC of a given predicate projects in \Last[a] only if one ``usually'' already knows that the CC is true when one acquires the belief that \Last[b], \poscite{schlenker_triggering_2021} analysis seems to predict that predicates fall into two groups: Those whose CCs one ``usually'' already knows to be true when one acquires the belief that \Last[b], and those for which this is not the case.
-If so, this analysis cannot predict the observed by-predicate projection variation.
 
@@
 
-\paragraph{By-operator variation.} \citealt{schlenker_triggering_2021} does not incorporate differential effects of entailment-cancelling operators on projection.
-Although subjective conditional probabilities are relativized to a local context, the local context under negation and in conditional antecedents is assumed to the global context (based on \citealt{heim_projection_1983,schlenker_local_2009}).
-Therefore, no differential effect of operator is predicted, contrary to what is observed.
-For instance, the conditional probability for $p$ given $E$ vs.\ that of $p$ given $(\emph{not } E)$ or that of $p$ given $(\emph{if } p, q)$ should not differ.
+\poscite{schlenker_triggering_2021} analysis does not incorporate differential effects of entailment-cancelling operators on projection.
+For instance, the local contexts $c'$ under negation and in conditional antecedents are both assumed to be the global context $c$ (based on \citealt{heim_projection_1983,schlenker_local_2009}).
+Therefore, the analysis does not make predictions about by-operator projection variation or about by-predicate variation in the effect of entailment-canceling operator.
+
+%For instance, the conditional probability for $p$ given $E$ vs.\ that of $p$ given $(\emph{not } E)$ or that of $p$ given $(\emph{if } p, q)$ should not differ.
 
@@
 
-	\subsubsection{Discussion: Contextual inferences and probabilistic triggering}
-
-		Our data showed gradient projection variation, based on a main effect of predicate, and an operator-predicate interaction.
+\subsubsection{Interim summary}
+
+There is currently no projection analysis on the market that predicts the by-predicate and by-operator variation we observed in the experiments reported on in section \ref{s2}.
+\citealt[585]{degen_are_2022} suggested that an empirically adequate analysis to capture the observed by-predicate variation requires consideration of ``more fine-grained distinctions [between clause-embedding predicates] that are based on the lexical meaning and discourse use of clause-embedding predicates''.
+In the next section, we offer some hypotheses based on the data from our experiments.
+
+%	\subsubsection{Discussion: Contextual inferences and probabilistic triggering}
+
+%		Our data showed gradient projection variation, based on a main effect of predicate, and an operator-predicate interaction.
 We found these effects for the CCs of clause-embedding predicates under four entailment-cancelling operators in out-of-the-blue contexts.
@@
 
-		The main effect of predicate and the operator-predicate interaction are both gradient effects, and cannot be explained by a categorical distinction between presuppositional and non-pre\-suppo\-si\-tion\-al predicates.
+%		The main effect of predicate and the operator-predicate interaction are both gradient effects, and cannot be explained by a categorical distinction between presuppositional and non-pre\-suppo\-si\-tion\-al predicates.
 The effects could rather be addressed in a probabilistic account, where various embedding contexts (i.e., combinations of entailment-cancelling operator and predicate) are associated with different probabilities that the embedded content projects.
 
-		Under the assumption of information-structural triggering, this is the probability that the embedded content is not at issue in the (possibly imagined) discourse context.
+%		Under the assumption of information-structural triggering, this is the probability that the embedded content is not at issue in the (possibly imagined) discourse context.
 Under a world-knowledge-based triggering account, the projection probability can be associated with the probability that the embedded content is already known in the discourse context.
@@
 
+This paper investigated variation in projection from under the four entailment-canceling operators that have traditionally been used in the family-of-sentences diagnostic for projection, namely negation, polar questions, epistemic modals, and conditional antecedents.
+The results of our experiments suggest that the projection of the contents of the clausal complements of clause-embedding predicates varies across these operators.
+As discussed, there is currently no projection analysis on the market that is able to predict the observed by-predicate variation or the by-operator variation.
+The results of our experiments also extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factives predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
+This results strengthens the conclusion of \citealt{degen_are_2022} that there is, to date, no empirical evidence for a coherent class of factive predicates.
+
 \bibliographystyle{sub-chicago}
@@
 To minimize the variability of the effect of the contents of these consequent clauses on the projection of the contents of the 20 complement clauses, the consequent clauses all consist of a uniquely named subject and an adjectival predication in the future tense ({\em will be}), and the adjectives all denote an emotion.
-We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner_norms_2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
+We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner-etal2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
 
@@
 Participants may judge B's response in (3b) to be less acceptable than in (3a) because of this missing content specification.
-This hypothesis is consistent with the results of \posscite{syrett_experimental_2015} Exp.~3, which suggests that content of a sentence-medial NRRCs can be the target of a direct denial, though the main clause content is preferred as the target of such a denial.
+This hypothesis is consistent with the results of \posscite{syrett-koev2015} Exp.~3, which suggests that content of a sentence-medial NRRCs can be the target of a direct denial, though the main clause content is preferred as the target of such a denial.
 That the group means on the control stimuli in our Exps.~2n, 2m, and 2c are still relatively low may be due to the fact that the one content that was specified is the at-issue main clause content.

### 2024-02-14 writing/SuB-paper/supplement.tex
tracking overleaf changes

@@
 
+\section{Analysis 1 --- By operator}
+
+Bayesian mixed effects beta regression using the
+
+\section{Analysis 2 --- By operator analysis for each predicate}
+
+The results are indicated in Figure~\ref{fig:op-pred-analysis}, which shows shows mean certainty ratings by entailment-canceling operator for the 20 predicates, here grouped by predicate.
+As in Figure~\ref{fig:op-pred-ratings}, (semi-)factive predicates are indicated in pink, and predicates are ordered by their overall mean certainty ratings.
+Non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
+
+Because...
+the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
+
+These observations were confirmed by a posthoc pairwise comparison of the estimated means for each control and target expression in the explicit ignorance context using the ‘emmeans’ package (Lenth 2023) in R (R Core Team 2016).
+The input to the pairwise comparison was a Bayesian mixed-effects beta regression model with weakly informative priors that was fit using the ‘brms’ package (B¨urkner 2017).
+
+The model predicted naturalness ratings8 from a fixed effect of expression (with treatment coding and ‘continue’ as reference level) and included random by-participant and by-item intercepts (where an item is a complement clause).
+The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated
+marginal mean differences between each of the expressions. We assume that two expressions differ in their naturalness in the explicit ignorance context if their HDI does not include 0
+
 \end{document}

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/1_summary-graphs.R

@@ -8,7 +8,7 @@ setwd(this.dir)
 
 # summary + graphing
 library(ggplot2)
-source('../../../../helpers.R')
+source('../../../helpers.R')
 cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
                 "#CC79A7", "#000000")
 graphcolor <- "#0072B2"
@@ -92,7 +92,7 @@ predicate_means %>% ggplot(aes(x = fct_reorder(predicate, projection, .fun = "me
         legend.position="bottom", text = element_text(size=12)
   )
 
-ggsave(f="../figures/predicate-graph-1.pdf", width=7, height=4.5)
+ggsave(f="../graphs/predicate-graph-1.pdf", width=7, height=4.5)
 
 
 # 4 projection by predicate and operator (Figure 3) ----------------------------
@@ -126,7 +126,7 @@ predicate_operator_means %>%
         legend.position="none", text = element_text(size=12)
   )
 
-ggsave("../figures/predicate-operator-graph-1.pdf", width=7, height=4.5)
+ggsave("../graphs/predicate-operator-graph-1.pdf", width=7, height=4.5)
 
 # 5 projection by predicate with violin plots, for each operator (fig. 4) ------
 
@@ -245,3 +245,4 @@ subset(predicate_operator_means, operator == "conditional") %>%
   )
 
 ggsave("../graphs/modal-predicate-graph.pdf", width=7, height=3.75)
+

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/3_analysis-predicates.R

@@ -142,6 +142,10 @@ levels(predicate_operator_means$predicate)
 pred_order <- levels(predicate_operator_means$predicate)
 textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), pinkk, "white")
 
+data %>% mutate(operator = fct_reorder(operator, projection, .fun = "mean"))
+levels(data$operator)
+
+
 # identify pairwise differences
 contrasts = data.frame(predicate = character(), contrast = character(), mean = numeric(), lower = numeric(), upper = numeric())
 contrasts

### 2024-02-14 writing/SuB-paper/projection-operators-SuB.tex
tracking udpates from overleaf wed afternoon

@@
     \usepackage{rotating}
+    \usepackage{longtable}
 
@@
 For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}).
-Based on \Next, he suggested that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%
+Based on \Next, he argued that the content of the complement (CC) of true factives consistently projects across the operators in \Last, but that of semi-factives does not always project from under polar questions, modals, or conditionals.%
 
@@
 
-    %    The factive/semi-factive distinction (iii) is further called into question by \citealt{beaver_have_2010}, who reports many naturally occurring examples of cognitive factive predicates, showing that their CC does not necessarily project (e.g., English \emph{be/become aware, know, discover, notice, find out, realize, accept the fact}; French \emph{trouver} \lq discover\rq).
-
-    %Experimental findings from \citet{djarv_cognitive_2018}, \citet{tonhauser_how_2018}, and \citet{degen_are_2022} can be evaluated with respect to whether they support Karttunen's distinction (iii), with mixed results.
-These studies observed by-predicate variation in polar questions, where more projection is expected for factives than semi-factives.
-	%
-	%\citet{djarv_cognitive_2018}, assessing acceptability of affirming the main clause while denying the CC, did find higher ratings for English \emph{be happy} and \emph{appreciate} (assumed to be factive) and \emph{be aware} than \emph{realize} (assumed to be semi-factive).
-This could be interpreted as limited support for (iii).
-But here, it is also not obvious how exactly this task relates to projection.\footnote{Say why}
-
-	%\citet{tonhauser_how_2018} and \citet{degen_are_2022} measured speaker commitments toward the CC of a broad range of English clause-embedding predicates more directly, by collecting judgments of how certain participants take a speaker to be of the contents under investigation.
-	%
-	%The observed differences between the various predicates do not align with theoretically assumed lexical classes.
-While \citet{degen_are_2022} demonstrate this for the distinction between purported factive and non-factive predicates, the results did not match the expectations from Karttunen's factive vs.
-semi-factive classification either.
-	%
-	%\citet{tonhauser_how_2018} found, e.g., that the CC of semi-factive \emph{realize} was as projective as that of factive \emph{be annoyed} and more than that of semi-factive \emph{discover}.
-\citet{degen_are_2022} observed that the CC of factive \textit{reveal} was less projective than that of semi-factive \textit{discover}, which, in turn, was less projective than that of factive \textit{know}.
-These findings also suggest that examining projection ratings for various predicates individually, rather than pre-grouping them based on theoretical assumptions, can provide more nuanced insights into by-predicate projection variation.
-
-	%Both \citet{smith_relationship_2014} \citet{sieker_projective_2022} found that projection is affected differently by entailment-cancelling operators.
-However, while Smith \& Hall found variation in the effect of the operator on projection between various projective contents, Sieker \& Solstad did not.
-	%
-	%These divergent results raise the question if they are due to cross-linguistic variation, task differences, or artifacts of the analysis.
-Finally, much previous research calls into question the hypothesis that by-predicate differences in the effect of operator on projection pattern according to Kattunen's factive/semi-factive distinction (\citealt{beaver_have_2010,tonhauser_how_2018,sieker_projective_2022,degen_are_2022}).
-
-    %Karttunen's distinction gives rise to three hypotheses about the projection associated with clause-embedding predicates, which we explore in this paper: \textbf{(i) by-operator variation:} entailment-cancelling operators differentially affect projection; \textbf{(ii) operator/predicate interaction:} the effect of operator differs between predicates; \textbf{(iii) factive/semi-factive distinction:} this difference can be addressed by characterizing lexical classes of (semi-)factive predicates.
-
-    %To address the three hypotheses, we conducted a set of experiments designed to assess projection across the four entailment-canceling operators in \ref{ex:family} for various clause-embedding predicates.
-We used the same projection measure as \citet{sieker_projective_2022} (the `certain that' task; see e.g., \citealp{tonhauser_how_2018,djarv_prosodic_2017,mahler_social_2020}) and applied it to the CC of 20 English predicates, including factives and semi-factives, and 15 non-factive predicates, given recent findings that their complements may also project, albeit to varying degrees (\citealt{degen_are_2022}).
-Our findings support (i) by-operator variation, and an operator/predicate interaction (ii), but not Karttunen's factive/semi-factive distinction (iii).
-
 \section{Experiments}\label{s2}
-
 	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
@@
 
-	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.} This diagnostic was also used in \citealt{sieker_projective_2022}.
+	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019,sieker_projective_2022}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.}
+    %This diagnostic was also used in \citealt{sieker_projective_2022}.
     Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.
@@
 
-		\subsubsection{Participants}
+		%\subsubsection{Participants}
+        \paragraph{Participants.}
 			We recruited 250-300 participants for each of the 12 experiments.
@@
 				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
-				\caption{A sample trial in Exps.~2. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
+				\caption{A sample trial in Exps.~1. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
 				\label{fig:trial}
@@
 
-  %which were main clauses with appositive relative clauses, which were also presented as utterances made by a named speaker.
-Here, the main clause content was targeted by the projection question.
-In the experiments with polar questions, the controls were expected to receive low certainty ratings, whereas in the experiments with declaratives, they were expected to receive high ratings.
-
-  %Each participant saw all 20 predicates, randomly paired with a unique embedded clause from a set of 20 contents, and a subject from a set of 20 proper names.
-%The predicates associated with projective content were embedded under (i) negation, (ii) polar questions, (iii) the epistemic possibility modal \textit{perhaps}, and (iv) in conditional antecedents in the four different sets of experiments.
-
         \subsubsection{Procedure}
-
-			Participants were told to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
-            %
-            On each trial, participants read an utterance and responded to the `certain that'   question.
-They gave their response on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other.
-A sample trial from Exps.~1 is given in Figure~\ref{fig:trial}.
-
-            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things:
-            %
-            On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection. This assumes that speaker commitment is gradient.
-            %
+            Participants were asked to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
+            %
+            On each trial, they read an utterance and gave a response to the `certain that' question on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other.
+A sample trial from Exps.~1 is shown in Figure~\ref{fig:trial}.
+
+            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things: On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection.
+This assumes that speaker commitment is interpreted in a gradient way.
+            % gradient.
             Conversely, higher certainty ratings could reflect the probability that an interpreter takes the speaker to be committed to the CC.
@@
 
-            %  we take high ratings of speaker certainty to reflect speaker greater speaker commitment towards the CC, that is, greater projection.
-
             After completing the experiment, participants filled out a short optional demographic survey.
@@
         \subsubsection{Data exclusion}
-
             Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}.
@@
 
-            %We analyze the data from $2,682$ self-reported native speakers of American English, after excluding participants who failed one of the following two attention-checks: their mean rating on six control items is more than 2 sd different from the group mean, or they always roughly selected the same point on the scale for target stimuli (identified by low variance, and then manually inspecting their response patterns).
-
     \subsection{Results and discussion}
@@
 
-        \mcdm{again needs to be consistent, sometimes Figure} Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates.
+        Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates.
 As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.
-
-			%When aggregating across predicates, we found an effect of \texttt{operator}.
-The distribution of projection ratings, as well as means and $95\%$ bootstrapped confidence intervals by operator are shown in Figure~\ref{fig:op-ratings}.
 
@@
 				\centering
-				\includegraphics[scale = .8]{operator-graph-1}
-				\caption{Mean certainty ratings by operator.
-Error bars indicate $95\%$ bootstrapped confidence intervals.
-Violin plots indicate the kernel probability density of the individual participants’ ratings.}
+				\includegraphics[scale = .8]{certainty-operator.pdf}
+				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the ratings.
+                % of the individual participants’ ratings.
+                }
 				\label{fig:op-ratings}
@@
 
-			%Mean projection ratings were higher for CCs in question-embeddings than under negation and modals, but lower than in conditional antecedents.
-
-   These generalizations are supported by Model \#1 reported in Table~\ref{t:op-model}.\mcdm{we don't know what Model \#1 is}
-
-			\begin{table}[ht]
+   These observations were confirmed by a posthoc pairwise comparison of the estimated means for each entailment-canceling operator using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}).
+The input to the pairwise comparison was a  Bayesian mixed-effects beta regression model with weakly informative priors that were fit using the \emph{brms} package (\citealt{buerkner2017}).
+The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson-verkuilen2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level and included a random by-predicate intercept.
+The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators.
+We assume that two operators differ in certainty ratings if their HDI does not include 0.\footnote{A more detailed report of the model output is provided in Supplement \ref{app:analysis-1}}
+
+Table \ref{t:pairwise} provides the output of the pairwise comparisons. As shown,
+
+, the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.\footnote{While the naturalness ratings collected from the participants range from 0 to 1, the greater than $|$1$|$  estimated marginal mean differences are the result of the beta regression using a logit link function for the mean parameter.} There is also variation between the factive predicates and the presuppositional controls: While all five factive predicates are more natural in explicit ignorance contexts than {\em continue, too} and {\em also}, {\em know} is additionally more natural than {\em again} and {\em stop}, and {\em see, discover} and {\em reveal} are additionally more natural than the {\em it-}cleft.
+Finally, the naturalness ratings in explicit ignorance contexts for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}.
+These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for categorical projection or for a categorical distinction between factive and nonfactive predicates.
+
+\begin{table}[!h]
 \centering
-					\hspace{-1.3em}
-					\begin{tabular}{llrrrr}
-						Model & & Estimate & Std. Error & t-value\\
-						\midrule
-						\#1 & Intercept: \emph{question} & 0.51 & 0.01 & 44.78 & ***\\
-						& operator: conditional & 0.05 & 0.01 & 5.30 & ***\\
-						& operator: modal & -0.04 & 0.01 & -4.45 & ***\\
-						& operator: negation & -0.03 & 0.01 & -4.67 & ***\\
-						\bottomrule
+\begin{tabular}{lrrr}
+ contrast & estimate & lower 95\% CI & upper 95\% CI \\\hline 
+ conditional - negation & 0.21 & 0.18 & 0.24 \\ 
+  conditional - question & 0.13 & 0.10 & 0.16 \\ 
+  modal - conditional & -0.26 & -0.29 & -0.23 \\ 
+  modal - negation & -0.05 & -0.08 & -0.02 \\ 
+  modal - question & -0.14 & -0.17 & -0.11 \\ 
+  negation - question & -0.09 & -0.12 & -0.06 \\ 
 \end{tabular}
-
-					\caption{Excerpt from the model output of a linear mixed effects model with fixed effects of operator; random effects: participant and item intercepts, fitted with \texttt{lme4, lmertest} in \texttt{R}.\label{t:op-model}}
+\caption{Pairwise differences between entailment-canceling operators.}\label{t:pairwise}
 \end{table}
@@
 		Like \citet{sieker_projective_2022} for German, we observe a small main effect of entailment-cancelling operator on projection.
-However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.\mcdm{does it make sense to compare different predicates?}
+However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.
+        % \mcdm{does it make sense to compare different predicates?}
         We note, however, that the differences in mean certainty ratings between the four entailment-canceling operators are very small.
@@
 
-    Figure~\ref{fig:op-pred-ratings} shows mean certainty ratings by entailment-canceling operator for the five (semi-)factive predicates (indicated in pink), and the 15 non-factive predicates, where predicates are ordered by their overall mean certainty ratings.
-
-    \begin{figure}[ht]
-        \centering
-		\includegraphics[width = \linewidth]{predicate-operator-graph-1}
-		\caption{Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals.
-\textcolor{pred-fcv-color}{(Semi-)factive} and non-factive predicates are ordered by mean certainty rating pooled over the four operators.
-        Embedding operator coded by letter and color:  \textcolor{op-n-color}{\texttt{N}: negation}, \textcolor{op-m-color}{\texttt{M}: modal \emph{perhaps}}, \textcolor{op-c-color}{\texttt{C}: conditional antecedents}, \textcolor{op-q-color}{\texttt{Q}: polar questions}.}
-		\label{fig:op-pred-ratings}
-	\end{figure}
-
+        Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the five (semi-)factive predicates, and the 15 non-factive predicates, where predicates are ordered by their overall mean certainty ratings.
+        %
         As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
 Further, the effect of operator differs between predicates.
-Examining the differences in certainty-ratings between the predicates for the four operators suggests that the observed variation does not align with widely assumed predicate classes.
-For instance, the five (semi-)factive predicates exhibit five different patterns:
-
-    i) the CC of \emph{be annoyed} projects more from under negation and questions than conditionals and modal {\em perhaps}; ii) the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under the modal {\em perhaps}; iii) the CC of \emph{see} projects most from under questions and conditionals, less from under negation, and least from under the modal \emph{perhaps}; iv) the CC of  \emph{discover} projects more from under conditionals and questions than negation and modal {\em perhaps}; and v) the CC of \emph{reveal} projects most from conditionals, less from questions, and least from negation and modal \emph{perhaps}.
-Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
-We also observe by-operator projection variation for non-factives predicates.
-Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}.
-Other non-factives predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.
-
-    These generalizations are confirmed by the statistical analysis.
-We fitted a Bayesian mixed-effects beta regression to the subset of the data containing the observations for each predicate, using the \texttt{brms} package (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}).
-Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect) on scaled\footnote{The beta regression model assumes a beta-distribution defined over the open unit interval $(0,1)$, excluding boundaries.
-Following \citet{degen_are_2022}, we scaled the response variable, which is distributed over the closed interval $[0,1]$ (including boundaries), using the method in \citealt{smithson_better_2006}.} projection ratings, with random by-item intercepts (where an item is a complement clause), and by-item slopes for operators.
-
-    For each predicate, the differences between operators were established by a pairwise comparison of posterior samples for the four operators using the \texttt{emmeans} package (\citealt{lenth_emmeans_2024}).
-This comparison yielded the $95\%$ high density interval (HDI) of estimated marginal mean differences for each pairwise combination of operators.
-We assume that two conditions differ when their HDI does not include $0$.
-
-    The results are indicated in Figure~\ref{fig:op-pred-analysis}, which shows shows mean certainty ratings by entailment-canceling operator for the 20 predicates, here grouped by predicate.
-As in Figure~\ref{fig:op-pred-ratings}, (semi-)factive predicates are indicated in pink, and predicates are ordered by their overall mean certainty ratings.
-Non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
-
-    \begin{figure}[t]
+Examining the observed variation suggests that it does not align with widely assumed predicate classes.
+
+         \begin{figure}[ht!]
             \centering
-		\includegraphics[width = \linewidth]{analysis.pdf}
+    		\includegraphics[width = .9\linewidth]{analysis.pdf}
     		\caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, non-factive) and operator (\textcolor{op-m-color}{modal},  \textcolor{op-n-color}{negation}, \textcolor{op-q-color}{polar questions}, \textcolor{op-c-color}{conditional antecedents}) with 95\% bootstrapped confidence intervals.
-Violin plots indicate kernel probability density of observations.
+Violin plots indicate kernel probability density of the individual participants' ratings.
+Facets for predicates are ordered by the predicate mean certainty rating (aggregating across operators).
 Below each facet, a line spanning two operators indicates a non-zero difference according to the pairwise comparison of operators.
@@
 
-    Because we found statistically robust non-zero differences by-operator differences for each of the predicates, and these differ beween predicates, the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
-
-    As a result, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator.
+        For instance, the five (semi-)factive predicates exhibit four different patterns:
+        %
+        First, the CC of \emph{be annoyed} projects
+        most from under questions, less from under negation, followed by conditionals, and least from under the modal \emph{perhaps} (\textcolor{op-q-color}{Q} $>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-c-color}{C} $>$ \textcolor{op-m-color}{M}).
+        % more from under negation and questions than conditionals and the modal {\em perhaps};
+        Second, the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under {\em perhaps} (\textcolor{op-q-color}{Q} $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-c-color}{C}$\} >$ \textcolor{op-m-color}{M}).
+        %
+        The CCs of \emph{discover} and \emph{see} exhibit a third pattern: They project most from under questions and conditionals, less from under negation, and least from under \emph{perhaps} ($\{$\textcolor{op-q-color}{Q}$,$ \textcolor{op-c-color}{C}$\}>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-m-color}{M}).
+        %
+        Finally, the CC of \emph{reveal} projects most from conditionals, less from questions, and least from under negation and \emph{perhaps} (\textcolor{op-c-color}{C} $>$ \textcolor{op-q-color}{Q}  $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-m-color}{M}$\}$).
+        %
+        Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
+
+        We also observe by-operator projection variation for non-factives predicates.
+Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}.
+Other non-factives predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.
+
+        These generalizations are confirmed by the statistical analysis.
+We fitted a Bayesian mixed-effects beta regression to the subset of the data containing the observations for each predicate (using \texttt{brms} in R).
+Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect) on (scaled) projection ratings, with random by-item intercepts and operator slopes.
+For each predicate, the differences between operators were established by a pairwise comparison of posterior estimates of marginal means for the four operators (using \texttt{emmeans}).
+We take those differences between posterior samples to be statistically robust where the $95\%$ (HDI) does not include $0$.\footnote{Further details about the model are provided in Supplement \ref{app:analysis-2}.} The results are indicated in Figure~\ref{fig:op-pred-analysis}, where non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
+
+        Because we found statistically robust non-zero differences by-operator differences for each of the predicates, and these differ between predicates, the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
+        %
+        Therefore, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator.
 However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find an effect here.
@@
 
-	Our results differ from those of \citealt{sieker_projective_2022}: While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
+        Our results, however, differ from those of \citealt{sieker_projective_2022}: While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
 As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic.
@@
 
-Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker10,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entail- ment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones.
+        Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entail- ment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones.
 This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates} embedded in polar questions.
@@
 
+\section{Analysis 1 --- Estimating operator effects on projection} \label{app:analysis-1}
+
+    Bayesian mixed effects beta regression for the full dataset.\footnote{The analysis script can be found at \url{https://github.com/judith-tonhauser/CommitmentBankPlus/blob/main/results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R}.}\\
+
+    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
+    % using method used in \cite{degen_are_2022}, from \cite{smithson_better_2006}, for proportional data.
+    $$y' = \frac{y * (n - 1) + 0.5}{n}$$
+
+    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\
+
+    \noindent\textbf{Random effects:}
+
+\section{Analysis 2 --- Estimating operator effects for each predicate} \label{app:analysis-2}
+
+    Bayesian mixed effects beta regression for the each subset of the dataset by predicate.\\
+
+    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression.\\
+
+    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\
+
+    \noindent\textbf{Random effects:} Item random intercepts and slopes by operator; No participant random effects, because each participant saw each predicate only once.\\
+
+    \noindent The models were fit using \texttt{brms}
+    % (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}), 
+    using low-information priors (\texttt{brms} defaults). The estimation used $4$ chains, $4,000$ iterations per chain, $700$ warm-up, total $13,200$ iterations saved for each model.
+
+    \subsection{Model output}
+
+    For each predicate, differences between levels of operator were established by considering the pairwise difference between estimated marginal means for each level of operator, using the \texttt{emmeans} package (\citealt{lenth_emmeans_2024}) in R (R Core Team 2016).
+The below table gives posterior means, as well as upper and lower values for 95\% highest density intervals (HDIs) for the derived variables representing the pairwise differences.
+
+    \begin{longtable}{llrrr}\toprule
+     predicate & contrast & mean & lower & upper \\ \midrule
+     acknowledge & conditional - negation & -0.50 & -0.70 & -0.31 \\ 
+      acknowledge & conditional - question & 0.19 & -0.03 & 0.42 \\ 
+      acknowledge & modal - conditional & -0.21 & -0.36 & -0.05 \\ 
+      acknowledge & modal - negation & -0.71 & -0.90 & -0.50 \\ 
+      acknowledge & modal - question & -0.01 & -0.23 & 0.21 \\ 
+      acknowledge & negation - question & 0.69 & 0.44 & 0.97 \\ \midrule
+
+      admit & conditional - negation & 0.58 & 0.40 & 0.76 \\ 
+      admit & conditional - question & 0.30 & 0.10 & 0.49 \\ 
+      admit & modal - conditional & -0.50 & -0.67 & -0.33 \\ 
+      admit & modal - negation & 0.08 & -0.08 & 0.22 \\ 
+      admit & modal - question & -0.21 & -0.37 & -0.05 \\ 
+      admit & negation - question & -0.28 & -0.46 & -0.12 \\ \midrule
+
+      announce & conditional - negation & 0.42 & 0.25 & 0.60 \\ 
+      announce & conditional - question & 0.32 & 0.14 & 0.48 \\ 
+      announce & modal - conditional & -0.48 & -0.63 & -0.34 \\ 
+      announce & modal - negation & -0.06 & -0.23 & 0.11 \\ 
+      announce & modal - question & -0.17 & -0.33 & -0.01 \\ 
+      announce & negation - question & -0.11 & -0.29 & 0.07 \\ \midrule
+
+      be\_annoyed & conditional - negation & -0.59 & -0.77 & -0.41 \\ 
+      be\_annoyed & conditional - question & -0.90 & -1.08 & -0.70 \\ 
+      be\_annoyed & modal - conditional & -0.33 & -0.48 & -0.16 \\ 
+      be\_annoyed & modal - negation & -0.92 & -1.08 & -0.76 \\ 
+      be\_annoyed & modal - question & -1.23 & -1.40 & -1.05 \\ 
+      be\_annoyed & negation - question & -0.31 & -0.50 & -0.12 \\ \midrule
+
+      be\_right & conditional - negation & -0.22 & -0.39 & -0.06 \\ 
+      be\_right & conditional - question & 0.40 & 0.21 & 0.57 \\ 
+      be\_right & modal - conditional & 0.43 & 0.28 & 0.58 \\ 
+      be\_right & modal - negation & 0.21 & 0.05 & 0.37 \\ 
+      be\_right & modal - question & 0.84 & 0.66 & 1.00 \\ 
+      be\_right & negation - question & 0.62 & 0.43 & 0.81 \\ \midrule
+
+      confess & conditional - negation & 0.65 & 0.48 & 0.82 \\ 
+      confess & conditional - question & 0.45 & 0.29 & 0.61 \\ 
+      confess & modal - conditional & -0.62 & -0.77 & -0.48 \\ 
+      confess & modal - negation & 0.03 & -0.14 & 0.18 \\ 
+      confess & modal - question & -0.17 & -0.32 & -0.02 \\ 
+      confess & negation - question & -0.20 & -0.38 & -0.03 \\ \midrule
+
+      confirm & conditional - negation & 0.77 & 0.59 & 0.95 \\ 
+      confirm & conditional - question & 0.55 & 0.34 & 0.76 \\ 
+      confirm & modal - conditional & 0.12 & -0.04 & 0.28 \\ 
+      confirm & modal - negation & 0.89 & 0.73 & 1.05 \\ 
+      confirm & modal - question & 0.67 & 0.48 & 0.85 \\ 
+      confirm & negation - question & -0.22 & -0.42 & -0.02 \\ \midrule
+
+      demonstrate & conditional - negation & 0.58 & 0.42 & 0.74 \\ 
+      demonstrate & conditional - question & 0.21 & 0.06 & 0.36 \\ 
+      demonstrate & modal - conditional & -0.17 & -0.31 & -0.03 \\ 
+      demonstrate & modal - negation & 0.41 & 0.25 & 0.56 \\ 
+      demonstrate & modal - question & 0.03 & -0.11 & 0.18 \\ 
+      demonstrate & negation - question & -0.37 & -0.53 & -0.21 \\ \midrule
+
+      discover & conditional - negation & 0.46 & 0.27 & 0.65 \\ 
+      discover & conditional - question & -0.15 & -0.33 & 0.03 \\ 
+      discover & modal - conditional & -0.64 & -0.79 & -0.48 \\ 
+      discover & modal - negation & -0.18 & -0.36 & -0.01 \\ 
+      discover & modal - question & -0.79 & -0.95 & -0.61 \\ 
+      discover & negation - question & -0.61 & -0.81 & -0.42 \\ \midrule
+
+      establish & conditional - negation & 0.44 & 0.26 & 0.62 \\ 
+      establish & conditional - question & 0.28 & 0.10 & 0.47 \\ 
+      establish & modal - conditional & -0.06 & -0.24 & 0.11 \\ 
+      establish & modal - negation & 0.37 & 0.22 & 0.52 \\ 
+      establish & modal - question & 0.22 & 0.07 & 0.37 \\ 
+      establish & negation - question & -0.15 & -0.32 & 0.01 \\ \midrule
+
+      hear & conditional - negation & 0.44 & 0.27 & 0.61 \\ 
+      hear & conditional - question & -0.03 & -0.19 & 0.13 \\ 
+      hear & modal - conditional & -0.92 & -1.06 & -0.76 \\ 
+      hear & modal - negation & -0.47 & -0.63 & -0.31 \\ 
+      hear & modal - question & -0.95 & -1.10 & -0.79 \\ 
+      hear & negation - question & -0.47 & -0.65 & -0.31 \\ \midrule
+
+      inform & conditional - negation & 0.12 & -0.07 & 0.30 \\ 
+      inform & conditional - question & -0.33 & -0.53 & -0.13 \\ 
+      inform & modal - conditional & -0.63 & -0.80 & -0.45 \\ 
+      inform & modal - negation & -0.51 & -0.67 & -0.35 \\ 
+      inform & modal - question & -0.96 & -1.12 & -0.79 \\ 
+      inform & negation - question & -0.45 & -0.62 & -0.26 \\ \midrule
+
+      know & conditional - negation & 0.13 & -0.06 & 0.31 \\ 
+      know & conditional - question & -0.56 & -0.79 & -0.35 \\ 
+      know & modal - conditional & -0.70 & -0.86 & -0.55 \\ 
+      know & modal - negation & -0.57 & -0.74 & -0.40 \\ 
+      know & modal - question & -1.26 & -1.49 & -1.04 \\ 
+      know & negation - question & -0.70 & -0.93 & -0.46 \\ \midrule
+
+      pretend & conditional - negation & -1.00 & -1.22 & -0.76 \\ 
+      pretend & conditional - question & -0.09 & -0.33 & 0.12 \\ 
+      pretend & modal - conditional & 0.20 & -0.03 & 0.42 \\ 
+      pretend & modal - negation & -0.80 & -0.97 & -0.64 \\ 
+      pretend & modal - question & 0.11 & -0.06 & 0.27 \\ 
+      pretend & negation - question & 0.91 & 0.73 & 1.08 \\ \midrule
+
+      prove & conditional - negation & 0.52 & 0.35 & 0.69 \\ 
+      prove & conditional - question & 0.38 & 0.20 & 0.54 \\ 
+      prove & modal - conditional & 0.24 & 0.09 & 0.40 \\ 
+      prove & modal - negation & 0.77 & 0.61 & 0.92 \\ 
+      prove & modal - question & 0.62 & 0.46 & 0.78 \\ 
+      prove & negation - question & -0.15 & -0.33 & 0.02 \\ \midrule
+
+      reveal & conditional - negation & 0.63 & 0.46 & 0.79 \\ 
+      reveal & conditional - question & 0.29 & 0.10 & 0.49 \\ 
+      reveal & modal - conditional & -0.60 & -0.77 & -0.45 \\ 
+      reveal & modal - negation & 0.02 & -0.13 & 0.16 \\ 
+      reveal & modal - question & -0.32 & -0.49 & -0.13 \\ 
+      reveal & negation - question & -0.34 & -0.52 & -0.16 \\ \midrule
+
+      say & conditional - negation & 0.30 & 0.04 & 0.54 \\ 
+      say & conditional - question & 1.23 & 1.01 & 1.44 \\ 
+      say & modal - conditional & -0.45 & -0.64 & -0.28 \\ 
+      say & modal - negation & -0.16 & -0.36 & 0.05 \\ 
+      say & modal - question & 0.77 & 0.59 & 0.95 \\ 
+      say & negation - question & 0.93 & 0.73 & 1.14 \\ \midrule
+
+      see & conditional - negation & 0.55 & 0.33 & 0.78 \\ 
+      see & conditional - question & -0.15 & -0.35 & 0.08 \\ 
+      see & modal - conditional & -0.89 & -1.09 & -0.67 \\ 
+      see & modal - negation & -0.33 & -0.50 & -0.17 \\ 
+      see & modal - question & -1.03 & -1.20 & -0.87 \\ 
+      see & negation - question & -0.70 & -0.89 & -0.52 \\ \midrule
+
+      suggest & conditional - negation & 0.25 & 0.08 & 0.42 \\ 
+      suggest & conditional - question & 0.61 & 0.44 & 0.77 \\ 
+      suggest & modal - conditional & -0.27 & -0.42 & -0.11 \\ 
+      suggest & modal - negation & -0.02 & -0.19 & 0.14 \\ 
+      suggest & modal - question & 0.34 & 0.18 & 0.49 \\ 
+      suggest & negation - question & 0.36 & 0.19 & 0.52 \\ \midrule
+
+      think & conditional - negation & -0.50 & -0.70 & -0.30 \\ 
+      think & conditional - question & 0.19 & -0.04 & 0.41 \\ 
+      think & modal - conditional & -0.21 & -0.36 & -0.05 \\ 
+      think & modal - negation & -0.71 & -0.91 & -0.52 \\ 
+      think & modal - question & -0.01 & -0.24 & 0.22 \\ 
+      think & negation - question & 0.70 & 0.44 & 0.97 \\ 
+      \end{longtable}
+
 \end{document}

### 2024-02-14 writing/SuB-paper/supplement.tex
tracking udpates from overleaf wed afternoon

--- 2024-02-14 writing/SuB-paper/supplement.tex
@@
 
+\section{Analysis 1 --- Estimating operator effects on projection}
+
+    Bayesian mixed effects beta regression for the full dataset.\footnote{The analysis script can be found at \url{https://github.com/judith-tonhauser/CommitmentBankPlus/blob/main/results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R}.}\\
+
+    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
+    % using method used in \cite{degen_are_2022}, from \cite{smithson_better_2006}, for proportional data.
+    $$y' = \frac{y * (n - 1) + 0.5}{n}$$
+
+    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\
+
+    \noindent\textbf{Random effects:}
+
+\section{Analysis 2 --- Estimating operator effects for each predicate}
+
+    Bayesian mixed effects beta regression for the each subset of the dataset by predicate.\\
+
+    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression.\\
+
+    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\
+
+    \noindent\textbf{Random effects:} Item random intercepts and slopes by operator; No participant random effects, because each participant saw each predicate only once.\\
+
+    \noindent The models were fit using \texttt{brms}
+    % (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}), 
+    using low-information priors (\texttt{brms} defaults). The estimation used $4$ chains, $4,000$ iterations per chain, $700$ warm-up, total $13,200$ iterations saved for each model.
+
+    \subsection{Model output}
+
+    For each predicate, differences between levels of operator were established by considering the pairwise difference between estimated marginal means for each level of operator, using the ‘emmeans’ package (Lenth 2023) in R (R Core Team 2016).
+The below table gives posterior means, as well as upper and lower values for 95\% highest density intervals (HDIs) for the derived variables representing the pairwise differences.
+
+    \begin{longtable}{llrrr}\toprule
+     predicate & contrast & mean & lower & upper \\ \midrule
+     acknowledge & conditional - negation & -0.50 & -0.70 & -0.31 \\ 
+      acknowledge & conditional - question & 0.19 & -0.03 & 0.42 \\ 
+      acknowledge & modal - conditional & -0.21 & -0.36 & -0.05 \\ 
+      acknowledge & modal - negation & -0.71 & -0.90 & -0.50 \\ 
+      acknowledge & modal - question & -0.01 & -0.23 & 0.21 \\ 
+      acknowledge & negation - question & 0.69 & 0.44 & 0.97 \\ \midrule
+      %
+      admit & conditional - negation & 0.58 & 0.40 & 0.76 \\ 
+      admit & conditional - question & 0.30 & 0.10 & 0.49 \\ 
+      admit & modal - conditional & -0.50 & -0.67 & -0.33 \\ 
+      admit & modal - negation & 0.08 & -0.08 & 0.22 \\ 
+      admit & modal - question & -0.21 & -0.37 & -0.05 \\ 
+      admit & negation - question & -0.28 & -0.46 & -0.12 \\ \midrule
+      %
+      announce & conditional - negation & 0.42 & 0.25 & 0.60 \\ 
+      announce & conditional - question & 0.32 & 0.14 & 0.48 \\ 
+      announce & modal - conditional & -0.48 & -0.63 & -0.34 \\ 
+      announce & modal - negation & -0.06 & -0.23 & 0.11 \\ 
+      announce & modal - question & -0.17 & -0.33 & -0.01 \\ 
+      announce & negation - question & -0.11 & -0.29 & 0.07 \\ \midrule
+      %
+      be\_annoyed & conditional - negation & -0.59 & -0.77 & -0.41 \\ 
+      be\_annoyed & conditional - question & -0.90 & -1.08 & -0.70 \\ 
+      be\_annoyed & modal - conditional & -0.33 & -0.48 & -0.16 \\ 
+      be\_annoyed & modal - negation & -0.92 & -1.08 & -0.76 \\ 
+      be\_annoyed & modal - question & -1.23 & -1.40 & -1.05 \\ 
+      be\_annoyed & negation - question & -0.31 & -0.50 & -0.12 \\ \midrule
+      %
+      be\_right & conditional - negation & -0.22 & -0.39 & -0.06 \\ 
+      be\_right & conditional - question & 0.40 & 0.21 & 0.57 \\ 
+      be\_right & modal - conditional & 0.43 & 0.28 & 0.58 \\ 
+      be\_right & modal - negation & 0.21 & 0.05 & 0.37 \\ 
+      be\_right & modal - question & 0.84 & 0.66 & 1.00 \\ 
+      be\_right & negation - question & 0.62 & 0.43 & 0.81 \\ \midrule
+      %
+      confess & conditional - negation & 0.65 & 0.48 & 0.82 \\ 
+      confess & conditional - question & 0.45 & 0.29 & 0.61 \\ 
+      confess & modal - conditional & -0.62 & -0.77 & -0.48 \\ 
+      confess & modal - negation & 0.03 & -0.14 & 0.18 \\ 
+      confess & modal - question & -0.17 & -0.32 & -0.02 \\ 
+      confess & negation - question & -0.20 & -0.38 & -0.03 \\ \midrule
+      %
+      confirm & conditional - negation & 0.77 & 0.59 & 0.95 \\ 
+      confirm & conditional - question & 0.55 & 0.34 & 0.76 \\ 
+      confirm & modal - conditional & 0.12 & -0.04 & 0.28 \\ 
+      confirm & modal - negation & 0.89 & 0.73 & 1.05 \\ 
+      confirm & modal - question & 0.67 & 0.48 & 0.85 \\ 
+      confirm & negation - question & -0.22 & -0.42 & -0.02 \\ \midrule
+
+      demonstrate & conditional - negation & 0.58 & 0.42 & 0.74 \\ 
+      demonstrate & conditional - question & 0.21 & 0.06 & 0.36 \\ 
+      demonstrate & modal - conditional & -0.17 & -0.31 & -0.03 \\ 
+      demonstrate & modal - negation & 0.41 & 0.25 & 0.56 \\ 
+      demonstrate & modal - question & 0.03 & -0.11 & 0.18 \\ 
+      demonstrate & negation - question & -0.37 & -0.53 & -0.21 \\ \midrule
+
+      discover & conditional - negation & 0.46 & 0.27 & 0.65 \\ 
+      discover & conditional - question & -0.15 & -0.33 & 0.03 \\ 
+      discover & modal - conditional & -0.64 & -0.79 & -0.48 \\ 
+      discover & modal - negation & -0.18 & -0.36 & -0.01 \\ 
+      discover & modal - question & -0.79 & -0.95 & -0.61 \\ 
+      discover & negation - question & -0.61 & -0.81 & -0.42 \\ \midrule
+
+      establish & conditional - negation & 0.44 & 0.26 & 0.62 \\ 
+      establish & conditional - question & 0.28 & 0.10 & 0.47 \\ 
+      establish & modal - conditional & -0.06 & -0.24 & 0.11 \\ 
+      establish & modal - negation & 0.37 & 0.22 & 0.52 \\ 
+      establish & modal - question & 0.22 & 0.07 & 0.37 \\ 
+      establish & negation - question & -0.15 & -0.32 & 0.01 \\ \midrule
+
+      hear & conditional - negation & 0.44 & 0.27 & 0.61 \\ 
+      hear & conditional - question & -0.03 & -0.19 & 0.13 \\ 
+      hear & modal - conditional & -0.92 & -1.06 & -0.76 \\ 
+      hear & modal - negation & -0.47 & -0.63 & -0.31 \\ 
+      hear & modal - question & -0.95 & -1.10 & -0.79 \\ 
+      hear & negation - question & -0.47 & -0.65 & -0.31 \\ \midrule
+
+      inform & conditional - negation & 0.12 & -0.07 & 0.30 \\ 
+      inform & conditional - question & -0.33 & -0.53 & -0.13 \\ 
+      inform & modal - conditional & -0.63 & -0.80 & -0.45 \\ 
+      inform & modal - negation & -0.51 & -0.67 & -0.35 \\ 
+      inform & modal - question & -0.96 & -1.12 & -0.79 \\ 
+      inform & negation - question & -0.45 & -0.62 & -0.26 \\ \midrule
+
+      know & conditional - negation & 0.13 & -0.06 & 0.31 \\ 
+      know & conditional - question & -0.56 & -0.79 & -0.35 \\ 
+      know & modal - conditional & -0.70 & -0.86 & -0.55 \\ 
+      know & modal - negation & -0.57 & -0.74 & -0.40 \\ 
+      know & modal - question & -1.26 & -1.49 & -1.04 \\ 
+      know & negation - question & -0.70 & -0.93 & -0.46 \\ \midrule
+
+      pretend & conditional - negation & -1.00 & -1.22 & -0.76 \\ 
+      pretend & conditional - question & -0.09 & -0.33 & 0.12 \\ 
+      pretend & modal - conditional & 0.20 & -0.03 & 0.42 \\ 
+      pretend & modal - negation & -0.80 & -0.97 & -0.64 \\ 
+      pretend & modal - question & 0.11 & -0.06 & 0.27 \\ 
+      pretend & negation - question & 0.91 & 0.73 & 1.08 \\ \midrule
+
+      prove & conditional - negation & 0.52 & 0.35 & 0.69 \\ 
+      prove & conditional - question & 0.38 & 0.20 & 0.54 \\ 
+      prove & modal - conditional & 0.24 & 0.09 & 0.40 \\ 
+      prove & modal - negation & 0.77 & 0.61 & 0.92 \\ 
+      prove & modal - question & 0.62 & 0.46 & 0.78 \\ 
+      prove & negation - question & -0.15 & -0.33 & 0.02 \\ \midrule
+
+      reveal & conditional - negation & 0.63 & 0.46 & 0.79 \\ 
+      reveal & conditional - question & 0.29 & 0.10 & 0.49 \\ 
+      reveal & modal - conditional & -0.60 & -0.77 & -0.45 \\ 
+      reveal & modal - negation & 0.02 & -0.13 & 0.16 \\ 
+      reveal & modal - question & -0.32 & -0.49 & -0.13 \\ 
+      reveal & negation - question & -0.34 & -0.52 & -0.16 \\ \midrule
+
+      say & conditional - negation & 0.30 & 0.04 & 0.54 \\ 
+      say & conditional - question & 1.23 & 1.01 & 1.44 \\ 
+      say & modal - conditional & -0.45 & -0.64 & -0.28 \\ 
+      say & modal - negation & -0.16 & -0.36 & 0.05 \\ 
+      say & modal - question & 0.77 & 0.59 & 0.95 \\ 
+      say & negation - question & 0.93 & 0.73 & 1.14 \\ \midrule
+
+      see & conditional - negation & 0.55 & 0.33 & 0.78 \\ 
+      see & conditional - question & -0.15 & -0.35 & 0.08 \\ 
+      see & modal - conditional & -0.89 & -1.09 & -0.67 \\ 
+      see & modal - negation & -0.33 & -0.50 & -0.17 \\ 
+      see & modal - question & -1.03 & -1.20 & -0.87 \\ 
+      see & negation - question & -0.70 & -0.89 & -0.52 \\ \midrule
+
+      suggest & conditional - negation & 0.25 & 0.08 & 0.42 \\ 
+      suggest & conditional - question & 0.61 & 0.44 & 0.77 \\ 
+      suggest & modal - conditional & -0.27 & -0.42 & -0.11 \\ 
+      suggest & modal - negation & -0.02 & -0.19 & 0.14 \\ 
+      suggest & modal - question & 0.34 & 0.18 & 0.49 \\ 
+      suggest & negation - question & 0.36 & 0.19 & 0.52 \\ \midrule
+
+      think & conditional - negation & -0.50 & -0.70 & -0.30 \\ 
+      think & conditional - question & 0.19 & -0.04 & 0.41 \\ 
+      think & modal - conditional & -0.21 & -0.36 & -0.05 \\ 
+      think & modal - negation & -0.71 & -0.91 & -0.52 \\ 
+      think & modal - question & -0.01 & -0.24 & 0.22 \\ 
+      think & negation - question & 0.70 & 0.44 & 0.97 \\ 
+      \end{longtable}
+
+\bibliographystyle{sub-chicago}
+\bibliography{projection.bib}
 \end{document}

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/1_summary-graphs.R

@@ -40,7 +40,7 @@ operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
 operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), 
                               y = projection, label = round(projection, digits = 2))) +
   geom_violin(data = data, scale="width", fill = graphcolor, 
-              alpha = .4, color = graphcolor) +
+              alpha = .4) +
   geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "black") +
   geom_point(size=0.5, color = "black") +
   geom_text(hjust = 0, nudge_x = -0.08, nudge_y = -0.06) + 

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/3_analysis-predicates.R

@@ -5,6 +5,7 @@
 library(tidyverse)
 this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
 setwd(this.dir)
+library(xtable)
 
 # analysis
 library(rstan)
@@ -85,7 +86,7 @@ for (current_pred in predicates) {
 # print fixed effects for each model
 for (current_pred in predicates) {
   print(current_pred)
-  print(fixef(eval(parse(text = paste("m.b.", current_pred, sep="")))))
+  print(kable(fixef(eval(parse(text = paste("m.b.", current_pred, sep="")))), "latex", booktabs = TRUE))
 }
 
 # 3 GET PAIRWISE DIFFERENCES ---------------------------------------------------
@@ -142,7 +143,7 @@ levels(predicate_operator_means$predicate)
 pred_order <- levels(predicate_operator_means$predicate)
 textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), pinkk, "white")
 
-data %>% mutate(operator = fct_reorder(operator, projection, .fun = "mean"))
+data <- data %>% mutate(operator = fct_reorder(operator, projection, .fun = "mean"))
 levels(data$operator)
 
 
@@ -163,7 +164,23 @@ for (p in predicates) {
 }
 View(contrasts)
 
-# show in table w colors
+
+# Get table with all contrasts for supplement
+print(xtable(contrasts),
+                  #only.contents = T,
+                  include.rownames=FALSE,
+                  include.colnames=TRUE,
+                  tabular.environment="longtable",
+                  floating=FALSE,
+                  hline.after = NULL,
+                  latex.environments=NULL,
+                  booktabs=TRUE,
+                  sanitize.text.function = function(x){x},
+                  comment = F
+)
+
+
+# show in figure
 # linetypes
 # 0 blank
 # 1 solid

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R
updated by-operator analysis

@@ -1,21 +1,21 @@
-##### ANALYSIS BY OPERATORS #####
+# SuB paper
+# by-operator analysis (section 2.2.1)
 
-# 1 load libraries, data, etc --------------------------------------------------
-
+# load libraries
 library(tidyverse)
-this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(this.dir)
-
-# analysis
-library(rstan)
-options(mc.cores=parallel::detectCores())
-rstan_options(auto_write=TRUE)
+#library(rstan)
+#options(mc.cores=parallel::detectCores())
+#rstan_options(auto_write=TRUE)
 library(brms)
 library(tidybayes)
 library(emmeans)
+library(xtable)
 
+# set working directory
+this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(this.dir)
 
-# LOAD DATA
+# load data 
 data <- read.csv("../../data_combined.csv", header = TRUE, sep = ",")
 data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), 
                         operator = as.factor(op), participant = as.factor(workerid), 
@@ -23,38 +23,110 @@ data <- data %>% mutate(projection = as.numeric(projective), predicate = as.fact
 levels(data$operator) <- c("conditional", "modal", "negation", "question")
 str(data)
 
+# Bayesian beta regression ----
 
-# 2 ANALYSIS -------------------------------------------------------------------
-# BAYESIAN MIXED EFFECTS BETA REGRESSION
+# set the reference level for operator
+data = data %>%
+  mutate(operator = fct_relevel(operator, "modal"))
+levels(data$operator)
 
-# Response: certainty ratings (rescale for beta regression)
+# rescale certainty ratings for beta regression
 # Scaling to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
-# using method used in Degen & Tonhauser (2022), from Smithson & Verkuilen (2006), for proportional data.
+# using method used in Smithson & Verkuilen (2006)
 # y' = (y * (n - 1) + 0.5) / n
 data$betaresponse <- (data$projection*(nrow(data)-1) + .5)/nrow(data)
-data$betaresponse <- as.numeric(data$betaresponse)
 summary(data$betaresponse)
 
 # beta regression w response: scaled projection ratings, fixed effect: operator
 # random effects: random intercepts for item, participant, slopes for both by operator
[…]

### 2024-02-14 writing/SuB-paper/projection-operators-SuB.tex
updated figures, tracking changes to manuscript

@@
 
-		\subsubsection{Materials}
-
-            The target sentences consisted of the 400 combinations of the 20 clause-embedding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}).
+		% \subsubsection{Materials}
+        \paragraph{Materials.}
+            The target sentences consisted of the 400 combinations of the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}).
 As mentioned above, the predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal \emph{perhaps} in Exps.~3, and in conditional antecedents in Exps.~4, for a total of 400 target stimuli in each of the four sets of experiments.
@@
 
-        \subsubsection{Procedure}
+        %\subsubsection{Procedure}
+        \paragraph{Procedure.}
             Participants were asked to imagine that they are at a party and that, when walking into the kitchen, they overhear somebody say something to somebody else.
@@
 On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of the uncertainty about possible interpretations.
-%, which suggests the latter kind of interpretation).
 In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation (for discussion, see also \citealt{grove_factivity_2023}).
@@
 
-        \subsubsection{Data exclusion}
-            Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}.
+        %\subsubsection{Data exclusion}
+        \noindent\textbf{Data exclusion}
+            was done based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.
+            % Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}.
 The data from 2,682 participants entered into the analysis.
@@
 				\includegraphics[scale = .8]{certainty-operator.pdf}
-				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals. Violin plots indicate the kernel probability density of the ratings.
+				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals and violin plots indicate the kernel probability density of the ratings.
                 % of the individual participants’ ratings.
@@
 
-   These observations were confirmed by a posthoc pairwise comparison of the estimated means for each entailment-canceling operator using the `emmeans' package (\citealt{emmeans}) in R (\citealt{r}).
-The input to the pairwise comparison was a  Bayesian mixed-effects beta regression model with weakly informative priors that were fit using the \emph{brms} package (\citealt{buerkner2017}).
-The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson-verkuilen2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level and included a random by-predicate intercept.
-The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators.
-We assume that two operators differ in certainty ratings if their HDI does not include 0.\footnote{A more detailed report of the model output is provided in Supplement \ref{app:analysis-1}}
-
-Table \ref{t:pairwise} provides the output of the pairwise comparisons. As shown,
-
-, the pairwise comparison suggests differences between the naturalness ratings of the CCs of the five factive predicates: The ratings of {\em be annoyed} are lower than those of {\em know}, which in turn are lower than those of {\em see} and {\em discover}, which in turn are lower than those of {\em reveal}.\footnote{While the naturalness ratings collected from the participants range from 0 to 1, the greater than $|$1$|$  estimated marginal mean differences are the result of the beta regression using a logit link function for the mean parameter.} There is also variation between the factive predicates and the presuppositional controls: While all five factive predicates are more natural in explicit ignorance contexts than {\em continue, too} and {\em also}, {\em know} is additionally more natural than {\em again} and {\em stop}, and {\em see, discover} and {\em reveal} are additionally more natural than the {\em it-}cleft.
-Finally, the naturalness ratings in explicit ignorance contexts for {\em see} and {\em discover} do not differ from those of the nonfactives {\em confess, inform}, and {\em acknowledge}, and those of {\em reveal} do not differ from those of the nonfactives {\em admit, hear}, and {\em prove}.
-These results suggest that naturalness ratings in explicit ignorance contexts do not provide empirical support for categorical projection or for a categorical distinction between factive and nonfactive predicates.
+        These generalizations are supported by a post-hoc pairwise comparison of the estimated means for each entailment-canceling operator using the \texttt{emmeans} package (\citealt{emmeans}) in R (\citealt{r}).
+The input to the pairwise comparison was a  Bayesian mixed-effects beta regression model that was fit using the \texttt{brms} package (\citealt{buerkner2017}) with weakly informative priors.
+The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson_better_2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level) and included a random by-predicate intercept.\footnote{See Supplement \ref{app:analysis-1} for more details on the model.} The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators.
+We assume that two operators differ in certainty ratings if the HDI of their pairwise comparison does not include 0.
+
+        Table \ref{t:pairwise} provides the output of the pairwise comparison on a logit scale.
+        %\footnote{While the certainty ratings collected from the participants range from 0 to 1, the smaller than 0 estimated marginal mean differences are the result of the beta regression using a logit link function for the mean parameter.}
+        As shown, the analysis suggests differences between each pair of operators, though the estimated marginal mean differences differ across the pairs.
+That is, certainty ratings are higher for CCs embedded in conditional antecedents than those embedded under polar questions; certainty ratings for CCs embedded under polar questions are higher than those for CCs embedded under negation; and finally certainty ratings for CCs embedded under negation are higher than those for CCs embedded under the epistemic modal \emph{perhaps}.
 
@@
         \end{tabular}
-\caption{Pairwise differences between entailment-canceling operators.}\label{t:pairwise}
+        \caption{Output of the pairwise comparison of entailment-canceling operators.
+The `contrast' column identifies the pair of entailment-canceling operators, the `estimate' column the estimated marginal mean difference, and the `lower 95\% CI' and `upper 95\% CI' columns provide the lower and upper bounds of the HDIs.}\label{t:pairwise}
         \end{table}
 
-		Like \citet{sieker_projective_2022} for German, we observe a small main effect of entailment-cancelling operator on projection.
-However, in contrast to their result that projection is strongest from under negation, our results suggest that projection is strongest from antecedents of conditionals.
-        % \mcdm{does it make sense to compare different predicates?}
-        We note, however, that the differences in mean certainty ratings between the four entailment-canceling operators are very small.
-This suggests that, when abstracting away from individual projective contents, projection from under the four entailment-canceling operators is very similar.
-In other words, when abstracting away from individual projective contents, the family of sentences really are a family.
+ These results suggest that certainty ratings for the CCs of of the English clause-embedding predicate we investigated vary by entailment-canceling operator.
+In contrast to \citet{sieker_projective_2022} for German, the results of our experiments do not suggest that projection is strongest from under negation.
+Recall, however, that they only investigated projection of the CCs of (semi-)factive predicates.
+Since there is by-predicate variation in the effect of entailment-canceling operator on projection (as we show in the next section), this difference between the results of their experiment and ours might very well be due to the types of predicates investigated.
+Finally, the differences in mean certainty ratings between the four entailment-canceling operators are very small.
+This suggests that, when abstracting away from individual predicates and contents, projection from under the four entailment-canceling operators is very similar.
+In other words, when abstracting away from individual  contents, the family of sentences really are a family.
 
@@
 
-        Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the five (semi-)factive predicates, and the 15 non-factive predicates, where predicates are ordered by their overall mean certainty ratings.
-        %
-        As shown, there is by-operator variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
+        Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the 20  predicates, with predicates ordered by their overall mean certainty rating.
+        %
+        As shown, there is by-operator projection variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
 Further, the effect of operator differs between predicates.
@@
 
-        These generalizations are confirmed by the statistical analysis.
-We fitted a Bayesian mixed-effects beta regression to the subset of the data containing the observations for each predicate (using \texttt{brms} in R).
-Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect) on (scaled) projection ratings, with random by-item intercepts and operator slopes.
-For each predicate, the differences between operators were established by a pairwise comparison of posterior estimates of marginal means for the four operators (using \texttt{emmeans}).
-We take those differences between posterior samples to be statistically robust where the $95\%$ (HDI) does not include $0$.\footnote{Further details about the model are provided in Supplement \ref{app:analysis-2}.} The results are indicated in Figure~\ref{fig:op-pred-analysis}, where non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
-
-        Because we found statistically robust non-zero differences by-operator differences for each of the predicates, and these differ between predicates, the analysis also supports the more general claims that there is by-operator variation for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
+        These generalizations are supported by the statistical analysis.
+We fitted Bayesian mixed-effects beta regression models for each the by-predicate subset of the data containing (using \emph{brms}, \citealt{burkner_brms_2017}; in R, \citealt{r}).
+Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect, treatment-coded with reference-level \lq modal\rq) on (scaled) certainty ratings, with by-item random intercepts and random operator slopes.\footnote{Further details about the models, their outputs, and the analysis are provided in Supplement \ref{app:analysis-2}.} For each predicate, the differences between operators were established by a pairwise comparison of posterior estimates of marginal means for the four operators (using \emph{emmeans}, \citealt{lenth_emmeans_2024}).
+Here we report those differences between posterior samples where the $95\%$ HDI does not include $0$.
+These are indicated in Figure~\ref{fig:op-pred-analysis}, where non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
+
+        More generally, we found by-operator differences for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
         %
         Therefore, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator.
-However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find an effect here.
+However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find a difference here.
 We hypothesize that this difference is due to the difference in projection diagnostic used.
 
-        Our results, however, differ from those of \citealt{sieker_projective_2022}: While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
+        Our results, however, differ from those of \citealt{sieker_projective_2022}.
+While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
 As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic.
@@
 
-        Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} As presuppositions are assumed to project from under entail- ment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones.
-This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-embedding predicates in \ref{ex:predicates} embedded in polar questions.
-Contrary to expectation,  \poscite{degen_are_2022} Exps.~1 found that the CCs of the five factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives predicates were as projective, or even more projective, than the CCs of some factive predicates.
+        Lexical approaches to projection assume that factive predicates are ones that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} Because presuppositions are assumed to project from under entailment-canceling operators, this definition predicts that factive predicates are distinguished from non-factives by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives.
+
+        % Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} Because presuppositions are assumed to project from under entailment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones.
+
+        This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} embedded in polar questions.
+Contrary to expectation, \posscite{degen_are_2022} Exps.~1 found that the CCs of the five (semi-)factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives were as projective, or even more projective, than those of some factive predicates.
 In short, projection of the CC from under polar questions did not categorically distinguish factive from non-factives predicates.
-Further support for this result came from the analysis of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{demarneffe-etal-sub23}), the VerbVeridicality dataset (\citealt{ross-pavlick2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).
+Further support for this result came from the analysis of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{de_marneffe_commitmentbank_2019}), the VerbVeridicality dataset (\citealt{ross_how_2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).
 
@@
             \begin{subfigure}[t]{0.57\textwidth}
-\includegraphics[width = 1\linewidth]{question-predicate-graph-1.pdf}
-\caption{Polar question}\label{fig:op-pred-ratings-q}
+            \caption{Polar question}
+            \includegraphics[width = 1\linewidth]{question-predicate-graph.pdf}
+            \label{fig:op-pred-ratings-q}
             \end{subfigure}
 
-\medskip
+            \vspace{-\baselineskip}
+            %\medskip
             \begin{subfigure}[t]{0.57\textwidth}
-\includegraphics[width = 1\linewidth]{modal-predicate-graph-1.pdf}
-\caption{Modal \emph{perhaps}}\label{fig:op-pred-ratings-m}
+            \caption{Modal \emph{perhaps}}
+            \includegraphics[width = 1\linewidth]{modal-predicate-graph.pdf}
+            \label{fig:op-pred-ratings-m}
             \end{subfigure}
 
-\medskip
+            \vspace{-\baselineskip}
+            %\medskip
             \begin{subfigure}[t]{0.57\textwidth}
-\includegraphics[width = 1\linewidth]{negation-predicate-graph-1.pdf}
-\caption{Negation}\label{fig:op-pred-ratings-n}
+            \caption{Negation}
+            \includegraphics[width = 1\linewidth]{negation-predicate-graph.pdf}
+            \label{fig:op-pred-ratings-n}
             \end{subfigure}
 
-\medskip
-
+            \vspace{-\baselineskip}
+            %\medskip
             \begin{subfigure}[t]{0.57\textwidth}
- \includegraphics[width = 1\linewidth]{conditional-predicate-graph-1.pdf}
-\caption{Conditional}\label{fig:op-pred-ratings-c}
+            \caption{Conditional}
+            \includegraphics[width = 1\linewidth]{conditional-predicate-graph.pdf}
+            \label{fig:op-pred-ratings-c}
             \end{subfigure}
-\caption{Mean certainty ratings by predicate (with factives in violet) under (a) polar question, (b) modal \emph{perhaps}, (c) negation, (d) conditional.
+
+            \vspace{-\baselineskip}
+            \caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, factive) for (a) polar questions, (b) the modal \emph{perhaps}, (c) negation, (d) conditional antecedents.
 Error bars indicate $95\%$ bootstrapped confidence intervals.
-Violin plots indicate kernel probability densities of individual ratings.}\label{fig:op-pred-ratings2}
+Violin plots indicate kernel probability densities of observations.}\label{fig:op-pred-ratings2}
         \end{figure}
 
-The results of the experiments reported on in this paper replicate this result.
+        The experiments reported on in this paper replicate this result.
+As shown in Fig.~\ref{fig:op-pred-ratings-q}, there is variation between the CCs of the five factive predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from non-factive predicates.
+Furthermore, our results suggest that this result can be extended to the three other entailment-canceling operators as well.
+As shown in Figs.~\ref{fig:op-pred-ratings-m}-\ref{fig:op-pred-ratings-c}, there is variation in the projection of the CCs of the five factive predicates from under \emph{perhaps}, negation, and conditional antecedents, and projection ratings in these conditions do not show a categorical difference between factive and non-factive predicates here, either.
+Instead, we see a gradient  effect of predicate on projection ratings in all four conditions.
+These results lend further support to the conclusion of \citealt{degen_are_2022} that there is, to date, no empirical evidence for a coherent class of factive predicates.
+
+        % The results of the experiments reported on in this paper replicate this result.
 As shown in Fig.~\ref{fig:op-pred-ratings-q}, there is variation between the CCs of the five factive predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from non-factives predicates.
@@
 
-%			Taking into account the effect of \texttt{predicate} on projection, we find gradient projection variation, like \citet{tonhauser_how_2018} and \citet{degen_are_2022}.
-			%
-			%This is illustrated in Figure~\ref{fig:op-pred-ratings}, which shows mean projection ratings for the 20 predicates grouped by embedding operator.
-Here, predicates on the x-axis are ordered by their mean rating across all operators (\emph{be annoyed} has the highest overall mean).
-
-			%Our results therefore provide further support from negation, modals, and conditionals for the result of \citet{degen_are_2022}, that projection does not categorically differentiate between (semi-)factive and non-factive predicates, while also replicating it for polar questions.
-The CCs of non-factive \emph{inform} and \emph{acknowledge}, for instance, are at least as projective as that of some (semi-)factive predicates.
-
 	\subsubsection{Summary}
@@
         The results of our experiments suggest that there is little by-operator variation when aggregating over clause-embedding predicates, but that the CCs of different clause embedding predicates exhibit by-operator projection variation.
-			%Based on our data, we conclude that embedded propositional content projects differently from under different entailment-cancelling operators. 
-    Crucially, the observed by-operator variation does not align with prior claims about differences between factive and semi-factive predicates in by-operator variation (e.g., \citealt{karttunen_observations_1971}).
+        %
+        Crucially, the effect of operator on projection differs by predicate, but not in ways that align with prior claims about differences between factive and semi-factive predicates (e.g., \citealt{karttunen_observations_1971}).
 Finally, the results of our experiments provide further support for the results of \citealt{degen_are_2022}, who did not find empirical support for a class of factive predicates based on the projection of the CC from under polar questions.
@@
 
-%   does not categorically distinguish between lexical classes (e.g., factives, semi-factives, veridical predicates), so future research appealing to these categories must clarify their definition.
-			%
-			%Finally, the effect of operator on projection differs by predicate.
-The theoretical implications of our findings are discussed in the following section.
-Importantly, accounts of projective content need to take into account predicate/operator interactions, and the gradient by-predicate variability.
-
 \section{Converging evidence for the by-operator variation}\label{s:converging}
 
-This section provides converging evidence for the by-operator variation based on the MegaVeridicality dataset (\citealt{white_role_2018}).
+    This section provides converging evidence for the by-operator variation based on the Mega\-Ve\-ri\-dicality dataset (\citealt{white_role_2018}).
 This dataset contains projection ratings for the CCs of 517 English clause-embedding predicates.
@@
 \section{Analysis 1 --- Estimating operator effects on projection} \label{app:analysis-1}
-
-    Bayesian mixed effects beta regression for the full dataset.\footnote{The analysis script can be found at \url{https://github.com/judith-tonhauser/CommitmentBankPlus/blob/main/results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R}.}\\
-
-    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,
-    % using method used in \cite{degen_are_2022}, from \cite{smithson_better_2006}, for proportional data.
+    Models: Bayesian mixed effects beta regression, with:
+    \begin{itemize}
+        \item \textbf{Repsonse variable:} projection ratings scaled for beta-regression to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$,  using method used in \cite{degen_are_2022}, from \cite{smithson_better_2006}, for proportional data.
             $$y' = \frac{y * (n - 1) + 0.5}{n}$$
 
-    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\
-
-    \noindent\textbf{Random effects:}
+        \item \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)
+
+        \item \textbf{Random effects:} Predicate random intercepts
+
+        \item Assuming a beta-distributed response variable, these effects were estimated for the mean ($\mu$) and precision ($\phi$) parameters of the beta-distribution.
+
+    \end{itemize}
+
+    \noindent The model was fit using \texttt{brms}
+    % (\citealt{burkner_brms_2017}) in R (\citealt{r_core_team_r_2014}), 
+    using low-information priors (\texttt{brms} defaults).
+The estimation used $4$ chains, $3,000$ iterations per chain, $500$ warm-up, total $10,000$ iterations saved.
+An excerpt from the model output is shown below, providing parameter estimates, and $95\%$ credible intervals (CIs) for the fixed effects.
+
+    \input{analysis1.model}
+
+    \footnote{The analysis script can be found at \url{https://github.com/judith-tonhauser/CommitmentBankPlus/blob/main/results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R}.}
 
@@
 
-    Bayesian mixed effects beta regression for the each subset of the dataset by predicate.\\
-
-    \noindent\textbf{Repsonse variable:} projection ratings scaled for beta-regression.\\
-
-    \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)\\
-
-    \noindent\textbf{Random effects:} Item random intercepts and slopes by operator; No participant random effects, because each participant saw each predicate only once.\\
+    Models: 20 Bayesian mixed effects beta regression models for the each subset of the dataset by predicate, with:
+    \begin{itemize}
+        \item \textbf{Repsonse variable:} projection ratings scaled for beta-regression.
+
+        \item \noindent\textbf{Fixed effect:} operator (modal, negation, question, conditional; with reference level: modal)
+
+        \item \textbf{Random effects:} Item random intercepts and slopes by operator % No participant random effects, because each participant saw each predicate only once.
+
+        \item These effects were estimated for the mean ($\mu$) and precision ($\phi$) parameters of the beta-distribution.
+
+    \end{itemize}
 
@@
     using low-information priors (\texttt{brms} defaults). The estimation used $4$ chains, $4,000$ iterations per chain, $700$ warm-up, total $13,200$ iterations saved for each model.
-
-    \subsection{Model output}
 

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/1_summary-graphs.R

@@ -13,7 +13,7 @@ cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00"
                 "#CC79A7", "#000000")
 graphcolor <- "#0072B2"
 pinkk <- "#CC79A7"
-pred_colors <- c(pinkk, "black")
+pred_colors <- c(pinkk, "gray")
 theme_set(theme_bw())
 
 # LOAD DATA
@@ -39,10 +39,9 @@ operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
 # plot
 operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), 
                               y = projection, label = round(projection, digits = 2))) +
-  geom_violin(data = data, scale="width", fill = graphcolor, 
-              alpha = .4) +
+  geom_violin(data = data, scale="width", color = "gray", fill = "gray", alpha = .4) +
   geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "black") +
-  geom_point(size=0.5, color = "black") +
+  geom_point(size=1, color = "black", shape = 21) +
   geom_text(hjust = 0, nudge_x = -0.08, nudge_y = -0.06) + 
   scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                      labels = c("0","0.2","0.4","0.6","0.8","1"),
@@ -50,7 +49,7 @@ operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean
   theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
   xlab("Operator")
 
-ggsave("../graphs/certainty-operator.pdf", width=3.3, height=5)
+ggsave("../graphs/certainty-operator.pdf", width=3, height=4)
 
 
 # 3 projection by predicate (all predicates, this is not in the paper) ---------
@@ -95,7 +94,7 @@ predicate_means %>% ggplot(aes(x = fct_reorder(predicate, projection, .fun = "me
 ggsave(f="../graphs/predicate-graph-1.pdf", width=7, height=4.5)
 
 
-# 4 projection by predicate and operator (Figure 3) ----------------------------
+# 4 projection by predicate and operator, used to be figure 3, now replaced ----
 
 # means and confidence intervals for projection rating by predicate / operator
 predicate_operator_means <- data %>% group_by(predicate, operator) %>%
[…]

### 2024-02-14 results/main/meta-analyses/1_projection/rscripts/3_analysis-predicates.R

@@ -18,11 +18,11 @@ library(emmeans)
 # summary + graphing
 library(ggplot2)
 library(ggh4x)
-source('../../../../helpers.R')
+source('../../../helpers.R')
 cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
                 "#CC79A7", "#000000")
 graphcolor <- "#0072B2"
-pinkk <- "#CC79A7"
+pinkk <- "#d694ba"
 pred_colors <- c(pinkk, "black")
 theme_set(theme_bw())
 
@@ -86,7 +86,8 @@ for (current_pred in predicates) {
 # print fixed effects for each model
 for (current_pred in predicates) {
   print(current_pred)
-  print(kable(fixef(eval(parse(text = paste("m.b.", current_pred, sep="")))), "latex", booktabs = TRUE))
+  # print(kable(fixef(eval(parse(text = paste("m.b.", current_pred, sep="")))), "latex", booktabs = TRUE))
+  print(fixef(eval(parse(text = paste("m.b.", current_pred, sep="")))))
 }
 
 # 3 GET PAIRWISE DIFFERENCES ---------------------------------------------------
@@ -244,12 +245,12 @@ strip <- strip_themed(background_x = elem_list_rect(fill = textcolors))
 # plot
 predicate_operator_means %>%
   ggplot() +
-  geom_point(aes(x = operator, y = projection), 
-             shape = 21, stroke = .5, size = .5, color = "black") +
-  geom_errorbar(aes(x = operator, y = projection, ymin = YMin, ymax = YMax), 
-                width = 0.5, color = "black", alpha = .6) +
-  geom_violin(data = data, aes(x = operator, y = projection, color = operator), 
+  geom_violin(data = data, aes(x = operator, y = projection, color = operator, fill = operator), 
               scale="width", alpha = .4) +
+  geom_point(aes(x = operator, y = projection, fill = operator), 
+             shape = 21, stroke = .5, size = 1.5, color = "black") +
+  geom_errorbar(aes(x = operator, y = projection, ymin = YMin, ymax = YMax), 
+                width = 0.2, color = "black") +
   geom_segment(data = contrasts, 
                aes(x = xstart, xend = xend, y = ystart, yend=yend, linetype = linetype)) +
   facet_wrap2(~ predicate, nrow = 4, strip = strip) +
@@ -263,6 +264,7 @@ predicate_operator_means %>%
[…]

### 2024-02-19 writing/SuB-paper/projection-operators-SuB.tex
SuB paper updates

@@
     The so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective in formal semantics/pragmatics (e.g., \citealt{chierchia_meaning_1990}).
-For instance, the clausal complement of \emph{discover} is diagnosed as as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.
+For instance, the clausal complement of \emph{discover} is diagnosed as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.
 
@@
 
-			The mean projection ratings by embedding context and predicate are presented in Figure~\ref{fig:figure4}, which shows that effect of embedding context on projection differs by predicate.
+	The mean projection ratings by embedding context and predicate are presented in Figure~\ref{fig:megaverid}, which shows that effect of embedding context on projection differs by predicate.
 
@@
 				\centering
-				\includegraphics[width = \linewidth]{mega-veridicality}
+				\includegraphics[width = \linewidth]{megaveridicality}
 				\caption{Mean projection ratings by embedding context and predicate in MegaVeridicality.}
-				\label{fig:figure4}
+				\label{fig:megaverid}
 			\end{figure}
@@
 
-The results of the experiments reported on in section \ref{s2} suggest that there is little by-operator projection variation when we aggregate over the clause-embedding predicates and their CCs, but that there is by-operator variation that cannot be neglected when we do not aggregate.
+    The results of the experiments reported on in section \ref{s2} suggest that there is little by-operator projection variation when aggregating observations for the CCs of the  clause-embedding predicates, but that there is by-operator variation that cannot be neglected when we do not aggregate.
 These results have two methodological implications.
 First, when initially investigating the projection of a content (or teaching projection to students), the family-of-sentences can indeed be treated as a family, as standard textbooks assume.
-However, it is advisable at this step already to apply the diagnostic with all four operators and attend to possible by-operator projection variation.
-
-A second methodological implication of the results of our experiments is that research on projection must consider which entailment-canceling operator to use, and report which entailment-canceling operator was used.
-This is because the results of the investigation may very well depend on the entailment-canceling operator under which the expressions and contents were embedded.
-For instance, two contents that exhibit projection variation from under polar questions might not exhibit such differences when embedded under a different entailment-canceling operator, and two contents that do not exhibit projection variation when embedded under negation, might exhibit variation when embedded under a different entailment-canceling operator.
+However, it is advisable to apply the diagnostic with all four operators and attend to possible by-operator projection variation.
+
+	A second methodological implication of the results of our experiments is that research on projection should take into account possible by-operator variation when choosing which entail\-ment-canceling operator to use.
+Further, results about projection variation should be relativized to the entailment-canceling operator under which the expressions and contents were embedded, as they may very well depend on this choice.
+For instance, two contents that exhibit projection variation when embedded under polar questions might not exhibit such differences from under a different entailment-canceling operator, and two contents that do not exhibit different projection across negation might exhibit variation when embedded under some other operator.
 
@@
 
-In this section, we consider whether contemporary analyses can capture the observed by-predicate projection variation and the interactions between predicate and entailment-canceling operator in projection ratings.
-%We consider three types of analyses: (i) analyses that assume that the CCs of factive predicates are lexically specified as presupposed  (\citealt{heim_projection_1983,van_der_sandt_presupposition_1992}), (ii) analyses on which backgrounded entailments are projective (\citealt{abrusan_predicting_2011,simons_best_2017}), and (iii) an analysis on which contextual entailments are presupposed
-%with a triggering algorithm based on world-knowledge that operates over contextual entailments of an utterance 
-%(\citealt{schlenker_triggering_2021}).
-
-%While all of these approaches meet limitations in accounting for our data, we argue that the observed gradient effects of predicate/operator interactions favor integrating insights from competing accounts, namely a contextual triggering algorithm (as in \citealt{abrusan_predicting_2011,simons_best_2017}) operating over contextual set of inferences (as in \citealt{schlenker_triggering_2021}).
-
-	%Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses.
+    This section discusses whether contemporary analyses can capture the observed by-predicate projection variation and the interactions between predicate and entailment-canceling operator in projection ratings.
+	% Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses.
 
@@
     These analyses do not predict the observed by-predicate variation, for several reasons.
-One reason is that there is variation between the factive predicates: For instance, the CC of \emph{be annoyed} is more projective than that of \emph{discover}, which in turn is more projective than that of \emph{reveal}.
+One reason is  variation among the factive predicates: For instance, when embedded under polar questions, negation, or \emph{perhaps,} the CC of \emph{be annoyed} is more projective than that of \emph{discover}, which in turn is more projective than that of \emph{reveal}.
+(In conditional antecedents, the order of \emph{be annoyed} and \emph{discover} is reversed, see Figure~\ref{fig:op-pred-ratings2}).
+    %
     This variation is not predicted because the CCs of factive predicates are invariably specified as presupposed, with no mechanism to predict projection variation between factive predicates.
@@
 
-%The assumption of categorical lexical triggering does not address the projection variation among purported factive predicates (see also discussion in \citealt{degen_are_2022}).
-It also cannot address the fact that the CCs of some non-factive predicates (e.g., \emph{inform, acknowledge}) project just as much, or more, than some (semi-)factive predicates (e.g., \emph{reveal, discover, see}).
-That is because, it only makes predictions for the projection of the CC of (semi-)factive predicates, whose CCs are analyzed as presuppositions.
-It therefore does not address the data for the non-veridical predicates in our study.
-		%
-%		Further, the categorical lexical distinctions are not sufficiently fine-grained to make predictions about the gradient by-predicate variation in the effect of operator on projection.
-		%
-%		This interaction effect can only be addressed by taking into account more fine-grained lexical distinctions than assuming that some predicates are presuppositional, while others are not.
-
-        %In our experiment, the stimulus utterances were presented out of the blue, leaving it to the participants to imagine a suitable discourse context.
-Therefore, none of the CCs should be interpreted as necessarily inconsistent with contextual information.
-		%
     The analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} also do not predict differential effects of entailment-cancelling operators on projection.
 This is because negation and conditional antecedents are given a semantics that derives their behavior as presuppositional holes, in the sense of \citealt{karttunen_observations_1971}.
-For instance in \citealt{heim_projection_1983}, presupposition triggers under negation or in conditional antecedents are evaluated relative to the global context, with the consequence that presuppositions are not expected to differentially project from under negation or conditional antecedents.
-%Therefore, presupposition triggers under these operators are predicted to behave like unembedded ones, with no expected differences.
-  While the analysis does not explicitly address epistemic possibility modals or polar questions, we might expect that they would be treated as presuppositional holes along the same lines, therefore also not predicting any by-operator projection variation, by by-predicate variation in the effect of entailment-canceling operators.
+For instance in \citealt{heim_projection_1983}, presupposition triggers under negation or in conditional antecedents are evaluated relative to a local context that is equal to the global context, with the consequence that presuppositions are not expected to project differently from under negation or conditional antecedents.
+While the analysis does not explicitly address epistemic possibility modals or polar questions, we might expect that they would be treated as presuppositional holes along the same lines, therefore also not predicting any by-operator projection variation, or by-predicate variation in the effect of entailment-canceling operators.
 	%
@@
     \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} do not assume that presuppositions are lexically specified.
-In \citealt{abrusan_predicting_2011}, a lexical entailment of an uttered sentence is a presupposition if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}) addressed by the utterance.
-For instance, the CC of B's utterance in \ref{ent}, that Phil's ballet class is canceled, is predicted to be a presupposition (and therefore to project) because it is a lexical entailment of the sentence embedded under the epistemic modal and not at-issue with respect to A's interrogative utterance.
+In \citealt{abrusan_predicting_2011}, a lexical entailment of a (possibly embedded) sentence is a presupposition, if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}) addressed by the utterance.
+For instance, the CC of B's utterance in \ref{ent}, that Phil's ballet class is canceled, is predicted to be a presupposition (and therefore to project) because it is a lexical entailment of the modal prejacent and not at-issue with respect to A's interrogative utterance.
 
@@
 If so, the Current Question entails that Phil's ballet class is canceled today, and the CC may therefore project under \poscite{simons_best_2017} analysis.
-
-%assume that projection is not determined lexically, but by a triggering mechanism, which operates over the entailments of an expression, and determines which of them projects.
-They assume entailments project that do not contribute to the main point invoked by the utterance.
-		% Instead of assuming that the possibility of projection is determined lexically, the accounts in \cite{abrusan_predicting_2011,simons_best_2017} and \citet{schlenker_triggering_2021} assume a triggering mechanism, which operates over a set of inferences associated with an utterance, and determines which of them projects.
-The assumed triggering mechanisms differ on two dimensions: (i) how the set of inferences is determined that constitutes the input to the triggering mechanism, and (ii) the algorithm by which these inferences are derived as projective or not.
-
 		%
@@
 
-		%The account assumes that an inference projects if it is entailed by the union of the alternatives (see also \citealt{abusch_presupposition_2010}).
-therefore predicting projection of the CC in \Last, but not \LLast.
-That is, under the assumption that $x$ \emph{was annoyed that} $p$ entails $p$.
-
-Like the analyses discussed in the previous section, the analyses in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} do not predict the observed by-predicate variation.
+    Under the analyses in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017}, CCs that are entailed based on the literal content of an utterance may or may not project, depending on whether they contribute to the main point of the utterance.
+This opens up the possibility for projection variation among veridical predicates (including those often characterized as factive), representing an improvement compared to the lexical specification of factivity assumed in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992}.
+However, these accounts still do not predict the observed by-predicate variation.
+    %
+    First, to make explicit predictions about the observed by-predicate variation, these types of analyses would need to make explicit assumptions about how different predicates contribute to how the question under discussion is chosen.
+    %
+    % The analyses in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} do not predict the observed by-predicate variation.
 First, if the CCs of factive predicates are entailed, they are invariably predicted to project, and so the observed differences between the CCs of different factive predicates are not predicted.
-Second, the analyses do not make predictions for non-entailed CCs, and hence they do not predict that the CCs of some non-factives predicates project at least as much or even more than those of some factive predicates.
-Neither of the analyses makes predictions about by-operator projection variation, nor about by-predicate variation in the effect of entailment-canceling operators.
-
-%  The triggering mechanisms assumed in \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017}, predict a potential of projection for inferences that are entailed by the literal content of the modal prejacent $p$ in \Last and \LLast, and similarly for propositions embedded under negation, and in conditional antecedents or polar questions.
-These accounts, therefore, predict a potential of projection for CCs that are entailed by the prejacent, due to the lexical entailments of veridical predicates, or due to entailments that are contributed compositionally by the literal content (see e.g., \citealt{roberts_i_2019}).
-However, they do not make systematic predictions for CCs, such as in \Last, that are not entailed.
-As a result, these accounts cannot address the fact that we see some amount of projection for all of the tested predicates.
-
-		%Because the main point of an utterance is often determined in the discourse context (\citealt{roberts_information_1996,roberts_information_2012}, \emph{et seq.}), these approaches incorporate discourse effects on projection.
-		%
-		% \citet{abrusan_predicting_2011} also makes predictions for utterances in out-of-the-blue contexts: Utterances are associated with a default main point that is grammatically determined by the time at which the matrix predicate is interpreted.
-		%
-		%The account in \citealt{simons_best_2017} could be extended to address our gradient data in out-of-the blue contexts, by assuming (i) that embedded contents that are not at-issue can project even if they are not entailed, and (ii) that linguistic expressions differ in how likely the contributed contents are interpreted as at-issue (as suggested in \citealt{tonhauser_how_2018}).
-		%
-		%In our data, for instance, the CC of \emph{discover} \Next[a] has received relatively high projection ratings, while that of \emph{confirm} \Next[b] has received medium to low ratings.
-To explain this difference, the information-structure based triggering approach would need to assume that the CC of \emph{discover} is less likely to be at issue than that of \emph{confirm}.
-
-	%	\ex. \a. Perhaps Cole discovered that Julian dances salsa.
-	%		\b. Perhaps Cole confirmed that Julian dances salsa.
-%			\z.
-%		\z.
-
-		%While \citet{tonhauser_how_2018} experimental evidence that projection ratings negatively correlate with at-issueness ratings in out-of-the-blue polar questions.
-\citet{sieker_projective_2022} argue that their data also shows the same correlation for modals, conditional antecedents, and negation for the German clause-embedding predicates they investigated.
-To address the gradient by-predicate variation theoretically, this approach would need to explicitly spell out how various predicates are associated with different likelihoods that their CCs is interpreted as at-issue.
-
-		%\paragraph{By-operator variation.} Neither of the analyses make predictions about by-operator variation.
-			%Since we found an operator/predicate interaction effect on projection, a further prediction of this type of approach would be that at-issueness ratings would show an operator-predicate interaction in this experimental setup as well.
-To address the interaction effect theoretically, an information structural triggering approach would need to make explicit predictions about how the meaning of entailment-cancelling operators affects at-issueness differentially for different predicates.
-This is certainly conceivable, when we take default assumptions about how likely embedded contents are interpreted at issue to be associated with the entire embedding context, including clause-embedding predicates, but also entailment-cancelling operators.
-However, this has not been spelled out explicitly and needs to be investigated in more detail.
+    %
+    Second, the analyses do not make systematic predictions for non-entailed CCs, and hence they do not predict that the CCs of some non-factives predicates project at least as much or even more than those of some factive predicates.
+
+    Both of these analyses set aside the effect of entailment-cancelling operators in determining the main point of an utterance.
+Therefore they do not make predictions about by-operator differences or by-predicate variation in the effect of entailment-canceling operators.
 
@@
 
-On the analysis proposed in \citealt{schlenker_triggering_2021}, the CC of a sentence S, like \ref{knowa}, that is uttered in a context $c$ is presupposed if the CC is presupposed by the sentence S$'$under the entailment-canceling operator, that is \ref{knowb}, in the local context $c'$.
-For the CC to be presupposed in \ref{knowb}, two conditions must be met: i) S$'$ contextually entails the CC relative to $c'$; and ii) If we consider ``a generic agent'' who believes the propositions in $c'$ and who has now learned about the truth of S$'$, then the probability that this generic agent already believed the CC is above a contextual threshold $\alpha$ (\S6.2); more colloquially, condition ii) requires that the generic agent ``typically antecedently believes'' the CC (p.6) upon interpreting S$'$ in $c'$.
+    Under the view put forward in \citealt{schlenker_triggering_2021}, the CC of a sentence $S$ like \ref{knowa}, is presupposed in a context $c$, if the CC is an epistemic precondition on the global meaning of $S$.
+Formally, Schlenker defines presuppositions relative to their local context (\S6.2).
+For instance, in \ref{knowa}, the negative prejacent $S'$, that is \ref{knowb}, is presupposed in the local context under negation $c'$, if two conditions are met:
+	i) $S'$ contextually entails the CC relative to $c'$;
+	and ii) If we consider ``a generic agent'' who believes the propositions in $c'$ and who has now learned about the truth of S$'$, then the probability that this generic agent already believed the CC is above a contextual threshold $\alpha$;
+	more colloquially, condition ii) requires that the generic agent ``typically antecedently believes'' the CC (p.6) upon interpreting S$'$ in $c'$.
+Based on \citealt{heim_projection_1983} and \citealt{schlenker_local_2009}, the the local context under negation $c'$ is assumed to be identical to the global context $c$.
+Therefore, epistemic preconditions relative to $c'$ will also apply to the global context.
+
 	Applied to \ref{know}, \poscite{schlenker_triggering_2021} analysis predicts that the CC  is presupposed if i) \ref{knowb} contextually entails that Julian dances salsa, and ii) if a generic agent would typically antecedently believe that Julian dances salsa upon interpreting \ref{knowb} in the minimal contexts we provided to our participants.
-(For \citealt{schlenker_triggering_2021}, the local context under negation $c'$ is identical to $c$.)
 
@@
 \citealt{schlenker_triggering_2021} also assumes the condition ii) is met: ``in many cases, one's knowledge of facts will precede one's knowledge of [Cole's] beliefs about them\ldots believing that [Julian dances salsa] is often an epistemic precondition for believing that'' Cole knows that Julian dances salsa (p.6).
-One might, however, challenge this assumption on the basis of the corpus study presented in \citealt{spenader02}, which showed that the CCs of the majority of the utterances of sentences with factive verbs (namely 81 out of 109) had to be accommodated (i.e., were not contextually entailed).
-In other words, utterance of sentences with the factive predicates investigated by \citet{spenader02}, which included \emph{know}, were ``generally used to communicate information the speaker thought was hearer-new'' (p.99).
+One might, however, challenge this assumption on the basis of the corpus study presented in \citealt{spenader_presuppositions_2002}, which showed that the CCs of the majority of the utterances of sentences with factive verbs (namely 81 out of 109) had to be accommodated (i.e., were not contextually entailed).
+In other words, utterance of sentences with the factive predicates investigated by \citet{spenader_presuppositions_2002}, which included \emph{know}, were ``generally used to communicate information the speaker thought was hearer-new'' (p.99).
 This result might therefore suggest that one cannot assume that a generic agent typically antecedently believes the CC of {\em know}.
 
-We assume that it is an open, empirical question which predicates are such that the probability of a generic agent antecedently believing the CC is above the contextual threshold $\alpha$ in the minimal contexts we provided our participants (and, of course, what that threshold might be).
-\citealt{schlenker_triggering_2021} appears to assume that there are two classes of predicates: those where the probability is above the threshold (including {\em know, inform}, and {\em announce}), and those where it is not (including {\em demonstrate} and {\em establish}; see p.12 and appendix I).
-Thus, one advantage of \poscite{schlenker_triggering_2021} analysis over the analyses reviewed in sections \ref{s:heim} and \ref{s:abrusan} is that it is able to predict the projection of the CCs of non-factives predicates (modulo the open questions about condition ii)).
-
-Does \poscite{schlenker_triggering_2021} analysis capture the observed by-predicate variation?
-The analysis correctly predicts that the CCs of \emph{know, inform}, and \emph{announce} are more projective than the CCs of \emph{demonstrate} and \emph{establish}, by virtue of the CCs of the former being possibly presupposed, in contrast to the CCs of the latter.
-It is not clear, however, that the analysis is able to predict the observed variation because the analysis -- even  though it does not divide predicates into factive and non-factives ones -- nevertheless suggests a binary, categorical distinction between predicates, such that the CCs of predicates with probabilities above the threshold may be presupposed (depending on the context), and those below the threshold are not.
-First, as shown in Figs.~\ref{fig:op-pred-ratings} and \ref{fig:op-pred-ratings2}, and discussed extensively in \citealt{degen_are_2022}, there is projection variation between the supposedly presupposed CCs that is not captured by an analysis that does not make more fine-grained distinctions between the meanings of clause-embedding predicates.
-Furthermore, it is not clear that the supposedly nonpresupposed CCs are not projective.
-Second, as discussed in \citealt{degen_are_2022}, the CCs of non-factives predicates like \emph{establish} and \emph{demonstrate} are projective when compared to nonprojective main clause content.
-\poscite{schlenker_triggering_2021} analysis does not make predictions for these CCs.
-In the experiments reported on in section \ref{s2}, the mean certainty rating of the CC of \emph{announce}, whose CC is assumed to possibly be presupposed, is .XX, and that of {\em demonstrate}, whose CC is not assumed to be presupposed, is .XX (both means are aggregated across entailment-canceling operators).
+    \poscite{schlenker_triggering_2021} analysis does not incorporate differential effects of entailment-cancelling operators on projection.
+As discussed for \citealt{heim_projection_1983}, above, the local contexts under negation and in conditional antecedents are both assumed to be the global context (and similarly in \citealt{schlenker_local_2009}).
+Therefore, the analysis does not make predictions about by-operator projection variation or about by-predicate variation in the effect of entailment-canceling operator.
+
+    Does the account capture the observed by-predicate variation?
+    %
+    Because the probabilities about previous beliefs and the threshold $\alpha$ are assumed to potentially vary based on the context, the formal proposal does not in principle preclude the possibility of by-predicate projection variation.
+However, the subjective conditional probabilities associated with expressions and the contextual threshold $\alpha$ are taken as given in the account, and therefore do not offer explicit or systematic predictions about projection ratings for different expressions.
+    %
+    We, therefore, assume that it is an open, empirical question which predicates are such that the probability of a generic agent antecedently believing the CC is above the contextual threshold $\alpha$ in the minimal contexts we provided our participants (and, of course, what that threshold might be).
+
+    Throughout the discussion, however, \citealt{schlenker_triggering_2021} appears to assume that there are two classes of predicates: those where the probability is usually above the threshold (including {\em know, inform}, and {\em announce}), and those where it is not (including {\em demonstrate} and {\em establish}; see p.12 and appendix I).
+    %
+    As this division does not fall along the lines of traditionally assumed classes of (non-)factives,
+    % Thus,
+    one advantage of \poscite{schlenker_triggering_2021} analysis over the analyses reviewed in sections \ref{s:heim} and \ref{s:abrusan} is that it predicts the projection of the CCs of (certain) non-factive predicates (modulo the open questions about condition ii)).
+
+    Based on these assumptions, the analysis correctly predicts that the CCs of \emph{know, inform}, and \emph{announce} are more projective than the CCs of \emph{demonstrate} and \emph{establish}, by virtue of the CCs of the former being usually presupposed, in contrast to the CCs of the latter.
+It is not clear, however, that the analysis is able to predict the observed variation, because the analysis -- even  though it does not divide predicates into factive and non-factives ones -- nevertheless imposes a binary, categorical distinction between predicates.
+    % such that the CCs of predicates with probabilities that are usually above the threshold are presupposed, and those below the threshold are not.
+    %
+    Crucially, the by-predicate variation reported in Sections \ref{s:by-pred}+\ref{s:factivity} is not captured by an analysis that does not make more fine-grained distinctions between the meanings of clause-embedding predicates, as discussed extensively in \citealt{degen_are_2022}.
+First, there is projection variation between the supposedly presupposed CCs (that of \emph{know} is more projective than the CC of \emph{inform}, which is more projective than that of \emph{announce}, under all four operators).
+    %
+    Furthermore, as discussed in \citealt{degen_are_2022}, the CCs of the supposedly nonpresupposed CCs (of \emph{establish} and \emph{demonstrate}) are projective when compared to nonprojective main clause content.
+In the experiments reported on in section \ref{s2}, the mean certainty rating of the CC of \emph{announce}, whose CC is assumed to possibly be presupposed, is $0.55$, and that of {\em demonstrate}, whose CC is not assumed to be presupposed, is $0.43$ (both means are aggregated across entailment-canceling operators).
 It is not clear whether this particular difference in mean certainty rating motivates analyzing the CC of \emph{announce} as presupposed in contrast to that of \emph{demonstrate}.
 
-%The triggering mechanism assumed in \citealt{schlenker_triggering_2021} operates on local contextual entailments of an expression, and determines which one of them projects based on probabilistic world knowledge about the truth of the invoked contents.
-For instance, the CC of the utterance with \emph{discover} in \Next[a], that Julian dances salsa, has the potential to project if it is entailed by the sentence in the scope of negation (namely \Next[b]) together with information that is available in its local context.
-Negation is treated as a presuppositional hole, and the local context under negation is taken to be the global context (as in  \citealt{heim_projection_1983} and \citealt{schlenker_local_2009}).
-
-	%		\ex. \a. Cole didn't discover that Julian dances salsa.
-	%			\b. Cole discovered that Julian dances salsa.
-	%		\z.
-
-			%The triggering mechanism states that an inference $p$ of some propositional expression $E$ pro\-jects if $p$ is an epistemic precondition for the truth of $E$ in its local context $c^\prime$.
-So, the CC of \Last[a] projects if the CC is an epistemic precondition for the truth of \Last[b].
-			%
-			%$p$ is an epistemic precondition of $E$ in $c^\prime$ if, usually, when one acquires the belief in $c^\prime$ that $E$, one already knows that $p$. 
-			%
-			%The CC of \Last[a] can be taken to be an epistemic precondition for \Last[b] based on the \lq subjective conditional probability\rq\ that a generic epistemic agent would already believe that Julian dances salsa when learning that \Last[b] is true.
-Specifically, the CC is taken to be an epistemic precondition if that probability reaches a contextually given threshold.
-
-%   In order to extend an account along these lines to address the out-of-the-blue utterances in our experiment, we might assume that these utterances are interpreted relative to an imagined plausible discourse context (for discussion, see \citealt{simons_best_2017}).
-
-   %It can therefore incorporate effects of contextual information on projection, such as the reliability of the source of a speech report (\citealt{rieh_credibility_2010,de_marneffe_did_2012}).
-
-%Due to the probabilistic nature and context-sensitivity of the triggering mechanism, the approach can make gradient predictions.
-However, the subjective conditional probabilities associated with expressions and their contextual inferences are taken as given in the account.
-In order to make concrete predictions about the gradient by-predicate variation, this account would need to make explicit how the relevant probabilities are derived based on different predicates.
-For instance, to explain the contrast that \emph{discover} receives higher projection ratings than \emph{confirm}, this account would assume that the situations in which one learns that a sentence like \emph{$x$ discovered that $p$} is true, are such that one already knew that $p$ is true more often than situations in which one learns the truth of a sentence like \emph{$x$ confirmed that $p$}.
-
-\poscite{schlenker_triggering_2021} analysis does not incorporate differential effects of entailment-cancelling operators on projection.
-For instance, the local contexts $c'$ under negation and in conditional antecedents are both assumed to be the global context $c$ (based on \citealt{heim_projection_1983,schlenker_local_2009}).
-Therefore, the analysis does not make predictions about by-operator projection variation or about by-predicate variation in the effect of entailment-canceling operator.
-
-%For instance, the conditional probability for $p$ given $E$ vs.\ that of $p$ given $(\emph{not } E)$ or that of $p$ given $(\emph{if } p, q)$ should not differ.
-
-%			In order to address the operator-predicate interaction effect, an approach based on epistemic preconditions would need to make explicit how the meaning of entailment-cancelling operators affects the conditional probabilities differentially for different predicates.
-			%
-%			The conditional probability that some embedded content is antecedently known would need to be computed relative to a global context, and take into account the entire embedding context, including the combination of entailment-cancelling operator and clause embedding predicate.
-
 \subsubsection{Interim summary}
@@
     There is currently no projection analysis on the market that predicts the by-predicate and by-operator variation we observed in the experiments reported on in section \ref{s2}.
-\citealt[585]{degen_are_2022} suggested that an empirically adequate analysis to capture the observed by-predicate variation requires consideration of ``more fine-grained distinctions [between clause-embedding predicates] that are based on the lexical meaning and discourse use of clause-embedding predicates''.
+\citealt{degen_are_2022} suggested that an empirically adequate analysis to capture the observed by-predicate variation requires consideration of ``more fine-grained distinctions [between clause-embedding predicates] that are based on the lexical meaning and discourse use of clause-embedding predicates'' (p.~585).
+Further, an account of the observed by-predicate variation in the effects of entailment-cancelling operators will need to take into account how semantic and pragmatic properties of predicate meanings interact with entailment-cancelling operators in varying ways.
 In the next section, we offer some hypotheses based on the data from our experiments.
-
-%	\subsubsection{Discussion: Contextual inferences and probabilistic triggering}
-
-%		Our data showed gradient projection variation, based on a main effect of predicate, and an operator-predicate interaction.
-We found these effects for the CCs of clause-embedding predicates under four entailment-cancelling operators in out-of-the-blue contexts.
-The fact that we see some amount of projection for all predicates (even non-veridical ones), can only be addressed by an account that assumes the potential of projection for contextual inferences, rather than only literal ones.
-
-%		The main effect of predicate and the operator-predicate interaction are both gradient effects, and cannot be explained by a categorical distinction between presuppositional and non-pre\-suppo\-si\-tion\-al predicates.
-The effects could rather be addressed in a probabilistic account, where various embedding contexts (i.e., combinations of entailment-cancelling operator and predicate) are associated with different probabilities that the embedded content projects.
-
-%		Under the assumption of information-structural triggering, this is the probability that the embedded content is not at issue in the (possibly imagined) discourse context.
-Under a world-knowledge-based triggering account, the projection probability can be associated with the probability that the embedded content is already known in the discourse context.
-These two notions are, of course, related.
-If some proposition is already known, raising an issue about whether $p$ is the case will be uninformative.
-Therefore, it is unclear whether these two approaches would make different predictions for our data.
-Both types of accounts would need to be extended to make concrete predictions for the contribution of the combinations of entailment-cancelling operators and predicates.
 
@@
 		\centering
-		% \hspace{-.5em}\resizebox{\linewidth}{!}{
-		\begin{tabular}{p{.205\linewidth} p{.18\linewidth} p{.18\linewidth} p{.18\linewidth} p{.18\linewidth}}
-			\footnotesize \hspace{2.1em} (a) Negation high &
-			\footnotesize \ (b) Negation low &
-			\footnotesize (c) Conditional high &
-			\footnotesize \ (d) Modal low
-			% & \footnotesize \  (e) Discover
-			\\
-			\vspace{-.6\baselineskip}
-			\includegraphics[width=.225 \textwidth, valign=T]{negation-high-1.pdf}
-			&
-			\vspace{-.6\baselineskip}
-			\includegraphics[width=.2\textwidth, valign=T]{negation-low-1.pdf}
-			&
-			\vspace{-.6\baselineskip}
-			\includegraphics[width=.2\textwidth, valign=T]{conditional-high-1.pdf}
-			&
-			\vspace{-.6\baselineskip}
-			\includegraphics[width=.2\textwidth, valign=T]{modal-low-1.pdf}
-			% &
-			% \vspace{-.6\baselineskip}
-			% \includegraphics[width=.2\textwidth, valign=T]{discover-1.pdf}
-			\\
-		\end{tabular}
-		% }
-
+		\includegraphics[width=\textwidth]{predicate-profiles.pdf}
 		\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).}
@@
 
-	%
+	%(a)
 	The non-veridical predicates \emph{pretend} and \emph{think} exhibit the `Negation high' pattern, shown in panel (a) of Figure~\ref{fig:patterns}.
 These are the only predicates that are most projective under negation compared to all other operators.
-This could be related to
-	assumptions that non-veridical doxastics like \emph{think} can often be interpreted in comparison to a stronger veridical alternative (e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
+This generalization, supported by the statistical analysis in Section \ref{s:by-pred} (see Figure \ref{fig:op-pred-analysis}), could be related to assumptions that non-veridical doxastics like \emph{think} can often be interpreted in comparison to a stronger veridical alternative (e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
 We tentatively hypothesize that this can lead to an inference that the CC is false in upward-monotone contexts, but not under negation.
@@
 
-	The inferential predicates \emph{confirm, demonstrate}, and \emph{prove} exhibit a `Negation low' pattern, shown in panel (b).
-These are most projective under modals and conditionals and least projective under negation.
-Here, we may hypothesize that a negative statement involving these predicates is often only taken as relevant (in a Gricean sense) when it is antecedently known that there was an attempt to confirm/demonstrate/prove the truth of the CC.
-This may result in an inference that this attempt failed and result in lower projection ratings under negation.
-
-	For \emph{admit, announce, confess}, and \emph{reveal}, the CC is most projective when embedded in conditional antecedents: This `Conditional high' pattern (c) may suggest that the discourse effect of a conditional interacts with these change-of-state communication predicates.
-
-	Finally, the predicates \emph{hear, inform,} and \emph{know} exhibit a `Modal low' pattern (d).
-The lexical meaning of these predicates, whose CCs are among the most projective, appears to interact with the modal adverb {\em perhaps}, yielding lower projection ratings.
+	%(b)
+	These latter two predicates, along with \emph{be right}, exhibit the `Question low' pattern, shown in panel (b).
+They are the only predicates which are least projective in polar questions (again, compare Figure \ref{fig:op-pred-analysis}).
+We tentatively suggest that these communicative predicates can interact with the pragmatics of polar questions in a way that can lead to an inference of incredulity towards the CC.
+However, there are also some differences between these predicates.
+The veridical \emph{be right} is most projective under modals (M $>$ C $>$ N $>$ Q), whereas \emph{say} and \emph{suggest} are most projective in conditional antecedents (C $>$ M,  N $>$ Q).
+
+	%(c+d)
+	The two u-shaped patterns in panels (c) and (d) are characterized by relatively low projection ratings with negation and high ratings with conditionals.
+While there is fine-grained variation with regards to which of the by-operator differences were statistically supported, (c) and (d) can be distinguished by their ratings for the modal.
+	%
+	The change-of-state communication predicates \emph{admit, announce, confess}, and \emph{reveal}, exhibit relatively low projection ratings of the CC when embedded under \emph{perhaps} (C $>$ Q  $>$ M, N).
+In contrast, the inferential predicates \emph{confirm, demonstrate, establish}, and \emph{prove} exhibit relatively high ratings with \emph{perhaps} (\emph{confirm:} C $>$ M, Q $>$ N; \emph{demonstrate:} C $>$ M, Q $>$ N; \emph{establish:} C, M $>$ Q, N; \emph{prove:} M $>$ C  $>$ Q, N).
+For an explanation of the high conditional ratings, one might examine ways in which the discourse effect of a conditional interacts with the change-of-state component of these inferential and communicative predicates.
+Our tentative suggestion for the relatively low negation ratings is that these predicates can be interpreted relative to contextual assumptions that lead to a neg-raising type inference more readily than others (so that, for instance, not announcing $p$ amounts to communicating not $p$, or not proving $p$ amounts to inferring not $p$).
+
+	%(e+f)
+	Finally, the two roof-shaped patterns in (e) and (f) are characterized by relatively high projection ratings with questions, and low ratings with {\em perhaps}.
+We can distinguish the two patterns by their relative ratings for conditionals.
+The predicates \emph{acknowledge, discover, hear,} and \emph{see}, which are associated with a change of some informational state, show relatively high ratings with conditionals (Q, C $>$ N $>$ M).
+The ratings for conditionals are relatively lower for the predicates \emph{be annoyed} (Q $>$ N $>$ C $>$ M), and \emph{inform}, and \emph{know} (Q $>$ N, C $>$ M), whose CCs are among the most projective.
 
@@
 To minimize the variability of the effect of the contents of these consequent clauses on the projection of the contents of the 20 complement clauses, the consequent clauses all consist of a uniquely named subject and an adjectival predication in the future tense ({\em will be}), and the adjectives all denote an emotion.
-We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner-etal2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
+We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner_norms_2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.
 
@@
 Participants may judge B's response in (3b) to be less acceptable than in (3a) because of this missing content specification.
-This hypothesis is consistent with the results of \posscite{syrett-koev2015} Exp.~3, which suggests that content of a sentence-medial NRRCs can be the target of a direct denial, though the main clause content is preferred as the target of such a denial.
+This hypothesis is consistent with the results of \posscite{syrett_experimental_2015} Exp.~3, which suggests that content of a sentence-medial NRRCs can be the target of a direct denial, though the main clause content is preferred as the target of such a denial.
 That the group means on the control stimuli in our Exps.~2n, 2m, and 2c are still relatively low may be due to the fact that the one content that was specified is the at-issue main clause content.

### 2024-02-19 results/MegaVeridicality-analysis/rscripts/analysis.R

@@ -68,9 +68,6 @@ table(mv1$veridicality_num)
 
 table(mv1$verb)
 
-factives <- c("discover", "know", "reveal", "see", "amuse", "annoy", "find_out", "forget", "learn",
-              "love", "notice", "realize", "recognize", "regret", "remember", "reveal", "understand")
-
 # create items for mv1 
 
 table(mv1$verb) #verb
@@ -84,7 +81,6 @@ str(mv1$conditional2)
 mv1$conditional2 <- as.character(mv1$conditional)
 mv1$conditional2[mv1$conditional2 == "True"] <- "conditional"
 mv1$conditional2[mv1$conditional2 == "False"] <- "matrix"
-
 table(mv1$conditional2) #conditional, matrix
 
 mv1$item <- paste(mv1$verb,mv1$frame,mv1$voice,mv1$polarity,mv1$conditional2, sep = "-")
@@ -122,42 +118,65 @@ table(mv1_tmp$embedding)
 p_means = mv1_tmp %>%
   group_by(verb, embedding) %>%
   summarize(Mean = mean(veridicality_num), CILow = ci.low(veridicality_num), CIHigh = ci.high(veridicality_num)) %>%
-  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, verb = fct_reorder(as.factor(verb),Mean))
-options(tibble.print_max = Inf)
-p_means
+  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, verb = as.factor(verb), embedding = as.factor(embedding))
+# options(tibble.print_max = Inf)
+
+p_means$verb <- fct_reorder(p_means$verb, p_means$Mean, .fun = "mean")
 levels(p_means$verb) # verbs sorted by projectivity mean 
 
-# create data subsets for the factives
-p_meansFactives <- droplevels(subset(p_means,p_means$verb %in% factives))
-p_meansFactives
-str(p_meansFactives$verb)
-levels(p_meansFactives$verb) # sorted by projectivity mean (pretend...be annoyed)
+# define subsets for selection
 
-p_means <- mutate(p_means, verb = fct_reorder(verb, Mean, .fun = mean))
-p_means <- mutate(p_means, embedding = fct_reorder(embedding, Mean, .fun = mean))
-p_meansFactives <- mutate(p_meansFactives, verb = fct_reorder(verb, Mean, .fun = mean))
-p_meansFactives <- mutate(p_meansFactives, embedding = fct_reorder(embedding, Mean, .fun = mean))
[…]

### 2024-02-19 results/main/meta-analyses/1_projection/rscripts/1_summary-graphs.R

@@ -8,6 +8,8 @@ setwd(this.dir)
 
 # summary + graphing
 library(ggplot2)
+library(ggh4x)
+library(ggrepel)
 source('../../../helpers.R')
 cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
                 "#CC79A7", "#000000")
@@ -68,6 +70,7 @@ data$VeridicalityGroup = as.factor(
   ifelse(data$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
 predicate_means$VeridicalityGroup = as.factor(
   ifelse(predicate_means$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
+
 pred_order <- levels(fct_reorder(predicate_means$predicate, predicate_means$projection, .fun = "mean"))
 textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), pinkk, "black")
 
@@ -91,7 +94,7 @@ predicate_means %>% ggplot(aes(x = fct_reorder(predicate, projection, .fun = "me
         legend.position="bottom", text = element_text(size=12)
   )
 
-ggsave(f="../graphs/predicate-graph-1.pdf", width=7, height=4.5)
+ggsave(f="../graphs/predicate-graph.pdf", width=7, height=4.5)
 
 
 # 4 projection by predicate and operator, used to be figure 3, now replaced ----
@@ -125,7 +128,7 @@ predicate_operator_means %>%
         legend.position="none", text = element_text(size=12)
   )
 
-ggsave("../graphs/predicate-operator-graph-1.pdf", width=7, height=4.5)
+ggsave("../graphs/predicate-operator-graph.pdf", width=7, height=4.5)
 
 # 5 projection by predicate with violin plots, for each operator (fig. 4) ------
 
@@ -245,3 +248,92 @@ subset(predicate_operator_means, operator == "conditional") %>%
 
 ggsave("../graphs/conditional-predicate-graph.pdf", width=7, height=3.75)
 
+
+
+# 6 predicate profiles
+
+negHigh <- c("pretend", "think")
+roof1 <- c("acknowledge", "see", "discover", "hear")
+roof2 <- c("inform", "know", "be_annoyed")
[…]

### 2024-02-19 results/main/meta-analyses/1_projection/rscripts/2_analysis-operators.R

@@ -40,17 +40,17 @@ summary(data$betaresponse)
 # beta regression w response: scaled projection ratings, fixed effect: operator
 # random effects: random intercepts for item, participant, slopes for both by operator
 # also estimating these effects on beta precision parameter
-beta_formula = bf(betaresponse ~ operator + (1 | predicate),
-                  phi ~ operator + (1 | predicate), # beta distribution's precision
+beta_formula = bf(betaresponse ~ operator + (1 + operator | predicate) + (1 | item),
+                  phi ~ operator  + (1 + operator | predicate) + (1 | item), # beta distribution's precision
                   family = Beta())
 
 m.b.operators <- brm(formula = beta_formula,
     family = Beta(),
     data = data,
-    cores = 4, iter = 3000, warmup = 500,
-    control = list(adapt_delta = .95,max_treedepth=15))
+    cores = 4, chains = 4, iter = 8000, warmup = 700,
+    control = list(adapt_delta = .95, max_treedepth=15))
 
-saveRDS(m.b.operators, "../models/beta-model-mixed-operators.rds")
+saveRDS(m.b.operators, "../models/beta-model-mixed-operators2.rds")
 summary(m.b.operators)
 
 # run posterior predictive checks
@@ -61,7 +61,7 @@ p1
 # pairwise comparison of operators ----
 
 # read the model
-m.b.operators <- readRDS("../models/beta-model-mixed-operators.rds")
+m.b.operators <- readRDS("../models/beta-model-mixed-operators2.rds")
 
 # draws of posterior distributions of estimated marginal means of pairwise differences
 pairwise <- m.b.operators %>%

### 2024-02-19 results/main/meta-analyses/1_projection/rscripts/3_analysis-predicates.R

@@ -142,7 +142,7 @@ levels(predicate_operator_means$operator)
 levels(predicate_operator_means$predicate)
 
 pred_order <- levels(predicate_operator_means$predicate)
-textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), pinkk, "white")
+textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "#cb7eab", "white")
 
 data <- data %>% mutate(operator = fct_reorder(operator, projection, .fun = "mean"))
 levels(data$operator)
@@ -164,7 +164,6 @@ for (p in predicates) {
   }
 }
 View(contrasts)
-
 
 # Get table with all contrasts for supplement
 print(xtable(contrasts),
@@ -200,6 +199,43 @@ contrasts <- contrasts %>% mutate(
 )
 
 table(contrasts$linetype)
+
+
+contrasts <- contrasts %>%
+  mutate(predicate = as.factor(predicate), contrast = as.factor(contrast), 
+         difftype =  case_when(
+           lower <= 0 & upper >= 0 ~ 0,
+           mean <= 0 & abs(mean) >= 1 ~ -1.5,
+           mean >= 0 & abs(mean) >= 1 ~ 1.5,
+           mean <= 0 & abs(mean) > 0.5 ~ -1,
+           mean >= 0 & abs(mean) > 0.5 ~ 1,
+           mean <= 0 & abs(mean) > 0 ~ -.5,
+           mean >= 0 & abs(mean) > 0 ~ .5,
+           TRUE ~ NA
+         ))
+           
+# clustering predicates by pairwise operator differences
+contrasts_wide <- data.frame(matrix(nrow = length(levels(contrasts$predicate)),
+                             ncol = length(levels(contrasts$contrast))),
+                      row.names = levels(contrasts$predicate))
+
+colnames(contrasts_wide) <- levels(contrasts$contrast)
+
+for (i in 1:length(levels(contrasts$predicate))) {
+  current_pred = levels(contrasts$predicate)[i]
+  for(j in 1:length(levels(contrasts$contrast))) {
+    current_op = levels(contrasts$contrast)[j]
+    contrasts_wide[current_pred, current_op] <- subset(contrasts,
+                                                predicate == current_pred &
+                                                  contrast == current_op)$difftype[1]
+  }
+}
+
[…]

### 2024-02-21 writing/SuB-paper/projection-operators-SuB.tex
tracking changes from overleaf

@@
 \usepackage{subcaption}
+\usepackage{booktabs}
 
@@
 \author{Hofmann \and de Marneffe  \and Tonhauser}
-%\title{Projection variability of clausal complements across different operators}
-\title{Projection variability: Is the family of sentences really a family?}
+%\title{Projection variation of clausal complements across different operators}
+\title{Projection variation: Is the family of sentences really a family?}
 \makeatletter
@@
 	\usepackage{xcolor}
-    \definecolor{op-n-color}{HTML}{56B4E9}
-    \definecolor{op-m-color}{HTML}{E69F00}
-    \definecolor{op-c-color}{HTML}{009E73}
-    \definecolor{op-q-color}{HTML}{0072B2}
+    \definecolor{op-n-color}{HTML}{009E73}
+    \definecolor{op-m-color}{HTML}{0072B2}
+    \definecolor{op-c-color}{HTML}{404040}
+    \definecolor{op-q-color}{HTML}{E69F00}
     \definecolor{pred-fcv-color}{HTML}{CC79A7}
@@
 
-{\large \textbf{Projection variability: Is the family of sentences really a family?}}\footnote{We thank Taylor Mahler for assistance in collecting the data presented here as well as valuable comments.
-We gratefully acknowledge the National Science foundation grant \#1452674 (to Marie-Catherine de Marneffe, Craige Roberts, and Judith Tonhauser), which provided financial support for this project.}\\
+{\large \textbf{Projection variation: Is the family of sentences really a family?}}\footnote{We thank Taylor Mahler for assistance in collecting the data presented here as well as valuable comments.
+We gratefully acknowledge the National Science Foundation grant BCS \#1452674 (to Marie-Catherine de Marneffe, Craige Roberts, and Judith Tonhauser), which provided financial support for this project.}\\
 Lisa HOFMANN --- \textit{University of Stuttgart}\\
@@
 
-\textbf{Abstract.} We present experimental evidence \textbf{(i)} that the projection of the content of clausal complements  varies across entailment-canceling operators, \textbf{(ii)} that the effect of operator varies between clause-embedding predicates, and \textbf{(iii)} we extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factives predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
-The observed variability is not captured by existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
-Our results suggest that an analysis must consider interactions between predicates and operators and raise important questions for future research on projection.
-
-\textbf{Keywords:} Projection variability, entailment-canceling operators, (non)factive predicates.
+\textbf{Abstract.}
+    Under the `family of sentences' diagnostic for projection, the projection of content is investigated by embedding the expression that contributes the content in the scope of negation, polar questions, epistemic possibility modals, and conditional antecedents.
+This paper reports on the results of a set of experiments designed to investigate whether there is variation in the projection of content from under these four types of entailment-canceling operators.
+The contents investigated are the contents of the complements of 20 English clause-embedding predicates.
+The results of the experiments suggest (i) that the by-operator variation is small when aggregating over the 20 contents, but (ii) that the effect of operator differs between the clause-embedding predicates.
+The results of these experiments also extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factive predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
+The observed by-predicate and by-operator variation is not captured by existing theoretical accounts of projection (e.g., \citealt{heim_projection_1983,van_der_sandt_presupposition_1992,abrusan_predicting_2011,schlenker_triggering_2021}).
+Our results suggest that an empirically adequate projection analysis must consider interactions between predicates and operators.
+
+\textbf{Keywords:} Projection variation, entailment-canceling operators, (non)factive predicates.
 
 \section{Introduction} \label{sec:intro}
-    The so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective in formal semantics/pragmatics (e.g., \citealt{chierchia_meaning_1990}).
-For instance, the clausal complement of \emph{discover} is diagnosed as projective content, if its content (here: that Julian dances salsa) follows not just from \ref{ex:simple}, but also from the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.
-
-    % In formal semantics/pragmatics, the so-called `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}).
-For instance, to diagnose the content of the clausal complement of \emph{discover} as projective content, a researcher would typically observe that the content that Julian dances salsa follows not just from \ref{ex:simple}, but also from the family-of-sentences variants of \ref{ex:simple} in \ref{ex:family}, where \ref{ex:simple} has been embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, polar questions \ref{ex:q}, epistemic possibility modals \ref{ex:mod}, or conditional antecedents \ref{ex:cond}.
+    The `family of sentences' diagnostic is the standard way of diagnosing whether a content is projective (e.g., \citealt{chierchia_meaning_1990}).
+For instance, in \ref{ex:simple}, the content of the clausal complement of \emph{discover} (that Julian dances salsa) is diagnosed as projective content, if it is typically implied not just by an utterance of \ref{ex:simple}, but also by utterances of the variants in \ref{ex:family}, where \ref{ex:simple} is embedded under an entailment-canceling operator, such as negation \ref{ex:neg}, a polar question \ref{ex:q}, an epistemic possibility modal \ref{ex:mod}, or in a conditional antecedent \ref{ex:cond}.
 
@@
 
-    Some research, however, suggests that entailment-cancelling operators may affect projection differentially.
+    Some research, however, suggests that entailment-canceling operators may affect projection differentially.
 For instance, \citet{karttunen_observations_1971} proposed distinguishing English factive predicates (e.g., \textit{regret}) from semi-factives (e.g., \textit{discover}).
@@
 In contrast, the content of appositive relative clauses and the preparatory content of \textit{win} showed the opposite pattern, and the existential presupposition of clefts showed no difference.
-    %However, their task involved asking participants how surprised they would be to learn the content under investigation after observing the utterance.
-Since other factors besides speaker commitment modulate discourse (un)expectedness (see e.g., \citealt{zimmermann_grammatical_2011,tonnis_german_2021}), it is unclear that they (only) measured projection.
 
     Second, \citet{sieker_projective_2022}  compared the projection of the CCs of German factives (\textit{wissen} `know', \textit{bereuen} `regret', \textit{ent\-hüllen} `reveal') and semi-factives (\textit{bemerken} `notice', \textit{entdecken} `discover', \textit{herausfinden} `find out') from under the four operators in \ref{ex:family}.
-The results suggest that the CCs project more from under negation than from under the other three operators.
+Their results suggest that the CCs project more from under negation than from under the other three operators.
 Contrary to what \citet{karttunen_observations_1971} suggested, a comparison of the factive and semi-factive predicates did not reveal that the CCs of factive predicates project more from under polar questions, modals, or antecedents of conditionals than the CCs of semi-factive predicates.\footnote{The factive/semi-factive distinction is also called into question by naturally occurring examples where the CCs of factive predicates do not project from under the four operators (see \citealt{beaver_have_2010,de_marneffe_did_2012,de_marneffe_commitmentbank_2019}).
-For experimental research on the distinction, see \citealt{djaerv-etal2016} and \citealt{djarv_cognitive_2018}.
+For experimental research on the distinction, see \citealt{djarv_cognitive_2018}.
     }
-
-    %They replicated \poscite{smith_relationship_2014} result that the CC of \textit{know} projects more from under negation than conditionals for German \textit{wissen} (`know').
-However, this finding was part of an overall pattern of higher projection ratings with negation than other operators.
-    %
-    %Grouping ratings for the four factive and semi-factive predicates, respectively, they found no interaction of operator with predicate type, and thus no evidence for the factive/semi-factive distinction (iii), or an operator/predicate interaction (ii).
-However, their analysis did not explore the impact of operators on each predicate individually.
 
@@
 
-    The results of our study suggest that the projection of the CCs of these 20 predicates is affected differently by the four entailment-cancelling operators in \ref{ex:family},  but not in a way that is consistent with \poscite{karttunen_observations_1971} factive/semi-factive distinction.
+    The results of our investigation suggest that the projection of the CCs of these 20 predicates is affected differently by the four entailment-canceling operators in \ref{ex:family},  but not in a way that is consistent with \poscite{karttunen_observations_1971} factive/semi-factive distinction.
 The results also replicate a result of \citealt{degen_are_2022}, namely that projection from under polar questions does not categorically distinguish factive and non-factive predicates.
@@
 \section{Experiments}\label{s2}
-	To assess the effect of entailment-cancelling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
+	To assess the effect of entailment-canceling operator and clause-embedding predicate on projection, we collected projection judgments for the CCs of the 20 clause-embedding predicates in four sets of experiments.
     %
     The predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal {\em perhaps} in Exps.~3, and in conditional antecedents in Exps.~4.
-Each set of experiments contained three experiments, that differed in an at-issueness measure used in a separate block.
-In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}
-
-	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,lorson_influence_2018,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019,sieker_projective_2022}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}; and discussion in \citealt{tonhauser_how_2018}.}
-    %This diagnostic was also used in \citealt{sieker_projective_2022}.
+Each set of experiments contained three experiments that differed in the at-issueness measure that was used in a separate block.
+In this paper, we limit our attention to the projection ratings collected in these twelve experiments.\footnote{\label{footnote-supplement}The experiments, data and analysis scripts, as well as the supplements referred to in this paper can be found in the following GitHub repository: \url{https://github.com/judith-tonhauser/CommitmentBankPlus}.}
+
+	In all twelve experiments, projection was measured with the `certain that' diagnostic, which has been used to measure projection with both polar interrogative and declarative sentences (see, e.g., \citealt{tonhauser_prosodic_2016,djarv_prosodic_2017,stevens_rational_2017,tonhauser_how_2018,mahler_does_2019,mahler_social_2020,de_marneffe_commitmentbank_2019,sieker_projective_2022}).\footnote{For other diagnostics of projection see, e.g.\ \citealt{smith_projection_2011,xue_correlation_2011}, and \citealt{tonhauser_toward_2013}, and discussion in \citealt{tonhauser_how_2018}.}
     Under this diagnostic, participants are presented with utterances like those in \Next, and asked to rate whether the (named) speaker is certain of the CC.
@@
 
-    We assume, following \citet{tonhauser_how_2018} and \citet{degen_are_2022}, that judgments of speaker certainty about the embedded content reflect speaker commitment to that content, that is, projection.
+    We assume, following \citealt{tonhauser_how_2018} and \citealt{degen_are_2022}, that judgments of speaker certainty about the embedded content reflect speaker commitment to that content, that is, projection.
 If a participant interprets utterances like \Last[a--d] in a way that the speaker (here, Christopher) is certain of the CC, the CC is assumed to project.
@@
 
-    %This setup allows us to test the three hypotheses outlined in Section \ref{sec:intro}.
-First, a main effect of operator on speaker certainty judgments for embedded contents would provide evidence for by-operator variation (i).
-Second, if there is a operator/predicate interaction (ii), we would expect the effect of operator to be different (in magnitude or direction) for different predicates.
-Finally, based on the factive/semi-factive distinction (iii), we would expect the CC of semi-factive predicates to receive high projection ratings under negation and lower ratings under the other operators, while the CC of factive predicates would consistently receive relatively high projection ratings throughout.
-
-	\subsection{Methods}
-
-		%\subsubsection{Participants}
     \paragraph{Participants.}
@@
 These participants were required to reside in the US, to be born in the US, to have English as their first language, and to have an approval rating of at least 99\%.
-See Supplement \ref{app:d-participants} (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.
+See Supplement D (in the repository linked to in footnote \ref{footnote-supplement}) for further information about the participants.
 
@@
 				\centering
+                    %\fbox{\includegraphics[scale=0.5]{task-1n-proj.eps}}
 				\fbox{\includegraphics[width = .6\linewidth]{task-1n-proj.eps}}
-				\caption{A sample trial in Exps.~1. In the corresponding trials in the other experiments, participants were presented with an utterance with a different entailment-cancelling operator.}
+				\caption{A sample trial from Exps.~1 (negation). In the other experiments, participants were presented with utterances with a different entailment-canceling operator.}
 				\label{fig:trial}
@@
         \paragraph{Materials.}
-            The target sentences consisted of the 400 combinations of the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement \ref{app:a-clauses}).
+            The target sentences consisted of the 400 combinations of the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} with 20 embedded clauses (provided in Supplement A).
 As mentioned above, the predicates were embedded under negation in Exps.~1, under polar questions in Exps.~2, under the epistemic possibility modal \emph{perhaps} in Exps.~3, and in conditional antecedents in Exps.~4, for a total of 400 target stimuli in each of the four sets of experiments.
 To assess whether participants were attending to the task, each experiment included six control stimuli.
-For details on the six control stimuli, see Supplement \ref{app:c-control}.
-
-            Each participant saw a random set of twenty-six stimuli: Each set contained one target stimulus for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same six control stimuli.\footnote{Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
+For details on the six control stimuli, see Supplement C.
+
+            Each participant saw a random set of 26 stimuli: Each set contained one target stimulus for each of the 20 clause-embedding predicates (each with a unique complement clause) and the same six control stimuli.\footnote{Each participant saw their set of 26 stimuli twice, once in the projection block and once in the at-issueness block.
 Block order was randomized.
-As mentioned above, we focus on the projection ratings here.} Trial order was randomized.
+As mentioned above, we focus here on the projection ratings.} Trial order was randomized.
 
@@
             %
-            On each trial, they read an utterance and gave a response to the `certain that' question on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) on the other.
-A sample trial from Exps.~1 is shown in Figure~\ref{fig:trial}.
-
-            Following \citet{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things: On the one hand, higher ratings could reflect greater speaker commitment towards the CC, and therefore greater projection.
+            On each trial, they read an utterance and gave a response to the `certain that' question on a slider marked `no' (coded as 0) at one end and `yes' (coded as 1) at the other.
+A sample trial %from Exps.~1
+            is shown in Figure~\ref{fig:trial}.            
+            Following \citealt{tonhauser_how_2018}, higher ratings of speaker certainty could reflect one of two things.
+First, higher certainty ratings could reflect greater speaker commitment towards the CC, and therefore greater projection.
 This assumes that speaker commitment is interpreted in a gradient way.
-            % gradient.
-            Conversely, higher certainty ratings could reflect the probability that an interpreter takes the speaker to be committed to the CC.
-On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of the uncertainty about possible interpretations.
-In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property, though our discussion of projection variability will be in line with the first interpretation (for discussion, see also \citealt{grove_factivity_2023}).
+Second, higher certainty ratings could reflect a higher probability that an interpreter takes the speaker to be committed to the CC.
+On this interpretation, speaker commitment may be a binary, categorical property and projection variation is a result of uncertainty about speaker commitment.
+In this paper, we remain agnostic about the underlying interpretation of projectivity as a gradient property (for discussion, see  \citealt{grove_factivity_2023}).
 
@@
         %\subsubsection{Data exclusion}
-        \noindent\textbf{Data exclusion}
-            was done based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}. The data from 2,682 participants entered into the analysis.
+        \noindent\textbf{Data exclusion.} Data were excluded based on self-declared non-native speaker status and other criteria given in Supplement D.
+The data from 2,682 participants entered into the analysis.
             % Data were excluded based on self-declared non-native speaker status and other criteria given in the Supplements \ref{app:d-participants}.
@@
 
-    \subsection{Results and discussion}
-
-        We first address the question of whether there is by-operator variation (section \ref{s:by-op}) and then the question of whether there is by-predicate variation in the observed by-operator variation (section \ref{s:by-pred}).
-Finally, in section \ref{s:factivity}, we relate the results to the result of \citealt{degen_are_2022}, that projection from under polar questions does not categorically distinguish factive and non-factive predicates.
-
-		%Our data reveals three key results: \textbf{(i)} There is projection variability by operator; \textbf{(ii)} There is by-predicate variation in the effect of operator on projection; and \textbf{(iii)} We find further support (from entailment-cancelling operators beyond polar questions) for \posscite{degen_are_2022} result that by-predicate projection variation is gradient in a way that cannot be predicted by categorical lexical classes.
-
-	\subsubsection{By-operator variation}\label{s:by-op}
+\section{Results and discussion}
+
+        We first address by-operator variation (Section \ref{s:by-op}) and then the question of whether there is by-predicate variation in the observed by-operator variation (Section \ref{s:by-pred}).
+Finally, in Section \ref{s:factivity}, we relate our results to those of \citealt{degen_are_2022}, that projection from under polar questions does not categorically distinguish factive and non-factive predicates.
+
+	\subsection{By-operator variation}\label{s:by-op}
 
         Figure~\ref{fig:op-ratings} shows the mean certainty ratings by entailment-canceling operator, aggregating over the clause-embedding predicates.
-As shown, there is projection variability by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.
+As shown, there is projection variation by operator: The CCs of the clause-embedding predicates were relatively more projective when embedded in the antecedent of a conditional than in a polar interrogative, where they were relatively more projective than when they were embedded under negation or the epistemic modal {\em perhaps}.
 
@@
 				\centering
-				\includegraphics[scale = .8]{certainty-operator.pdf}
-				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals and violin plots indicate the kernel probability density of the ratings.
-                % of the individual participants’ ratings.
+				\includegraphics[scale = .7]{certainty-operator.pdf}
+				\caption{Mean certainty ratings by operator. Error bars indicate $95\%$ bootstrapped confidence intervals. Violin plots indicate the kernel probability density of participants' individual ratings.
                 }
@@
 
-        These generalizations are supported by a post-hoc pairwise comparison of the estimated means for each entailment-canceling operator using the \texttt{emmeans} package (\citealt{emmeans}) in R (\citealt{r}).
+        These observations are supported by a post-hoc pairwise comparison of the estimated means for each entailment-canceling operator using the \texttt{emmeans} package (\citealt{emmeans}) in R (\citealt{r}).
 The input to the pairwise comparison was a Bayesian mixed-effects beta regression model that was fit using the \texttt{brms} package (\citealt{buerkner2017}) with weakly informative priors.
-The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson_better_2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level) and included a random by-predicate intercept.\footnote{See Supplement \ref{app:analysis-1} for more details on the model.} The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators.
+The model predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson_better_2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level) and included a random by-predicate intercept.\footnote{See Supplement E for details on the model.} The output of the pairwise comparison were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators.
 We assume that two operators differ in certainty ratings if the HDI of their pairwise comparison does not include 0.
 
-        Table \ref{t:pairwise} provides the output of the pairwise comparison on a logit scale.
-        %\footnote{While the certainty ratings collected from the participants range from 0 to 1, the smaller than 0 estimated marginal mean differences are the result of the beta regression using a logit link function for the mean parameter.}
-        As shown, the analysis suggests differences between each pair of operators, though the estimated marginal mean differences differ across the pairs.
-That is, certainty ratings are higher for CCs embedded in conditional antecedents than those embedded under polar questions; certainty ratings for CCs embedded under polar questions are higher than those for CCs embedded under negation; and finally certainty ratings for CCs embedded under negation are higher than those for CCs embedded under the epistemic modal \emph{perhaps}.
+        Table~\ref{t:pairwise} provides the output of the pairwise comparison on a logit scale.
+        As shown, the analysis suggests differences between each pair of operators.
+That is, certainty ratings are higher for CCs embedded in conditional antecedents than those embedded under polar questions, certainty ratings for CCs embedded under polar questions are higher than those for CCs embedded under negation, and finally certainty ratings for CCs embedded under negation are higher than those for CCs embedded under the epistemic modal \emph{perhaps}.
 
@@
         \begin{tabular}{lrrr}
-         contrast & estimate & lower 95\% CI & upper 95\% CI \\\hline 
+        \toprule
+         contrast & estimate & lower 95\% CrI & upper 95\% CrI \\\midrule
          conditional - negation & 0.21 & 0.18 & 0.24 \\ 
@@
           modal - question & -0.14 & -0.17 & -0.11 \\ 
-          negation - question & -0.09 & -0.12 & -0.06 \\ 
+          negation - question & -0.09 & -0.12 & -0.06 \\ \bottomrule
         \end{tabular}
         \caption{Output of the pairwise comparison of entailment-canceling operators.
-The `contrast' column identifies the pair of entailment-canceling operators, the `estimate' column the estimated marginal mean difference, and the `lower 95\% CI' and `upper 95\% CI' columns provide the lower and upper bounds of the HDIs.}\label{t:pairwise}
+The `contrast' column identifies entailment-canceling operators pairs, `estimate' the estimated marginal mean difference, and `lower/upper 95\% CrI' provide the lower/upper bounds of the HDIs.}\label{t:pairwise}
         \end{table}
 
- These results suggest that certainty ratings for the CCs of of the English clause-embedding predicate we investigated vary by entailment-canceling operator.
-In contrast to \citet{sieker_projective_2022} for German, the results of our experiments do not suggest that projection is strongest from under negation.
-Recall, however, that they only investigated projection of the CCs of (semi-)factive predicates.
-Since there is by-predicate variation in the effect of entailment-canceling operator on projection (as we show in the next section), this difference between the results of their experiment and ours might very well be due to the types of predicates investigated.
+        These results suggest that certainty ratings for the CCs of the English clause-embedding predicates we investigated vary by entailment-canceling operator.
+In contrast to \citealt{sieker_projective_2022} for German, the results of our experiments do not suggest that projection is strongest from under negation.
+Recall, however, that they only investigated projection of the CCs of (semi-) factive predicates.
+Since there is by-predicate variation in the effect of entailment-canceling operator on projection (as we show in the next section), this difference between the results of their experiment and ours might be due to the types of predicates investigated.
 Finally, the differences in mean certainty ratings between the four entailment-canceling operators are very small.
@@
 
-	\subsubsection{By-predicate variation in the effect of entailment-canceling operator}\label{s:by-pred}
+	\subsection{By-predicate variation in the effect of entailment-canceling operator}\label{s:by-pred}
 
         Figure~\ref{fig:op-pred-analysis} shows mean certainty ratings by entailment-canceling operator for the 20 predicates, with predicates ordered by their overall mean certainty rating.
-        %
-        As shown, there is by-operator projection variation for the CCs of all 20 clause-embedding predicate, that is, there is not a single predicate for which there is no by-operator variation.
+As shown, there is by-operator projection variation for each predicate.
 Further, the effect of operator differs between predicates.
-Examining the observed variation suggests that it does not align with widely assumed predicate classes.
+For instance, the five (semi-)factive predicates exhibit four different patterns.
+        %
+        First, the CC of \emph{be annoyed} projects
+        most from under questions, less from under negation, followed by conditionals, and least from under the modal \emph{perhaps} (\textcolor{op-q-color}{Q} $>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-c-color}{C} $>$ \textcolor{op-m-color}{M}).
+        % more from under negation and questions than conditionals and the modal {\em perhaps};
+        Second, the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under {\em perhaps} (\textcolor{op-q-color}{Q} $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-c-color}{C}$\} >$ \textcolor{op-m-color}{M}).
+        %
+        The CCs of \emph{discover} and \emph{see} exhibit a third pattern: They project most from under questions and conditionals, less from under negation, and least from under \emph{perhaps} ($\{$\textcolor{op-q-color}{Q}$,$ \textcolor{op-c-color}{C}$\}>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-m-color}{M}).
+        %
+        Finally, the CC of \emph{reveal} projects most from conditionals, less from questions, and least from under negation and \emph{perhaps} (\textcolor{op-c-color}{C} $>$ \textcolor{op-q-color}{Q}  $> \{$\textcolor{op-m-color}{M}$, $\textcolor{op-n-color}{N}$\}$).
+        %
+        Thus, the CCs of none of our (semi-)factive predicts project uniformly from under all four entailment-canceling operators (contrary to what \citealt{karttunen_observations_1971} suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
 
@@
     		\caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, non-factive) and operator (\textcolor{op-m-color}{modal},  \textcolor{op-n-color}{negation}, \textcolor{op-q-color}{polar questions}, \textcolor{op-c-color}{conditional antecedents}) with 95\% bootstrapped confidence intervals.
-Violin plots indicate kernel probability density of the individual participants' ratings.
-Facets for predicates are ordered by the predicate mean certainty rating (aggregating across operators).
+Violin plots indicate kernel probability density of individual participants' ratings.
+Predicate facets are ordered by the predicate mean certainty rating (aggregating across operators).
 Below each facet, a line spanning two operators indicates a non-zero difference according to the pairwise comparison of operators.
-The linetype indicates whether the difference is
-            at least $1$ (solid line: —),
-            below $1$ but at least $0.5$ (dashed line: – –),
-            or below $0.5$ but above $0$ (dotted line: . . . )}
+The line type indicates whether the difference $d$ is $\geq 1$ (solid line: —), $0.5 \le d < 1$ (dashed line: – –), $0 \le d < 0.5$ (dotted line: .
+.
+.
+).}
+            %at least $1$ (solid line: —),
+            %below $1$ but at least $0.5$ (dashed line: – –),
+            %or below $0.5$ but above $0$ (dotted line: . . . )}
     		\label{fig:op-pred-analysis}
@@
 
-        For instance, the five (semi-)factive predicates exhibit four different patterns:
-        %
-        First, the CC of \emph{be annoyed} projects
-        most from under questions, less from under negation, followed by conditionals, and least from under the modal \emph{perhaps} (\textcolor{op-q-color}{Q} $>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-c-color}{C} $>$ \textcolor{op-m-color}{M}).
-        % more from under negation and questions than conditionals and the modal {\em perhaps};
-        Second, the CC of \emph{know} projects most from under questions, less from conditionals and negation, and least from under {\em perhaps} (\textcolor{op-q-color}{Q} $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-c-color}{C}$\} >$ \textcolor{op-m-color}{M}).
-        %
-        The CCs of \emph{discover} and \emph{see} exhibit a third pattern: They project most from under questions and conditionals, less from under negation, and least from under \emph{perhaps} ($\{$\textcolor{op-q-color}{Q}$,$ \textcolor{op-c-color}{C}$\}>$ \textcolor{op-n-color}{N} $>$ \textcolor{op-m-color}{M}).
-        %
-        Finally, the CC of \emph{reveal} projects most from conditionals, less from questions, and least from under negation and \emph{perhaps} (\textcolor{op-c-color}{C} $>$ \textcolor{op-q-color}{Q}  $> \{$\textcolor{op-n-color}{N}$,$ \textcolor{op-m-color}{M}$\}$).
-        %
-        Thus, contrary to what \citealt{karttunen_observations_1971} proposed, there is no predicate whose CC projects uniformly from under all four entailment-canceling operators (as he suggested for factive predicates) and the purported semi-factive predicates  \emph{discover} and \emph{reveal} do not project more from under negation than the other three entailment-canceling operators.
-
-        We also observe by-operator projection variation for non-factives predicates.
-Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CC of \emph{acknowledge} exhibit the same pattern as that of the CC of \emph{discover}.
-Other non-factives predicates exhibit other patterns: For instance, the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the antecedents of conditionals than the other three operators.
-
-        These generalizations are supported by the statistical analysis.
-We fitted Bayesian mixed-effects beta regression models for each the by-predicate subset of the data containing (using \emph{brms}, \citealt{burkner_brms_2017}; in R, \citealt{r}).
-Each of the 20 models used low-information priors and estimated the effect of the \lq operator\rq\ (fixed effect, treatment-coded with reference-level \lq modal\rq) on (scaled) certainty ratings, with by-item random intercepts and random operator slopes.\footnote{Further details about the models, their outputs, and the analysis are provided in Supplement \ref{app:analysis-2}.} For each predicate, the differences between operators were established by a pairwise comparison of posterior estimates of marginal means for the four operators (using \emph{emmeans}, \citealt{lenth_emmeans_2024}).
-Here we report those differences between posterior samples where the $95\%$ HDI does not include $0$.
-These are indicated in Figure~\ref{fig:op-pred-analysis}, where non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet.
-
-        More generally, we found by-operator differences for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
-        %
-        Therefore, our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator.
+        We also observe by-operator projection variation for non-factive predicates.
+Some of this variation aligns with that observed for factive predicates: For instance, the CC of \emph{inform} exhibits the same pattern as the CC of \emph{know}, and the CCs of \emph{hear} and \emph{see} the same pattern as that of \emph{discover}.
+Other non-factive predicates exhibit other patterns: the CCs of \emph{admit}, \emph{confess} and \emph{announce} project most from the conditional antecedents than the other three operators.
+
+        These observations are supported by post-hoc pairwise comparisons of the estimated means for each entailment-canceling operator within each predicate using the \texttt{emmeans} package (\citealt{emmeans}).
+The input to the pairwise comparisons for each predicate were 20 Bayesian mixed-effects beta regression models that were fit using the \texttt{brms} package (\citealt{buerkner2017}) with weakly informative priors.
+The models for each predicate predicted certainty ratings\footnote{To model the certainty ratings using a beta regression, the ratings were first transformed from the interval [0,1] to the interval (0,1) using the method proposed in \citealt{smithson_better_2006}.} from a fixed effect of entailment-canceling operator (with treatment coding and `modal' as the reference level) and included a random by-item intercept and a random slope for operator by item.\footnote{See Supplement F for further details on the models.} The output of the pairwise comparisons were 95\% highest density intervals (HDIs) of estimated marginal mean differences between each of the operators for each predicate.
+We assume that two operators differ in certainty ratings for a given predicate if the HDI of their pairwise comparison does not include 0.
+These non-zero differences between two operators are indicated by lines spanning the two operators below each predicate facet in Figure~\ref{fig:op-pred-analysis}.
+
+        %More generally, we found by-operator differences for the CC of each of the 20 clause-embedding predicates, and that the effect of operator differs between predicates.
+        %
+        Our findings align with those of \citealt{smith_relationship_2014}, who also observed by-expression variation in the effect of operator.
 However, while they found that the CC of \emph{know} projects more from under negation than the antecedent of a conditional, we did not find a difference here.
 We hypothesize that this difference is due to the difference in projection diagnostic used.
-
-        Our results, however, differ from those of \citealt{sieker_projective_2022}.
-While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggested five different patterns of by-operator variation for the five (semi)factive predicates.
-As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the diagnostic.
-Other factors that varied between our experiments are the language under investigated (German vs.\ English), the clause-embedding predicates investigated, and the CCs that the predicates were paired with.
+Our results differ, however, from those of \citealt{sieker_projective_2022}.
+While their work did not find differences in by-operator projection variation between factive and semi-factive predicates, our results suggest four different patterns of by-operator variation for the five (semi)factive predicates.
+As \citealt{sieker_projective_2022} also used the `certain that' diagnostic for projection, this difference in results is not likely due to the projection diagnostic.
+Other factors that varied between our experiments are the language under investigation (German vs.\ English), the clause-embedding predicates investigated, and the CCs that the predicates were paired with.
 Future research will need to establish which of these factors are implicated in the observed differences.
@@
 
-	\subsubsection{Factive vs.\ non-factives predicates}\label{s:factivity}
-
-        Lexical approaches to projection assume that factive predicates are ones that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} Because presuppositions are assumed to project from under entailment-canceling operators, this definition predicts that factive predicates are distinguished from non-factives by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives.
-
-        % Factive predicates are defined as predicates that presuppose the CC, while the CC of non-factives predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.} Because presuppositions are assumed to project from under entailment-canceling operators, this definition leads one to expect that factive predicates are distinguished from non-factives ones by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives ones.
-
+	\subsection{Factive vs.\ non-factive predicates}\label{s:factivity}
+
+        Lexical approaches to projection assume that factive predicates are ones that presuppose the CC, while the CC of non-factive predicates is not presupposed (e.g., \citealt{kiparsky_fact_1970,karttunen_observations_1971,schlenker_local_2010,abrusan_predicting_2011}).\footnote{Some of these works additionally assume that the CC of factive predicates is entailed.
+For details see \citealt{degen_are_2022}.} Because presuppositions are assumed to typically project from under entailment-canceling operators, this definition predicts that factive predicates are distinguished from non-factives by the projection of their CCs: The CCs of factive predicates are expected to be categorically more projective than those of non-factives.
 This expectation was investigated in \citealt{degen_are_2022} based on the 20 clause-em\-bed\-ding predicates in \ref{ex:predicates} embedded in polar questions.
-Contrary to expectation, \posscite{degen_are_2022} Exps.~1 found that the CCs of the five (semi-)factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives were as projective, or even more projective, than those of some factive predicates.
-In short, projection of the CC from under polar questions did not categorically distinguish factive from non-factives predicates.
-Further support for this result came from the analysis of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{de_marneffe_commitmentbank_2019}), the VerbVeridicality dataset (\citealt{ross_how_2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).
+Contrary to expectation, \poscite{degen_are_2022} Exps.~1 found that the CCs of the five (semi-)factive predicates varied in projection, that the CCs of the 15 non-factive predicates were projective compared to the non-projective main clause contents, and that the CCs of some non-factives were as projective, or even more projective, than those of some factive predicates.
+In short, projection of the CC from under polar questions did not categorically distinguish factive from non-factive predicates.
+Further support for this result came from analyses of projection ratings in three additional datasets, namely the CommitmentBank (\citealt{de_marneffe_commitmentbank_2019}), the VerbVeridicality dataset (\citealt{ross_how_2019}), and the MegaVeridicality  dataset (\citealt{white_role_2018}).
+
+        The results of the experiments reported on in this paper replicate \poscite{degen_are_2022} result.
+As shown in Figure~\ref{fig:op-pred-ratings-q}, there is variation between the five \textcolor{pred-fcv-color}{factive} predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from non-factive predicates.
+Furthermore, our results suggest that this result can be extended to the three other entailment-canceling operators.
+As shown in Figures~\ref{fig:op-pred-ratings-m}-\ref{fig:op-pred-ratings-c}, there is variation in the projection of the CCs of the five factive predicates from under \emph{perhaps}, negation, and conditional antecedents, and projection ratings in these conditions do not show a categorical difference between factive and non-factive predicates either.
+These results lend further support to the conclusion of \citealt{degen_are_2022} that there is, to date, no empirical evidence for a coherent class of factive predicates.
+
+        %Instead, we see a gradient  effect of predicate on projection ratings in all four conditions.
 
@@
             \centering
-            \begin{subfigure}[t]{0.57\textwidth}
+            \begin{subfigure}[t]{0.49\textwidth}
             \caption{Polar question}
@@
             \end{subfigure}
-
-            \vspace{-\baselineskip}
+            %
+            %\vspace{-\baselineskip}
             %\medskip
-            \begin{subfigure}[t]{0.57\textwidth}
+            \begin{subfigure}[t]{0.49\textwidth}
             \caption{Modal \emph{perhaps}}
@@
             %\medskip
-            \begin{subfigure}[t]{0.57\textwidth}
+            \begin{subfigure}[t]{0.49\textwidth}
             \caption{Negation}
@@
             \end{subfigure}
-
-            \vspace{-\baselineskip}
+            %
+            %\vspace{-\baselineskip}
             %\medskip
-            \begin{subfigure}[t]{0.57\textwidth}
+            \begin{subfigure}[t]{0.49\textwidth}
             \caption{Conditional}
@@
             \vspace{-\baselineskip}
-            \caption{Mean certainty ratings by predicate (\textcolor{pred-fcv-color}{(semi-)factive}, factive) for (a) polar questions, (b) the modal \emph{perhaps}, (c) negation, (d) conditional antecedents.
+            \caption{Mean certainty ratings by predicate, with (semi-)factive predicates in \textcolor{pred-fcv-color}{violet}, for (a) polar questions, (b) the modal \emph{perhaps}, (c) negation, (d) conditional antecedents.
 Error bars indicate $95\%$ bootstrapped confidence intervals.
-Violin plots indicate kernel probability densities of observations.}\label{fig:op-pred-ratings2}
+Violin plots indicate kernel probability densities of the individual participants’ ratings.}\label{fig:op-pred-ratings2}
         \end{figure}
-
-        The experiments reported on in this paper replicate this result.
-As shown in Fig.~\ref{fig:op-pred-ratings-q}, there is variation between the CCs of the five factive predicates in the polar question condition, and projection from under polar questions does not categorically distinguish factive from non-factive predicates.
-Furthermore, our results suggest that this result can be extended to the three other entailment-canceling operators as well.
-As shown in Figs.~\ref{fig:op-pred-ratings-m}-\ref{fig:op-pred-ratings-c}, there is variation in the projection of the CCs of the five factive predicates from under \emph{perhaps}, negation, and conditional antecedents, and projection ratings in these conditions do not show a categorical difference between factive and non-factive predicates here, either.
-Instead, we see a gradient  effect of predicate on projection ratings in all four conditions.
-These results lend further support to the conclusion of \citealt{degen_are_2022} that there is, to date, no empirical evidence for a coherent class of factive predicates.
 
@@
 
-	\subsubsection{Summary}
+	\subsection{Summary}
 
@@
 Our experiments suggest that projection of the CC from under negation, the antecedent of conditionals and epistemic modals also do not provide empirical support for a natural class of factive predicates.
-
-        Before discussing the methodological and theoretical implications of these results in section \ref{s:general}, we provide converging evidence from a different dataset in section \ref{s:converging}.
-
-\section{Converging evidence for the by-operator variation}\label{s:converging}
-
-    This section provides converging evidence for the by-operator variation based on the Mega\-Ve\-ri\-dicality dataset (\citealt{white_role_2018}).
+Before discussing the methodological and theoretical implications of these results in Section~\ref{s:general}, we provide converging evidence from a different dataset in Section~\ref{s:converging}.
+
+\section{Converging evidence for the by-predicate variation in the effect of operator}\label{s:converging}
+
+    We provide converging evidence for the by-predicate variation in the effect of operator based on the Mega\-Ve\-ri\-dicality dataset (\citealt{white_role_2018}).
 This dataset contains projection ratings for the CCs of 517 English clause-embedding predicates.
-The stimuli that participants rated consisted of combinations of these predicates with what White and Rawlins (2018) referred to as ‘low content arguments’, as shown in \Next for \emph{know}.
-The predicates were embedded under negation in stimuli like \Next[a], in the antecedent of a conditional and a polar question in stimuli like \Next[b], and under negation, in the antecedent of a conditional, and in a question in stimuli like \Next[c].
-To assess projection, participants were asked to respond to the question \emph{Did that thing happen?} for stimuli like \Next[a] and to respond to the question posed by stimuli like \Next[b] and \Next[c].
-The response options were ‘yes’, ‘maybe or maybe not’, and ‘no’.
+As shown in \Next for \emph{know}, the predicates were combined with (what the authors refer to as) ``low lexical content''.
+%Indeed, the stimuli that participants rated consisted of combinations of these predicates with what White and Rawlins 2018 referred to as ‘low content arguments’, as shown in \Next for \emph{know}.
+%The predicates were embedded under negation \Next[a], in the antecedent of a conditional and a polar question \Next[b], and under negation, in the antecedent of a conditional, and in a question \Next[c].
+To assess projection, participants were asked to respond to the question \emph{did that thing happen?} %for stimuli like \Next[a] and to respond to the question posed by stimuli like \Next[b] and \Next[c].
+    %The response options were ‘yes’, ‘maybe or maybe not’, and ‘no’.
+    The predicates were embedded under negation as in \Next[a], in the antecedent of a conditional as in \Next[b], and under negation in the antecedent of a conditional as in \Next[c].
+To assess projection, participants were asked to respond to the question \emph{did that thing happen?}.\footnote{In \ref{mvb} and \ref{mvc}, the participants responded to a question presented in the consequent of the conditional.
+Thus, the predicates in these two conditions may also be taken to be embedded in a polar question.} %for stimuli like \Next[a] and to respond to the question posed by stimuli like \Next[b] and \Next[c].
+    The response options were ‘yes’ (indicating projection), ‘maybe or maybe not’, and ‘no’ (no projection).
 
@@
 			\ex. \a. Somebody didn’t know that a particular thing happened. Did that thing happen?
-				\b. If somebody knows that a particular thing happened, did that thing happen?
-				\b. If somebody didn’t know that a particular thing happened, did that thing happen?
-				\z.
-			\z.
-
-    To investigate by-operator projection variability, we recoded `yes' responses as 1, `no' responses as -1, and `maybe or maybe not' responses as 0.
-We then calculated the mean projection rating for 16 predicates under the three types of operator combinations shown in \Last.
-The 16 predicates include ones also used in our experiment as purported factives (\textit{be annoyed, know, reveal}) and semi-factives (\textit{discover, see}), and eleven further predicates commonly characterized as factive or semi-factive (\textit{amuse, find out, forget, learn, love, notice, realize, recognize, regret, remember, understand}).
+				\b.\label{mvb} If somebody knows that a particular thing happened, did that thing happen?
+				\b.\label{mvc} If somebody didn’t know that a particular thing happened, did that thing happen?
+				\z.
+			\z.
+
+    To investigate by-predicate projection variation, we recoded the responses as 1 (`yes'), -1 (no), and 0 `maybe or maybe not').
+We calculated the mean projection ratings for 25 predicates under the three types of operator combinations shown in \Last.
+We used the 14 non-factive and 5 factive predicates from our experiments that are in MegaVeridicality (\textit{be right} is not included).
+As \citealt{djarv_cognitive_2018} suggested that the factive/semi-factive distinction can be understood as a difference between cognitive and emotive predicates, we included the emotive (\textit{disappoint}, \textit{surprise}) and cognitive (\textit{realize}, \textit{find out}) predicates from their experiments that are in MegaVeridicality.
+We also added two other predicates suggested by Karttunen as factive (\textit{regret}) and semi-factive (\textit{notice}).
+
+           	\begin{figure}
+				\centering
+				\includegraphics[width = \linewidth]{megaveridicality}
+				\caption{Mean projection ratings by entailment-canceling operator(s) and clause-embedding predicate in the MegaVeridicality dataset.}
+				\label{fig:megaverid}
+			\end{figure}
+
+    Figure \ref{fig:megaverid} shows mean projection ratings by embedding operator(s) and predicate.
+As shown, the effect of operator varies by predicate: For many predicates, ratings are (at least numerically) lower when embedded under negation, or under negation and in a conditional antecedent, than when embedded in a conditional antecedent.
+These observations suggest that there is by-predicate variation in the effect of entailment-canceling operator even when projection ratings are collected with a different measure and different materials than in our experiments.
 
@@
 
-	The mean projection ratings by embedding context and predicate are presented in Figure~\ref{fig:megaverid}, which shows that effect of embedding context on projection differs by predicate.
-
-	For instance,\footnote{Give some examples of differences here, but first double check if lines are labeled correctly. Also angle labels on x-axis.}
-
 	% the predicates \textit{see} and \textit{learn} have the same mean ratings for both negative and conditional contexts.
@@
 
-			\begin{figure}[ht]
-				\centering
-				\includegraphics[width = \linewidth]{megaveridicality}
-				\caption{Mean projection ratings by embedding context and predicate in MegaVeridicality.}
-				\label{fig:megaverid}
-			\end{figure}
-
-	It is worth noting that the task assesses global projection only in case of \Last[a], but not \Last[b+c], where the question \textit{did it happen?} is embedded in the conditional consequent.
+	%It is worth noting that the task assesses global projection only in case of \Last[a], but not \Last[b+c], where the question \textit{did it happen?}\ is embedded in the conditional consequent.
 It assesses whether the projective inference holds in the conditional local context.
-Nevertheless, the data provides supporting evidence for by-predicate variation in the effect of operators on projection.
-Crucially, the by-predicate variation in the effect of negation in the global context is different from the by-predicate variation in the effect of negation in the conditional context.
+
+    %Nevertheless, the data provides supporting evidence for by-predicate variation in the effect of operators on projection.
+%Crucially, the by-predicate variation in the effect of negation in the global context is different from the by-predicate variation in the effect of negation in the conditional context.
+    %Crucially, the by-predicate variation in the effect of negation alone is different from the by-predicate variation in the effect of negation in the conditional context.
 
@@
 
-    In this section, we point out methodological implications of our results (section \ref{s:method-impl}), discuss whether contemporary projection analyses can capture the observed variation (section \ref{s:analysis}), and speculate about lexical differences between the clause-embedding predicates that might predict the by-operator variation observed (section \ref{s:lex}).
+    We now point out methodological implications of our results (Section~\ref{s:method-impl}), discuss whether contemporary projection analyses can capture the observed variation (Section~\ref{s:analysis}), and speculate about lexical differences between the clause-embedding predicates that might predict the by-operator variation observed (Section~\ref{s:lex}).
 
@@
 
-    The results of the experiments reported on in section \ref{s2} suggest that there is little by-operator projection variation when aggregating observations for the CCs of the  clause-embedding predicates, but that there is by-operator variation that cannot be neglected when we do not aggregate.
+    The results of the experiments reported on in Section~\ref{s2} suggest that there is little by-operator projection variation when aggregating observations for the CCs of the  clause-embedding predicates, but that there is by-operator variation that cannot be neglected when we do not aggregate.
 These results have two methodological implications.
@@
 Further, results about projection variation should be relativized to the entailment-canceling operator under which the expressions and contents were embedded, as they may very well depend on this choice.
-For instance, two contents that exhibit projection variation when embedded under polar questions might not exhibit such differences from under a different entailment-canceling operator, and two contents that do not exhibit different projection across negation might exhibit variation when embedded under some other operator.
+For instance, two contents that exhibit projection variation when embedded under polar questions might not exhibit such differences with a different entailment-canceling operator, and two contents that do not exhibit projection variation from under negation might exhibit variation when embedded under some other operator.
 
@@
 
-    This section discusses whether contemporary analyses can capture the observed by-predicate projection variation and the interactions between predicate and entailment-canceling operator in projection ratings.
+    This section discusses whether contemporary analyses can capture the observed by-predicate projection variation and the interactions between predicate and entailment-canceling operator.
 	% Our results---that projection is modulated by entailment-canceling operators and that there is by-predicate variation in the effect of operator on projection---are not captured by contemporary projection analyses.
@@
 
-    On the analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992}, the CCs of factive predicates are lexically specified as presupposed, in contrast to the CCs of non-factives predicates.
+    On the analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992}, the CCs of factive predicates are lexically specified as presupposed, in contrast to the CCs of non-factive predicates.
 Presupposed CCs project to the global context, except when that would produce an inconsistency, in which case they are accommodated to the local context of the operator.
@@
 One reason is  variation among the factive predicates: For instance, when embedded under polar questions, negation, or \emph{perhaps,} the CC of \emph{be annoyed} is more projective than that of \emph{discover}, which in turn is more projective than that of \emph{reveal}.
-(In conditional antecedents, the order of \emph{be annoyed} and \emph{discover} is reversed, see Figure~\ref{fig:op-pred-ratings2}).
+(In conditional antecedents, the order of \emph{be annoyed} and \emph{discover} is reversed, see Figure~\ref{fig:op-pred-ratings2}.)
     %
     This variation is not predicted because the CCs of factive predicates are invariably specified as presupposed, with no mechanism to predict projection variation between factive predicates.
-(Recourse to local accommodation is not possible because the stimuli in our experiments were presented in minimal contexts that were not inconsistent with the CCs.)  A second reason is that the CCs of some non-factives predicates (e.g., \emph{inform, acknowledge}) are just as projective as those of some factive predicates, or even more projective.
-Analyses like those in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} fail to make predictions about the projection of the CCs of non-factives predicates.
+(Recourse to local accommodation is not possible because the stimuli in our experiments were presented in minimal contexts that were not inconsistent with the CCs.)  A second reason is that the CCs of some non-factive predicates (e.g., \emph{inform, acknowledge}) are just as projective as those of some factive predicates, or even more projective.
+Analyses like those in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} fail to make predictions about the projection of the CCs of non-factive predicates.
 Thus, as discussed in detail in \citealt{degen_are_2022}, the factive/non-factive distinction is not sufficiently fine-grained to adequately predict the observed by-predicate projection variation.
 
-    The analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} also do not predict differential effects of entailment-cancelling operators on projection.
+    The analyses in \citealt{heim_projection_1983} and \citealt{van_der_sandt_presupposition_1992} also do not predict differential effects of entailment-canceling operators on projection.
 This is because negation and conditional antecedents are given a semantics that derives their behavior as presuppositional holes, in the sense of \citealt{karttunen_observations_1971}.
@@
 
-    \citealt{abrusan_predicting_2011} and \citealt{simons_best_2017} do not assume that presuppositions are lexically specified.
-In \citealt{abrusan_predicting_2011}, a lexical entailment of a (possibly embedded) sentence is a presupposition, if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_1996,roberts_information_2012}) addressed by the utterance.
+    \citet{abrusan_predicting_2011} and \citet{simons_best_2017} do not assume that presuppositions are lexically specified.
+In \citealt{abrusan_predicting_2011}, a lexical entailment of a (possibly embedded) sentence is a presupposition, if it is about a time that is not the event time of the matrix predication and it is not at-issue with respect to the Question Under Discussion (QUD, see \citealt{roberts_information_2012}) addressed by the utterance.
 For instance, the CC of B's utterance in \ref{ent}, that Phil's ballet class is canceled, is predicted to be a presupposition (and therefore to project) because it is a lexical entailment of the modal prejacent and not at-issue with respect to A's interrogative utterance.
@@
 
-    In \citealt{simons_best_2017}, the CC of a clause-embedding predicate projects if it is entailed by the Current Question of the utterance (where the Current Question is the question that is congruent with the utterance).\footnote{The Current Question is defined in \citealt[194]{simons_best_2017} as follows: ``The Current Question for an utterance is a privileged subset of the focal alternative set of the uttered sentence (given a structural analysis of that sentence, including focus marking)'' which meets the conditions that ``i) the proposition expressed is a member of the Current Question and ii) the Current Question has at least one additional member.''} In \ref{ent},  the Current Question of B's utterance might be the set of propositions \{Phil forgot that his ballet class is canceled today, Phil is aware that his ballet class is canceled today\}.
+    In \citealt{simons_best_2017}, the CC of a clause-embedding predicate projects if it is entailed by the Current Question of the utterance (where, roughly, the Current Question is the question that is congruent with the utterance).\footnote{The Current Question is defined in \citealt[194]{simons_best_2017} as follows: ``The Current Question for an utterance is a privileged subset of the focal alternative set of the uttered sentence (given a structural analysis of that sentence, including focus marking)'' which meets the conditions that ``i) the proposition expressed is a member of the Current Question and ii) the Current Question has at least one additional member.''} In \ref{ent},  the Current Question of B's utterance might be the set of propositions \{Phil forgot that his ballet class is canceled today, Phil is aware that his ballet class is canceled today\}.
 If so, the Current Question entails that Phil's ballet class is canceled today, and the CC may therefore project under \poscite{simons_best_2017} analysis.
-		%
-		%In such analyses, every utterance addresses a question, which is determined by its alternative-structure, which, can be indicated by focus-marking.
-For example, in a modalized utterance with \emph{be annoyed}, the CC can be focused \Next[b], which leads to the alternative structure in \Next[c], allowing the utterance to address a question like \Next[a].
-
-		%\ex. \a. What was Cole annoyed about today?
-		%	\b. Perhaps Cole was annoyed [$_F$ that Julian dances salsa].
-		%	\b. $\{p \mid \exists q(p = \lambda w.$\emph{Cole was annoyed that} $q$ \emph{in} $w)\}$
-		%	\z.
-		%\z.
-
-		%In contrast, focus on the subject \Next[b] leads to the alternative structure in \Next[c], allowing the utterance to address a question like \Next[a].
-
-		%\ex. \a. Who was annoyed today?
-		%	\b. Perhaps [$_F$ Cole] was annoyed that Julian dances salsa.
-		%	\b. $\{p \mid \exists x(p = \lambda w.x$ \emph{was annoyed in} $w$ \emph{that Julian dances salsa}$)\}$
-		%	\z.
-		%\z.
 
@@
     %
-    Second, the analyses do not make systematic predictions for non-entailed CCs, and hence they do not predict that the CCs of some non-factives predicates project at least as much or even more than those of some factive predicates.
-
-    Both of these analyses set aside the effect of entailment-cancelling operators in determining the main point of an utterance.
+    Second, the analyses do not make systematic predictions for non-entailed CCs, and hence they do not predict that the CCs of some non-factive predicates project at least as much or even more than those of some factive predicates.
+
+    Both of these analyses set aside the effect of entailment-canceling operators in determining the main point of an utterance.
 Therefore they do not make predictions about by-operator differences or by-predicate variation in the effect of entailment-canceling operators.
@@
 
-    Under the view put forward in \citealt{schlenker_triggering_2021}, the CC of a sentence $S$ like \ref{knowa}, is presupposed in a context $c$, if the CC is an epistemic precondition on the global meaning of $S$.
-Formally, Schlenker defines presuppositions relative to their local context (\S6.2).
-For instance, in \ref{knowa}, the negative prejacent $S'$, that is \ref{knowb}, is presupposed in the local context under negation $c'$, if two conditions are met:
+    Under the view put forward in \citealt{schlenker_triggering_2021}, the CC of a sentence $S$ like \ref{knowa}, is presupposed in a context $c$, if the CC 
+    is presupposed by the sentence $S'$ under the entailment-canceling operator, that is \ref{knowb} in the local context $c'$.
+    %is an epistemic precondition on the global meaning of $S$. 
+    %Formally, Schlenker defines presuppositions relative to their local context (\S6.2).
+    For the CC to be presupposed in \ref{knowb}, two conditions must be met %for $S'$, the prejacent of negation in \ref{knowb}:
 	i) $S'$ contextually entails the CC relative to $c'$;
@@
 Therefore, epistemic preconditions relative to $c'$ will also apply to the global context.
-
-	Applied to \ref{know}, \poscite{schlenker_triggering_2021} analysis predicts that the CC  is presupposed if i) \ref{knowb} contextually entails that Julian dances salsa, and ii) if a generic agent would typically antecedently believe that Julian dances salsa upon interpreting \ref{knowb} in the minimal contexts we provided to our participants.
+Applied to \ref{know}, \poscite{schlenker_triggering_2021} analysis predicts that the CC  is presupposed if i) \ref{knowb} contextually entails that Julian dances salsa, and ii) a generic agent ``typically antecedently believes'' that Julian dances salsa upon interpreting \ref{knowb} in the contexts provided to our participants.
 
@@
 \citealt{schlenker_triggering_2021} also assumes the condition ii) is met: ``in many cases, one's knowledge of facts will precede one's knowledge of [Cole's] beliefs about them\ldots believing that [Julian dances salsa] is often an epistemic precondition for believing that'' Cole knows that Julian dances salsa (p.6).
-One might, however, challenge this assumption on the basis of the corpus study presented in \citealt{spenader_presuppositions_2002}, which showed that the CCs of the majority of the utterances of sentences with factive verbs (namely 81 out of 109) had to be accommodated (i.e., were not contextually entailed).
-In other words, utterance of sentences with the factive predicates investigated by \citet{spenader_presuppositions_2002}, which included \emph{know}, were ``generally used to communicate information the speaker thought was hearer-new'' (p.99).
+One might, however, challenge this assumption on the basis of the corpus investigation presented in \citealt{spenader_presuppositions_2002}, which showed that the CCs of the majority of the utterances of sentences with factive verbs (namely 81 out of 109) had to be accommodated (i.e., were not contextually entailed).
+In other words, utterances of sentences with the factive predicates investigated by \citet{spenader_presuppositions_2002}, which included \emph{know}, were ``generally used to communicate information the speaker thought was hearer-new'' (p.99).
 This result might therefore suggest that one cannot assume that a generic agent typically antecedently believes the CC of {\em know}.
 
-    \poscite{schlenker_triggering_2021} analysis does not incorporate differential effects of entailment-cancelling operators on projection.
-As discussed for \citealt{heim_projection_1983}, above, the local contexts under negation and in conditional antecedents are both assumed to be the global context (and similarly in \citealt{schlenker_local_2009}).
+    \poscite{schlenker_triggering_2021} analysis does not incorporate differential effects of entailment-canceling operators on projection.
+As discussed for \citealt{heim_projection_1983} above, the local contexts under negation and in conditional antecedents are both assumed to be the global context (and similarly in \citealt{schlenker_local_2009}).
 Therefore, the analysis does not make predictions about by-operator projection variation or about by-predicate variation in the effect of entailment-canceling operator.
@@
     %
-    Because the probabilities about previous beliefs and the threshold $\alpha$ are assumed to potentially vary based on the context, the formal proposal does not in principle preclude the possibility of by-predicate projection variation.
-However, the subjective conditional probabilities associated with expressions and the contextual threshold $\alpha$ are taken as given in the account, and therefore do not offer explicit or systematic predictions about projection ratings for different expressions.
+    The account takes the subjective conditional probabilities associated with expressions and the contextual threshold $\alpha$ as given, and therefore does not offer explicit or systematic predictions about projection ratings for different expressions.
     %
@@
     % Thus,
-    one advantage of \poscite{schlenker_triggering_2021} analysis over the analyses reviewed in sections \ref{s:heim} and \ref{s:abrusan} is that it predicts the projection of the CCs of (certain) non-factive predicates (modulo the open questions about condition ii)).
+    one advantage of \poscite{schlenker_triggering_2021} analysis over the analyses reviewed in Sections~\ref{s:heim} and \ref{s:abrusan} is that it predicts the projection of the CCs of (certain) non-factive predicates (modulo the open questions about condition ii).
 
     Based on these assumptions, the analysis correctly predicts that the CCs of \emph{know, inform}, and \emph{announce} are more projective than the CCs of \emph{demonstrate} and \emph{establish}, by virtue of the CCs of the former being usually presupposed, in contrast to the CCs of the latter.
-It is not clear, however, that the analysis is able to predict the observed variation, because the analysis -- even  though it does not divide predicates into factive and non-factives ones -- nevertheless imposes a binary, categorical distinction between predicates.
+It is not clear, however, that the analysis is able to predict the observed variation, because the analysis -- even  though it does not divide predicates into factive and non-factive ones -- nevertheless imposes a binary, categorical distinction between predicates.
     % such that the CCs of predicates with probabilities that are usually above the threshold are presupposed, and those below the threshold are not.
     %
-    Crucially, the by-predicate variation reported in Sections \ref{s:by-pred}+\ref{s:factivity} is not captured by an analysis that does not make more fine-grained distinctions between the meanings of clause-embedding predicates, as discussed extensively in \citealt{degen_are_2022}.
+    Crucially, the by-predicate variation reported in Sections~\ref{s:by-pred} and \ref{s:factivity} is not captured by an analysis that does not make more fine-grained distinctions between the meanings of clause-embedding predicates, as discussed extensively in \citealt{degen_are_2022}.
 First, there is projection variation between the supposedly presupposed CCs (that of \emph{know} is more projective than the CC of \emph{inform}, which is more projective than that of \emph{announce}, under all four operators).
@@
     Furthermore, as discussed in \citealt{degen_are_2022}, the CCs of the supposedly nonpresupposed CCs (of \emph{establish} and \emph{demonstrate}) are projective when compared to nonprojective main clause content.
-In the experiments reported on in section \ref{s2}, the mean certainty rating of the CC of \emph{announce}, whose CC is assumed to possibly be presupposed, is $0.55$, and that of {\em demonstrate}, whose CC is not assumed to be presupposed, is $0.43$ (both means are aggregated across entailment-canceling operators).
+In the experiments reported on in Section~\ref{s2}, the mean certainty rating of the CC of \emph{announce}, whose CC is assumed to possibly be presupposed, is $.55$, and that of {\em demonstrate}, whose CC is not assumed to be presupposed, is $.43$ (both means are aggregated across entailment-canceling operators).
 It is not clear whether this particular difference in mean certainty rating motivates analyzing the CC of \emph{announce} as presupposed in contrast to that of \emph{demonstrate}.
@@
 
-    There is currently no projection analysis on the market that predicts the by-predicate and by-operator variation we observed in the experiments reported on in section \ref{s2}.
+    There is currently no projection analysis on the market that predicts the by-predicate and by-operator variation we observed in the experiments from Section~\ref{s2}.
 \citealt{degen_are_2022} suggested that an empirically adequate analysis to capture the observed by-predicate variation requires consideration of ``more fine-grained distinctions [between clause-embedding predicates] that are based on the lexical meaning and discourse use of clause-embedding predicates'' (p.~585).
-Further, an account of the observed by-predicate variation in the effects of entailment-cancelling operators will need to take into account how semantic and pragmatic properties of predicate meanings interact with entailment-cancelling operators in varying ways.
-In the next section, we offer some hypotheses based on the data from our experiments.
+Further, an account of the observed by-predicate variation in the effects of entailment-canceling operators will need to consider how semantic and pragmatic properties of predicate meanings interact with entailment-canceling operators.
+We offer some suggestions based on the data from our experiments in the next section.
 
@@
 
-    Can the observed interactions between predicate and entailment-canceling operator be predicted from lexical semantic/pragmatic properties of the predicates?
+The results of our experiments suggest that there is by-predicate variation in the effect of entailment-canceling operator.
+In this section, we suggest that the observed variation may be predictable, at least in part, from lexical semantic and pragmatic properties shared by subsets of the 20 clause-embedding predicates featured in our experiments.
+Specifically, we observe that the projection of the CCs of the 20 predicates we investigated exhibit (what we call) `projection patterns', such that the CCs of several predicates exhibit the same or at least a similar by-operator projection variation.
+The six patterns we identified based on our results are shown in Figure~\ref{fig:patterns}, which shows mean certainty ratings for the four operators by projection pattern.
+
+    %Can the observed interactions between predicate and entailment-canceling operator be predicted from lexical semantic/pragmatic properties of the predicates?
 This is a pressing question for future research, to which our data offer some tentative answers.
 
-    We can find some initial generalizations over lexical properties, indicated in Figure~\ref{fig:patterns}, which gives the mean certainty ratings for the four operators by predicate, identifying four groups of predicates that show similar by-operator variation.
+    %We can find some initial generalizations over lexical properties, indicated in Figure~\ref{fig:patterns}, which gives the mean certainty ratings for the four operators by predicate, identifying six groups of predicates that show similar by-operator variation.
 
@@
 		\includegraphics[width=\textwidth]{predicate-profiles.pdf}
-		\caption{Mean certainty ratings by operator with $95\%$ bootstrapped confidence intervals, for some groups of predicates (\lq predicate patterns\rq).}
+		\caption{Mean certainty ratings by operator and predicate, by projection pattern. Error bars indicate $95\%$ bootstrapped confidence intervals.}
 		\label{fig:patterns}
@@
 
-	%(a)
-	The non-veridical predicates \emph{pretend} and \emph{think} exhibit the `Negation high' pattern, shown in panel (a) of Figure~\ref{fig:patterns}.
+We suggest that these patterns are not accidental, but that the predicates that share a pattern also share lexical semantic and pragmatic meanings.
+For instance, the non-veridical predicates \emph{pretend} and \emph{think} exhibit the `Negation high' pattern, shown in panel (a) of Figure~\ref{fig:patterns}.
 These are the only predicates that are most projective under negation compared to all other operators.
-This generalization, supported by the statistical analysis in Section \ref{s:by-pred} (see Figure \ref{fig:op-pred-analysis}), could be related to assumptions that non-veridical doxastics like \emph{think} can often be interpreted in comparison to a stronger veridical alternative (e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
+This generalization, supported by the statistical analysis in Section~\ref{s:by-pred} (Figure~\ref{fig:op-pred-analysis}), might be derivable from the observation that there is an anti-veridical inference for the CC of both \emph{think} and \emph{pretend}.
+For \emph{think}, this anti-veridical inference can arise from the fact that \emph{think} has a veridical alternative in \emph{know}
+(e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
+We tentatively hypothesize that this can lead to an inference that the CC is false in upward-monotone contexts, but not under negation.
+For \emph{pretend}, one might either assume a similar alternative or investigate whether this predicate entails that the speaker assumes that the CC is false.
+
+%assumptions that non-veridical doxastics like \emph{think} can often be interpreted in comparison to a stronger veridical alternative e.g., \citealt{heim_artikel_1991,chemla_epistemic_2008}).
 We tentatively hypothesize that this can lead to an inference that the CC is false in upward-monotone contexts, but not under negation.
@@
 
-	%(b)
-	These latter two predicates, along with \emph{be right}, exhibit the `Question low' pattern, shown in panel (b).
-They are the only predicates which are least projective in polar questions (again, compare Figure \ref{fig:op-pred-analysis}).
-We tentatively suggest that these communicative predicates can interact with the pragmatics of polar questions in a way that can lead to an inference of incredulity towards the CC.
+The predicates \emph{suggest, be right}, and \emph{say} (`Question low' pattern, panel (b)) are the only predicates which are least projective in polar questions (again see Figure \ref{fig:op-pred-analysis}).
+We tentatively suggest that these predicates can interact with the pragmatics of polar questions in a way that can lead to an inference of incredulity towards the CC.
 However, there are also some differences between these predicates.
-The veridical \emph{be right} is most projective under modals (M $>$ C $>$ N $>$ Q), whereas \emph{say} and \emph{suggest} are most projective in conditional antecedents (C $>$ M,  N $>$ Q).
+The veridical \emph{be right} is most projective under modals (M $>$ C $>$ N $>$ Q), whereas \emph{say} and \emph{suggest} are most projective in conditional antecedents (C $>$ \{M,  N\} $>$ Q).
 
 	%(c+d)
-	The two u-shaped patterns in panels (c) and (d) are characterized by relatively low projection ratings with negation and high ratings with conditionals.
-While there is fine-grained variation with regards to which of the by-operator differences were statistically supported, (c) and (d) can be distinguished by their ratings for the modal.
-	%
-	The change-of-state communication predicates \emph{admit, announce, confess}, and \emph{reveal}, exhibit relatively low projection ratings of the CC when embedded under \emph{perhaps} (C $>$ Q  $>$ M, N).
-In contrast, the inferential predicates \emph{confirm, demonstrate, establish}, and \emph{prove} exhibit relatively high ratings with \emph{perhaps} (\emph{confirm:} C $>$ M, Q $>$ N; \emph{demonstrate:} C $>$ M, Q $>$ N; \emph{establish:} C, M $>$ Q, N; \emph{prove:} M $>$ C  $>$ Q, N).
-For an explanation of the high conditional ratings, one might examine ways in which the discourse effect of a conditional interacts with the change-of-state component of these inferential and communicative predicates.
-Our tentative suggestion for the relatively low negation ratings is that these predicates can be interpreted relative to contextual assumptions that lead to a neg-raising type inference more readily than others (so that, for instance, not announcing $p$ amounts to communicating not $p$, or not proving $p$ amounts to inferring not $p$).
+The two $\cup$-shaped patterns in panels (c) and (d) are characterized by relatively low projection ratings with negation and high ratings with conditionals.
+While there is fine-grained variation with regard to which the by-operator differences were statistically supported, (c) and (d) can be distinguished by their ratings for the modal.
+The predicates \emph{admit}, \emph{announce}, \emph{confess} and \emph{reveal}, exhibit relatively low projection ratings of the CC when embedded under \emph{perhaps} (C $>$ Q  $>$ \{M, N\}).
+In contrast, the predicates \emph{confirm, demonstrate, establish}, and \emph{prove} exhibit relatively high ratings with \emph{perhaps} (\emph{confirm}, \textit{prove}: M $>$ C $>$ Q $>$ N; \emph{demonstrate:} C $>$ \{M, Q\} $>$ N; \emph{establish:} \{C, M\} $>$ \{Q, N\}).
+For an explanation of the high conditional ratings, one might examine ways in which the discourse effect of a conditional interacts with a change-of-state meaning component of these inferential and communicative predicates.
+A possible explanation for the relatively low negation ratings is that these predicates can be interpreted relative to contextual assumptions that lead to a neg-raising type inference more readily than others (so that, for instance, not announcing $p$ amounts to communicating not $p$, or not proving $p$ amounts to inferring not $p$).
 
 	%(e+f)
-	Finally, the two roof-shaped patterns in (e) and (f) are characterized by relatively high projection ratings with questions, and low ratings with {\em perhaps}.
+	Finally, the two $\cap$-shaped patterns in (e) and (f) are characterized by relatively high projection ratings with questions, and low ratings with {\em perhaps}.
 We can distinguish the two patterns by their relative ratings for conditionals.
-The predicates \emph{acknowledge, discover, hear,} and \emph{see}, which are associated with a change of some informational state, show relatively high ratings with conditionals (Q, C $>$ N $>$ M).
-The ratings for conditionals are relatively lower for the predicates \emph{be annoyed} (Q $>$ N $>$ C $>$ M), and \emph{inform}, and \emph{know} (Q $>$ N, C $>$ M), whose CCs are among the most projective.
+The predicates \emph{acknowledge, discover, hear,} and \emph{see}, which are associated with a change of some informational state, show relatively high ratings with conditionals (\{Q, C\} $>$ N $>$ M).
+The ratings for conditionals are relatively lower for the predicates \emph{be annoyed} (Q $>$ N $>$ C $>$ M), and \emph{inform}, and \emph{know} (Q $>$ \{N, C\} $>$ M), whose CCs are among the most projective.
+
+    Although the patterns we have identified here are tentative and only based on few predicates, we believe that future investigations into shared projection patterns across a wider range of predicates is a fruitful enterprise for future investigations of projection inferences.
 
@@
 As discussed, there is currently no projection analysis on the market that is able to predict the observed by-predicate variation or the by-operator variation.
-The results of our experiments also extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factives predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
+The results of our experiments also extend a result of \citealt{degen_are_2022}, that projection ratings in polar questions do not categorically distinguish factive from non-factive predicates, to cases with negation, the epistemic possibility modal \textit{perhaps}, and conditional antecedents.
 This results strengthens the conclusion of \citealt{degen_are_2022} that there is, to date, no empirical evidence for a coherent class of factive predicates.
@@
 \bibliography{projection}
+
+\end{document}
 
@@
 Each of the 20 clauses was paired with a unique consequent clause, as shown in the list below.
-To minimize the variability of the effect of the contents of these consequent clauses on the projection of the contents of the 20 complement clauses, the consequent clauses all consist of a uniquely named subject and an adjectival predication in the future tense ({\em will be}), and the adjectives all denote an emotion.
+To minimize the variation of the effect of the contents of these consequent clauses on the projection of the contents of the 20 complement clauses, the consequent clauses all consist of a uniquely named subject and an adjectival predication in the future tense ({\em will be}), and the adjectives all denote an emotion.
 We selected the 20 emotion-denoting adjectives based on the valence and arousal values reported in \citealt{warriner_norms_2013}: 10 of the adjectives had a positive valence, and 10 had a negative valence; all 20 adjective had an arousal value between 4.7 and 6.5.

### 2024-02-21 results/main/meta-analyses/1_projection/rscripts/1_summary-graphs.R

@@ -8,14 +8,15 @@ setwd(this.dir)
 
 # summary + graphing
 library(ggplot2)
-library(ggh4x)
-library(ggrepel)
+# library(ggh4x)
+# library(ggrepel)
 source('../../../helpers.R')
-cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
-                "#CC79A7", "#000000")
-graphcolor <- "#0072B2"
-pinkk <- "#CC79A7"
-pred_colors <- c(pinkk, "gray")
+# created separate color blind-friendly palettes
+palette1 <- c("#0072B2", "#009E73", "#E69F00", "#404040")
+factives_color <- "#CC79A7"
+palette2 <- c("#332288", "#404040", "#AA4499", "#56B4E9")
+# palette3 <- c("#404040", "#6A3D9A", "#33A02C", "#E69F00")
+pred_colors <- c(factives_color, "gray")
 theme_set(theme_bw())
 
 # LOAD DATA
@@ -24,6 +25,8 @@ data <- data %>% mutate(projection = as.numeric(projective), predicate = as.fact
                         operator = as.factor(op), participant = as.factor(workerid), 
                         item = as.factor(content), .keep = "none")
 levels(data$operator) <- c("conditional", "modal", "negation", "question")
+data <- data %>% mutate(predicate = fct_reorder(predicate, projection), 
+                        operator = fct_reorder(operator, projection))
 str(data)
 
 
@@ -32,26 +35,31 @@ str(data)
 # projection means and 95% CIs by operator 
 operator_means <- data %>% group_by(operator) %>% 
   summarize(Mean = mean(projection), CILow = ci.low(projection), 
-            CIHigh = ci.high(projection)) %>%
+            CIHigh = ci.high(projection)) %>% ungroup() %>%
   mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
-         operator = fct_reorder(as.factor(operator), Mean)) %>% ungroup()
-operator_means
-operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
+         operator = fct_reorder(as.factor(operator), Mean)) %>% 
+  mutate(projection = Mean, .drop = "used")
+
+str(operator_means)
 
 # plot
 operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), 
                               y = projection, label = round(projection, digits = 2))) +
[…]

### 2024-02-21 results/main/meta-analyses/1_projection/rscripts/3_analysis-predicates.R

@@ -19,11 +19,12 @@ library(emmeans)
 library(ggplot2)
 library(ggh4x)
 source('../../../helpers.R')
-cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", 
-                "#CC79A7", "#000000")
-graphcolor <- "#0072B2"
-pinkk <- "#d694ba"
-pred_colors <- c(pinkk, "black")
+# created separate color blind-friendly palettes
+palette1 <- c("#0072B2", "#009E73", "#E69F00", "#404040")
+factives_color <- "#CC79A7"
+palette2 <- c("#332288", "#404040", "#AA4499", "#56B4E9")
+# palette3 <- c("#404040", "#6A3D9A", "#33A02C", "#E69F00")
+pred_colors <- c(factives_color, "gray")
 theme_set(theme_bw())
 
 # LOAD DATA
@@ -299,8 +300,8 @@ predicate_operator_means %>%
                      labels = c("0",".2",".4",".6",".8","1"),
                      name = "Certainty rating") +
   scale_linetype_identity() +
-  scale_colour_manual(values = cbbPalette, name = "Operator") +
-  scale_fill_manual(values = cbbPalette, name = "Operator") +
+  scale_colour_manual(values = palette1, name = "Operator") +
+  scale_fill_manual(values = palette1, name = "Operator") +
   theme(legend.position="top") +
   theme(panel.grid.minor = element_blank()) +
   # theme(strip.background = element_rect(fill="white")) +
